### 2022-02-25 paper.tex
init

%!TEX encoding = UTF-8 Unicode 
\documentclass[12pt]{article} %try amsproc, amsart
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper, margin=2.5cm}                   % ... or a4paper or a5paper or ... 
% \geometry{landscape}                % Activate for for rotated page geometry
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{setspace}
\setstretch{1.1}

\usepackage{hyperref}
\hyperbaseurl{}
\urlstyle{same}
\usepackage[ruled]{algorithm2e}
\newcommand{\lIfElse}[3]{\lIf{#1}{#2 \textbf{else}~#3}}

\usepackage{times}
\usepackage{letltxmacro}
\usepackage[shortlabels]{enumitem}
\usepackage{kotex}
\usepackage{fancyhdr}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

% Language selector
\newif\ifEN
\ENtrue % Change to false to export Korean
%\ENfalse

\ifEN
{
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\newtheorem{conjecture}{Conjecture}
}
\else {
\newtheorem{theorem}{정리}
\newtheorem{lemma}{기본정리}
\newtheorem{corollary}{따름정리}
\newtheorem{proposition}{제의}
\theoremstyle{definition}
\newtheorem{example}{예}
\newtheorem{definition}{정의}
\newtheorem{assumption}{가정}
\newtheorem{conjecture}{추측}
}
\fi

%\ifEN {} \else \renewcommand*{\proofname}{증명\@}  \fi

\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\fancyhf{}
\rhead{Uncertainty in college admissions markets \\~}
\lhead{Max Kapur \\ \date{\today}}
\cfoot{\thepage}

\ifEN{ \title{How to apply to college} }
\else {\title{How to apply to college}} \fi
\author{Max Kapur}
\date{\today}                % Activate to display a given date or no date

\begin{document}

\maketitle
\section*{Abstract}
This paper considers the maximization of the expected maximum value of a portfolio of random variables subject to a budget constraint. We refer to this as the optimal college application problem. When each variable's cost, or each college's application fee, is identical, we show that the optimal portfolios are nested in the budget constraint, yielding an exact polynomial-time algorithm. When colleges differ in their application fees, the problem is NP-complete. For this more general problem, we provide two dynamic programs that produce exact solutions in pseudopolynomial time and a polynomial-time approximation scheme. 
\pagebreak

\tableofcontents
\pagebreak

\ifEN \section{Introduction}  \else \section{서론} \fi
We consider the optimal college application strategy for a single student. As Chao (2014) remarked, this represents a somewhat subtle portfolio optimization problem. The traditional Markowitz model trades off the expected value across all assets with a risk term, obtaining a concave maximization problem with linear constraints. But college applicants maximize the observed value of their \emph{best} asset: If a student is admitted to her $j$th choice, then she is indifferent to whether or not she gets into her $(j+1)$th choice. As a result, the valuation function that students maximize is \emph{convex} in the expected utility associated with individual applications. Risk management is implicit in the college application problem because, in a typical admissions market, college preferability is negatively correlated with competitiveness. Thus, students must negotiate a tradeoff between highly attractive, selective “reach schools” and less preferable “safety schools” where admission is a safer bet. Finally, the combinatorial nature of the college application problem makes it difficult to solve using the gradient-based techniques used in continuous portfolio optimization.

Chao estimated her model (which considers application as a \emph{cost} rather than a constraint) by clustering the schools so that $m=8$, a scale at which enumeration is possible. However, subject to certain assumptions on the quality of the data available to students in their decisionmaking process, an optimal application portfolio for a single student can be computed in time polynomial in $h$ and $m$, as we show presently.

% Nestedness property and paper

\ifEN \section{Model}  \else \section{모형} \fi \label{optimalcollegeappstrat}

\ifEN \section{Homogeneous application costs}  \else \subsection{} \fi
Consider a college admissions market with $m$ schools, $\mathcal{C} = \{ 1 \dots m\}$. The $j$th school is named $c_j$. By government regulation, students are allowed to apply to no more than $h$ schools. (In the Korean case, $m=202$ and $h=3$.) We consider the optimal application strategy for a single student, whom we will call Alma.

For $j=1\dots m$, let $t_j > 0$ denote the utility that Alma receives from attending $c_j$, and let $f_j$ denote the probability that she is admitted if she applies. Let the random variable $Z_j$ equal one if Alma gets into $c_j$ and zero otherwise.  We assume that Alma’s admissions outcome at each school is independent of her outcome at the other schools.\footnote{This assumption is appropriate when $f$ gives the admissions probabilities \emph{specifically} for Alma, as estimated from her academic abilities. It is inappropriate to take a school's overall acceptance rate, which reflects the admissions probability for a \emph{generic} student, as $f_j$.} Thus $Z$ is a vector of independent Bernoulli variables with probabilities given by $f$. Let $c_0$ denote Alma's outcome if she does not get into any college, with utility $t_0$ and $f_0 = 1$. Without loss of generality, we assume that $t_{j-1} \leq t_j$ for $j=1 \dots m$.(We may assume that $t_0 \leq t_1$ because schools for which $t_j < t_0$ can be trivially excluded from consideration.)
%For the \emph{generic} student, when $f_j$ is simply $c_j$'s acceptance rate, admissions outcomes will almost certainly correlate across schools.

Let $\mathcal{X} \subseteq \mathcal{C}$ denote the set of schools to which Alma applies, called her \emph{application portfolio.}The expected utility Alma receives from $\mathcal{X}$ is called the portfolio’s \emph{valuation}. %Let $x$ denote the same encoded as a binary vector, where $x_j=1 \iff j\in \mathcal{X}$ for $j=1\dots m$. 
\begin{definition}[Portfolio valuation function]
$v(\mathcal{X}) = \operatorname{E}\left[ \max\{t_j Z_j : j \in \mathcal{X}\} \right]$.
\end{definition}
It is helpful to define the random variable $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$ as the utility achieved by the portfolio, so that $v(\mathcal{X}) = \operatorname{E}[X]$.

Let $p_j(\mathcal{X})$ denote the probability that Alma attends $c_j$. Alma attends $c_j$ if and only if she \emph{applies} to $c_j$, is \emph{admitted} to $c_j$, and is \emph{rejected} from any school she prefers to $c_j$; that is, any school with higher index. Hence, for $j= 0\dots m$,
\begin{align}
p_j(\mathcal{X}) &= 
\begin{cases}
\displaystyle f_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - f_{j’}), \quad & j \in \{0\}\cup\mathcal{X}\\
0, \quad & \text{otherwise}
\end{cases} \\
%\iff\quad p_j(x) &= x_j  f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’})
\end{align}
%with the understanding that $x_0 = 1$ and
where the empty product equals one. The following proposition follows immediately.

\begin{proposition}[Closed form of portfolio valuation function]
\begin{align}
v(\mathcal{X}) &= \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - f_{j’}) \Bigr)  \label{closedformportfoliovaluationX}%, \quad \text{or equivalently,}\\
%\qquad v(x) &= t_0 \prod_{j=1}^m (1 - f_{j} x_j) + \sum_{j=1}^m \Bigl( x_j t_j f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’}) \Bigr) \label{closedformportfoliovaluationx}
\end{align}
\end{proposition}
%\begin{proof}Computing $v(\mathcal{X}) = \sum_{j=0}^m  t_j p_j(\mathcal{X})$ yields \eqref{closedformportfoliovaluationX}. Next, because $1 - f_j x_j = 1$ if $x_j = 0$, we may define $p_j$ equivalently as $p_j(x) = x_j  f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’})$ to obtain \eqref{closedformportfoliovaluationx}. 
%\end{proof}

Next, we show that without loss of generality, we may assume that $t_0 = 0$ (or any constant).
\begin{theorem} \label{assumetzerozero}
Let $\bar t_j = t_j - \gamma$ for $j = 0 \dots m$. Then $v(\mathcal{X}; \bar t_j) = v(\mathcal{X};  t_j) -  \gamma$ regardless of $\mathcal{X}$. 
\end{theorem}
\begin{proof}
By definition, $\sum_{j=0}^m p_j(\mathcal{X}) = \sum_{j \in \{0\}\cup\mathcal{X}} p_j(\mathcal{X}) = 1$. Therefore
\begin{align}
v(\mathcal{X}; \bar t_j) &= \sum_{j\in \{0\}\cup\mathcal{X}}  \bar t_j p_j(\mathcal{X})
=\sum_{j\in \{0\}\cup\mathcal{X}} (t_j - \gamma) p_j(\mathcal{X}) \\
&=\sum_{j\in \{0\}\cup\mathcal{X}} t_j p_j(\mathcal{X})  - \gamma 
= v(\mathcal{X}; t_j) - \gamma
\end{align}
which completes the proof. 
\end{proof}

%Let us express the optimal college application problem as an INLP.
\begin{definition}[Alma’s problem]
Alma's optimal college application portfolio is given by the solution to the following integer nonlinear program:
\begin{align}
\begin{split}
\text{maximize}\quad &  v(\mathcal{X}) \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - f_{j’})\\
\text{subject to}\quad & \mathcal{X}\subseteq\mathcal{C}, \quad|\mathcal{X}| \leq h 
\end{split}
\end{align}
\end{definition}

\ifEN \subsection{Approximation properties of a na\"ive solution}  \else \subsection{모형화} \fi
Notice that for a given school $c_j$, the expected utility associated with applying to $c_j$ is simply $\operatorname{E}[t_j Z_j] = t_j f_j$. It is therefore tempting to adopt the following greedy strategy, which turns out to be inoptimal.
\begin{definition}[Na\"ive algorithm for Alma’s problem] \label{naivealgorithm}
Apply to the $h$ schools having the highest expected utility $t_j f_j$.
\end{definition}
The basic error of this algorithm is that it maximizes $\operatorname{E}\left[\sum t_j Z_j \right]$ instead of $\operatorname{E}\left[\max \{t_j Z_j\} \right]$. The latter is what Alma is truly concerned with, since in the end she can attend only one school. The following example shows that the na\"ive algorithm can produce a suboptimal solution.
\begin{example}
Suppose $m=3$, $q=2$, and
\begin{align*}
t &= (70, 80, 90) \\
f &= (0.4, 0.4, 0.3) \\
\implies t * f &= (28, 32, 27)
\end{align*}
The na\"ive algorithm picks $\mathcal{T} = \{1, 2\}$ with 
\[v(\mathcal{T}) = 70(0.4)(1-0.4) + 80(0.4) = 48.8\]
But $\mathcal{X} = \{2, 3\}$ with
\[v(\mathcal{X}) = 80(0.4)(1-0.3) + 90(0.3) = 49.4\]
is the optimal solution. 
\end{example}

In fact, the na\"ive algorithm is a $(1/h)$-approximation algorithm for Alma’s problem, as expressed in the following theorem.

\begin{theorem} \label{oneoverhopt}
When the application limit is $h$, let $\mathcal{X}_h$ denote the optimal portfolio, and $\mathcal{T}_h$ the set of the $h$ schools having the largest values of $t_j f_j$. Then $v(\mathcal{T}_h) / v(\mathcal{X}_h) \geq 1/h$. 
\end{theorem}
\begin{proof}
Because $\mathcal{T}_h$ maximizes the quantity $\operatorname{E}\bigl[ \sum_{j \in \mathcal{T}_h}\{ t_j Z_j \}\bigr]$, we have
\begin{align} \label{oneoverhopt}
\begin{split}
v(\mathcal{X}_h) &= \operatorname{E}\bigl[ \max_{j \in \mathcal{X}_h}\{ t_j Z_j \}\bigr] \leq \operatorname{E}\bigl[ \sum_{j \in \mathcal{X}_h}\{ t_j Z_j \}\bigr] \leq \operatorname{E}\bigl[ \sum_{j \in \mathcal{T}_h}\{ t_j Z_j \}\bigr] \\
&= h  \operatorname{E}\bigl[ \tfrac{1}{h} \sum_{j \in \mathcal{T}_h}\{ t_j Z_j \}\bigr]
\leq h  \operatorname{E}\bigl[ \max_{j \in \mathcal{T}_h}\{ t_j Z_j \}\bigr]
= h\,v(\mathcal{T}_h)
\end{split}
\end{align}
where the final inequality follows from the concavity of the $\max\{\}$ operator.
\end{proof}

The following example establishes the tightness of the approximation factor.

\begin{example} \label{tightexampleforoneoverhopt}
Pick any $h$ and let $m = 2h$. For a small constant $\varepsilon \in (0, 1)$, let
\begin{align*}
t &= \Bigl(\,\underbrace{1,  \dots, 1}_{h},\;
\underbrace{\varepsilon^{-1}, \varepsilon^{-2}, \dots, \varepsilon^{-(h-1)}, \varepsilon^{-h}}_{h}\,\Bigr) \\
\text{and}\quad f &= \Bigl(\,\underbrace{1,  \dots, 1}_{h},\;
\underbrace{\varepsilon^{1}, \varepsilon^{2}, \dots, \varepsilon^{h-1}, \varepsilon^{h}}_{h}\,\Bigr)
\end{align*}
Since all $t_j f_j = 1$, the na\"ive algorithm can choose $\mathcal{T}_h = \{1, \dots, h\}$, with $v(\mathcal{T}_h) = 1$. But the optimal solution is $\mathcal{X}_h = \{h+1, \dots, m\}$, with
\begin{equation*}
v(\mathcal{X}_h) = \sum_{j= h +1}^m \Bigl( t_j f_j \prod_{j' = j+1}^m (1 - f_{j'}) \Bigr) =  \sum_{j= 1}^h  (1 - \varepsilon)^{j} \approx h
\end{equation*}
Thus, as $\epsilon$ approaches zero, we have $v(\mathcal{T}_h) / v(\mathcal{X}_h) \to 1/h$. (The optimality of $\mathcal{X}_h$ follows from the fact that it achieves the upper bound of Theorem \ref{oneoverhopt}.)
\end{example}

Hope is not lost. We can still find the optimal solution in time polynomial in $h$ and $m$, as we will now show.

\ifEN \subsection{The nestedness property}  \else \subsection{풀이} \fi
It turns out that the solution to Alma's problem possesses a special structure: An optimal portfolio of size $h+1$ includes an optimal portfolio of size $h$ as a subset. To show this result, we need to introduce some new notation. For an application portfolio $\mathcal{X}$, the random variable $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$ represents the utility achieved by the portfolio, so that $v(\mathcal{X}) = \operatorname{E}[X]$.
%It is also helpful to define $\Delta v(\mathcal{X}, j)$ as the utility gained by adding $c_j$ to the portfolio $\mathcal{X}$. That is, 
%\begin{align} \Delta v(\mathcal{X}, j) &= v(\mathcal{X} \cup \{j\})  - v(\mathcal{X}) \\
%&= (1 - f_j) v(\mathcal{X}) + f_j \operatorname{E}[\max\{X, t_j\}] - v(\mathcal{X}) \\
%&= f_j \operatorname{E}[\max\{0, t_j - X\}]  \label{emaxtjminusX}
%\end{align}
%
%\begin{lemma}
%For all $\mathcal{X}\subseteq \mathcal{C}$ and $i, j \in \mathcal{C}$, $\Delta v(\mathcal{X}, j) \geq \Delta v(\mathcal{X} \cup \{i\}, j)$. 
%\end{lemma}
%\begin{proof}
%If a new school $i$ is added to $\mathcal{X}$, then the expected value of $X$ can only increase. Likewise, the probability that Alma is admitted to at least one school in $\mathcal{X}\cup\{i\}$ that she prefers to $j$ must be higher than in $\mathcal{X}$. Therefore, the maximum in \eqref{emaxtjminusX} is more likely to occur in the second entry, and achieves a lower value. 
%\end{proof}
%Now we may present the main result.

\begin{theorem}[Nestedness of optimal application portfolios] \label{nestedapplication}
Let $\mathcal{X}_h$ denote Alma’s optimal application portfolio when the application limit is $h$. If each $\mathcal{X}_h$ is unique, then
\begin{equation}
\mathcal{X}_1 \subset \mathcal{X}_2\subset \dots \subset \mathcal{X}_m
\end{equation}
If the optimal portfolios are not unique, then there is a sequence of optimal portfolios satisfying the above.
\end{theorem}
\begin{proof} By induction on $h$. Applying Theorem \ref{assumetzerozero}, we assume that $t_0 = 0$.

(Base case.) First, we will show that $\mathcal{X}_1 \subset \mathcal{X}_2$. To get a contradiction, suppose that the optima are $\mathcal{X}_1 = \{j\}$ and $\mathcal{X}_2 = \{k, l\}$, where we may assume that $t_k \leq t_l$. Optimality requires that
\begin{equation}v(\mathcal{X}_1 )  = f_j t_j > v(\{k\}) = f_k t_k\end{equation}
and
\begin{align}
v(\mathcal{X}_2) =  f_k (1- f_l) t_k + f_l t_l &> v(\{j, l\}) \\
& = f_j (1- f_l) t_j + (1- f_j) f_l t_l + f_j f_l \max\{t_j, t_l\} \\
&\geq  f_j (1- f_l) t_j + (1- f_j) f_l t_l + f_j f_l  t_l \\
&= f_j (1- f_l) t_j + f_l t_l  \\
&\geq f_k (1- f_l) t_k + f_l t_l  = v(\mathcal{X}_2)
\end{align}
which is a contradiction.

(Inductive step.) Assume that $\mathcal{X}_1 \subset \cdots \subset \mathcal{X}_h$, and we will show $\mathcal{X}_h \subset \mathcal{X}_{h+1}$. Let $k = \argmax\{ t_k: k \in \mathcal{X}_{h+1}\}$ and write $\mathcal{X}_{h+1} = \mathcal{Y}_{h} \cup \{k\}$.

Suppose $k \notin \mathcal{X}_h$.  To get a contradiction, assume that $v(\mathcal{Y}_h) < v(\mathcal{X}_h)$. Then
\begin{align}
\begin{split}
v(\mathcal{X}_{h+1})&= v(\mathcal{Y}_{h} \cup \{k\}) \\
&= (1 - f_k) v(\mathcal{Y}_h) + f_k t_k \\
& < (1 - f_k) v(\mathcal{X}_h) + f_k \operatorname{E}\bigl[ \max\{t_k, X_h\}\bigr]\\
&=  v(\mathcal{X}_h\cup \{k\})
\end{split}
\end{align}
contradicts the optimality of $\mathcal{X}_{h+1}$.

Now suppose that $k \in \mathcal{X}_h$. We can write $\mathcal{X}_h = \mathcal{Y}_{h-1} \cup \{k\}$, where $ \mathcal{Y}_{h-1}$ is some portfolio of size $h-1$. It suffices to show that $ \mathcal{Y}_{h-1} \subset \mathcal{Y}_h$. By definition, $\mathcal{Y}_{h-1}$ (respectively, $\mathcal{Y}_{h}$) maximizes the function $v(\mathcal{Y}\cup\{k\})$ over portfolios of size $h-1$ (respectively, $h$) that do not include $k$. That is, $\mathcal{Y}_{h-1}$ and $\mathcal{Y}_h$ are the optimal \emph{complements} to the singleton portfolio $\{k\}$.

We will use the function $w(\mathcal{Y})$ to grade portfolios $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$ according to how well they complement $\{k\}$. To construct $w(\mathcal{Y})$, let $\tilde t_j$ denote the expected utility Alma receives from school $c_j$ \emph{given} that she has been admitted to $c_j$ and applied to $c_k$. For $j < k$, including $j = 0$, this is $\tilde t_j = t_j (1- f_k) + t_k f_k$; for $j > k $, this is $\tilde t_j = t_j$. This means that 
\begin{equation}\label{Vyastildet}
v(\mathcal{Y}\cup\{k\}) = \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y})\end{equation}
The transformation to $\tilde t$ does not change the order of the $t_j$-values. Therefore, the expression on the right side of \eqref{Vyastildet} is itself a portfolio valuation function. In the corresponding market, $t$ is replaced by $\tilde t$ and $\mathcal{C}$ is replaced by $\mathcal{C}\setminus\{k\}$. Now, we obtain $w(\mathcal{Y})$ through one more transformation: Define $\bar t_j = \tilde t_j - \tilde t_0$ so that $t_0 = 0$ and let
\begin{equation}  \label{wYvXminusconst}
w(\mathcal{Y})
= \sum_{j \in \{0\} \cup \mathcal{Y}} \bar t_j p_j(\mathcal{Y})
= \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y})- \tilde t_0
= v(\mathcal{Y}\cup\{k\}) -  t_k f_k \end{equation}
where the second equality follows from Theorem \ref{assumetzerozero}. This identity says that the optimal complements to $\{k\}$, given by $\mathcal{Y}_{h-1}$ and $\mathcal{Y}_h$, are themselves optimal portfolios of size $h-1$ and $h$ for the market whose objective function is $w(\mathcal{Y})$. Since $\bar t_0 = 0$ in the latter market, the inductive hypothesis implies that $\mathcal{Y}_{h-1} \subset \mathcal{Y}_h$, which completes the proof.\footnote{We thank Yim Seho for discovering this critical transformation.}
\end{proof}

\subsection{Polynomial-time solution}
Applying the result above yields an efficient algorithm for the optimal portfolio: Start with the empty set and add schools one at a time, maximizing $v(\mathcal{X}\cup \{k\})$ at each addition. Sorting $t$ is  $O(m \log m)$.  At each of the $h$ iterations, there are $O(m)$ candidates for $k$, and computing $v(\mathcal{X}\cup \{k\})$ is $O(h)$ using \eqref{closedformportfoliovaluationX}; therefore, the time complexity of this algorithm is $O(h^2 m + m \log m)$. 
%\begin{algorithm}[H] 
%  %\DontPrintSemicolon
%\caption{Optimal portfolio algorithm for Alma’s problem when $h$ is small.} \label{algorithmforsmallh}
%\KwData{Utility values $t \in[0, \infty)^m$, admissions probabilities $f \in [0, 1]^m$, application limit $h \leq m$.}
%Index schools in ascending order by $t$\;
%$\mathcal{X} \gets \O$\;
%\For{$i=1\dots h$}
%{
%    $j \gets \argmax_k\bigl\{v(\mathcal{X}\cup\{k\}) : k \notin \mathcal{X}\bigr\}$\;
%    $\mathcal{X} \gets \mathcal{X}\cup\{j\}$ \;
%}
%\Return{$\mathcal{X}$}
%\end{algorithm}
%
%\begin{theorem}[Validity of Algorithm \ref{algorithmforsmallh}]
%Algorithm \ref{algorithmforsmallh} produces an optimal application portfolio in $O(h^2 m + m \log m)$ time.
%\end{theorem}
%\begin{proof}
%Optimality follows from Theorem \ref{nestedapplication}. Sorting $t$ is  $O(m \log m)$. At each of the $h$ iterations, there are $O(m)$ candidates for $k$, and computing $v(\mathcal{X}\cup \{k\})$ is $O(h)$ using \eqref{closedformportfoliovaluationX}; therefore, the time complexity of the main loop is $O(h^2 m)$. 
%\end{proof}

We reduce the computation time to $O(hm)$ by taking advantage of the transformation from the inductive step in the proof of Theorem \ref{nestedapplication}. Once school $k$ is added to $\mathcal{X}$, we remove it from the set $\mathcal{C}\setminus \mathcal{X}$ of candidates, and update the $t_j$-values of the remaining schools according to the following transformation:
\begin{align}\label{howtotransformfj}
\bar t_j = 
\begin{cases}
t_j (1 - f_k), \quad & t_j \leq t_k \\
t_j - t_k f_k, \quad& t_j > t_k
\end{cases}
\end{align}
It is easy to verify that this is the composition of the two transformations (from $t$ to $\tilde t$, and from $\tilde t$ to $\bar t$) given in the proof. Now, the \emph{next} school added must be the optimal singleton portfolio in the modified market. But the optimal singleton portfolio consists simply of the school with the highest value of $f_j \bar t_j$. Therefore, by updating the $t_j$-values at each iteration according to \eqref{howtotransformfj}, we eliminate the need to compute $v(\mathcal{X})$ entirely. Moreover, this algorithm does not require the schools to be indexed in ascending order by $t_j$, which removes the $O(m\log m)$ sorting cost.

The algorithm below outputs a list $\mathtt{X}$ of the $h$ schools to which Alma should apply. The schools appear in the order of entry such that when the algorithm is run with $h=m$, the optimal portfolio of size $h$ is given by $\mathcal{X}_h = \{\mathtt{X[1]}, \dots, \mathtt{X[h]}\}$. The entries of the list $\mathtt{V}$ give the valuation thereof.

\begin{algorithm}[H] 
%\DontPrintSemicolon
\caption{Optimal portfolio algorithm for Alma’s problem.} \label{algorithmforlargeh}
\KwData{Utility values $t \in[0, \infty)^m$, admissions probabilities $f \in [0, 1]^m$, application limit $h \leq m$.}
%Index schools in ascending order by $t$\;
$\mathcal{C} \gets \{1 \dots m\}$\;
$\mathtt{X, V} \gets $ empty lists\;
%$\mathtt{L} \gets $, an empty heap of named tuples $\mathtt{C}$ ordered by $\mathtt{C.f}\cdot\mathtt{C.t}$\;
%\lFor{$j=1\dots m$}
%{$\operatorname{append!}\bigl(\mathtt{L}, (j, f_j, t_j)\bigr)$}
\For{$i=1\dots h$}
{
    $k \gets \argmax_{j \in \mathcal{C}}\{f_j t_j\}$\;
    $\mathcal{C} \gets \mathcal{C} \setminus \{k\}$\;
    $\operatorname{append!}(\mathtt{X}, k)$\;
     \lIfElse{$i=1$}{$\operatorname{append!}(\mathtt{V}, f_k t_k)$}
     {$\operatorname{append!}(\mathtt{V}, \mathtt{V[i-1]} + f_k t_k)$}
    \For{$j \in \mathcal{C}$}
	{
	\lIfElse{$t_j \leq t_k$}{$t_j \gets  t_j (1 -  f_k)$}{$t_j \gets  t_j -  f_k t_k$}
	}
%    \For{$\mathtt{E} \in \mathtt{H}: 1\leq \mathtt{E.j} < \mathtt{D.j}$}
%	{
%	Update $\mathtt{E.t} \gets  \mathtt{E.t} \cdot(1 -  \mathtt{D.f})$\;
%    }
%    \For{$\mathtt{E} \in \mathtt{H}:  \mathtt{D.j} <  \mathtt{E.j} \leq m$}
%	{
%	Update $\mathtt{E.t} \gets  \mathtt{E.t} -  \mathtt{D.f}\cdot\mathtt{D.t} $\;
%    }
}
\Return{$\mathtt{X, V}$}
\end{algorithm}

\begin{theorem}[Validity of Algorithm \ref{algorithmforlargeh}]
Algorithm \ref{algorithmforlargeh} produces an optimal application portfolio for Alma's problem in $O(h m)$ time.
\end{theorem}
\begin{proof}
Optimality follows from the proof of Theorem \ref{nestedapplication}. Suppose $\mathcal{C}$ is stored as a list. Then at each of the $h$ iterations of the main loop, finding the top school costs $O(m)$, and the $t_j$-values of the remaining $O(m)$ schools are each updated in unit time. Therefore, the overall time complexity is $O(h m)$.
\end{proof}

In our numerical experiments, we found it effective to store $\mathcal{C}$ as a binary max heap rather than a list. The heap is ordered according to the criterion $i \geq j \iff f_i t_i \geq f_j t_j$. Nominally, using a heap increases the cost of the main loop from $O(h m)$ to $O(hm \log m)$ because the heap is rebalanced when each $t_j$-value is updated. However, typical problem instances do not achieve this upper bound because the order of the $f_j t_j$-values changes only slightly between iterations. The cost of updating each $t_j$-value can be reduced to unit time using a Fibonacci heap (Fredman and Tarjan 1987), yielding the same overall computation time. 
%The computation time can be further reduced by observing that the school added at each iteration must be maximal in the tuple $(f_j, t_j)$. That is, if $f_j \geq f_k \wedge t_j \geq t_k$ holds for two candidate schools $j$ and $k$, then $k$ cannot be the uniquely optimal addition, since admission to $j$ is both weakly preferable and weakly more probable. Therefore, we may restrict our attention to the subset of schools \emph{not} dominated by another in this way.
%\begin{proposition}[Efficient frontier for Alma's problem]
%Given a set of schools $\mathcal{D}$, define the set
%\begin{equation}
%\operatorname{EF}(\mathcal{D})  = \{j \in \mathcal{D}: f_j \geq f_k \vee t_j \geq t_k, \forall k \in \mathcal{D}\}
%\end{equation}
%as the \emph{efficient frontier} of $\mathcal{D}$. Then
%\begin{equation} l \in \operatorname{EF}(\mathcal{C}\setminus  \mathcal{X}_{h-1}) \end{equation}
%where $l$ is the school added at iteration $h$ of Algorithm \ref{dynamicprogram}.
%\end{proposition}

\subsection{Properties of the optimal portfolios}
The nestedness property implies that Alma's expected utility is a concave function of $h$.

\begin{theorem}[Optimal portfolio valuation concave in $h$] \label{concavityinh}
For $h = 2 \dots (m-1)$, \begin{equation}v(\mathcal{X}_h) - v(\mathcal{X}_{h-1}) \geq v(\mathcal{X}_{h+1}) - v(\mathcal{X}_{h})\end{equation} 
\end{theorem}
\begin{proof}
We will prove the equivalent expression $2 v(\mathcal{X}_h) \geq v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1})$. Applying Theorem \ref{nestedapplication}, we write $\mathcal{X}_h = \mathcal{X}_{h-1} \cup\{j\}$ and $\mathcal{X}_{h+1} = \mathcal{X}_{h-1} \cup\{j, k\}$. Define the random variables $X_i$ as above. If $t_k \leq t_j$, then 
\begin{align}
\begin{split}
2 v(\mathcal{X}_h) &= v(\mathcal{X}_{h-1} \cup\{j\}) + v(\mathcal{X}_{h-1} \cup\{j\}) \\
&\geq v(\mathcal{X}_{h-1} \cup\{k\}) + v(\mathcal{X}_{h-1} \cup\{j\}) \\
&= v(\mathcal{X}_{h-1} \cup\{k\}) + (1 - f_j) v(\mathcal{X}_{h-1}) + f_j \operatorname{E}[\max\{t_j, X_{h-1}\}] \\
&= v(\mathcal{X}_{h-1} \cup\{k\}) - f_j v(\mathcal{X}_{h-1}) + f_j \operatorname{E}[\max\{t_j, X_{h-1}\}] + v(\mathcal{X}_{h-1})  \\
&\geq v(\mathcal{X}_{h-1} \cup\{k\})  - f_j v(\mathcal{X}_{h-1}\cup\{k\}) + f_j \operatorname{E}[\max\{t_j, X_{h-1}\}]+ v(\mathcal{X}_{h-1})\\
&= (1 - f_j) v(\mathcal{X}_{h-1} \cup\{k\})  + f_j \operatorname{E}[\max\{t_j, X_{h-1}\}]+ v(\mathcal{X}_{h-1})\\
&=  v(\mathcal{X}_{h-1} \cup\{j, k\}) + v(\mathcal{X}_{h-1})\\
&=  v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1})
\end{split} 
\end{align}
The first inequality follows from the optimality of $\mathcal{X}_h$, while the second follows from the fact that adding $k$ to $\mathcal{X}_{h-1}$ can only increase its valuation.

If $t_k \geq t_j$, then the steps are analogous:
\begin{align}
\begin{split}
2 v(\mathcal{X}_h) &= v(\mathcal{X}_{h-1} \cup\{j\}) + v(\mathcal{X}_{h-1} \cup\{j\}) \\
&\geq v(\mathcal{X}_{h-1} \cup\{k\}) + v(\mathcal{X}_{h-1} \cup\{j\}) \\
&= (1 - f_k) v(\mathcal{X}_{h-1}) + f_k \operatorname{E}[\max\{t_k, X_{h-1}\}] +  v(\mathcal{X}_{h-1} \cup\{j\})  \\
&= v(\mathcal{X}_{h-1}) - f_k  v(\mathcal{X}_{h-1}) + f_k \operatorname{E}[\max\{t_k, X_{h-1}\}] +  v(\mathcal{X}_{h-1} \cup\{j\})  \\
&\geq v(\mathcal{X}_{h-1}) - f_k  v(\mathcal{X}_{h-1}\cup\{j\}) + f_k \operatorname{E}[\max\{t_k, X_{h-1}\}] +  v(\mathcal{X}_{h-1} \cup\{j\})  \\
&= v(\mathcal{X}_{h-1}) + (1 - f_k) v(\mathcal{X}_{h-1}\cup\{j\}) + f_k \operatorname{E}[\max\{t_k, X_{h-1}\}]  \\
&= v(\mathcal{X}_{h-1}) + v(\mathcal{X}_{h-1} \cup\{j, k\})
\end{split}\\
&= v(\mathcal{X}_{h-1})  + v(\mathcal{X}_{h+1}) \qedhere
\end{align}
\end{proof}

It follows that when $\mathcal{X}_h$ is the optimal $h$-portfolio, for a given market, $v(\mathcal{X}_h)$ is $O(h)$. Example \ref{tightexampleforoneoverhopt}, in which $v(\mathcal{X}_h)$ can be made arbitrarily close to $h$, establishes the tightness of this bound.

To wrap up, we provide an example showing that if the entries of $Z$ are dependent, then the optimal solution may violate the nestedness property of Theorem \ref{nestedapplication}.
\begin{example}
Let $t = (3, 3, 4)$, $Z_1 \sim \operatorname{Bernoulli}(0.5)$, $Z_2 = 1 - Z_1$,  and $Z_3 \sim \operatorname{Bernoulli}(0.5)$. Then it is easy to verify that the unique optimal portfolios are $\mathcal{X}_1 = \{3\}$ and $\mathcal{X}_2 = \{1, 2\}$. 
\end{example}

\section{Heterogeneous application costs}
In this section, we turn to the more general problem in which the constant $g_j$ represents the \emph{cost} of applying to $c_j$ and the student, whom we now call Ellis, has a \emph{budget} of $H$ to spend on college applications. Applying Theorem \ref{assumetzerozero}, we assume $t_0 = 0$ and disregard $c_0$ throughout.

\begin{definition}[Ellis's problem]
Ellis's optimal college application portfolio is given by the solution to the following integer nonlinear program:
\begin{align}
\begin{split}
\text{maximize}\quad & v(x) = \sum_{j=1}^m x_j t_j f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’}) \\
\text{subject to}\quad & \sum_{j=1}^m g_j x_j \leq H \\
&x_j \in \{0, 1\}, \quad j = 1\dots m
\end{split}
\end{align}
\end{definition}

The optima for Ellis's problem are not necessarily nested, nor is the number of schools in the optimal portfolio necessarily increasing in $H$. For example, if
$f = (0.5, 0.5, 0.5)$, $t = (1, 1, 1003)$, and $g = (1, 1, 3)$,
then it is evident that the optimal portfolio for $H = 2$ is $\{1, 2\}$ while that for $H = 3$ is $\{3\}$.

\subsection{NP-completeness}
In fact, Ellis’s problem is NP-complete, as we will show by a transformation from the binary knapsack problem, which is known to be NP-complete (Garey and Johnson 1979).

\begin{definition}[Decision form of knapsack problem]
An \emph{instance} consists of a set $\mathcal{B}$ of $m$ objects; utility values $u_j \in \mathbb{N}$ and weight $w_j \in \mathbb{N}$ for each $j \in \mathcal{B}$; and target utility $U\in \mathbb{N}$ and knapsack capacity $W\in \mathbb{N}$. The instance is called a \emph{yes-instance} if and only if there exists a set $\mathcal{B’} \subseteq \mathcal{B}$ having $\sum_{j \in \mathcal{B’}} u_j \geq U$ and  $\sum_{j \in \mathcal{B’}} w_j \leq W$.
\end{definition}

\begin{definition}[Decision form of Ellis’s problem]
An \emph{instance} consists of an instance of Ellis’s problem and a target valuation $V$. The instance is called a \emph{yes-instance} if and only if there exists a portfolio $\mathcal{X} \subseteq \mathcal{C}$ having $v(\mathcal{X}) \geq V$ and  $\sum_{j \in \mathcal{X}} g_j \leq H$.
\end{definition}

\begin{theorem}
The decision form of Ellis’s problem is NP-complete.
\end{theorem}

\begin{proof}It is obvious that the problem is in NP.

Consider an instance of the knapsack problem, and we will construct an instance of Ellis’s problem that is a yes-instance if and only if the corresponding knapsack instance is a yes-instance. Without loss of generality, we may assume that the the objects in $\mathcal{B}$ are indexed in increasing order of $u_j$, that each $u_j >0$, and that the knapsack instance admits a feasible solution other than the empty set.

Let $U_{\mathrm{max}} = \sum_{j \in \mathcal{B}} u_j$ and $\delta = {1} /{m U_{\mathrm{max}}} > 0$, and construct an instance of Ellis’s problem with $\mathcal{C} = \mathcal{B}$, $H = W$, all $f_j = \delta$, and $t_j = u_j / \delta$ for all $j$. Clearly, $\mathcal{X} \subseteq \mathcal{C}$ is feasible for Ellis’s problem if and only if it is feasible for the knapsack instance. Now, we observe that if $\mathcal{X}$ is nonempty,
\begin{align}
\begin{split}
\sum_{j \in \mathcal{X}} u_j &=  \sum_{j \in \mathcal{X}} t_j f_j 
> \sum_{j \in \mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - f_{j’}) \Bigr)
%= \sum_{j \in \mathcal{X}} \Bigl( u_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - \delta) \Bigr)
= v(\mathcal{X}) \\
&= \sum_{j \in \mathcal{X}} \Bigl( u_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - \delta) \Bigr)
\geq (1 - \delta)^m \sum_{j \in \mathcal{X}} u_j \\
&\geq (1 - m\delta) \sum_{j \in \mathcal{X}} u_j 
\geq \sum_{j \in \mathcal{X}} u_j  - m\delta U_{\mathrm{max}}
= \sum_{j \in \mathcal{X}} u_j  - 1
\end{split}
\end{align}
This means that given a nonempty portfolio $\mathcal{X}$, its value for the corresponding knapsack instance is the smallest integer greater than $v( \mathcal{X})$. That is, $\sum_{j \in \mathcal{X}} u_j \geq U$ if and only if $v(\mathcal{X}) \geq U -1$. Taking $V = U-1$ completes the transformation and concludes the proof.  \end{proof}

An intuitive extension of the greedy algorithm for Alma's problem is to iteratively add to $\mathcal{X}$ the school $k$ for which $[ v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) ] / g_k$ is largest. However, this algorithm offers no guaranteed approximation ratio. 
% as the following example suggests ...The construction above shows that the objective function of Ellis's problem can approximate that of a knapsack problem with arbitrary precision, and this algorithm

\subsection{Pseudopolynomial-time dynamic programs}

\subsubsection{Dynamic program based on application costs}
We will provide an algorithmic solution to Ellis's problem that runs in $O(Hm + m\log m)$ time and $O(Hm)$ space. The algorithm resembles the dynamic programming algorithm for the binary knapsack problem. Because we cannot assume that $H \leq m$ (as was the case in Alma's problem), this represents a pseudopolynomial-time solution (Dantzig 1957; Garey and Johnson 1979; \emph{Wikipedia}, s.v. ``Knapsack problem'').

With minimal loss of generality, we assume that $g_j \in \mathbb{N}$ for $j = 1\dots m$ and $H \in\mathbb{N}$. For $j = 0 \dots m$ and $h = 0 \dots H$, let $\mathcal{X}[j, h]$ denote the optimal portfolio using only the schools $\{ 1, \dots, j\}$ and costing no more than $h$, and let $V[j,h] = v(\mathcal{X}[j, h])$.  It is clear that if $j=0$ or $h=0$, then $\mathcal{X}[j, h] = \O$ and $V[j, h] = 0$.  For convenience, we also define $V[j, h] = -\infty$ for all $h < 0$.

For the remaining indices, $\mathcal{X}[j, h]$ either contains $j$ or not. If it does not contain $j$, then $\mathcal{X}[j, h] = \mathcal{X}[j-1, h]$. On the other hand, if  $\mathcal{X}[j, h]$ contains $j$, then its value is $(1 - f_j) v(\mathcal{X}[j, h]\setminus \{j\}) + f_j t_j$. This requires that $\mathcal{X}[j, h]\setminus \{j\}$ make optimal use of the remaining budget over the remaining schools; that is, $\mathcal{X}[j, h] = \mathcal{X}[j-1, h - g_j] \cup\{j\}$. From these observations, we obtain the following Bellman equation for $ j = 1\dots m$ and $h = 1\dots H$:
\begin{align}
V[j, h] = \max\bigl\{ V[j-1, h], (1 - f_j) V[j-1, h-g_j] + f_j t_j \bigr\}
\end{align}
with the convention that $ -\infty \cdot 0 = -\infty$. The corresponding optimal portfolios can be computed by observing that $\mathcal{X}[j, h]$ contains $j$ if and only if $V[j, h]> V[j-1, h]$. The optimal solution is given by $\mathcal{X}[m, H]$. The algorithm below performs these computations and outputs the optimal portfolio $\mathcal{X}$.

\begin{algorithm}[H] 
%\DontPrintSemicolon
\caption{Dynamic program for Ellis's problem with integral application costs.} \label{ellisDP1}
\KwData{Utility values $t \in[0, \infty)^m$, admissions probabilities $f \in [0, 1]^m$, application costs $g \in \mathbb{N}^m$, budget $H \in\mathbb{N}$.}
Index schools in ascending order by $t$\;
\SetKwProg{Fn}{function}{ do}{end}
\Fn{$V(j, h)$}{
	\lIf{$h < g_j$}{\Return{$-\infty$}}
	\lElseIf{$j = 0$ or $h=0$}{\Return{$0$}}
	\lElse{\Return{$\max\bigl\{ V(j-1, h), (1 - f_j) V(j-1, h-g_j) + f_j t_j \bigr\}$}}
}
%\lFor{$j=1\dots m$}
%{$\operatorname{push!}\bigl(\mathtt{H}, (j, f_j, t_j)\bigr)$}
%\For{$i=1\dots h$}
%{
%    $\mathtt{D} \gets \operatorname{pop!}(\mathtt{H})$\;
%    $\operatorname{append!}(\mathtt{X}, \mathtt{D.j})$\;
%     \lIfElse{$i=1$}{$\operatorname{append!}(\mathtt{V}, \mathtt{D.f}\cdot\mathtt{D.t})$}
%     {$\operatorname{append!}(\mathtt{V}, \mathtt{V[i-1]} + \mathtt{D.f}\cdot\mathtt{D.t})$}
%    \For{$\mathtt{E} \in \mathtt{H}$}
%	{
%	\lIfElse{$\mathtt{E.j} < \mathtt{D.j}$}{$\mathtt{E.t} \gets  \mathtt{E.t} \cdot(1 -  \mathtt{D.f})$}{$\mathtt{E.t} \gets  \mathtt{E.t} -  \mathtt{D.f}\cdot\mathtt{D.t}$}
%	}
%}
$h \gets H$\;
$\mathcal{X} \gets \O$\;
\For{$j = m, m-1, \dots, 1$}{
	\If{$V(j-1, h) < V(j, h)$}{
		$\mathcal{X} \gets \mathcal{X}\cup\{j\}$\; 
		$h \gets h - g_j$\;
	}
}
\Return{$\mathcal{X}$}
\end{algorithm}

\begin{theorem}[Validity of Algorithm \ref{ellisDP1}]
Algorithm \ref{ellisDP1} produces an optimal application portfolio for Ellis's problem in $O(H m + m \log m)$ time and $O(H m)$ space.
\end{theorem}

\begin{proof}
Optimality follows from the foregoing discussion. Sorting by $t_j$ is $O(m \log m)$.  The number of indices at which $V(j, h)$ is nontrivially defined is $O(Hm)$. To prevent $V(j, h)$ from being evaluated more than once at a given index, its values can be stored in a dictionary as they are computed.
\end{proof}

\subsubsection{Dynamic program based on portfolio valuations}
As with the knapsack problem, Ellis's problem admits a complementary dynamic program that iterates on the value of the cheapest portfolio instead of on the cost of the most valuable portfolio.

With minimal loss of generality, we assume that all $t_j\in \mathbb{N}$ and $f_j \in (0, 1] \cap \mathbb{Q}$. Let $D$ denote the product of the denominators of the $f_j$-values. Since $v(\mathcal{X})$ is linear in $t_j$, by rescaling the $t_j$-values, we may assume that each $t_j$ is divisible by $D$. Therefore, for any portfolio $\mathcal{X}\subseteq\mathcal{X}$, $v(\mathcal{X})$ is a natural number between zero and $t_m$. %Mention size of t_m

For integers $0 \leq j \leq m$ and $v \leq t_m$, let $\mathcal{W}[j, v]$ denote the least expensive portfolio using only the schools $\{ 1, \dots, j\}$ and having value at least $v$, if such a portfolio exists. Let $G[j, v] = v(\mathcal{W}[j, v])$ denote the corresponding value, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist. It is clear that if $v \leq 0$, then $\mathcal{W}[j, v] = \O$ and $G[j, h] = 0$. If $j = 0$ and $v > 0$, then $G[j, h] = \infty$.  For the remaining indices (where $j, v > 0$), we claim that
\begin{align} \label{recursionrelationforcostmindp}
G[j, v] &=
\begin{cases}
\infty, \quad & t_j < v \\
\min\bigl\{G[j-1, v], g_j + G[j-1, v - \Delta] \bigr\}, \quad & t_j \geq v 
%\begin{cases}
%\min\bigl\{G[j-1, v], g_j + G[j-1, v - \Delta] \bigr\}, \quad &f_j < 1 \\
%\min\bigl\{G[j-1, v], g_j \bigr\}, \quad &f_j = 1 \text{ and } f_j t_j \geq v\\
%G[j-1, v], \quad &f_j = 1 \text{ and } f_j t_j < v
\end{cases}\\
\text{where}\qquad
\Delta &= 
\begin{cases}
\lfloor \frac{f_j}{1 - f_j} (t_j - v)\rfloor, \quad & f_j < 1\\
\infty, &f_j = 1
\end{cases}
\end{align}
In the $t_j < v$ case, any feasible portfolio must be composed of schools with utility less than $v$, and therefore its valuation can not equal $v$, meaning that $\mathcal{W}[j, v]$ is undefined. In the $t_j \geq v$ case, the first argument to $\min\{\}$ says simply that omitting $j$ and choosing $\mathcal{W}[j-1, v]$ is a feasible choice for $\mathcal{W}[j, v]$. If, on the other hand, $j \in \mathcal{W}[j, v]$, then
\begin{equation}v(\mathcal{W}[j, v]) = (1 - f_j )v(\mathcal{W}[j, v]\setminus \{j\}) + f_j t_j\end{equation}
Therefore, the subportfolio $\mathcal{W}[j, v]\setminus \{j\}$ must have a value of at least $v - \Delta$, where $\Delta$ satisfies $v = (1 - f_j )(v - \Delta) + f_j t_j $. When $f_j < 1$, the solution to this equation is $ \Delta = \frac{f_j}{1 - f_j} (t_j - v)$. This value may not be an integer; however, because $v(\mathcal{X})$ can take only integer values, if some portfolio has $v(\mathcal{X})\geq v - \Delta$, then $v(\mathcal{X})\geq \lceil v - \Delta \rceil$ as well, which allows us to apply the floor operator to $\Delta$. When $f_j = 1$, then the singleton $\{j\}$ has $v(\{j\}) \geq v$, so
\begin{equation}G[j, v] = \min\bigl\{G[j-1, v], g_j \bigr\}\end{equation}
Defining $\Delta = \infty$ in this case ensures that $G[j-1, v-\Delta] = 0$ as required.
%we have $t_j < v$, and since no portfolio can be more valueable than the largest $t_j$ value of its elements, we must have $G[j, v] =\infty$. Defining $\Delta$ as zero in this guess automatically chooses $G[j, v] = G[j-1, v] = \infty$ for the same reason.
%
%In the $f_j t_j \geq v$ case, the singleton $\{j\}$ is feasible for $j$ and $v$, so $\mathcal{W}[j, v]$ is simply the cheaper of $\{j\}$ and $.\mathcal{W}[j-1, v]$.

Once $G[j, v]$ has been calculated at each index, the corresponding portfolio can be found by applying the observation that $
\mathcal{W}[j, v]$ contains $j$ if and only if $G[j, v] < G[j-1, v]$. Then solution to Ellis's problem is obtained by applying this observation to the optimal objective value $\max\{ w: G[m, w] \leq H\}$.

\begin{algorithm}[H] 
%\DontPrintSemicolon
\caption{Dynamic program for Ellis's problem with integral utility parameters and rational admissions probabilities.} \label{ellisDP2}
\KwData{Utility values $t \in \mathbb{N}^m$, admissions probabilities $f \in \bigl((0, 1] \cap \mathbb{Q}\bigr)^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$.}
Index schools in ascending order by $t$\;
Multiply each $t_j$ by $D$, the product of the denominators of the $f_j$\;
\SetKwProg{Fn}{function}{ do}{end}
\Fn{$G(j, v)$}{
	\lIf{$v \leq 0$}{\Return{$0$}}
	\lElseIf{$j=0$ or $t_j < v$}{\Return{$\infty$}}
	\Else {\Return{$\min\bigl\{ G(j-1, v), g_j + G(j-1, v -  \Delta) \bigr\}$}}
}
%	\uIf{$f_j < 1$}{\Return{$\min\bigl\{ G(j-1, v), g_j + G(j-1, v -  \lfloor \frac{f_j}{1 - f_j} (t_j - v)\rfloor ) \bigr\}$}}
%	\uElseIf{$f_j = 1$ and $f_j t_j \geq v$}{\Return{$\min\bigl\{G[j-1, v], g_j \bigr\}$}}
%	\lElse{\Return{$G(j-1, v)$}}
%	}
%\lFor{$j=1\dots m$}
%{$\operatorname{push!}\bigl(\mathtt{H}, (j, f_j, t_j)\bigr)$}
%\For{$i=1\dots h$}
%{
%    $\mathtt{D} \gets \operatorname{pop!}(\mathtt{H})$\;
%    $\operatorname{append!}(\mathtt{X}, \mathtt{D.j})$\;
%     \lIfElse{$i=1$}{$\operatorname{append!}(\mathtt{V}, \mathtt{D.f}\cdot\mathtt{D.t})$}
%     {$\operatorname{append!}(\mathtt{V}, \mathtt{V[i-1]} + \mathtt{D.f}\cdot\mathtt{D.t})$}
%    \For{$\mathtt{E} \in \mathtt{H}$}
%	{
%	\lIfElse{$\mathtt{E.j} < \mathtt{D.j}$}{$\mathtt{E.t} \gets  \mathtt{E.t} \cdot(1 -  \mathtt{D.f})$}{$\mathtt{E.t} \gets  \mathtt{E.t} -  \mathtt{D.f}\cdot\mathtt{D.t}$}
%	}
%}
$v \gets  \max\{ w: G(m, w) \leq H\}$\;
$\mathcal{X} \gets \O$\;
\For{$j = m, m-1, \dots, 1$}{
	\If{$G(j, v)< \infty$ and $G(j, v) < G(j-1, v)$}{
		$\mathcal{X} \gets \mathcal{X}\cup\{j\}$\; 
		$v \gets v - \Delta$\;
	}
}
\Return{$\mathcal{X}$}
\end{algorithm}

\begin{theorem}[Validity of Algorithm \ref{ellisDP2}]
Algorithm \ref{ellisDP2} produces an optimal application portfolio for Ellis's problem in $O(D m + m \log m)$ time and $O(D m)$ space, where $D$ is the product of the denominators of the $f_j$-values.
\end{theorem}

\begin{proof}
Optimality follows from the foregoing discussion. The computational complexity follows by symmetry with Algorithm \ref{ellisDP1}. 
\end{proof}

\subsection{A polynomial-time approximation scheme}
By rounding the $f_j$ values down to the nearest multiple of $1/K$, where $K \in \mathbb{N}$, Algorithm \ref{ellisDP2} yields a polynomial-time approximation scheme for Ellis's problem. Here we assume only that is each $t_j$ is integral.

\begin{algorithm}[H] 
%\DontPrintSemicolon
\caption{Polynomial-time approximation scheme for Ellis's problem.} \label{ellisPTAS}
\KwData{Utility values $t \in \mathbb{N}^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$, tolerance $\varepsilon \in (0, 1)$.}
$K \gets \bigl\lceil\,{m}/{\sqrt[m]{\varepsilon f_m}}\,\bigr\rceil$\;
Round each $f_j$ down to the nearest multiple of $1/K$\;
\Return{} the output of Algorithm \ref{ellisDP2}
\end{algorithm}

\begin{proof}
The initialization steps trivial. By rounding, we ensure that $D \leq K^m$; therefore, the time required by Algorithm \ref{ellisDP2} is
\begin{equation}
O\left(Dm + m\log m \right) = 
\end{equation}
\end{proof}

\pagebreak
\ifEN \section{References} \else \section{참고문헌} \fi
\noindent

\parskip 0em
\leftskip 2em
\parindent -2em
Budish, Eric. 2011. ``The Combinatorial Assignment Problem: Approximate Competitive Equilibrium from Equal Incomes.'' \emph{Journal of Political Economy} 119 (6): 1061--1103. \url{https://doi.org/10.1086/664613}.

Dantzig, George B. 1957. ``Discrete-Variable Extremum Problems.'' \emph{Operations Research} 5 (2): 266--88.

Fisher, Marshall, George Nemhauser, and Laurence Wolsey. 1978. ``An analysis of approximations for maximizing submodular set functions—I.'' \emph{Mathematical Programming} 14: 265--94.

Fredman, Michael Lawrence and Robert Tarjan. 1987. ``Fibonacci heaps and their uses in improved network optimization algorithms.'' \emph{Journal of the Association for Computing Machinery} 34 (3): 596--615.

Fu, Chao. 2014. ``Equilibrium Tuition, Applications, Admissions, and Enrollment in the College Market.'' \emph{Journal of Political Economy} 122 (2): 225--81. \url{https://doi.org/10.1086/675503}.

Garey, Michael and David Johnson. 1979. \emph{Computers and Intractability: A Guide to the Theory of NP-Completeness.} New York: W. H. Freeman and Company.

Othman, Abraham, Eric Budish, and Tuomas Sandholm. 2010. ``Finding Approximate Competitive Equilibria: Efficient and Fair
Course Allocation.'' In \emph{Proceedings of 9th International Conference on Autonomous Agents and Multiagent Systems.} New York: ACM. \url{https://dl.acm.org/doi/abs/10.5555/1838206.1838323}.

Rozanov, Mark and Arie Tamir. 2020. ``The nestedness property of the convex ordered median location problem on a tree.'' \emph{Discrete Optimization} 36: 100581. \url{https://doi.org/10.1016/j.disopt.2020.100581}.

% p. 70: FPTAS for knapsack, should work for Ellis's problem
Vazirani, Vijay. 2001. \emph{Approximation Algorithms.} Berlin: Springer.

\end{document}

### 2022-02-25 paper.tex
towards FPTAS

@@
 When each variable's cost, or each college's application fee, is identical, we show that the optimal portfolios are nested in the budget constraint, yielding an exact polynomial-time algorithm.
-When colleges differ in their application fees, the problem is NP-complete.
-For this more general problem, we provide two dynamic programs that produce exact solutions in pseudopolynomial time and a polynomial-time approximation scheme.
+When colleges differ in their application fees, we show that the problem is NP-complete.
+We provide two dynamic programs for this more general problem.
+The first produces an exact solution in pseudopolynomial time, and the second yields a fully polynomial-time approximation scheme.
 \pagebreak
@@
 Applying Theorem \ref{assumetzerozero}, we assume $t_0 = 0$ and disregard $c_0$ throughout.
+We also assume with trivial loss of generality that all $g_j \leq H$.
 
@@
 
-\subsection{Pseudopolynomial-time dynamic programs}
-
-\subsubsection{Dynamic program based on application costs}
-We will provide an algorithmic solution to Ellis's problem that runs in $O(Hm + m\log m)$ time and $O(Hm)$ space.
+\subsection{Pseudopolynomial-time dynamic program}
+In this subsection, we provide an algorithmic solution to Ellis's problem that runs in $O(Hm + m\log m)$ time and $O(Hm)$ space.
 The algorithm resembles the dynamic programming algorithm for the binary knapsack problem.
@@
 
-\subsubsection{Dynamic program based on portfolio valuations}
-As with the knapsack problem, Ellis's problem admits a complementary dynamic program that iterates on the value of the cheapest portfolio instead of on the cost of the most valuable portfolio.
-
-With minimal loss of generality, we assume that all $t_j\in \mathbb{N}$ and $f_j \in (0, 1] \cap \mathbb{Q}$.
+%\subsubsection{Dynamic program based on portfolio valuations}
+%
+%With minimal loss of generality, we assume that all $t_j\in \mathbb{N}$ and $f_j \in (0, 1] \cap \mathbb{Q}$.
 Let $D$ denote the product of the denominators of the $f_j$-values.
@@
 %Mention size of t_m
+%
+%For integers $0 \leq j \leq m$ and $v \leq t_m$, let $\mathcal{W}[j, v]$ denote the least expensive portfolio using only the schools $\{ 1, \dots, j\}$ and having value at least $v$, if such a portfolio exists.
+Let $G[j, v] = v(\mathcal{W}[j, v])$ denote the corresponding value, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist.
+It is clear that if $v \leq 0$, then $\mathcal{W}[j, v] = \O$ and $G[j, h] = 0$.
+If $j = 0$ and $v > 0$, then $G[j, h] = \infty$.
+For the remaining indices (where $j, v > 0$), we claim that
+%\begin{align} \label{recursionrelationforcostmindp}
+%G[j, v] &=
+%\begin{cases}
+%\infty, \quad & t_j < v \\
+%\min\bigl\{G[j-1, v], g_j + G[j-1, v - \Delta] \bigr\}, \quad & t_j \geq v 
+%%\begin{cases}
+%%\min\bigl\{G[j-1, v], g_j + G[j-1, v - \Delta] \bigr\}, \quad &f_j < 1 \\
+%%\min\bigl\{G[j-1, v], g_j \bigr\}, \quad &f_j = 1 \text{ and } f_j t_j \geq v\\
+%%G[j-1, v], \quad &f_j = 1 \text{ and } f_j t_j < v
+%\end{cases}\\
+%\text{where}\qquad
+%\Delta &= 
+%\begin{cases}
+%\lfloor \frac{f_j}{1 - f_j} (t_j - v)\rfloor, \quad & f_j < 1\\
+%\infty, &f_j = 1
+%\end{cases}
+%\end{align}
+%In the $t_j < v$ case, any feasible portfolio must be composed of schools with utility less than $v$, and therefore its valuation can not equal $v$, meaning that $\mathcal{W}[j, v]$ is undefined.
+In the $t_j \geq v$ case, the first argument to $\min\{\}$ says simply that omitting $j$ and choosing $\mathcal{W}[j-1, v]$ is a feasible choice for $\mathcal{W}[j, v]$.
+If, on the other hand, $j \in \mathcal{W}[j, v]$, then
+%\begin{equation}v(\mathcal{W}[j, v]) = (1 - f_j )v(\mathcal{W}[j, v]\setminus \{j\}) + f_j t_j\end{equation}
+%Therefore, the subportfolio $\mathcal{W}[j, v]\setminus \{j\}$ must have a value of at least $v - \Delta$, where $\Delta$ satisfies $v = (1 - f_j )(v - \Delta) + f_j t_j $.
+When $f_j < 1$, the solution to this equation is $ \Delta = \frac{f_j}{1 - f_j} (t_j - v)$.
+This value may not be an integer; however, because $v(\mathcal{X})$ can take only integer values, if some portfolio has $v(\mathcal{X})\geq v - \Delta$, then $v(\mathcal{X})\geq \lceil v - \Delta \rceil$ as well, which allows us to apply the floor operator to $\Delta$.
+When $f_j = 1$, then the singleton $\{j\}$ has $v(\{j\}) \geq v$, so
+%\begin{equation}G[j, v] = \min\bigl\{G[j-1, v], g_j \bigr\}\end{equation}
+%Defining $\Delta = \infty$ in this case ensures that $G[j-1, v-\Delta] = 0$ as required.
+%%we have $t_j < v$, and since no portfolio can be more valueable than the largest $t_j$ value of its elements, we must have $G[j, v] =\infty$.
+Defining $\Delta$ as zero in this guess automatically chooses $G[j, v] = G[j-1, v] = \infty$ for the same reason.
+%%
+%%In the $f_j t_j \geq v$ case, the singleton $\{j\}$ is feasible for $j$ and $v$, so $\mathcal{W}[j, v]$ is simply the cheaper of $\{j\}$ and $.\mathcal{W}[j-1, v]$.
+%
+%Once $G[j, v]$ has been calculated at each index, the corresponding portfolio can be found by applying the observation that $
+%\mathcal{W}[j, v]$ contains $j$ if and only if $G[j, v] < G[j-1, v]$.
+Then solution to Ellis's problem is obtained by applying this observation to the optimal objective value $\max\{ w: G[m, w] \leq H\}$.
+%
+%
+%
+%\begin{algorithm}[H] 
+%%\DontPrintSemicolon
+%\caption{Dynamic program for Ellis's problem with integral utility parameters and rational admissions probabilities.} \label{ellisDP2}
+%\KwData{Utility values $t \in \mathbb{N}^m$, admissions probabilities $f \in \bigl((0, 1] \cap \mathbb{Q}\bigr)^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$.}
+%Index schools in ascending order by $t$\;
+%Multiply each $t_j$ by $D$, the product of the denominators of the $f_j$\;
+%\SetKwProg{Fn}{function}{ do}{end}
+%\Fn{$G(j, v)$}{
+%	\lIf{$v \leq 0$}{\Return{$0$}}
+%	\lElseIf{$j=0$ or $t_j < v$}{\Return{$\infty$}}
+%	\Else {\Return{$\min\bigl\{ G(j-1, v), g_j + G(j-1, v -  \Delta) \bigr\}$}}
+%}
+%%	\uIf{$f_j < 1$}{\Return{$\min\bigl\{ G(j-1, v), g_j + G(j-1, v -  \lfloor \frac{f_j}{1 - f_j} (t_j - v)\rfloor ) \bigr\}$}}
+%%	\uElseIf{$f_j = 1$ and $f_j t_j \geq v$}{\Return{$\min\bigl\{G[j-1, v], g_j \bigr\}$}}
+%%	\lElse{\Return{$G(j-1, v)$}}
+%%	}
+%%\lFor{$j=1\dots m$}
+%%{$\operatorname{push!}\bigl(\mathtt{H}, (j, f_j, t_j)\bigr)$}
+%%\For{$i=1\dots h$}
+%%{
+%%    $\mathtt{D} \gets \operatorname{pop!}(\mathtt{H})$\;
+%%    $\operatorname{append!}(\mathtt{X}, \mathtt{D.j})$\;
+%%     \lIfElse{$i=1$}{$\operatorname{append!}(\mathtt{V}, \mathtt{D.f}\cdot\mathtt{D.t})$}
+%%     {$\operatorname{append!}(\mathtt{V}, \mathtt{V[i-1]} + \mathtt{D.f}\cdot\mathtt{D.t})$}
+%%    \For{$\mathtt{E} \in \mathtt{H}$}
+%%	{
+%%	\lIfElse{$\mathtt{E.j} < \mathtt{D.j}$}{$\mathtt{E.t} \gets  \mathtt{E.t} \cdot(1 -  \mathtt{D.f})$}{$\mathtt{E.t} \gets  \mathtt{E.t} -  \mathtt{D.f}\cdot\mathtt{D.t}$}
+%%	}
+%%}
+%$v \gets  \max\{ w: G(m, w) \leq H\}$\;
+%$\mathcal{X} \gets \O$\;
+%\For{$j = m, m-1, \dots, 1$}{
+%	\If{$G(j, v)< \infty$ and $G(j, v) < G(j-1, v)$}{
+%		$\mathcal{X} \gets \mathcal{X}\cup\{j\}$\; 
+%		$v \gets v - \Delta$\;
+%	}
+%}
+%\Return{$\mathcal{X}$}
+%\end{algorithm}
+%
+%\begin{theorem}[Validity of Algorithm \ref{ellisDP2}]
+%Algorithm \ref{ellisDP2} produces an optimal application portfolio for Ellis's problem in $O(D m + m \log m)$ time and $O(D m)$ space, where $D$ is the product of the denominators of the $f_j$-values.
+%\end{theorem}
+%
+%\begin{proof}
+%Optimality follows from the foregoing discussion. The computational complexity follows by symmetry with Algorithm \ref{ellisDP1}. 
+%\end{proof}
+
+\subsection{Fully polynomial-time approximation scheme}
+As with the knapsack problem, Ellis's problem admits a complementary dynamic program that iterates on the value of the cheapest portfolio instead of on the cost of the most valuable portfolio.
+We will use this algorithm as the basis for a fully polynomial-time approximation scheme for Ellis's problem.
+%Now, we modify Algorithm \ref{ellisDP2} into a fully polynomial-time approximation scheme for Ellis's problem.
+Here we assume only that is each $t_j$ is integral.
+We will represent the approximate value of each portfolio using a fixed-point decimal with a precision of $P$, where $P$ is the number of digits to retain after the decimal point.
+Since $t_m$ is a trivial upper bound on the value of any portfolio, the set of possible portfolio valuations is now finite: $\mathcal{V} = \{0, 1\times 10^{-P}, 2\times 10^{-P}, \dots, t_m - 1\times 10^{-P}, t_m\}$.
+Its cardinality is $|\mathcal{V} | = P t_m$.
+Let $r[x] = \lfloor 10^P x \rfloor 10^{-P}$ denote the value of $x$ rounded down to its nearest fixed-point representation.
 
 For integers $0 \leq j \leq m$ and $v \leq t_m$, let $\mathcal{W}[j, v]$ denote the least expensive portfolio using only the schools $\{ 1, \dots, j\}$ and having value at least $v$, if such a portfolio exists.
-Let $G[j, v] = v(\mathcal{W}[j, v])$ denote the corresponding value, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist.
+Let $G[j, v] = r\bigl[v(\mathcal{W}[j, v])\bigr]$ denote the fixed-point representation of the corresponding value, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist.
 It is clear that if $v \leq 0$, then $\mathcal{W}[j, v] = \O$ and $G[j, h] = 0$.
@@
 \text{where}\qquad
-\Delta &= 
+\Delta_j (v) &= 
 \begin{cases}
-\lfloor \frac{f_j}{1 - f_j} (t_j - v)\rfloor, \quad & f_j < 1\\
+r\left[\frac{f_j}{1 - f_j} (t_j - v)\right], \quad & f_j < 1\\
 \infty, &f_j = 1
-\end{cases}
+\end{cases} \label{deltajvdef}
 \end{align}
@@
 If, on the other hand, $j \in \mathcal{W}[j, v]$, then
-\begin{equation}v(\mathcal{W}[j, v]) = (1 - f_j )v(\mathcal{W}[j, v]\setminus \{j\}) + f_j t_j\end{equation}
+\begin{equation} \label{solvemeforvwjvminusj}
+v(\mathcal{W}[j, v]) = (1 - f_j )v(\mathcal{W}[j, v]\setminus \{j\}) + f_j t_j\end{equation}
 Therefore, the subportfolio $\mathcal{W}[j, v]\setminus \{j\}$ must have a value of at least $v - \Delta$, where $\Delta$ satisfies $v = (1 - f_j )(v - \Delta) + f_j t_j $.
 When $f_j < 1$, the solution to this equation is $ \Delta = \frac{f_j}{1 - f_j} (t_j - v)$.
-This value may not be an integer; however, because $v(\mathcal{X})$ can take only integer values, if some portfolio has $v(\mathcal{X})\geq v - \Delta$, then $v(\mathcal{X})\geq \lceil v - \Delta \rceil$ as well, which allows us to apply the floor operator to $\Delta$.
+By rounding this value down, we ensure that the true valuation of $\mathcal{W}[j, v]$ is \emph{at least} $v - \Delta$.
+
 When $f_j = 1$, then the singleton $\{j\}$ has $v(\{j\}) \geq v$, so
@@
 \mathcal{W}[j, v]$ contains $j$ if and only if $G[j, v] < G[j-1, v]$.
-Then solution to Ellis's problem is obtained by applying this observation to the optimal objective value $\max\{ w: G[m, w] \leq H\}$.
+Then an approximate solution to Ellis's problem is obtained by applying this observation to the approximate optimal objective value $\max\{ w: G[m, w] \leq H\}$.
 
@@
 %\DontPrintSemicolon
-\caption{Dynamic program for Ellis's problem with integral utility parameters and rational admissions probabilities.} \label{ellisDP2}
-\KwData{Utility values $t \in \mathbb{N}^m$, admissions probabilities $f \in \bigl((0, 1] \cap \mathbb{Q}\bigr)^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$.}
+\caption{Polynomial-time approximation scheme for Ellis's problem.} \label{ellisDP3}
+\KwData{Utility values $t \in \mathbb{N}^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$, tolerance $\varepsilon \in (0, 1)$.}
 Index schools in ascending order by $t$\;
-Multiply each $t_j$ by $D$, the product of the denominators of the $f_j$\;
+$P \gets \bigl\lceil\log_{10}\left(m / \varepsilon f_m t_m\right)\bigr\rceil$\;
 \SetKwProg{Fn}{function}{ do}{end}
@@
 	\lElseIf{$j=0$ or $t_j < v$}{\Return{$\infty$}}
-	\Else {\Return{$\min\bigl\{ G(j-1, v), g_j + G(j-1, v -  \Delta) \bigr\}$}}
+	\Else {\Return{$\min\Bigl\{ G(j-1, v), g_j + G\bigl(j-1, v -  \Delta_j(v)\bigr) \Bigr\}$}}
 }
@@
 %}
-$v \gets  \max\{ w: G(m, w) \leq H\}$\;
+$v^* \gets  \max\{ w: G(m, w) \leq H\}$\;
+$v \gets v^*$\;
 $\mathcal{X} \gets \O$\;
@@
 		$\mathcal{X} \gets \mathcal{X}\cup\{j\}$\; 
-		$v \gets v - \Delta$\;
+		$v \gets v -  \Delta_j(v)$\;
 	}
@@
 
-\begin{theorem}[Validity of Algorithm \ref{ellisDP2}]
-Algorithm \ref{ellisDP2} produces an optimal application portfolio for Ellis's problem in $O(D m + m \log m)$ time and $O(D m)$ space, where $D$ is the product of the denominators of the $f_j$-values.
+\begin{theorem}[Validity of Algorithm \ref{ellisDP3}]
+Algorithm \ref{ellisDP3} produces a $(1 - \varepsilon)$-optimal application portfolio for Ellis's problem.% in $O(D m + m \log m)$ time and $O(D m)$ space, where $D$ is the product of the denominators of the $f_j$-values.
 \end{theorem}
@@
 \begin{proof}
-Optimality follows from the foregoing discussion. The computational complexity follows by symmetry with Algorithm \ref{ellisDP1}. 
+%Let $\mathcal{W}$ denote the output of Algorithm \ref{ellisDP3} and $\mathcal{X}$ the true optimum. In the main loop, each time the target valuation $v$ is updated using
+Because $\Delta_j(v)$ is rounded down, if $j \in \mathcal{W}[j, v]$, then the true value of $(1 - f_j) v\bigl(\mathcal{W}[j-1, v- \Delta_j(v)]\bigr) + f_j t_j$ may exceed the fixed-point value $v$ of $\mathcal{W}[j, v]$, but not by more than $10^{-P}$.
+This error accumulates with each school added to $\mathcal{W}$, but the number of additions is at most $m$.
+Therefore, where $v^*(\mathcal{W})$ denotes the fixed-point valuation of $\mathcal{W}$ as recorded in the algorithm,
+\begin{equation}
+v(\mathcal{W}) - v^*(\mathcal{W}) \leq m 10^{-P} 
+\end{equation}
+We can define $v^*(\mathcal{X})$ analogously as the fixed-point valuation of $\mathcal{X}$ when its elements are added in index order and its value is updated and rounded down at each addition in accordance with \eqref{solvemeforvwjvminusj}.
+For the same reason as above,
+$v(\mathcal{X}) - v^*(\mathcal{X}) \leq m 10^{-P}$.
+
+Finally, by the trivial assumption that all $g_j \leq H$, we know $v(\mathcal{X}) \geq v(\{m\}) = f_m t_m$. From this we obtain
+\begin{equation}
+\begin{split}
+v(\mathcal{W}) &\geq v^*(\mathcal{W}) \geq v^*(\mathcal{X})
+\geq v(\mathcal{X}) - m 10^{-P} \\
+&\geq \left(1 - \frac{m 10^{-P}}{f_m t_m}\right) v(\mathcal{X})
+\geq \left(1 - \varepsilon\right) v(\mathcal{X})
+\end{split}
+\end{equation}
+which completes the proof. 
 \end{proof}
 
-\subsection{A polynomial-time approximation scheme}
-By rounding the $f_j$ values down to the nearest multiple of $1/K$, where $K \in \mathbb{N}$, Algorithm \ref{ellisDP2} yields a polynomial-time approximation scheme for Ellis's problem.
-Here we assume only that is each $t_j$ is integral.
+\begin{theorem}[Computational complexity of Algorithm \ref{ellisDP3}]
+Algorithm \ref{ellisDP3} uses $O(m /\varepsilon + m \log m)$ time and $O(m/\varepsilon)$ space, where $D$ is the product of the denominators of the $f_j$-values.
+\end{theorem}
 
-\begin{algorithm}[H] 
-%\DontPrintSemicolon
-\caption{Polynomial-time approximation scheme for Ellis's problem.} \label{ellisPTAS}
-\KwData{Utility values $t \in \mathbb{N}^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$, tolerance $\varepsilon \in (0, 1)$.}
-$K \gets \bigl\lceil\,{m}/{\sqrt[m]{\varepsilon f_m}}\,\bigr\rceil$\;
-Round each $f_j$ down to the nearest multiple of $1/K$\;
-\Return{} the output of Algorithm \ref{ellisDP2}
-\end{algorithm}
+% To make this fully polynomial, need to get rid of the f_m t_m business ...
+\begin{proof}
+In the worst case, the number of indices at which $G(j, v)$ is evaluated is $O(|\mathcal{V}| m)$. Since
+\begin{equation}
+|\mathcal{V}| = t_m 10^{P} = t_m10^ { \bigl\lceil\log_{10}\left(m / \varepsilon f_m t_m\right)\bigr\rceil}
+%\leq {\text{const.}}/{\varepsilon}
+\end{equation}
+is $O(1/ \varepsilon)$, adding the $O(m \log m)$-time required to sort $t$ yields the result.
+By recording the values of $G(j, v)$ in a dictionary as they are computed, we prevent redundant recursive calls at a cost of $O(|\mathcal{V}| m) = O(m / \varepsilon)$ space.
+\end{proof}
 
-\begin{proof}
-The initialization steps trivial. By rounding, we ensure that $D \leq K^m$; therefore, the time required by Algorithm \ref{ellisDP2} is
-\begin{equation}
-O\left(Dm + m\log m \right) = 
-\end{equation}
-\end{proof}
+Since these bounds are polynomial in $m$ and $1 / \varepsilon$, Algorithm \ref{ellisDP3} is a fully polynomial-time approximation scheme for Ellis's problem.
 

### 2022-02-25 paper.tex
fixed FPTAS bound

@@
 \urlstyle{same}
-\usepackage[ruled]{algorithm2e}
+\usepackage[ruled,linesnumbered]{algorithm2e}
 \newcommand{\lIfElse}[3]{\lIf{#1}{#2 \textbf{else}~#3}}
@@
 An intuitive extension of the greedy algorithm for Alma's problem is to iteratively add to $\mathcal{X}$ the school $k$ for which $[ v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) ] / g_k$ is largest.
-However, this algorithm offers no guaranteed approximation ratio.
+However, this algorithm offers no guaranteed approximation ratio, as the following example shows.
+\begin{example}
+Let $t = (2, 2022)$, $f = (1, 1)$, $g = (1, 10000)$, and $H = 10000$. Then the greedy approximation algorithm produces a clearly inoptimal solution.
+\end{example} 
 % as the following example suggests ...The construction above shows that the objective function of Ellis's problem can approximate that of a knapsack problem with arbitrary precision, and this algorithm
@@
 \subsection{Pseudopolynomial-time dynamic program}
-In this subsection, we provide an algorithmic solution to Ellis's problem that runs in $O(Hm + m\log m)$ time and $O(Hm)$ space.
-The algorithm resembles the dynamic programming algorithm for the binary knapsack problem.
+In this subsection, we assume, with minimal loss of generality, that $g_j \in \mathbb{N}$ for $j = 1\dots m$ and $H \in\mathbb{N}$, and provide an algorithmic solution to Ellis's problem that runs in $O(Hm + m\log m)$ time and $O(Hm)$ space.
+The algorithm resembles a familiar dynamic programming algorithm for the binary knapsack problem.
 Because we cannot assume that $H \leq m$ (as was the case in Alma's problem), this represents a pseudopolynomial-time solution (Dantzig 1957; Garey and Johnson 1979; \emph{Wikipedia}, s.v.
@@
 
-With minimal loss of generality, we assume that $g_j \in \mathbb{N}$ for $j = 1\dots m$ and $H \in\mathbb{N}$.
 For $j = 0 \dots m$ and $h = 0 \dots H$, let $\mathcal{X}[j, h]$ denote the optimal portfolio using only the schools $\{ 1, \dots, j\}$ and costing no more than $h$, and let $V[j,h] = v(\mathcal{X}[j, h])$.
@@
 Optimality follows from the foregoing discussion.
-Sorting by $t_j$ is $O(m \log m)$.
+Sorting $t$ is $O(m \log m)$.
 The number of indices at which $V(j, h)$ is nontrivially defined is $O(Hm)$.
-To prevent $V(j, h)$ from being evaluated more than once at a given index, its values can be stored in a dictionary as they are computed.
+To prevent redundant calls in the recursive function $V(j, h)$, its values must be stored in a dictionary as they are computed; this dictionary’s size is $O(Hm)$.
 \end{proof}
@@
 As with the knapsack problem, Ellis's problem admits a complementary dynamic program that iterates on the value of the cheapest portfolio instead of on the cost of the most valuable portfolio.
-We will use this algorithm as the basis for a fully polynomial-time approximation scheme for Ellis's problem.
+We will use this algorithm as the basis for a fully polynomial-time approximation scheme for Ellis's problem that uses $O(m^3 / \varepsilon)$ time and space.
+Here we assume, with minimal loss of generality, that each $t_j$ is integral.
 %Now, we modify Algorithm \ref{ellisDP2} into a fully polynomial-time approximation scheme for Ellis's problem.
-Here we assume only that is each $t_j$ is integral.
-We will represent the approximate value of each portfolio using a fixed-point decimal with a precision of $P$, where $P$ is the number of digits to retain after the decimal point.
-Since $t_m$ is a trivial upper bound on the value of any portfolio, the set of possible portfolio valuations is now finite: $\mathcal{V} = \{0, 1\times 10^{-P}, 2\times 10^{-P}, \dots, t_m - 1\times 10^{-P}, t_m\}$.
-Its cardinality is $|\mathcal{V} | = P t_m$.
+
+We will represent the approximate value of portfolios using a fixed-point decimal with a precision of $P$, where $P$ is the number of digits to retain after the decimal point.
 Let $r[x] = \lfloor 10^P x \rfloor 10^{-P}$ denote the value of $x$ rounded down to its nearest fixed-point representation.
+Since $\bar U = \sum_{j\in \mathcal{C}} f_j t_j$ is an upper bound on the value of any portfolio, the set of possible approximate portfolio valuations $\mathcal{V}$ is finite:
+\begin{equation}
+\mathcal{V} = \Bigl\{0, 1\times 10^{-P}, 2\times 10^{-P}, \dots, r\bigl[\bar U- 1\times 10^{-P}\bigr], r\bigl[\bar U\bigr], r\bigl[\bar U+ 1\times 10^{-P}\bigr]\Bigr\}
+\end{equation}
+Then $|\mathcal{V} | = \bar U \times 10^P + 2$.
 
-For integers $0 \leq j \leq m$ and $v \leq t_m$, let $\mathcal{W}[j, v]$ denote the least expensive portfolio using only the schools $\{ 1, \dots, j\}$ and having value at least $v$, if such a portfolio exists.
-Let $G[j, v] = r\bigl[v(\mathcal{W}[j, v])\bigr]$ denote the fixed-point representation of the corresponding value, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist.
-It is clear that if $v \leq 0$, then $\mathcal{W}[j, v] = \O$ and $G[j, h] = 0$.
-If $j = 0$ and $v > 0$, then $G[j, h] = \infty$.
+For the remainder of this subsection, unless otherwise specified, the word \emph{value} refers to a portfolio’s value within the fixed-point framework, with the understanding that this is an approximation.
+We will account for the approximation error below when we prove the dynamic program’s validity.
+
+For integers $0 \leq j \leq m$ and $v \in [-\infty, 0) \cup \mathcal{V}$, let $\mathcal{W}[j, v]$ denote the least expensive portfolio that uses only schools $\{ 1, \dots, j\}$ and has value at least $v$, if such a portfolio exists.
+Denote its value by $G[j, v] = r\bigl[v(\mathcal{W}[j, v])\bigr]$, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist.
+It is clear that if $v \leq 0$, then $\mathcal{W}[j, v] = \O$ and $G[j, h] = 0$, and that if $j = 0$ and $v > 0$, then $G[j, h] = \infty$.
 For the remaining indices (where $j, v > 0$), we claim that
@@
 \infty, \quad & t_j < v \\
-\min\bigl\{G[j-1, v], g_j + G[j-1, v - \Delta] \bigr\}, \quad & t_j \geq v 
+\min\bigl\{G[j-1, v], g_j + G[j-1, v - \Delta_j(v)] \bigr\}, \quad & t_j \geq v 
 %\begin{cases}
@@
 In the $t_j < v$ case, any feasible portfolio must be composed of schools with utility less than $v$, and therefore its valuation can not equal $v$, meaning that $\mathcal{W}[j, v]$ is undefined.
-In the $t_j \geq v$ case, the first argument to $\min\{\}$ says simply that omitting $j$ and choosing $\mathcal{W}[j-1, v]$ is a feasible choice for $\mathcal{W}[j, v]$.
+In the $t_j \geq v$ case, the first argument to $\min\{\}$ says simply that omitting $j$ and choosing $\mathcal{W}[j-1, v]$ is a permissible choice for $\mathcal{W}[j, v]$.
 If, on the other hand, $j \in \mathcal{W}[j, v]$, then
@@
 
-Once $G[j, v]$ has been calculated at each index, the corresponding portfolio can be found by applying the observation that $
-\mathcal{W}[j, v]$ contains $j$ if and only if $G[j, v] < G[j-1, v]$.
-Then an approximate solution to Ellis's problem is obtained by applying this observation to the approximate optimal objective value $\max\{ w: G[m, w] \leq H\}$.
+Once $G[j, v]$ has been calculated at each index, the associated portfolio can be found by applying the observation that $\mathcal{W}[j, v]$ contains $j$ if and only if $G[j, v] < G[j-1, v]$.
+Then an approximate solution to Ellis's problem is obtained by computing the optimal objective value $\max\{ w: G[m, w] \leq H\}$ and corresponding portfolio.
 
@@
 %\DontPrintSemicolon
-\caption{Polynomial-time approximation scheme for Ellis's problem.} \label{ellisDP3}
+\caption{Fully polynomial-time approximation scheme for Ellis's problem.} \label{ellisDP3}
 \KwData{Utility values $t \in \mathbb{N}^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$, tolerance $\varepsilon \in (0, 1)$.}
 Index schools in ascending order by $t$\;
-$P \gets \bigl\lceil\log_{10}\left(m / \varepsilon f_m t_m\right)\bigr\rceil$\;
+Set precision $P \gets \bigl\lceil\log_{10}\left(m^2 / \varepsilon \bar U\right)\bigr\rceil$\;
 \SetKwProg{Fn}{function}{ do}{end}
@@
 	\lElseIf{$j=0$ or $t_j < v$}{\Return{$\infty$}}
-	\Else {\Return{$\min\Bigl\{ G(j-1, v), g_j + G\bigl(j-1, v -  \Delta_j(v)\bigr) \Bigr\}$}}
+	\lElse {\Return{$\min\bigl\{ G(j-1, v), g_j + G\bigl(j-1, v -  \Delta_j(v)\bigr) \bigr\}$}}
 }
-%	\uIf{$f_j < 1$}{\Return{$\min\bigl\{ G(j-1, v), g_j + G(j-1, v -  \lfloor \frac{f_j}{1 - f_j} (t_j - v)\rfloor ) \bigr\}$}}
-%	\uElseIf{$f_j = 1$ and $f_j t_j \geq v$}{\Return{$\min\bigl\{G[j-1, v], g_j \bigr\}$}}
-%	\lElse{\Return{$G(j-1, v)$}}
-%	}
-%\lFor{$j=1\dots m$}
-%{$\operatorname{push!}\bigl(\mathtt{H}, (j, f_j, t_j)\bigr)$}
-%\For{$i=1\dots h$}
-%{
-%    $\mathtt{D} \gets \operatorname{pop!}(\mathtt{H})$\;
-%    $\operatorname{append!}(\mathtt{X}, \mathtt{D.j})$\;
-%     \lIfElse{$i=1$}{$\operatorname{append!}(\mathtt{V}, \mathtt{D.f}\cdot\mathtt{D.t})$}
-%     {$\operatorname{append!}(\mathtt{V}, \mathtt{V[i-1]} + \mathtt{D.f}\cdot\mathtt{D.t})$}
-%    \For{$\mathtt{E} \in \mathtt{H}$}
-%	{
-%	\lIfElse{$\mathtt{E.j} < \mathtt{D.j}$}{$\mathtt{E.t} \gets  \mathtt{E.t} \cdot(1 -  \mathtt{D.f})$}{$\mathtt{E.t} \gets  \mathtt{E.t} -  \mathtt{D.f}\cdot\mathtt{D.t}$}
-%	}
-%}
-$v^* \gets  \max\{ w: G(m, w) \leq H\}$\;
-$v \gets v^*$\;
+$v\gets  \max\{ w: G(m, w) \leq H\}$\; \label{vrecordedhere}
 $\mathcal{X} \gets \O$\;
@@
 \begin{theorem}[Validity of Algorithm \ref{ellisDP3}]
-Algorithm \ref{ellisDP3} produces a $(1 - \varepsilon)$-optimal application portfolio for Ellis's problem.% in $O(D m + m \log m)$ time and $O(D m)$ space, where $D$ is the product of the denominators of the $f_j$-values.
+Algorithm \ref{ellisDP3} produces a $(1 - \varepsilon)$-optimal application portfolio for Ellis's problem. It uses $O(m^3 /\varepsilon)$ time and $O(m^3/\varepsilon)$ space.
+% in $O(D m + m \log m)$ time and $O(D m)$ space, where $D$ is the product of the denominators of the $f_j$-values.
 \end{theorem}
@@
 \begin{proof}
-%Let $\mathcal{W}$ denote the output of Algorithm \ref{ellisDP3} and $\mathcal{X}$ the true optimum. In the main loop, each time the target valuation $v$ is updated using
-Because $\Delta_j(v)$ is rounded down, if $j \in \mathcal{W}[j, v]$, then the true value of $(1 - f_j) v\bigl(\mathcal{W}[j-1, v- \Delta_j(v)]\bigr) + f_j t_j$ may exceed the fixed-point value $v$ of $\mathcal{W}[j, v]$, but not by more than $10^{-P}$.
+(Optimality.) Let $\mathcal{W}$ denote the output of Algorithm \ref{ellisDP3} and $\mathcal{X}$ the true optimum.
+We know that $v(\mathcal{X}) \leq \bar U$, and because each singleton portfolio is feasible, $\mathcal{X}$ must be more valuable than the average singleton portfolio; that is, $v(\mathcal{X}) \geq \bar U / m$.
+
+% In the main loop, each time the target valuation $v$ is updated using
+Because $\Delta_j(v)$ is rounded down in the recursion relation defined by \eqref{recursionrelationforcostmindp} and \eqref{deltajvdef}, if $j \in \mathcal{W}[j, v]$, then the true value of $(1 - f_j) v\bigl(\mathcal{W}[j-1, v- \Delta_j(v)]\bigr) + f_j t_j$ may exceed the fixed-point value $v$ of $\mathcal{W}[j, v]$, but not by more than $10^{-P}$.
 This error accumulates with each school added to $\mathcal{W}$, but the number of additions is at most $m$.
-Therefore, where $v^*(\mathcal{W})$ denotes the fixed-point valuation of $\mathcal{W}$ as recorded in the algorithm,
-\begin{equation}
-v(\mathcal{W}) - v^*(\mathcal{W}) \leq m 10^{-P} 
-\end{equation}
+Therefore, where $v^*(\mathcal{W})$ denotes the fixed-point valuation of $\mathcal{W}$ recorded in line \ref{vrecordedhere} of the algorithm,
+$v(\mathcal{W}) - v^*(\mathcal{W}) \leq m 10^{-P}$.
+
 We can define $v^*(\mathcal{X})$ analogously as the fixed-point valuation of $\mathcal{X}$ when its elements are added in index order and its value is updated and rounded down at each addition in accordance with \eqref{solvemeforvwjvminusj}.
-For the same reason as above,
-$v(\mathcal{X}) - v^*(\mathcal{X}) \leq m 10^{-P}$.
+By the same logic,
+$v(\mathcal{X}) - v^*(\mathcal{X}) \leq m 10^{-P}$. The optimality of $\mathcal{W}$ in the fixed-point environment implies that $v^*(\mathcal{W}) \geq v^*(\mathcal{X})$.
 
-Finally, by the trivial assumption that all $g_j \leq H$, we know $v(\mathcal{X}) \geq v(\{m\}) = f_m t_m$. From this we obtain
+Applying these observations, we have
+%the fact that $\bar U / m \leq v(\mathcal{X} \leq \bar U$, we have
 \begin{equation}
@@
 v(\mathcal{W}) &\geq v^*(\mathcal{W}) \geq v^*(\mathcal{X})
-\geq v(\mathcal{X}) - m 10^{-P} \\
-&\geq \left(1 - \frac{m 10^{-P}}{f_m t_m}\right) v(\mathcal{X})
+\geq v(\mathcal{X}) - m 10^{-P}
+%\\&
+\geq \left(1 - \frac{m^2 10^{-P}}{\bar U}\right) v(\mathcal{X})
 \geq \left(1 - \varepsilon\right) v(\mathcal{X})
@@
 \end{equation}
-which completes the proof. 
-\end{proof}
+which establishes the approximation bound.
 
-\begin{theorem}[Computational complexity of Algorithm \ref{ellisDP3}]
-Algorithm \ref{ellisDP3} uses $O(m /\varepsilon + m \log m)$ time and $O(m/\varepsilon)$ space, where $D$ is the product of the denominators of the $f_j$-values.
-\end{theorem}
-
-% To make this fully polynomial, need to get rid of the f_m t_m business ...
-\begin{proof}
-In the worst case, the number of indices at which $G(j, v)$ is evaluated is $O(|\mathcal{V}| m)$. Since
+(Computational resources.) In the worst case, the number of indices at which $G(j, v)$ is evaluated is $O(|\mathcal{V}| m)$. Since
 \begin{equation}
-|\mathcal{V}| = t_m 10^{P} = t_m10^ { \bigl\lceil\log_{10}\left(m / \varepsilon f_m t_m\right)\bigr\rceil}
-%\leq {\text{const.}}/{\varepsilon}
+|\mathcal{V}| = \bar U \times 10^{P} + 2 = \bar U \times 10^ { \bigl\lceil\log_{10}\left(m ^2/ \varepsilon \bar U\right)\bigr\rceil} + 2
+\leq\frac{m^2}{\varepsilon} \times \text{const.}
 \end{equation}
-is $O(1/ \varepsilon)$, adding the $O(m \log m)$-time required to sort $t$ yields the result.
-By recording the values of $G(j, v)$ in a dictionary as they are computed, we prevent redundant recursive calls at a cost of $O(|\mathcal{V}| m) = O(m / \varepsilon)$ space.
+is $O(m^2/ \varepsilon)$, the time complexity is as promised.
+To prevent redundant recursive calls to $G(j, v)$, we need to record its values in a dictionary as they are computed; the space required by the dictionary is $O(|\mathcal{V}| m) = O(m^3 / \varepsilon)$.
 \end{proof}

### 2022-02-27 paper.tex
introduction

@@
 When colleges differ in their application fees, we show that the problem is NP-complete.
-We provide two dynamic programs for this more general problem.
+We provide two dynamic programs for this more general setup.
 The first produces an exact solution in pseudopolynomial time, and the second yields a fully polynomial-time approximation scheme.
@@
 \ifEN \section{Introduction}  \else \section{서론} \fi
-We consider the optimal college application strategy for a single student.
+This paper considers the optimization problem
+%\begin{definition}[Optimal college application portfolio] \label{generalproblemstatement}
+\begin{align}
+\begin{split}
+\text{maximize}\quad & \operatorname{E}\left[\max\bigr\{t_0,
+\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\right] \\
+\text{subject to}\quad & \mathcal{X} \subseteq \mathcal{C}, ~~\sum_{j\in \mathcal{X}} g_j \leq H
+\end{split}
+\end{align}
+%\end{definition}
+where $\mathcal{C} = \{ 1 \dots m\}$ is an index set, $H$ is a budget parameter, and for $j = 1 \dots m$, $g_j > 0$ is a cost parameter, $Z_j$ is a random, independent Bernoulli variable with probability $f_j$, and $t_j  > 0$ is a utility parameter.
+Without loss of generality, we assume that the $t_j$-values are weakly increasing in $j$, that $0 \leq t_0 \leq t_1$, and that each $g_j \leq H$.
+
+We refer to this problem as the \emph{optimal college application portfolio,} as follows: Consider a college admissions market with $m$ college.
+The $j$th college is named $c_j$.
+Consider a single student, and let $f_j$ denote her probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not.
+It is appropriate to assume that the $Z_j$ are probabilistically independent as long as $f_j$ are probabilities estimated specifically for this student (as opposed to generic acceptance rates).
+Each $t_j$-value indicates the utility associated with attending $c_j$, and $t_0$ gives the utility associated with not attending college, an option named $c_0$.
+The application fee for $c_j$ is $g_j$, and $H$ represents the student's total budget to spend on application fees.
+Then the student's objective is to maximize the expected utility associated with the best school she gets into, and her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which the student applies.
+
 As Chao (2014) remarked, this represents a somewhat subtle portfolio optimization problem.
@@
 Finally, the combinatorial nature of the college application problem makes it difficult to solve using the gradient-based techniques used in continuous portfolio optimization.
+Chao estimated her model (which considers application as a \emph{cost} rather than a constraint) by clustering the schools so that $m=8$, a scale at which enumeration is possible.
+Our study pursues a more general solution.
 
-Chao estimated her model (which considers application as a \emph{cost} rather than a constraint) by clustering the schools so that $m=8$, a scale at which enumeration is possible.
-However, subject to certain assumptions on the quality of the data available to students in their decisionmaking process, an optimal application portfolio for a single student can be computed in time polynomial in $h$ and $m$, as we show presently.
+% Further references to nestedness and greedy approximation algos
+
+\subsection{Structure of this paper}
+Section \label{preliminaries} introduces some additional notation and assumptions that can be imposed with trivial loss of generality.
+
+In section \ref{homogappcosts}, we solve the special case where each $g_j = 1$ and $H$ is an integer $h \leq m$.
+This case mirrors the centralized college application process in Korea, where there is no application fee, but students are allowed to apply to only three schools during the main admissions cycle.
+We show that an intuitive heuristic is in fact a $1/h$-approximation algorithm.
+Then, we show that the optimal portfolios are nested in the budget constraint, which yields an exact algorithm that runs in $O(hm)$-time.
+
+In section \ref{hetappcosts}, we turn to the scenario in which colleges differ in their application fees.
+We show that the decision form of the portfolio optimization problem is NP-complete through a reduction from the binary knapsack problem.
+We provide two dynamic programs for this more general setup.
+The first iterates on total expenditures and produces an exact solution in pseudopolynomial time, namely $O(Hm + m \log m)$.
+The second iterates on truncated portfolio valuations and yields a fully polynomial-time approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in $O(m^3 / \varepsilon)$ time.
+% subject to certain assumptions on the quality of the data available to students in their decisionmaking process, an optimal application portfolio for a single student can be computed in time polynomial in $h$ and $m$, as we show presently.
+
+% \footnote{This assumption is appropriate when $f$ gives the admissions probabilities \emph{specifically} for Alma, as estimated from her academic abilities.
+It is inappropriate to take a school's overall acceptance rate, which reflects the admissions probability for a \emph{generic} student, as $f_j$.}
 
@@
 
-\ifEN \section{Model}  \else \section{모형} \fi \label{optimalcollegeappstrat}
+\section{Notation and preliminaries} \label{preliminaries}
+Before discussing the optimal solution, we organize a few preliminary results.
 
-\ifEN \section{Homogeneous application costs}  \else \subsection{} \fi
-Consider a college admissions market with $m$ schools, $\mathcal{C} = \{ 1 \dots m\}$.
-The $j$th school is named $c_j$.
-By government regulation, students are allowed to apply to no more than $h$ schools.
-(In the Korean case, $m=202$ and $h=3$.) We consider the optimal application strategy for a single student, whom we will call Alma.
-
-For $j=1\dots m$, let $t_j > 0$ denote the utility that Alma receives from attending $c_j$, and let $f_j$ denote the probability that she is admitted if she applies.
-Let the random variable $Z_j$ equal one if Alma gets into $c_j$ and zero otherwise.
-We assume that Alma’s admissions outcome at each school is independent of her outcome at the other schools.\footnote{This assumption is appropriate when $f$ gives the admissions probabilities \emph{specifically} for Alma, as estimated from her academic abilities.
-It is inappropriate to take a school's overall acceptance rate, which reflects the admissions probability for a \emph{generic} student, as $f_j$.} Thus $Z$ is a vector of independent Bernoulli variables with probabilities given by $f$.
-Let $c_0$ denote Alma's outcome if she does not get into any college, with utility $t_0$ and $f_0 = 1$.
-Without loss of generality, we assume that $t_{j-1} \leq t_j$ for $j=1 \dots m$.(We may assume that $t_0 \leq t_1$ because schools for which $t_j < t_0$ can be trivially excluded from consideration.)
-%For the \emph{generic} student, when $f_j$ is simply $c_j$'s acceptance rate, admissions outcomes will almost certainly correlate across schools.
-
-Let $\mathcal{X} \subseteq \mathcal{C}$ denote the set of schools to which Alma applies, called her \emph{application portfolio.}The expected utility Alma receives from $\mathcal{X}$ is called the portfolio’s \emph{valuation}.
+The set $\mathcal{X} \subseteq \mathcal{C}$ of schools to which a student applies is called her \emph{application portfolio.} The expected utility the student receives from $\mathcal{X}$ is called its \emph{valuation}.
 %Let $x$ denote the same encoded as a binary vector, where $x_j=1 \iff j\in \mathcal{X}$ for $j=1\dots m$.
 \begin{definition}[Portfolio valuation function]
-$v(\mathcal{X}) = \operatorname{E}\left[ \max\{t_j Z_j : j \in \mathcal{X}\} \right]$.
+$v(\mathcal{X}) =  \operatorname{E}\left[\max\bigr\{t_0,
+\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\right]$.
 \end{definition}
-It is helpful to define the random variable $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$ as the utility achieved by the portfolio, so that $v(\mathcal{X}) = \operatorname{E}[X]$.
+It is helpful to define the random variable $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$ as the utility achieved by the schools in the portfolio, so that when $t_0 = 0$, $v(\mathcal{X}) = \operatorname{E}[X]$.
 
-Let $p_j(\mathcal{X})$ denote the probability that Alma attends $c_j$.
-Alma attends $c_j$ if and only if she \emph{applies} to $c_j$, is \emph{admitted} to $c_j$, and is \emph{rejected} from any school she prefers to $c_j$; that is, any school with higher index.
+Let $p_j(\mathcal{X})$ denote the probability that the student attends $c_j$.
+This occurs if and only if she \emph{applies} to $c_j$, is \emph{admitted} to $c_j$, and is \emph{rejected} from any school she prefers to $c_j$; that is, any school with higher index.
 Hence, for $j= 0\dots m$,
@@
 \begin{cases}
-\displaystyle f_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - f_{j’}), \quad & j \in \{0\}\cup\mathcal{X}\\
+\displaystyle f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}), \quad & j \in \{0\}\cup\mathcal{X}\\
 0, \quad & \text{otherwise}
-\end{cases} \\
-%\iff\quad p_j(x) &= x_j  f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’})
+\end{cases} 
 \end{align}
-%with the understanding that $x_0 = 1$ and
 where the empty product equals one. The following proposition follows immediately.
@@
 \begin{align}
-v(\mathcal{X}) &= \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - f_{j’}) \Bigr)  \label{closedformportfoliovaluationX}%, \quad \text{or equivalently,}\\
+v(\mathcal{X}) &= \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)  \label{closedformportfoliovaluationX}%, \quad \text{or equivalently,}\\
 %\qquad v(x) &= t_0 \prod_{j=1}^m (1 - f_{j} x_j) + \sum_{j=1}^m \Bigl( x_j t_j f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’}) \Bigr) \label{closedformportfoliovaluationx}
@@
 
-Next, we show that without loss of generality, we may assume that $t_0 = 0$ (or any constant).
+Next, we show that without loss of generality, we may assume that $t_0 = 0$ (or any constant less than $t_1$).
 \begin{theorem} \label{assumetzerozero}
@@
 
-%Let us express the optimal college application problem as an INLP.
+\ifEN \section{Homogeneous application costs}  \else \subsection{} \fi \label{homogappcosts}
+Let us consider the special case in which $g_j = 1$ and $H$ is a constant $h \leq m$.
+This case is similar to the centralized college admissions process in Korea, where there is no application fee, but by law, students are allowed to apply to no more than $h$ schools.
+(In the Korean case, $m=202$ and $h=3$.) Applying Theorem \ref{assumetzerozero}, we assume that $t_0 = 0$.
+
+In this section, we will call the applicant Alma, and refer to the corresponding optimization problem as Alma's problem.
+
 \begin{definition}[Alma’s problem]
-Alma's optimal college application portfolio is given by the solution to the following integer nonlinear program:
+Alma's optimal college application portfolio is given by the solution to the following combinatorial optimization problem:
 \begin{align}
 \begin{split}
-\text{maximize}\quad &  v(\mathcal{X}) \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - f_{j’})\\
+\text{maximize}\quad &  v(\mathcal{X}) = \sum_{j\in
+%\{0\}\cup
+\mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i})\\
 \text{subject to}\quad & \mathcal{X}\subseteq\mathcal{C}, \quad|\mathcal{X}| \leq h 
@@
 \end{definition}
+%This problem can be formulated as an integer nonlinear program in $m$ binary variables ..
 
 \ifEN \subsection{Approximation properties of a na\"ive solution}  \else \subsection{모형화} \fi
-Notice that for a given school $c_j$, the expected utility associated with applying to $c_j$ is simply $\operatorname{E}[t_j Z_j] = t_j f_j$.
-It is therefore tempting to adopt the following greedy strategy, which turns out to be inoptimal.
+The expected utility associated with a single school $c_j$ is simply $\operatorname{E}[t_j Z_j] = t_j f_j$. It is therefore tempting to adopt the following strategy, which turns out to be inoptimal.
 \begin{definition}[Na\"ive algorithm for Alma’s problem] \label{naivealgorithm}
@@
 \begin{example}
-Suppose $m=3$, $q=2$, and
-\begin{align*}
-t &= (70, 80, 90) \\
-f &= (0.4, 0.4, 0.3) \\
-\implies t * f &= (28, 32, 27)
-\end{align*}
-The na\"ive algorithm picks $\mathcal{T} = \{1, 2\}$ with 
-\[v(\mathcal{T}) = 70(0.4)(1-0.4) + 80(0.4) = 48.8\]
+Suppose $m=3$, $h=2$,  $ t= (70, 80, 90) $, and $f = (0.4, 0.4, 0.3)$. 
+%\implies t * f &= (28, 32, 27)
+%\begin{align*}
+%t &= (70, 80, 90) \\
+%f &= (0.4, 0.4, 0.3) \\
+%\implies t * f &= (28, 32, 27)
+%\end{align*}
+Then the na\"ive algorithm picks $\mathcal{T} = \{1, 2\}$ with 
+$v(\mathcal{T}) = 70(0.4)(1-0.4) + 80(0.4) = 48.8$.
 But $\mathcal{X} = \{2, 3\}$ with
-\[v(\mathcal{X}) = 80(0.4)(1-0.3) + 90(0.3) = 49.4\]
+$v(\mathcal{X}) = 80(0.4)(1-0.3) + 90(0.3) = 49.4$
 is the optimal solution. 
@@
 \begin{equation*}
-v(\mathcal{X}_h) = \sum_{j= h +1}^m \Bigl( t_j f_j \prod_{j' = j+1}^m (1 - f_{j'}) \Bigr) =  \sum_{j= 1}^h  (1 - \varepsilon)^{j} \approx h
+v(\mathcal{X}_h) = \sum_{j= h +1}^m \Bigl( t_j f_j \prod_{j' = j+1}^m (1 - f_{j'}) \Bigr) =  \sum_{j= 1}^h  (1 - \varepsilon)^{j} \approx h.
 \end{equation*}
@@
 It turns out that the solution to Alma's problem possesses a special structure: An optimal portfolio of size $h+1$ includes an optimal portfolio of size $h$ as a subset.
-To show this result, we need to introduce some new notation.
+% Incl above
+%To show this result, we need to introduce some new notation.
 For an application portfolio $\mathcal{X}$, the random variable $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$ represents the utility achieved by the portfolio, so that $v(\mathcal{X}) = \operatorname{E}[X]$.
+
+% Unused
 %It is also helpful to define $\Delta v(\mathcal{X}, j)$ as the utility gained by adding $c_j$ to the portfolio $\mathcal{X}$. That is, 
@@
 \begin{theorem}[Nestedness of optimal application portfolios] \label{nestedapplication}
-Let $\mathcal{X}_h$ denote Alma’s optimal application portfolio when the application limit is $h$. If each $\mathcal{X}_h$ is unique, then
+There exists a sequence of portfolios $\{\mathcal{X}_h\}_{h=1}^m$ such that each $\mathcal{X}_h$ is an optimal application portfolio when the application limit is $h$, and the sequence satisfies the nestedness relation
 \begin{equation}
-\mathcal{X}_1 \subset \mathcal{X}_2\subset \dots \subset \mathcal{X}_m
+\mathcal{X}_1 \subset \mathcal{X}_2\subset \dots \subset \mathcal{X}_m.
 \end{equation}
-If the optimal portfolios are not unique, then there is a sequence of optimal portfolios satisfying the above.
 \end{theorem}
@@
 \begin{equation}\label{Vyastildet}
-v(\mathcal{Y}\cup\{k\}) = \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y})\end{equation}
+v(\mathcal{Y}\cup\{k\}) = \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y}).\end{equation}
 The transformation to $\tilde t$ does not change the order of the $t_j$-values.
@@
 \subsection{Properties of the optimal portfolios}
-The nestedness property implies that Alma's expected utility is a concave function of $h$.
+The nestedness property implies that Alma's expected utility is a discretely concave function of $h$.
 
 \begin{theorem}[Optimal portfolio valuation concave in $h$] \label{concavityinh}
-For $h = 2 \dots (m-1)$, \begin{equation}v(\mathcal{X}_h) - v(\mathcal{X}_{h-1}) \geq v(\mathcal{X}_{h+1}) - v(\mathcal{X}_{h})\end{equation} 
+For $h = 2 \dots (m-1)$, \begin{equation}v(\mathcal{X}_h) - v(\mathcal{X}_{h-1}) \geq v(\mathcal{X}_{h+1}) - v(\mathcal{X}_{h}).\end{equation} 
 \end{theorem}
@@
 &=  v(\mathcal{X}_{h-1} \cup\{j, k\}) + v(\mathcal{X}_{h-1})\\
-&=  v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1})
+&=  v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1}).
 \end{split} 
@@
 
-\section{Heterogeneous application costs}
+\section{Heterogeneous application costs}  \label{hetappcosts}
 In this section, we turn to the more general problem in which the constant $g_j$ represents the \emph{cost} of applying to $c_j$ and the student, whom we now call Ellis, has a \emph{budget} of $H$ to spend on college applications.
 Applying Theorem \ref{assumetzerozero}, we assume $t_0 = 0$ and disregard $c_0$ throughout.
-We also assume with trivial loss of generality that all $g_j \leq H$.
 
 \begin{definition}[Ellis's problem]
-Ellis's optimal college application portfolio is given by the solution to the following integer nonlinear program:
+Ellis's optimal college application portfolio is given by the solution to the following combinatorial optimization problem.
 \begin{align}
 \begin{split}
-\text{maximize}\quad & v(x) = \sum_{j=1}^m x_j t_j f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’}) \\
-\text{subject to}\quad & \sum_{j=1}^m g_j x_j \leq H \\
-&x_j \in \{0, 1\}, \quad j = 1\dots m
+\text{maximize}\quad &  v(\mathcal{X}) = \sum_{j\in
+%\{0\}\cup
+\mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)\\
+\text{subject to}\quad & \mathcal{X}\subseteq\mathcal{C},~~\sum_{j \in \mathcal{X}} g_j \leq H 
 \end{split}
@@
 Clearly, $\mathcal{X} \subseteq \mathcal{C}$ is feasible for Ellis’s problem if and only if it is feasible for the knapsack instance.
-Now, we observe that if $\mathcal{X}$ is nonempty,
+Now, we observe that for any nonempty $\mathcal{X}$,
 \begin{align}
@@
 \geq \sum_{j \in \mathcal{X}} u_j  - m\delta U_{\mathrm{max}}
-= \sum_{j \in \mathcal{X}} u_j  - 1
+= \sum_{j \in \mathcal{X}} u_j  - 1.
 \end{split}
 \end{align}
-This means that given a nonempty portfolio $\mathcal{X}$, its value for the corresponding knapsack instance is the smallest integer greater than $v( \mathcal{X})$.
+This means that the value of an application portfolio $\mathcal{X}$ for the corresponding knapsack instance is the smallest integer greater than $v( \mathcal{X})$.
 That is, $\sum_{j \in \mathcal{X}} u_j \geq U$ if and only if $v(\mathcal{X}) \geq U -1$.
@@
 \subsection{Pseudopolynomial-time dynamic program}
-In this subsection, we assume, with minimal loss of generality, that $g_j \in \mathbb{N}$ for $j = 1\dots m$ and $H \in\mathbb{N}$, and provide an algorithmic solution to Ellis's problem that runs in $O(Hm + m\log m)$ time and $O(Hm)$ space.
-The algorithm resembles a familiar dynamic programming algorithm for the binary knapsack problem.
-Because we cannot assume that $H \leq m$ (as was the case in Alma's problem), this represents a pseudopolynomial-time solution (Dantzig 1957; Garey and Johnson 1979; \emph{Wikipedia}, s.v.
+In this subsection, we assume, with a small loss of generality, that $g_j \in \mathbb{N}$ for $j = 1\dots m$ and $H \in\mathbb{N}$, and provide an algorithmic solution to Ellis's problem that runs in $O(Hm + m\log m)$ time and $O(Hm)$ space.
+The algorithm resembles a familiar dynamic programming algorithm for the binary knapsack problem (Dantzig 1957; \emph{Wikipedia}, s.v.
 ``Knapsack problem'').
+Because we cannot assume that $H \leq m$ (as was the case in Alma's problem), this represents a pseudopolynomial-time solution (Garey and Johnson 1979).
 
@@
 We will use this algorithm as the basis for a fully polynomial-time approximation scheme for Ellis's problem that uses $O(m^3 / \varepsilon)$ time and space.
-Here we assume, with minimal loss of generality, that each $t_j$ is integral.
+Here we assume, with a small loss of generality, that each $t_j$ is a natural number.
 %Now, we modify Algorithm \ref{ellisDP2} into a fully polynomial-time approximation scheme for Ellis's problem.
@@
 Let $r[x] = \lfloor 10^P x \rfloor 10^{-P}$ denote the value of $x$ rounded down to its nearest fixed-point representation.
-Since $\bar U = \sum_{j\in \mathcal{C}} f_j t_j$ is an upper bound on the value of any portfolio, the set of possible approximate portfolio valuations $\mathcal{V}$ is finite:
+Since $\bar U = \sum_{j\in \mathcal{C}} f_j t_j$ is an upper bound on the value of any portfolio, and since we will ensure that each fixed-point approximation is an underestimate of the portfolio's true value, the set $\mathcal{V}$ of possible valuations possible in the fixed-point framework is finite:
 \begin{equation}
@@
 For integers $0 \leq j \leq m$ and $v \in [-\infty, 0) \cup \mathcal{V}$, let $\mathcal{W}[j, v]$ denote the least expensive portfolio that uses only schools $\{ 1, \dots, j\}$ and has value at least $v$, if such a portfolio exists.
-Denote its value by $G[j, v] = r\bigl[v(\mathcal{W}[j, v])\bigr]$, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist.
+Denote its value by $G[j, v]$, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist.
 It is clear that if $v \leq 0$, then $\mathcal{W}[j, v] = \O$ and $G[j, h] = 0$, and that if $j = 0$ and $v > 0$, then $G[j, h] = \infty$.
@@
 \begin{equation} \label{solvemeforvwjvminusj}
-v(\mathcal{W}[j, v]) = (1 - f_j )v(\mathcal{W}[j, v]\setminus \{j\}) + f_j t_j\end{equation}
+v(\mathcal{W}[j, v]) = (1 - f_j )v(\mathcal{W}[j, v]\setminus \{j\}) + f_j t_j.\end{equation}
 Therefore, the subportfolio $\mathcal{W}[j, v]\setminus \{j\}$ must have a value of at least $v - \Delta$, where $\Delta$ satisfies $v = (1 - f_j )(v - \Delta) + f_j t_j $.
@@
 When $f_j = 1$, then the singleton $\{j\}$ has $v(\{j\}) \geq v$, so
-\begin{equation}G[j, v] = \min\bigl\{G[j-1, v], g_j \bigr\}\end{equation}
-Defining $\Delta = \infty$ in this case ensures that $G[j-1, v-\Delta] = 0$ as required.
+\begin{equation}G[j, v] = \min\bigl\{G[j-1, v], g_j \bigr\}.\end{equation}
+Defining $\Delta = \infty$ in this case ensures that $G[j-1, v-\infty] = 0$ as required.
 %we have $t_j < v$, and since no portfolio can be more valueable than the largest $t_j$ value of its elements, we must have $G[j, v] =\infty$.
@@
 }
-$v\gets  \max\{ w: G(m, w) \leq H\}$\; \label{vrecordedhere}
+$v\gets  \max\{ w \in \mathcal{V} : G(m, w) \leq H\}$\; \label{vrecordedhere}
 $\mathcal{X} \gets \O$\;
@@
 This error accumulates with each school added to $\mathcal{W}$, but the number of additions is at most $m$.
-Therefore, where $v^*(\mathcal{W})$ denotes the fixed-point valuation of $\mathcal{W}$ recorded in line \ref{vrecordedhere} of the algorithm,
-$v(\mathcal{W}) - v^*(\mathcal{W}) \leq m 10^{-P}$.
+Therefore, where $v'(\mathcal{W})$ denotes the fixed-point valuation of $\mathcal{W}$ recorded in line \ref{vrecordedhere} of the algorithm,
+$v(\mathcal{W}) - v'(\mathcal{W}) \leq m 10^{-P}$.
 
-We can define $v^*(\mathcal{X})$ analogously as the fixed-point valuation of $\mathcal{X}$ when its elements are added in index order and its value is updated and rounded down at each addition in accordance with \eqref{solvemeforvwjvminusj}.
+We can define $v'(\mathcal{X})$ analogously as the fixed-point valuation of $\mathcal{X}$ when its elements are added in index order and its value is updated and rounded down at each addition in accordance with \eqref{solvemeforvwjvminusj}.
 By the same logic,
-$v(\mathcal{X}) - v^*(\mathcal{X}) \leq m 10^{-P}$. The optimality of $\mathcal{W}$ in the fixed-point environment implies that $v^*(\mathcal{W}) \geq v^*(\mathcal{X})$.
+$v(\mathcal{X}) - v'(\mathcal{X}) \leq m 10^{-P}$. The optimality of $\mathcal{W}$ in the fixed-point environment implies that $v'(\mathcal{W}) \geq v'(\mathcal{X})$.
 
@@
 \begin{split}
-v(\mathcal{W}) &\geq v^*(\mathcal{W}) \geq v^*(\mathcal{X})
+v(\mathcal{W}) &\geq v'(\mathcal{W}) \geq v'(\mathcal{X})
 \geq v(\mathcal{X}) - m 10^{-P}

### 2022-02-28 src/OptimalApplication.jl
make a Julia project; unit tests for all functions

@@ -1,0 +1,381 @@
+module OptimalApplication
+
+using Combinatorics: combinations, multiset_combinations
+using DataStructures
+# For implementing the FPTAS for Ellis's problem
+using FixedPointNumbers
+import Base.isless
+
+
+export 
+    valuation,
+    applicationorder,
+    optimalportfolio_dynamicprogram,
+    optimalportfolio_enumerate,
+    optimalportfolio_valuationtable,
+    optimalportfolio_fptas
+
+
+
+"""
+    College(f, t, ft)
+
+Contains a college's admissions probability `f`, utility value `t`,
+and their product `ft`.
+"""
+struct College
+    f::Float64
+    t::Float64
+    ft::Float64
+
+    College(f::Float64, t::Float64) = new(f, t, f * t)
+end
+
+# Overload isless() so that the heap is ordered by expected utility.
+isless(c1::College, c2::College) = isless(c1.ft, c2.ft)
+
+
+"""
+    applicationorder(f, t, h=m; datastructure=:heap)
+
+Produce the optimal order of application for the market defined by admissions probabilities `f`
+and utility values `t`. Stores college data in a binary `:heap` or as a `:dict`. 
+"""
+function applicationorder(
+    f::Vector{Float64},
+    t::Vector{Float64},
+    h = nothing::Union{Int64,Nothing};
+    datastructure = :heap::Symbol)
+    m = length(f)
+    @assert m == length(t)
+    @assert issorted(t)
+    if isnothing(h)
+        h = m
+    else
+        @assert 0 < h ≤ m
+    end
+
+    if datastructure == :heap
+        mkt = MutableBinaryMaxHeap{College}()
+        for j in 1:m
+            push!(mkt, College(f[j], t[j]))
+        end
+
+        onheap = trues(m)
+        apporder = zeros(Int, h)
+        v = zeros(h)
+        for j in 1:h
+            c_k, k = top_with_handle(mkt)
+            v[j] = get(v, j - 1, 0) + c_k.ft
+
+            pop!(mkt)
+
+            apporder[j] = k
+            onheap[k] = false
+            for i in 1:k-1
+                onheap[i] && update!(mkt, i, College(mkt[i].f, mkt[i].t * (1 - c_k.f)))
+            end
+            for i in k+1:m
+                onheap[i] && update!(mkt, i, College(mkt[i].f, mkt[i].t - c_k.ft))
+            end
+        end
[…]

### 2022-02-28 test/runtests.jl

@@ -1,0 +1,73 @@
+using OptimalApplication
+using Test
+
+# Number of markets to generate for each test
+n_markets = 3
+
+function make_correlated_market(m)
+    φ = 0.7
+    t = rand(m) |> sort
+    f = 1 .- (φ * t + (1 - φ) * rand(m))
+    return f, t
+end
+
+@testset verbose = true "Homogeneous application costs" begin
+    m = 20
+    h = m ÷ 2
+
+    for _ in 1:n_markets
+        f, t = make_correlated_market(m)
+
+        X, vX = optimalportfolio_enumerate(f, t, h)
+        sort!(X)
+        W, vW = applicationorder(f, t, h; datastructure = :heap)
+        Y, vY = applicationorder(f, t, h; datastructure = :dict)
+        @test X == sort(W)
+        @test vX ≈ vW[h]
+        @test X == sort(Y)
+        @test vX ≈ vY[h]
+    end
+end
+
+@testset verbose = true "Heterogeneous application costs" begin
+    m = 20
+
+    @testset "Exact algorithms" begin
+        for _ in 1:n_markets
+            f, t = make_correlated_market(m)
+            g = rand(1:5, m)
+            H = sum(g) ÷ 2
+
+            X, vX = optimalportfolio_enumerate(f, t, g, H)
+            sort!(X)
+            W, VW = optimalportfolio_valuationtable(f, t, g, H)
+            Y, vY = optimalportfolio_dynamicprogram(f, t, g, H)
+            # Without memoization
+            Z, vZ = optimalportfolio_dynamicprogram(f, t, g, H, false)
+
+            @test X == sort(W)
+            @test vX ≈ VW[m, H]
+            @test X == sort(Y)
+            @test vX ≈ vY
+            @test X == sort(Z)
+            @test vX ≈ vZ
+        end
+    end
+
+    m = 100
+    ε = 0.1
+
+    @testset "FPTAS" begin
+        for _ in 1:n_markets
+            f, t = make_correlated_market(m)
+            t = round.(Int, 100 * t)
+            g = rand(1:5, m)
+            H = sum(g) ÷ 2
+
+            W, vW = optimalportfolio_fptas(f, t, g, H, ε)
+            Y, vY = optimalportfolio_dynamicprogram(f, t, g, H)
+
+            @test vW / vY ≥ 1 - ε
+        end
+    end
+end

### 2022-02-28 benchmark/benchmark.jl
type annotations, additional tests, benchmark outline

@@ -1,0 +1,113 @@
+using OptimalApplication
+using DataFrames
+using Random
+using StatsBase
+using Base.Threads
+
+# An all-day benchmark; tweak parameters of run with caution.
+
+# Number of times to repeat each computation, where min of these is reported as time
+n_reps = 3
+
+function make_correlated_market(m)
+    A = 10
+    t = ceil.(Int, 10 * randexp(m))
+    sort!(t)
+    f = inv.(t .+ 10 * rand(m))
+    g = rand(5:10, m)
+    H = sum(g) ÷ 2
+    return f, t, g, H
+end
+
+function samecosts(ds, M, n_markets)
+    sizes = Int64[]
+    times = Float64[]
+
+    @threads for i in 1:n_markets
+        println("  i = $i of $n_markets")
+        for m in M
+            f, t, = make_correlated_market(m)
+            push!(sizes, m)
+            push!(times, minimum(@elapsed applicationorder(f, t; datastructure = ds) for r in 1:n_reps))
+        end
+    end
+
+    return DataFrame("m" => sizes, "time" => times)
+end
+
+
+function dp_H(M, n_markets)
+    sizes = Int64[]
+    times = Float64[]
+
+    @threads for i in 1:n_markets
+        println("  i = $i of $n_markets")
+        for m in M
+            f, t, g, H = make_correlated_market(m)
+            push!(sizes, m)
+            push!(times, minimum(@elapsed optimalportfolio_dynamicprogram(f, t, g, H) for r in 1:n_reps))
+        end
+    end
+
+    return DataFrame("m" => sizes, "time" => times)
+end
+
+
+function fptas(epsilon, M, n_markets)
+    sizes = Int64[]
+    times = Float64[]
+
+    @threads for i in 1:n_markets
+        println("  i = $i of $n_markets")
+        for m in M
+            f, t, g, H = make_correlated_market(m)
+            push!(sizes, m)
+            push!(times, minimum(@elapsed optimalportfolio_fptas(f, t, g, H, epsilon) for r in 1:n_reps))
+        end
+    end
+
+    return DataFrame("m" => sizes, "time" => times)
+end
+
+
+
+function table1()
+    println("== Homogeneous cost algorithms ==")
+    M = [5, 50, 500, 5000]
+    n_markets = 50
+
+    println("Timing with dictionary")
+    het_dict = samecosts(:dict, M, n_markets)
+
[…]

### 2022-02-28 src/OptimalApplication.jl

@@ -25,10 +25,10 @@ and their product `ft`.
 """
 struct College
     f::Float64
-    t::Float64
+    t::Real
     ft::Float64
 
-    College(f::Float64, t::Float64) = new(f, t, f * t)
+    College(f::Float64, t::Real) = new(f, t, f * t)
 end
 
 # Overload isless() so that the heap is ordered by expected utility.
@@ -42,10 +42,10 @@ Produce the optimal order of application for the market defined by admissions pr
 and utility values `t`. Stores college data in a binary `:heap` or as a `:dict`. 
 """
 function applicationorder(
-    f::Vector{Float64},
-    t::Vector{Float64},
-    h = nothing::Union{Int64,Nothing};
-    datastructure = :heap::Symbol)
+        f::Vector{Float64},
+        t::Vector{<:Real},
+        h = nothing::Union{Int64,Nothing};
+        datastructure = :heap::Symbol)::Tuple{Vector{Int64}, Vector{Float64}}
     m = length(f)
     @assert m == length(t)
     @assert issorted(t)
@@ -61,7 +61,7 @@ function applicationorder(
             push!(mkt, College(f[j], t[j]))
         end
 
-        onheap = trues(m)
+        onheap = Set(1:m)
         apporder = zeros(Int, h)
         v = zeros(h)
         for j in 1:h
@@ -71,12 +71,13 @@ function applicationorder(
             pop!(mkt)
 
             apporder[j] = k
-            onheap[k] = false
-            for i in 1:k-1
-                onheap[i] && update!(mkt, i, College(mkt[i].f, mkt[i].t * (1 - c_k.f)))
-            end
-            for i in k+1:m
-                onheap[i] && update!(mkt, i, College(mkt[i].f, mkt[i].t - c_k.ft))
+            delete!(onheap, k)
+            for i in onheap
+                if i < k
+                    update!(mkt, i, College(mkt[i].f, mkt[i].t * (1 - c_k.f)))
+                else
+                    update!(mkt, i, College(mkt[i].f, mkt[i].t - c_k.ft))
+                end
             end
         end
     else
@@ -85,7 +86,6 @@ function applicationorder(
             push!(mkt, j => College(f[j], t[j]))
         end
 
-        indict = trues(m)
         apporder = zeros(Int, h)
         v = zeros(h)
         for j in 1:h
[…]

### 2022-02-28 test/runtests.jl

@@ -1,73 +1,80 @@
 using OptimalApplication
 using Test
+using Random
 
 # Number of markets to generate for each test
 n_markets = 3
 
 function make_correlated_market(m)
-    φ = 0.7
-    t = rand(m) |> sort
-    f = 1 .- (φ * t + (1 - φ) * rand(m))
-    return f, t
+    A = 10
+    t = ceil.(Int, 10 * randexp(m))
+    sort!(t)
+    f = inv.(t .+ 10 * rand(m))
+    g = rand(5:10, m)
+    H = sum(g) ÷ 2
+    return f, t, g, H
 end
 
-@testset verbose = true "Homogeneous application costs" begin
-    m = 20
-    h = m ÷ 2
+@testset verbose = true "OptimalApplication.jl" begin
+    @testset verbose = true "Same app. costs" begin
+        m = 20
+        h = m ÷ 2
 
-    for _ in 1:n_markets
-        f, t = make_correlated_market(m)
-
-        X, vX = optimalportfolio_enumerate(f, t, h)
-        sort!(X)
-        W, vW = applicationorder(f, t, h; datastructure = :heap)
-        Y, vY = applicationorder(f, t, h; datastructure = :dict)
-        @test X == sort(W)
-        @test vX ≈ vW[h]
-        @test X == sort(Y)
-        @test vX ≈ vY[h]
-    end
-end
-
-@testset verbose = true "Heterogeneous application costs" begin
-    m = 20
-
-    @testset "Exact algorithms" begin
         for _ in 1:n_markets
-            f, t = make_correlated_market(m)
-            g = rand(1:5, m)
-            H = sum(g) ÷ 2
+            f, t, = make_correlated_market(m)
+            if rand() > 0.5
+                t = Float64.(t)
+            end
 
-            X, vX = optimalportfolio_enumerate(f, t, g, H)
+            X, vX = optimalportfolio_enumerate(f, t, h)
             sort!(X)
-            W, VW = optimalportfolio_valuationtable(f, t, g, H)
-            Y, vY = optimalportfolio_dynamicprogram(f, t, g, H)
-            # Without memoization
-            Z, vZ = optimalportfolio_dynamicprogram(f, t, g, H, false)
-
+            W, vW = applicationorder(f, t, h; datastructure = :heap)
+            Y, vY = applicationorder(f, t, h; datastructure = :dict)
             @test X == sort(W)
-            @test vX ≈ VW[m, H]
[…]

### 2022-02-28 paper/paper.tex
tests, use smaller int types where safe

--- 2022-02-27 paper.tex
@@
 As Chao (2014) remarked, this represents a somewhat subtle portfolio optimization problem.
-The traditional Markowitz model trades off the expected value across all assets with a risk term, obtaining a concave maximization problem with linear constraints.
-But college applicants maximize the observed value of their \emph{best} asset: If a student is admitted to her $j$th choice, then she is indifferent to whether or not she gets into her $(j+1)$th choice.
+Traditional portfolio optimization models weigh the expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Meucci 2005).
+But college applicants maximize the observed value of their \emph{best} asset: If a student is admitted to her $j$th choice, then she is indifferent as to whether she gets into her $(j+1)$th choice.
 As a result, the valuation function that students maximize is \emph{convex} in the expected utility associated with individual applications.
 Risk management is implicit in the college application problem because, in a typical admissions market, college preferability is negatively correlated with competitiveness.
-Thus, students must negotiate a tradeoff between highly attractive, selective “reach schools” and less preferable “safety schools” where admission is a safer bet.
+That is, students  negotiate a tradeoff between highly attractive, selective “reach schools” and less preferable “safety schools” where admission is a safer bet.
 Finally, the combinatorial nature of the college application problem makes it difficult to solve using the gradient-based techniques used in continuous portfolio optimization.
@@
 
-% Further references to nestedness and greedy approximation algos
+Our study takes special interest in the validity of greedy solution algorithms, such as those that iteratively add the asset that yields the greatest increase in the objective function until the budget is exhausted.
+For certain classes of optimization problems, such as maximizing a submodular set function over a cardinality constraint, the greedy algorithm is known to be a good approximate solution and exact under certain additional assumptions (Fisher et al.
+1978).
+For other problems, most notably the binary knapsack problem, the most obvious greedy algorithm can be made to perform arbitrarily poorly (Vazirani 2001).
+
+We show analogous results for the college application problem: In the special case where each $g_j = 1$, the optimal portfolio satisfy a nestedness property that is equivalent to the validity of the greedy algorithm.
+As Rozanov and Tamir (2020) remark, the nestedness property is useful not only for obtaining solution algorithms, but in the implementation of an optimal policy under uncertain information.
+In the United States, many college applications are due at the beginning of November, and it is typical for students to begin working on their applications during the prior summer because colleges increasingly expect students to tailor their essays to the target school.
+Therefore, students may not know how many schools they can afford to apply to (in terms of both time and monetary costs) until late October.
+The nestedness property implies that students can prioritize their application strategy by working on applications in the order that they appear in the sequence of optimal application portfolios.
+
+Unfortunately, the nestedness property does not hold in the general case, nor does the greedy algorithm offers any performance guarantee.
+We offer an alternative approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in fully polynomial time.
 
 \subsection{Structure of this paper}
-Section \label{preliminaries} introduces some additional notation and assumptions that can be imposed with trivial loss of generality.
+Section \ref{preliminaries} introduces some additional notation and assumptions that can be imposed with trivial loss of generality.
 
@@
 
+In section \ref{numericalexperiments}, we present the results of computational experiments that confirm the validity and time complexity results established in the previous two sections.
+
 % \footnote{This assumption is appropriate when $f$ gives the admissions probabilities \emph{specifically} for Alma, as estimated from her academic abilities.
@@
 \section{Notation and preliminaries} \label{preliminaries}
-Before discussing the optimal solution, we organize a few preliminary results.
+Before discussing the solution algorithms, we will introduce some additional notation and a few preliminary results that will come in handy.
 
-The set $\mathcal{X} \subseteq \mathcal{C}$ of schools to which a student applies is called her \emph{application portfolio.} The expected utility the student receives from $\mathcal{X}$ is called its \emph{valuation}.
+We refer to the set $\mathcal{X} \subseteq \mathcal{C}$ of schools to which a student applies is called her \emph{application portfolio.} The expected utility the student receives from $\mathcal{X}$ is called its \emph{valuation}.
 %Let $x$ denote the same encoded as a binary vector, where $x_j=1 \iff j\in \mathcal{X}$ for $j=1\dots m$.
@@
 It is helpful to define the random variable $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$ as the utility achieved by the schools in the portfolio, so that when $t_0 = 0$, $v(\mathcal{X}) = \operatorname{E}[X]$.
+Similar pairs of variables with italic and script names such as $\mathcal{Y}_h$ and $Y_h$ carry an analogous meaning.
 
-Let $p_j(\mathcal{X})$ denote the probability that the student attends $c_j$.
+Given an application portfolio, let $p_j(\mathcal{X})$ denote the probability that the student attends $c_j$.
 This occurs if and only if she \emph{applies} to $c_j$, is \emph{admitted} to $c_j$, and is \emph{rejected} from any school she prefers to $c_j$; that is, any school with higher index.
@@
 \begin{align}
-v(\mathcal{X}) &= \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)  \label{closedformportfoliovaluationX}%, \quad \text{or equivalently,}\\
+v(\mathcal{X}) &= \sum_{j=0}^m t_j p_j(\mathcal{X}) = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)  \label{closedformportfoliovaluationX}%, \quad \text{or equivalently,}\\
 %\qquad v(x) &= t_0 \prod_{j=1}^m (1 - f_{j} x_j) + \sum_{j=1}^m \Bigl( x_j t_j f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’}) \Bigr) \label{closedformportfoliovaluationx}
@@
 \ifEN \section{Homogeneous application costs}  \else \subsection{} \fi \label{homogappcosts}
-Let us consider the special case in which $g_j = 1$ and $H$ is a constant $h \leq m$.
+In this section, we derive a polynomial-time algorithm for the special case in which $g_j = 1$ and $H$ is a constant $h \leq m$.
 This case is similar to the centralized college admissions process in Korea, where there is no application fee, but by law, students are allowed to apply to no more than $h$ schools.
-(In the Korean case, $m=202$ and $h=3$.) Applying Theorem \ref{assumetzerozero}, we assume that $t_0 = 0$.
-
-In this section, we will call the applicant Alma, and refer to the corresponding optimization problem as Alma's problem.
+(In the Korean case, $m=202$ and $h=3$.) Applying Theorem \ref{assumetzerozero}, we assume that $t_0 = 0$ unless otherwise noted.
+Throughout this section, we will call the applicant Alma, and refer to the corresponding optimization problem as Alma's problem.
 
@@
 \begin{theorem}[Nestedness of optimal application portfolios] \label{nestedapplication}
-There exists a sequence of portfolios $\{\mathcal{X}_h\}_{h=1}^m$ such that each $\mathcal{X}_h$ is an optimal application portfolio when the application limit is $h$, and the sequence satisfies the nestedness relation
+There exists a sequence of portfolios $\{\mathcal{X}_h\}_{h=1}^m$ satisfying the nestedness relation 
 \begin{equation}
@@
 \end{equation}
+such that each $\mathcal{X}_h$ is an optimal application portfolio when the application limit is $h$.
 \end{theorem}
@@
 
-In our numerical experiments, we found it effective to store $\mathcal{C}$ as a binary max heap rather than a list.
+In our numerical experiments,
+%whose results are reported in section \ref{numericalexperiments},
+we found it effective to store $\mathcal{C}$ as a binary max heap rather than a list.
 The heap is ordered according to the criterion $i \geq j \iff f_i t_i \geq f_j t_j$.
@@
 The cost of updating each $t_j$-value can be reduced to unit time using a Fibonacci heap (Fredman and Tarjan 1987), yielding the same overall computation time.
-%The computation time can be further reduced by observing that the school added at each iteration must be maximal in the tuple $(f_j, t_j)$.
-That is, if $f_j \geq f_k \wedge t_j \geq t_k$ holds for two candidate schools $j$ and $k$, then $k$ cannot be the uniquely optimal addition, since admission to $j$ is both weakly preferable and weakly more probable.
-Therefore, we may restrict our attention to the subset of schools \emph{not} dominated by another in this way.
-%\begin{proposition}[Efficient frontier for Alma's problem]
-%Given a set of schools $\mathcal{D}$, define the set
-%\begin{equation}
-%\operatorname{EF}(\mathcal{D})  = \{j \in \mathcal{D}: f_j \geq f_k \vee t_j \geq t_k, \forall k \in \mathcal{D}\}
-%\end{equation}
-%as the \emph{efficient frontier} of $\mathcal{D}$. Then
-%\begin{equation} l \in \operatorname{EF}(\mathcal{C}\setminus  \mathcal{X}_{h-1}) \end{equation}
-%where $l$ is the school added at iteration $h$ of Algorithm \ref{dynamicprogram}.
-%\end{proposition}
 
@@
 
-To wrap up, we provide an example showing that if the entries of $Z$ are dependent, then the optimal solution may violate the nestedness property of Theorem \ref{nestedapplication}.
-\begin{example}
-Let $t = (3, 3, 4)$, $Z_1 \sim \operatorname{Bernoulli}(0.5)$, $Z_2 = 1 - Z_1$,  and $Z_3 \sim \operatorname{Bernoulli}(0.5)$.
+%To wrap up, we provide an example showing that if the entries of $Z$ are dependent, then the optimal solution may violate the nestedness property of Theorem \ref{nestedapplication}.
+%\begin{example}
+%Let $t = (3, 3, 4)$, $Z_1 \sim \operatorname{Bernoulli}(0.5)$, $Z_2 = 1 - Z_1$,  and $Z_3 \sim \operatorname{Bernoulli}(0.5)$.
 Then it is easy to verify that the unique optimal portfolios are $\mathcal{X}_1 = \{3\}$ and $\mathcal{X}_2 = \{1, 2\}$.
-\end{example}
+%\end{example}
 
@@
 The optima for Ellis's problem are not necessarily nested, nor is the number of schools in the optimal portfolio necessarily increasing in $H$. For example, if
-$f = (0.5, 0.5, 0.5)$, $t = (1, 1, 1003)$, and $g = (1, 1, 3)$,
+$f = (0.5, 0.5, 0.5)$, $t = (1, 1, 219)$, and $g = (1, 1, 3)$,
 then it is evident that the optimal portfolio for $H = 2$ is $\{1, 2\}$ while that for $H = 3$ is $\{3\}$.
@@
 > \sum_{j \in \mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - f_{j’}) \Bigr)
-%= \sum_{j \in \mathcal{X}} \Bigl( u_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - \delta) \Bigr)
 = v(\mathcal{X}) \\
@@
 An intuitive extension of the greedy algorithm for Alma's problem is to iteratively add to $\mathcal{X}$ the school $k$ for which $[ v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) ] / g_k$ is largest.
-However, this algorithm offers no guaranteed approximation ratio, as the following example shows.
+However, the construction above shows that the objective function of Ellis's problem can approximate that of a knapsack problem with arbitrary precision.
+Therefore, in pathological examples such as the following, the greedy algorithm can achieve an arbitrarily poor approximation ratio.
 \begin{example}
-Let $t = (2, 2022)$, $f = (1, 1)$, $g = (1, 10000)$, and $H = 10000$. Then the greedy approximation algorithm produces a clearly inoptimal solution.
+Let $t = (10, 2021)$, $f = (1, 1)$, $g = (1, 500)$, and $H = 500$. Then the greedy approximation algorithm produces the clearly inoptimal solution $\mathcal{X} = \{1\}$. 
 \end{example} 
@@
 Index schools in ascending order by $t$\;
-\SetKwProg{Fn}{function}{ do}{end}
-\Fn{$V(j, h)$}{
-	\lIf{$h < g_j$}{\Return{$-\infty$}}
-	\lElseIf{$j = 0$ or $h=0$}{\Return{$0$}}
-	\lElse{\Return{$\max\bigl\{ V(j-1, h), (1 - f_j) V(j-1, h-g_j) + f_j t_j \bigr\}$}}
-}
-%\lFor{$j=1\dots m$}
-%{$\operatorname{push!}\bigl(\mathtt{H}, (j, f_j, t_j)\bigr)$}
-%\For{$i=1\dots h$}
-%{
-%    $\mathtt{D} \gets \operatorname{pop!}(\mathtt{H})$\;
-%    $\operatorname{append!}(\mathtt{X}, \mathtt{D.j})$\;
-%     \lIfElse{$i=1$}{$\operatorname{append!}(\mathtt{V}, \mathtt{D.f}\cdot\mathtt{D.t})$}
-%     {$\operatorname{append!}(\mathtt{V}, \mathtt{V[i-1]} + \mathtt{D.f}\cdot\mathtt{D.t})$}
-%    \For{$\mathtt{E} \in \mathtt{H}$}
-%	{
-%	\lIfElse{$\mathtt{E.j} < \mathtt{D.j}$}{$\mathtt{E.t} \gets  \mathtt{E.t} \cdot(1 -  \mathtt{D.f})$}{$\mathtt{E.t} \gets  \mathtt{E.t} -  \mathtt{D.f}\cdot\mathtt{D.t}$}
-%	}
+%\SetKwProg{Fn}{function}{ do}{end}
+%\Fn{$V(j, h)$}{
+%	\lIf{$h < g_j$}{\Return{$-\infty$}}
+%	\lElseIf{$j = 0$ or $h=0$}{\Return{$0$}}
+%	\lElse{\Return{$\max\bigl\{ V(j-1, h), (1 - f_j) V(j-1, h-g_j) + f_j t_j \bigr\}$}}
 %}
+Fill a lookup table with the entries of $V[j, h]$\; \label{Vcreatedlookuptable}
 $h \gets H$\;
@@
 \For{$j = m, m-1, \dots, 1$}{
-	\If{$V(j-1, h) < V(j, h)$}{
+	\If{$V[j-1, h] < V[j, h]$}{
 		$\mathcal{X} \gets \mathcal{X}\cup\{j\}$\; 
@@
 Sorting $t$ is $O(m \log m)$.
-The number of indices at which $V(j, h)$ is nontrivially defined is $O(Hm)$.
-To prevent redundant calls in the recursive function $V(j, h)$, its values must be stored in a dictionary as they are computed; this dictionary’s size is $O(Hm)$.
+The bottleneck step is the creation of the lookup table for $V[j, h]$ in line \ref{Vcreatedlookuptable}.
+Each entry is generated in unit time, and the size of the table is $O(Hm)$.
 \end{proof}
@@
 \begin{equation}
-\mathcal{V} = \Bigl\{0, 1\times 10^{-P}, 2\times 10^{-P}, \dots, r\bigl[\bar U- 1\times 10^{-P}\bigr], r\bigl[\bar U\bigr], r\bigl[\bar U+ 1\times 10^{-P}\bigr]\Bigr\}
+\mathcal{V} = \Bigl\{0, 1\times 10^{-P}, 2\times 10^{-P}, \dots, r\bigl[\bar U- 1\times 10^{-P}\bigr], r\bigl[\bar U\bigr]\Bigr\}
 \end{equation}
-Then $|\mathcal{V} | = \bar U \times 10^P + 2$.
+Then $|\mathcal{V} | = \bar U \times 10^P + 1$.
 
@@
 Set precision $P \gets \bigl\lceil\log_{10}\left(m^2 / \varepsilon \bar U\right)\bigr\rceil$\;
-\SetKwProg{Fn}{function}{ do}{end}
-\Fn{$G(j, v)$}{
-	\lIf{$v \leq 0$}{\Return{$0$}}
-	\lElseIf{$j=0$ or $t_j < v$}{\Return{$\infty$}}
-	\lElse {\Return{$\min\bigl\{ G(j-1, v), g_j + G\bigl(j-1, v -  \Delta_j(v)\bigr) \bigr\}$}}
-}
-$v\gets  \max\{ w \in \mathcal{V} : G(m, w) \leq H\}$\; \label{vrecordedhere}
+%\SetKwProg{Fn}{function}{ do}{end}
+%\Fn{$G(j, v)$}{
+%	\lIf{$v \leq 0$}{\Return{$0$}}
+%	\lElseIf{$j=0$ or $t_j < v$}{\Return{$\infty$}}
+%	\lElse {\Return{$\min\bigl\{ G(j-1, v), g_j + G\bigl(j-1, v -  \Delta_j(v)\bigr) \bigr\}$}}
+%}
+Fill a lookup table with the entries of $G[j, h]$\; \label{createdlookuptable}
+$v\gets  \max\{ w \in \mathcal{V} : G[m, w] \leq H\}$\; \label{vrecordedhere}
 $\mathcal{X} \gets \O$\;
 \For{$j = m, m-1, \dots, 1$}{
-	\If{$G(j, v)< \infty$ and $G(j, v) < G(j-1, v)$}{
+	\If{$G[j, v]< \infty$ and $G[j, v] < G[j-1, v]$}{
 		$\mathcal{X} \gets \mathcal{X}\cup\{j\}$\; 
@@
 \begin{theorem}[Validity of Algorithm \ref{ellisDP3}]
-Algorithm \ref{ellisDP3} produces a $(1 - \varepsilon)$-optimal application portfolio for Ellis's problem. It uses $O(m^3 /\varepsilon)$ time and $O(m^3/\varepsilon)$ space.
+Algorithm \ref{ellisDP3} produces a $(1 - \varepsilon)$-optimal application portfolio for Ellis's problem in $O(m^3 /\varepsilon)$ time. % and $O(m^3/\varepsilon)$ space.
 % in $O(D m + m \log m)$ time and $O(D m)$ space, where $D$ is the product of the denominators of the $f_j$-values.
@@
 
-We can define $v'(\mathcal{X})$ analogously as the fixed-point valuation of $\mathcal{X}$ when its elements are added in index order and its value is updated and rounded down at each addition in accordance with \eqref{solvemeforvwjvminusj}.
+We can define $v'(\mathcal{X})$ analogously as the fixed-point valuation of $\mathcal{X}$ when its elements are added in index order and its value is updated and rounded down to the nearest multiple of $10^{-P}$ at each addition in accordance with \eqref{solvemeforvwjvminusj}.
 By the same logic,
@@
 
-(Computational resources.) In the worst case, the number of indices at which $G(j, v)$ is evaluated is $O(|\mathcal{V}| m)$. Since
+(Computational resources.) The bottleneck step is the creation of the lookup table in line \ref{createdlookuptable}, whose size is $O(|\mathcal{V}| m)$. Since
 \begin{equation}
-|\mathcal{V}| = \bar U \times 10^{P} + 2 = \bar U \times 10^ { \bigl\lceil\log_{10}\left(m ^2/ \varepsilon \bar U\right)\bigr\rceil} + 2
+|\mathcal{V}| = \bar U \times 10^{P} + 1 = \bar U \times 10^ { \bigl\lceil\log_{10}\left(m ^2/ \varepsilon \bar U\right)\bigr\rceil} + 1
 \leq\frac{m^2}{\varepsilon} \times \text{const.}
@@
 is $O(m^2/ \varepsilon)$, the time complexity is as promised.
-To prevent redundant recursive calls to $G(j, v)$, we need to record its values in a dictionary as they are computed; the space required by the dictionary is $O(|\mathcal{V}| m) = O(m^3 / \varepsilon)$.
+%To prevent redundant recursive calls to $G(j, v)$, we need to record its values in a dictionary as they are computed; the space required by the dictionary is $O(|\mathcal{V}| m) = O(m^3 / \varepsilon)$.
 \end{proof}
 
-Since these bounds are polynomial in $m$ and $1 / \varepsilon$, Algorithm \ref{ellisDP3} is a fully polynomial-time approximation scheme for Ellis's problem.
+Since these bounds are polynomial in $m$ and $1 / \varepsilon$, Algorithm \ref{ellisDP3} is a fully polynomial-time approximation scheme for Ellis's problem (Vazirani 2001).
+
+Algorithms \ref{ellisDP1} and \ref{ellisDP3} can be written using recursive functions instead; however, since each function references itself \emph{twice,} the function values at each index must be recorded in a lookup table or otherwise memoized to prevent an exponential number of calls from forming on the stack.
+
+\section{(WIP) Numerical experiments} \label{numericalexperiments}
+In this section, we present the results of numerical experiments designed to confirm the time complexity results established above.
+In both experiments, markets were generated by drawing $t_j$ independently from an exponential distribution with a scale parameter of ten and rounding up to the nearest integer.
+To achieve partial negative correlation between $t_j$ and $f_j$, we then set $f_j = 1 / (t_j + 10 + Q)$, where $Q$ is a uniform draw from interval $[0, 1)$.
+In Experiment 1, which concerns Algorithm \ref{algorithmforlargeh}, we take each $g_j = 1$ and set $H = h = \lfloor m/ 2 \rfloor$.
+In Experiment 2, which concerns Algorithms \ref{ellisDP1} and \ref{ellisDP3}, each $g_j$ is drawn uniformly from the set $\{5, \dots, 10\}$ and we take $H = \lfloor \frac{1}{2} \sum g_j \rfloor$.
+At each combination of the experimental variables, we generated 50 markets, and each computation was repeated three times, with the fastest of the three  recorded as the computation time.
+Therefore, each cell of each table represents 150 computations.
+We report the mean and standard deviation across the 50 markets.
+Where applicable, we do not count the time required to sort the entries of $t$.
+
+The dynamic programs were implemented using recursive functions and memoization.
+Our implementation of Algorithm \ref{ellisDP3} differed from the text in that portfolio valuations were recorded as fixed-point \emph{binary} numbers instead, with the definitions of $P$ and $\mathcal{V}$ modified accordingly.
 
@@
 
+Meucci, Attilio. 2005. \emph{Risk and Asset Allocation.} Berlin: Springer-Verlag, 2005.
+
 Othman, Abraham, Eric Budish, and Tuomas Sandholm. 2010. ``Finding Approximate Competitive Equilibria: Efficient and Fair

### 2022-02-28 benchmark/benchmark.jl

@@ -1,8 +1,7 @@
 using OptimalApplication
 using DataFrames
 using Random
-using StatsBase
-using Base.Threads
+using Statistics
 
 # An all-day benchmark; tweak parameters of run with caution.
 
@@ -23,12 +22,12 @@ function samecosts(ds, M, n_markets)
     sizes = Int64[]
     times = Float64[]
 
-    @threads for i in 1:n_markets
+    for i in 1:n_markets
         println("  i = $i of $n_markets")
         for m in M
             f, t, = make_correlated_market(m)
             push!(sizes, m)
-            push!(times, minimum(@elapsed applicationorder(f, t; datastructure = ds) for r in 1:n_reps))
+            push!(times, 1000 * minimum(@elapsed applicationorder(f, t; datastructure = ds) for r in 1:n_reps))
         end
     end
 
@@ -40,12 +39,12 @@ function dp_H(M, n_markets)
     sizes = Int64[]
     times = Float64[]
 
-    @threads for i in 1:n_markets
+    for i in 1:n_markets
         println("  i = $i of $n_markets")
         for m in M
             f, t, g, H = make_correlated_market(m)
             push!(sizes, m)
-            push!(times, minimum(@elapsed optimalportfolio_dynamicprogram(f, t, g, H) for r in 1:n_reps))
+            push!(times, 1000 * minimum(@elapsed optimalportfolio_dynamicprogram(f, t, g, H) for r in 1:n_reps))
         end
     end
 
@@ -57,12 +56,12 @@ function fptas(epsilon, M, n_markets)
     sizes = Int64[]
     times = Float64[]
 
-    @threads for i in 1:n_markets
+    for i in 1:n_markets
         println("  i = $i of $n_markets")
         for m in M
             f, t, g, H = make_correlated_market(m)
             push!(sizes, m)
-            push!(times, minimum(@elapsed optimalportfolio_fptas(f, t, g, H, epsilon) for r in 1:n_reps))
+            push!(times, 1000 * minimum(@elapsed optimalportfolio_fptas(f, t, g, H, epsilon) for r in 1:n_reps))
         end
     end
 
@@ -77,37 +76,38 @@ function table1()
     n_markets = 50
 
     println("Timing with dictionary")
-    het_dict = samecosts(:dict, M, n_markets)
+    grouped_het_dict = groupby(samecosts(:dict, M, n_markets), :m)
 
[…]

### 2022-02-28 src/OptimalApplication.jl

@@ -7,7 +7,7 @@ using FixedPointNumbers
 import Base.isless
 
 
-export 
+export
     valuation,
     applicationorder,
     optimalportfolio_dynamicprogram,
@@ -15,6 +15,8 @@ export
     optimalportfolio_valuationtable,
     optimalportfolio_fptas
 
+
+const IntTypes = [Int8, Int16, Int32, Int64]
 
 
 """
@@ -42,10 +44,10 @@ Produce the optimal order of application for the market defined by admissions pr
 and utility values `t`. Stores college data in a binary `:heap` or as a `:dict`. 
 """
 function applicationorder(
-        f::Vector{Float64},
-        t::Vector{<:Real},
-        h = nothing::Union{Int64,Nothing};
-        datastructure = :heap::Symbol)::Tuple{Vector{Int64}, Vector{Float64}}
+    f::Vector{Float64},
+    t::Vector{<:Real},
+    h = nothing::Union{Int64,Nothing};
+    datastructure = :heap::Symbol)::Tuple{Vector{Int64},Vector{Float64}}
     m = length(f)
     @assert m == length(t)
     @assert issorted(t)
@@ -116,9 +118,9 @@ Returns the expected value of the portfolio `X`, a vector of school indices, on 
 the market defined by admissions probabilities `f` and utility values `t`. 
 """
 function valuation(
-        X::Vector{Int64},
-        f::Vector{<:Real},
-        t::Vector{<:Real})::Float64
+    X::Vector{Int64},
+    f::Vector{<:Real},
+    t::Vector{<:Real})::Float64
     @assert issorted(t)
     sort!(X)
     h = length(X)
@@ -147,9 +149,9 @@ Produce the optimal portfolio of size `h` on the market defined by admissions pr
 and utility values `t`. Solves by enumeration. 
 """
 function optimalportfolio_enumerate(
-        f::Vector{<:Real},
-        t::Vector{<:Real},
-        h::Int64)::Tuple{Vector{Int64},Float64}
+    f::Vector{<:Real},
+    t::Vector{<:Real},
+    h::Int64)::Tuple{Vector{Int64},Float64}
     m = length(t)
     X = zeros(Int64, h)
     v = 0.0
[…]

### 2022-02-28 src/OptimalApplication.jl
fixed T1 bound

@@ -338,7 +338,7 @@ function optimalportfolio_fptas(
     P = ceil(Int64, log2(m^2 / (ε * Ū)))
     infty = sumg + 1
 
-    T1 = IntTypes[findfirst(Ū * 2^P < typemax(T) for T in IntTypes)]
+    T1 = IntTypes[findfirst(4 * Ū * 2^P < typemax(T) for T in IntTypes)]
     T2 = IntTypes[findfirst(m < typemax(T) for T in IntTypes)]
     T3 = IntTypes[findfirst(infty < typemax(T) for T in IntTypes)]
 

### 2022-03-01 benchmark/benchmark.jl
fixed DomainError when comparing Fixed{} to Int; add @test_throws

@@ -3,10 +3,17 @@ using DataFrames
 using Random
 using Statistics
 
-# An all-day benchmark; tweak parameters of run with caution.
+fullscale = false
+
+# An all-day benchmark; tweak parameters with caution.
+# Set fullscale = false to run a smaller benchmark to check formatting etc.
 
 # Number of times to repeat each computation, where min of these is reported as time
-n_reps = 3
+n_reps = fullscale ? 3 : 1
+
+function printheader(s)
+    printstyled(s*"\n", bold=true, color=80)
+end
 
 function make_correlated_market(m)
     A = 10
@@ -71,9 +78,9 @@ end
 
 
 function table1()
-    println("== Homogeneous cost algorithms ==")
-    M = [5, 50, 500, 5000]
-    n_markets = 50
+    printheader("Homogeneous-cost algorithms")
+    M = fullscale ? [5, 50, 500, 5000] : [5, 10]
+    n_markets = fullscale ? 50 : 10
 
     println("Timing with dictionary")
     grouped_het_dict = groupby(samecosts(:dict, M, n_markets), :m)
@@ -90,9 +97,9 @@ end
 
 
 function table2()
-    println("== Heterogeneous cost algorithms ==")
-    M = [5, 50, 500]
-    n_markets = 50
+    printheader("Heterogeneous-cost algorithms")
+    M = fullscale ? [5, 50, 500] : [5, 10]
+    n_markets = fullscale ? 50 : 10
     epsilons = [0.5, 0.05]
 
     println("Timing dynamic program")
@@ -109,5 +116,5 @@ function table2()
 end
 
 display(table1())
-println()
+println("\n")
 display(table2())

### 2022-03-01 src/OptimalApplication.jl

@@ -2,7 +2,6 @@ module OptimalApplication
 
 using Combinatorics: combinations, multiset_combinations
 using DataStructures
-# For implementing the FPTAS for Ellis's problem
 using FixedPointNumbers
 import Base.isless
 
@@ -37,6 +36,73 @@ end
 isless(c1::College, c2::College) = isless(c1.ft, c2.ft)
 
 
+function iscoherentmarket(f::Vector{<:Real}, t::Vector{<:Real})
+    @assert length(f) == length(t)
+    @assert all(0 .< f .≤ 1)
+    @assert issorted(t)
+end
+
+
+function iscoherentmarket(f::Vector{<:Real}, t::Vector{<:Real}, g::Vector{<:Real})
+    @assert length(f) == length(t)
+    @assert length(f) == length(g)
+    @assert all(0 .< f .≤ 1)
+    @assert issorted(t)
+end
+
+
+
+function isnontrivialmarket(f::Vector{<:Real}, t::Vector{<:Real}, h::Int)
+    iscoherentmarket(f, t)
+    @assert 0 < h ≤ length(t)
+end
+
+
+function isnontrivialmarket(
+        f::Vector{<:Real},
+        t::Vector{<:Real},
+        g::Vector{<:Real},
+        H::Real)
+    iscoherentmarket(f, t, g)
+    @assert 0 < H ≤ sum(g)
+    @assert all(0 .< g .≤ H)
+end
+
+
+
+"""
+    valuation(X, f, t)
+
+Returns the expected value of the portfolio `X`, a vector of school indices, on the
+the market defined by admissions probabilities `f` and utility values `t`. 
+"""
+function valuation(
+        X::Vector{Int64},
+        f::Vector{<:Real},
+        t::Vector{<:Real})::Float64
+    isempty(X) && return 0.0
+    iscoherentmarket(f, t)
+
+    sort!(X)
+    h = length(X)
+
+    if h > 1
+        res = 0.0
+        cp = reverse(cumprod(1 .- reverse(f[X[2:end]])))
+
+        for j in 1:h-1
+            res += t[X[j]] * f[X[j]] * cp[j]
+        end
+
+        res += t[X[end]] * f[X[end]]
+
+        return res
+    else
+        return f[X[1]] * t[X[1]]
+    end
+end
+
+
 """
     applicationorder(f, t, h=m; datastructure=:heap)
 
[…]

### 2022-03-01 test/runtests.jl

@@ -77,4 +77,53 @@ end
             end
         end
     end
+
+    @testset verbose = true "Large problems" begin
+        f, t, g, H = make_correlated_market(500)
+
+        W, vW = optimalportfolio_fptas(f, t, g, H, 0.5)
+        Y, vY = optimalportfolio_dynamicprogram(f, t, g, H)
+        @test vW / vY ≥ 0.5
+
+        f, t, g, H = make_correlated_market(5000)
+
+        X, V = applicationorder(f, t, 2500)
+        @test !isnothing(X)
+    end
+
+    @testset verbose = true "Trivial and malformed markets" begin
+        # Dim mismatch
+        f = [0.1, 0.1]
+        t = [1, 2, 4]
+        g = [2, 2]
+        H = 3
+
+        @test_throws AssertionError optimalportfolio_enumerate(f, t, 1)
+        @test_throws AssertionError optimalportfolio_enumerate(f, t, g, H)
+
+        # t not sorted
+        f = [0.1, 0.1]
+        t = [7.0, 4.0]
+        g = [2, 2]
+        H = 3
+
+        @test_throws AssertionError optimalportfolio_enumerate(f, t, 1)
+        @test_throws AssertionError optimalportfolio_enumerate(f, t, g, H)
+
+        # f not in (0, 1]
+        f = [5, 1]
+        t = [4, 7]
+
+        @test_throws AssertionError optimalportfolio_enumerate(f, t, 1)
+        @test_throws AssertionError optimalportfolio_enumerate(f, t, g, H)
+
+        # Some g[j] > H
+        f = [0.1, 0.1]
+        t = [4, 7]
+        g = [10, 10]
+        H = 5
+
+        @test_throws AssertionError optimalportfolio_fptas(f, t, g, H, 0.5)
+        @test_throws AssertionError optimalportfolio_dynamicprogram(f, t, g, H)
+    end
 end

### 2022-03-01 paper/paper.tex
intro

@@
 \fancyhf{}
-\rhead{Uncertainty in college admissions markets \\~}
+\rhead{How to apply to college \\~}
 \lhead{Max Kapur \\ \date{\today}}
@@
 \ifEN \section{Introduction}  \else \section{서론} \fi
-This paper considers the optimization problem
+This paper considers the portfolio optimization problem
 %\begin{definition}[Optimal college application portfolio] \label{generalproblemstatement}
@@
 \begin{split}
-\text{maximize}\quad & \operatorname{E}\left[\max\bigr\{t_0,
-\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\right] \\
+\text{maximize}\quad & \operatorname{E}\Bigl[\max\bigr\{t_0,
+\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\Bigr] \\
 \text{subject to}\quad & \mathcal{X} \subseteq \mathcal{C}, ~~\sum_{j\in \mathcal{X}} g_j \leq H
@@
 %\end{definition}
-where $\mathcal{C} = \{ 1 \dots m\}$ is an index set, $H$ is a budget parameter, and for $j = 1 \dots m$, $g_j > 0$ is a cost parameter, $Z_j$ is a random, independent Bernoulli variable with probability $f_j$, and $t_j  > 0$ is a utility parameter.
+where $\mathcal{C} = \{ 1 \dots m\}$ is an index set, $H$ is a budget parameter, and for $j = 1 \dots m$, $g_j > 0$ is a cost parameter, $Z_j$ is a random, independent Bernoulli variable with probability $f_j \in (0, 1]$, and $t_j  > 0$ is a utility parameter.
 Without loss of generality, we assume that the $t_j$-values are weakly increasing in $j$, that $0 \leq t_0 \leq t_1$, and that each $g_j \leq H$.
 
-We refer to this problem as the \emph{optimal college application portfolio,} as follows: Consider a college admissions market with $m$ college.
+We refer to this problem as the \emph{optimal college application portfolio,} as follows: Consider a college market with $m$ colleges.
 The $j$th college is named $c_j$.
-Consider a single student, and let $f_j$ denote her probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not.
+Consider a single prospective student in this market, and let each $t_j$-value indicate the utility she associates with attending $c_j$, where her utility is $t_0$ if she does not attend college.
+Let $g_j$ denote the application fee for $c_j$ and $H$ the student's total budget to spend on application fees.
+Finally, let $f_j$ denote her probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not.
 It is appropriate to assume that the $Z_j$ are probabilistically independent as long as $f_j$ are probabilities estimated specifically for this student (as opposed to generic acceptance rates).
-Each $t_j$-value indicates the utility associated with attending $c_j$, and $t_0$ gives the utility associated with not attending college, an option named $c_0$.
-The application fee for $c_j$ is $g_j$, and $H$ represents the student's total budget to spend on application fees.
-Then the student's objective is to maximize the expected utility associated with the best school she gets into, and her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which the student applies.
+Then the student's objective is to maximize the expected utility associated with the best school she gets into within this budget.
+Her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which she applies.
 
-As Chao (2014) remarked, this represents a somewhat subtle portfolio optimization problem.
+As Chao (2014) remarked, college application represents a somewhat subtle portfolio optimization problem.
 Traditional portfolio optimization models weigh the expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Meucci 2005).
@@
 
-Our study takes special interest in the validity of greedy solution algorithms, such as those that iteratively add the asset that yields the greatest increase in the objective function until the budget is exhausted.
+We take special interest in the validity of greedy solution algorithms, such as that that iteratively adds the asset that yields the greatest increase in the objective function until the budget is exhausted.
+% Greedy algorithms produce a nested, parametric family of solutions. As Rozanov ...
 For certain classes of optimization problems, such as maximizing a submodular set function over a cardinality constraint, the greedy algorithm is known to be a good approximate solution and exact under certain additional assumptions (Fisher et al.
 1978).
-For other problems, most notably the binary knapsack problem, the most obvious greedy algorithm can be made to perform arbitrarily poorly (Vazirani 2001).
+For other problems, notably the binary knapsack problem, the most intuitive greedy algorithm can be made to perform arbitrarily poorly (Vazirani 2001).
 
 We show analogous results for the college application problem: In the special case where each $g_j = 1$, the optimal portfolio satisfy a nestedness property that is equivalent to the validity of the greedy algorithm.
-As Rozanov and Tamir (2020) remark, the nestedness property is useful not only for obtaining solution algorithms, but in the implementation of an optimal policy under uncertain information.
+As Rozanov and Tamir (2020) remark, the nestedness property is useful not only for obtaining an optimal solution, but in the implementation of an optimal policy under uncertain information.
 In the United States, many college applications are due at the beginning of November, and it is typical for students to begin working on their applications during the prior summer because colleges increasingly expect students to tailor their essays to the target school.
@@
 Unfortunately, the nestedness property does not hold in the general case, nor does the greedy algorithm offers any performance guarantee.
-We offer an alternative approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in fully polynomial time.
+Instead, we offer a pseudopolynomial-time algorithm that is tractable for typical college market instances, as well as an approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in fully polynomial time.
 
@@
 
-\section{Notation and preliminaries} \label{preliminaries}
+\section{Notation and preliminary results} \label{preliminaries}
 Before discussing the solution algorithms, we will introduce some additional notation and a few preliminary results that will come in handy.
@@
 In both experiments, markets were generated by drawing $t_j$ independently from an exponential distribution with a scale parameter of ten and rounding up to the nearest integer.
-To achieve partial negative correlation between $t_j$ and $f_j$, we then set $f_j = 1 / (t_j + 10 + Q)$, where $Q$ is a uniform draw from interval $[0, 1)$.
+To achieve partial negative correlation between $t_j$ and $f_j$, we then set $f_j = 1 / (t_j + 10 + Q)$, where $Q$ is drawn uniformly from the interval $[0, 1)$.
 In Experiment 1, which concerns Algorithm \ref{algorithmforlargeh}, we take each $g_j = 1$ and set $H = h = \lfloor m/ 2 \rfloor$.

### 2022-03-01 benchmark/benchmark.jl
rm dependency on FixedPointNumbers using scaled Ints instead

@@ -2,8 +2,9 @@ using OptimalApplication
 using DataFrames
 using Random
 using Statistics
+using Base.Threads
 
-fullscale = false
+fullscale = true
 
 # An all-day benchmark; tweak parameters with caution.
 # Set fullscale = false to run a smaller benchmark to check formatting etc.
@@ -12,7 +13,7 @@ fullscale = false
 n_reps = fullscale ? 3 : 1
 
 function printheader(s)
-    printstyled(s*"\n", bold=true, color=80)
+    printstyled(s*"\n", bold=true, color=222)
 end
 
 function make_correlated_market(m)
@@ -26,59 +27,59 @@ function make_correlated_market(m)
 end
 
 function samecosts(ds, M, n_markets)
-    sizes = Int64[]
-    times = Float64[]
+    sizes = zeros(Int64, n_markets, length(M))
+    times = zeros(Float64, n_markets, length(M))
 
-    for i in 1:n_markets
+    @threads for i in 1:n_markets
         println("  i = $i of $n_markets")
-        for m in M
+        for (j, m) in enumerate(M)
             f, t, = make_correlated_market(m)
-            push!(sizes, m)
-            push!(times, 1000 * minimum(@elapsed applicationorder(f, t; datastructure = ds) for r in 1:n_reps))
+            sizes[i, j] = m
+            times[i, j] = 1000 * minimum(@elapsed applicationorder(f, t; datastructure = ds) for r in 1:n_reps)
         end
     end
 
-    return DataFrame("m" => sizes, "time" => times)
+    return DataFrame("m" => sizes[:], "time" => times[:])
 end
 
 
 function dp_H(M, n_markets)
-    sizes = Int64[]
-    times = Float64[]
+    sizes = zeros(Int64, n_markets, length(M))
+    times = zeros(Float64, n_markets, length(M))
 
-    for i in 1:n_markets
+    @threads for i in 1:n_markets
         println("  i = $i of $n_markets")
-        for m in M
+        for (j, m) in enumerate(M)
             f, t, g, H = make_correlated_market(m)
-            push!(sizes, m)
-            push!(times, 1000 * minimum(@elapsed optimalportfolio_dynamicprogram(f, t, g, H) for r in 1:n_reps))
+            sizes[i, j] = m
[…]

### 2022-03-01 src/OptimalApplication.jl

@@ -2,7 +2,6 @@ module OptimalApplication
 
 using Combinatorics: combinations, multiset_combinations
 using DataStructures
-using FixedPointNumbers
 import Base.isless
 
 
@@ -365,38 +364,38 @@ function optimalportfolio_fptas(
     isnontrivialmarket(f, t, g, H)
     @assert 0 < ε < 1
     m = length(f)
-    infty = sum(g) + 1
 
-    ft = f .* t
+    P = ceil(Int, log2(m^2 / (ε * sum(f .* t))))
+
     omf = 1 .- f
 
-    Ū = sum(ft)
-    P = ceil(Int64, log2(m^2 / (ε * Ū)))
+    t *= 2^P
+    ft = f .* t
 
-    # Avoid DomainError for Fixed{Int8, 8}, Fixed{Int16, 16} etc
-    T1 = IntTypes[findfirst(2^(P-1) < typemax(T) && Ū + eps(Fixed{T, P}) < typemax(Fixed{T, P}) for T in IntTypes)]
-    T2 = IntTypes[findfirst(m < typemax(T) for T in IntTypes)]
-    T3 = IntTypes[findfirst(infty < typemax(T) for T in IntTypes)]
+    Ū = ceil(Int, sum(ft))
 
-    FP = Fixed{T1,P}
-    Ū = FP(Ū + eps(FP))
-    G_dict = Dict{Tuple{T2,FP},T3}()
+    infty = sum(g) + 1
 
-    # Want to assert v::FP here but gives a strange error
-    function G(j::Int, v::Real)
+    vType = IntTypes[findfirst(Ū < typemax(T) for T in IntTypes)] # Ū + 1 ?
+    jType = IntTypes[findfirst(m < typemax(T) for T in IntTypes)]
+    gType = IntTypes[findfirst(infty < typemax(T) for T in IntTypes)]
+
+    Ū = vType(Ū)
+    G_dict = Dict{Tuple{jType,vType},gType}()
+
+    function G(j::Integer, v::Integer)
         haskey(G_dict, (j, v)) && return G_dict[(j, v)]
 
         if v ≤ 0
             return 0
-        elseif j == 0 || t[j] < floor(Int, v) || v ≥ Ū
-            # push!(G_dict, (j, v::FP) => infty)
-            return infty # G_dict[(j, v)]
+        elseif j == 0 || t[j] < v || v ≥ Ū
+            return infty
         else
             if f[j] < 1
                 # Clamping prevents over/underflow: for any v≤0 or v>Ū the function
[…]

### 2022-03-01 test/runtests.jl

@@ -91,7 +91,7 @@ end
         @test !isnothing(X)
     end
 
-    @testset verbose = true "Trivial and malformed markets" begin
+    @testset verbose = true "Bad markets" begin
         # Dim mismatch
         f = [0.1, 0.1]
         t = [1, 2, 4]

### 2022-03-02 benchmark/benchmark.jl
parallelize benchmark

@@ -116,7 +116,9 @@ function table2()
         ["fptas_$(epsilons[i])_$f" => combine(fptas_times[i], :time => f)[!, Symbol("time_$f")] for f in [mean, std], i in 1:length(epsilons)]...)
 end
 
-println()
-display(table1())
-println("\n")
-display(table2())
+@time begin
+    println()
+    display(table1())
+    println("\n")
+    display(table2())
+end

### 2022-03-02 paper/paper.tex
benchmark and assoc writeup

@@
 % Language selector
-\newif\ifEN
-\ENtrue % Change to false to export Korean
-%\ENfalse
+\newif\ifen
+%\enfalse % Change to false to export Korean
 
-\ifEN
+\ifen
 {
@@
 
-%\ifEN {} \else \renewcommand*{\proofname}{증명\@}  \fi
+\ifen {} \else \renewcommand*{\proofname}{증명\@}  \fi
 
@@
 
-\ifEN{ \title{How to apply to college} }
+\ifen { \title{How to apply to college} }
 \else {\title{How to apply to college}} \fi
@@
 \maketitle
+
 \section*{Abstract}
@@
 The first produces an exact solution in pseudopolynomial time, and the second yields a fully polynomial-time approximation scheme.
+
+\section*{요약}
+본 논문은 다수의 확률 변수로 구성된 포트폴리오의 기대 최적값을 예산 조건 하에서 최대화하는 문제를 고려한다.
+이를 최적 대학 지원 문제라고 부른다.
+각 확률 변수의 비용, 즉 각 대학의 지원비가 동일한 경우, 최적 포트폴리오는 예산 제약식으로 결정된 포함 사슬 관계를 가짐을 보이고 이에 따라 다항 시간 해법을 제시한다.
+대학의 지원비가 서로 다른 경우, 문제가 NP-complete함을 증명한다.
+일반적인 상황을 위해 두 가지 동적 계획 해법을 도출한다.
+첫째는 의사 다항 시간 안에 정확한 해를 출력하며 둘째는 전체 다항 시간 근사 해법이다.
+
 \pagebreak
@@
 \tableofcontents
+
 \pagebreak
-
-\ifEN \section{Introduction}  \else \section{서론} \fi
-This paper considers the portfolio optimization problem
+\ifen \section{Introduction}  \else \section{서론} \fi
+\ifen This paper considers the portfolio optimization problem
+\else 본 논문은 다음과 같은 최적화 문제를 고려한다.\fi
 %\begin{definition}[Optimal college application portfolio] \label{generalproblemstatement}
@@
 %\end{definition}
-where $\mathcal{C} = \{ 1 \dots m\}$ is an index set, $H$ is a budget parameter, and for $j = 1 \dots m$, $g_j > 0$ is a cost parameter, $Z_j$ is a random, independent Bernoulli variable with probability $f_j \in (0, 1]$, and $t_j  > 0$ is a utility parameter.
-Without loss of generality, we assume that the $t_j$-values are weakly increasing in $j$, that $0 \leq t_0 \leq t_1$, and that each $g_j \leq H$.
+\ifen where $\mathcal{C} = \{ 1 \dots m\}$ is an index set, $H$ is a budget parameter, and for $j = 1 \dots m$, $g_j > 0$ is a cost parameter, $Z_j$ is a random, independent Bernoulli variable with probability $f_j$.
+\else 단,  $\mathcal{C} = \{ 1 \dots m\}$은 지표 집합이며 $H$는 예산 모수이다. 각 $j = 1 \dots m$에 대해 $g_j > 0$는 비용 모수이며 $Z_j$는 확률 $f_j$를 가지는 서로 독립적인 Bernoulli 변수이다.\fi
 
+\ifen
 We refer to this problem as the \emph{optimal college application portfolio,} as follows: Consider a college market with $m$ colleges.
@@
 Her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which she applies.
+\else
+다음 해석에 따라 이를 `최적 대학 지원 포트폴리오 문제'라고 부른다.
+$m$개의 대학교를 가지는 입학 시장을 고려하자.
+$j$번째 학교의 이름은 $c_j$이다.
+이 시장에 참여하는 한 학생을 고려해서 그가 $c_j$에 진학하면 $t_j$ 단위의 효용이 발생한다고 하자.
+단, 어떤 대학에도 진학하지 않는 경우 그의 효용은 $t_0$이다.
+$c_j$의 지원비가 $g_j$이며 학생이 지원비에 쓸 수 있는 예산이 $H$라고 하자.
+마지막, $c_j$에 지원하면 학생이 합격할 확률이 $f_j$이라고 하자.
+따라서 합격하면 $Z_j = 1$, 합격 안 하면 $Z_j = 0$이 된다.
+$f_j$가 (학교의 전체적인 합격률이 아니라) 바로 이 학생의 합격 확률일 때 $Z_j$의 확률적 독립성을 적절한 가정이다.
+그러면 학생의 목표는 주어진 예산 안에서 학생이 합격하는 가장 좋은 학교의 기대 효용을 최대화하는 것이다.
+위 문제의 최적해가 $\mathcal{X}$일 때, 학생의 최적 대학 지원 전략은 $\mathcal{X}$에 지원하는 것이다.
+\fi
 
@@
 Finally, the combinatorial nature of the college application problem makes it difficult to solve using the gradient-based techniques used in continuous portfolio optimization.
-Chao estimated her model (which considers application as a \emph{cost} rather than a constraint) by clustering the schools so that $m=8$, a scale at which enumeration is possible.
+Chao estimated her model (which considers application as a \emph{cost} rather than a constraint) by clustering the schools so that $m=8$, a scale at which enumeration is tractable.
 Our study pursues a more general solution.
@@
 We take special interest in the validity of greedy solution algorithms, such as that that iteratively adds the asset that yields the greatest increase in the objective function until the budget is exhausted.
-% Greedy algorithms produce a nested, parametric family of solutions. As Rozanov ...
-For certain classes of optimization problems, such as maximizing a submodular set function over a cardinality constraint, the greedy algorithm is known to be a good approximate solution and exact under certain additional assumptions (Fisher et al.
+Greedy algorithms produce a \emph{nested} family of solutions, parameterized by the budget $H$: If $H \leq H'$, then the greedy solution for budget $H$ is a subset of the greedy solution for budget $H'$.
+As Rozanov and Tamir (2020) remark, this nestedness property is useful not only for obtaining an optimal solution, but in the implementation thereof under uncertain information.
+For example, in the United States, many college applications are due at the beginning of November, and it is typical for students to begin working on their applications during the prior summer because colleges expect students to tailor their essays to the target school.
+However, students may not know how many schools they can afford to apply to until late October.
+The nestedness property---or equivalently, the validity of a greedy algorithm---implies that even in the absence of complete budget information, students can begin to carry out the optimal application strategy by writing essays for schools in the order that they enter the maximal portfolio.
+
+For certain classes of optimization problems, such as maximizing a submodular set function over a cardinality constraint, a greedy algorithm is known to be a good approximate solution and exact under certain additional assumptions (Fisher et al.
 1978).
 For other problems, notably the binary knapsack problem, the most intuitive greedy algorithm can be made to perform arbitrarily poorly (Vazirani 2001).
-
 We show analogous results for the college application problem: In the special case where each $g_j = 1$, the optimal portfolio satisfy a nestedness property that is equivalent to the validity of the greedy algorithm.
-As Rozanov and Tamir (2020) remark, the nestedness property is useful not only for obtaining an optimal solution, but in the implementation of an optimal policy under uncertain information.
-In the United States, many college applications are due at the beginning of November, and it is typical for students to begin working on their applications during the prior summer because colleges increasingly expect students to tailor their essays to the target school.
-Therefore, students may not know how many schools they can afford to apply to (in terms of both time and monetary costs) until late October.
-The nestedness property implies that students can prioritize their application strategy by working on applications in the order that they appear in the sequence of optimal application portfolios.
-
+This case mirrors the centralized college application process in Korea, where there is no application fee, but students are allowed to apply to only three schools during the main admissions cycle.
 Unfortunately, the nestedness property does not hold in the general case, nor does the greedy algorithm offers any performance guarantee.
@@
 
-\subsection{Structure of this paper}
+\ifen \subsection{Structure of this paper} \else \subsection{본 논문의 구성}\fi
 Section \ref{preliminaries} introduces some additional notation and assumptions that can be imposed with trivial loss of generality.
@@
 In section \ref{homogappcosts}, we solve the special case where each $g_j = 1$ and $H$ is an integer $h \leq m$.
-This case mirrors the centralized college application process in Korea, where there is no application fee, but students are allowed to apply to only three schools during the main admissions cycle.
 We show that an intuitive heuristic is in fact a $1/h$-approximation algorithm.
@@
 
-\section{Notation and preliminary results} \label{preliminaries}
+\pagebreak
+\ifen \section{Notation and preliminary results} \else \section{표기법과 예비 결과}\fi\label{preliminaries}
 Before discussing the solution algorithms, we will introduce some additional notation and a few preliminary results that will come in handy.
 
-We refer to the set $\mathcal{X} \subseteq \mathcal{C}$ of schools to which a student applies is called her \emph{application portfolio.} The expected utility the student receives from $\mathcal{X}$ is called its \emph{valuation}.
+For the remainder of the paper, unless otherwise noted, we assume with trivial loss of generality that each $g_j \leq H$, each $f_j \in (0, 1]$, and $t_0 < t_1 \leq \cdots \leq t_m$.
+Below, we will show how to transform an arbitrary instance so that $t_0 = 0$, in which case each $t_j > 0$.
+
+We refer to the set $\mathcal{X} \subseteq \mathcal{C}$ of schools to which a student applies as her \emph{application portfolio.} The expected utility the student receives from $\mathcal{X}$ is called its \emph{valuation}.
 %Let $x$ denote the same encoded as a binary vector, where $x_j=1 \iff j\in \mathcal{X}$ for $j=1\dots m$.
@@
 
-\ifEN \section{Homogeneous application costs}  \else \subsection{} \fi \label{homogappcosts}
+\ifen \section{Homogeneous application costs}  \else \subsection{동일한 지원 비용} \fi \label{homogappcosts}
 In this section, we derive a polynomial-time algorithm for the special case in which $g_j = 1$ and $H$ is a constant $h \leq m$.
@@
 
-\ifEN \subsection{Approximation properties of a na\"ive solution}  \else \subsection{모형화} \fi
+\ifen \subsection{Approximation properties of a na\"ive solution}  \else \subsection{Na\"ive 해법의 근사 성질} \fi
 The expected utility associated with a single school $c_j$ is simply $\operatorname{E}[t_j Z_j] = t_j f_j$. It is therefore tempting to adopt the following strategy, which turns out to be inoptimal.
@@
 
+\begin{corollary}
+The function $v(\mathcal{X})$ is not submodular. 
+\end{corollary}
+\begin{proof}
+If $v(\mathcal{X})$ is submodular, then theorem 4.2 of Fisher et al.
+(1978) implies that the na\"ive algorithm achieves an optimality ratio of $v(\mathcal{T}_h) / v(\mathcal{X}_h)  \geq 1 - \left(\frac{h-1}{h}\right)^h$.
+Example \ref{tightexampleforoneoverhopt} provides a counterexample.
+\end{proof}
+
 Hope is not lost. We can still find the optimal solution in time polynomial in $h$ and $m$, as we will now show.
 
-\ifEN \subsection{The nestedness property}  \else \subsection{풀이} \fi
+\ifen \subsection{The nestedness property}  \else \subsection{포함 사슬 관계 성질} \fi
 It turns out that the solution to Alma's problem possesses a special structure: An optimal portfolio of size $h+1$ includes an optimal portfolio of size $h$ as a subset.
@@
 
-\subsection{Polynomial-time solution}
+\ifen \subsection{Polynomial-time solution} \else \subsection{다항 시간 해법}\fi
 Applying the result above yields an efficient algorithm for the optimal portfolio: Start with the empty set and add schools one at a time, maximizing $v(\mathcal{X}\cup \{k\})$ at each addition.
@@
 
-\subsection{Properties of the optimal portfolios}
+\ifen \subsection{Properties of the optimal portfolios} \else\subsection{최적 포트폴리오의 성질} \fi
 The nestedness property implies that Alma's expected utility is a discretely concave function of $h$.
@@
 
-\section{Heterogeneous application costs}  \label{hetappcosts}
+\pagebreak
+\ifen \section{Heterogeneous application costs} \else\section{다양한 지원 비용} \fi\label{hetappcosts}
 In this section, we turn to the more general problem in which the constant $g_j$ represents the \emph{cost} of applying to $c_j$ and the student, whom we now call Ellis, has a \emph{budget} of $H$ to spend on college applications.
@@
 
-\subsection{NP-completeness}
+\ifen \subsection{NP-completeness} \else \subsection{NP-completeness}\fi
 In fact, Ellis’s problem is NP-complete, as we will show by a transformation from the binary knapsack problem, which is known to be NP-complete (Garey and Johnson 1979).
@@
 
-\subsection{Pseudopolynomial-time dynamic program}
+\ifen \subsection{Pseudopolynomial-time dynamic program} \else\subsection{의사 다항 시간 동적 계획} \fi
 In this subsection, we assume, with a small loss of generality, that $g_j \in \mathbb{N}$ for $j = 1\dots m$ and $H \in\mathbb{N}$, and provide an algorithmic solution to Ellis's problem that runs in $O(Hm + m\log m)$ time and $O(Hm)$ space.
@@
 
-\subsection{Fully polynomial-time approximation scheme}
+\ifen \subsection{Fully polynomial-time approximation scheme}\else \subsection{전체 다항 시간 근사 해법}\fi
 As with the knapsack problem, Ellis's problem admits a complementary dynamic program that iterates on the value of the cheapest portfolio instead of on the cost of the most valuable portfolio.
@@
 By rounding this value down, we ensure that the true valuation of $\mathcal{W}[j, v]$ is \emph{at least} $v - \Delta$.
-
-When $f_j = 1$, then the singleton $\{j\}$ has $v(\{j\}) \geq v$, so
+When $t_j \geq v$ and $f_j = 1$, the singleton $\{j\}$ has $v(\{j\}) \geq v$, so
 \begin{equation}G[j, v] = \min\bigl\{G[j-1, v], g_j \bigr\}.\end{equation}
-Defining $\Delta = \infty$ in this case ensures that $G[j-1, v-\infty] = 0$ as required.
+Defining $\Delta_j(v) = \infty$ in this case ensures that $g_j + G[j-1, v-\Delta_j(v)] = g_j+ G[j-1, v-\infty] = g_j $ as required.
 %we have $t_j < v$, and since no portfolio can be more valueable than the largest $t_j$ value of its elements, we must have $G[j, v] =\infty$.
@@
 Once $G[j, v]$ has been calculated at each index, the associated portfolio can be found by applying the observation that $\mathcal{W}[j, v]$ contains $j$ if and only if $G[j, v] < G[j-1, v]$.
-Then an approximate solution to Ellis's problem is obtained by computing the optimal objective value $\max\{ w: G[m, w] \leq H\}$ and corresponding portfolio.
+Then an approximate solution to Ellis's problem is obtained by computing the largest achievable objective value $\max\{ w: G[m, w] \leq H\}$ and corresponding portfolio.
 
@@
 
-(Computational resources.) The bottleneck step is the creation of the lookup table in line \ref{createdlookuptable}, whose size is $O(|\mathcal{V}| m)$. Since
+(Computation time.) The bottleneck step is the creation of the lookup table in line \ref{createdlookuptable}, whose size is $O(|\mathcal{V}| m)$. Since
 \begin{equation}
@@
 
-\section{(WIP) Numerical experiments} \label{numericalexperiments}
+\pagebreak
+\ifen \section{(WIP) Numerical experiments} \else \section{계산적 실험}\fi\label{numericalexperiments}
 In this section, we present the results of numerical experiments designed to confirm the time complexity results established above.
 In both experiments, markets were generated by drawing $t_j$ independently from an exponential distribution with a scale parameter of ten and rounding up to the nearest integer.
-To achieve partial negative correlation between $t_j$ and $f_j$, we then set $f_j = 1 / (t_j + 10 + Q)$, where $Q$ is drawn uniformly from the interval $[0, 1)$.
+To achieve partial negative correlation between $t_j$ and $f_j$, we then set $f_j = 1 / (t_j + 10 Q)$, where $Q$ is drawn uniformly from the interval $[0, 1)$.
 In Experiment 1, which concerns Algorithm \ref{algorithmforlargeh}, we take each $g_j = 1$ and set $H = h = \lfloor m/ 2 \rfloor$.
@@
 
-The dynamic programs were implemented using recursive functions and memoization.
-Our implementation of Algorithm \ref{ellisDP3} differed from the text in that portfolio valuations were recorded as fixed-point \emph{binary} numbers instead, with the definitions of $P$ and $\mathcal{V}$ modified accordingly.
+\begin{figure}[h!]
+  \begin{center}\includegraphics[width=0.75\textwidth]{./plots/samplemarket.pdf}\end{center}
+  \caption{A typical randomly-generated instance with $m=60$ schools. }
+\end{figure}
+
+The dynamic programs, namely Algorithms \ref{ellisDP1} and \ref{ellisDP3}, were implemented using recursive functions and memoization.
+Our implementation of Algorithm \ref{ellisDP3} differed slightly from that given above: We represented portfolio valuations in \emph{binary} rather than decimal, with the definitions of $P$ and $\mathcal{V}$ modified accordingly, and instead of fixed-point numbers, we worked in integers by multiplying each $t_j$-value by $2^P$.
+These modifications yield a substantial performance improvement without changing the fundamental algorithm design or complexity analysis.
 
 \pagebreak
-\ifEN \section{References} \else \section{참고문헌} \fi
+\ifen \section{Conclusion} \else \section{결론}\fi\label{conclusion}
+
+\pagebreak
+\ifen \section{References} \else \section{참고문헌} \fi
 \noindent

### 2022-03-02 benchmark/benchmark.jl

@@ -1,19 +1,19 @@
 using OptimalApplication
 using DataFrames
 using Random
-using Statistics
+using StatsBase
 using Base.Threads
 
-fullscale = true
+fullscale = false
 
-# An all-day benchmark; tweak parameters with caution.
+# A long benchmark; tweak parameters with caution.
 # Set fullscale = false to run a smaller benchmark to check formatting etc.
 
 # Number of times to repeat each computation, where min of these is reported as time
 n_reps = fullscale ? 3 : 1
 
 function printheader(s)
-    printstyled(s*"\n", bold=true, color=222)
+    printstyled(s * "\n", bold = true, color = 222)
 end
 
 function make_correlated_market(m)
@@ -26,99 +26,82 @@ function make_correlated_market(m)
     return f, t, g, H
 end
 
-function samecosts(ds, M, n_markets)
+function benchmark1()
+    printheader("Benchmark 1: Homogeneous-cost algorithms")
+    M = fullscale ? [5, 50, 500, 5000] : [5, 10]
+    n_markets = fullscale ? 50 : 10
+
     sizes = zeros(Int64, n_markets, length(M))
-    times = zeros(Float64, n_markets, length(M))
+    times_dict = zeros(Float64, n_markets, length(M))
+    times_heap = zeros(Float64, n_markets, length(M))
 
     @threads for i in 1:n_markets
         println("  i = $i of $n_markets")
         for (j, m) in enumerate(M)
             f, t, = make_correlated_market(m)
             sizes[i, j] = m
-            times[i, j] = 1000 * minimum(@elapsed applicationorder(f, t; datastructure = ds) for r in 1:n_reps)
+            times_dict[i, j] = 1000 * minimum(@elapsed applicationorder(f, t; datastructure = :dict) for r in 1:n_reps)
+            times_heap[i, j] = 1000 * minimum(@elapsed applicationorder(f, t; datastructure = :heap) for r in 1:n_reps)
         end
     end
 
-    return DataFrame("m" => sizes[:], "time" => times[:])
-end
+    df = DataFrame("m" => sizes[:], "time_dict" => times_dict[:], "time_heap" => times_heap[:])
 
-
-function dp_H(M, n_markets)
-    sizes = zeros(Int64, n_markets, length(M))
-    times = zeros(Float64, n_markets, length(M))
-
-    @threads for i in 1:n_markets
[…]

### 2022-03-03 paper/paper.tex
branch-and-bound algorithm; not yet integrated into package

@@
 \newif\ifen
-%\enfalse % Change to false to export Korean
+\entrue % Change to false to export Korean
 
@@
 Let $g_j$ denote the application fee for $c_j$ and $H$ the student's total budget to spend on application fees.
-Finally, let $f_j$ denote her probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not.
+Lastly, let $f_j$ denote her probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not.
 It is appropriate to assume that the $Z_j$ are probabilistically independent as long as $f_j$ are probabilities estimated specifically for this student (as opposed to generic acceptance rates).
@@
 
+\ifen
 As Chao (2014) remarked, college application represents a somewhat subtle portfolio optimization problem.
-Traditional portfolio optimization models weigh the expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Meucci 2005).
-But college applicants maximize the observed value of their \emph{best} asset: If a student is admitted to her $j$th choice, then she is indifferent as to whether she gets into her $(j+1)$th choice.
+In computational finance, traditional portfolio optimization models weigh the expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Meucci 2005).
+But college applicants maximize the value of their \emph{best} asset: If a student is admitted to her $j$th choice, then she is indifferent as to whether she gets into her $(j+1)$th choice.
 As a result, the valuation function that students maximize is \emph{convex} in the expected utility associated with individual applications.
 Risk management is implicit in the college application problem because, in a typical admissions market, college preferability is negatively correlated with competitiveness.
-That is, students  negotiate a tradeoff between highly attractive, selective “reach schools” and less preferable “safety schools” where admission is a safer bet.
+That is, students negotiate a tradeoff between highly attractive, selective “reach schools” and less preferable “safety schools” where admission is a safer bet (Kim 2015).
 Finally, the combinatorial nature of the college application problem makes it difficult to solve using the gradient-based techniques used in continuous portfolio optimization.
@@
 Our study pursues a more general solution.
+\else
+Chao (2014)가 주장한 바처럼 대학 지원 전략은 약간 미묘한 포트폴리오 최적화 문제이다.
+재정학 분야에서 고전적 포트폴리오 최적화 모형은 전체 자산의 총이익에서 위험회피 항을 뺌으로 선형식으로 제약된 오목 최대화 문제를 이룬다 (Meucci 2005).
+그러나 대학 지원자는 가치가 제일 높을 단 한 자산의 가치를 최대화하고자 한다.
+어떤 학생이 자신의 $j$위 학교에 합격하면, $(j+1)$위 학교에 합격하든 불합격하든 무관심하기 때문이다.
+이에 따라 학생이 최대화하려는 함수는 각 지원의 기대 효용에 대해 오목이 아닌 볼록 함수이다.
+또한 전형적인 입학 시장에서 대학의 효용과 합격 확률이 서로 반비례하므로 대학 지원 문제에 위험 관리가 내포 되어 있다.
+선호도가 높으며 붙기 어려운 “상향학교”(reach school)와 선호도가 낮으며 붙기 쉬운 “안정학교”(safety school) 사이의 균형을 고려해야 하기 때문이다 (김민희 2015).
+마지막, 대학 지원의 조합적인 본성 때문에 연속적인 포트폴리오 최적화 문제에서 흔히 사용하는 기울기 방법으로 풀기가 어렵다.
+Chao는 지원을 제약 조건 대신 비용으로 모형화했으며 모형을 추정하기 위해 $m=8$이 되도록 학교를 먼저 클러스터로 모았다.
+이는 열거법을 통해 문제를 쉽게 풀 수 있는 규모이었지만 본 연구는 더 일반적인 해법을 목표로서 추구한다.
+\fi
 
+\ifen
 We take special interest in the validity of greedy solution algorithms, such as that that iteratively adds the asset that yields the greatest increase in the objective function until the budget is exhausted.
 Greedy algorithms produce a \emph{nested} family of solutions, parameterized by the budget $H$: If $H \leq H'$, then the greedy solution for budget $H$ is a subset of the greedy solution for budget $H'$.
-As Rozanov and Tamir (2020) remark, this nestedness property is useful not only for obtaining an optimal solution, but in the implementation thereof under uncertain information.
+As Rozanov and Tamir (2020) remark, the knowledge that the optima are nested aids not only in computing the optimal solution, but in the implementation thereof under uncertain information.
 For example, in the United States, many college applications are due at the beginning of November, and it is typical for students to begin working on their applications during the prior summer because colleges expect students to tailor their essays to the target school.
@@
 The nestedness property---or equivalently, the validity of a greedy algorithm---implies that even in the absence of complete budget information, students can begin to carry out the optimal application strategy by writing essays for schools in the order that they enter the maximal portfolio.
+\else
+본 연구에서 특히 탐욕 해법의 타당성에 관심을 기울인다.
+예를 들어 예산이 다 소비될 때까지 목적 함수를 가장 빨리 증가시키는 학교를 차례대로 선택하는 알고리즘은 일종의 탐욕 해법이다.
+탐욕 해법 예산 $H$으로 모수화된 해의 순서를 유도하며 `포함 사슬 관계` (nestedness) 성질로 연결된다.
+$H \leq H'$일 때, 예산 $H$에 해당하는 탐욕 해는 예산 $H'$에 해당하는 탐욕 해의 부분집합이다.
+Rozanov과 Tamir (2020)이 주장하듯, 최적해가 포함 사슬 관계를 가짐을 알면 최적해를 구하는 데일 뿐더러 정보가 불확실한 상황에서 최적해를 이행하는 데에도 유용하다.
+가령 많은 미국 대학의 지원 기한은 11월 초인데 지원 에세이를 학교의 취향을 맞춰서 작성해야 하므로 여름부터 원서를 작성하는 학생이 많다.
+그러나 시간이나 지원비에 대한 예산을 10월 말까지 모를 수도 있다.
+포함 사슬 관계 성질, 또는 탐욕 해법의 타당성은 완전한 예산 정보가 없음에도 학교가 최적 포트폴리오에 진입하는 순서대로 원서를 작성하는 식으로 최적 전략을 이행할 수 있다고 의미한다.
+\fi
 
@@
 \parindent -2em
+\ifen \else
+김민희. 2015. ``{[대입 수시 전략]} 총 6번의 기회 \textellipsis `상향·소신·안정' 분산 지원하라.'' 중앙일보, 8월 26일. \url{https://www.joongang.co.kr/article/18524069}.\fi
+
 Budish, Eric.
@@
 Dantzig, George B. 1957. ``Discrete-Variable Extremum Problems.'' \emph{Operations Research} 5 (2): 266--88.
+
+\ifen Kim, Minhee.
+2015.
+``[College application strategy] Six chances total\dots divide applications across reach, target, and safety schools'' (in Korean).
+Jungang Ilbo, Aug.
+26.
+\url{https://doi.org/10.1086/664613}\fi
 

### 2022-03-03 benchmark/benchmark.jl

@@ -2,9 +2,10 @@ using OptimalApplication
 using DataFrames
 using Random
 using StatsBase
+using Statistics
 using Base.Threads
 
-fullscale = false
+fullscale = true
 
 # A long benchmark; tweak parameters with caution.
 # Set fullscale = false to run a smaller benchmark to check formatting etc.
@@ -25,6 +26,7 @@ function make_correlated_market(m)
     H = sum(g) ÷ 2
     return f, t, g, H
 end
+
 
 function benchmark1()
     printheader("Benchmark 1: Homogeneous-cost algorithms")
@@ -56,7 +58,6 @@ function benchmark1()
     end
     return combine(groupby(df, :m), kv_pairs...), df
 end
-
 
 
 function benchmark2()
@@ -98,10 +99,10 @@ function benchmark2()
 end
 
 
-
 @time begin
     println()
     display(benchmark1()[1])
     println("\n")
     display(benchmark2()[1])
+    println("\n")
 end

### 2022-03-03 src/OptimalApplication.jl

@@ -414,9 +414,7 @@ function optimalportfolio_fptas(
     v = vType(0)
     v_UB = Ū
 
-    nit = 0
-    while nit < 500 && v + 1 < v_UB
-        nit += 1
+    while v + 1 < v_UB
         mid = (v + v_UB) ÷ 2
 
         if G(m, mid) > H

### 2022-03-03 src/VariedCostsBnB.jl

@@ -1,0 +1,224 @@
+using Random
+using OptimalApplication
+import Base.hash
+import Base.isless
+
+function make_correlated_market(m)
+    A = 10
+    t = ceil.(Int, 10 * randexp(m))
+    sort!(t)
+    f = inv.(t .+ 10 * rand(m))
+    g = rand(5:10, m)
+    H = sum(g) ÷ 2
+    return f, t, g, H
+end
+
+
+struct VariedCostsMarket
+    f::Vector{<:Real}
+    t::Vector{<:Real}
+    g::Vector{<:Real}
+    H::Real
+    ft::Vector{<:Real}      # = f .* t
+    omf::Vector{<:Real}     # = 1 .- f
+
+    function VariedCostsMarket(
+        f::Vector{<:Real},
+        t::Vector{<:Real},
+        g::Vector{<:Real},
+        H::Real)
+
+        return new(f, t, g, H,
+            f .* t,
+            1 .- f)
+    end
+end
+
+
+mutable struct Node{jType<:Integer}
+    I::Set{jType}
+    N::Set{jType}
+    t̄::Dict{jType,<:AbstractFloat}
+    H̄::Real
+    v_I::AbstractFloat
+    v_LP::AbstractFloat
+    children::Set{UInt64}
+    isleaf::Bool
+
+    function Node(I::Set, N::Set, t̄::Dict, H̄::Real, v_I::Real, mkt::VariedCostsMarket)
+        # For generating a new node with its LP relaxation value and empty child set
+
+        # Leaf node
+        if isempty(N)
+            return new{eltype(I)}(I, N, t̄, H̄, v_I, v_I, Set{UInt64}(), true)
+        end
+
+        j_order = collect(N)
+        j_order[:] = j_order[sortperm(collect(mkt.f[j] * t̄[j] / mkt.g[j] for j in j_order), rev = true)]
+
+        v_LP = v_I
+        H_left = H̄
+        for j in j_order
+            if mkt.g[j] ≤ H_left
+                H_left -= mkt.g[j]
+                v_LP += mkt.f[j] * t̄[j]
+            else
+                v_LP += mkt.f[j] * t̄[j] * H_left / mkt.g[j]
+                @goto done
+            end
+        end
+        # Need to use goto here for the case in which all of N fits within H̄
+        @label done
+
+        return new{eltype(I)}(I, N, t̄, H̄, v_I, v_LP, Set{UInt64}(), false)
+    end
+end
+
[…]

### 2022-03-04 paper/paper.tex
write up branch and bound; improvements to translation and intro; small optimizations in BnB code

@@
 \newtheorem{definition}{Definition}
+\newtheorem{problem}{Problem}
 \newtheorem{assumption}{Assumption}
@@
 \newtheorem{definition}{정의}
+\newtheorem{problem}{문제}
 \newtheorem{assumption}{가정}
@@
 When colleges differ in their application fees, we show that the problem is NP-complete.
-We provide two dynamic programs for this more general setup.
-The first produces an exact solution in pseudopolynomial time, and the second yields a fully polynomial-time approximation scheme.
+We provide three algorithms for this more general setup.
+The first is a branch-and-bound routine.
+The second is an dynamic program that produces an exact solution in pseudopolynomial time.
+The third is a fully polynomial-time approximation scheme.
 
@@
 대학의 지원비가 서로 다른 경우, 문제가 NP-complete함을 증명한다.
-일반적인 상황을 위해 두 가지 동적 계획 해법을 도출한다.
-첫째는 의사 다항 시간 안에 정확한 해를 출력하며 둘째는 전체 다항 시간 근사 해법이다.
+일반적인 상황을 위해 세 가지 해법을 도출한다.
+첫째는 분지한계 기반 해법이다.
+둘째는 의사 다항 시간 안에 정확한 해를 출력하는 동적 계획이다.
+셋째는 전체 다항 시간 근사 해법이다.
 
@@
 Let $g_j$ denote the application fee for $c_j$ and $H$ the student's total budget to spend on application fees.
-Lastly, let $f_j$ denote her probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not.
-It is appropriate to assume that the $Z_j$ are probabilistically independent as long as $f_j$ are probabilities estimated specifically for this student (as opposed to generic acceptance rates).
+Lastly, let $f_j$ denote the student's probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not.
+It is appropriate to assume that the $Z_j$ are statistically independent as long as $f_j$ are probabilities estimated specifically for this student (as opposed to generic acceptance rates).
 Then the student's objective is to maximize the expected utility associated with the best school she gets into within this budget.
-Her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which she applies.
+Therefore, her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which she applies.
 \else
@@
 As Chao (2014) remarked, college application represents a somewhat subtle portfolio optimization problem.
-In computational finance, traditional portfolio optimization models weigh the expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Meucci 2005).
+In computational finance, traditional portfolio optimization models weigh the sum of expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Meucci 2005).
 But college applicants maximize the value of their \emph{best} asset: If a student is admitted to her $j$th choice, then she is indifferent as to whether she gets into her $(j+1)$th choice.
@@
 Chao (2014)가 주장한 바처럼 대학 지원 전략은 약간 미묘한 포트폴리오 최적화 문제이다.
-재정학 분야에서 고전적 포트폴리오 최적화 모형은 전체 자산의 총이익에서 위험회피 항을 뺌으로 선형식으로 제약된 오목 최대화 문제를 이룬다 (Meucci 2005).
+재정학 분야에서 고전적 포트폴리오 최적화 모형은 전체 자산에 대한 기대 총이익에서 위험회피 항을 뺌으로 선형식으로 제약된 오목 최대화 문제를 이룬다 (Meucci 2005).
 그러나 대학 지원자는 가치가 제일 높을 단 한 자산의 가치를 최대화하고자 한다.
-어떤 학생이 자신의 $j$위 학교에 합격하면, $(j+1)$위 학교에 합격하든 불합격하든 무관심하기 때문이다.
-이에 따라 학생이 최대화하려는 함수는 각 지원의 기대 효용에 대해 오목이 아닌 볼록 함수이다.
-또한 전형적인 입학 시장에서 대학의 효용과 합격 확률이 서로 반비례하므로 대학 지원 문제에 위험 관리가 내포 되어 있다.
-선호도가 높으며 붙기 어려운 “상향학교”(reach school)와 선호도가 낮으며 붙기 쉬운 “안정학교”(safety school) 사이의 균형을 고려해야 하기 때문이다 (김민희 2015).
-마지막, 대학 지원의 조합적인 본성 때문에 연속적인 포트폴리오 최적화 문제에서 흔히 사용하는 기울기 방법으로 풀기가 어렵다.
+어떤 학생이 자신의 $j$위 학교에 합격하면, $(j+1)$위 학교에 합격하든 불합격하든 무관심하기 때문에 다른 상황이다.
+이에 따라 학생이 최대화하려는 함수는 각 지원의 기대 효용에 대해 오목이 아닌 볼록 함수다.
+또한 전형적인 입학 시장에서 대학의 효용과 합격 확률이 서로 반비례하므로 대학 지원 문제에 위험 관리가 내포되어 있다.
+특히 선호도가 높으며 붙기 어려운 “상향학교”(reach school)와 선호도가 낮으며 붙기 쉬운 “안정학교”(safety school) 사이의 균형을 고려해야 한다 (김민희 2015).
+마지막, 대학 지원의 조합적인 본성으로 인해 연속적인 포트폴리오 최적화 문제에서 흔히 사용하는 기울기 방법으로는 풀기가 어렵다.
 Chao는 지원을 제약 조건 대신 비용으로 모형화했으며 모형을 추정하기 위해 $m=8$이 되도록 학교를 먼저 클러스터로 모았다.
-이는 열거법을 통해 문제를 쉽게 풀 수 있는 규모이었지만 본 연구는 더 일반적인 해법을 목표로서 추구한다.
+이는 열거법을 통해 문제를 쉽게 풀 수 있는 규모이지만 본 연구는 더 일반적인 해법을 목표로서 추구한다.
 \fi
@@
 본 연구에서 특히 탐욕 해법의 타당성에 관심을 기울인다.
-예를 들어 예산이 다 소비될 때까지 목적 함수를 가장 빨리 증가시키는 학교를 차례대로 선택하는 알고리즘은 일종의 탐욕 해법이다.
-탐욕 해법 예산 $H$으로 모수화된 해의 순서를 유도하며 `포함 사슬 관계` (nestedness) 성질로 연결된다.
-$H \leq H'$일 때, 예산 $H$에 해당하는 탐욕 해는 예산 $H'$에 해당하는 탐욕 해의 부분집합이다.
-Rozanov과 Tamir (2020)이 주장하듯, 최적해가 포함 사슬 관계를 가짐을 알면 최적해를 구하는 데일 뿐더러 정보가 불확실한 상황에서 최적해를 이행하는 데에도 유용하다.
-가령 많은 미국 대학의 지원 기한은 11월 초인데 지원 에세이를 학교의 취향을 맞춰서 작성해야 하므로 여름부터 원서를 작성하는 학생이 많다.
+예를 들어 예산이 다 소비될 때까지 목적 함수를 가장 크게 증가시키는 학교를 차례대로 선택하는 알고리즘은 일종의 탐욕 해법이다.
+탐욕 해법은 예산 $H$으로 모수화된 해의 순서를 유도하며 해는 `포함 사슬 관계'(nestedness) 성질에 따라 서로에 연결된다.
+즉 $H \leq H'$일 때, 예산 $H$에 해당하는 탐욕 해는 예산 예산 $H'$에 해당하는 탐욕 해의 부분집합이다.
+Rozanov과 Tamir (2020)가 주장하듯, 최적해가 포함 사슬 관계를 가짐을 알면 최적해를 구하는 데일 뿐더러 정보가 불확실한 상황에서 최적해를 이행하는 데에도 유용하다.
+가령, 많은 미국 대학의 지원 기한은 11월 초인데 지원 에세이를 학교의 취향을 맞춰서 작성해야 하므로 여름부터 원서를 작성하는 학생이 많다.
 그러나 시간이나 지원비에 대한 예산을 10월 말까지 모를 수도 있다.
-포함 사슬 관계 성질, 또는 탐욕 해법의 타당성은 완전한 예산 정보가 없음에도 학교가 최적 포트폴리오에 진입하는 순서대로 원서를 작성하는 식으로 최적 전략을 이행할 수 있다고 의미한다.
+포함 사슬 관계 성질, 또는 탐욕 해법의 타당성은 완전한 예산 정보가 없음에도 학교가 최적 포트폴리오에 진입하는 순서대로 원서를 작성하는 식으로 최적 전략을 이행해 낼 수 있다고 의미한다.
 \fi
@@
 For other problems, notably the binary knapsack problem, the most intuitive greedy algorithm can be made to perform arbitrarily poorly (Vazirani 2001).
-We show analogous results for the college application problem: In the special case where each $g_j = 1$, the optimal portfolio satisfy a nestedness property that is equivalent to the validity of the greedy algorithm.
+We show that the college application problem that the college application problem is analogous to the knapsack problem in many respects: In the special case where each $g_j = 1$, the optimal portfolio satisfy a nestedness property that is equivalent to the validity of the greedy algorithm.
 This case mirrors the centralized college application process in Korea, where there is no application fee, but students are allowed to apply to only three schools during the main admissions cycle.
@@
 We show that the decision form of the portfolio optimization problem is NP-complete through a reduction from the binary knapsack problem.
-We provide two dynamic programs for this more general setup.
-The first iterates on total expenditures and produces an exact solution in pseudopolynomial time, namely $O(Hm + m \log m)$.
-The second iterates on truncated portfolio valuations and yields a fully polynomial-time approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in $O(m^3 / \varepsilon)$ time.
+We provide three algorithms for this more general setup.
+The first is a branch-and-bound routine.
+The second is a dynamic program that iterates on total expenditures and produces an exact solution in pseudopolynomial time, namely $O(Hm + m \log m)$.
+The third is a dynamic program that iterates on truncated portfolio valuations.
+It yields a fully polynomial-time approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in $O(m^3 / \varepsilon)$ time.
 % subject to certain assumptions on the quality of the data available to students in their decisionmaking process, an optimal application portfolio for a single student can be computed in time polynomial in $h$ and $m$, as we show presently.
@@
 
-\begin{definition}[Alma’s problem]
+\begin{problem}[Alma’s problem]
 Alma's optimal college application portfolio is given by the solution to the following combinatorial optimization problem:
@@
 \end{align}
-\end{definition}
+\end{problem}
 %This problem can be formulated as an integer nonlinear program in $m$ binary variables ..
@@
 \ifen \section{Heterogeneous application costs} \else\section{다양한 지원 비용} \fi\label{hetappcosts}
-In this section, we turn to the more general problem in which the constant $g_j$ represents the \emph{cost} of applying to $c_j$ and the student, whom we now call Ellis, has a \emph{budget} of $H$ to spend on college applications.
-Applying Theorem \ref{assumetzerozero}, we assume $t_0 = 0$ and disregard $c_0$ throughout.
-
-\begin{definition}[Ellis's problem]
+Now we turn to the more general problem in which the constant $g_j$ represents the \emph{cost} of applying to $c_j$ and the student, whom we now call Ellis, has a \emph{budget} of $H$ to spend on college applications.
+Applying Theorem \ref{assumetzerozero}, we assume $t_0 = 0$ and disregard the outside option throughout.
+
+\begin{problem}[Ellis's problem]
 Ellis's optimal college application portfolio is given by the solution to the following combinatorial optimization problem.
@@
 \end{align}
-\end{definition}
-
+\end{problem}
+
+In this section, we show that this problem is NP-complete, then provide three algorithmic solutions: an exact branch-and-bound routine, an exact dynamic program, and a fully polynomial-time approximation scheme.
+
+\ifen \subsection{NP-completeness} \else \subsection{NP-completeness}\fi
 The optima for Ellis's problem are not necessarily nested, nor is the number of schools in the optimal portfolio necessarily increasing in $H$. For example, if
@@
 then it is evident that the optimal portfolio for $H = 2$ is $\{1, 2\}$ while that for $H = 3$ is $\{3\}$.
-
-\ifen \subsection{NP-completeness} \else \subsection{NP-completeness}\fi
 In fact, Ellis’s problem is NP-complete, as we will show by a transformation from the binary knapsack problem, which is known to be NP-complete (Garey and Johnson 1979).
 
-\begin{definition}[Decision form of knapsack problem]
-An \emph{instance} consists of a set $\mathcal{B}$ of $m$ objects; utility values $u_j \in \mathbb{N}$ and weight $w_j \in \mathbb{N}$ for each $j \in \mathcal{B}$; and target utility $U\in \mathbb{N}$ and knapsack capacity $W\in \mathbb{N}$.
+\begin{problem}[Decision form of knapsack problem]
+An \emph{instance} consists of a set $\mathcal{B}$ of $m$ objects, utility values $u_j \in \mathbb{N}$ and weight $w_j \in \mathbb{N}$ for each $j \in \mathcal{B}$, and target utility $U\in \mathbb{N}$ and knapsack capacity $W\in \mathbb{N}$.
 The instance is called a \emph{yes-instance} if and only if there exists a set $\mathcal{B’} \subseteq \mathcal{B}$ having $\sum_{j \in \mathcal{B’}} u_j \geq U$ and  $\sum_{j \in \mathcal{B’}} w_j \leq W$.
-\end{definition}
-
-\begin{definition}[Decision form of Ellis’s problem]
+\end{problem}
+
+\begin{problem}[Decision form of Ellis’s problem]
 An \emph{instance} consists of an instance of Ellis’s problem and a target valuation $V$.
 The instance is called a \emph{yes-instance} if and only if there exists a portfolio $\mathcal{X} \subseteq \mathcal{C}$ having $v(\mathcal{X}) \geq V$ and  $\sum_{j \in \mathcal{X}} g_j \leq H$.
-\end{definition}
+\end{problem}
 
@@
 \end{align}
-This means that the value of an application portfolio $\mathcal{X}$ for the corresponding knapsack instance is the smallest integer greater than $v( \mathcal{X})$.
+This means that the utility of an application portfolio $\mathcal{X}$ in the corresponding knapsack instance is the smallest integer greater than $v( \mathcal{X})$.
 That is, $\sum_{j \in \mathcal{X}} u_j \geq U$ if and only if $v(\mathcal{X}) \geq U -1$.
@@
 
+\ifen \subsection{Branch-and-bound algorithm} \else \subsection{분지한계법} \fi
+A traditional approach to knapsack problems is the branch-and-bound framework, which generates subproblems in which the values of one or more decision variables are fixed and uses an upper bound on the objective function to exclude, or \emph{fathom,} branches of the tree of subproblems that cannot yield a solution better than the best solution on hand (Martello and Toth 1990; Kellerer et al.
+2004).
+In this subsection, we present an integer formulation of Ellis's problem and a linear program (LP) that bounds the objective value from above.
+We tighten the LP bound for specific subproblems by reusing the conditional transformation of the $t_j$-values from Algorithm \ref{algorithmforlargeh}.
+A branch-and-bound routine emerges naturally from these ingredients.
+
+To begin, let us characterize the portfolio $\mathcal{X}$ as the binary vector $x \in \{0, 1\}^m$, where $x_j = 1 $ if and only if $j \in \mathcal\{X\}$.
+Then it is not difficult to see that Ellis's problem is equivalent to the following integer nonlinear program.
+\begin{problem}[Integer NLP for Ellis's problem] \label{integernlp}
+\begin{align}
+\begin{split}
+\text{maximize}\quad &  v(x) = \sum_{j=1}^m \Bigl( t_j f_j  x_j \prod_{i > j} (1 - f_{i} x_i) \Bigr)\\
+\text{subject to}\quad & \sum_{j=1}^m g_j x_j \leq H, ~~ x_i \in \{0, 1\}^m
+\end{split}
+\end{align}
+\end{problem}
+Since the product in $v(x)$ does not exceed one, the following LP relaxation is an upper bound on the valuation of the optimal portfolio.
+\begin{problem}[LP relaxation for Ellis's problem] \label{LPrelaxation}
+\begin{align}
+\begin{split}
+\text{maximize}\quad &  v_{\mathrm{LP}}(x) = \sum_{j=1}^m  t_j f_j  x_j \\
+\text{subject to}\quad & \sum_{j=1}^m g_j x_j \leq H, ~~ x \in [0, 1]^m
+\end{split}
+\end{align}
+\end{problem}
+Problem \ref{LPrelaxation} is a continuous knapsack problem, which is easily solved in $O(m \log m)$ time by the following greedy algorithm: Start with $x = \mathbf{0}$.
+While $H > 0$, select the school $k$ for which $t_j f_j / g_j$ is highest, set $x_k \gets \min\{1, H / g_k\}$, and set $H \gets H - g_k$ (Dantzig 1957).
+
+In our branch-and-bound architecture, a \emph{node} is characterized by a three-way partition of schools $\mathcal{C}= \mathcal{I} \cup \mathcal{O} \cup \mathcal{N}$ satisfying $\sum_{j \in \mathcal{I}} g_j \leq H$.
+$\mathcal{I}$ consists of schools that are ``in'' the application portfolio, $\mathcal{O}$ consists of those that are ``out,'' and $\mathcal{N}$ consists of those that are ``negotiable.''
+%This partition induces a constant $\gamma$, transformed $t_j$-values $\bar t$, the residual budget $\mathcal{H}$, and a set $\mathcal{L}$ of pointers to the node's children, if any.
+The choice of partition induces a pair of subproblems. The first subproblem is an instance of Problem \ref{integernlp}, namely
+\begin{align} \label{nodenlpsubproblem}
+\begin{split}
+\text{maximize}\quad &  w(y) = \gamma + \sum_{j \in \mathcal{N}} \Bigl( \bar t_j f_j  y_j \prod_{\substack{i \in \mathcal{N}:\\i > j}} (1 - f_{i} y_i) \Bigr)\\
+\text{subject to}\quad & \sum_{j \in \mathcal{N}} g_j y_j \leq \bar H; ~~ y_j \in \{0, 1\},~~j \in \mathcal{N}.
+\end{split}
+\end{align}
+The second is the corresponding instance of Problem \ref{LPrelaxation}:
+\begin{align} \label{nodelprelaxation}
+\begin{split}
+\text{maximize}\quad &  w_{\mathrm{LP}}(y)= \gamma + \sum_{j \in \mathcal{N}}  \bar t_j f_j  y_j \\
+\text{subject to}\quad & \sum_{j \in \mathcal{N}} g_j y_j \leq \bar H; ~~ y_j \in [0, 1],~~j \in \mathcal{N}
+\end{split}
+\end{align}
+In both subproblems, $\bar H = H - \sum_{j\in \mathcal{I}} g_j$ denotes the residual budget.
+The parameters $\gamma$ and $\bar t$ are obtained by iteratively applying the transformation \eqref{howtotransformfj} to the schools in $\mathcal{I}$.
+For each $j\in \mathcal{I}$,  we increment $\gamma$ by the current value of $f_j \bar t_j$, eliminate $j$ from the market, and update the remaining $\bar t_j$-values using \eqref{howtotransformfj}.
+
+Given a node $n  = (\mathcal{I}, \mathcal{O}, \mathcal{N})$, its children are generated as follows.
+Every node has zero, one, or two children.
+In the typical case, we select a school $j \in \mathcal{N}$ for which $g_j \leq \bar H$ and generate one child by moving $j$ to $\mathcal{I}$, and another child by moving $j$ it to $\mathcal{O}$.
+Equivalently, we set $x_j = 1$ in one child and $x_j = 0$ in the other.
+In principle, any school in $\mathcal{N}$ can be chosen for $j$, but as a greedy heuristic, we choose the school for which the ratio $t_j f_j / g_j$ is highest.
+Notice that this method of generating children ensures that each node's $\mathcal{I}$-set differs from its parent's by at most a single school, so the constant $\gamma$ and transformed $\bar t_j$-values for the new node can be obtained by a single application of \eqref{howtotransformfj}.
+
+There are two atypical cases.
+First, if every school in $\mathcal{N}$ has $g_j > \bar H$, then there is no school that can be added to $\mathcal{I}$ in a feasible portfolio, and the optimal portfolio on this branch is $\mathcal{I}$ itself.
+In this case, we generate only one child by moving all the schools from $\mathcal{N}$ to $\mathcal{O}$.
+Second, if $\mathcal{N} = \O$, then the node has zero children, and as no further branching is possible, the node is called a \emph{leaf.}
+
+In the branch-and-bound algorithm, we store nodes in a tree $\mathfrak{T}$, which contains nodes alongside pointers to their children.
+Each time a node $n= (\mathcal{I}, \mathcal{O}, \mathcal{N})$ is generated, we record the values $v_{\mathcal{I}}[n] = v(\mathcal{I})$ and $v^*_{\mathrm{LP}}[n]$, the optimal objective value of the LP relaxation \eqref{nodelprelaxation}.
+Because $\mathcal{I}$ is a feasible portfolio, $v_{\mathcal{I}}[n]$ is a lower bound on the optimal objective value.
+Moreover, by the argument given in the proof of Theorem \ref{nestedapplication}, the objective function of  \eqref{nodenlpsubproblem} is identical to the function $v(\mathcal{I} \cup \mathcal{Y})$, where $\mathcal{Y} = \{ j: y_j = 1\}$.
+This means that the optimal objective value $v_{\mathrm{LP}}^*$  is an upper bound on the valuation of any portfolio that contains $\mathcal{I}$ as a subset and does not include any school in $\mathcal{O}$, and hence on the valuation of any portfolio on this branch.
+Therefore, if upon generating a new node $n_2$, we discover that its objective value $v_{\mathcal{I}}[n_2]$ is greater than $v_{\mathrm{LP}}^*[n_1]$ for some other node $n_1$, then $n_1$ and all its descendants can be eliminated from consideration.
+
+The algorithm is initialized by populating $\mathfrak{T}$ with the root node $n_0 = (\O, \O, \mathcal{C})$.
+The algorithm terminates when all the nodes that remain on the tree are either leaf nodes, or have had their children explored.
+
+\begin{algorithm}[H] 
+%\DontPrintSemicolon
+\caption{Branch and bound for Ellis's problem.} \label{ellisbnb}
+\KwData{Utility values $t \in(0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)$.}
+Root node $n_0 \gets (\O, \O, \mathcal{C})$\;
+Current lower bound $L \gets 0$\;
+Initialize tree $\mathfrak{T} \gets \{ n_0\}$\;
+\While{not finished}{
+	$\mathfrak{S} \gets$ the set of non-leaf nodes whose children have not yet been generated\;
+	\uIf{$\mathfrak{S} = \O$}{
+		$\mathcal{X} \gets \mathcal{I}$ from node associated with current value of $L$\;
+		\Return{$\mathcal{X}, L$}
+	}
+	\Else{
+		$n \gets \argmax\{ v_{\mathrm{LP}}^*[n] : n \in \mathfrak{S}\}$\;
+		Generate children of $n$\;
+		\For{each child $n'$ of $n$}{
+			$L \gets \max\{L, v_{\mathcal{I}}[n']\}$\;
+		}
+	}
+	\For{$n''$ in $\mathfrak{T}$} {
+		\lIf{$L >  v_{\mathrm{LP}}^*[n'']$}{fathom $n''$}
+	}
+}
+\end{algorithm}
+
+\begin{theorem}[Validity of Algorithm \ref{ellisbnb}]
+Algorithm \ref{ellisbnb} produces an optimal application portfolio for Ellis's problem.
+\end{theorem}
+
+\begin{proof}
+The discussion above implies that the algorithm only fathoms branches when justified; therefore, if the algorithm terminates, then an optimal solution exists among the leaves of the tree, and it is therefore returned.
+
+To show that the algorithm does not cycle, it suffices to show that no node is generated twice.
+Suppose not: that two distinct nodes $n_1$ and $n_2$ share the same partition $(\mathcal{I}_{12}, \mathcal{O}_{12}, \mathcal{N}_{12})$.
+Trace each node's lineage up the tree and let $n$ denote the \emph{first} node at which the lineages meet.
+$n$ must have two children, or else its sole child is a common ancestor of $n_1$ and $n_2$, and one of these children, say $n_3$, must be an ancestor of $n_1$ while the other, say $n_4$, is an ancestor of $n_2$.
+Write $n_3 = (\mathcal{I}_{3}, \mathcal{O}_{3}, \mathcal{N}_{3})$ and $n_4 = (\mathcal{I}_{4}, \mathcal{O}_{4}, \mathcal{N}_{4})$.
+By the node-generation rule, there is a school $j$ in $\mathcal{I}_3 \cap \mathcal{O}_4$, and the $\mathcal{I}$-set (respectively, $\mathcal{O}$-set) for any descendant of $\mathcal{I}_3$ (respectively, $\mathcal{O}_4$) is a superset of  $\mathcal{I}_3$ (respectively, $\mathcal{O}_4$).
+Therefore, $j \in \mathcal{I}_{12} \cap \mathcal{O}_{12}$, meaning that $(\mathcal{I}_{12}, \mathcal{O}_{12}, \mathcal{N}_{12})$ is not a partition of $\mathcal{C}$, a contradiction.
+\end{proof}
+
+The branch-and-bound algorithm is an interesting benchmark, but as a kind of enumeration algorithm, its computation time grows rapidly in the problem size, and unlike the approximation scheme we propose later on, there is no guaranteed bound on the approximation error after a fixed number of iterations.
+In our numerical experiments, Algorithm \ref{ellisbnb} was impractical for instances containing more than about $m=40$ schools.
+
 \ifen \subsection{Pseudopolynomial-time dynamic program} \else\subsection{의사 다항 시간 동적 계획} \fi
@@
 If it does not contain $j$, then $\mathcal{X}[j, h] = \mathcal{X}[j-1, h]$.
-On the other hand, if  $\mathcal{X}[j, h]$ contains $j$, then its value is $(1 - f_j) v(\mathcal{X}[j, h]\setminus \{j\}) + f_j t_j$.
+On the other hand, if  $\mathcal{X}[j, h]$ contains $j$, then its valuation is $(1 - f_j) v(\mathcal{X}[j, h]\setminus \{j\}) + f_j t_j$.
 This requires that $\mathcal{X}[j, h]\setminus \{j\}$ make optimal use of the remaining budget over the remaining schools; that is, $\mathcal{X}[j, h] = \mathcal{X}[j-1, h - g_j] \cup\{j\}$.
@@
 \caption{Dynamic program for Ellis's problem with integral application costs.} \label{ellisDP1}
-\KwData{Utility values $t \in[0, \infty)^m$, admissions probabilities $f \in [0, 1]^m$, application costs $g \in \mathbb{N}^m$, budget $H \in\mathbb{N}$.}
+\KwData{Utility values $t \in[0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in \mathbb{N}^m$, budget $H \in\mathbb{N}$.}
 Index schools in ascending order by $t$\;
@@
 
-We will represent the approximate value of portfolios using a fixed-point decimal with a precision of $P$, where $P$ is the number of digits to retain after the decimal point.
+We will represent approximate portfolio valuations using a fixed-point decimal with a precision of $P$, where $P$ is the number of digits to retain after the decimal point.
 Let $r[x] = \lfloor 10^P x \rfloor 10^{-P}$ denote the value of $x$ rounded down to its nearest fixed-point representation.
-Since $\bar U = \sum_{j\in \mathcal{C}} f_j t_j$ is an upper bound on the value of any portfolio, and since we will ensure that each fixed-point approximation is an underestimate of the portfolio's true value, the set $\mathcal{V}$ of possible valuations possible in the fixed-point framework is finite:
+Since $\bar U = \sum_{j\in \mathcal{C}} f_j t_j$ is an upper bound on the valuation of any portfolio, and since we will ensure that each fixed-point approximation is an underestimate of the portfolio's true valuation, the set $\mathcal{V}$ of possible valuations possible in the fixed-point framework is finite:
 \begin{equation}
@@
 
-For the remainder of this subsection, unless otherwise specified, the word \emph{value} refers to a portfolio’s value within the fixed-point framework, with the understanding that this is an approximation.
+For the remainder of this subsection, unless otherwise specified, the word \emph{valuation} refers to a portfolio’s valuation within the fixed-point framework, with the understanding that this is an approximation.
 We will account for the approximation error below when we prove the dynamic program’s validity.
 
-For integers $0 \leq j \leq m$ and $v \in [-\infty, 0) \cup \mathcal{V}$, let $\mathcal{W}[j, v]$ denote the least expensive portfolio that uses only schools $\{ 1, \dots, j\}$ and has value at least $v$, if such a portfolio exists.
-Denote its value by $G[j, v]$, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist.
+For integers $0 \leq j \leq m$ and $v \in [-\infty, 0) \cup \mathcal{V}$, let $\mathcal{W}[j, v]$ denote the least expensive portfolio that uses only schools $\{ 1, \dots, j\}$ and has valuation at least $v$, if such a portfolio exists.
+Denote its cost by $G[j, v]$, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist.
 It is clear that if $v \leq 0$, then $\mathcal{W}[j, v] = \O$ and $G[j, h] = 0$, and that if $j = 0$ and $v > 0$, then $G[j, h] = \infty$.
@@
 v(\mathcal{W}[j, v]) = (1 - f_j )v(\mathcal{W}[j, v]\setminus \{j\}) + f_j t_j.\end{equation}
-Therefore, the subportfolio $\mathcal{W}[j, v]\setminus \{j\}$ must have a value of at least $v - \Delta$, where $\Delta$ satisfies $v = (1 - f_j )(v - \Delta) + f_j t_j $.
+Therefore, the subportfolio $\mathcal{W}[j, v]\setminus \{j\}$ must have a valuation of at least $v - \Delta$, where $\Delta$ satisfies $v = (1 - f_j )(v - \Delta) + f_j t_j $.
 When $f_j < 1$, the solution to this equation is $ \Delta = \frac{f_j}{1 - f_j} (t_j - v)$.
@@
 % In the main loop, each time the target valuation $v$ is updated using
-Because $\Delta_j(v)$ is rounded down in the recursion relation defined by \eqref{recursionrelationforcostmindp} and \eqref{deltajvdef}, if $j \in \mathcal{W}[j, v]$, then the true value of $(1 - f_j) v\bigl(\mathcal{W}[j-1, v- \Delta_j(v)]\bigr) + f_j t_j$ may exceed the fixed-point value $v$ of $\mathcal{W}[j, v]$, but not by more than $10^{-P}$.
+Because $\Delta_j(v)$ is rounded down in the recursion relation defined by \eqref{recursionrelationforcostmindp} and \eqref{deltajvdef}, if $j \in \mathcal{W}[j, v]$, then the true value of $(1 - f_j) v\bigl(\mathcal{W}[j-1, v- \Delta_j(v)]\bigr) + f_j t_j$ may exceed the fixed-point valuation $v$ of $\mathcal{W}[j, v]$, but not by more than $10^{-P}$.
 This error accumulates with each school added to $\mathcal{W}$, but the number of additions is at most $m$.
@@
 
-We can define $v'(\mathcal{X})$ analogously as the fixed-point valuation of $\mathcal{X}$ when its elements are added in index order and its value is updated and rounded down to the nearest multiple of $10^{-P}$ at each addition in accordance with \eqref{solvemeforvwjvminusj}.
+We can define $v'(\mathcal{X})$ analogously as the fixed-point valuation of $\mathcal{X}$ when its elements are added in index order and its valuation is updated and rounded down to the nearest multiple of $10^{-P}$ at each addition in accordance with \eqref{solvemeforvwjvminusj}.
 By the same logic,
@@
 
+Kellerer, Hans, Ulrich Pferschy, and David Pisinger. 2004. \emph{Knapsack Problems.} Berlin: Springer.
+
 \ifen Kim, Minhee.
@@
 
+Martello, Silvano and Paolo Toth. 1990. \emph{Knapsack Problems: Algorithms and Computer Implementations.} New York: John Wiley \& Sons.
+
 Meucci, Attilio. 2005. \emph{Risk and Asset Allocation.} Berlin: Springer-Verlag, 2005.

### 2022-03-04 benchmark/benchmark.jl

@@ -39,7 +39,7 @@ end
 
 function benchmark1()
     printheader("Benchmark 1: Homogeneous-cost algorithms")
-    M = fullscale ? [5, 50, 500, 5000] : [5, 10]
+    M = fullscale ? 4 .^ (2:6) : [5, 10]
 
     sizes = zeros(Int64, n_markets, length(M))
     times_dict = zeros(Float64, n_markets, length(M))
@@ -70,8 +70,8 @@ end
 
 function benchmark2()
     printheader("Benchmark 2: Heterogeneous-cost algorithms")
-    M = fullscale ? [5, 50, 500] : [5, 10]
-    bnbcutoff = fullscale ? 32 : 6
+    M = fullscale ? 2 .^ (4:9) : [5, 10]
+    bnbcutoff = fullscale ? 33 : 6
     epsilons = [0.5, 0.05]
 
     dtype = Union{Float64,Missing}

### 2022-03-04 src/VariedCostsBnB.jl

@@ -55,31 +55,26 @@ With respect to the data in `mkt`, generates the child(ren) of node `nd` and wri
 their hashes to `nd.children`.
 """
 function generatechildren!(nd::Node, mkt::VariedCostsMarket)
-    newI = copy(nd.I)
-    newN = copy(nd.N)
-
     fltr = filter(j -> mkt.g[j] ≤ nd.H̄, nd.N)
 
     # No way to add any school to this: just skip to the leaf node
     if isempty(fltr)
-        child = Node(newI, Set{eltype(nd.I)}(), nd.t̄, nd.H̄, nd.v_I, mkt)
+        child = Node(nd.I, Set{eltype(nd.I)}(), nd.t̄, nd.H̄, nd.v_I, mkt)
         push!(nd.children, hash(child))
         return (child,)
     end
 
-    # School we will branch on.
-    # You may be tempted to divide by g[j] here, but this is unwise, 
-    # because the optimal addition on this branch is always the affordable
-    # school with highest ft[j]. Different from knapsack.
-    i = argmax(j -> mkt.f[j] * nd.t̄[j], fltr)
+    # School we will branch on. In principle it can by any school in fltr.
+    i = argmax(j -> mkt.f[j] * nd.t̄[j] / mkt.g[j], fltr)
 
-    delete!(newN, i)
+
+    newN = setdiff(nd.N, i)
 
     # Node without i
     t̄2 = copy(nd.t̄)
     delete!(t̄2, i)
 
-    child2 = Node(newI, newN, t̄2, nd.H̄, nd.v_I, mkt)
+    child2 = Node(nd.I, newN, t̄2, nd.H̄, nd.v_I, mkt)
 
     if mkt.g[i] < nd.H̄
         # Node with i
@@ -93,7 +88,7 @@ function generatechildren!(nd::Node, mkt::VariedCostsMarket)
         end
         delete!(t̄1, i)
 
-        child1 = Node(union(newI, i), newN, t̄1, nd.H̄ - mkt.g[i],
+        child1 = Node(union(nd.I, i), newN, t̄1, nd.H̄ - mkt.g[i],
             nd.v_I + mkt.f[i] * nd.t̄[i],
             mkt)
 
@@ -112,7 +107,7 @@ function generatechildren!(nd::Node, mkt::VariedCostsMarket)
         end
         delete!(t̄1, i)
 
[…]

### 2022-03-07 paper/paper.tex
rm notebooks, translation in intro, bnb

@@
 \newif\ifen
-\entrue % Change to false to export Korean
+%\entrue % Change to false to export Korean
 
@@
 
-\ifen {} \else \renewcommand*{\proofname}{증명\@}  \fi
+\ifen {} \else
+\renewcommand*{\proofname}{증명\@} 
+\renewcommand{\figurename}{그림}
+\fi
+
+\numberwithin{equation}{section}
+%\numberwithin{theorem}{section}
+%\numberwithin{lemma}{section}
+%\numberwithin{corollary}{section}
+%\numberwithin{proposition}{section}
+%\numberwithin{example}{section}
+%\numberwithin{definition}{section}
+%\numberwithin{problem}{section}
+%\numberwithin{assumption}{section}
+%\numberwithin{conjecture}{section}
 
@@
 \ifen \section{Introduction}  \else \section{서론} \fi
-\ifen This paper considers the portfolio optimization problem
+\ifen This paper considers the following optimization problem:
 \else 본 논문은 다음과 같은 최적화 문제를 고려한다.\fi
 %\begin{definition}[Optimal college application portfolio] \label{generalproblemstatement}
-\begin{align}
+\begin{align} \label{headlineproblem}
 \begin{split}
-\text{maximize}\quad & \operatorname{E}\Bigl[\max\bigr\{t_0,
+\text{maximize}\quad & v(\mathcal{X}) =  \operatorname{E}\Bigl[\max\bigr\{t_0,
 \max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\Bigr] \\
@@
 %\end{definition}
-\ifen where $\mathcal{C} = \{ 1 \dots m\}$ is an index set, $H$ is a budget parameter, and for $j = 1 \dots m$, $g_j > 0$ is a cost parameter, $Z_j$ is a random, independent Bernoulli variable with probability $f_j$.
-\else 단,  $\mathcal{C} = \{ 1 \dots m\}$은 지표 집합이며 $H$는 예산 모수이다. 각 $j = 1 \dots m$에 대해 $g_j > 0$는 비용 모수이며 $Z_j$는 확률 $f_j$를 가지는 서로 독립적인 Bernoulli 변수이다.\fi
+\ifen Here $\mathcal{C} = \{ 1 \dots m\}$ is an index set; $H > 0$ is a budget parameter; for $j = 1 \dots m$, $g_j > 0$ is a cost parameter and $Z_j$ is a random, independent Bernoulli variable with probability $f_j$; and for $j = 0\dots m$, $t_j$ is a utility parameter.
+\else 단,  $\mathcal{C} = \{ 1 \dots m\}$은 지표 집합이며 $H$는 예산 모수이다. 각 $j = 1 \dots m$에 대해 $g_j > 0$는 비용 모수이며 $Z_j$는 확률 $f_j$를 가지는 서로 독립적인 Bernoulli 변수이다. 각 $j = 0 \dots m$에 대해 $t_j$는 효용 모수이다. \fi
 
@@
 $j$번째 학교의 이름은 $c_j$이다.
-이 시장에 참여하는 한 학생을 고려해서 그가 $c_j$에 진학하면 $t_j$ 단위의 효용이 발생한다고 하자.
+이 시장에 참가하는 한 학생을 고려해서 그가 $c_j$에 진학하면 $t_j$ 단위의 효용이 발생한다고 하자.
 단, 어떤 대학에도 진학하지 않는 경우 그의 효용은 $t_0$이다.
@@
 As Chao (2014) remarked, college application represents a somewhat subtle portfolio optimization problem.
-In computational finance, traditional portfolio optimization models weigh the sum of expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Meucci 2005).
+In computational finance, traditional portfolio optimization models weigh the sum of expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Markowitz 1952; Meucci 2005).
 But college applicants maximize the value of their \emph{best} asset: If a student is admitted to her $j$th choice, then she is indifferent as to whether she gets into her $(j+1)$th choice.
@@
 Chao (2014)가 주장한 바처럼 대학 지원 전략은 약간 미묘한 포트폴리오 최적화 문제이다.
-재정학 분야에서 고전적 포트폴리오 최적화 모형은 전체 자산에 대한 기대 총이익에서 위험회피 항을 뺌으로 선형식으로 제약된 오목 최대화 문제를 이룬다 (Meucci 2005).
+재정학 분야에서 고전적 포트폴리오 최적화 모형은 전체 자산에 대한 기대 총이익에서 위험회피 항을 뺌으로 선형식으로 제약된 오목 최대화 문제를 이룬다 (Markowitz 1952; Meucci 2005).
 그러나 대학 지원자는 가치가 제일 높을 단 한 자산의 가치를 최대화하고자 한다.
-어떤 학생이 자신의 $j$위 학교에 합격하면, $(j+1)$위 학교에 합격하든 불합격하든 무관심하기 때문에 다른 상황이다.
+어떤 학생이 자신의 $j$위 학교에 합격하면, $(j+1)$위 학교에 합격하든 불합격하든 무관심하므로 다른 상황이다.
 이에 따라 학생이 최대화하려는 함수는 각 지원의 기대 효용에 대해 오목이 아닌 볼록 함수다.
@@
 Chao는 지원을 제약 조건 대신 비용으로 모형화했으며 모형을 추정하기 위해 $m=8$이 되도록 학교를 먼저 클러스터로 모았다.
-이는 열거법을 통해 문제를 쉽게 풀 수 있는 규모이지만 본 연구는 더 일반적인 해법을 목표로서 추구한다.
+이는 열거법을 통해 문제를 쉽게 풀 수 있는 규모이지만 본 연구는 더 일반적인 해법을 목표로 추구한다.
 \fi
@@
 \ifen
-We take special interest in the validity of greedy solution algorithms, such as that that iteratively adds the asset that yields the greatest increase in the objective function until the budget is exhausted.
-Greedy algorithms produce a \emph{nested} family of solutions, parameterized by the budget $H$: If $H \leq H'$, then the greedy solution for budget $H$ is a subset of the greedy solution for budget $H'$.
+The integer formulation of the college application problem, given below, can also be viewed as a kind of binary knapsack problem with a polynomial objective function of degree $m$.
+Our branch-and-bound and dynamic programming algorithms closely resemble existing algorithms for knapsack problems (Martello and Toth 1990, \S\,2.5--6).
+In fact, by manipulating the admissions probabilities, the objective function can be made to approximate a linear function of the characteristic vector to an arbitrary degree of accuracy, a fact that we exploit in our NP-completeness proof.
+Previous research has introduced various forms of stochasticity to the knapsack problem, including variants in which each item's utility takes a known probability distribution (Steinberg and Parks 1979; Carraway et al.
+1993) and an online context in which the weight of each item is observed after insertion into the knapsack (Dean et al.
+2008).
+Our problem superficially resembles the preference-order knapsack problem considered by Steinberg and Parks and Carraway et al., but these models lack the college application problem's unique ``maximax'' form.
+Additionally, unlike those models, we do not attempt to replace the scalar objective function with a preference order over outcome distributions, which introduces methodological issues concerning competing notions of stochastic dominance (Sniedovich 1980).
+We take for granted the student's preferences over \emph{outcomes} (as encoded in the $t_j$-values), and are interested instead in an efficient computational approach to the well-defined problem above.
+\else
+대학 지원 문제의 정수 모형은 $m$차 다항식을 목적함수로 갖춘 일종의 이진 배낭 문제로 볼 수도 있다.
+본 논문에서 제시하는 분지한계법과 동적 계획 해법은 배낭 문제를 위한 기존 알고리즘하고 비슷하다 (Martello와 Toth 1990, \S\,2.5--6).
+입학 확률을 적당히 조정하면 원래 목적함수를 특성벡터의 선형 함수에 대한 임의로 정확한 근사로 변환할 수 있으며 NP-completeness 증명에서 이 사실을 활용한다.
+배낭 문제에 확률성을 도입한 선행 연구 중 각 상품의 효용이 정해진 확률 분포로 결정되는 모형(Steinberg와 Parks 1979; Carraway 외 1993), 그리고 배낭에 삽입한 다음에 상품의 무게가 관측되는 온라인 모형(Dean et al.
+2008)과 같은 바가 있다.
+본 연구가 고려하는 문제는 Steinberg와 Parks 그리고 Carraway 외가 살펴본 ``우선순위 배낭 문제''와는 피상적으로 비슷하지만, 우선순위 배낭 문제는 대학 지원 문제의 특색인 ``maximax'' 형태를 가지지는 않는다.
+또한 우선순위 모형과 달리, 본 연구에서 스칼라-값 목적함수를 `결과 분포'의 선호 순위로 대체하고자 노력하지 않는다.
+확률적 지배와 관련된 방법론적인 문제가 생기기 때문이다 (Sniedovich 1980).
+대신 $t_j$-값으로 유도된 학생의 `결과'에 대한 선호 순위를 그대로 받아들이고 위에서 정의한, 모호하지 않은 문제를 위한 효율적인 계산법에 집중한다.
+\fi
+
+\ifen
+We take special interest in the validity of greedy optimization algorithms for the set function $v(\mathcal{X})$, such as the algorithm that iteratively adds the school that elicits the greatest increase in the objective function until the budget is exhausted.
+Greedy algorithms produce a \emph{nested} family of solutions parameterized by the budget $H$: If $H \leq H'$, then the greedy solution for budget $H$ is a subset of the greedy solution for budget $H'$.
 As Rozanov and Tamir (2020) remark, the knowledge that the optima are nested aids not only in computing the optimal solution, but in the implementation thereof under uncertain information.
-For example, in the United States, many college applications are due at the beginning of November, and it is typical for students to begin working on their applications during the prior summer because colleges expect students to tailor their essays to the target school.
+For example, in the United States, many college applications are due at the beginning of November, and it is typical for students to begin working on their applications during the prior summer because colleges reward students who tailor their essays to the target school.
 However, students may not know how many schools they can afford to apply to until late October.
-The nestedness property---or equivalently, the validity of a greedy algorithm---implies that even in the absence of complete budget information, students can begin to carry out the optimal application strategy by writing essays for schools in the order that they enter the maximal portfolio.
+The nestedness property---or equivalently, the validity of a greedy algorithm---implies that even in the absence of complete budget information, students can begin to carry out the optimal application strategy by writing essays for schools in the order that they enter the optimal portfolio.
 \else
 본 연구에서 특히 탐욕 해법의 타당성에 관심을 기울인다.
-예를 들어 예산이 다 소비될 때까지 목적 함수를 가장 크게 증가시키는 학교를 차례대로 선택하는 알고리즘은 일종의 탐욕 해법이다.
-탐욕 해법은 예산 $H$으로 모수화된 해의 순서를 유도하며 해는 `포함 사슬 관계'(nestedness) 성질에 따라 서로에 연결된다.
-즉 $H \leq H'$일 때, 예산 $H$에 해당하는 탐욕 해는 예산 예산 $H'$에 해당하는 탐욕 해의 부분집합이다.
-Rozanov과 Tamir (2020)가 주장하듯, 최적해가 포함 사슬 관계를 가짐을 알면 최적해를 구하는 데일 뿐더러 정보가 불확실한 상황에서 최적해를 이행하는 데에도 유용하다.
-가령, 많은 미국 대학의 지원 기한은 11월 초인데 지원 에세이를 학교의 취향을 맞춰서 작성해야 하므로 여름부터 원서를 작성하는 학생이 많다.
+예를 들어 예산이 다 소비될 때까지 목적함수를 가장 많이 증가시키는 학교를 차례대로 선택하는 알고리즘은 일종의 탐욕 해법이다.
+탐욕 해법은 예산 $H$으로 모수화된 해의 순서를 유도하며 해는 `포함 사슬 관계'(nestedness)로 서로에 연결된다.
+즉 $H \leq H'$일 때, 예산 $H$에 해당하는 탐욕 해는 예산 $H'$에 해당하는 탐욕 해의 부분집합이다.
+Rozanov와 Tamir (2020)가 주장하듯, 최적해가 포함 사슬 관계를 맺음을 알면 최적해를 구하는 데일뿐더러 정보가 불확실한 상황에서 최적해를 이행하는 데에도 유용하다.
+가령, 많은 미국 대학의 지원 기한은 11월 초인데 학업 계획서를 학교의 취향에 맞춰서 작성해야 하므로 여름부터 원서를 작성하는 학생이 많다.
 그러나 시간이나 지원비에 대한 예산을 10월 말까지 모를 수도 있다.
@@
 
+\ifen
 For certain classes of optimization problems, such as maximizing a submodular set function over a cardinality constraint, a greedy algorithm is known to be a good approximate solution and exact under certain additional assumptions (Fisher et al.
@@
 For other problems, notably the binary knapsack problem, the most intuitive greedy algorithm can be made to perform arbitrarily poorly (Vazirani 2001).
-We show that the college application problem that the college application problem is analogous to the knapsack problem in many respects: In the special case where each $g_j = 1$, the optimal portfolio satisfy a nestedness property that is equivalent to the validity of the greedy algorithm.
+We show results for the college application problem that mirror those for knapsack: In the special case where each $g_j = 1$, the optimal portfolio satisfy a nestedness property that is equivalent to the validity of the greedy algorithm.
 This case mirrors the centralized college application process in Korea, where there is no application fee, but students are allowed to apply to only three schools during the main admissions cycle.
 Unfortunately, the nestedness property does not hold in the general case, nor does the greedy algorithm offers any performance guarantee.
-Instead, we offer a pseudopolynomial-time algorithm that is tractable for typical college market instances, as well as an approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in fully polynomial time.
+Instead, we offer a branch-and-bound routine, a pseudopolynomial-time algorithm that is tractable for typical college market instances, and an approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in fully polynomial time.
+\else
+한편, 탐욕 알고리즘이 좋은 근사 해법이며 다른 가정을 추가하면 정확한 알고리즘인 되는 최적화 모형 몇몇이 있으며 대표적인 바는 카디널리티 제약 위에 자모(子模) 집합 함수(submodular set function)를 최대화하는 문제이다 (Fisher et al.
+1978).
+반면에 이진 배낭 문제 같은 경우에서는 가장 직관적인 탐욕 알고리즘이 무한히 비최적인 해를 출력하는 인스턴스를 만들 수 있다 (Vazirani 2001).
+대학 지원 문제에 대해 배낭 문제와 매우 유사한 결과를 도출할 수 있다.
+모든 $g_j=1$인 특수한 경우에서, 최적 포트폴리오가 탐욕 알고리즘의 타당성과 동등한 포함 사슬 관계를 만족하는 것을 보인다.
+이 경우는 지원비가 없으며 정시 모집 기간에 학교 3개에만 지원할 수 있는 한국 입학 과정과 같다.
+불행하게도 일반적인 경우에서 포함 사슬 관계가 성립하지 않을뿐더러 탐욕 알고리즘에 대해 최소한 근사 계수를 보장할 수 없다.
+대신 분지한계법, 전형적인 입학 시장 인스턴스에 대해 효율적인 의사 다항 시간 동적 계획, 그리고 전체 다항 시간 안에 $(1 - \varepsilon)$-근사해를 출력하는 알고리즘과 같은 3가지 해법은 제안한다.
+
+\fi
 
 \ifen \subsection{Structure of this paper} \else \subsection{본 논문의 구성}\fi
-Section \ref{preliminaries} introduces some additional notation and assumptions that can be imposed with trivial loss of generality.
-
-In section \ref{homogappcosts}, we solve the special case where each $g_j = 1$ and $H$ is an integer $h \leq m$.
+\ifen Section \ref{preliminaries} introduces some additional notation and assumptions that can be imposed with trivial loss of generality. 
+\else \ref{preliminaries}절은 표기법은 명시하고 일방성을 잃지 않는 추가의 가정을 소개한다.\fi
+
+\ifen
+In section \ref{homogappcosts}, we consider the special case where each $g_j = 1$ and $H$ is an integer $h \leq m$.
 We show that an intuitive heuristic is in fact a $1/h$-approximation algorithm.
 Then, we show that the optimal portfolios are nested in the budget constraint, which yields an exact algorithm that runs in $O(hm)$-time.
-
+\else
+\ref{homogappcosts}절에서 각 $g_j = 1$이고 $H$가 $h \leq m$인 자연수인 특수한 경우를 고려한다. 직관적인 휴리스틱 해법이 실제로 $1/h$-근사 해법임을 보인다. 그다음에 최적 포트폴리오가 예산 제약에 대한 포함 사슬 관계를 보이고 이에 따라 $O(hm)$-시간 정확한 해법을 도출한다.
+\fi
+
+\ifen
 In section \ref{hetappcosts}, we turn to the scenario in which colleges differ in their application fees.
-We show that the decision form of the portfolio optimization problem is NP-complete through a reduction from the binary knapsack problem.
+We show that the decision form of the portfolio optimization problem is NP-complete through a polynomial reduction from the binary knapsack problem.
 We provide three algorithms for this more general setup.
@@
 The second is a dynamic program that iterates on total expenditures and produces an exact solution in pseudopolynomial time, namely $O(Hm + m \log m)$.
-The third is a dynamic program that iterates on truncated portfolio valuations.
+The third is a different dynamic program that iterates on truncated portfolio valuations.
 It yields a fully polynomial-time approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in $O(m^3 / \varepsilon)$ time.
-% subject to certain assumptions on the quality of the data available to students in their decisionmaking process, an optimal application portfolio for a single student can be computed in time polynomial in $h$ and $m$, as we show presently.
-
+\else
+\ref{hetappcosts}절은 각 학교의 지원비가 다를 수 있는 상황으로 계속한다.
+이진 배낭 문제에서의 다항 변환을 통해서 포트폴리오 최적화 문제가 NP-complete임을 증명한다.
+일반적인 상황을 위해 3개의 해법은 제안한다.
+첫째는 분지한계법이다.
+둘째는 총 지출액으로 탐색해서 $O(Hm + m \log m)$과 같은 의사 다항 시간 동적 계획이다.
+셋째는 분수를 내린 포트폴리오 가치로 탐색하는 다른 동적 계획이다.
+이를 통해서 $O(m^3 / \varepsilon)$-시간 안에 $(1 - \varepsilon)$-근사해를 출력하는 전체 다항 시간 근사 해법(fully polynomial-time approximation scheme)을 얻는다.
+\fi
+
+\ifen
 In section \ref{numericalexperiments}, we present the results of computational experiments that confirm the validity and time complexity results established in the previous two sections.
 
-% \footnote{This assumption is appropriate when $f$ gives the admissions probabilities \emph{specifically} for Alma, as estimated from her academic abilities.
-It is inappropriate to take a school's overall acceptance rate, which reflects the admissions probability for a \emph{generic} student, as $f_j$.}
-
-% Nestedness property and paper
+A brief conclusion follows.
+\else
+\ref{numericalexperiments}절에서 위에서 도출한 타당성과 계산 복잡성을 확인하는 계산적 실험 결과를 정리한다.
+
+간단한 결론이 뒤를 이은다.
+\fi
 
@@
 \begin{align}
-v(\mathcal{X}) &= \sum_{j=0}^m t_j p_j(\mathcal{X}) = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)  \label{closedformportfoliovaluationX}%, \quad \text{or equivalently,}\\
+v(\mathcal{X}) &= \sum_{j=0}^m t_j p_j(\mathcal{X}) = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)  \label{closedformportfoliovaluationX}%, \quad \text{or equivalently,}\\
 %\qquad v(x) &= t_0 \prod_{j=1}^m (1 - f_{j} x_j) + \sum_{j=1}^m \Bigl( x_j t_j f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’}) \Bigr) \label{closedformportfoliovaluationx}
@@
 
-\ifen \section{Homogeneous application costs}  \else \subsection{동일한 지원 비용} \fi \label{homogappcosts}
+\ifen \section{Homogeneous application costs}  \else \section{동일한 지원 비용} \fi \label{homogappcosts}
 In this section, we derive a polynomial-time algorithm for the special case in which $g_j = 1$ and $H$ is a constant $h \leq m$.
@@
 %\{0\}\cup
-\mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i})\\
+\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)\\
 \text{subject to}\quad & \mathcal{X}\subseteq\mathcal{C}, \quad|\mathcal{X}| \leq h 
@@
 \ifen \subsection{Approximation properties of a na\"ive solution}  \else \subsection{Na\"ive 해법의 근사 성질} \fi
-The expected utility associated with a single school $c_j$ is simply $\operatorname{E}[t_j Z_j] = t_j f_j$. It is therefore tempting to adopt the following strategy, which turns out to be inoptimal.
+The expected utility associated with a single school $c_j$ is simply $\operatorname{E}[t_j Z_j] = f_j t_j$. It is therefore tempting to adopt the following strategy, which turns out to be inoptimal.
 \begin{definition}[Na\"ive algorithm for Alma’s problem] \label{naivealgorithm}
-Apply to the $h$ schools having the highest expected utility $t_j f_j$.
+Apply to the $h$ schools having the highest expected utility $f_j t_j$.
 \end{definition}
@@
 \begin{theorem} \label{oneoverhopt}
-When the application limit is $h$, let $\mathcal{X}_h$ denote the optimal portfolio, and $\mathcal{T}_h$ the set of the $h$ schools having the largest values of $t_j f_j$.
+When the application limit is $h$, let $\mathcal{X}_h$ denote the optimal portfolio, and $\mathcal{T}_h$ the set of the $h$ schools having the largest values of $f_j t_j$.
 Then $v(\mathcal{T}_h) / v(\mathcal{X}_h) \geq 1/h$.
@@
 \end{align*}
-Since all $t_j f_j = 1$, the na\"ive algorithm can choose $\mathcal{T}_h = \{1, \dots, h\}$, with $v(\mathcal{T}_h) = 1$. But the optimal solution is $\mathcal{X}_h = \{h+1, \dots, m\}$, with
+Since all $f_j t_j = 1$, the na\"ive algorithm can choose $\mathcal{T}_h = \{1, \dots, h\}$, with $v(\mathcal{T}_h) = 1$. But the optimal solution is $\mathcal{X}_h = \{h+1, \dots, m\}$, with
 \begin{equation*}
-v(\mathcal{X}_h) = \sum_{j= h +1}^m \Bigl( t_j f_j \prod_{j' = j+1}^m (1 - f_{j'}) \Bigr) =  \sum_{j= 1}^h  (1 - \varepsilon)^{j} \approx h.
+v(\mathcal{X}_h) = \sum_{j= h +1}^m \Bigl( f_j t_j \prod_{j' = j+1}^m (1 - f_{j'}) \Bigr) =  \sum_{j= 1}^h  (1 - \varepsilon)^{j} \approx h.
 \end{equation*}
@@
 
-\ifen \subsection{The nestedness property}  \else \subsection{포함 사슬 관계 성질} \fi
+\ifen \subsection{The nestedness property}  \else \subsection{포함 사슬 관계} \fi
 It turns out that the solution to Alma's problem possesses a special structure: An optimal portfolio of size $h+1$ includes an optimal portfolio of size $h$ as a subset.
@@
 To construct $w(\mathcal{Y})$, let $\tilde t_j$ denote the expected utility Alma receives from school $c_j$ \emph{given} that she has been admitted to $c_j$ and applied to $c_k$.
-For $j < k$, including $j = 0$, this is $\tilde t_j = t_j (1- f_k) + t_k f_k$; for $j > k $, this is $\tilde t_j = t_j$.
+For $j < k$, including $j = 0$, this is $\tilde t_j = (1- f_k) t_j + f_k t_k$; for $j > k $, this is $\tilde t_j = t_j$.
 This means that
@@
 = \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y})- \tilde t_0
-= v(\mathcal{Y}\cup\{k\}) -  t_k f_k \end{equation}
+= v(\mathcal{Y}\cup\{k\}) -  f_k t_k \end{equation}
 where the second equality follows from Theorem \ref{assumetzerozero}.
@@
 \begin{cases}
-t_j (1 - f_k), \quad & t_j \leq t_k \\
-t_j - t_k f_k, \quad& t_j > t_k
+(1 - f_k) t_j, \quad & t_j \leq t_k \\
+t_j - f_k t_k, \quad& t_j > t_k
 \end{cases}
@@
 
+\ifen \subsection{A small example} \else \subsection{작은 예} \fi
+Let us consider a fictional market consisting of $m=8$ schools. The market data, along with the optimal solutions for each $h \leq m$, appear in Table \ref{planetsdata}.
+
+Below are shown the first several iterations of Algorithm \ref{algorithmforlargeh}.
+The values of $f_j$, $t_j$, and their product are recorded only for the schools remaining in $\mathcal{C}$.
+$f * t$, where $(f * t)_j = f_j t_j$, denotes the entrywise product of $f$ and $t$.
+The school added at each iteration, underlined, is the one whose $f_j t_j$-value is greatest.
+\begin{align*}
+\text{Iteration 1:}
+&&C &= \{1, 2, 3, 4, 5, 6, 7, 8\} \\
+&&f &= \{0.39, 0.33, 0.24, 0.24, 0.05, 0.03, 0.1, 0.12\} \\
+&&t &= \{200, 250, 300, 350, 400, 450, 500, 550\} \\
+&&f * t &= \{78.0, 82.5, 72.0, \underline{84.0}, 20.0, 13.5, 50.0, 66.0\} 
+~~\tag*{\(\implies \mathcal{X}_3 = \{4\} \)}\\
+\text{Iteration 2:}
+&&C &= \{1, 2, 3, 5, 6, 7, 8\} \\
+&&f &= \{0.39, 0.33, 0.24, 0.05, 0.03, 0.1, 0.12\} \\
+&&t &= \{152, 190, 228, 316, 366, 416, 466\} \\
+&&f * t &= \{59.28, \underline{62.7}, 54.72, 15.8, 10.98, 41.6, 55.92\} 
+~~\tag*{\(\implies \mathcal{X}_3 = \{4, 2\} \)}\\
+\text{Iteration 3:}
+&&C &= \{1, 3, 5, 6, 7, 8\} \\
+&&f &= \{0.39, 0.24, 0.05, 0.03, 0.1, 0.12\} \\
+&&t &= \{101.84, 165.3, 253.3, 303.3, 353.3, 403.3\} \\
+&&f * t &= \{39.718, 39.672, 12.665, 9.099, 35.33, \underline{48.396}\} 
+~~\tag*{\(\implies \mathcal{X}_3 = \{4, 2, 8\} \)}\\
+&&&\cdots\\
+\text{Iteration 8:}
+&&C &= \{6\}, f = \{0.03\}, t = \{177.622\}, f * t= \{\underline{5.329}\}
+~~\tag*{\(\implies \mathcal{X}_3 = \{4, 2, 8, 1, 7, 3, 5, 6\} \)}
+\end{align*}
+The output of the algorithm is $\mathtt{X} = [4, 2, 8, 1, 7, 3, 5, 6]$, and the optimal $h$-portfolio consists of its first $h$ entries.
+The ``priority'' column of Table \ref{planetsdata} shows the inverse permutation of $\mathtt{X}$, which is the minimum value of $h$ for which the school is included in the optimal portfolio.
+Figure \ref{planetsplot} shows the value of the optimal portfolio as a function of $h$.
+The concave shape of the plot suggests the result of Theorem \ref{concavityinh}.
+
+\begin{table} \centering
+\begin{tabular}{clcccc}
+\ifen\textbf{Index $j$} & \textbf{School $c_j$} & \textbf{Utility $t_j$} & \textbf{Admit prob. $f_j$} & \textbf{Priority} & \textbf{$v(\mathcal{X}_h)$} \\ \hline
+\else 
+\textbf{지표 $j$} & \textbf{학교 $c_j$} & \textbf{효용 $t_j$} & \textbf{합격 확률 $f_j$} & \textbf{지원 순위} & \textbf{$v(\mathcal{X}_h)$} \\ \hline  \fi
+\\[-.75em]
+1 & \ifen Mercury University   \else  수성대       \fi        & 200 & 0.39 & 4   & 230.0   \\
+2 & \ifen Venus University     \else  금성대       \fi        & 250 & 0.33 & 2   & 146.7  \\
+3 & \ifen Mars University      \else  화성대       \fi        & 300 & 0.24 & 6   & 281.5  \\
+4 & \ifen Jupiter University   \else  목성대       \fi        & 350 & 0.24 & 1   & \phantom{0}84.0  \\
+5 & \ifen Saturn University    \else  토성대       \fi        & 400 & 0.05 & 7   & 288.8  \\
+6 & \ifen Uranus University    \else  천왕성대     \fi        & 450 & 0.03 & 8   & 294.1  \\
+7 & \ifen Neptune University   \else  해왕성대     \fi        & 500 & 0.10 & 5   & 257.7  \\
+8 & \ifen Pluto College        \else  명왕성대       \fi        & 550 & 0.12 & 3   & 195.1              
+\end{tabular}
+\caption{Market data and optimal application portfolios for a fictional market with $m=8$ schools.
+By the nestedness property (Theorem \ref{nestedapplication}), the optimal portfolio when the application limit is $h$ consists of the $h$ schools having priority number $h$ or less.
+\label{planetsdata}}
+\end{table}
+
+\begin{figure}[h!] 
+ \centering
+ \includegraphics[width=0.95\textwidth]{./plots/h_v-example.pdf}
+  \caption{  \label{planetsplot}
+  \ifen Application limit $h$ versus the optimal portfolio valuation $v^* = v(\mathcal{X}_h)$ in a fictional market with $m=8$ schools.
+  \else $m=8$개의 학교로 구성된 가상 입학 시상에서, 각 지원 제한 $h$에 해당하는 최적 포트폴리오의 가치 $v^* = v(\mathcal{X}_h)$.\fi}
+\end{figure}
+
 \pagebreak
@@
 %\{0\}\cup
-\mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)\\
+\mathcal{X}} \Bigl(f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)\\
 \text{subject to}\quad & \mathcal{X}\subseteq\mathcal{C},~~\sum_{j \in \mathcal{X}} g_j \leq H 
@@
 \begin{split}
-\sum_{j \in \mathcal{X}} u_j &=  \sum_{j \in \mathcal{X}} t_j f_j 
-> \sum_{j \in \mathcal{X}} \Bigl( t_j f_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - f_{j’}) \Bigr)
+\sum_{j \in \mathcal{X}} u_j &=  \sum_{j \in \mathcal{X}} f_j t_j
+> \sum_{j \in \mathcal{X}} \Bigl( f_j t_j \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - f_{j’}) \Bigr)
 = v(\mathcal{X}) \\
@@
 \ifen \subsection{Branch-and-bound algorithm} \else \subsection{분지한계법} \fi
-A traditional approach to knapsack problems is the branch-and-bound framework, which generates subproblems in which the values of one or more decision variables are fixed and uses an upper bound on the objective function to exclude, or \emph{fathom,} branches of the tree of subproblems that cannot yield a solution better than the best solution on hand (Martello and Toth 1990; Kellerer et al.
+A traditional approach to knapsack problems is the branch-and-bound framework, which generates subproblems in which the values of one or more decision variables are fixed and uses an upper bound on the objective function to exclude, or \emph{fathom,} branches of the decision tree that cannot yield a solution better than the best solution on hand (Martello and Toth 1990; Kellerer et al.
 2004).
@@
 \begin{split}
-\text{maximize}\quad &  v(x) = \sum_{j=1}^m \Bigl( t_j f_j  x_j \prod_{i > j} (1 - f_{i} x_i) \Bigr)\\
+\text{maximize}\quad &  v(x) = \sum_{j=1}^m \Bigl(f_j t_j  x_j \prod_{i > j} (1 - f_{i} x_i) \Bigr)\\
 \text{subject to}\quad & \sum_{j=1}^m g_j x_j \leq H, ~~ x_i \in \{0, 1\}^m
@@
 \begin{split}
-\text{maximize}\quad &  v_{\mathrm{LP}}(x) = \sum_{j=1}^m  t_j f_j  x_j \\
+\text{maximize}\quad &  v_{\mathrm{LP}}(x) = \sum_{j=1}^m  f_j t_j x_j \\
 \text{subject to}\quad & \sum_{j=1}^m g_j x_j \leq H, ~~ x \in [0, 1]^m
@@
 Problem \ref{LPrelaxation} is a continuous knapsack problem, which is easily solved in $O(m \log m)$ time by the following greedy algorithm: Start with $x = \mathbf{0}$.
-While $H > 0$, select the school $k$ for which $t_j f_j / g_j$ is highest, set $x_k \gets \min\{1, H / g_k\}$, and set $H \gets H - g_k$ (Dantzig 1957).
-
-In our branch-and-bound architecture, a \emph{node} is characterized by a three-way partition of schools $\mathcal{C}= \mathcal{I} \cup \mathcal{O} \cup \mathcal{N}$ satisfying $\sum_{j \in \mathcal{I}} g_j \leq H$.
+While $H > 0$, select the school $k$ for which $f_j t_j / g_j$ is highest, set $x_k \gets \min\{1, H / g_k\}$, and set $H \gets H - g_k$ (Dantzig 1957).
+
+In our branch-and-bound framework, a \emph{node} is characterized by a three-way partition of schools $\mathcal{C}= \mathcal{I} \cup \mathcal{O} \cup \mathcal{N}$ satisfying $\sum_{j \in \mathcal{I}} g_j \leq H$.
 $\mathcal{I}$ consists of schools that are ``in'' the application portfolio, $\mathcal{O}$ consists of those that are ``out,'' and $\mathcal{N}$ consists of those that are ``negotiable.''
@@
 \begin{split}
-\text{maximize}\quad &  w(y) = \gamma + \sum_{j \in \mathcal{N}} \Bigl( \bar t_j f_j  y_j \prod_{\substack{i \in \mathcal{N}:\\i > j}} (1 - f_{i} y_i) \Bigr)\\
-\text{subject to}\quad & \sum_{j \in \mathcal{N}} g_j y_j \leq \bar H; ~~ y_j \in \{0, 1\},~~j \in \mathcal{N}.
+\text{maximize}\quad &  v(x) = \gamma + \sum_{j \in \mathcal{N}} \Bigl(f_j \bar t_j x_j \prod_{\substack{i \in \mathcal{N}:\\i > j}} (1 - f_{i} x_i) \Bigr)\\
+\text{subject to}\quad & \sum_{j \in \mathcal{N}} g_j x_j \leq \bar H; ~~ x_j \in \{0, 1\},~~j \in \mathcal{N}.
 \end{split}
@@
 \begin{split}
-\text{maximize}\quad &  w_{\mathrm{LP}}(y)= \gamma + \sum_{j \in \mathcal{N}}  \bar t_j f_j  y_j \\
-\text{subject to}\quad & \sum_{j \in \mathcal{N}} g_j y_j \leq \bar H; ~~ y_j \in [0, 1],~~j \in \mathcal{N}
+\text{maximize}\quad &  w_{\mathrm{LP}}(x)= \gamma + \sum_{j \in \mathcal{N}}  f_j \bar t_j  x_j \\
+\text{subject to}\quad & \sum_{j \in \mathcal{N}} g_j x_j \leq \bar H; ~~ x_j \in [0, 1],~~j \in \mathcal{N}
 \end{split}
@@
 Given a node $n  = (\mathcal{I}, \mathcal{O}, \mathcal{N})$, its children are generated as follows.
-Every node has zero, one, or two children.
+Every node has two, one, or zero children.
 In the typical case, we select a school $j \in \mathcal{N}$ for which $g_j \leq \bar H$ and generate one child by moving $j$ to $\mathcal{I}$, and another child by moving $j$ it to $\mathcal{O}$.
 Equivalently, we set $x_j = 1$ in one child and $x_j = 0$ in the other.
-In principle, any school in $\mathcal{N}$ can be chosen for $j$, but as a greedy heuristic, we choose the school for which the ratio $t_j f_j / g_j$ is highest.
+In principle, any school in $\mathcal{N}$ can be chosen for $j$, but as a greedy heuristic, we choose the school for which the ratio $f_j \bar t_j / g_j$ is highest.
 Notice that this method of generating children ensures that each node's $\mathcal{I}$-set differs from its parent's by at most a single school, so the constant $\gamma$ and transformed $\bar t_j$-values for the new node can be obtained by a single application of \eqref{howtotransformfj}.
@@
 
-In the branch-and-bound algorithm, we store nodes in a tree $\mathfrak{T}$, which contains nodes alongside pointers to their children.
+\begin{example}
+Consider a market in which $t = (20, 40, 60, 80, 100)$, $f = (0.5, 0.5, 0.5, 0.5, 0.5)$, $g = (3, 2, 3, 2, 3)$, and $h = 8$, and the node $n =  (\mathcal{I}, \mathcal{O}, \mathcal{N}) = (\{2, 5\}, \{1\}, \{3, 4\})$.
+Let us compute the two subproblems associated with $n$ and generate its children.
+To compute the subproblems, we first simple disregard school 1.
+Next, to eliminate school 2, we apply \eqref{howtotransformfj} to obtain
+$\{\bar{\bar t}_3, \bar{\bar t}_4, \bar{\bar t}_5\} = \{ (1 - f_2) t_3, (1 - f_2) t_4, (1 -f_2) t_5\} = \{30, 40, 50 \}$ and $\bar{\bar \gamma} = f_2 t_2 = 20$.
+We eliminate school 5 by applying \eqref{howtotransformfj} to $\bar{ \bar t}$ to obtain $\{\bar t_3, \bar t_4\} =  \{\bar{\bar t}_3 - f_5 \bar{\bar t}_5, \bar{\bar t}_4-  f_5 \bar{\bar t}_5\}= \{5, 15\}$ and $\gamma = \bar{\bar \gamma} + f_5 \bar{\bar t}_5 = 35$.
+Finally, $\bar H = H - g_2 - g_5 = 3$.
+Problems \ref{nodenlpsubproblem} and \ref{nodelprelaxation} are now easily obtained by substitution.
+
+Since all the schools in $\mathcal{N}$ have $g_j \leq \bar H$, $n$ has two children.
+Applying the node-generation rule, school 4 has the highest $f_j \bar t_j / g_j$-ratio, so the children are $n_1 = (\{2, 4, 5\}, \{1\}, \{3\})$ and $n_2 = (\{2, 5\}, \{1, 4\}, \{3\})$.
+\end{example}
+
+To implement the branch-and-bound algorithm, we represent the set of candidate nodes---namely, nonleaf nodes whose children have not yet been generated---by $\mathfrak{T}$.
 Each time a node $n= (\mathcal{I}, \mathcal{O}, \mathcal{N})$ is generated, we record the values $v_{\mathcal{I}}[n] = v(\mathcal{I})$ and $v^*_{\mathrm{LP}}[n]$, the optimal objective value of the LP relaxation \eqref{nodelprelaxation}.
@@
 The algorithm is initialized by populating $\mathfrak{T}$ with the root node $n_0 = (\O, \O, \mathcal{C})$.
-The algorithm terminates when all the nodes that remain on the tree are either leaf nodes, or have had their children explored.
+At each iteration, it selects the node $n \in \mathfrak{T}$ having the highest $v_{\mathrm{LP}}^*$-value, generates its children, and removes $n$ from the candidate set.
+Next, the children $n'$ of $n$ are inspected; if one of them yields a new optimal solution, then we mark it as the best candidate and fathom any nodes $n''$ for which  $v_{\mathcal{I}}[n'] >  v_{\mathrm{LP}}^*[n'']$.
+When no nodes remain in the candidate set, the algorithm has explored every branch, so it terminates and returns the best candidate.
 
@@
 Root node $n_0 \gets (\O, \O, \mathcal{C})$\;
-Current lower bound $L \gets 0$\;
-Initialize tree $\mathfrak{T} \gets \{ n_0\}$\;
+Current lower bound $L \gets 0$ and best solution $\mathcal{X} \gets \O$\;
+Initialize candidate set $\mathfrak{T} \gets \{ n_0\}$\;
 \While{not finished}{
-	$\mathfrak{S} \gets$ the set of non-leaf nodes whose children have not yet been generated\;
-	\uIf{$\mathfrak{S} = \O$}{
-		$\mathcal{X} \gets \mathcal{I}$ from node associated with current value of $L$\;
-		\Return{$\mathcal{X}, L$}
+	\lIf{$\mathfrak{T} = \O$}{\Return{$\mathcal{X}, L$}}
+	\Else{
+		$n \gets \argmax\{ v_{\mathrm{LP}}^*[n] : n \in \mathfrak{T}\}$\;
+		Remove $n$ from $\mathfrak{T}$\;
+%		Generate children of $n$\;
+		\For{each child $n'$ of $n$}{
+			\If{$L < v_{\mathcal{I}}[n']$}{
+				$L \gets v_{\mathcal{I}}[n']$\;
+				Update $\mathcal{X}$ to the $\mathcal{I}$-set associated with $n'$\;
 			}
-	\Else{
-		$n \gets \argmax\{ v_{\mathrm{LP}}^*[n] : n \in \mathfrak{S}\}$\;
-		Generate children of $n$\;
-		\For{each child $n'$ of $n$}{
-			$L \gets \max\{L, v_{\mathcal{I}}[n']\}$\;
-		}
+			\lIf{$n'$ is not a leaf}{add $n'$ to $\mathfrak{T}$}
 		}
 		\For{$n''$ in $\mathfrak{T}$} {
-		\lIf{$L >  v_{\mathrm{LP}}^*[n'']$}{fathom $n''$}
+			\lIf{$L >  v_{\mathrm{LP}}^*[n'']$}{remove $n''$ from $\mathfrak{T}$}
+		}
 	}
@@
 \begin{proof}
-The discussion above implies that the algorithm only fathoms branches when justified; therefore, if the algorithm terminates, then an optimal solution exists among the leaves of the tree, and it is therefore returned.
-
-To show that the algorithm does not cycle, it suffices to show that no node is generated twice.
+The discussion above implies that the as if the algorithm terminates, then an optimal solution exists among the leaves of the tree, and it is therefore returned.
+
+Since the set of possible nodes is finite, to show that the algorithm does not cycle, it suffices to show that no node is generated twice.
 Suppose not: that two distinct nodes $n_1$ and $n_2$ share the same partition $(\mathcal{I}_{12}, \mathcal{O}_{12}, \mathcal{N}_{12})$.
@@
 The branch-and-bound algorithm is an interesting benchmark, but as a kind of enumeration algorithm, its computation time grows rapidly in the problem size, and unlike the approximation scheme we propose later on, there is no guaranteed bound on the approximation error after a fixed number of iterations.
-In our numerical experiments, Algorithm \ref{ellisbnb} was impractical for instances containing more than about $m=40$ schools.
+Moreover, when there are many schools in $\mathcal{N}$, the LP upper bound may be much higher than $v_{\mathcal{I}}[n']$.
+This means that significant fathoming operations do not occur until the algorithm has explored deep into the tree, at which point the bulk of the computational effort has already been exhausted.
+In our numerical experiments, Algorithm \ref{ellisbnb} was ineffective on instances larger than about $m=30$ schools.
 
@@
 
-(Computation time.) The bottleneck step is the creation of the lookup table in line \ref{createdlookuptable}, whose size is $O(|\mathcal{V}| m)$. Since
+(Computation time.) The bottleneck step is the creation of the lookup table in line \ref{createdlookuptable}, whose size is $m \times |\mathcal{V}|$. Since
 \begin{equation}
@@
 To achieve partial negative correlation between $t_j$ and $f_j$, we then set $f_j = 1 / (t_j + 10 Q)$, where $Q$ is drawn uniformly from the interval $[0, 1)$.
+A typical instance is shown in Figure \ref{samplemarket}.
+
 In Experiment 1, which concerns Algorithm \ref{algorithmforlargeh}, we take each $g_j = 1$ and set $H = h = \lfloor m/ 2 \rfloor$.
-In Experiment 2, which concerns Algorithms \ref{ellisDP1} and \ref{ellisDP3}, each $g_j$ is drawn uniformly from the set $\{5, \dots, 10\}$ and we take $H = \lfloor \frac{1}{2} \sum g_j \rfloor$.
-At each combination of the experimental variables, we generated 50 markets, and each computation was repeated three times, with the fastest of the three  recorded as the computation time.
-Therefore, each cell of each table represents 150 computations.
+In Experiment 2, which concerns Algorithms \ref{ellisbnb}, \ref{ellisDP1}, and \ref{ellisDP3}, each $g_j$ is drawn uniformly from the set $\{5, \dots, 10\}$ and we take $H = \lfloor \frac{1}{2} \sum g_j \rfloor$.
+
+At each combination of the experimental variables, we generated ten markets, and each computation was repeated three times, with the fastest of the three  recorded as the computation time.
+Therefore, each cell of each table represents 30 computations.
 We report the mean and standard deviation across the 50 markets.
@@
 \begin{figure}[h!] 
-  \begin{center}\includegraphics[width=0.75\textwidth]{./plots/samplemarket.pdf}\end{center}
-  \caption{A typical randomly-generated instance with $m=60$ schools. }
+\centering
+\includegraphics[width=0.75\textwidth]{./plots/samplemarket.pdf}
+  \caption{\label{samplemarket}
+ \ifen A typical randomly-generated instance with $m=60$ schools and its optimal application portfolio.
+The application costs $g_j$ were drawn uniformly from $\{5, \dots, 10\}$.
+The optimal portfolio was computed using Algorithm \ref{ellisDP1}.
+ \else $m=60$개의 학교로 구성된 전형적인 무작위로 생성한 인스턴스와 해당 최적 포트폴리오. 지원비 $g_j$는 $\{5, \dots, 10\}$에서 균일한 확률로 선정했으며 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
 \end{figure}
@@
 \ifen \else
-김민희. 2015. ``{[대입 수시 전략]} 총 6번의 기회 \textellipsis `상향·소신·안정' 분산 지원하라.'' 중앙일보, 8월 26일. \url{https://www.joongang.co.kr/article/18524069}.\fi
+김민희. 2015. ``{[대입 수시 전략]} 총 6번의 기회 $\cdots$ `상향·소신·안정' 분산 지원하라.'' 중앙일보, 8월 26일. \url{https://www.joongang.co.kr/article/18524069}.\fi
 
@@
 
+Carraway, Robert, Robert Schmidt, and Lawrence Weatherford.
+1993.
+``An algorithm for maximizing target achievement in the stochastic knapsack problem with normal returns.'' \emph{Naval Research Logistics} 40 (2): 161--73.
+\url{https://doi.org/10.1002/nav.3220400203}.
+
 Dantzig, George B. 1957. ``Discrete-Variable Extremum Problems.'' \emph{Operations Research} 5 (2): 266--88.
+
+Dean, Brian, Michel Goemans, and Jan Vondr\'ak.
+2008.
+``Approximating the Stochastic Knapsack Problem: The Benefit of Adaptivity.'' \emph{Mathematics of Operations Research} 33 (4): 945--64.
+\url{https://doi.org/10.1287/moor.1080.0330}.
 
@@
 
+Markowitz, Harry. 1952. ``Portfolio Selection.'' \emph{The Journal of Finance} 7 (1): 77--91. \url{https://www.jstor.org/stable/2975974}.
+
 Martello, Silvano and Paolo Toth. 1990. \emph{Knapsack Problems: Algorithms and Computer Implementations.} New York: John Wiley \& Sons.
@@
 
-% p. 70: FPTAS for knapsack, should work for Ellis's problem
+Sniedovich, Moshe.
+1980.
+``Preference Order Stochastic Knapsack Problems: Methodological Issues.'' \emph{The Journal of the Operational Research Society} 31 (11): 1025--32.
+\url{https://www.jstor.org/stable/2581283}.
+
+Steinberg, E.
+and M.
+S.
+Parks.
+1979.
+``A Preference Order Dynamic Program for a Knapsack Problem with Stochastic Rewards.'' \emph{The Journal of the Operational Research Society} 30 (2): 141--47.
+\url{https://www.jstor.org/stable/3009295}.
+
 Vazirani, Vijay. 2001. \emph{Approximation Algorithms.} Berlin: Springer.

### 2022-03-07 src/FastSolvers.jl

@@ -1,4 +1,3 @@
-
 """
     College(f, t, ft, omf)
 
@@ -32,7 +31,7 @@ application costs and the corresponding portfolio valuations.
 """
 function applicationorder(
     mkt::SameCostsMarket;
-    datastructure = :heap::Symbol)::Tuple{Vector{Int64},Vector{Float64}}
+    datastructure = :heap::Symbol, verbose=false::Bool)::Tuple{Vector{Int64},Vector{Float64}}
 
     apporder = zeros(Int, mkt.h)
     v = zeros(mkt.h)
@@ -43,7 +42,16 @@ function applicationorder(
         end
 
         onheap = Set(1:mkt.m)
+    
         for j in 1:mkt.h
+            if verbose
+                K = sort(collect(onheap))
+                println("C = ", K)
+                println("f = ", [mkt_heap[j].f for j in K])
+                println("t = ", [mkt_heap[j].t for j in K])
+                println("ft= ", [mkt_heap[j].ft for j in K])
+            end
+
             c_k, k = top_with_handle(mkt_heap)
             v[j] = get(v, j - 1, 0) + c_k.ft
 
@@ -64,8 +72,16 @@ function applicationorder(
         for j in 1:mkt.m
             push!(mkt_dict, j => College(mkt.f[j], mkt.t[j], mkt.ft[j], mkt.omf[j]))
         end
-
+        
         for j in 1:mkt.h
+            if verbose
+                K = sort(collect(keys(mkt_dict)))
+                println("C = ", K)
+                println("f = ", [mkt_dict[j].f for j in K])
+                println("t = ", [mkt_dict[j].t for j in K])
+                println("ft= ", [mkt_dict[j].ft for j in K])
+            end
+
             c_k, k = findmax(mkt_dict)
             v[j] = get(v, j - 1, 0) + c_k.ft
 

### 2022-03-07 src/OptimalApplication.jl

@@ -3,6 +3,7 @@ module OptimalApplication
 using Combinatorics: combinations, multiset_combinations
 using DataStructures
 import Base.isless
+import Base.hash
 
 
 export

### 2022-03-07 src/VariedCostsBnB.jl

@@ -128,10 +128,16 @@ function optimalportfolio_branchbound(mkt; maxit = 1000000::Integer, verbose = f
     LB = 0
     LB_node = rootnode
 
-    tree = MutableBinaryMaxHeap{Node{eltype(C)}}()
-    treekeys = Set{Int64}()
+    # Commented out is an implementation that stores tree as a heap. The dict is faster because
+    # fathoming nodes is slow on the heap. 
 
-    push!(treekeys, push!(tree, rootnode))
+    # tree = MutableBinaryMaxHeap{Node{eltype(C)}}()        
+    # treekeys = Set{Int64}()
+
+    # push!(treekeys, push!(tree, rootnode))
+
+    tree = Dict{UInt64,Node}()
+    push!(tree, hash(rootnode) => rootnode)
 
     for k in 1:maxit
         verbose && @show k, LB, length(tree)
@@ -141,9 +147,12 @@ function optimalportfolio_branchbound(mkt; maxit = 1000000::Integer, verbose = f
             return collect(LB_node.I), LB
         else
             # Select the node with the highest UB
-            thisnode, thisnodehandle = top_with_handle(tree)
-            pop!(tree)
-            delete!(treekeys, thisnodehandle)
+            # thisnode, thisnodehandle = top_with_handle(tree)
+            # pop!(tree)
+            # delete!(treekeys, thisnodehandle)
+        
+            thisnode, thisnodehandle = findmax(tree)
+            delete!(tree, thisnodehandle)
         end
 
         children = generatechildren(thisnode, mkt)
@@ -156,15 +165,17 @@ function optimalportfolio_branchbound(mkt; maxit = 1000000::Integer, verbose = f
                 LB = child.v_I
             end
 
-            isempty(child.N) || push!(treekeys, push!(tree, child))
+            # isempty(child.N) || push!(treekeys, push!(tree, child))
+            isempty(child.N) || push!(tree, hash(child) => child)
         end
 
         if newLB
-            for k in treekeys
+            # for k in treekeys
+            for k in keys(tree)
                 if tree[k].v_LP < LB
                     verbose && println("Fathoming node $k")
                     delete!(tree, k)
-                    delete!(treekeys, k)
+                    # delete!(treekeys, k)
                 end
             end
         end

### 2022-03-08 paper/paper.tex
translations, better printing in benchmark

@@
 %!TEX encoding = UTF-8 Unicode 
-\documentclass[12pt]{article} %try amsproc, amsart
+\documentclass[11pt]{article} %try amsproc, amsart
 \usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
@@
 
-\usepackage{times}
+%\usepackage{times}
 \usepackage{letltxmacro}
@@
 \renewcommand{\figurename}{그림}
-\fi
-
-\numberwithin{equation}{section}
+\renewcommand{\tablename}{표}
+\fi
+
+%\numberwithin{equation}{section}
 %\numberwithin{theorem}{section}
@@
 
-\pagestyle{fancy}
-\renewcommand{\headrulewidth}{0pt}
-\fancyhf{}
-\rhead{How to apply to college \\~}
-\lhead{Max Kapur \\ \date{\today}}
-\cfoot{\thepage}
-
 \ifen { \title{How to apply to college} }
@@
 \date{\today}                % Activate to display a given date or no date
+
+\pagestyle{fancy}
+\renewcommand{\headrulewidth}{0pt}
+\fancyhf{}
+\rhead{How to apply to college \\}
+\lhead{Max Kapur \\ \date{\today}}
+\cfoot{\thepage}
 
@@
 Risk management is implicit in the college application problem because, in a typical admissions market, college preferability is negatively correlated with competitiveness.
-That is, students negotiate a tradeoff between highly attractive, selective “reach schools” and less preferable “safety schools” where admission is a safer bet (Kim 2015).
+That is, students negotiate a tradeoff between attractive, selective “reach schools” and less preferable “safety schools” where admission is a safer bet (Kim 2015).
 Finally, the combinatorial nature of the college application problem makes it difficult to solve using the gradient-based techniques used in continuous portfolio optimization.
@@
 또한 전형적인 입학 시장에서 대학의 효용과 합격 확률이 서로 반비례하므로 대학 지원 문제에 위험 관리가 내포되어 있다.
-특히 선호도가 높으며 붙기 어려운 “상향학교”(reach school)와 선호도가 낮으며 붙기 쉬운 “안정학교”(safety school) 사이의 균형을 고려해야 한다 (김민희 2015).
-마지막, 대학 지원의 조합적인 본성으로 인해 연속적인 포트폴리오 최적화 문제에서 흔히 사용하는 기울기 방법으로는 풀기가 어렵다.
-Chao는 지원을 제약 조건 대신 비용으로 모형화했으며 모형을 추정하기 위해 $m=8$이 되도록 학교를 먼저 클러스터로 모았다.
+특히 선호도가 높으며 붙기 어려운 “상향” 지원 학교(reach school)와 선호도가 낮으며 붙기 쉬운 “안정” 지원 학교(safety school) 사이의 균형을 고려해야 한다 (김민희 2015).
+마지막, 대학 지원의 조합적인 본성으로 인해 연속적인 포트폴리오 최적화 문제에서 흔히 사용하는 기울기 해법으로는 풀기가 어렵다.
+Chao는 지원 비용을 제약 조건 대신 목적함수의 한 항으로 모형화했으며 모형을 추정하기 위해 $m=8$이 되도록 학교를 먼저 클러스터로 모았다.
 이는 열거법을 통해 문제를 쉽게 풀 수 있는 규모이지만 본 연구는 더 일반적인 해법을 목표로 추구한다.
@@
 \ifen
-The integer formulation of the college application problem, given below, can also be viewed as a kind of binary knapsack problem with a polynomial objective function of degree $m$.
+The integer formulation of the college application problem can also be viewed as a kind of binary knapsack problem with a polynomial objective function of degree $m$.
 Our branch-and-bound and dynamic programming algorithms closely resemble existing algorithms for knapsack problems (Martello and Toth 1990, \S\,2.5--6).
@@
 Our problem superficially resembles the preference-order knapsack problem considered by Steinberg and Parks and Carraway et al., but these models lack the college application problem's unique ``maximax'' form.
-Additionally, unlike those models, we do not attempt to replace the scalar objective function with a preference order over outcome distributions, which introduces methodological issues concerning competing notions of stochastic dominance (Sniedovich 1980).
+Additionally, unlike those models, we do not attempt to replace the scalar objective function with a preference order over \emph{outcome distributions,} which introduces technical issues concerning competing notions of stochastic dominance (Sniedovich 1980).
 We take for granted the student's preferences over \emph{outcomes} (as encoded in the $t_j$-values), and are interested instead in an efficient computational approach to the well-defined problem above.
@@
 대학 지원 문제의 정수 모형은 $m$차 다항식을 목적함수로 갖춘 일종의 이진 배낭 문제로 볼 수도 있다.
-본 논문에서 제시하는 분지한계법과 동적 계획 해법은 배낭 문제를 위한 기존 알고리즘하고 비슷하다 (Martello와 Toth 1990, \S\,2.5--6).
+본 논문에서 제시하는 분지한계법과 동적 계획 해법은 배낭 문제를 위한 기존 알고리즘과 매우 비슷하다 (Martello와 Toth 1990, \S\,2.5--6).
 입학 확률을 적당히 조정하면 원래 목적함수를 특성벡터의 선형 함수에 대한 임의로 정확한 근사로 변환할 수 있으며 NP-completeness 증명에서 이 사실을 활용한다.
-배낭 문제에 확률성을 도입한 선행 연구 중 각 상품의 효용이 정해진 확률 분포로 결정되는 모형(Steinberg와 Parks 1979; Carraway 외 1993), 그리고 배낭에 삽입한 다음에 상품의 무게가 관측되는 온라인 모형(Dean et al.
+배낭 문제에 확률성을 도입한 선행 연구 중, 각 상품의 효용이 정해진 확률 분포로 결정되는 모형(Steinberg와 Parks 1979; Carraway 외 1993), 그리고 배낭에 삽입한 다음에 상품의 무게가 관측되는 온라인 모형(Dean et al.
 2008)과 같은 바가 있다.
-본 연구가 고려하는 문제는 Steinberg와 Parks 그리고 Carraway 외가 살펴본 ``우선순위 배낭 문제''와는 피상적으로 비슷하지만, 우선순위 배낭 문제는 대학 지원 문제의 특색인 ``maximax'' 형태를 가지지는 않는다.
-또한 우선순위 모형과 달리, 본 연구에서 스칼라-값 목적함수를 `결과 분포'의 선호 순위로 대체하고자 노력하지 않는다.
-확률적 지배와 관련된 방법론적인 문제가 생기기 때문이다 (Sniedovich 1980).
-대신 $t_j$-값으로 유도된 학생의 `결과'에 대한 선호 순위를 그대로 받아들이고 위에서 정의한, 모호하지 않은 문제를 위한 효율적인 계산법에 집중한다.
+본 연구가 고려하는 문제는 Steinberg와 Parks 그리고 Carraway 외가 살펴본 ``우선순위 배낭 문제''와는 피상적으로 비슷하지만, 우선순위 배낭 문제는 대학 지원 문제의 특색인 `maximax' 형태를 가지지는 않는다.
+또한 우선순위 모형과 달리, 본 연구에서 스칼라값 목적함수를 `결과 분포'의 선호 순위로 대체하고자 노력하지 않는다.
+확률적 지배와 관련된 기술적인 문제가 생기기 때문이다 (Sniedovich 1980).
+대신 $t_j$-값으로 유도된 학생의 `결과'에 대한 선호 순위를 그대로 받아들이고 위에서 정의한, 모호하지 않은 문제를 위한 효율적인 계산법을 지향한다.
 \fi
@@
 예를 들어 예산이 다 소비될 때까지 목적함수를 가장 많이 증가시키는 학교를 차례대로 선택하는 알고리즘은 일종의 탐욕 해법이다.
-탐욕 해법은 예산 $H$으로 모수화된 해의 순서를 유도하며 해는 `포함 사슬 관계'(nestedness)로 서로에 연결된다.
+탐욕 해법은 예산 $H$으로 모수화된 해의 순서를 유도하며 그의 원소들은 `포함 사슬 관계'(nestedness)로 서로에 연결된다.
 즉 $H \leq H'$일 때, 예산 $H$에 해당하는 탐욕 해는 예산 $H'$에 해당하는 탐욕 해의 부분집합이다.
@@
 가령, 많은 미국 대학의 지원 기한은 11월 초인데 학업 계획서를 학교의 취향에 맞춰서 작성해야 하므로 여름부터 원서를 작성하는 학생이 많다.
-그러나 시간이나 지원비에 대한 예산을 10월 말까지 모를 수도 있다.
-포함 사슬 관계 성질, 또는 탐욕 해법의 타당성은 완전한 예산 정보가 없음에도 학교가 최적 포트폴리오에 진입하는 순서대로 원서를 작성하는 식으로 최적 전략을 이행해 낼 수 있다고 의미한다.
+그러나 지원 예산은 10월 말까지 모를 수도 있다.
+포함 사슬 관계, 또는 탐욕 해법의 타당성은 완전한 예산 정보가 없음에도 학교가 최적 포트폴리오에 진입하는 순서대로 원서를 작성하는 식으로 최적 전략을 이행해 낼 수 있다고 의미한다.
 \fi
@@
 \else
-한편, 탐욕 알고리즘이 좋은 근사 해법이며 다른 가정을 추가하면 정확한 알고리즘인 되는 최적화 모형 몇몇이 있으며 대표적인 바는 카디널리티 제약 위에 자모(子模) 집합 함수(submodular set function)를 최대화하는 문제이다 (Fisher et al.
+한편, 탐욕 알고리즘이 좋은 근사 해법이며 다른 가정을 추가하면 정확한 알고리즘이 되는 최적화 모형 몇몇이 알려져 있다.
+대표적인 바는 카디널리티 제약 위에 자모(子模) 집합 함수(submodular set function)를 최대화하는 문제이다 (Fisher et al.
 1978).
@@
 대학 지원 문제에 대해 배낭 문제와 매우 유사한 결과를 도출할 수 있다.
-모든 $g_j=1$인 특수한 경우에서, 최적 포트폴리오가 탐욕 알고리즘의 타당성과 동등한 포함 사슬 관계를 만족하는 것을 보인다.
+모든 $g_j=1$인 특수한 경우에서, 최적 포트폴리오가 탐욕 알고리즘의 타당성과 동등한 포함 사슬 관계를 만족하는 것을 증명한다.
 이 경우는 지원비가 없으며 정시 모집 기간에 학교 3개에만 지원할 수 있는 한국 입학 과정과 같다.
-불행하게도 일반적인 경우에서 포함 사슬 관계가 성립하지 않을뿐더러 탐욕 알고리즘에 대해 최소한 근사 계수를 보장할 수 없다.
+그러나 일반적인 경우에서 포함 사슬 관계가 성립하지 않을뿐더러 탐욕 알고리즘에 대해 최소한 어떤 근사 계수를 보장할 수 없다.
 대신 분지한계법, 전형적인 입학 시장 인스턴스에 대해 효율적인 의사 다항 시간 동적 계획, 그리고 전체 다항 시간 안에 $(1 - \varepsilon)$-근사해를 출력하는 알고리즘과 같은 3가지 해법은 제안한다.
-
 \fi
@@
 \ifen Section \ref{preliminaries} introduces some additional notation and assumptions that can be imposed with trivial loss of generality. 
-\else \ref{preliminaries}절은 표기법은 명시하고 일방성을 잃지 않는 추가의 가정을 소개한다.\fi
+\else \ref{preliminaries}절은 표기법은 명시하고 일방성을 잃지 않는 추가의 가정을 정리한다.\fi
 
@@
 \else
-\ref{homogappcosts}절에서 각 $g_j = 1$이고 $H$가 $h \leq m$인 자연수인 특수한 경우를 고려한다. 직관적인 휴리스틱 해법이 실제로 $1/h$-근사 해법임을 보인다. 그다음에 최적 포트폴리오가 예산 제약에 대한 포함 사슬 관계를 보이고 이에 따라 $O(hm)$-시간 정확한 해법을 도출한다.
+\ref{homogappcosts}절에서 각 $g_j = 1$이고 $H$가 $h \leq m$인 자연수인 특수한 경우를 고려한다. 직관적인 휴리스틱 해법이 실제로 $1/h$-근사 해법임을 보인다. 그다음에 최적 포트폴리오의 예산 제약에 대한 포함 사슬 관계를 증명하고 이를 이용해서 $O(hm)$-시간 정확한 해법을 도출한다.
 \fi
@@
 \else
-\ref{hetappcosts}절은 각 학교의 지원비가 다를 수 있는 상황으로 계속한다.
+\ref{hetappcosts}절에서 각 학교의 지원비가 다를 수 있는 상황을 다룬다.
 이진 배낭 문제에서의 다항 변환을 통해서 포트폴리오 최적화 문제가 NP-complete임을 증명한다.
@@
 첫째는 분지한계법이다.
-둘째는 총 지출액으로 탐색해서 $O(Hm + m \log m)$과 같은 의사 다항 시간 동적 계획이다.
-셋째는 분수를 내린 포트폴리오 가치로 탐색하는 다른 동적 계획이다.
+둘째는 총 지출액으로 탐색하는 $O(Hm + m \log m)$과 같은 의사 다항 시간 동적 계획이다.
+셋째는 분수값은 내린 포트폴리오 가치로 탐색하는 또다른 동적 계획이다.
 이를 통해서 $O(m^3 / \varepsilon)$-시간 안에 $(1 - \varepsilon)$-근사해를 출력하는 전체 다항 시간 근사 해법(fully polynomial-time approximation scheme)을 얻는다.
@@
 \else
-\ref{numericalexperiments}절에서 위에서 도출한 타당성과 계산 복잡성을 확인하는 계산적 실험 결과를 정리한다.
+\ref{numericalexperiments}절에서 위에서 도출한 타당성과 계산 복잡성을 확인하는 계산 실험 결과를 정리한다.
 
@@
 \ifen \section{Notation and preliminary results} \else \section{표기법과 예비 결과}\fi\label{preliminaries}
-Before discussing the solution algorithms, we will introduce some additional notation and a few preliminary results that will come in handy.
+\ifen Before discussing the solution algorithms, we will introduce some additional notation and a few preliminary results that will come in handy.
 
@@
 Below, we will show how to transform an arbitrary instance so that $t_0 = 0$, in which case each $t_j > 0$.
-
+\else
+해법은 의논하기 전, 본 절에서 추가적인 표기법은 명시하고 몇 가지 유용한 예비 결과를 정리하고자 한다.
+
+남은 논문에서 다른 언급이 없으면, 일반성을 잃지 않고 각 $g_j \leq H$, 각 $f_j \in (0, 1]$, 그리고 $t_0 < t_1 \leq \cdots \leq t_m$이라고 가정한다. 뒤에서 임의의 인스턴스를 $t_0 = 0$이 되도록 변환하는 방법을 제시하는데 이를 적용하면 $t_j > 0$인 가정을 세울 수 있다. 
+\fi
+
+\ifen
 We refer to the set $\mathcal{X} \subseteq \mathcal{C}$ of schools to which a student applies as her \emph{application portfolio.} The expected utility the student receives from $\mathcal{X}$ is called its \emph{valuation}.
 %Let $x$ denote the same encoded as a binary vector, where $x_j=1 \iff j\in \mathcal{X}$ for $j=1\dots m$.
-\begin{definition}[Portfolio valuation function]
+\else
+학생이 지원하는 학교 집합 $\mathcal{X} \subseteq \mathcal{C}$를 `지원 포트폴리오'라고 부르자. $\mathcal{X}$에서 학생이 받는 기대 효용을 포트폴리오의 `가치'라고 한다.
+\fi 
+\begin{definition}[\ifen Portfolio valuation function\else 포트폴리오 가치 함수\fi]
 $v(\mathcal{X}) =  \operatorname{E}\left[\max\bigr\{t_0,
@@
 \end{definition}
+\ifen
 It is helpful to define the random variable $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$ as the utility achieved by the schools in the portfolio, so that when $t_0 = 0$, $v(\mathcal{X}) = \operatorname{E}[X]$.
-Similar pairs of variables with italic and script names such as $\mathcal{Y}_h$ and $Y_h$ carry an analogous meaning.
+Similar pairs of variables with script and italic names such as $\mathcal{Y}_h$ and $Y_h$ carry an analogous meaning.
+\else
+편의상 포트폴리오에 속한 학교가 실제로 이루는 효용을 확률 변수 $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$로 정의한다.
+그러면 $t_0 = 0$이면 $v(\mathcal{X}) = \operatorname{E}[X]$임을 알 수 있다.
+비슷한 식으로 $\mathcal{Y}_h$와 $Y_h$ 같은 스크립트체와 이탈릭체로 표기한 변수 쌍은 유사한 의미를 가진다.
+\fi
 
@@
 
-\begin{proposition}[Closed form of portfolio valuation function]
+\begin{proposition}[\ifen Closed form of portfolio valuation function\else 포트폴리오 가치 함수의 다른 식\fi]
 \begin{align}
@@
 \begin{align}
+\begin{split}
 v(\mathcal{X}; \bar t_j) &= \sum_{j\in \{0\}\cup\mathcal{X}}  \bar t_j p_j(\mathcal{X})
@@
 = v(\mathcal{X}; t_j) - \gamma
+\end{split}
 \end{align}
@@
 \end{equation*}
-Thus, as $\epsilon$ approaches zero, we have $v(\mathcal{T}_h) / v(\mathcal{X}_h) \to 1/h$.
+Thus, as $\varepsilon$ approaches zero, we have $v(\mathcal{T}_h) / v(\mathcal{X}_h) \to 1/h$.
 (The optimality of $\mathcal{X}_h$ follows from the fact that it achieves the upper bound of Theorem \ref{oneoverhopt}.)
@@
 \ifen \subsection{The nestedness property}  \else \subsection{포함 사슬 관계} \fi
-It turns out that the solution to Alma's problem possesses a special structure: An optimal portfolio of size $h+1$ includes an optimal portfolio of size $h$ as a subset.
+\ifen It turns out that the solution to Alma's problem possesses a special structure: An optimal portfolio of size $h+1$ includes an optimal portfolio of size $h$ as a subset.
+\else 알마의 문제의 최적해에는 특별한 구조가 있다. 크기 $h+1$에 대한 최적 포트폴리오는 항상 크기 $h$에 대한 최적 포트폴리오를 부분집합으로 포함한다.\fi
 % Incl above
@@
 
-\begin{theorem}[Nestedness of optimal application portfolios] \label{nestedapplication}
-There exists a sequence of portfolios $\{\mathcal{X}_h\}_{h=1}^m$ satisfying the nestedness relation 
+\begin{theorem}[\ifen Nestedness of optimal application portfolios\else 최적 포트폴리오의 포함 사슬 관계\fi] \label{nestedapplication}
+\ifen There exists a sequence of portfolios $\{\mathcal{X}_h\}_{h=1}^m$ satisfying the nestedness relation \fi
 \begin{equation}
@@
 \end{equation}
-such that each $\mathcal{X}_h$ is an optimal application portfolio when the application limit is $h$.
+\ifen such that each $\mathcal{X}_h$ is an optimal application portfolio when the application limit is $h$.
+\else  각 $\mathcal{X}_h$가 지원 제한 $h$에 대한 최적 포트폴리오며 위의 포함 사슬 관계를 만족하는 포트톨리오 수열  $\{\mathcal{X}_h\}_{h=1}^m$가 존재한다.\fi
 \end{theorem}
-\begin{proof} By induction on $h$. Applying Theorem \ref{assumetzerozero}, we assume that $t_0 = 0$.
+
+\begin{proof}\ifen
+By induction on $h$. Applying Theorem \ref{assumetzerozero}, we assume that $t_0 = 0$.
 
@@
 Optimality requires that
+\else$h$에 대해 귀납법을 적용한다. 정리 \ref{assumetzerozero}에 따라 $t_0 = 0$이라고 가정한다.
+
+(기본 경우.) 우선 $\mathcal{X}_1 \subset \mathcal{X}_2$임을 증명하자. 모순을 보이기 위해 최적해가 $\mathcal{X}_1 = \{j\}$와 $\mathcal{X}_2 = \{k, l\}$이라고 하자. 여기서 $t_k \leq t_l$이라고 가정할 수 있다. 최적성에 따라 다음이 성립한다:
+\fi
 \begin{equation}v(\mathcal{X}_1 )  = f_j t_j > v(\{k\}) = f_k t_k\end{equation}
-and
+\ifen and \else 그리고\fi
 \begin{align}
+\begin{split}
 v(\mathcal{X}_2) =  f_k (1- f_l) t_k + f_l t_l &> v(\{j, l\}) \\
@@
 &\geq f_k (1- f_l) t_k + f_l t_l  = v(\mathcal{X}_2)
-\end{align}
-which is a contradiction.
-
-(Inductive step.) Assume that $\mathcal{X}_1 \subset \cdots \subset \mathcal{X}_h$, and we will show $\mathcal{X}_h \subset \mathcal{X}_{h+1}$.
+\end{split}
+\end{align}
+\ifen which is a contradiction. 
+\else 이는 모순이다.\fi
+
+\ifen (Inductive step.) Assume that $\mathcal{X}_1 \subset \cdots \subset \mathcal{X}_h$, and we will show $\mathcal{X}_h \subset \mathcal{X}_{h+1}$.
 Let $k = \argmax\{ t_k: k \in \mathcal{X}_{h+1}\}$ and write $\mathcal{X}_{h+1} = \mathcal{Y}_{h} \cup \{k\}$.
-
-Suppose $k \notin \mathcal{X}_h$.  To get a contradiction, assume that $v(\mathcal{Y}_h) < v(\mathcal{X}_h)$. Then
+\else (추론.) $\mathcal{X}_1 \subset \cdots \subset \mathcal{X}_h$라고 가정하고 $\mathcal{X}_h \subset \mathcal{X}_{h+1}$임을 보이자.
+$k = \argmax\{ t_k: k \in \mathcal{X}_{h+1}\}$으로 정의해서 $\mathcal{X}_{h+1} = \mathcal{Y}_{h} \cup \{k\}$으로 표현하자.\fi
+
+\ifen Suppose $k \notin \mathcal{X}_h$. To get a contradiction, assume that $v(\mathcal{Y}_h) < v(\mathcal{X}_h)$. Then
+\else $k \notin \mathcal{X}_h$인 경우를 고려하자. 모순을 보이기 위해 $v(\mathcal{Y}_h) < v(\mathcal{X}_h)$라고 하면 \fi
 \begin{align}
@@
 \end{align}
-contradicts the optimality of $\mathcal{X}_{h+1}$.
-
-Now suppose that $k \in \mathcal{X}_h$.
+\ifen contradicts the optimality of $\mathcal{X}_{h+1}$.
+\else 이는 $\mathcal{X}_{h+1}$의 최적성을 위반한다.\fi
+
+\ifen Now suppose that $k \in \mathcal{X}_h$.
 We can write $\mathcal{X}_h = \mathcal{Y}_{h-1} \cup \{k\}$, where $ \mathcal{Y}_{h-1}$ is some portfolio of size $h-1$.
@@
 That is, $\mathcal{Y}_{h-1}$ and $\mathcal{Y}_h$ are the optimal \emph{complements} to the singleton portfolio $\{k\}$.
-
-We will use the function $w(\mathcal{Y})$ to grade portfolios $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$ according to how well they complement $\{k\}$.
+\else 이제 $k \in \mathcal{X}_h$인 경우를 고려하자.
+그러면 크기 $h-1$인 어떤 포트폴리오 $\mathcal{Y}_{h-1}$에 대해 $\mathcal{X}_h = \mathcal{Y}_{h-1} \cup \{k\}$으로 표현할 수 있다.
+$\mathcal{Y}_{h-1} \subset \mathcal{Y}_h$임을 보이면 충분하다.
+정의에 따라 $\mathcal{Y}_{h-1}$ ($\mathcal{Y}_{h}$)는 크기가 $h-1$ ($h$)인 포트폴리오 중에 함수 $v(\mathcal{Y}\cup\{k\})$를 최대화한다.
+다시 말해 $\mathcal{Y}_{h-1}$과 $\mathcal{Y}_h$는 각각 단일 원소 포트폴리오 $\{k\}$와 최적으로 `보완적인' 학교 집합이다.
+\fi
+
+\ifen We will use the function $w(\mathcal{Y})$ to grade portfolios $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$ according to how well they complement $\{k\}$.
 To construct $w(\mathcal{Y})$, let $\tilde t_j$ denote the expected utility Alma receives from school $c_j$ \emph{given} that she has been admitted to $c_j$ and applied to $c_k$.
@@
 This means that
+\else 함수 $w(\mathcal{Y})$를 이용해서 $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$인 포트폴리오와 $\{k\}$ 사이의 \mbox{`보완성'을} 평가하자.
+$w(\mathcal{Y})$를 구성하기 위해, 알마가 $c_j$에 합격한 상황에서 $c_k$에 지원했을 때 $c_j$에서 받는 조건부 기대 효용을 $\tilde t_j$라고 하자.
+$j = 0$을 포함한 $j < k$에 대해 이값은 $\tilde t_j = t_j (1- f_k) + t_k f_k$이며, $j > k $에 대해 $\tilde t_j = t_j$임을 알 수 있다.
+따라서 \fi
 \begin{equation}\label{Vyastildet}
 v(\mathcal{Y}\cup\{k\}) = \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y}).\end{equation}
-The transformation to $\tilde t$ does not change the order of the $t_j$-values.
+\ifen The transformation to $\tilde t$ does not change the order of the $t_j$-values.
 Therefore, the expression on the right side of \eqref{Vyastildet} is itself a portfolio valuation function.
@@
 Now, we obtain $w(\mathcal{Y})$ through one more transformation: Define $\bar t_j = \tilde t_j - \tilde t_0$ so that $t_0 = 0$ and let
+\else $\tilde t$로 변환하면 $t_j$의 순서가 변하지 않기 때문에 \eqref{Vyastildet}의 우변 그 자체가 포트폴리오 가치 함수이다.
+$t$를 $\tilde t$로, $\mathcal{C}$를 $\mathcal{C}\setminus \{k\}$로 대체하면 대응하는 시장이 된다.
+이제 또 하나의 변환을 적용한다.
+$t_0 = 0$이 되도록 $\bar t_j = \tilde t_j - \tilde t_0$으로 정의하고 $w(\mathcal{Y})$를 다음과 같이 얻을 수 있다.
+\fi
 \begin{equation}  \label{wYvXminusconst}
@@
 = v(\mathcal{Y}\cup\{k\}) -  f_k t_k \end{equation}
-where the second equality follows from Theorem \ref{assumetzerozero}.
+\ifen where the second equality follows from Theorem \ref{assumetzerozero}.
 This identity says that the optimal complements to $\{k\}$, given by $\mathcal{Y}_{h-1}$ and $\mathcal{Y}_h$, are themselves optimal portfolios of size $h-1$ and $h$ for the market whose objective function is $w(\mathcal{Y})$.
-Since $\bar t_0 = 0$ in the latter market, the inductive hypothesis implies that $\mathcal{Y}_{h-1} \subset \mathcal{Y}_h$, which completes the proof.\footnote{We thank Yim Seho for discovering this critical transformation.}
+Since $\bar t_0 = 0$ in the latter market, the inductive hypothesis implies that $\mathcal{Y}_{h-1} \subset \mathcal{Y}_h$, which completes the proof.\footnote{We thank Yim Seho for discovering this useful transformation.}
+\else 두번째 등호는 정리 \ref{assumetzerozero}에 따라 성립한다.
+이 등호는, $\{k\}$와 최적으로 보완적인 $\mathcal{Y}_{h-1}$과 $\mathcal{Y}_h$는 목적함수가 $w(\mathcal{Y})$인 시장에서 각각 크기가 $h-1$과 $h$인 최적 포트폴리오임을 의미한다.
+새로운 시장에서 $\bar t_0 = 0$이므로 귀납법 가설을 적용하면 $\mathcal{Y}_{h-1} \subset \mathcal{Y}_h$임을 알 수 있다.
+따라서 증명이 완성된다.\footnote{유용한 변환을 발견한 임세호에게 감사를 표한다.} \fi
 \end{proof}
@@
 \ifen \subsection{Polynomial-time solution} \else \subsection{다항 시간 해법}\fi
-Applying the result above yields an efficient algorithm for the optimal portfolio: Start with the empty set and add schools one at a time, maximizing $v(\mathcal{X}\cup \{k\})$ at each addition.
+\ifen Applying the result above yields an efficient greedy algorithm for the optimal portfolio: Start with the empty set and add schools one at a time, maximizing $v(\mathcal{X}\cup \{k\})$ at each addition.
 Sorting $t$ is  $O(m \log m)$.
 At each of the $h$ iterations, there are $O(m)$ candidates for $k$, and computing $v(\mathcal{X}\cup \{k\})$ is $O(h)$ using \eqref{closedformportfoliovaluationX}; therefore, the time complexity of this algorithm is $O(h^2 m + m \log m)$.
+\else 위의 결과를 적용하면 효율적인 최적 포트폴리오 탐욕 해법을 얻는다: 공집합으로 시작하고 $v(\mathcal{X}\cup \{k\})$를 최대화하는 학교를 차례대로 추가한다.
+$t$를 배열하는 시간을 $O(m \log m)$이다.
+모든 $h$개의 반복단계에서 $k$의 후보자의 개수는 $O(m)$이며 \eqref{closedformportfoliovaluationX}을(를) 사용하면 $v(\mathcal{X}\cup \{k\})$를 계산하는 시간은 $O(h)$이다.
+따라서 이 해법의 계산 시간이 $O(h^2 m + m \log m)$임을 알 수 있다.\fi
 %\begin{algorithm}[H] 
@@
 %\end{proof}
-
+\ifen 
 We reduce the computation time to $O(hm)$ by taking advantage of the transformation from the inductive step in the proof of Theorem \ref{nestedapplication}.
 Once school $k$ is added to $\mathcal{X}$, we remove it from the set $\mathcal{C}\setminus \mathcal{X}$ of candidates, and update the $t_j$-values of the remaining schools according to the following transformation:
+\else
+정리 \ref{nestedapplication} 증명의 추론 단계에서 제시한 변환을 활용하면 계산 시간을 $O(hm)$으로 감소시킬 수 있다. $\mathcal{X}$에 $k$를 추가하면, 후보자 집합인 $\mathcal{C}\setminus \mathcal{X}$에서 $k$를 제거하고 남은 학교의 $t_j$-값은 다음처럼 수정한다.
+\fi
 \begin{align}\label{howtotransformfj}
@@
 \end{align}
+\ifen
 It is easy to verify that this is the composition of the two transformations (from $t$ to $\tilde t$, and from $\tilde t$ to $\bar t$) given in the proof.
@@
 Moreover, this algorithm does not require the schools to be indexed in ascending order by $t_j$, which removes the $O(m\log m)$ sorting cost.
-
+\else
+이것이 증명에서 정의한 두개의 변환($t$에서 $\tilde t$로, 그리고 $\tilde t$에서 $\bar t$로)의 합성임을 쉽게 확인할 수 있다.
+그러면 다음으로 추가하는 학교는 수정된 시장의 최적 단일 원소 포트폴리오가 되어야 한다.
+그런데 최적 단일 원소 포트폴리오는 단순히 $f_j \bar t_j$가 가장 높은 학교로 이루어진다.
+따라서 각 반복단계에서 \eqref{howtotransformfj}을(를) 이용해서 $t_j$-값을 수정하면 $v(\mathcal{X})$를 계산할 필요가 완전히 없어진다.
+게다가 이 알고리즘에서 $t_j$를 배열할 필요가 없으므로 $O(m\log m)$인 배열 비용이 사라진다.
+\fi
+
+ \ifen 
 The algorithm below outputs a list $\mathtt{X}$ of the $h$ schools to which Alma should apply.
@@
 The entries of the list $\mathtt{V}$ give the valuation thereof.
-
-\begin{algorithm}[H] 
+\else
+아래 알고리즘은 알마가 지원하는 $h$개의 학교를 목록 $\mathtt{X}$로 출력한다.
+그의 원소들은 진입하는 순서대로 등장한다.
+즉, $h=m$으로 알고리즘을 돌리면 크기가 $h$인 최적 포트폴리오는 $\mathcal{X}_h = \{\mathtt{X[1]}, \dots, \mathtt{X[h]}\}$와 같다.
+또한 목록 $\mathtt{V}$의 원소들은 해당 포트폴리오의 가치다.
+\fi
+
+\ifen {
+\begin{algorithm}[h] 
 %\DontPrintSemicolon
 \caption{Optimal portfolio algorithm for Alma’s problem.} \label{algorithmforlargeh}
-\KwData{Utility values $t \in[0, \infty)^m$, admissions probabilities $f \in [0, 1]^m$, application limit $h \leq m$.}
+\KwData{Utility values $t \in[0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application limit $h \leq m$.}
 %Index schools in ascending order by $t$\;
@@
 $\mathtt{X, V} \gets $ empty lists\;
-%$\mathtt{L} \gets $, an empty heap of named tuples $\mathtt{C}$ ordered by $\mathtt{C.f}\cdot\mathtt{C.t}$\;
-%\lFor{$j=1\dots m$}
-%{$\operatorname{append!}\bigl(\mathtt{L}, (j, f_j, t_j)\bigr)$}
 \For{$i=1\dots h$}
@@
 	{
-	\lIfElse{$t_j \leq t_k$}{$t_j \gets  t_j (1 -  f_k)$}{$t_j \gets  t_j -  f_k t_k$}
+	\lIfElse{$t_j \leq t_k$}{$t_j \gets (1 -  f_k) t_j $}{$t_j \gets  t_j -  f_k t_k$}
 	}
-%    \For{$\mathtt{E} \in \mathtt{H}: 1\leq \mathtt{E.j} < \mathtt{D.j}$}
-%	{
-%	Update $\mathtt{E.t} \gets  \mathtt{E.t} \cdot(1 -  \mathtt{D.f})$\;
-%    }
-%    \For{$\mathtt{E} \in \mathtt{H}:  \mathtt{D.j} <  \mathtt{E.j} \leq m$}
-%	{
-%	Update $\mathtt{E.t} \gets  \mathtt{E.t} -  \mathtt{D.f}\cdot\mathtt{D.t} $\;
-%    }
+}
+\Return{$\mathtt{X, V}$}
+} \else {
+\begin{algorithm}[H] 
+%\DontPrintSemicolon
+\caption{알마의 문제를 위한 최적 포트폴리오 알고리즘.} \label{algorithmforlargeh}
+\KwData{효용 모수 $t \in[0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 제한 $h \leq m$.}
+%$t$의 순서대로 학교의 지표를 짓는다\;
+$\mathcal{C} \gets \{1 \dots m\}$\;
+$\mathtt{X, V} \gets $ 빈 목록\;
+\For{$i=1\dots h$}
+{
+    $k \gets \argmax_{j \in \mathcal{C}}\{f_j t_j\}$\;
+    $\mathcal{C} \gets \mathcal{C} \setminus \{k\}$\;
+    $\operatorname{append!}(\mathtt{X}, k)$\;
+     \lIfElse{$i=1$}{$\operatorname{append!}(\mathtt{V}, f_k t_k)$}
+     {$\operatorname{append!}(\mathtt{V}, \mathtt{V[i-1]} + f_k t_k)$}
+    \For{$j \in \mathcal{C}$}
+	{
+	\lIfElse{$t_j \leq t_k$}{$t_j \gets (1 -  f_k) t_j $}{$t_j \gets  t_j -  f_k t_k$}
+	}
 }
@@
 
-\begin{theorem}[Validity of Algorithm \ref{algorithmforlargeh}]
-Algorithm \ref{algorithmforlargeh} produces an optimal application portfolio for Alma's problem in $O(h m)$ time.
+\begin{theorem}[\ifen Validity of Algorithm \ref{algorithmforlargeh}\else 알고리즘 \ref{algorithmforlargeh}의 타당성\fi] \label{validityofalmaalgorithm}
+\ifen Algorithm \ref{algorithmforlargeh} produces an optimal application portfolio for Alma's problem in $O(h m)$-time.
+\else 알고리즘 \ref{algorithmforlargeh}은(는) $O(h m)$-시간 안에 알마의 문제를 위한 최적 지원 포트폴리오를 출력한다.\fi
 \end{theorem}
 \begin{proof}
+\ifen
 Optimality follows from the proof of Theorem \ref{nestedapplication}.
@@
 Therefore, the overall time complexity is $O(h m)$.
+\else
+최적성은 정리 \ref{nestedapplication}의 증명에 따라 성립한다. $\mathcal{C}$를 목록 구현으로 저장한다고 하자. 그러면 각 $h$개의 반복단계에서 최적 추가 학교를 구하는 시간은 $O(m)$이며 나머지 $O(m)$개 학교의 $t_j$-값을 각 단위 시간으로 수정할 수 있다. 따라서 전체 시간 복잡도가 $O(h m)$이다.
+\fi
 \end{proof}
 
-In our numerical experiments,
-%whose results are reported in section \ref{numericalexperiments},
-we found it effective to store $\mathcal{C}$ as a binary max heap rather than a list.
+\ifen 
+In our numerical experiments, whose results are reported in section \ref{numericalexperiments},
+we also explore the possibility of storing $\mathcal{C}$ as a binary max heap rather than a list.
 The heap is ordered according to the criterion $i \geq j \iff f_i t_i \geq f_j t_j$.
@@
 The cost of updating each $t_j$-value can be reduced to unit time using a Fibonacci heap (Fredman and Tarjan 1987), yielding the same overall computation time.
+\else
+(\ref{numericalexperiments}절에서 결과를 제시할) 수리 실험에서 $\mathcal{C}$를 목록 대신 이진 최대 힙 구현으로 저장하는 방법도 고려한다.
+단, 힙의 순서는 $i \geq j \iff f_i t_i \geq f_j t_j$의 조건으로 정의한다.
+기술적으로, $t_j$-값을 수정할 때마다 힙의 구조를 다시 복구해야 하므로 이는 주 반복단계의 계산 비용을 $O(h m)$에서 $O(hm \log m)$으로 증가시킨다.
+그러나 전형적인 시장에서는, 반복단계 사이의 $f_j t_j$-값의 순서가 크게 변하지 않으므로 이 상한은 자주 발생하지 않는다.
+또한 피보나치 힙을 사용하면 $t_j$-값을 수정하는 시간을 단위 시간으로 줄여서 동등한 계산 시간을 얻어 낼 수 있다 (Fredman and Tarjan 1987).
+\fi
 
 \ifen \subsection{Properties of the optimal portfolios} \else\subsection{최적 포트폴리오의 성질} \fi
-The nestedness property implies that Alma's expected utility is a discretely concave function of $h$.
-
-\begin{theorem}[Optimal portfolio valuation concave in $h$] \label{concavityinh}
-For $h = 2 \dots (m-1)$, \begin{equation}v(\mathcal{X}_h) - v(\mathcal{X}_{h-1}) \geq v(\mathcal{X}_{h+1}) - v(\mathcal{X}_{h}).\end{equation} 
+\ifen The nestedness property implies that Alma's expected utility is a discretely concave function of $h$.
+\else 포함 사슬 관계 성질은 알마의 기대 효용이 $h$의 이산 오목 함수임을 의미한다.\fi
+
+\begin{theorem}[\ifen Optimal portfolio valuation concave in $h$\else 최적 포트폴리오 가치의 $h$-오목성\fi] \label{concavityinh}
+\ifen For $h = 2 \dots (m-1)$,
+\else $h = 2 \dots (m-1)$에 대해,\fi
+\begin{equation}v(\mathcal{X}_h) - v(\mathcal{X}_{h-1}) \geq v(\mathcal{X}_{h+1}) - v(\mathcal{X}_{h}).\end{equation} 
 \end{theorem}
 \begin{proof}
-We will prove the equivalent expression $2 v(\mathcal{X}_h) \geq v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1})$.
+\ifen We will prove the equivalent expression $2 v(\mathcal{X}_h) \geq v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1})$.
 Applying Theorem \ref{nestedapplication}, we write $\mathcal{X}_h = \mathcal{X}_{h-1} \cup\{j\}$ and $\mathcal{X}_{h+1} = \mathcal{X}_{h-1} \cup\{j, k\}$.
-Define the random variables $X_i$ as above.
 If $t_k \leq t_j$, then
+\else $2 v(\mathcal{X}_h) \geq v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1})$ 같은 동등한 부등식을 증명하자.
+정리 \ref{nestedapplication}을(를) 적용하면 $ \mathcal{X}_h = \mathcal{X}_{h-1} \cup\{j\}$ 그리고 $\mathcal{X}_{h+1} = \mathcal{X}_{h-1} \cup\{j, k\}$으로 표현할 수 있다.
+$t_k \leq t_j$인 경우, \fi
 \begin{align}
@@
 \end{align}
-The first inequality follows from the optimality of $\mathcal{X}_h$, while the second follows from the fact that adding $k$ to $\mathcal{X}_{h-1}$ can only increase its valuation.
-
-If $t_k \geq t_j$, then the steps are analogous:
+\ifen The first inequality follows from the optimality of $\mathcal{X}_h$, while the second follows from the fact that adding $k$ to $\mathcal{X}_{h-1}$ can only increase its valuation.
+\else 첫번째 부등식은 $\mathcal{X}_h$의 최적성에 따르며, 두번째 부등식은 $\mathcal{X}_{h-1}$에 $j$를 더하면 가치가 증가할 수 밖에 없기 때문이다.\fi
+
+\ifen If $t_k \geq t_j$, then the steps are analogous:
+\else  $t_k \geq t_j$인 경우는 유사하다:\fi
 \begin{align}
@@
 
-It follows that when $\mathcal{X}_h$ is the optimal $h$-portfolio, for a given market, $v(\mathcal{X}_h)$ is $O(h)$.
+\ifen
+It follows that when $\mathcal{X}_h$ is the optimal $h$-portfolio for a given market, $v(\mathcal{X}_h)$ is $O(h)$.
 Example \ref{tightexampleforoneoverhopt}, in which $v(\mathcal{X}_h)$ can be made arbitrarily close to $h$, establishes the tightness of this bound.
+\else
+위 결과는 $\mathcal{X}_h$가 어떤 시장의 최적 $h$-포트폴리오일 때, $v(\mathcal{X}_h)$가 $O(h)$ 함수라고 의미한다. 예\ref{tightexampleforoneoverhopt}에서 $v(\mathcal{X}_h)$를 $h$에 임의로 가깝게 조정할 수 있으므로 이 상한이 타이트함을 알 수 있다.
+\fi
 
@@
 \ifen \subsection{A small example} \else \subsection{작은 예} \fi
-Let us consider a fictional market consisting of $m=8$ schools. The market data, along with the optimal solutions for each $h \leq m$, appear in Table \ref{planetsdata}.
-
+\ifen 
+Let us consider a fictional admissions market consisting of $m=8$ schools. The school data, along with the optimal solutions for each $h \leq m$, appear in Table \ref{planetsdata}.
+\else 
+$m = 8$개의 학교로 구성된 가상 입학 시장을 고려하자. 학교 자료와 각 $h \leq m$에 대응하는 최적해는 표 \ref{planetsdata}에서 나타난다.
+\fi
+
+\ifen
 Below are shown the first several iterations of Algorithm \ref{algorithmforlargeh}.
@@
 The school added at each iteration, underlined, is the one whose $f_j t_j$-value is greatest.
+\else
+아래에서 알고리즘 \ref{algorithmforlargeh}의 몇 반복단계가 나타난다.
+$f_j$, $t_j$, 그리고 그 곱의 값은 $\mathcal{C}$에 남아 있는 학교에 해당하는 값만 기록한다.
+$(f * t)_j = f_j t_j$로 정의된 $f * t$는 $f$와 $t$의 원소 당 곱을 의미한다.
+각 반복단계에서 최적해에 추가하는 학교는 $f_j t_j$-값이 가장 높은 학교이며 이를 밑줄로 강조한다.
+\fi
 \begin{align*}
@@
 \end{align*}
+\ifen 
 The output of the algorithm is $\mathtt{X} = [4, 2, 8, 1, 7, 3, 5, 6]$, and the optimal $h$-portfolio consists of its first $h$ entries.
@@
 The concave shape of the plot suggests the result of Theorem \ref{concavityinh}.
-
-\begin{table} \centering
-\begin{tabular}{clcccc}
+\else
+알고리즘은 $\mathtt{X} = [4, 2, 8, 1, 7, 3, 5, 6]$를 출력하며 최적 $h$-포트폴리오는 그의 첫 $h$개의 원소로 이루어진다.
+표 \ref{planetsdata}의 ``지원 순위'' 열은 $\mathtt{X}$의 역순열(inverse permutation)이며 이는 해당 학교가 최적 포트폴리오에 포함되는 최소 $h$-값을 의미한다.
+그림 \ref{planetsplot}은 최적 포트폴리오의 가치를 $h$의 함수로 나타낸다.
+곡선의 오목성은 정리 \ref{concavityinh}의 결과를 시사한다.
+\fi
+
+\begin{table}[h!] \centering
+\begin{tabular}{r|lcccc}
 \ifen\textbf{Index $j$} & \textbf{School $c_j$} & \textbf{Utility $t_j$} & \textbf{Admit prob. $f_j$} & \textbf{Priority} & \textbf{$v(\mathcal{X}_h)$} \\ \hline
@@
 \end{tabular}
-\caption{Market data and optimal application portfolios for a fictional market with $m=8$ schools.
+\caption{\label{planetsdata}
+\ifen College data and optimal application portfolios for a fictional market with $m=8$ schools.
 By the nestedness property (Theorem \ref{nestedapplication}), the optimal portfolio when the application limit is $h$ consists of the $h$ schools having priority number $h$ or less.
-\label{planetsdata}}
+\else
+$m=8$개의 학교로 이루어진 가상 입학 시장의 대학교 자료와 최적 지원 포트폴리오. 포함 사슬 관계 성질(정리 \ref{nestedapplication})에 따라, 지원 제한이 $h$일 때 최적 포트폴리오는 지원 순위가 $h$ 이하인 $h$개의 학교로 구성된다.
+\fi}
 \end{table}
@@
 
-\begin{algorithm}[H] 
+\begin{algorithm}[h] 
 %\DontPrintSemicolon
@@
 This means that significant fathoming operations do not occur until the algorithm has explored deep into the tree, at which point the bulk of the computational effort has already been exhausted.
-In our numerical experiments, Algorithm \ref{ellisbnb} was ineffective on instances larger than about $m=30$ schools.
+In our numerical experiments, Algorithm \ref{ellisbnb} was ineffective on instances larger than about $m=20$ schools, and therefore it does not represent a significant improvement over na\"ive enumeration.
 
@@
 
-\ifen \subsection{Fully polynomial-time approximation scheme}\else \subsection{전체 다항 시간 근사 해법}\fi
+\ifen \subsection{Fully polynomial-time approximation scheme}\else \subsection{전체 다항 시간 근사 해법}\fi \label{fptashead}
 As with the knapsack problem, Ellis's problem admits a complementary dynamic program that iterates on the value of the cheapest portfolio instead of on the cost of the most valuable portfolio.
@@
 \pagebreak
-\ifen \section{(WIP) Numerical experiments} \else \section{계산적 실험}\fi\label{numericalexperiments}
-In this section, we present the results of numerical experiments designed to confirm the time complexity results established above.
-In both experiments, markets were generated by drawing $t_j$ independently from an exponential distribution with a scale parameter of ten and rounding up to the nearest integer.
-To achieve partial negative correlation between $t_j$ and $f_j$, we then set $f_j = 1 / (t_j + 10 Q)$, where $Q$ is drawn uniformly from the interval $[0, 1)$.
+\ifen \section{Numerical experiments} \else \section{계산적 실험}\fi\label{numericalexperiments}
+In this section, we present the results of numerical experiments designed to test the time complexity results established above.
+
+\ifen \subsection{Experimental procedure} \else \subsection{실험 방법}\fi
+\def\nmarkets{20}
+The experimental procedure was as follows.
+First, to generate synthetic markets, we drew the $t_j$-values independently from an exponential distribution with a scale parameter of ten and rounding up to the nearest integer.
+To achieve partial negative correlation between $t_j$ and $f_j$, we then set $f_j = 1 / (t_j + 10\,Q)$, where $Q$ is drawn uniformly from the interval $[0, 1)$.
+In the first experiment, which concerns Alma's problem, we set each $g_j = 1$ and $H = h = \lfloor m/ 2 \rfloor$.
+In the second experiment, which concerns Ellis's problem, each $g_j$ is drawn uniformly from the set $\{5, \dots, 10\}$ and we set $H = \lfloor \frac{1}{2} \sum g_j \rfloor$.
 A typical instance is shown in Figure \ref{samplemarket}.
 
-In Experiment 1, which concerns Algorithm \ref{algorithmforlargeh}, we take each $g_j = 1$ and set $H = h = \lfloor m/ 2 \rfloor$.
-In Experiment 2, which concerns Algorithms \ref{ellisbnb}, \ref{ellisDP1}, and \ref{ellisDP3}, each $g_j$ is drawn uniformly from the set $\{5, \dots, 10\}$ and we take $H = \lfloor \frac{1}{2} \sum g_j \rfloor$.
-
-At each combination of the experimental variables, we generated ten markets, and each computation was repeated three times, with the fastest of the three  recorded as the computation time.
-Therefore, each cell of each table represents 30 computations.
-We report the mean and standard deviation across the 50 markets.
+For each combination of the experimental variables, we generated \nmarkets~markets, and the optimal portfolio was computed three times, with the fastest of the three repetitions recorded as the computation time.
+We then report the mean and standard deviation across the \nmarkets~markets.
+Therefore, each cell of each table below represents a statistic over {\the\numexpr 3 * \nmarkets \relax}~computations.
 Where applicable, we do not count the time required to sort the entries of $t$.
+
+\ifen \subsection{Summary of results} \else \subsection{결과 정리}\fi
+In our first experiment, we compare the performance of Algorithm \ref{algorithmforlargeh} for homogeneous-cost markets of various sizes when the set of candidate schools $\mathcal{C}$ is stored as a list and as a binary max heap ordered by the $f_j \bar t_j$-values.
+The results appear in Table \ref{experiment1results}.
+Although using a max heap increases the nominal runtime from $O(hm)$ to $O(hm \log m)$, we hypothesized that in typical instances, the order of the heap would change only slightly between iterations, yielding an overall savings because of the lower costs of extracting the maximal school.
+Our results indicate that due to its lower overhead cost, the list implementation is faster; however, the ratio between the average times of the two implementations stays roughly constant at about 1.5, suggesting that a more effective heap implementation could be competitive in certain classes of problem instances.
+Overall, the rate of growth is quadratic in $m$, which accords with the $O(hm)$ time complexity result of Theorem \ref{validityofalmaalgorithm}.
+
+In the second experiment, we turn to the general problem, and compare the performance of the exact algorithms (Algorithms \ref{ellisbnb} and \ref{ellisDP1}) and the approximation scheme (Algorithm \ref{ellisDP3}) at tolerances 0.5 and 0.05.
+The results, which appear in Table \ref{experiment2results}, agree with the time complexity analyses presented above.
+Overall, we found the exact dynamic program to be the fastest algorithm, a result that echoes the results of computational studies on knapsack problems.
+The branch-and-bound algorithm proved impractical for even medium-sized instances.
+%cite computational studies knapsack
+
+\newcommand{\missing}{---}
+\newcommand{\lastptofcaption}{\ifen
+For each value of $m$, \nmarkets~markets were generated, and the computation time was recorded as fastest of three repetitions of the algorithm.
+The table shows the average time (standard deviation) in milliseconds over the \nmarkets~instances.
+\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복해서 그 중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (평균편차) 시간을 밀리초 단위로 나타난다.\fi}
+
+\begin{table}[h!] \centering
+\begin{tabular}{r|rr|rr}
+\ifen 
+ \textbf{\begin{tabular}[r]{@{}r@{}}Number of\\schools $m$\end{tabular}}& \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{algorithmforlargeh}\\with list\end{tabular}}}  & \multicolumn{2}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{algorithmforlargeh}\\with heap\end{tabular}}}\\ \hline
+ \else
+ \textbf{\begin{tabular}[r]{@{}r@{}}학교의\\개수 $m$\end{tabular}}& \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{algorithmforlargeh}\\(목록 구현)\end{tabular}}}  & \multicolumn{2}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{algorithmforlargeh}\\(힙 구현)\end{tabular}}}\\ \hline
+ \fi
+%16         & \phantom{000}0.0712748          &    (\phantom{000}0.00755644)              & \phantom{000}0.0977929      &     (\phantom{000}0.00367772)            \\
+%64         & \phantom{000}0.95445\phantom{00}          &    (\phantom{000}0.0822217\phantom{0})              & \phantom{000}1.28911\phantom{00}      &     (\phantom{000}0.0517071\phantom{0})            \\
+%256        & \phantom{00}14.8587\phantom{000}          &    (\phantom{000}0.725739\phantom{00})              & \phantom{00}20.3465\phantom{000}      &     (\phantom{000}0.708838\phantom{00})            \\
+%1024       & \phantom{0}287.057\phantom{0000}          &    (\phantom{00}81.9166\phantom{0000})              & \phantom{0}390.925\phantom{0000}      &     (\phantom{0}110.501\phantom{00000})            \\
+%4096       & 4826.75\phantom{00000}          &    (1432.85\phantom{000000})              & 7304.67\phantom{00000}      &     (2071.38\phantom{000000})
+16         & 0.07          &    (0.01)              & 0.01      &     (0.00)            \\
+64         & 0.95          &    (0.08)              & 1.29      &     (0.05)            \\
+256        & 14.86         &    (0.73)              & 20.35     &     (0.71)            \\
+1024       & 287.06          &    (81.92)              & 390.93    &     (110.50)            \\
+4096       & 4826.75         &    (1432.85)              & 7304.67     &     (2071.38)                                     
+\end{tabular}
+\caption{\label{experiment1results}
+\ifen Time to compute an optimal portfolio for an admissions market with homogeneous application costs using Algorithm \ref{algorithmforlargeh} when $\mathcal{C}$ is stored as a list and as a heap.
+\lastptofcaption~In every case, $h = m/2$.
+\else 알고리즘 \ref{algorithmforlargeh}에서 $\mathcal{C}$를 목록 또는 힙 구현으로 저장했을 때, 
+동일 지원 비용 입학 시장의 최적 지원 포트폴리오를 계산하는 시간. \lastptofcaption~모든 경우, $h = m/2$. \fi}
+\end{table}
+
+\begin{table}[h!] \centering
+\begin{tabular}{r|rr|rr|rr|rr}
+\ifen
+\textbf{\begin{tabular}[r]{@{}r@{}}Number of \\schools $m$\end{tabular}}&\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisbnb}:\\Branch \& bound\end{tabular}}}  & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP1}:\\App.
+costs DP\end{tabular}}}  &\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.5$\end{tabular}}}  & \multicolumn{2}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.05$\end{tabular}}}   \\ \hline
+\else
+\textbf{\begin{tabular}[r]{@{}r@{}}학교의\\개수 $m$\end{tabular}}&\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘  \ref{ellisbnb}:\\분지한계법\end{tabular}}}  & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP1}:\\지원비 동적 계획\end{tabular}}}  &\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.5$\end{tabular}}}  & \multicolumn{2}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.05$\end{tabular}}}   \\ \hline
+\fi
+    5 &         0.05 &     (0.02) &       0.01 &      (0.00) &    0.09 &  (0.01) &     0.26 &         (0.08) \\
+   10 &            — &        (—) &       0.13 &      (0.03) &    0.78 &  (0.07) &     2.89 &         (0.29)            
+\end{tabular}%
+\caption{\label{experiment2results}
+\ifen Time to compute an optimal or $(1 - \varepsilon)$-optimal portfolio for an admissions market with heterogeneous application costs using the three algorithms developed in section \ref{hetappcosts}.
+%branch-and-bound routine (Algorithm \ref{ellisbnb}), exact dynamic program (Algorithm \ref{ellisDP1}), and the fully polynomial-time approximation scheme (Algorithm \ref{ellisDP3}).
+The branch-and-bound algorithm is impractical for large markets. \lastptofcaption
+\else  %분지한계법 (알고리즘 \ref{ellisbnb}), 정확한 동적 계획 (알고리즘 \ref{ellisDP1}), 그리고 전체 다항 시간 근사 해법 (알고리즘 \ref{ellisDP3})을
+ \ref{hetappcosts}절에서 도출한 3개의 알고리즘을 사용할 때, 다양한 지원 비용으로 갖춘 입학 시장의 최적 또는 $(1- \varepsilon)$-최적 포트폴리오를 계산하는 시간. 분지한계법은 큰 시장에서 비실용적이다. \lastptofcaption \fi}
+\end{table}
 
@@
 \centering
-\includegraphics[width=0.75\textwidth]{./plots/samplemarket.pdf}
+\includegraphics[width=0.85\textwidth]{./plots/samplemarket.pdf}
   \caption{\label{samplemarket}
@@
 The optimal portfolio was computed using Algorithm \ref{ellisDP1}.
- \else $m=60$개의 학교로 구성된 전형적인 무작위로 생성한 인스턴스와 해당 최적 포트폴리오. 지원비 $g_j$는 $\{5, \dots, 10\}$에서 균일한 확률로 선정했으며 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
+ \else $m=60$개의 학교로 구성된 전형적인 무작위로 생성한 인스턴스와 해당 최적 포트폴리오. 지원비 $g_j$는 $\{5, \dots, 10\}$에서 균일한 확률로 선정했으며 기호의 넓이와 비례된다. 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
 \end{figure}
 
-The dynamic programs, namely Algorithms \ref{ellisDP1} and \ref{ellisDP3}, were implemented using recursive functions and memoization.
-Our implementation of Algorithm \ref{ellisDP3} differed slightly from that given above: We represented portfolio valuations in \emph{binary} rather than decimal, with the definitions of $P$ and $\mathcal{V}$ modified accordingly, and instead of fixed-point numbers, we worked in integers by multiplying each $t_j$-value by $2^P$.
-These modifications yield a substantial performance improvement without changing the fundamental algorithm design or complexity analysis.
+\ifen \subsection{Implementation notes} \else \subsection{알고리즘의 이행}\fi
+We chose to implement our algorithms in the Julia language (v1.7.0) because its system of type inference allows the compiler to optimize for differential cases such as when the $t_j$-values are integers or floating-point numbers, and because Julia offers convenient macros for parallel computing (Bezanson et al.
+2017).
+We made extensive use of the DataStructures.jl package (v0.18.11).
+The code is available at \url{https://github.com/maxkapur/OptimalApplication}.
+
+The dynamic programs, namely Algorithms \ref{ellisDP1} and \ref{ellisDP3}, were implemented using recursive functions and memoization rather than a full lookup table.
+Our implementation of Algorithm \ref{ellisDP3} also differed from that described in section \ref{fptashead} in that we represented portfolio valuations in \emph{binary} rather than decimal, with the definitions of $P$ and $\mathcal{V}$ modified accordingly, and instead of fixed-point numbers, we worked in integers by multiplying each $t_j$-value by $2^P$.
+These modifications yielded a substantial performance improvement without changing the fundamental algorithm design or complexity analysis.
 
@@
 김민희. 2015. ``{[대입 수시 전략]} 총 6번의 기회 $\cdots$ `상향·소신·안정' 분산 지원하라.'' 중앙일보, 8월 26일. \url{https://www.joongang.co.kr/article/18524069}.\fi
+
+Bezanson, Jeff, Alan Edelman, Stefan Karpinski, and Viral B. Shah. 2017. ``Julia: A Fresh Approach to Numerical Computing.'' \emph{SIAM Review} 59: 65–98. \url{https://doi.org/10.1137/141000671}.
 

### 2022-03-08 benchmark/benchmark.jl

@@ -1,9 +1,11 @@
 using OptimalApplication
 using DataFrames
+using PrettyTables
 using Random
 using StatsBase
 using Statistics
 using Base.Threads
+import Printf: @sprintf
 
 fullscale = false
 
@@ -50,8 +52,8 @@ function benchmark1()
         for (j, m) in enumerate(M)
             mkt = randSCM(m)
             sizes[i, j] = m
-            times_dict[i, j] = 1000 * minimum(@elapsed applicationorder(mkt; datastructure = :dict) for r in 1:n_reps)
-            times_heap[i, j] = 1000 * minimum(@elapsed applicationorder(mkt; datastructure = :heap) for r in 1:n_reps)
+            times_dict[i, j] = minimum(@elapsed applicationorder(mkt; datastructure = :dict) for r in 1:n_reps)
+            times_heap[i, j] = minimum(@elapsed applicationorder(mkt; datastructure = :heap) for r in 1:n_reps)
         end
     end
 
@@ -90,12 +92,12 @@ function benchmark2()
             mkt = randVCM(m)
             sizes[i, j] = m
             if m ≤ bnbcutoff
-                times_bnb[i, j] = 1000 * minimum(@elapsed optimalportfolio_branchbound(mkt; maxit=40000) for r in 1:n_reps)
+                times_bnb[i, j] = minimum(@elapsed optimalportfolio_branchbound(mkt; maxit = 40000) for r in 1:n_reps)
             end
-            times_dp[i, j] = 1000 * minimum(@elapsed optimalportfolio_dynamicprogram(mkt) for r in 1:n_reps)
+            times_dp[i, j] = minimum(@elapsed optimalportfolio_dynamicprogram(mkt) for r in 1:n_reps)
 
             for (k, epsilon) in enumerate(epsilons)
-                times_fptas[k][i, j] = 1000 * minimum(@elapsed optimalportfolio_fptas(mkt, epsilon) for r in 1:n_reps)
+                times_fptas[k][i, j] = minimum(@elapsed optimalportfolio_fptas(mkt, epsilon) for r in 1:n_reps)
             end
         end
     end
@@ -117,10 +119,21 @@ function benchmark2()
 end
 
 
+function fmter(v, i, j)
+    j == 1 && return v
+
+    v = ismissing(v) ? "—" : @sprintf "%1.2f" 1000*v
+
+    return isodd(j) ? "($v)" : v
+end
+
+
 @time begin
     println()
[…]

### 2022-03-08 src/VariedCostsBnB.jl

@@ -45,7 +45,7 @@ hash(nd::Node) = hash((nd.I, nd.N))
 # Depth-first strategy:
 # isless(nd1::Node, nd2::Node) = isless(nd1.v_I, nd2.v_I)
 # Breadth-first strategy:
-isless(nd1::Node, nd2::Node) = isless(nd1.v_LP, nd2.v_LP)
+# isless(nd1::Node, nd2::Node) = isless(nd1.v_LP, nd2.v_LP)
 
 
 
@@ -65,8 +65,8 @@ function generatechildren(nd::Node, mkt::VariedCostsMarket)
     end
 
     # School we will branch on. In principle it can by any school in fltr.
-    i = argmax(j -> mkt.f[j] * nd.t̄[j] / mkt.g[j], fltr)
-
+    # i = argmax(j -> mkt.f[j] * nd.t̄[j] / mkt.g[j], fltr)
+    i = argmax(j -> mkt.f[j] * nd.t̄[j], fltr)
 
     newN = setdiff(nd.N, i)
 
@@ -151,8 +151,12 @@ function optimalportfolio_branchbound(mkt; maxit = 1000000::Integer, verbose = f
             # pop!(tree)
             # delete!(treekeys, thisnodehandle)
         
-            thisnode, thisnodehandle = findmax(tree)
+            thisnodehandle, thisnode = argmax(hash_nd -> hash_nd[2].v_LP, tree)
             delete!(tree, thisnodehandle)
+            
+            # Another option: Select the node with best obj value. Works pretty bad. 
+            # thisnodehandle, thisnode = argmax(hash_nd -> hash_nd[2].v_I, tree)
+            # delete!(tree, thisnodehandle)
         end
 
         children = generatechildren(thisnode, mkt)

### 2022-03-10 paper/paper.tex
translations in bnb section

@@
 \newif\ifen
-%\entrue % Change to false to export Korean
+\entrue % Change to false to export Korean
 
@@
 \section*{요약}
-본 논문은 다수의 확률 변수로 구성된 포트폴리오의 기대 최적값을 예산 조건 하에서 최대화하는 문제를 고려한다.
-이를 최적 대학 지원 문제라고 부른다.
-각 확률 변수의 비용, 즉 각 대학의 지원비가 동일한 경우, 최적 포트폴리오는 예산 제약식으로 결정된 포함 사슬 관계를 가짐을 보이고 이에 따라 다항 시간 해법을 제시한다.
-대학의 지원비가 서로 다른 경우, 문제가 NP-complete함을 증명한다.
-일반적인 상황을 위해 세 가지 해법을 도출한다.
+본 논문은 다수의 확률 변수로 구성된 포트폴리오의 기대 최대값을 예산 조건 하에서 최대화하는 문제를 고려한다.
+이를 대학 지원 최적화 문제라고 부른다.
+각 확률 변수의 비용, 즉 각 대학의 지원 비용이 동일한 경우, 최적 포트폴리오는 예산 제약식으로 결정된 포함 사슬 관계 성질을 가짐을 보이고 이를 바탕으로 다항 시간 해법을 제시한다.
+대학의 지원 비용이 서로 다른 경우, 문제가 NP-complete함을 증명한다.
+세 가지 일반적인 해법을 도출한다.
 첫째는 분지한계 기반 해법이다.
-둘째는 의사 다항 시간 안에 정확한 해를 출력하는 동적 계획이다.
-셋째는 전체 다항 시간 근사 해법이다.
+둘째는 의사 다항 시간 안에 정확한 해를 출력하는 동적 계획 해법이다.
+마지막으로 다른 동적 계획 기반으로 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme)이 존재함을 보인다.
 
@@
 \ifen Here $\mathcal{C} = \{ 1 \dots m\}$ is an index set; $H > 0$ is a budget parameter; for $j = 1 \dots m$, $g_j > 0$ is a cost parameter and $Z_j$ is a random, independent Bernoulli variable with probability $f_j$; and for $j = 0\dots m$, $t_j$ is a utility parameter.
-\else 단,  $\mathcal{C} = \{ 1 \dots m\}$은 지표 집합이며 $H$는 예산 모수이다. 각 $j = 1 \dots m$에 대해 $g_j > 0$는 비용 모수이며 $Z_j$는 확률 $f_j$를 가지는 서로 독립적인 Bernoulli 변수이다. 각 $j = 0 \dots m$에 대해 $t_j$는 효용 모수이다. \fi
+\else 단,  $\mathcal{C} = \{ 1 \dots m\}$은 지표 집합이며 $H$는 예산을 나타내는 모수이다. 각 $j = 1 \dots m$에 대해 $g_j > 0$는 비용 모수이며 $Z_j$는 확률 $f_j$를 가지는 서로 독립적인 Bernoulli 변수이다. 각 $j = 0 \dots m$에 대해 $t_j$는 효용 모수이다. \fi
 
@@
 Let $g_j$ denote the application fee for $c_j$ and $H$ the student's total budget to spend on application fees.
-Lastly, let $f_j$ denote the student's probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not.
+(Alternatively, the budget constraint may represent the time needed to complete each application, or a legal limit on the number of applications permitted.) Lastly, let $f_j$ denote the student's probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not.
 It is appropriate to assume that the $Z_j$ are statistically independent as long as $f_j$ are probabilities estimated specifically for this student (as opposed to generic acceptance rates).
@@
 \else
-다음 해석에 따라 이를 `최적 대학 지원 포트폴리오 문제'라고 부른다.
+다음 해석에 따라 이를 `대학 지원 포트폴리오 최적화 문제'라고 부른다.
 $m$개의 대학교를 가지는 입학 시장을 고려하자.
 $j$번째 학교의 이름은 $c_j$이다.
-이 시장에 참가하는 한 학생을 고려해서 그가 $c_j$에 진학하면 $t_j$ 단위의 효용이 발생한다고 하자.
-단, 어떤 대학에도 진학하지 않는 경우 그의 효용은 $t_0$이다.
-$c_j$의 지원비가 $g_j$이며 학생이 지원비에 쓸 수 있는 예산이 $H$라고 하자.
-마지막, $c_j$에 지원하면 학생이 합격할 확률이 $f_j$이라고 하자.
+어떤 학생이 $c_j$에 입학하게 되면 효용 $t_j$를 얻게 된다고 하자.
+단, 어떤 대학에도 진학하지 않는 경우 그 학생의 효용은 $t_0$이다.
+$c_j$의 지원 비용이 $g_j$이며 학생이 지원에 쓸 수 있는 예산이 $H$라고 하자.
+(예산 제약 조건은 또한 원서를 작성하는 시간이나 지원 개수에 대한 법적인 제한을 나타낼 수도 있다.) 마지막으로, $c_j$에 지원하면 학생이 합격할 확률이 $f_j$이라고 하자.
 따라서 합격하면 $Z_j = 1$, 합격 안 하면 $Z_j = 0$이 된다.
-$f_j$가 (학교의 전체적인 합격률이 아니라) 바로 이 학생의 합격 확률일 때 $Z_j$의 확률적 독립성을 적절한 가정이다.
-그러면 학생의 목표는 주어진 예산 안에서 학생이 합격하는 가장 좋은 학교의 기대 효용을 최대화하는 것이다.
-위 문제의 최적해가 $\mathcal{X}$일 때, 학생의 최적 대학 지원 전략은 $\mathcal{X}$에 지원하는 것이다.
+$f_j$가 (학교의 전체적인 합격률이 아니라) 바로 이 학생의 합격 확률일 때 $Z_j$의 확률적 독립성은 적절한 가정이다.
+그러면 학생의 목표는 주어진 예산 안에서 학생이 합격하는 학교들에서 얻는 효용 중 기대 최대값을 최대화하는 것이다.
+위 문제의 최적해가 $\mathcal{X}$일 때, 학생의 최적 대학 지원 전략은 $\mathcal{X}$에  속한 학교로 지원하는 것이다.
 \fi
@@
 In computational finance, traditional portfolio optimization models weigh the sum of expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Markowitz 1952; Meucci 2005).
-But college applicants maximize the value of their \emph{best} asset: If a student is admitted to her $j$th choice, then she is indifferent as to whether she gets into her $(j+1)$th choice.
+But college applicants maximize the expected value of their \emph{best} asset: If a student is admitted to her $j$th choice, then she is indifferent as to whether she gets into her $(j+1)$th choice.
 As a result, the valuation function that students maximize is \emph{convex} in the expected utility associated with individual applications.
@@
 재정학 분야에서 고전적 포트폴리오 최적화 모형은 전체 자산에 대한 기대 총이익에서 위험회피 항을 뺌으로 선형식으로 제약된 오목 최대화 문제를 이룬다 (Markowitz 1952; Meucci 2005).
-그러나 대학 지원자는 가치가 제일 높을 단 한 자산의 가치를 최대화하고자 한다.
-어떤 학생이 자신의 $j$위 학교에 합격하면, $(j+1)$위 학교에 합격하든 불합격하든 무관심하므로 다른 상황이다.
-이에 따라 학생이 최대화하려는 함수는 각 지원의 기대 효용에 대해 오목이 아닌 볼록 함수다.
-또한 전형적인 입학 시장에서 대학의 효용과 합격 확률이 서로 반비례하므로 대학 지원 문제에 위험 관리가 내포되어 있다.
+그러나 대학 지원자는 가치가 제일 높은 단일 자산의 기대 가치를 최대화하고자 한다.
+어떤 학생이 자신이 $j$번째로 선호하는 학교에 합격하면 $(j+1)$번째 학교의 합격 여부는 무관한 상황이 된다.
+이는 학생이 최대화하려는 함수를 각 지원 발송의 기대 효용에 대해 오목이 아닌 볼록 함수로 만든다.
+또한 입학 시장에서는 전형적으로 대학의 효용과 합격 확률이 서로 반비례하므로 대학 지원 문제는 위험 관리를 포함하게 된다.
 특히 선호도가 높으며 붙기 어려운 “상향” 지원 학교(reach school)와 선호도가 낮으며 붙기 쉬운 “안정” 지원 학교(safety school) 사이의 균형을 고려해야 한다 (김민희 2015).
 마지막, 대학 지원의 조합적인 본성으로 인해 연속적인 포트폴리오 최적화 문제에서 흔히 사용하는 기울기 해법으로는 풀기가 어렵다.
-Chao는 지원 비용을 제약 조건 대신 목적함수의 한 항으로 모형화했으며 모형을 추정하기 위해 $m=8$이 되도록 학교를 먼저 클러스터로 모았다.
-이는 열거법을 통해 문제를 쉽게 풀 수 있는 규모이지만 본 연구는 더 일반적인 해법을 목표로 추구한다.
+Chao는 지원 비용을 제약 조건 대신 목적함수의 한 항으로 모형화했으며, 모형의 모수를 추정하기 위해 $m=8$이 되도록 학교들의 클러스터를 먼저 구성했다.
+이는 열거법을 통해 문 쉽게 풀 수 있는 규모이지만 본 연구는 더 일반적인 $m$에 대한 해법을 추구한다.
 \fi
@@
 Our problem superficially resembles the preference-order knapsack problem considered by Steinberg and Parks and Carraway et al., but these models lack the college application problem's unique ``maximax'' form.
-Additionally, unlike those models, we do not attempt to replace the scalar objective function with a preference order over \emph{outcome distributions,} which introduces technical issues concerning competing notions of stochastic dominance (Sniedovich 1980).
+Additionally, unlike those models, we do not attempt to replace the real-valued objective function with a preference order over \emph{outcome distributions,} which introduces technical issues concerning competing notions of stochastic dominance (Sniedovich 1980).
 We take for granted the student's preferences over \emph{outcomes} (as encoded in the $t_j$-values), and are interested instead in an efficient computational approach to the well-defined problem above.
@@
 본 논문에서 제시하는 분지한계법과 동적 계획 해법은 배낭 문제를 위한 기존 알고리즘과 매우 비슷하다 (Martello와 Toth 1990, \S\,2.5--6).
-입학 확률을 적당히 조정하면 원래 목적함수를 특성벡터의 선형 함수에 대한 임의로 정확한 근사로 변환할 수 있으며 NP-completeness 증명에서 이 사실을 활용한다.
-배낭 문제에 확률성을 도입한 선행 연구 중, 각 상품의 효용이 정해진 확률 분포로 결정되는 모형(Steinberg와 Parks 1979; Carraway 외 1993), 그리고 배낭에 삽입한 다음에 상품의 무게가 관측되는 온라인 모형(Dean et al.
-2008)과 같은 바가 있다.
-본 연구가 고려하는 문제는 Steinberg와 Parks 그리고 Carraway 외가 살펴본 ``우선순위 배낭 문제''와는 피상적으로 비슷하지만, 우선순위 배낭 문제는 대학 지원 문제의 특색인 `maximax' 형태를 가지지는 않는다.
-또한 우선순위 모형과 달리, 본 연구에서 스칼라값 목적함수를 `결과 분포'의 선호 순위로 대체하고자 노력하지 않는다.
-확률적 지배와 관련된 기술적인 문제가 생기기 때문이다 (Sniedovich 1980).
-대신 $t_j$-값으로 유도된 학생의 `결과'에 대한 선호 순위를 그대로 받아들이고 위에서 정의한, 모호하지 않은 문제를 위한 효율적인 계산법을 지향한다.
+입학 확률을 적당히 조정하면 특성벡터의 선형 함수를 원래 목적함수로 원하는 만큼의 정확성을 가지고 근사할 수 있으며 NP-completeness 증명에서 이 성질을 활용한다.
+배낭 문제에 확률성을 도입한 선행 연구 중, 각 상품의 효용이 정해진 확률 분포로 결정되는 모형(Steinberg와 Parks 1979; Carraway 외 1993), 그리고 배낭에 삽입한 다음에 상품의 무게를 관측할 수 있는 온라인 모형 (Dean et al.
+2008) 등이 있다.
+본 연구가 고려하는 문제는 Steinberg와 Parks 그리고 Carraway 외가 고려한 ``우선순위 배낭 문제''와는 유사성을 가지지만, 우선순위 배낭 문제는 대학 지원 문제의 특색인 `maximax' 형태를 가지지는 않는다.
+또한 우선순위 모형과 달리, 본 연구에서 실수값 목적함수를 선호 순위를 정의해야 하는 `결과 분포'로 대체할 필요가 없다.
+확률적 우위에 대한 상반된 개념들의 문제를 야기시키기 때문이다 (Sniedovich 1980).
+대신 $t_j$-값으로 유도된 학생의 `결과'에 대한 선호 순위를 받아들여 위에서 정의한 것처럼 명확히 정의된 문제를 위한 효율적인 계산법을 지향한다.
 \fi
@@
 \else
-본 연구에서 특히 탐욕 해법의 타당성에 관심을 기울인다.
-예를 들어 예산이 다 소비될 때까지 목적함수를 가장 많이 증가시키는 학교를 차례대로 선택하는 알고리즘은 일종의 탐욕 해법이다.
-탐욕 해법은 예산 $H$으로 모수화된 해의 순서를 유도하며 그의 원소들은 `포함 사슬 관계'(nestedness)로 서로에 연결된다.
-즉 $H \leq H'$일 때, 예산 $H$에 해당하는 탐욕 해는 예산 $H'$에 해당하는 탐욕 해의 부분집합이다.
-Rozanov와 Tamir (2020)가 주장하듯, 최적해가 포함 사슬 관계를 맺음을 알면 최적해를 구하는 데일뿐더러 정보가 불확실한 상황에서 최적해를 이행하는 데에도 유용하다.
+본 연구에서 특히 탐욕 해법의 가능성에 관심을 기울인다.
+예를 들어 예산이 다 소비될 때까지 목적함수를 가장 많이 증가시키는 학교를 차례대로 추가하는 알고리즘은 일종의 탐욕 해법이다.
+탐욕 해법은 예산 $H$으로 모수화된 해의 순서를 유도하며 그의 원소들은 `포함 사슬 관계'(nestedness)로 연결한다.
+즉 $H \leq H'$일 때, 예산 $H$에 해당하는 탐욕 해는 예산 $H'$에 해당하는 탐욕 해의 부분집합이 된다.
+Rozanov와 Tamir (2020)가 주장하듯, 최적해가 포함 사슬 관계를 가지면 최적해를 구하는 것뿐 아니라 정보가 불확실한 상황에서 최적해를 구현하는 데에도 유용하다.
 가령, 많은 미국 대학의 지원 기한은 11월 초인데 학업 계획서를 학교의 취향에 맞춰서 작성해야 하므로 여름부터 원서를 작성하는 학생이 많다.
 그러나 지원 예산은 10월 말까지 모를 수도 있다.
-포함 사슬 관계, 또는 탐욕 해법의 타당성은 완전한 예산 정보가 없음에도 학교가 최적 포트폴리오에 진입하는 순서대로 원서를 작성하는 식으로 최적 전략을 이행해 낼 수 있다고 의미한다.
+포함 사슬 관계, 또는 탐욕 해법의 타당성은 완전한 예산 정보가 없어도 학교가 최적 포트폴리오에 진입하는 순서대로 원서를 작성하면 최적 전략을 구현해 낼 수 있음을 의미한다.
 \fi
@@
 \else
-한편, 탐욕 알고리즘이 좋은 근사 해법이며 다른 가정을 추가하면 정확한 알고리즘이 되는 최적화 모형 몇몇이 알려져 있다.
-대표적인 바는 카디널리티 제약 위에 자모(子模) 집합 함수(submodular set function)를 최대화하는 문제이다 (Fisher et al.
+탐욕 알고리즘이 좋은 근사해거나 정확한 알고리즘이 되는 최적화 모형들이 알려져 있다.
+집합 크기 제약 하에서 하위 모듈 집합 함수(submodular set function)를 최대화하는 문제가 대표적인 예다 (Fisher et al.
 1978).
-반면에 이진 배낭 문제 같은 경우에서는 가장 직관적인 탐욕 알고리즘이 무한히 비최적인 해를 출력하는 인스턴스를 만들 수 있다 (Vazirani 2001).
-대학 지원 문제에 대해 배낭 문제와 매우 유사한 결과를 도출할 수 있다.
+반면에 이진 배낭 문제 같은 경우에서는 가장 직관적인 탐욕 알고리즘이 최적과 거리가 먼 해를 출력하는 예를 만들 수 있다 (Vazirani 2001).
+대학 지원 문제와 배낭 문제가 서로 매우 유사함을 보인다.
 모든 $g_j=1$인 특수한 경우에서, 최적 포트폴리오가 탐욕 알고리즘의 타당성과 동등한 포함 사슬 관계를 만족하는 것을 증명한다.
-이 경우는 지원비가 없으며 정시 모집 기간에 학교 3개에만 지원할 수 있는 한국 입학 과정과 같다.
-그러나 일반적인 경우에서 포함 사슬 관계가 성립하지 않을뿐더러 탐욕 알고리즘에 대해 최소한 어떤 근사 계수를 보장할 수 없다.
-대신 분지한계법, 전형적인 입학 시장 인스턴스에 대해 효율적인 의사 다항 시간 동적 계획, 그리고 전체 다항 시간 안에 $(1 - \varepsilon)$-근사해를 출력하는 알고리즘과 같은 3가지 해법은 제안한다.
+이 경우는 지원 비용이 없으며 정시 모집 기간에 학교 3개에만 지원할 수 있는 한국 입학 과정과 같다.
+그러나 일반적인 경우에서 포함 사슬 관계가 성립하지 않을뿐더러 탐욕 알고리즘이 어떤 근사 계수를 보장할 수 없을 보일 수 있다.
+대신 분지한계법, 전형적인 입학 시장 인스턴스에 대해 효율적인 의사 다항 시간 동적 계획, 그리고 완전 다항 시간 안에 $(1 - \varepsilon)$-근사해를 출력하는 알고리즘 등 3가지 해법을 제안한다.
 \fi
@@
 \ifen Section \ref{preliminaries} introduces some additional notation and assumptions that can be imposed with trivial loss of generality. 
-\else \ref{preliminaries}절은 표기법은 명시하고 일방성을 잃지 않는 추가의 가정을 정리한다.\fi
+\else \ref{preliminaries}절은 표기법과 일반성을 제한하지 않는 편의적 가정을 제시한다.\fi
 
@@
 \else
-\ref{homogappcosts}절에서 각 $g_j = 1$이고 $H$가 $h \leq m$인 자연수인 특수한 경우를 고려한다. 직관적인 휴리스틱 해법이 실제로 $1/h$-근사 해법임을 보인다. 그다음에 최적 포트폴리오의 예산 제약에 대한 포함 사슬 관계를 증명하고 이를 이용해서 $O(hm)$-시간 정확한 해법을 도출한다.
+\ref{homogappcosts}절에서 각 $g_j = 1$이고 $H$가 $h \leq m$이고 자연수인 특수한 경우를 고려한다. 직관적인 휴리스틱 해법이 실제로 $1/h$-근사 해법임을 증명한다. 그다음에 최적 포트폴리오의 예산 제약에 대한 포함 사슬 관계를 보이고 이를 이용해서 $O(hm)$-시간 정확한 해법을 도출한다.
 \fi
@@
 \else
-\ref{hetappcosts}절에서 각 학교의 지원비가 다를 수 있는 상황을 다룬다.
+\ref{hetappcosts}절에서 각 학교의 지원 비용이 동일하지 않은 일반적인 상황을 다룬다.
 이진 배낭 문제에서의 다항 변환을 통해서 포트폴리오 최적화 문제가 NP-complete임을 증명한다.
-일반적인 상황을 위해 3개의 해법은 제안한다.
+또한 3개의 일반적인 해법을 제안한다.
 첫째는 분지한계법이다.
-둘째는 총 지출액으로 탐색하는 $O(Hm + m \log m)$과 같은 의사 다항 시간 동적 계획이다.
-셋째는 분수값은 내린 포트폴리오 가치로 탐색하는 또다른 동적 계획이다.
-이를 통해서 $O(m^3 / \varepsilon)$-시간 안에 $(1 - \varepsilon)$-근사해를 출력하는 전체 다항 시간 근사 해법(fully polynomial-time approximation scheme)을 얻는다.
+둘째는 총 지출액으로 탐색하는 $O(Hm + m \log m)$과 같은 의사 다항 시간 동적 계획 해법이다.
+셋째는 실수값을 분수값으로 내림한 포트폴리오 가치를 탐색하는 또다른 동적 계획 해법이다.
+이를 통해서 $O(m^3 / \varepsilon)$-시간 안에 $(1 - \varepsilon)$-근사해를 출력하는 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme)을 얻는다.
 \fi
@@
 
-간단한 결론이 뒤를 이은다.
+그리고 간단한 결론을 제시한다.
 \fi
@@
 
+\ifen
 Given an application portfolio, let $p_j(\mathcal{X})$ denote the probability that the student attends $c_j$.
@@
 Hence, for $j= 0\dots m$,
+\else
+주어진 지원 포트폴리오에 대해 학생이 $c_j$로 진학할 확률을 $p_j(\mathcal{X})$라고 하자. 이 사건이 발생하는 필수충분 조건은 $c_j$에 지원하고, $c_j$에서 합격받고, $c_j$보다 선호하는 (즉, 지표가 더 높은) 학교에서 불합격받는 것이다. 따라서  $j= 0\dots m$에 대해
+\fi
 \begin{align}
@@
 \displaystyle f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}), \quad & j \in \{0\}\cup\mathcal{X}\\
-0, \quad & \text{otherwise}
+0, \quad & \text{\ifen otherwise\else otherwise.\fi}
 \end{cases} 
 \end{align}
+\ifen
 where the empty product equals one. The following proposition follows immediately.
+\else
+단, 공집합의 곱은 1이다. 다음 제의는 직설적인 결과다.
+\fi
 
@@
 
+\ifen
 Next, we show that without loss of generality, we may assume that $t_0 = 0$ (or any constant less than $t_1$).
+\else
+그다음에, 일반성을 제한하지 않고 $t_0 = 0$(또는 $t_1$보다 작은 임의의 상수)이라고 가정할 수 있음을 보이자.
+\fi
 \begin{theorem} \label{assumetzerozero}
@@
 \begin{proof}
+\ifen
 By definition, $\sum_{j=0}^m p_j(\mathcal{X}) = \sum_{j \in \{0\}\cup\mathcal{X}} p_j(\mathcal{X}) = 1$. Therefore
+\else
+정의에 따라 $\sum_{j=0}^m p_j(\mathcal{X}) = \sum_{j \in \{0\}\cup\mathcal{X}} p_j(\mathcal{X}) = 1$이다. 그러면 다음 수식으로 증명을 완성할 수 있다.
+\fi
 \begin{align}
@@
 = v(\mathcal{X}; t_j) - \gamma
-\end{split}
-\end{align}
-which completes the proof. 
+\end{split} \ifen\else\qedhere\fi
+\end{align}
+\ifen which completes the proof.\fi
 \end{proof}
@@
 \ifen \section{Homogeneous application costs}  \else \section{동일한 지원 비용} \fi \label{homogappcosts}
-In this section, we derive a polynomial-time algorithm for the special case in which $g_j = 1$ and $H$ is a constant $h \leq m$.
+\ifen In this section, we focus on the special case in which each $g_j = 1$ and $H$ is a natural number $h \leq m$.
+We show that an intuitive heuristic is in fact a $1/h$-approximation algorithm, then derive an exact polynomial-time solution algorithm.
 This case is similar to the centralized college admissions process in Korea, where there is no application fee, but by law, students are allowed to apply to no more than $h$ schools.
@@
 Throughout this section, we will call the applicant Alma, and refer to the corresponding optimization problem as Alma's problem.
-
-\begin{problem}[Alma’s problem]
-Alma's optimal college application portfolio is given by the solution to the following combinatorial optimization problem:
+\else
+본 절에서 모든 $g_j = 1$이며 $H$가 자연수 $h \leq m$인 특수한 경우에 집중한다.
+직관적인 휴리스틱 해법이 실제로 $1/h$-근사 해법임을 보인 다음에 정확한 다항 시간 해법을 도출한다.
+이 특수한 경우는 지원 비용이 없으며 국가 법에 따라 최대 $h$개의 학교에 지원할 수 있는 한국의 중심화된 대학 지원 과정과 비슷하다.
+(한국 경우,  $m=202$이고 $h=3$이다.) 정리 \ref{assumetzerozero}을(를) 적용해서 다른 언급이 없으면 $t_0 = 0$임을 가정하자.
+본 절 내내 지원자를 `알마'라고 부르고 해당 최적화 문제를 `알마의 문제'라고 부른다.
+\fi
+
+\begin{problem}[\ifen Alma’s problem\else 알마의 문제\fi]
+\ifen Alma's optimal college application portfolio is given by the solution to the following combinatorial optimization problem:
+\else
+알마의 최적 대학 지원 포트폴리오는 다음 조합 최적화 문제의 최적해이다.
+\fi
 \begin{align}
@@
 
-\ifen \subsection{Approximation properties of a na\"ive solution}  \else \subsection{Na\"ive 해법의 근사 성질} \fi
+\ifen \subsection{Approximation properties of a na\"ive solution}  \else \subsection{나이브 해법의 근사 성질} \fi
+\ifen
 The expected utility associated with a single school $c_j$ is simply $\operatorname{E}[t_j Z_j] = f_j t_j$. It is therefore tempting to adopt the following strategy, which turns out to be inoptimal.
-\begin{definition}[Na\"ive algorithm for Alma’s problem] \label{naivealgorithm}
+\else 
+단일 학교 $c_j$에 해당하는 기대 효용이 단순히 $\operatorname{E}[t_j Z_j] = f_j t_j$이므로 다음과 같은 알고리즘이 매력적으로 보일 수 있지만 사실은 최적해가 아니다.
+\fi
+\begin{definition}[\ifen Na\"ive algorithm for Alma’s problem\else 알마의 문제를 위한 나이브 해법\fi] \label{naivealgorithm}
+\ifen 
 Apply to the $h$ schools having the highest expected utility $f_j t_j$.
+\else
+기대 효용 $f_j t_j$가 가장 높은 $h$개의 학교로 지원한다.
+\fi
 \end{definition}
+\ifen 
 The basic error of this algorithm is that it maximizes $\operatorname{E}\left[\sum t_j Z_j \right]$ instead of $\operatorname{E}\left[\max \{t_j Z_j\} \right]$.
@@
 The following example shows that the na\"ive algorithm can produce a suboptimal solution.
+\else
+위 알고리즘의 기본 실수는 $\operatorname{E}\left[\max \{t_j Z_j\} \right]$ 대신  $\operatorname{E}\left[\sum t_j Z_j \right]$을 최대화하는 것이다. 결국에는 단 한 학교로만 진학할 수 있으므로 알마의 목적은 후자가 아닌 전자다. %학교 개별 효용의 기대 최대값이다. 
+다음 예는 나이브 해법이 최적해가 아닌 해를 출력할 수 있음을 보인다.
+\fi
 \begin{example}
+\ifen 
 Suppose $m=3$, $h=2$,  $ t= (70, 80, 90) $, and $f = (0.4, 0.4, 0.3)$. 
@@
 is the optimal solution. 
+\else
+$m=3$, $h=2$,  $ t= (70, 80, 90) $, 그리고 $f = (0.4, 0.4, 0.3)$이라고 하자.
+그러면 나이브 해법은 $\mathcal{T} = \{1, 2\}$를 선택하며 그의 기대 효용은 $v(\mathcal{T}) = 70(0.4)(1-0.4) + 80(0.4) = 48.8$이다.
+그러나 $\mathcal{X} = \{2, 3\}$를 선택하면 기대 효용은 $v(\mathcal{X}) = 80(0.4)(1-0.3) + 90(0.3) = 49.4$이고 이는 최적해다.
+\fi
 \end{example}
 
+\ifen
 In fact, the na\"ive algorithm is a $(1/h)$-approximation algorithm for Alma’s problem, as expressed in the following theorem.
-
-\begin{theorem} \label{oneoverhopt}
+\else
+다음 정리는 나이브 해법이 알마의 문제를 위한 $(1/h)$-근사 해법임을 의미한다.
+\fi
+
+\begin{theorem}[\ifen Accuracy of the na\"ive algorithm\else 나이브 해법의 정확성\fi] \label{oneoverhopt}
+\ifen 
 When the application limit is $h$, let $\mathcal{X}_h$ denote the optimal portfolio, and $\mathcal{T}_h$ the set of the $h$ schools having the largest values of $f_j t_j$.
 Then $v(\mathcal{T}_h) / v(\mathcal{X}_h) \geq 1/h$.
+\else
+지원 제한이 $h$일 때, 최적 포트폴리오가 $\mathcal{X}_h$고 기대 효용 $f_j t_j$가 가장 높은 $h$개의 학교의 집합이 $\mathcal{T}_h$라고 하자. 그러면 $v(\mathcal{T}_h) / v(\mathcal{X}_h) \geq 1/h$이다.
+\fi
 \end{theorem}
 \begin{proof}
+\ifen
 Because $\mathcal{T}_h$ maximizes the quantity $\operatorname{E}\bigl[ \sum_{j \in \mathcal{T}_h}\{ t_j Z_j \}\bigr]$, we have
+\else
+$\mathcal{T}_h$는 $\operatorname{E}\bigl[ \sum_{j \in \mathcal{T}_h}\{ t_j Z_j \}\bigr]$를 최대화하므로,
+\fi
 \begin{align} \label{oneoverhopt}
@@
 \leq h  \operatorname{E}\bigl[ \max_{j \in \mathcal{T}_h}\{ t_j Z_j \}\bigr]
-= h\,v(\mathcal{T}_h)
+= h\,v(\mathcal{T}_h).
 \end{split}
 \end{align}
+\ifen 
 where the final inequality follows from the concavity of the $\max\{\}$ operator.
+\else
+단, 마지막 부등호는 $\max\{\}$ 연산의 볼록성에 따라 성립한다.
+\fi
 \end{proof}
 
+\ifen
 The following example establishes the tightness of the approximation factor. 
+\else
+다음 예는 근사 계수가 타이트(tight)함을 보인다.
+\fi
 
 \begin{example} \label{tightexampleforoneoverhopt}
-Pick any $h$ and let $m = 2h$. For a small constant $\varepsilon \in (0, 1)$, let
+\ifen
+Pick any $h$ and let $m = 2h$. For a small constant $\varepsilon \in (0, 1)$, construct the market as follows.
+\else
+임의의 $h$를 택하고 $m = 2h$라고 하자. 작은 상수  $\varepsilon \in (0, 1)$에 대해 시장을 다음처럼 구성한다.
+\fi
 \begin{align*}
@@
 \end{align*}
+\ifen
 Since all $f_j t_j = 1$, the na\"ive algorithm can choose $\mathcal{T}_h = \{1, \dots, h\}$, with $v(\mathcal{T}_h) = 1$. But the optimal solution is $\mathcal{X}_h = \{h+1, \dots, m\}$, with
+\else
+모든 $f_j t_j = 1$이므로 나이브 해법은 $\mathcal{T}_h = \{1, \dots, h\}$를 출력할 수 있으며 $v(\mathcal{T}_h) = 1$이다. 그러나 최적해는 $\mathcal{X}_h = \{h+1, \dots, m\}$이며 그의 기대 효용은 다음과 같다.
+\fi
 \begin{equation*}
-v(\mathcal{X}_h) = \sum_{j= h +1}^m \Bigl( f_j t_j \prod_{j' = j+1}^m (1 - f_{j'}) \Bigr) =  \sum_{j= 1}^h  (1 - \varepsilon)^{j} \approx h.
+v(\mathcal{X}_h) = \sum_{j= h +1}^m \Bigl( f_j t_j \prod_{j' = j+1}^m (1 - f_{j'}) \Bigr) =  \sum_{j= 1}^h  (1 - \varepsilon)^{j} \approx h\ifen.\fi
 \end{equation*}
+\ifen
 Thus, as $\varepsilon$ approaches zero, we have $v(\mathcal{T}_h) / v(\mathcal{X}_h) \to 1/h$.
 (The optimality of $\mathcal{X}_h$ follows from the fact that it achieves the upper bound of Theorem \ref{oneoverhopt}.)
+\else
+따라서 $\varepsilon$이 0에 가까워지면서 $v(\mathcal{T}_h) / v(\mathcal{X}_h) \to 1/h$이 된다.  ($\mathcal{X}_h$의 최적성은 정리 \ref{oneoverhopt}이(가) 제시하는 상한을 실천하기 때문에 성립한다.)
+\fi
 \end{example}
@@
 \begin{corollary}
+\ifen
 The function $v(\mathcal{X})$ is not submodular. 
+\else
+$v(\mathcal{X})$는 하위 모듈 함수가 아니다.
+\fi
 \end{corollary}
 \begin{proof}
+\ifen
 If $v(\mathcal{X})$ is submodular, then theorem 4.2 of Fisher et al.
@@
 Example \ref{tightexampleforoneoverhopt} provides a counterexample.
+\else
+$v(\mathcal{X})$가 하위 모듈 함수라면 Fisher 외(1978)의 정리 4.2는 나이브 해법이 $v(\mathcal{T}_h) / v(\mathcal{X}_h)  \geq 1 - \left(\frac{h-1}{h}\right)^h$ 같은 근사 비율을 이룬다고 의미한다. 예 \ref{tightexampleforoneoverhopt}는 반례다.
+\fi
 \end{proof}
 
-Hope is not lost. We can still find the optimal solution in time polynomial in $h$ and $m$, as we will now show.
+\ifen
+Although the na\"ive algorithm is inoptimal, we can still find the optimal solution in time polynomial in $h$ and $m$, as we will now show.
+\else
+나이브 해법은 최적해가 아니지만 $h$와 $m$의 다항 시간 안에 최적해를 구할 수 있음을 보이자.
+\fi
 
@@
 새로운 시장에서 $\bar t_0 = 0$이므로 귀납법 가설을 적용하면 $\mathcal{Y}_{h-1} \subset \mathcal{Y}_h$임을 알 수 있다.
-따라서 증명이 완성된다.\footnote{유용한 변환을 발견한 임세호에게 감사를 표한다.} \fi
+따라서 증명이 완성된다.\footnote{이 유용한 변환을 발견한 임세호에게 감사를 표한다.} \fi
 \end{proof}
@@
 \Return{$\mathtt{X, V}$}
+\end{algorithm}
 } \else {
@@
 \end{algorithm}
+} \fi
 
@@
 \fi
+\newcommand{\Iteration}{\ifen Iteration~\else 반복단계~\fi}
 \begin{align*}
-\text{Iteration 1:}
+\text{\Iteration 1:}
 &&C &= \{1, 2, 3, 4, 5, 6, 7, 8\} \\
@@
 ~~\tag*{\(\implies \mathcal{X}_3 = \{4\} \)}\\
-\text{Iteration 2:}
+\text{\Iteration 2:}
 &&C &= \{1, 2, 3, 5, 6, 7, 8\} \\
@@
 ~~\tag*{\(\implies \mathcal{X}_3 = \{4, 2\} \)}\\
-\text{Iteration 3:}
+\text{\Iteration 3:}
 &&C &= \{1, 3, 5, 6, 7, 8\} \\
@@
 &&&\cdots\\
-\text{Iteration 8:}
+\text{\Iteration 8:}
 &&C &= \{6\}, f = \{0.03\}, t = \{177.622\}, f * t= \{\underline{5.329}\}
@@
 \else
-알고리즘은 $\mathtt{X} = [4, 2, 8, 1, 7, 3, 5, 6]$를 출력하며 최적 $h$-포트폴리오는 그의 첫 $h$개의 원소로 이루어진다.
-표 \ref{planetsdata}의 ``지원 순위'' 열은 $\mathtt{X}$의 역순열(inverse permutation)이며 이는 해당 학교가 최적 포트폴리오에 포함되는 최소 $h$-값을 의미한다.
+알고리즘의 출력은 $\mathtt{X} = [4, 2, 8, 1, 7, 3, 5, 6]$이며 최적 $h$-포트폴리오는 그의 첫 $h$개의 원소로 이루어진다.
+표 \ref{planetsdata}에서 등장하는 ``지원 순위'' 자료는 $\mathtt{X}$의 역순열(inverse permutation)이며 이는 해당 학교가 최적 포트폴리오에 포함되는 최소 $h$-값을 의미한다.
 그림 \ref{planetsplot}은 최적 포트폴리오의 가치를 $h$의 함수로 나타낸다.
@@
   \ifen Application limit $h$ versus the optimal portfolio valuation $v^* = v(\mathcal{X}_h)$ in a fictional market with $m=8$ schools.
-  \else $m=8$개의 학교로 구성된 가상 입학 시상에서, 각 지원 제한 $h$에 해당하는 최적 포트폴리오의 가치 $v^* = v(\mathcal{X}_h)$.\fi}
+The concave shape suggests the result of Theorem \ref{concavityinh}.
+  \else $m=8$개의 학교로 구성된 가상 입학 시상에서, 각 지원 제한 $h$에 해당하는 최적 포트폴리오의 가치 $v^* = v(\mathcal{X}_h)$. 곡선의 오목성은 정리 \ref{concavityinh}의 결과를 시사한다.\fi}
 \end{figure}
@@
 \pagebreak
-\ifen \section{Heterogeneous application costs} \else\section{다양한 지원 비용} \fi\label{hetappcosts}
+\ifen \section{Heterogeneous application costs} \else\section{동일하지 않은 지원 비용} \fi\label{hetappcosts}
+\ifen
 Now we turn to the more general problem in which the constant $g_j$ represents the \emph{cost} of applying to $c_j$ and the student, whom we now call Ellis, has a \emph{budget} of $H$ to spend on college applications.
-Applying Theorem \ref{assumetzerozero}, we assume $t_0 = 0$ and disregard the outside option throughout.
-
-\begin{problem}[Ellis's problem]
+Applying Theorem \ref{assumetzerozero}, we assume $t_0 = 0$ throughout.
+\else
+이제 $g_j$가 $c_j$에 지원하는 비용을 나타내며 학생이 지원에 쓸 수 있는 예산이 $H$인 일반적인 문제를 고려한다. 이 문제를 풀고자 하는 학생을 엘리스라고 부르자. 본 절 내내 정리 \ref{assumetzerozero}을(를) 적용해서 $t_0 = 0$임을 가정하자.
+\fi
+
+\begin{problem}[\ifen Ellis's problem\else 엘리스의 문제\fi]
+\ifen
 Ellis's optimal college application portfolio is given by the solution to the following combinatorial optimization problem.
+\else 
+엘리스의 최적 대학 지원 포트폴리오는 다음 조합 최적화 문제의 최적해이다.
+\fi
 \begin{align}
@@
 
-In this section, we show that this problem is NP-complete, then provide three algorithmic solutions: an exact branch-and-bound routine, an exact dynamic program, and a fully polynomial-time approximation scheme.
+\ifen
+In this section, we show that this problem is NP-complete, then provide three algorithmic solutions: an exact branch-and-bound routine, an exact dynamic program, and a fully polynomial-time approximation scheme (FPTAS).
+\else
+본 절에서 이 문제가 NP-complete임을 보이고 3가지 해법을 제시한다: 정확한 분지한계법 해법, 정확한 동적 계획 해법, 그리고 완전 다항 시간 근사 해법 (fully polynomial-time approximation scheme, FPTAS).
+\fi
 
 \ifen \subsection{NP-completeness} \else \subsection{NP-completeness}\fi
+\ifen
 The optima for Ellis's problem are not necessarily nested, nor is the number of schools in the optimal portfolio necessarily increasing in $H$. For example, if
@@
 In fact, Ellis’s problem is NP-complete, as we will show by a transformation from the binary knapsack problem, which is known to be NP-complete (Garey and Johnson 1979).
-
-\begin{problem}[Decision form of knapsack problem]
+\else
+엘리스 문제의 최적해는 필수적으로 포함 사슬 관계를 가지지 않으며 최적 포트폴리오에 속한 학교의 개수는 $H$에 대해 감소할 수도 있다. 예를 들어
+$f = (0.5, 0.5, 0.5)$, $t = (1, 1, 219)$, 그리고 $g = (1, 1, 3)$이면 $H = 2$에 대응하는 최적해가 $\{1, 2\}$이며 $H = 3$에 대응하는 최적해는 $\{3\}$임을 쉽게 알 수 있다.
+사실은 엘리스의 문제가 NP-complete이며 이를 이진 배낭 문제에서의 변환을 통해 증명할 수 있다.
+배낭 문제의 NP-completeness가 이미 알려져 있다 (Garey and Johnson 1979).
+\fi
+
+\begin{problem}[\ifen Decision form of knapsack problem\else 배낭 문제의 결정 형태\fi]
+\ifen 
 An \emph{instance} consists of a set $\mathcal{B}$ of $m$ objects, utility values $u_j \in \mathbb{N}$ and weight $w_j \in \mathbb{N}$ for each $j \in \mathcal{B}$, and target utility $U\in \mathbb{N}$ and knapsack capacity $W\in \mathbb{N}$.
 The instance is called a \emph{yes-instance} if and only if there exists a set $\mathcal{B’} \subseteq \mathcal{B}$ having $\sum_{j \in \mathcal{B’}} u_j \geq U$ and  $\sum_{j \in \mathcal{B’}} w_j \leq W$.
+\else
+`인스턴스'는 $m$개 상품으로 이루어진 집합 $\mathcal{B}$, 각 $j \in \mathcal{B}$에 대한 효용값 $u_j \in \mathbb{N}$와 무게 $w_j \in \mathbb{N}$, 목적 효용 $U\in \mathbb{N}$, 그리고 배낭 용량 $W\in \mathbb{N}$로 구성된다.
+인스턴스가 `예-인스턴스'가 될 필수충분 조건은 $\sum_{j \in \mathcal{B’}} u_j \geq U$와  $\sum_{j \in \mathcal{B’}} w_j \leq W$를 만족하는 집합 $\mathcal{B’} \subseteq \mathcal{B}$가 존재하는 것이다.
+\fi
 \end{problem}
 
-\begin{problem}[Decision form of Ellis’s problem]
+\begin{problem}[\ifen Decision form of Ellis's problem\else 엘리스 문제의 결정 형태\fi] \label{ellisdecisionform}
+\ifen
 An \emph{instance} consists of an instance of Ellis’s problem and a target valuation $V$.
 The instance is called a \emph{yes-instance} if and only if there exists a portfolio $\mathcal{X} \subseteq \mathcal{C}$ having $v(\mathcal{X}) \geq V$ and  $\sum_{j \in \mathcal{X}} g_j \leq H$.
+\else
+`인스턴스'는 엘리스 문제의 인스턴스와 목적 포트폴리오 가치 $V$로 구성된다. 인스턴스가 `예-인스턴스'가 될 필수충분 조건은 $v(\mathcal{X}) \geq V$와 $\sum_{j \in \mathcal{X}} g_j \leq H$를 만족하는 포트폴리오 $\mathcal{X} \subseteq \mathcal{C}$가 존재하는 것이다.
+\fi
 \end{problem}
@@
 \begin{theorem}
+\ifen
 The decision form of Ellis’s problem is NP-complete.
+\else
+엘리스 문제의 결정 형태는 NP-complete다.
+\fi
 \end{theorem}
 
-\begin{proof}It is obvious that the problem is in NP.
-
-Consider an instance of the knapsack problem, and we will construct an instance of Ellis’s problem that is a yes-instance if and only if the corresponding knapsack instance is a yes-instance.
-Without loss of generality, we may assume that the the objects in $\mathcal{B}$ are indexed in increasing order of $u_j$, that each $u_j >0$, and that the knapsack instance admits a feasible solution other than the empty set.
-
-Let $U_{\mathrm{max}} = \sum_{j \in \mathcal{B}} u_j$ and $\delta = {1} /{m U_{\mathrm{max}}} > 0$, and construct an instance of Ellis’s problem with $\mathcal{C} = \mathcal{B}$, $H = W$, all $f_j = \delta$, and $t_j = u_j / \delta$ for all $j$.
+\begin{proof}
+\ifen It is obvious that the problem is in NP.
+\else 문제가 NP에 속하는 것은 명백하다.\fi
+
+\ifen
+Consider an instance of the knapsack problem, and we will construct an instance of Problem \ref{ellisdecisionform} that is a yes-instance if and only if the corresponding knapsack instance is a yes-instance.
+Without loss of generality, we may assume that the the objects in $\mathcal{B}$ are indexed in increasing order of $u_j$, that each $u_j >0$, and that each $w_j \leq W$.
+\else
+임의의 배낭 문제 인스턴스를 택하고, 예-인스턴스 여부가 동치인 문제 \ref{ellisdecisionform}의 인스턴스를 만들자. 일반성을 잃지 않고 $\mathcal{B}$에 속한 상품이 $u_j$가 증가하는 순서대로 배열되며 모든 $u_j >0$이고 $w_j \leq W$임을 가정할 수 있다.
+\fi
+
+\ifen
+Let $U_{\mathrm{max}} = \sum_{j \in \mathcal{B}} u_j$ and $\delta = {1} /{m U_{\mathrm{max}}} > 0$, and construct an instance of Ellis’s problem with $\mathcal{C} = \mathcal{B}$, $H = W$, all $f_j = \delta$, and each $t_j = u_j / \delta$.
 Clearly, $\mathcal{X} \subseteq \mathcal{C}$ is feasible for Ellis’s problem if and only if it is feasible for the knapsack instance.
 Now, we observe that for any nonempty $\mathcal{X}$,
+\else
+$U_{\mathrm{max}} = \sum_{j \in \mathcal{B}} u_j$와 $\delta = {1} /{m U_{\mathrm{max}}} > 0$일 때, $\mathcal{C} = \mathcal{B}$, $H = W$,  모든 $f_j = \delta$, 그리고 각 $t_j = u_j / \delta$인 엘리스 문제의 인스턴스를 고려하자.
+$\mathcal{X} \subseteq \mathcal{C}$가 엘리스 문제의 가능해가 되는 것은 배낭 문제 인스턴스의 가능해가 되는 것과 동치임을 쉽게 알 수 있다.
+그러면 공집합이 아닌 임의의 $\mathcal{X}$ 에 대해,
+\fi
 \begin{align}
@@
 \end{align}
+\ifen
 This means that the utility of an application portfolio $\mathcal{X}$ in the corresponding knapsack instance is the smallest integer greater than $v( \mathcal{X})$.
@@
 Taking $V = U-1$ completes the transformation and concludes the proof.
+\else
+이는 지원 포트폴리오 $\mathcal{X}$가 해당 배낭 문제에서 일으키는 효용이 $v( \mathcal{X})$보다 크면서 가장 작은 정수임을 의미한다.
+즉, $\sum_{j \in \mathcal{X}} u_j \geq U$가 성립할 필수충분 조건은 $v(\mathcal{X}) \geq U -1$이다.
+$V = U-1$으로 정의해서 변환을 완성하면 증명이 완료된다.
+\fi
 \end{proof}
 
+\ifen 
 An intuitive extension of the greedy algorithm for Alma's problem is to iteratively add to $\mathcal{X}$ the school $k$ for which $[ v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) ] / g_k$ is largest.
 However, the construction above shows that the objective function of Ellis's problem can approximate that of a knapsack problem with arbitrary precision.
-Therefore, in pathological examples such as the following, the greedy algorithm can achieve an arbitrarily poor approximation ratio.
+Therefore, in pathological examples such as the following, the greedy algorithm can achieve an arbitrarily small approximation ratio.
+\else
+알마 문제를 위한 탐욕 해법의 직관적인 확장은 $[ v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) ] / g_k$-값이 가장 높은 학교 $k$를  $\mathcal{X}$에 차례대로 추가하는 것이다.
+그러나 위에서 도출한 변환은 엘리스 문제의 목적함수를 배낭 문제의 목적함수의 원하는 만큼 정확한 근사로 만들 수 있음을 의미한다.
+따라서 다음 같은 병례에서 탐욕 해법은 임의로 낮은 근사 비율을 일으킬 수 있다.
+\fi
 \begin{example}
+\ifen
 Let $t = (10, 2021)$, $f = (1, 1)$, $g = (1, 500)$, and $H = 500$. Then the greedy approximation algorithm produces the clearly inoptimal solution $\mathcal{X} = \{1\}$. 
+\else
+ $t = (10, 2021)$, $f = (1, 1)$, $g = (1, 500)$, 그리고 $H = 500$이라고 하자. 그러면 탐욕 해법은 $\mathcal{X} = \{1\}$을 출력하며 이는 분명히 최적해가 아니다.
+\fi
 \end{example} 
@@
 \ifen \subsection{Branch-and-bound algorithm} \else \subsection{분지한계법} \fi
-A traditional approach to knapsack problems is the branch-and-bound framework, which generates subproblems in which the values of one or more decision variables are fixed and uses an upper bound on the objective function to exclude, or \emph{fathom,} branches of the decision tree that cannot yield a solution better than the best solution on hand (Martello and Toth 1990; Kellerer et al.
+\ifen 
+A traditional approach to integer optimization problems is the branch-and-bound framework, which generates subproblems in which the values of one or more decision variables are fixed and uses an upper bound on the objective function to exclude, or \emph{fathom,} branches of the decision tree that cannot yield a solution better than the best solution on hand (Martello and Toth 1990; Kellerer et al.
 2004).
@@
 A branch-and-bound routine emerges naturally from these ingredients.
-
+\else
+정수 최적화의 고전적인 기법 중 분지한계법(branch and bound)이 있다.
+이는 하나 이상의 결정 변수를 고정시키고 해당 부문제를 탐색하는 해법이며, 목적함수의 상한(bound)을 활용해서 기존 해보다 좋은 해를 생성할 수 없는 결정 나무의 가지(branch)를 제외시켜 결정 공간을 줄인다 (Martello and Toth 1990; Kellerer et al.
+2004).
+본 항에서, 엘리스 문제의 정수 모형과 그의 상한이 되는 선형 계획을 제시한다.
+알고리즘 \ref{algorithmforlargeh}에서 개발한 $t_j$-값의 변환을 활용해서 부문제의 선형 완화 문제에 의한 상한을 더 타이트하게 만든다.
+이 기반으로 분지한계법 알고리즘을 도출한다.
+\fi
+
+\ifen
 To begin, let us characterize the portfolio $\mathcal{X}$ as the binary vector $x \in \{0, 1\}^m$, where $x_j = 1 $ if and only if $j \in \mathcal\{X\}$.
 Then it is not difficult to see that Ellis's problem is equivalent to the following integer nonlinear program.
-\begin{problem}[Integer NLP for Ellis's problem] \label{integernlp}
+\else
+우선 포트폴리오 $\mathcal{X}$를 이진 특성 벡터 $x \in \{0, 1\}^m$로 표현하자. 단, $x_j = 1 $인 것은 $j \in \mathcal\{X\}$과 동치다. 그러면 엘리스의 문제가 다음 비선형 정수 최적화 문제와 동등함을 볼 수 있다.
+\fi
+\begin{problem}[\ifen Integer NLP for Ellis's problem\else 엘리스의 문제를 위한 비선형 정수 계획\fi] \label{integernlp}
 \begin{align}
@@
 \end{problem}
+\ifen
 Since the product in $v(x)$ does not exceed one, the following LP relaxation is an upper bound on the valuation of the optimal portfolio.
-\begin{problem}[LP relaxation for Ellis's problem] \label{LPrelaxation}
+\else
+$v(x)$에서 등장하는 곱은 1을 넘을 수 없으므로 다음 선형 완화 문제는 최적 포트폴리오 가치의 상한이 돈다.
+\fi
+\begin{problem}[\ifen LP relaxation for Ellis's problem\else 엘리스의 문제를 위한 선형 완화 문제\fi] \label{LPrelaxation}
 \begin{align}
@@
 \end{problem}
-Problem \ref{LPrelaxation} is a continuous knapsack problem, which is easily solved in $O(m \log m)$ time by the following greedy algorithm: Start with $x = \mathbf{0}$.
-While $H > 0$, select the school $k$ for which $f_j t_j / g_j$ is highest, set $x_k \gets \min\{1, H / g_k\}$, and set $H \gets H - g_k$ (Dantzig 1957).
-
+\ifen
+Problem \ref{LPrelaxation} is a continuous knapsack problem, which is easily solved in $O(m \log m)$-time by the following greedy algorithm: Start with $x = \mathbf{0}$.
+While $H > 0$, select the school $j$ for which $f_j t_j / g_j$ is highest, set $x_j \gets \min\{1, H / g_j\}$, and set $H \gets H - g_j$ (Dantzig 1957).
+\else
+문제 \ref{LPrelaxation}은(는) ``연속 배낭 문제''라고 불리며 다음 탐욕 알고리즘을 적용하면 $O(m \log m)$-시간 안에 쉽게 풀 수 있다: $x = \mathbf{0}$로 초기화한다.
+$H > 0$가 성립하는 동안, $f_j t_j / g_j$-값이 가장 높은 학교 $j$를 택한다.
+$x_j \gets \min\{1, H / g_j\}$ 그리고 $H \gets H - g_j$ 같이 수정해서 반복한다 (Dantzig 1957).
+\fi
+
+\ifen 
 In our branch-and-bound framework, a \emph{node} is characterized by a three-way partition of schools $\mathcal{C}= \mathcal{I} \cup \mathcal{O} \cup \mathcal{N}$ satisfying $\sum_{j \in \mathcal{I}} g_j \leq H$.
-$\mathcal{I}$ consists of schools that are ``in'' the application portfolio, $\mathcal{O}$ consists of those that are ``out,'' and $\mathcal{N}$ consists of those that are ``negotiable.''
+$\mathcal{I}$ consists of schools that are ``in'' the application portfolio ($x_j = 1$), $\mathcal{O}$ consists of those that are ``out'' ($x_j = 0$), and $\mathcal{N}$ consists of those that are ``negotiable.''
 %This partition induces a constant $\gamma$, transformed $t_j$-values $\bar t$, the residual budget $\mathcal{H}$, and a set $\mathcal{L}$ of pointers to the node's children, if any.
 The choice of partition induces a pair of subproblems. The first subproblem is an instance of Problem \ref{integernlp}, namely
+\else
+본 연구의 분지한계법 설계에서, `마디'(node)는 학교를 세 군으로 나눈 분할 $\mathcal{C}= \mathcal{I} \cup \mathcal{O} \cup \mathcal{N}$로 묘사한다.
+단, $\sum_{j \in \mathcal{I}} g_j \leq H$을 만족한다.
+$\mathcal{I}$는 지원 포트폴리오에 속한 학교(`in': $x_j = 1$), $\mathcal{O}$는 속하지 않은 학교(`out': $x_j = 0$), 그리고 $\mathcal{N}$은 미결의 학교(`negotiable')로 이루어진다.
+분할이 결정되면 부문제의 한 쌍이 결정해진다.
+첫번째 부문제는 다음 같은 문제 \ref{integernlp}의 인스턴스다.
+\fi
 \begin{align} \label{nodenlpsubproblem}
@@
 \end{align}
-The second is the corresponding instance of Problem \ref{LPrelaxation}:
+\ifen The second is the corresponding instance of Problem \ref{LPrelaxation}:
+\else 두번째 부문제는 대응하는 문제 \ref{LPrelaxation} 인스턴스다.\fi
 \begin{align} \label{nodelprelaxation}
@@
 \end{align}
+\ifen
 In both subproblems, $\bar H = H - \sum_{j\in \mathcal{I}} g_j$ denotes the residual budget.
@@
 For each $j\in \mathcal{I}$,  we increment $\gamma$ by the current value of $f_j \bar t_j$, eliminate $j$ from the market, and update the remaining $\bar t_j$-values using \eqref{howtotransformfj}.
-
+(An example is provided below.)
+\else
+각 부문제에서 $\bar H = H - \sum_{j\in \mathcal{I}} g_j$는 잔여 예산을 의미한다.
+$\mathcal{I}$에 속한 학교에 \eqref{howtotransformfj}의 변환을 적용해서 모수 $\gamma$와 $\bar t$를 얻을 수 있다.
+즉, 각 $j\in \mathcal{I}$에 대해 $\gamma$를 $f_j \bar t_j$ 만큼 증가시키고, 시장에서 $j$를 소거하고, \eqref{howtotransformfj}에 따라 남은 $\bar t_j$-값을 수정한다.
+(아래에서 예를 제시한다.)
+\fi
+
+\ifen 
 Given a node $n  = (\mathcal{I}, \mathcal{O}, \mathcal{N})$, its children are generated as follows.
@@
 Equivalently, we set $x_j = 1$ in one child and $x_j = 0$ in the other.
-In principle, any school in $\mathcal{N}$ can be chosen for $j$, but as a greedy heuristic, we choose the school for which the ratio $f_j \bar t_j / g_j$ is highest.
+In principle, any school can be chosen for $j$, but as a greedy heuristic, we choose the school for which the ratio $f_j \bar t_j / g_j$ is highest.
 Notice that this method of generating children ensures that each node's $\mathcal{I}$-set differs from its parent's by at most a single school, so the constant $\gamma$ and transformed $\bar t_j$-values for the new node can be obtained by a single application of \eqref{howtotransformfj}.
-
+\else
+마디  $n  = (\mathcal{I}, \mathcal{O}, \mathcal{N})$이 주어질 때, 그의 새끼 마디를 다음처럼 생성한다.
+모든 마디에는 2, 1, 혹은 0개의 새끼가 있다.
+전형적인 경우에서, $g_j \leq \bar H$를 만족하는 학교 $j \in \mathcal{N}$ 선택한다.
+$j$를 $\mathcal{I}$로 이동시켜서 한 새끼를 얻고, $j$를 $\mathcal{O}$로 이동시켜서 다른 새끼를 얻는다.
+즉, 한 새끼 마디에서 $x_j = 1$로 고정시키도 다른 새끼 마디에서 $x_j = 0$으로 고정시킨다.
+일반적으로 $j$는 임의의 학교가 될 수 있지만 탐욕적인 휴리스틱으로 $f_j \bar t_j / g_j$의 비율이 가장 높은 학교를 선택한다.
+이런 식으로 새끼를 생성하면 각 마디의 $\mathcal{I}$-집합은 자신의 어미 마디의 $\mathcal{I}$-집합과 최대 한 학교로만 다르다.
+따라서 새로운 마디를 생성할 때, \eqref{howtotransformfj}을(를) 한번만 적용하면 상수 $\gamma$와 변화된 $\bar t_j$-값을 얻을 수 있다.
+\fi
+
+\ifen
 There are two atypical cases.
@@
 Second, if $\mathcal{N} = \O$, then the node has zero children, and as no further branching is possible, the node is called a \emph{leaf.}
+\else
+이례적인 경우 2가지 있다.
+첫번째는, $\mathcal{N}$에 속한 모든 학교에 대해 $g_j > \bar H$이면 가능성을 유지하며 $\mathcal{I}$에 추가할 수 있는 학교가 없으므로 본 가지의 최적 포트폴리오는 단지 $\mathcal{I}$ 자체다.
+이때 $\mathcal{N}$에 원소를 모두 $\mathcal{O}$로 이동시켜서 단일 새끼를 생성한다.
+두번째는, $\mathcal{N} = \O$인 마디는 새끼가 없다.
+더 이상 분지할 수 없으므로 이 마디를 `잎 마디'(leaf)라고 한다.
+\fi
 
 \begin{example}
-Consider a market in which $t = (20, 40, 60, 80, 100)$, $f = (0.5, 0.5, 0.5, 0.5, 0.5)$, $g = (3, 2, 3, 2, 3)$, and $h = 8$, and the node $n =  (\mathcal{I}, \mathcal{O}, \mathcal{N}) = (\{2, 5\}, \{1\}, \{3, 4\})$.
-Let us compute the two subproblems associated with $n$ and generate its children.
-To compute the subproblems, we first simple disregard school 1.
-Next, to eliminate school 2, we apply \eqref{howtotransformfj} to obtain
+\ifen
+Consider a market in which $t = (20, 40, 60, 80, 100)$, $f = (0.5, 0.5, 0.5, 0.5, 0.5)$, $g = (3, 2, 3, 2, 3)$, and $H = 8$, and the node $n =  (\mathcal{I}, \mathcal{O}, \mathcal{N}) = (\{2, 5\}, \{1\}, \{3, 4\})$.
+Let us compute the two subproblems associated with $n$ and identify its children.
+To compute the subproblems, we first simply disregard $c_1$.
+Next, to eliminate $c_2$, we apply \eqref{howtotransformfj} to $t$ to obtain
 $\{\bar{\bar t}_3, \bar{\bar t}_4, \bar{\bar t}_5\} = \{ (1 - f_2) t_3, (1 - f_2) t_4, (1 -f_2) t_5\} = \{30, 40, 50 \}$ and $\bar{\bar \gamma} = f_2 t_2 = 20$.
-We eliminate school 5 by applying \eqref{howtotransformfj} to $\bar{ \bar t}$ to obtain $\{\bar t_3, \bar t_4\} =  \{\bar{\bar t}_3 - f_5 \bar{\bar t}_5, \bar{\bar t}_4-  f_5 \bar{\bar t}_5\}= \{5, 15\}$ and $\gamma = \bar{\bar \gamma} + f_5 \bar{\bar t}_5 = 35$.
+We eliminate $c_5$ by again applying \eqref{howtotransformfj} to $\bar{ \bar t}$ to obtain $\{\bar t_3, \bar t_4\} =  \{\bar{\bar t}_3 - f_5 \bar{\bar t}_5, \bar{\bar t}_4-  f_5 \bar{\bar t}_5\}= \{5, 15\}$ and $\gamma = \bar{\bar \gamma} + f_5 \bar{\bar t}_5 = 35$.
 Finally, $\bar H = H - g_2 - g_5 = 3$.
-Problems \ref{nodenlpsubproblem} and \ref{nodelprelaxation} are now easily obtained by substitution.
-
-Since all the schools in $\mathcal{N}$ have $g_j \leq \bar H$, $n$ has two children.
-Applying the node-generation rule, school 4 has the highest $f_j \bar t_j / g_j$-ratio, so the children are $n_1 = (\{2, 4, 5\}, \{1\}, \{3\})$ and $n_2 = (\{2, 5\}, \{1, 4\}, \{3\})$.
+Now problems \eqref{nodenlpsubproblem} and \eqref{nodelprelaxation} are easily obtained by substitution.
+\else
+ $t = (20, 40, 60, 80, 100)$, $f = (0.5, 0.5, 0.5, 0.5, 0.5)$, $g = (3, 2, 3, 2, 3)$, 그리고 $H = 8$인 시장과 마디 $n =  (\mathcal{I}, \mathcal{O}, \mathcal{N}) = (\{2, 5\}, \{1\}, \{3, 4\})$이 주어질 때 $n$에 대응하는 부문제와 새끼 마디를 구하자.
+부문제를 계산하기 위해 먼저 $c_1$을 단순히 무시한다.
+$c_2$를 소거하기 위해 $t$에 \eqref{howtotransformfj}에 적용해서
+$\{\bar{\bar t}_3, \bar{\bar t}_4, \bar{\bar t}_5\} = \{ (1 - f_2) t_3, (1 - f_2) t_4, (1 -f_2) t_5\} = \{30, 40, 50 \}$과 $\bar{\bar \gamma} = f_2 t_2 = 20$을 얻는다.
+$c_5$를 소거하기 위해 $\bar{\bar t}$에 \eqref{howtotransformfj}에 다시 적용해서 $\{\bar t_3, \bar t_4\} =  \{\bar{\bar t}_3 - f_5 \bar{\bar t}_5, \bar{\bar t}_4-  f_5 \bar{\bar t}_5\}= \{5, 15\}$와 $\gamma = \bar{\bar \gamma} + f_5 \bar{\bar t}_5 = 35$를 얻는다.
+마지막으로, $\bar H = H - g_2 - g_5 = 3$이다.
+이제 \eqref{nodenlpsubproblem}과(와) \eqref{nodelprelaxation}인 부문제를 쉽게 구할 수 있다.
+\fi
+
+\ifen
+Since at least one of the schools in $\mathcal{N}$ has $g_j \leq \bar H$, $n$ has two children.
+Applying the node-generation rule, $c_4$ has the highest $f_j \bar t_j / g_j$-ratio, so the children are $n_1 = (\{2, 4, 5\}, \{1\}, \{3\})$ and $n_2 = (\{2, 5\}, \{1, 4\}, \{3\})$.
+\else
+$\mathcal{N}$에 속한 학교 중에 $g_j \leq \bar H$인 학교가 있으므로 $n$은 새끼 마디 2개 있다.
+$f_j \bar t_j / g_j$-비율이 가장 높은 학교는 $c_4$이므로 새끼 마디를 생성하는 법칙을 적용하면 $n_1 = (\{2, 4, 5\}, \{1\}, \{3\})$과 $n_2 = (\{2, 5\}, \{1, 4\}, \{3\})$ 같은 새끼를 얻는다.
+\fi
 \end{example}
 
+\ifen
 To implement the branch-and-bound algorithm, we represent the set of candidate nodes---namely, nonleaf nodes whose children have not yet been generated---by $\mathfrak{T}$.
@@
 Because $\mathcal{I}$ is a feasible portfolio, $v_{\mathcal{I}}[n]$ is a lower bound on the optimal objective value.
-Moreover, by the argument given in the proof of Theorem \ref{nestedapplication}, the objective function of  \eqref{nodenlpsubproblem} is identical to the function $v(\mathcal{I} \cup \mathcal{Y})$, where $\mathcal{Y} = \{ j: y_j = 1\}$.
-This means that the optimal objective value $v_{\mathrm{LP}}^*$  is an upper bound on the valuation of any portfolio that contains $\mathcal{I}$ as a subset and does not include any school in $\mathcal{O}$, and hence on the valuation of any portfolio on this branch.
-Therefore, if upon generating a new node $n_2$, we discover that its objective value $v_{\mathcal{I}}[n_2]$ is greater than $v_{\mathrm{LP}}^*[n_1]$ for some other node $n_1$, then $n_1$ and all its descendants can be eliminated from consideration.
-
+Moreover, by the argument given in the proof of Theorem \ref{nestedapplication}, the objective function of  \eqref{nodenlpsubproblem} is identical to the function $v(\mathcal{I} \cup \mathcal{X})$.
+This means that $v_{\mathrm{LP}}^*[n]$  is an upper bound on the valuation of any portfolio that contains $\mathcal{I}$ as a subset and does not include any school in $\mathcal{O}$, and hence on the valuation of any portfolio on this branch.
+Therefore, if upon generating a new node $n_2$, we discover that its objective value $v_{\mathcal{I}}[n_2]$ is greater than $v_{\mathrm{LP}}^*[n_1]$ for some other node $n_1$, then we can disregard $n_1$ and all its descendants.
+\else
+분지한계법을 실행하기 위해 후보 마디, 즉 잎이 아니면서 새끼를 아직 생성하지 않은 마디의 집합을  $\mathfrak{T}$라고 부르자.
+마디 $n= (\mathcal{I}, \mathcal{O}, \mathcal{N})$를 생성할 때마다, $v_{\mathcal{I}}[n] = v(\mathcal{I})$과 선형 완화 문제 \eqref{nodelprelaxation}의 최대값 $v^*_{\mathrm{LP}}[n]$를 기록한다.
+$\mathcal{I}$는 예산에 가능한 포트폴리오이므로 $v_{\mathcal{I}}[n]$는 원래 목적함수 최대값의 하한이 된다.
+또한 정리 \ref{nestedapplication}의 증명 과정에서 보였듯, \eqref{nodenlpsubproblem}의 목적함수는 함수 $v(\mathcal{I} \cup \mathcal{X})$과 동치다.
+이는 $v_{\mathrm{LP}}^*[n]$는 $\mathcal{I}$를 부분집합으로 포함하며 $\mathcal{O}$에 속한 학교를 포함하지 않은, 즉 이 가지에 있는 모든 포트폴리오의 가치에 대한 상한임을 의미한다.
+따라서 새로운 마디 $n_2$를 생성했을 때, 만약 그의 목적함수값 $v_{\mathcal{I}}[n_2]$가 어떤 다른 마디 $n_1$에 대응하는 $v_{\mathrm{LP}}^*[n_1]$보다 크다면, $n_1$과 그의 모든 후손을 무시할 수 있다.
+\fi
+
+\ifen
 The algorithm is initialized by populating $\mathfrak{T}$ with the root node $n_0 = (\O, \O, \mathcal{C})$.
-At each iteration, it selects the node $n \in \mathfrak{T}$ having the highest $v_{\mathrm{LP}}^*$-value, generates its children, and removes $n$ from the candidate set.
+At each iteration, it selects the node $n \in \mathfrak{T}$ having the highest $v_{\mathrm{LP}}^*[n]$-value, generates its children, and removes $n$ from the candidate set.
 Next, the children $n'$ of $n$ are inspected; if one of them yields a new optimal solution, then we mark it as the best candidate and fathom any nodes $n''$ for which  $v_{\mathcal{I}}[n'] >  v_{\mathrm{LP}}^*[n'']$.
-When no nodes remain in the candidate set, the algorithm has explored every branch, so it terminates and returns the best candidate.
+When no nodes remain in the candidate set, the algorithm has explored every branch, so it returns the best candidate and terminates.
+\else
+$\mathfrak{T}$에 뿌리 마디 $n_0 = (\O, \O, \mathcal{C})$를 넣어서 알고리즘을 초기화한다.
+각 반복단계에서 $v_{\mathrm{LP}}^*[n]$-값이 가장 높은 마디 $n \in \mathfrak{T}$을 선택한다.
+$n$의 새끼를 생성하고 후보 집합에서 제거한다.
+그다음에 $n$의 각 새끼 마디 $n'$을 검증한다.
+그중에 새로운 최적해를 발견하면 이를 가장 좋은 후보로 표하고  $v_{\mathcal{I}}[n'] >  v_{\mathrm{LP}}^*[n'']$인 마디 $n''$을 후보 집합에서 제거한다.
+후보 집합이 공집합이 되면 알고리즘이 모든 가지를 탐색했으므로 가장 좋은 후보를 출력하고 종료한다.
+\fi
 
@@
 
-\ifen \subsection{Fully polynomial-time approximation scheme}\else \subsection{전체 다항 시간 근사 해법}\fi \label{fptashead}
+\ifen \subsection{Fully polynomial-time approximation scheme}\else \subsection{완전 다항 시간 근사 해법}\fi \label{fptashead}
 As with the knapsack problem, Ellis's problem admits a complementary dynamic program that iterates on the value of the cheapest portfolio instead of on the cost of the most valuable portfolio.
@@
 We will represent approximate portfolio valuations using a fixed-point decimal with a precision of $P$, where $P$ is the number of digits to retain after the decimal point.
-Let $r[x] = \lfloor 10^P x \rfloor 10^{-P}$ denote the value of $x$ rounded down to its nearest fixed-point representation.
+Let $r[x] =  10^{-P}\lfloor 10^P x \rfloor$ denote the value of $x$ rounded down to its nearest fixed-point representation.
 Since $\bar U = \sum_{j\in \mathcal{C}} f_j t_j$ is an upper bound on the valuation of any portfolio, and since we will ensure that each fixed-point approximation is an underestimate of the portfolio's true valuation, the set $\mathcal{V}$ of possible valuations possible in the fixed-point framework is finite:
@@
 \begin{table}[h!] \centering
-\begin{tabular}{r|rr|rr}
+\begin{tabular}{r|r@{~}r|r@{~}r}
 \ifen 
@@
 %4096       & 4826.75\phantom{00000}          &    (1432.85\phantom{000000})              & 7304.67\phantom{00000}      &     (2071.38\phantom{000000})
-16         & 0.07          &    (0.01)              & 0.01      &     (0.00)            \\
-64         & 0.95          &    (0.08)              & 1.29      &     (0.05)            \\
-256        & 14.86         &    (0.73)              & 20.35     &     (0.71)            \\
-1024       & 287.06          &    (81.92)              & 390.93    &     (110.50)            \\
-4096       & 4826.75         &    (1432.85)              & 7304.67     &     (2071.38)                                     
+    16 &    0.07 &    (0.01) &     0.09 &    (0.02) \\
+    64 &    0.88 &    (0.16) &     1.18 &    (0.21) \\
+   256 &   12.51 &    (2.20) &    18.39 &    (2.90) \\
+  1024 &  281.79 &   (58.56) &   387.97 &   (96.41) \\
+  4096 & 4661.75 & (1152.54) &  7218.04 & (1733.91)                                  
 \end{tabular}
@@
 \begin{table}[h!] \centering
-\begin{tabular}{r|rr|rr|rr|rr}
-\ifen
-\textbf{\begin{tabular}[r]{@{}r@{}}Number of \\schools $m$\end{tabular}}&\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisbnb}:\\Branch \& bound\end{tabular}}}  & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP1}:\\App.
+\begin{tabular}{r|r@{~}r|r@{~}r|r@{~}r|r@{~}r}
+\ifen
+\textbf{\begin{tabular}[r]{@{}r@{}}Number of\\schools $m$\end{tabular}}&\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisbnb}:\\Branch \& bound\end{tabular}}}  & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP1}:\\App.
 costs DP\end{tabular}}}  &\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.5$\end{tabular}}}  & \multicolumn{2}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.05$\end{tabular}}}   \\ \hline
 \else
-\textbf{\begin{tabular}[r]{@{}r@{}}학교의\\개수 $m$\end{tabular}}&\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘  \ref{ellisbnb}:\\분지한계법\end{tabular}}}  & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP1}:\\지원비 동적 계획\end{tabular}}}  &\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.5$\end{tabular}}}  & \multicolumn{2}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.05$\end{tabular}}}   \\ \hline
-\fi
-    5 &         0.05 &     (0.02) &       0.01 &      (0.00) &    0.09 &  (0.01) &     0.26 &         (0.08) \\
-   10 &            — &        (—) &       0.13 &      (0.03) &    0.78 &  (0.07) &     2.89 &         (0.29)            
+\textbf{\begin{tabular}[r]{@{}r@{}}학교의\\개수 $m$\end{tabular}}&\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘  \ref{ellisbnb}:\\분지한계법\end{tabular}}}  & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP1}: 지원\\비용 동적 계획\end{tabular}}}  &\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.5$\end{tabular}}}  & \multicolumn{2}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.05$\end{tabular}}}   \\ \hline
+\fi
+   8 &          0.44 &       (0.42) &         0.06 &      (0.02) &                0.39 &             (0.16) &                 1.66 &              (0.64) \\
+  16 &       3298.08 &    (5285.45) &         0.48 &      (0.16) &                2.64 &             (0.96) &                19.20 &              (7.05) \\
+  32 &             — &          (—) &         2.30 &      (0.83) &               19.01 &             (9.18) &               184.00 &             (70.15) \\
+  64 &             — &          (—) &         8.33 &      (3.52) &              101.05 &            (59.47) &              2089.30 &            (961.43) \\
+ 128 &             — &          (—) &        36.61 &     (15.36) &              481.61 &           (210.68) &              8626.29 &           (2615.38) \\
+ 256 &             — &          (—) &       205.57 &     (52.98) &             3106.71 &           (957.95) &             69291.93 &          (34881.04) \\
+ 512 &             — &          (—) &      1136.18 &    (396.24) &            16613.99 &          (6241.99) &            206020.31 &          (45217.42) 
 \end{tabular}%
@@
 The branch-and-bound algorithm is impractical for large markets. \lastptofcaption
-\else  %분지한계법 (알고리즘 \ref{ellisbnb}), 정확한 동적 계획 (알고리즘 \ref{ellisDP1}), 그리고 전체 다항 시간 근사 해법 (알고리즘 \ref{ellisDP3})을
+\else  %분지한계법 (알고리즘 \ref{ellisbnb}), 정확한 동적 계획 (알고리즘 \ref{ellisDP1}), 그리고 완전 다항 시간 근사 해법 (알고리즘 \ref{ellisDP3})을
  \ref{hetappcosts}절에서 도출한 3개의 알고리즘을 사용할 때, 다양한 지원 비용으로 갖춘 입학 시장의 최적 또는 $(1- \varepsilon)$-최적 포트폴리오를 계산하는 시간. 분지한계법은 큰 시장에서 비실용적이다. \lastptofcaption \fi}
@@
 The optimal portfolio was computed using Algorithm \ref{ellisDP1}.
- \else $m=60$개의 학교로 구성된 전형적인 무작위로 생성한 인스턴스와 해당 최적 포트폴리오. 지원비 $g_j$는 $\{5, \dots, 10\}$에서 균일한 확률로 선정했으며 기호의 넓이와 비례된다. 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
+ \else $m=60$개의 학교로 구성된 전형적인 무작위로 생성한 인스턴스와 해당 최적 포트폴리오. 지원 비용 $g_j$는 $\{5, \dots, 10\}$에서 균일한 확률로 선정했으며 기호의 넓이와 비례된다. 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
 \end{figure}
@@
 \ifen \subsection{Implementation notes} \else \subsection{알고리즘의 이행}\fi
-We chose to implement our algorithms in the Julia language (v1.7.0) because its system of type inference allows the compiler to optimize for differential cases such as when the $t_j$-values are integers or floating-point numbers, and because Julia offers convenient macros for parallel computing (Bezanson et al.
+We chose to implement our algorithms in the Julia language (v1.7.0) because its system of type inference allows the compiler to optimize for differential cases such as when the $t_j$-values are integers or floating-point numbers.
+Julia also offers convenient macros for parallel computing, which enabled us to solve more and larger problems in the benchmark (Bezanson et al.
 2017).

### 2022-03-10 benchmark/benchmark.jl

@@ -7,7 +7,7 @@ using Statistics
 using Base.Threads
 import Printf: @sprintf
 
-fullscale = false
+fullscale = true
 
 # A long benchmark; tweak parameters with caution.
 # Set fullscale = false to run a smaller benchmark to check formatting etc.
@@ -16,7 +16,7 @@ fullscale = false
 n_reps = fullscale ? 3 : 2
 
 # Number of markets to test at each intersection of the experimental variables
-n_markets = fullscale ? 10 : 2
+n_markets = fullscale ? 20 : 2
 
 function printheader(s)
     printstyled(s * "\n", bold = true, color = 222)

### 2022-03-11 paper/paper.tex
translations in first dynamic program

@@
 \caption{Optimal portfolio algorithm for Alma’s problem.} \label{algorithmforlargeh}
-\KwData{Utility values $t \in[0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application limit $h \leq m$.}
+\KwData{Utility values $t \in(0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application limit $h \leq m$.}
 %Index schools in ascending order by $t$\;
@@
 \caption{알마의 문제를 위한 최적 포트폴리오 알고리즘.} \label{algorithmforlargeh}
-\KwData{효용 모수 $t \in[0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 제한 $h \leq m$.}
+\KwData{효용 모수 $t \in(0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 제한 $h \leq m$.}
 %$t$의 순서대로 학교의 지표를 짓는다\;
@@
 
+\ifen {
 \begin{algorithm}[h] 
@@
 Current lower bound $L \gets 0$ and best solution $\mathcal{X} \gets \O$\;
-Initialize candidate set $\mathfrak{T} \gets \{ n_0\}$\;
+Candidate set $\mathfrak{T} \gets \{ n_0\}$\;
 \While{not finished}{
@@
 		}
-		\For{$n''$ in $\mathfrak{T}$} {
+		\For{$n'' \in \mathfrak{T}$} {
 			\lIf{$L >  v_{\mathrm{LP}}^*[n'']$}{remove $n''$ from $\mathfrak{T}$}
@@
 \end{algorithm}
-
-\begin{theorem}[Validity of Algorithm \ref{ellisbnb}]
-Algorithm \ref{ellisbnb} produces an optimal application portfolio for Ellis's problem.
+} \else {
+\begin{algorithm}[h] 
+%\DontPrintSemicolon
+\caption{엘리스의 문제를 위한 분지한계법.} \label{ellisbnb}
+\KwData{효용 모수 $t \in(0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 비용 $g \in (0, \infty)^m$, 예산 $H \in (0, \infty)$.}
+뿌리 마디 $n_0 \gets (\O, \O, \mathcal{C})$\;
+현재 하한 $L \gets 0$ 및 최적해 $\mathcal{X} \gets \O$\;
+후보 집합 $\mathfrak{T} \gets \{ n_0\}$\;
+\While{종료되지 않음}{
+	\lIf{$\mathfrak{T} = \O$}{\Return{$\mathcal{X}, L$}}
+	\Else{
+		$n \gets \argmax\{ v_{\mathrm{LP}}^*[n] : n \in \mathfrak{T}\}$\;
+		$\mathfrak{T}$에서 $n$을 제거한다\;
+%		Generate children of $n$\;
+		\For{$n$의 각 새끼 마디 $n'$}{
+			\If{$L < v_{\mathcal{I}}[n']$}{
+				$L \gets v_{\mathcal{I}}[n']$\;
+				$\mathcal{X}$을 $n'$에 대응하는 $\mathcal{I}$-집합으로 수정한다\;
+			}
+			\lIf{$n'$이 잎 마디 아님}{$\mathfrak{T}$에 $n'$을 추가한다}
+		}
+		\For{$n'' \in \mathfrak{T}$} {
+			\lIf{$L >  v_{\mathrm{LP}}^*[n'']$}{$\mathfrak{T}$에서 $n''$을 제거한다}
+		}
+	}
+}
+\end{algorithm}
+} \fi
+
+\begin{theorem}[\ifen Validity of Algorithm \ref{ellisbnb}\else 알고리즘 \ref{ellisbnb}의 타당성\fi]
+\ifen Algorithm \ref{ellisbnb} produces an optimal application portfolio for Ellis's problem.
+\else 알고리즘 \ref{ellisbnb}은(는) 엘리스 문제의 최적 지원 포트폴리오를 출력한다.
 \end{theorem}
@@
 \begin{proof}
-The discussion above implies that the as if the algorithm terminates, then an optimal solution exists among the leaves of the tree, and it is therefore returned.
-
-Since the set of possible nodes is finite, to show that the algorithm does not cycle, it suffices to show that no node is generated twice.
+\ifen
+Because an optimal solution exists among the leaves of the tree, the discussion above implies that as long as the algorithm terminates, it returns an optimal solution.
+
+To show that the algorithm does not cycle, it suffices to show that no node is generated twice.
 Suppose not: that two distinct nodes $n_1$ and $n_2$ share the same partition $(\mathcal{I}_{12}, \mathcal{O}_{12}, \mathcal{N}_{12})$.
@@
 Therefore, $j \in \mathcal{I}_{12} \cap \mathcal{O}_{12}$, meaning that $(\mathcal{I}_{12}, \mathcal{O}_{12}, \mathcal{N}_{12})$ is not a partition of $\mathcal{C}$, a contradiction.
+\else
+나무의 잎 마디 중 최적해가 반드시 존재하므로 위의 논의에 따라 알고리즘이 종료한다면 최적해를 출력하는 것을 알 수 있다.
+
+알고리즘에서 회로가 생기지 않음을 보이기 위해 어떤 한 마디가 두번 생성되지 않는 것을 증명하면 충분하다.
+모순을 보이기 위해 같은 분할 $(\mathcal{I}_{12}, \mathcal{O}_{12}, \mathcal{N}_{12})$를 가지는 상이한 마디 $n_1$과 $n_2$가 생성된다고 하자.
+나무에 올라가면서 두 마디의 가계가 서로 만나는 첫 마디를 $n$이라고 하자.
+$n$은 새끼 마디 하나만 가지면 그 자체가  $n_1$과 $n_2$의 공통 선조가 되며 이는 모순이므로 $n$ 새끼의 개수가 2임을 알 수 있다.
+또한 그 중 하나는 $n_1$의 선조며 다른 하나는 $n_2$의 선조다.
+각각 $n_3 = (\mathcal{I}_{3}, \mathcal{O}_{3}, \mathcal{N}_{3})$ 그리고 $n_4 = (\mathcal{I}_{4}, \mathcal{O}_{4}, \mathcal{N}_{4})$라고 하자.
+마디 생성 규칙에 따라 $\mathcal{I}_3 \cap \mathcal{O}_4$에 속한 학교 $j$가 존재하며,  $\mathcal{I}_3$ ($\mathcal{O}_4$)의 모든 후손 마디의 $\mathcal{I}$-집합 ($\mathcal{O}$-집합)은  $\mathcal{I}_3$ ($\mathcal{O}_4$)의 확대 집합이다.
+따라서 $j \in \mathcal{I}_{12} \cap \mathcal{O}_{12}$가 되며 이는 $(\mathcal{I}_{12}, \mathcal{O}_{12}, \mathcal{N}_{12})$가 $\mathcal{C}$의 분할이 아님을 의미한다.
+모순.
+\fi
 \end{proof}
 
+\ifen
 The branch-and-bound algorithm is an interesting benchmark, but as a kind of enumeration algorithm, its computation time grows rapidly in the problem size, and unlike the approximation scheme we propose later on, there is no guaranteed bound on the approximation error after a fixed number of iterations.
@@
 In our numerical experiments, Algorithm \ref{ellisbnb} was ineffective on instances larger than about $m=20$ schools, and therefore it does not represent a significant improvement over na\"ive enumeration.
+\else
+분지한계법은 기술적으로 유리한 기준점이지만, 일종의 열거 해법이므로 계산 시간이 문제 크기와 지수적으로 늘어난다.
+그리고 뒤에서 제안하는 근사 해법과 달리, 주어진 반복한계의 개수에 대한 정확성 보장이 없다.
+또한 $\mathcal{N}$에 학교가 많을 때, 선형 완화 문제 상한은 $v_{\mathcal{I}}[n']$보다 매우 높을 수도 있다.
+이는 해법이 나무에 어느 정도 깊게 파고들어야 마디를 제외시킬 수 있음을 의미하며, 그때 계산 노력의 대부분은 이미 쏟은 것이다.
+수리 실험에서, 알고리즘 \ref{ellisbnb}은(는) 약 $m=20$개의 학교를 넘는 인스턴스에 대해 비효율적이었으며 이는 단순한 열거법에 비해 큰 개선이 아니다.
+\fi
 
 \ifen \subsection{Pseudopolynomial-time dynamic program} \else\subsection{의사 다항 시간 동적 계획} \fi
-In this subsection, we assume, with a small loss of generality, that $g_j \in \mathbb{N}$ for $j = 1\dots m$ and $H \in\mathbb{N}$, and provide an algorithmic solution to Ellis's problem that runs in $O(Hm + m\log m)$ time and $O(Hm)$ space.
+\ifen
+In this subsection, we assume, with a small loss of generality, that $g_j \in \mathbb{N}$ for $j = 1\dots m$ and $H \in\mathbb{N}$, and provide an algorithmic solution to Ellis's problem that runs in $O(Hm + m\log m)$-time.
 The algorithm resembles a familiar dynamic programming algorithm for the binary knapsack problem (Dantzig 1957; \emph{Wikipedia}, s.v.
 ``Knapsack problem'').
-Because we cannot assume that $H \leq m$ (as was the case in Alma's problem), this represents a pseudopolynomial-time solution (Garey and Johnson 1979).
-
+Because we cannot assume that $H \leq m$ (as was the case in Alma's problem), this represents a pseudopolynomial-time solution (Garey and Johnson 1979, \S\,4.2).
+\else
+본 절에서 일반성을 약간 제한해서 $g_j \in \mathbb{N}$ for $j = 1\dots m$과 $H \in\mathbb{N}$임을 가정한다.
+이때 엘리스의 문제를 위한 $O(Hm + m\log m)$-시간 해법을 제시한다.
+해법은 이진 배낭 문제를 위한 익숙한 동적 계획 해법과 비슷하다 (Dantzig 1957; \emph{Wikipedia}, s.v.
+``Knapsack problem'').
+알마의 문제와 달리, 여기에서 $H \leq m$임을 가정할 수 없으므로 이는 의사 다항 시간 해법이라고 한다 (Garey and Johnson 1979, \S\,4.2).
+\fi
+
+\ifen 
 For $j = 0 \dots m$ and $h = 0 \dots H$, let $\mathcal{X}[j, h]$ denote the optimal portfolio using only the schools $\{ 1, \dots, j\}$ and costing no more than $h$, and let $V[j,h] = v(\mathcal{X}[j, h])$.
@@
 For convenience, we also define $V[j, h] = -\infty$ for all $h < 0$.
-
+\else
+$j = 0 \dots m$와 $h = 0 \dots H$에 대해, $\{ 1, \dots, j\}$에 속한 학교만 사용하면서 최대한 $h$의 지원 비용을 지불하는 최적 포트폴리오를 $\mathcal{X}[j, h]$라고하자.
+또한  $V[j,h] = v(\mathcal{X}[j, h])$이다.
+$j=0$ 혹은 $h=0$이면 $\mathcal{X}[j, h] = \O$이고 $V[j, h] = 0$이 되는 것은 분명하다.
+편의상 $h < 0$이면 $V[j, h] = -\infty$이라고 정의하자.
+\fi
+
+\ifen
 For the remaining indices, $\mathcal{X}[j, h]$ either contains $j$ or not.
@@
 From these observations, we obtain the following Bellman equation for $ j = 1\dots m$ and $h = 1\dots H$:
+\else
+남은 지표에 대해, $\mathcal{X}[j, h]$는 $j$를 포함하거나 포함하지 않는다.
+$j$를 포함하지 않는다면 $\mathcal{X}[j, h] = \mathcal{X}[j-1, h]$이다.
+$\mathcal{X}[j, h]$가 $j$를 포함한다면, 그의 가치는  $(1 - f_j) v(\mathcal{X}[j, h]\setminus \{j\}) + f_j t_j$다.
+따라서 $\mathcal{X}[j, h]\setminus \{j\}$는 남은 예산과 남은 학교에 대한 최적 포트폴리오다.
+즉,  $\mathcal{X}[j, h] = \mathcal{X}[j-1, h - g_j] \cup\{j\}$이다.
+이 관찰에 따라 $ j = 1\dots m$와 $h = 1\dots H$에 대해 다음 같은 Bellman 식을 얻는다.
+\fi
 \begin{align}
@@
 \end{align}
+\ifen 
 with the convention that $ -\infty \cdot 0 = -\infty$.
@@
 The algorithm below performs these computations and outputs the optimal portfolio $\mathcal{X}$.
-
+\else
+단, $ -\infty \cdot 0 = -\infty$으로 처리한다.
+그러면 $\mathcal{X}[j, h]$가  $j$를 포함할 필수충분 조건이 $V[j, h]> V[j-1, h]$임을 관찰하면 대응되는 최적 포트폴리오를 쉽게 계산할 수 있으며 전역 최적해는 $\mathcal{X}[m, H]$이다.
+아래 알고리즘은 이 계산을 수행하고 최적 포트폴리오 $\mathcal{X}$를 출력한다.
+\fi
+
+\ifen {
 \begin{algorithm}[h] 
@@
 \caption{Dynamic program for Ellis's problem with integral application costs.} \label{ellisDP1}
-\KwData{Utility values $t \in[0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in \mathbb{N}^m$, budget $H \in\mathbb{N}$.}
+\KwData{Utility values $t \in(0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in \mathbb{N}^m$, budget $H \in\mathbb{N}$.}
 Index schools in ascending order by $t$\;
@@
 %}
-Fill a lookup table with the entries of $V[j, h]$\; \label{Vcreatedlookuptable}
+Fill a lookup table with the values of $V[j, h]$\; \label{Vcreatedlookuptable}
 $h \gets H$\;
@@
 \end{algorithm}
-
-\begin{theorem}[Validity of Algorithm \ref{ellisDP1}]
-Algorithm \ref{ellisDP1} produces an optimal application portfolio for Ellis's problem in $O(H m + m \log m)$ time and $O(H m)$ space.
+} \else {
+\begin{algorithm}[h] 
+%\DontPrintSemicolon
+\caption{정수 지원 비용의 엘리스 문제를 위한 동적 계획 해법.} \label{ellisDP1}
+\KwData{효용 모수 $t \in(0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 비용 $g \in \mathbb{N}^m$, 예산 $H \in\mathbb{N}$.}
+$t$의 순서대로 학교를 배열한다\;
+%\SetKwProg{Fn}{function}{ do}{end}
+%\Fn{$V(j, h)$}{
+%	\lIf{$h < g_j$}{\Return{$-\infty$}}
+%	\lElseIf{$j = 0$ or $h=0$}{\Return{$0$}}
+%	\lElse{\Return{$\max\bigl\{ V(j-1, h), (1 - f_j) V(j-1, h-g_j) + f_j t_j \bigr\}$}}
+%}
+$V[j, h]$-값으로 표를 채운다\; \label{Vcreatedlookuptable}
+$h \gets H$\;
+$\mathcal{X} \gets \O$\;
+\For{$j = m, m-1, \dots, 1$}{
+	\If{$V[j-1, h] < V[j, h]$}{
+		$\mathcal{X} \gets \mathcal{X}\cup\{j\}$\; 
+		$h \gets h - g_j$\;
+	}
+}
+\Return{$\mathcal{X}$}
+\end{algorithm}
+}\fi
+
+\begin{theorem}[\ifen Validity of Algorithm \ref{ellisDP1}\else 알고리즘 \ref{ellisDP1}의 타당성\fi]
+\ifen 
+Algorithm \ref{ellisDP1} produces an optimal application portfolio for Ellis's problem in $O(H m + m \log m)$-time.
+\else
+알고리즘 \ref{ellisDP1}은(는) $O(H m + m \log m)$-시간 안의 엘리스 문제의 최적 포트폴리오를 출력한다.
+\fi
 \end{theorem}
@@
 \begin{proof}
+\ifen
 Optimality follows from the foregoing discussion.
@@
 Each entry is generated in unit time, and the size of the table is $O(Hm)$.
+\else
+위 논의에 따라 최적성이 성립한다. $t$를 배열하는 시간은 $O(m \log m)$이다. 목병 단계는 \ref{Vcreatedlookuptable}줄에서 $V[j, h]$의 표를 만드는 것이다. 각 원소를 단위 시간 안에 생성할 수 있으며 표의 크기가 $O(Hm)$다.
+\fi
 \end{proof}

### 2022-03-14 paper/paper.tex
typo
finish translations so far; rm false statement about submodularity

@@
 
-\ifen { \title{How to apply to college} }
-\else {\title{How to apply to college}} \fi
-\author{Max Kapur}
-\date{\today}                % Activate to display a given date or no date
+\ifen
+\newcommand{\thetitle}{How to apply to college}
+\newcommand{\therunningtitle}{\thetitle}
+\newcommand{\thedate}{\today}
+\newcommand{\theauthor}{Max Kapur}
+\else
+\newcommand{\thetitle}{대학 지원 최적화 문제}
+\newcommand{\therunningtitle}{\thetitle}
+\newcommand{\thedate}{\today}
+\newcommand{\theauthor}{Max Kapur}
+\fi
+
+\title{\thetitle}
+\author{\theauthor}
+\date{\thedate}
 
@@
 \fancyhf{}
-\rhead{How to apply to college \\}
-\lhead{Max Kapur \\ \date{\today}}
+\lhead{\theauthor}
+\chead{\thedate}
+\rhead{\thetitle}
 \cfoot{\thepage}
@@
 
-\section*{Abstract}
+\begin{abstract}
 This paper considers the maximization of the expected maximum value of a portfolio of random variables subject to a budget constraint.
@@
 
-\section*{요약}
+\ifen
+\else The full text of this paper is available in English at \url{https://github.com/maxkapur/OptimalApplication}.\fi
+\end{abstract}
+
+\renewcommand{\abstractname}{요약}
+\begin{abstract}
 본 논문은 다수의 확률 변수로 구성된 포트폴리오의 기대 최대값을 예산 조건 하에서 최대화하는 문제를 고려한다.
@@
 
+\ifen
+\else 본 논문은 \url{https://github.com/maxkapur/OptimalApplication}에서 전문을 한글로 공유한다. \fi
+\end{abstract}
 \pagebreak
@@
 입학 확률을 적당히 조정하면 특성벡터의 선형 함수를 원래 목적함수로 원하는 만큼의 정확성을 가지고 근사할 수 있으며 NP-completeness 증명에서 이 성질을 활용한다.
-배낭 문제에 확률성을 도입한 선행 연구 중, 각 상품의 효용이 정해진 확률 분포로 결정되는 모형(Steinberg와 Parks 1979; Carraway 외 1993), 그리고 배낭에 삽입한 다음에 상품의 무게를 관측할 수 있는 온라인 모형 (Dean et al.
-2008) 등이 있다.
+배낭 문제에 확률성을 도입한 선행 연구 중, 각 상품의 효용이 정해진 확률 분포로 결정되는 모형(Steinberg와 Parks 1979; Carraway 외 1993), 그리고 배낭에 삽입한 다음에 상품의 무게를 관측할 수 있는 온라인 모형 (Dean 외 2008) 등이 있다.
 본 연구가 고려하는 문제는 Steinberg와 Parks 그리고 Carraway 외가 고려한 ``우선순위 배낭 문제''와는 유사성을 가지지만, 우선순위 배낭 문제는 대학 지원 문제의 특색인 `maximax' 형태를 가지지는 않는다.
@@
 탐욕 알고리즘이 좋은 근사해거나 정확한 알고리즘이 되는 최적화 모형들이 알려져 있다.
-집합 크기 제약 하에서 하위 모듈 집합 함수(submodular set function)를 최대화하는 문제가 대표적인 예다 (Fisher et al.
-1978).
+집합 크기 제약 하에서 하위 모듈 집합 함수(submodular set function)를 최대화하는 문제가 대표적인 예다 (Fisher 외 1978).
 반면에 이진 배낭 문제 같은 경우에서는 가장 직관적인 탐욕 알고리즘이 최적과 거리가 먼 해를 출력하는 예를 만들 수 있다 (Vazirani 2001).
@@
 
-\pagebreak
+%\pagebreak
 \ifen \section{Notation and preliminary results} \else \section{표기법과 예비 결과}\fi\label{preliminaries}
@@
 \else
-지원 제한이 $h$일 때, 최적 포트폴리오가 $\mathcal{X}_h$고 기대 효용 $f_j t_j$가 가장 높은 $h$개의 학교의 집합이 $\mathcal{T}_h$라고 하자. 그러면 $v(\mathcal{T}_h) / v(\mathcal{X}_h) \geq 1/h$이다.
+지원 제한이 $h$일 때, 최적 포트폴리오가 $\mathcal{X}_h$이고 기대 효용 $f_j t_j$가 가장 높은 $h$개의 학교의 집합이 $\mathcal{T}_h$라고 하자. 그러면 $v(\mathcal{T}_h) / v(\mathcal{X}_h) \geq 1/h$이다.
 \fi
@@
 
-\begin{corollary}
-\ifen
-The function $v(\mathcal{X})$ is not submodular. 
-\else
-$v(\mathcal{X})$는 하위 모듈 함수가 아니다.
-\fi
-\end{corollary}
-\begin{proof}
-\ifen
-If $v(\mathcal{X})$ is submodular, then theorem 4.2 of Fisher et al.
+%% Wrong: The Fisher theorem would apply to the greedy, i.e. correct algorithm. 
+%\begin{corollary}
+%\ifen
+%The function $v(\mathcal{X})$ is not submodular. 
+%\else
+%$v(\mathcal{X})$는 하위 모듈 함수가 아니다.
+%\fi
+%\end{corollary}
+%\begin{proof}
+%\ifen
+%If $v(\mathcal{X})$ is submodular, then theorem 4.2 of Fisher et al.
 (1978) implies that the na\"ive algorithm achieves an optimality ratio of $v(\mathcal{T}_h) / v(\mathcal{X}_h)  \geq 1 - \left(\frac{h-1}{h}\right)^h$.
 Example \ref{tightexampleforoneoverhopt} provides a counterexample.
-\else
-$v(\mathcal{X})$가 하위 모듈 함수라면 Fisher 외(1978)의 정리 4.2는 나이브 해법이 $v(\mathcal{T}_h) / v(\mathcal{X}_h)  \geq 1 - \left(\frac{h-1}{h}\right)^h$ 같은 근사 비율을 이룬다고 의미한다. 예 \ref{tightexampleforoneoverhopt}는 반례다.
-\fi
-\end{proof}
-
-\ifen
-Although the na\"ive algorithm is inoptimal, we can still find the optimal solution in time polynomial in $h$ and $m$, as we will now show.
-\else
-나이브 해법은 최적해가 아니지만 $h$와 $m$의 다항 시간 안에 최적해를 구할 수 있음을 보이자.
+%\else
+%$v(\mathcal{X})$가 하위 모듈 함수라면 Fisher 외(1978)의 정리 4.2는 나이브 해법이 $v(\mathcal{T}_h) / v(\mathcal{X}_h)  \geq 1 - \left(\frac{h-1}{h}\right)^h$ 같은 근사 비율을 이룬다고 의미한다. 예 \ref{tightexampleforoneoverhopt}는 반례다.
+%\fi
+%\end{proof}
+
+\ifen
+Although the na\"ive algorithm is inoptimal, we can still find the optimal solution in $O(hm)$-time, as we will now show.
+\else
+나이브 해법은 최적해가 아니지만 $O(hm)$-시간 안에 최적해를 구할 수 있다.
 \fi
@@
 \else
-위 결과는 $\mathcal{X}_h$가 어떤 시장의 최적 $h$-포트폴리오일 때, $v(\mathcal{X}_h)$가 $O(h)$ 함수라고 의미한다. 예\ref{tightexampleforoneoverhopt}에서 $v(\mathcal{X}_h)$를 $h$에 임의로 가깝게 조정할 수 있으므로 이 상한이 타이트함을 알 수 있다.
+위 결과는 $\mathcal{X}_h$가 어떤 시장의 최적 $h$-포트폴리오일 때, $v(\mathcal{X}_h)$가 $O(h)$ 함수라고 의미한다. 예 \ref{tightexampleforoneoverhopt}에서 $v(\mathcal{X}_h)$를 $h$에 임의로 가깝게 조정할 수 있으므로 이 상한이 타이트함을 알 수 있다.
 \fi
@@
 
-\pagebreak
+%\pagebreak
 \ifen \section{Heterogeneous application costs} \else\section{동일하지 않은 지원 비용} \fi\label{hetappcosts}
@@
 \else
-엘리스 문제의 결정 형태는 NP-complete다.
+엘리스 문제의 결정 형태는 NP-complete하다.
 \fi
@@
 정수 최적화의 고전적인 기법 중 분지한계법(branch and bound)이 있다.
-이는 하나 이상의 결정 변수를 고정시키고 해당 부문제를 탐색하는 해법이며, 목적함수의 상한(bound)을 활용해서 기존 해보다 좋은 해를 생성할 수 없는 결정 나무의 가지(branch)를 제외시켜 결정 공간을 줄인다 (Martello and Toth 1990; Kellerer et al.
-2004).
+이는 하나 이상의 결정 변수를 고정시키고 해당 부문제를 탐색하는 해법이며, 목적함수의 상한(bound)을 활용해서 기존 해보다 좋은 해를 생성할 수 없는 결정 나무의 가지(branch)를 제외시켜 결정 공간을 줄인다 (Martello and Toth 1990; Kellerer 외 2004).
 본 항에서, 엘리스 문제의 정수 모형과 그의 상한이 되는 선형 계획을 제시한다.
@@
 \else
-$v(x)$에서 등장하는 곱은 1을 넘을 수 없으므로 다음 선형 완화 문제는 최적 포트폴리오 가치의 상한이 돈다.
+$v(x)$에서 등장하는 곱은 1을 넘을 수 없으므로 다음 선형 완화 문제는 최적 포트폴리오 가치의 상한이 된다.
 \fi
@@
 문제 \ref{LPrelaxation}은(는) ``연속 배낭 문제''라고 불리며 다음 탐욕 알고리즘을 적용하면 $O(m \log m)$-시간 안에 쉽게 풀 수 있다: $x = \mathbf{0}$로 초기화한다.
-$H > 0$가 성립하는 동안, $f_j t_j / g_j$-값이 가장 높은 학교 $j$를 택한다.
-$x_j \gets \min\{1, H / g_j\}$ 그리고 $H \gets H - g_j$ 같이 수정해서 반복한다 (Dantzig 1957).
+$H > 0$임이 성립하는 동안, $f_j t_j / g_j$-값이 가장 높은 학교 $j$를 택한다.
+$x_j \gets \min\{1, H / g_j\}$ 그리고 $H \gets H - g_j$로 수정해서 반복한다 (Dantzig 1957).
 \fi
@@
 본 연구의 분지한계법 설계에서, `마디'(node)는 학교를 세 군으로 나눈 분할 $\mathcal{C}= \mathcal{I} \cup \mathcal{O} \cup \mathcal{N}$로 묘사한다.
-단, $\sum_{j \in \mathcal{I}} g_j \leq H$을 만족한다.
+단, $\mathcal{I}$는 $\sum_{j \in \mathcal{I}} g_j \leq H$을 만족한다.
 $\mathcal{I}$는 지원 포트폴리오에 속한 학교(`in': $x_j = 1$), $\mathcal{O}$는 속하지 않은 학교(`out': $x_j = 0$), 그리고 $\mathcal{N}$은 미결의 학교(`negotiable')로 이루어진다.
@@
 \ifen Algorithm \ref{ellisbnb} produces an optimal application portfolio for Ellis's problem.
-\else 알고리즘 \ref{ellisbnb}은(는) 엘리스 문제의 최적 지원 포트폴리오를 출력한다.
+\else 알고리즘 \ref{ellisbnb}은(는) 엘리스 문제의 최적 지원 포트폴리오를 출력한다. \fi
 \end{theorem}
@@
 $j$를 포함하지 않는다면 $\mathcal{X}[j, h] = \mathcal{X}[j-1, h]$이다.
-$\mathcal{X}[j, h]$가 $j$를 포함한다면, 그의 가치는  $(1 - f_j) v(\mathcal{X}[j, h]\setminus \{j\}) + f_j t_j$다.
+$\mathcal{X}[j, h]$가 $j$를 포함한다면, 그의 가치는  $(1 - f_j) v(\mathcal{X}[j, h]\setminus \{j\}) + f_j t_j$이다.
 따라서 $\mathcal{X}[j, h]\setminus \{j\}$는 남은 예산과 남은 학교에 대한 최적 포트폴리오다.
@@
 %}
-$V[j, h]$-값으로 표를 채운다\; \label{Vcreatedlookuptable}
+$V[j, h]$의 값으로 표를 채운다\; \label{Vcreatedlookuptable}
 $h \gets H$\;
@@
 \else
-위 논의에 따라 최적성이 성립한다. $t$를 배열하는 시간은 $O(m \log m)$이다. 목병 단계는 \ref{Vcreatedlookuptable}줄에서 $V[j, h]$의 표를 만드는 것이다. 각 원소를 단위 시간 안에 생성할 수 있으며 표의 크기가 $O(Hm)$다.
+위 논의에 따라 최적성이 성립한다. $t$를 배열하는 시간은 $O(m \log m)$이다. 목병 단계는 \ref{Vcreatedlookuptable}줄에서 $V[j, h]$의 표를 만드는 것이다. 각 원소를 단위 시간 안에 생성할 수 있으며 표의 크기가 $O(Hm)$이다.
 \fi
@@
 
-%\subsubsection{Dynamic program based on portfolio valuations}
-%
-%With minimal loss of generality, we assume that all $t_j\in \mathbb{N}$ and $f_j \in (0, 1] \cap \mathbb{Q}$.
-Let $D$ denote the product of the denominators of the $f_j$-values.
-Since $v(\mathcal{X})$ is linear in $t_j$, by rescaling the $t_j$-values, we may assume that each $t_j$ is divisible by $D$.
-Therefore, for any portfolio $\mathcal{X}\subseteq\mathcal{X}$, $v(\mathcal{X})$ is a natural number between zero and $t_m$.
-%Mention size of t_m
-%
-%For integers $0 \leq j \leq m$ and $v \leq t_m$, let $\mathcal{W}[j, v]$ denote the least expensive portfolio using only the schools $\{ 1, \dots, j\}$ and having value at least $v$, if such a portfolio exists.
-Let $G[j, v] = v(\mathcal{W}[j, v])$ denote the corresponding value, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist.
-It is clear that if $v \leq 0$, then $\mathcal{W}[j, v] = \O$ and $G[j, h] = 0$.
-If $j = 0$ and $v > 0$, then $G[j, h] = \infty$.
-For the remaining indices (where $j, v > 0$), we claim that
-%\begin{align} \label{recursionrelationforcostmindp}
-%G[j, v] &=
-%\begin{cases}
-%\infty, \quad & t_j < v \\
-%\min\bigl\{G[j-1, v], g_j + G[j-1, v - \Delta] \bigr\}, \quad & t_j \geq v 
-%%\begin{cases}
-%%\min\bigl\{G[j-1, v], g_j + G[j-1, v - \Delta] \bigr\}, \quad &f_j < 1 \\
-%%\min\bigl\{G[j-1, v], g_j \bigr\}, \quad &f_j = 1 \text{ and } f_j t_j \geq v\\
-%%G[j-1, v], \quad &f_j = 1 \text{ and } f_j t_j < v
-%\end{cases}\\
-%\text{where}\qquad
-%\Delta &= 
-%\begin{cases}
-%\lfloor \frac{f_j}{1 - f_j} (t_j - v)\rfloor, \quad & f_j < 1\\
-%\infty, &f_j = 1
-%\end{cases}
-%\end{align}
-%In the $t_j < v$ case, any feasible portfolio must be composed of schools with utility less than $v$, and therefore its valuation can not equal $v$, meaning that $\mathcal{W}[j, v]$ is undefined.
-In the $t_j \geq v$ case, the first argument to $\min\{\}$ says simply that omitting $j$ and choosing $\mathcal{W}[j-1, v]$ is a feasible choice for $\mathcal{W}[j, v]$.
-If, on the other hand, $j \in \mathcal{W}[j, v]$, then
-%\begin{equation}v(\mathcal{W}[j, v]) = (1 - f_j )v(\mathcal{W}[j, v]\setminus \{j\}) + f_j t_j\end{equation}
-%Therefore, the subportfolio $\mathcal{W}[j, v]\setminus \{j\}$ must have a value of at least $v - \Delta$, where $\Delta$ satisfies $v = (1 - f_j )(v - \Delta) + f_j t_j $.
-When $f_j < 1$, the solution to this equation is $ \Delta = \frac{f_j}{1 - f_j} (t_j - v)$.
-This value may not be an integer; however, because $v(\mathcal{X})$ can take only integer values, if some portfolio has $v(\mathcal{X})\geq v - \Delta$, then $v(\mathcal{X})\geq \lceil v - \Delta \rceil$ as well, which allows us to apply the floor operator to $\Delta$.
-When $f_j = 1$, then the singleton $\{j\}$ has $v(\{j\}) \geq v$, so
-%\begin{equation}G[j, v] = \min\bigl\{G[j-1, v], g_j \bigr\}\end{equation}
-%Defining $\Delta = \infty$ in this case ensures that $G[j-1, v-\Delta] = 0$ as required.
-%%we have $t_j < v$, and since no portfolio can be more valueable than the largest $t_j$ value of its elements, we must have $G[j, v] =\infty$.
-Defining $\Delta$ as zero in this guess automatically chooses $G[j, v] = G[j-1, v] = \infty$ for the same reason.
-%%
-%%In the $f_j t_j \geq v$ case, the singleton $\{j\}$ is feasible for $j$ and $v$, so $\mathcal{W}[j, v]$ is simply the cheaper of $\{j\}$ and $.\mathcal{W}[j-1, v]$.
-%
-%Once $G[j, v]$ has been calculated at each index, the corresponding portfolio can be found by applying the observation that $
-%\mathcal{W}[j, v]$ contains $j$ if and only if $G[j, v] < G[j-1, v]$.
-Then solution to Ellis's problem is obtained by applying this observation to the optimal objective value $\max\{ w: G[m, w] \leq H\}$.
-%
-%
-%
-%\begin{algorithm}[H] 
-%%\DontPrintSemicolon
-%\caption{Dynamic program for Ellis's problem with integral utility parameters and rational admissions probabilities.} \label{ellisDP2}
-%\KwData{Utility values $t \in \mathbb{N}^m$, admissions probabilities $f \in \bigl((0, 1] \cap \mathbb{Q}\bigr)^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$.}
-%Index schools in ascending order by $t$\;
-%Multiply each $t_j$ by $D$, the product of the denominators of the $f_j$\;
-%\SetKwProg{Fn}{function}{ do}{end}
-%\Fn{$G(j, v)$}{
-%	\lIf{$v \leq 0$}{\Return{$0$}}
-%	\lElseIf{$j=0$ or $t_j < v$}{\Return{$\infty$}}
-%	\Else {\Return{$\min\bigl\{ G(j-1, v), g_j + G(j-1, v -  \Delta) \bigr\}$}}
-%}
-%%	\uIf{$f_j < 1$}{\Return{$\min\bigl\{ G(j-1, v), g_j + G(j-1, v -  \lfloor \frac{f_j}{1 - f_j} (t_j - v)\rfloor ) \bigr\}$}}
-%%	\uElseIf{$f_j = 1$ and $f_j t_j \geq v$}{\Return{$\min\bigl\{G[j-1, v], g_j \bigr\}$}}
-%%	\lElse{\Return{$G(j-1, v)$}}
-%%	}
-%%\lFor{$j=1\dots m$}
-%%{$\operatorname{push!}\bigl(\mathtt{H}, (j, f_j, t_j)\bigr)$}
-%%\For{$i=1\dots h$}
-%%{
-%%    $\mathtt{D} \gets \operatorname{pop!}(\mathtt{H})$\;
-%%    $\operatorname{append!}(\mathtt{X}, \mathtt{D.j})$\;
-%%     \lIfElse{$i=1$}{$\operatorname{append!}(\mathtt{V}, \mathtt{D.f}\cdot\mathtt{D.t})$}
-%%     {$\operatorname{append!}(\mathtt{V}, \mathtt{V[i-1]} + \mathtt{D.f}\cdot\mathtt{D.t})$}
-%%    \For{$\mathtt{E} \in \mathtt{H}$}
-%%	{
-%%	\lIfElse{$\mathtt{E.j} < \mathtt{D.j}$}{$\mathtt{E.t} \gets  \mathtt{E.t} \cdot(1 -  \mathtt{D.f})$}{$\mathtt{E.t} \gets  \mathtt{E.t} -  \mathtt{D.f}\cdot\mathtt{D.t}$}
-%%	}
-%%}
-%$v \gets  \max\{ w: G(m, w) \leq H\}$\;
-%$\mathcal{X} \gets \O$\;
-%\For{$j = m, m-1, \dots, 1$}{
-%	\If{$G(j, v)< \infty$ and $G(j, v) < G(j-1, v)$}{
-%		$\mathcal{X} \gets \mathcal{X}\cup\{j\}$\; 
-%		$v \gets v - \Delta$\;
-%	}
-%}
-%\Return{$\mathcal{X}$}
-%\end{algorithm}
-%
-%\begin{theorem}[Validity of Algorithm \ref{ellisDP2}]
-%Algorithm \ref{ellisDP2} produces an optimal application portfolio for Ellis's problem in $O(D m + m \log m)$ time and $O(D m)$ space, where $D$ is the product of the denominators of the $f_j$-values.
-%\end{theorem}
-%
-%\begin{proof}
-%Optimality follows from the foregoing discussion. The computational complexity follows by symmetry with Algorithm \ref{ellisDP1}. 
-%\end{proof}
-
 \ifen \subsection{Fully polynomial-time approximation scheme}\else \subsection{완전 다항 시간 근사 해법}\fi \label{fptashead}
+\ifen 
 As with the knapsack problem, Ellis's problem admits a complementary dynamic program that iterates on the value of the cheapest portfolio instead of on the cost of the most valuable portfolio.
-We will use this algorithm as the basis for a fully polynomial-time approximation scheme for Ellis's problem that uses $O(m^3 / \varepsilon)$ time and space.
+We will use this algorithm as the basis for a fully polynomial-time approximation scheme (FPTAS) for Ellis's problem that uses $O(m^3 / \varepsilon)$-time and -space.
 Here we assume, with a small loss of generality, that each $t_j$ is a natural number.
 %Now, we modify Algorithm \ref{ellisDP2} into a fully polynomial-time approximation scheme for Ellis's problem.
-
+\else
+엘리스의 문제는 배낭 문제와 같이, 가치가  가장 높은 포트폴리오의 비용 대신 비용이 가장 낮은 포트폴리오의 가치를 탐색하는 보완적인 동적 계획이 존재한다.
+이를 기반으로  $O(m^3 / \varepsilon)$-시간과 -공간을 사용하는 엘리스 문제의 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme, FPTAS)을 도출할 수 있다.
+여기서 일반성을 약간 제안해서 각 $t_j$가 자연수임을 가정한다.
+\fi
+
+\ifen 
 We will represent approximate portfolio valuations using a fixed-point decimal with a precision of $P$, where $P$ is the number of digits to retain after the decimal point.
@@
 Since $\bar U = \sum_{j\in \mathcal{C}} f_j t_j$ is an upper bound on the valuation of any portfolio, and since we will ensure that each fixed-point approximation is an underestimate of the portfolio's true valuation, the set $\mathcal{V}$ of possible valuations possible in the fixed-point framework is finite:
+\else
+포티폴리오의 근사적 가치를 정확도 $P$로 구성된 고정소수점 십진수(fixed-point decimal)로 나타내자.
+다, $P$는 소수점 뒤에 등장하는 숫자의 수다.
+이때 $x$를 가장 가까운 고정소수점 십진수로 내림한 것을 $r[x] =  10^{-P}\lfloor 10^P x \rfloor$라고 하자.
+임의의 포트폴리오의 가치가 $\bar U = \sum_{j\in \mathcal{C}} f_j t_j$를 넘을 수 없으며, 모든 포트폴리오 가치의 고정소수점 십진수 근사가 실제값보다 작게 나타내도록 한다.
+따라서 고정소수점 환경에서 발생할 수 있는 포트폴리오 가치로 이루어진 집합 $\mathcal{V}$는 유한 집합이다.
+\fi
 \begin{equation}
@@
 \end{equation}
-Then $|\mathcal{V} | = \bar U \times 10^P + 1$.
-
+\ifen Then $|\mathcal{V} | = \bar U \times 10^P + 1$.
+\else 그러면 $|\mathcal{V} | = \bar U \times 10^P + 1$이다.\fi
+
+\ifen
 For the remainder of this subsection, unless otherwise specified, the word \emph{valuation} refers to a portfolio’s valuation within the fixed-point framework, with the understanding that this is an approximation.
 We will account for the approximation error below when we prove the dynamic program’s validity.
-
+\else
+본 항 나머지에서, 다른 언급이 없으면 포트폴리오의 ``가치''라고 하면 고정소수점 환경 안에서의 가치를 의미하며 이것이 근사값임을 유념한다. 나중에 동적 계획의 타당성을 증명할 때 근사 오차를 처리한다.
+\fi
+
+\ifen
 For integers $0 \leq j \leq m$ and $v \in [-\infty, 0) \cup \mathcal{V}$, let $\mathcal{W}[j, v]$ denote the least expensive portfolio that uses only schools $\{ 1, \dots, j\}$ and has valuation at least $v$, if such a portfolio exists.
-Denote its cost by $G[j, v]$, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist.
+Denote its cost by $G[j, v] = \sum_{j\in \mathcal{W}[j, v]} g_j$, where $G[j, v] = \infty$ if $\mathcal{W}[j, v]$ does not exist.
 It is clear that if $v \leq 0$, then $\mathcal{W}[j, v] = \O$ and $G[j, h] = 0$, and that if $j = 0$ and $v > 0$, then $G[j, h] = \infty$.
 For the remaining indices (where $j, v > 0$), we claim that
+\else
+정수 $0 \leq j \leq m$ 그리고 $v \in [-\infty, 0) \cup \mathcal{V}$에 대해, $\{ 1, \dots, j\}$에 속한 학교만 사용하면 최소한 $v$의 가치를 가지는 포트폴리오가 존재한다면 그 중 가장 비용이 낮은 포트폴리오를  $\mathcal{W}[j, v]$라고 하자.
+그의 비용을 $G[j, v] = \sum_{j\in \mathcal{W}[j, v]} g_j$라고 하자.
+단, $\mathcal{W}[j, v]$가 존재하지 않으면 $G[j, v] = \infty$으로 처리한다.
+$v \leq 0$이면 $\mathcal{W}[j, v] = \O$이고 $G[j, h] = 0$이며, $j = 0$이고 $v > 0$이면 $G[j, h] = \infty$인 것이 분명하다.
+남은 지표(즉 $j, v > 0$)는 다음을 만족한다.
+\fi
 \begin{align} \label{recursionrelationforcostmindp}
@@
 r\left[\frac{f_j}{1 - f_j} (t_j - v)\right], \quad & f_j < 1\\
-\infty, &f_j = 1
+\infty, &f_j = 1\ifen.\fi
 \end{cases} \label{deltajvdef}
 \end{align}
+\ifen
 In the $t_j < v$ case, any feasible portfolio must be composed of schools with utility less than $v$, and therefore its valuation can not equal $v$, meaning that $\mathcal{W}[j, v]$ is undefined.
@@
 If, on the other hand, $j \in \mathcal{W}[j, v]$, then
+\else
+$t_j < v$인 경우, 가능한 포트폴리오는 효용이 $v$보다 작은 학교로 구성되므로 그의 가치는 $v$를 넘을 수가 없고 $\mathcal{W}[j, v]$가 정의되지 않는다.
+$t_j \geq v$인 경우, $\min\{\}$의 첫 입력값은 $j$를 제외하고 $\mathcal{W}[j-1, v]$로 지원하는 것이 가능한 선택임을 의미한다.
+반면에 $j \in \mathcal{W}[j, v]$이라면 다음이 성립한다.
+\fi
 \begin{equation} \label{solvemeforvwjvminusj}
-v(\mathcal{W}[j, v]) = (1 - f_j )v(\mathcal{W}[j, v]\setminus \{j\}) + f_j t_j.\end{equation}
+v(\mathcal{W}[j, v]) = (1 - f_j )v(\mathcal{W}[j, v]\setminus \{j\}) + f_j t_j\ifen.\fi\end{equation}
+\ifen
 Therefore, the subportfolio $\mathcal{W}[j, v]\setminus \{j\}$ must have a valuation of at least $v - \Delta$, where $\Delta$ satisfies $v = (1 - f_j )(v - \Delta) + f_j t_j $.
@@
 When $t_j \geq v$ and $f_j = 1$, the singleton $\{j\}$ has $v(\{j\}) \geq v$, so
+\else
+이는 부분 포트폴리오 $\mathcal{W}[j, v]\setminus \{j\}$가 최소한 $v - \Delta$의 가치를 가져야 한다고 의미한다.
+단,  $\Delta$는 $v = (1 - f_j )(v - \Delta) + f_j t_j $를 만족한다.
+$f_j < 1$일 때, 이 식의 해는 $ \Delta = \frac{f_j}{1 - f_j} (t_j - v)$이다.
+이값을 내림함으로써 $\mathcal{W}[j, v]$가 `최소한' $v - \Delta$임을 보장한다.
+$t_j \geq v$이고 $f_j = 1$일 때, 단일 원소 집합 $\{j\}$에 대해 $v(\{j\}) \geq v$이므로
+\fi
 \begin{equation}G[j, v] = \min\bigl\{G[j-1, v], g_j \bigr\}.\end{equation}
+\ifen
 Defining $\Delta_j(v) = \infty$ in this case ensures that $g_j + G[j-1, v-\Delta_j(v)] = g_j+ G[j-1, v-\infty] = g_j $ as required.
+\else
+이때 $\Delta_j(v) = \infty$으로 정의하면 $g_j + G[j-1, v-\Delta_j(v)] = g_j+ G[j-1, v-\infty] = g_j $와 같은 적절한 값이 나온다.
+\fi
 %we have $t_j < v$, and since no portfolio can be more valueable than the largest $t_j$ value of its elements, we must have $G[j, v] =\infty$.
@@
 
+\ifen
 Once $G[j, v]$ has been calculated at each index, the associated portfolio can be found by applying the observation that $\mathcal{W}[j, v]$ contains $j$ if and only if $G[j, v] < G[j-1, v]$.
 Then an approximate solution to Ellis's problem is obtained by computing the largest achievable objective value $\max\{ w: G[m, w] \leq H\}$ and corresponding portfolio.
-
+\else
+각 지표에 대해 $G[j, v]$를 계산한 다음에,  $\mathcal{W}[j, v]$가 $j$를 포함할 필수충분 조건인 $G[j, v] < G[j-1, v]$를 적용하면 해당 최적 포트폴리오를 구할 수 있다.
+그러면 가능하면서 최대적인 목적함수값 $\max\{ w: G[m, w] \leq H\}$와 해당 포트폴리오를 계산하면 엘리스 문제의 근사해를 얻을 수 있다.
+\fi
+
+\ifen
 \begin{algorithm}[h] 
@@
 \end{algorithm}
-
-\begin{theorem}[Validity of Algorithm \ref{ellisDP3}]
-Algorithm \ref{ellisDP3} produces a $(1 - \varepsilon)$-optimal application portfolio for Ellis's problem in $O(m^3 /\varepsilon)$ time. % and $O(m^3/\varepsilon)$ space.
-% in $O(D m + m \log m)$ time and $O(D m)$ space, where $D$ is the product of the denominators of the $f_j$-values.
+\else
+\begin{algorithm}[h] 
+%\DontPrintSemicolon
+\caption{엘리스의 문제를 위한 완전 다항 시간 근사 해법.} \label{ellisDP3}
+\KwData{효용 모수 $t \in \mathbb{N}^m$, 합격 확률 $f \in (0, 1]^m$, 지원 비용 $g \in (0, \infty)^m$, 예산 $H \in (0, \infty)^m$, 허용치 $\varepsilon \in (0, 1)$.}
+$t$의 순서대로 학교를 배열한다\;
+정확도 $P \gets \bigl\lceil\log_{10}\left(m^2 / \varepsilon \bar U\right)\bigr\rceil$\;
+%\SetKwProg{Fn}{function}{ do}{end}
+%\Fn{$G(j, v)$}{
+%	\lIf{$v \leq 0$}{\Return{$0$}}
+%	\lElseIf{$j=0$ or $t_j < v$}{\Return{$\infty$}}
+%	\lElse {\Return{$\min\bigl\{ G(j-1, v), g_j + G\bigl(j-1, v -  \Delta_j(v)\bigr) \bigr\}$}}
+%}
+$G[j, h]$의 값으로 표를 채운다\; \label{createdlookuptable}
+$v\gets  \max\{ w \in \mathcal{V} : G[m, w] \leq H\}$\; \label{vrecordedhere}
+$\mathcal{X} \gets \O$\;
+\For{$j = m, m-1, \dots, 1$}{
+	\If{$G[j, v]< \infty$이고 $G[j, v] < G[j-1, v]$}{
+		$\mathcal{X} \gets \mathcal{X}\cup\{j\}$\; 
+		$v \gets v -  \Delta_j(v)$\;
+	}
+}
+\Return{$\mathcal{X}$}
+\end{algorithm}
+\fi
+
+\begin{theorem}[\ifen Validity of Algorithm \ref{ellisDP3}\else 알고리즘 \ref{ellisDP3}의 타당성\fi]
+\ifen
+Algorithm \ref{ellisDP3} produces a $(1 - \varepsilon)$-optimal application portfolio for Ellis's problem in $O(m^3 /\varepsilon)$-time. % and $O(m^3/\varepsilon)$ space.
+\else
+알고리즘 \ref{ellisDP3}은(는) $O(m^3 /\varepsilon)$-시간 안에 엘리스의 문제를 위한 $(1 - \varepsilon)$-근사 최적 포트폴리오를 출력한다.
+\fi
 \end{theorem}
@@
 \begin{proof}
+\ifen
 (Optimality.) Let $\mathcal{W}$ denote the output of Algorithm \ref{ellisDP3} and $\mathcal{X}$ the true optimum.
 We know that $v(\mathcal{X}) \leq \bar U$, and because each singleton portfolio is feasible, $\mathcal{X}$ must be more valuable than the average singleton portfolio; that is, $v(\mathcal{X}) \geq \bar U / m$.
-
-% In the main loop, each time the target valuation $v$ is updated using
+\else
+(최적성.) 알고리즘 \ref{ellisDP3}의 출력이 $\mathcal{W}$이며 실제 최적해가 $\mathcal{X}$라고 하자.
+$v(\mathcal{X}) \leq \bar U$이고, 모든 단일 원소 포트폴리오가 가능해므로  $\mathcal{X}$는 평균적인 단일 원서 포트폴리오보다 가치 높을 수 밖에 없다.
+즉,  $v(\mathcal{X}) \geq \bar U / m$이다.
+\fi
+
+\ifen
 Because $\Delta_j(v)$ is rounded down in the recursion relation defined by \eqref{recursionrelationforcostmindp} and \eqref{deltajvdef}, if $j \in \mathcal{W}[j, v]$, then the true value of $(1 - f_j) v\bigl(\mathcal{W}[j-1, v- \Delta_j(v)]\bigr) + f_j t_j$ may exceed the fixed-point valuation $v$ of $\mathcal{W}[j, v]$, but not by more than $10^{-P}$.
-This error accumulates with each school added to $\mathcal{W}$, but the number of additions is at most $m$.
+This error accumulates additively with each school added to $\mathcal{W}$, but the number of additions is at most $m$.
 Therefore, where $v'(\mathcal{W})$ denotes the fixed-point valuation of $\mathcal{W}$ recorded in line \ref{vrecordedhere} of the algorithm,
 $v(\mathcal{W}) - v'(\mathcal{W}) \leq m 10^{-P}$.
-
+\else
+\eqref{recursionrelationforcostmindp}과(와) \eqref{deltajvdef}(으)로 정의된 점화식에 따라 $\Delta_j(v)$를 내림하기 때문에, $j \in \mathcal{W}[j, v]$이면 $(1 - f_j) v\bigl(\mathcal{W}[j-1, v- \Delta_j(v)]\bigr) + f_j t_j$의 실제값은 $\mathcal{W}[j, v]$의 고정소수점값인 $v$보다 클 수 있는데 최대한 $10^{-P}$ 만큼의 차이가 발생하다.
+또한 이 오차는 $\mathcal{W}$에 학교를 추가할 때마다 가산적으로 누적되는데 추가하는 학교의 개수는 최대한 $m$이다.
+따라서 알고리즘의 \ref{vrecordedhere}줄에서 기록하는 $\mathcal{W}$의 값이 $v'(\mathcal{W})$일 때, $v(\mathcal{W}) - v'(\mathcal{W}) \leq m 10^{-P}$이 성립한다.
+\fi
+
+\ifen
 We can define $v'(\mathcal{X})$ analogously as the fixed-point valuation of $\mathcal{X}$ when its elements are added in index order and its valuation is updated and rounded down to the nearest multiple of $10^{-P}$ at each addition in accordance with \eqref{solvemeforvwjvminusj}.
@@
 $v(\mathcal{X}) - v'(\mathcal{X}) \leq m 10^{-P}$. The optimality of $\mathcal{W}$ in the fixed-point environment implies that $v'(\mathcal{W}) \geq v'(\mathcal{X})$. 
-
+\else
+$\mathcal{X}$의 원소를 지표 순서대로 추가했을 때, 각 학교를 추가하면 \eqref{solvemeforvwjvminusj}에 따라 가치를 수정하고 가장 가까운 $10^{-P}$의 곱으로 내림한 값으로 $v'(\mathcal{X})$를 $v'(\mathcal{W})$와 유사하게 정의할 수 있다.
+위와 같은 논리에 따라 $v(\mathcal{X}) - v'(\mathcal{X}) \leq m 10^{-P}$이다.
+또한 $\mathcal{W}$가 고정소수점 환경에서 최적이므로 $v'(\mathcal{W}) \geq v'(\mathcal{X})$임을 알 수 있다.
+\fi
+
+\ifen
 Applying these observations, we have
+\else
+위 관찰을 적용하면 다음 부등식을 도출할 수 있다.
+\fi
 %the fact that $\bar U / m \leq v(\mathcal{X} \leq \bar U$, we have
@@
 \end{equation}
+\ifen
 which establishes the approximation bound. 
-
+\else
+이는 정리에서 말한 근사 계수다.
+\fi
+
+\ifen
 (Computation time.) The bottleneck step is the creation of the lookup table in line \ref{createdlookuptable}, whose size is $m \times |\mathcal{V}|$. Since
+\else
+(계산 시간.) \ref{createdlookuptable}줄에서 표를 채우는 것이 목병 단계이며 표의 크기는 $m \times |\mathcal{V}|$이다. 그러면
+\fi
 \begin{equation}
 |\mathcal{V}| = \bar U \times 10^{P} + 1 = \bar U \times 10^ { \bigl\lceil\log_{10}\left(m ^2/ \varepsilon \bar U\right)\bigr\rceil} + 1
-\leq\frac{m^2}{\varepsilon} \times \text{const.}
+\leq\frac{m^2}{\varepsilon} \times \text{\ifen const.\else 상수\fi}
 \end{equation}
+\ifen
 is $O(m^2/ \varepsilon)$, the time complexity is as promised.
-%To prevent redundant recursive calls to $G(j, v)$, we need to record its values in a dictionary as they are computed; the space required by the dictionary is $O(|\mathcal{V}| m) = O(m^3 / \varepsilon)$.
+\else
+는 $O(m^2/ \varepsilon)$이므로 시간 복잡성을 정리와 같다.
+\fi%To prevent redundant recursive calls to $G(j, v)$, we need to record its values in a dictionary as they are computed; the space required by the dictionary is $O(|\mathcal{V}| m) = O(m^3 / \varepsilon)$.
 \end{proof}
 
-Since these bounds are polynomial in $m$ and $1 / \varepsilon$, Algorithm \ref{ellisDP3} is a fully polynomial-time approximation scheme for Ellis's problem (Vazirani 2001).
-
-Algorithms \ref{ellisDP1} and \ref{ellisDP3} can be written using recursive functions instead; however, since each function references itself \emph{twice,} the function values at each index must be recorded in a lookup table or otherwise memoized to prevent an exponential number of calls from forming on the stack.
-
-\pagebreak
-\ifen \section{Numerical experiments} \else \section{계산적 실험}\fi\label{numericalexperiments}
-In this section, we present the results of numerical experiments designed to test the time complexity results established above.
+\ifen
+Since its time complexity is polynomial in $m$ and $1 / \varepsilon$, Algorithm \ref{ellisDP3} is an FPTAS for Ellis's problem (Vazirani 2001).
+
+Algorithms \ref{ellisDP1} and \ref{ellisDP3} can be written using recursive functions instead of lookup tables.
+However, since each function references itself \emph{twice,} the function values at each index must be recorded in a lookup table or otherwise memoized to prevent an exponential number of calls from forming on the stack.
+\else
+계산 시간이 $m$과 $1 / \varepsilon$의 다항식으로 제한되므로  알고리즘 \ref{ellisDP3}은(는) 엘리스의 문제를 위한 FPTAS다 (Vazirani 2001).
+
+알고리즘 \ref{ellisDP1}과(와) 알고리즘 \ref{ellisDP3}을(를) 표 대신 재귀 함수 기반으로 구성할 수 있다. 그러나 각 재귀 함수가 스스로를 `두번' 호출하므로 지수적으로 늘어나는 호출 수를 방화하기 위해 각 지표에 대응하는 함수값을 표에 저장하거나 다른 식으로 기록해야 한다.
+\fi
+
+%\pagebreak
+\ifen \section{Numerical experiments} \else \section{계산 실험}\fi\label{numericalexperiments}
+\ifen
+In this section, we present the results of numerical experiments designed to test the practical efficacy of the algorithms derived above. 
+\else
+본 절에서 위에서 도출한 알고리즘의 현실성을 탐구하고자 계산 실험의 결과를 제시한다.
+\fi
+
+\ifen \subsection{Implementation notes} \else \subsection{알고리즘의 실행}\fi
+\ifen
+We chose to implement our algorithms in the Julia language (v1.7.0) because its system of type inference allows the compiler to optimize for differential cases such as when the $t_j$-values are integers or floating-point numbers.
+Julia also offers convenient macros for parallel computing, which enabled us to solve more and larger problems in the benchmark (Bezanson et al.
+2017).
+We made extensive use of the DataStructures.jl package (v0.18.11).
+The code is available at \url{https://github.com/maxkapur/OptimalApplication}.
+\else
+모든 알고리즘을 쥴리아 (Julia, v1.7.0) 언어로 실행했다.
+쥴리아는 $t_j$-값이 정수 또는 부동소수점 수와 같은 특수한 경우에 대해 컴파일러가 자동적으로 최적화하는 자료형 추론(type inference) 기능이 있어서 유용한 선택이었다.
+또한 병렬 연산을 쉽게 관리할 수 있는 매크로를 제공하므로 계산 실험의 규모를 증가할 수 있었다 (Bezanson 외 2017).
+DataStructures.jl 패키지(v0.18.11)도 활용했다.
+실험 코드는  \url{https://github.com/maxkapur/OptimalApplication}에서 공유한다.
+\fi
+
+\ifen
+The dynamic programs, namely Algorithms \ref{ellisDP1} and \ref{ellisDP3}, were implemented using recursive functions and dictionary memoization rather than a full lookup table.
+Our implementation of Algorithm \ref{ellisDP3} also differed from that described in subsection \ref{fptashead} in that we represented portfolio valuations in \emph{binary} rather than decimal, with the definitions of $P$ and $\mathcal{V}$ modified accordingly, and instead of fixed-point numbers, we worked in integers by multiplying each $t_j$-value by $2^P$.
+These modifications yielded a substantial performance improvement without changing the fundamental algorithm design or complexity analysis.
+\else
+동적 계획 해법인 \ref{ellisDP1}과(와) 알고리즘 \ref{ellisDP3}을(를) 표 대신 사전 기록 및 재귀 함수로 실행했다.
+알고리즘 \ref{ellisDP3}은(는) \ref{fptashead}항에서 설명한 것과 달리, 포트폴리오의 가치를 십진수 대신 `이진수'로 나타냈으며 $P$와  $\mathcal{V}$의 정의를 적절히 수정했다.
+그리고 고정소수점 대신 $t_j$-값을 $2^P$으로 곱해서 정수로 구성했다.
+이런 변경은 기본 알고리즘 설계와 계산 시간 분석은 변하지 않는데 성능은 많이 개선한다.
+\fi
 
@@
 \def\nmarkets{20}
+\ifen
 The experimental procedure was as follows.
@@
 A typical instance is shown in Figure \ref{samplemarket}.
-
+\else
+실험 방법은 다음과 같다.
+먼저, 가산 시장을 생성하기 위해 $t_j$-값을 평균이 10인 지수 분포에서 동립적으로 관측하고 정수로 올림했다.
+그다음에, $t_j$와 $f_j$가 서로 부분적으로 반비례하게 만들도록 $f_j = 1 / (t_j + 10\,Q)$로 정의했다.
+단, $Q$는 $[0, 1)$에서 균일한 확률로 관측한다.
+알마의 문제를 고려하는 첫번째 실험에서, 모든 $g_j = 1$ 그리고 $H = h = \lfloor m/ 2 \rfloor$으로 처리했다.
+엘리스의 문제를 고려하는 두번째 실험에서, 모든  $g_j$를 $\{5, \dots, 10\}$에서 균일한 확률로 관측했으며 $H = \lfloor \frac{1}{2} \sum g_j \rfloor$으로 처리했다.
+그림 \ref{samplemarket}에서 전형적인 인스턴스가 나타난다.
+\fi
+
+\ifen
+The experimental variables were the market size $m$, the choice of algorithm, and (for Algorithm \ref{ellisDP3}) the tolerance $\varepsilon$.
 For each combination of the experimental variables, we generated \nmarkets~markets, and the optimal portfolio was computed three times, with the fastest of the three repetitions recorded as the computation time.
@@
 Therefore, each cell of each table below represents a statistic over {\the\numexpr 3 * \nmarkets \relax}~computations.
-Where applicable, we do not count the time required to sort the entries of $t$.
+Where applicable, we did not count the time required to sort the entries of $t$.
+\else
+실험 변수로는 시장의 크기 $m$, 해법 선택, 그리고 (알고리즘 \ref{ellisDP3}인 경우) 허용치 $\varepsilon$을 고려했다.
+실험 변수의 각 조합에 대해,  \nmarkets 개의 시장을 생성했으며 최적 포트폴리오를 3번 계산 한 것 중 가장 빠른 것으로 계산 시간으로 기록했다.
+\nmarkets 개의 시장에 대한 평균과 평균편차를 표에 나타낸다.
+따라서 아래 표의 각 칸은 {\the\numexpr 3 * \nmarkets \relax}개의 계한에 대한 통계다.
+$t$를 배열하는 시간은 고려하지 않았다.
+\fi
 
 \ifen \subsection{Summary of results} \else \subsection{결과 정리}\fi
+\ifen
 In our first experiment, we compare the performance of Algorithm \ref{algorithmforlargeh} for homogeneous-cost markets of various sizes when the set of candidate schools $\mathcal{C}$ is stored as a list and as a binary max heap ordered by the $f_j \bar t_j$-values.
@@
 Although using a max heap increases the nominal runtime from $O(hm)$ to $O(hm \log m)$, we hypothesized that in typical instances, the order of the heap would change only slightly between iterations, yielding an overall savings because of the lower costs of extracting the maximal school.
-Our results indicate that due to its lower overhead cost, the list implementation is faster; however, the ratio between the average times of the two implementations stays roughly constant at about 1.5, suggesting that a more effective heap implementation could be competitive in certain classes of problem instances.
+Our results indicate that the list implementation is faster.
+However, the ratio between the average times of the two implementations stays roughly constant at about 1.5, suggesting that a more effective heap implementation could be competitive in certain classes of problem instances.
 Overall, the rate of growth is quadratic in $m$, which accords with the $O(hm)$ time complexity result of Theorem \ref{validityofalmaalgorithm}.
-
+\else
+첫번째 실험에서, 동일한 지원 비용으로 정의된 크기가 다양한 시장에 대해 알고리즘 \ref{algorithmforlargeh}의 성능을 고려했다.
+이를 $\mathcal{C}$를 목록 혹은 $f_j \bar t_j$-값에 따라 배열된 힙 구현으로 구성하는 경우로 나눴으며 실험 결과는 표 \ref{experiment1results}에서 등장한다.
+이론적으로 최대 힙을 사용하면 계산 시간이 $O(hm)$에서 $O(hm \log m)$으로 늘어나지만, 전형적인 인스턴스에서 반복단계 사이에 힙의 순서가 많이 변하지 않으므로 최대 학교를 구하는 시간을 줄임에 따라 전체적인 계산 시간이 감소할 것으로 가설했다.
+실험 결과에 따라 목록 구현이 더 빠른 해법이었다.
+그러나, 두개의 방법 사이에 계산 시간 비율은 1.5같은 상수에 가까우므로 개선된 힙 해법이 어떤 문제 인스턴스에 대해 더 빠를 가능성이 있어 보인다.
+전체적으로 계산 시간은 $m$의 제곱으로 늘아나며 이는 정리 \ref{validityofalmaalgorithm}에서 말한 $O(hm)$-시간과 동의한다.
+\fi
+
+\ifen
 In the second experiment, we turn to the general problem, and compare the performance of the exact algorithms (Algorithms \ref{ellisbnb} and \ref{ellisDP1}) and the approximation scheme (Algorithm \ref{ellisDP3}) at tolerances 0.5 and 0.05.
-The results, which appear in Table \ref{experiment2results}, agree with the time complexity analyses presented above.
-Overall, we found the exact dynamic program to be the fastest algorithm, a result that echoes the results of computational studies on knapsack problems.
+The results, which appear in Table \ref{experiment2results}, broadly agree with the time complexity analyses presented above.
+Overall, we found the exact dynamic program to be the fastest algorithm, while the FPTAS was rather slow, a result that echoes the results of computational studies on knapsack problems (Martello and Toth 1990, \S\,2.10).
 The branch-and-bound algorithm proved impractical for even medium-sized instances.
 %cite computational studies knapsack
+\else
+두번째 실험에서 일반적 문제를 고려한다.
+정확한 해법 (알고리즘 \ref{ellisbnb} 및 알고리즘 \ref{ellisDP1}) 그리고 허용치를 0.5와 0.05로 설정한 근사 해법 (알고리즘 \ref{ellisDP3})의 성능을 비교한다.
+실험 결과는 표 \ref{experiment1results}에서 나타나며 위에서 도출한 시간 복잡성 결과와 대략 같다.
+정확한 동적계획 해법이 전책적으로 가장 빠르며 FPTAS가 약간 느린 것으로 나온 결과는 배낭 문제 해법을 비교한 기존 연구와 비슷하다 (Martello와 Toth 1990, \S\,2.10).
+분지한계법은 크기가 중간인 인스턴스에 대해서도 비효율적이었다.
+\fi
 
@@
 The table shows the average time (standard deviation) in milliseconds over the \nmarkets~instances.
-\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복해서 그 중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (평균편차) 시간을 밀리초 단위로 나타난다.\fi}
+\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복해서 그 중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (평균편차) 시간이 밀리초 단위로 나타난다.\fi}
 
@@
   \caption{\label{samplemarket}
- \ifen A typical randomly-generated instance with $m=60$ schools and its optimal application portfolio.
+ \ifen A typical randomly-generated instance with $m=64$ schools and its optimal application portfolio.
 The application costs $g_j$ were drawn uniformly from $\{5, \dots, 10\}$.
 The optimal portfolio was computed using Algorithm \ref{ellisDP1}.
- \else $m=60$개의 학교로 구성된 전형적인 무작위로 생성한 인스턴스와 해당 최적 포트폴리오. 지원 비용 $g_j$는 $\{5, \dots, 10\}$에서 균일한 확률로 선정했으며 기호의 넓이와 비례된다. 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
+ \else $m=64$개의 학교로 구성된 전형적인 무작위로 생성한 인스턴스와 해당 최적 포트폴리오. 지원 비용 $g_j$는 $\{5, \dots, 10\}$에서 균일한 확률로 선정했으며 기호의 넓이와 비례된다. 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
 \end{figure}
 
-\ifen \subsection{Implementation notes} \else \subsection{알고리즘의 이행}\fi
-We chose to implement our algorithms in the Julia language (v1.7.0) because its system of type inference allows the compiler to optimize for differential cases such as when the $t_j$-values are integers or floating-point numbers.
-Julia also offers convenient macros for parallel computing, which enabled us to solve more and larger problems in the benchmark (Bezanson et al.
-2017).
-We made extensive use of the DataStructures.jl package (v0.18.11).
-The code is available at \url{https://github.com/maxkapur/OptimalApplication}.
-
-The dynamic programs, namely Algorithms \ref{ellisDP1} and \ref{ellisDP3}, were implemented using recursive functions and memoization rather than a full lookup table.
-Our implementation of Algorithm \ref{ellisDP3} also differed from that described in section \ref{fptashead} in that we represented portfolio valuations in \emph{binary} rather than decimal, with the definitions of $P$ and $\mathcal{V}$ modified accordingly, and instead of fixed-point numbers, we worked in integers by multiplying each $t_j$-value by $2^P$.
-These modifications yielded a substantial performance improvement without changing the fundamental algorithm design or complexity analysis.
-
-\pagebreak
+%\pagebreak
 \ifen \section{Conclusion} \else \section{결론}\fi\label{conclusion}

### 2022-03-16 paper/paper.tex
finish translations so far; rm false statement about submodularity
better heap implementation, docstrings, readme

@@
 
-\ifen
-\else 본 논문은 \url{https://github.com/maxkapur/OptimalApplication}에서 전문을 한글로 공유한다. \fi
+\ifen 본 논문은 \url{https://github.com/maxkapur/OptimalApplication}에서 전문을 한글로 공유한다. \fi
 \end{abstract}
@@
 \ifen 
+It is also possible to store $\mathcal{C}$ as a binary max heap rather than a list.
+The heap is ordered according to the criterion $i \geq j \iff f_i t_i \geq f_j t_j$, and by draining the heap and reheapifying at the end of each iteration, the computation time remains $O(hm)$.
 In our numerical experiments, whose results are reported in section \ref{numericalexperiments},
-we also explore the possibility of storing $\mathcal{C}$ as a binary max heap rather than a list.
-The heap is ordered according to the criterion $i \geq j \iff f_i t_i \geq f_j t_j$.
-Nominally, using a heap increases the cost of the main loop from $O(h m)$ to $O(hm \log m)$ because the heap is rebalanced when each $t_j$-value is updated.
-However, typical problem instances do not achieve this upper bound because the order of the $f_j t_j$-values changes only slightly between iterations.
-The cost of updating each $t_j$-value can be reduced to unit time using a Fibonacci heap (Fredman and Tarjan 1987), yielding the same overall computation time.
-\else
-(\ref{numericalexperiments}절에서 결과를 제시할) 수리 실험에서 $\mathcal{C}$를 목록 대신 이진 최대 힙 구현으로 저장하는 방법도 고려한다.
-단, 힙의 순서는 $i \geq j \iff f_i t_i \geq f_j t_j$의 조건으로 정의한다.
-기술적으로, $t_j$-값을 수정할 때마다 힙의 구조를 다시 복구해야 하므로 이는 주 반복단계의 계산 비용을 $O(h m)$에서 $O(hm \log m)$으로 증가시킨다.
-그러나 전형적인 시장에서는, 반복단계 사이의 $f_j t_j$-값의 순서가 크게 변하지 않으므로 이 상한은 자주 발생하지 않는다.
-또한 피보나치 힙을 사용하면 $t_j$-값을 수정하는 시간을 단위 시간으로 줄여서 동등한 계산 시간을 얻어 낼 수 있다 (Fredman and Tarjan 1987).
+we found the heap implementation to be faster than the list implementation. 
+\else
+$\mathcal{C}$는 목록 대신 이진 최대 힙 구현으로 저장하는 방법도 고려한다.
+단, 힙의 순서는 $i \geq j \iff f_i t_i \geq f_j t_j$의 조건으로 정의하며, 각 반복단계의 마지막 단계에서 힙을 비우고 다시 힙화(heapify)하면 $O(hm)$ 계산 시간을 유지할 수 있다.
+(\ref{numericalexperiments}절에서 결과를 제시할) 수리 실험에서 힙 구현으로 구성한 알고리즘은 목록보다 빨랐다.
 \fi
@@
 The results appear in Table \ref{experiment1results}.
-Although using a max heap increases the nominal runtime from $O(hm)$ to $O(hm \log m)$, we hypothesized that in typical instances, the order of the heap would change only slightly between iterations, yielding an overall savings because of the lower costs of extracting the maximal school.
 Our results indicate that the list implementation is faster.
@@
 이를 $\mathcal{C}$를 목록 혹은 $f_j \bar t_j$-값에 따라 배열된 힙 구현으로 구성하는 경우로 나눴으며 실험 결과는 표 \ref{experiment1results}에서 등장한다.
-이론적으로 최대 힙을 사용하면 계산 시간이 $O(hm)$에서 $O(hm \log m)$으로 늘어나지만, 전형적인 인스턴스에서 반복단계 사이에 힙의 순서가 많이 변하지 않으므로 최대 학교를 구하는 시간을 줄임에 따라 전체적인 계산 시간이 감소할 것으로 가설했다.
 실험 결과에 따라 목록 구현이 더 빠른 해법이었다.
@@
 \fi
+
+%\ifen
+%In our first experiment, we compare the performance of Algorithm \ref{algorithmforlargeh} for homogeneous-cost markets of various sizes when the set of candidate schools $\mathcal{C}$ is stored as a list and as a binary max heap ordered by the $f_j \bar t_j$-values.
+The results appear in Table \ref{experiment1results}.
+Our results indicate that the heap implementation is faster.
+Overall, the rate of growth is quadratic in $m$, which accords with the $O(hm)$ time complexity result of Theorem \ref{validityofalmaalgorithm}.
+%\else
+%첫번째 실험에서, 동일한 지원 비용으로 정의된 크기가 다양한 시장에 대해 알고리즘 \ref{algorithmforlargeh}의 성능을 고려했다.
+이를 $\mathcal{C}$를 목록 혹은 $f_j \bar t_j$-값에 따라 배열된 힙 구현으로 구성하는 경우로 나눴으며 실험 결과는 표 \ref{experiment1results}에서 등장한다.
+실험 결과에 따라 힙 구현이 더 빠른 해법이었다.
+전체적으로 계산 시간은 $m$의 제곱으로 늘아나며 이는 정리 \ref{validityofalmaalgorithm}에서 말한 $O(hm)$-시간과 동의한다.
+%\fi
 
@@
  \fi
-%16         & \phantom{000}0.0712748          &    (\phantom{000}0.00755644)              & \phantom{000}0.0977929      &     (\phantom{000}0.00367772)            \\
-%64         & \phantom{000}0.95445\phantom{00}          &    (\phantom{000}0.0822217\phantom{0})              & \phantom{000}1.28911\phantom{00}      &     (\phantom{000}0.0517071\phantom{0})            \\
-%256        & \phantom{00}14.8587\phantom{000}          &    (\phantom{000}0.725739\phantom{00})              & \phantom{00}20.3465\phantom{000}      &     (\phantom{000}0.708838\phantom{00})            \\
-%1024       & \phantom{0}287.057\phantom{0000}          &    (\phantom{00}81.9166\phantom{0000})              & \phantom{0}390.925\phantom{0000}      &     (\phantom{0}110.501\phantom{00000})            \\
-%4096       & 4826.75\phantom{00000}          &    (1432.85\phantom{000000})              & 7304.67\phantom{00000}      &     (2071.38\phantom{000000})
    16 &     0.07 &     (0.01) &        0.09 &    (0.02) \\

### 2022-03-16 benchmark/benchmark.jl

@@ -7,7 +7,7 @@ using Statistics
 using Base.Threads
 import Printf: @sprintf
 
-fullscale = true
+fullscale = false
 
 # A long benchmark; tweak parameters with caution.
 # Set fullscale = false to run a smaller benchmark to check formatting etc.
@@ -44,7 +44,7 @@ function benchmark1()
     M = fullscale ? 4 .^ (2:6) : [5, 10]
 
     sizes = zeros(Int64, n_markets, length(M))
-    times_dict = zeros(Float64, n_markets, length(M))
+    times_list = zeros(Float64, n_markets, length(M))
     times_heap = zeros(Float64, n_markets, length(M))
 
     @threads for i in 1:n_markets
@@ -52,12 +52,12 @@ function benchmark1()
         for (j, m) in enumerate(M)
             mkt = randSCM(m)
             sizes[i, j] = m
-            times_dict[i, j] = minimum(@elapsed applicationorder(mkt; datastructure = :dict) for r in 1:n_reps)
+            times_list[i, j] = minimum(@elapsed applicationorder(mkt; datastructure = :list) for r in 1:n_reps)
             times_heap[i, j] = minimum(@elapsed applicationorder(mkt; datastructure = :heap) for r in 1:n_reps)
         end
     end
 
-    df = DataFrame("m" => sizes[:], "time_dict" => times_dict[:], "time_heap" => times_heap[:])
+    df = DataFrame("m" => sizes[:], "time_list" => times_list[:], "time_heap" => times_heap[:])
 
     kv_pairs = Pair[]
     for c in names(df)

### 2022-03-16 src/FastSolvers.jl

@@ -5,17 +5,18 @@ Contains a college's admissions probability `f`, utility value `t`,
 their product `ft`, and `1 - f = omf`. Used only by `applicationorder()`.
 """
 struct College
+    j::Int
     f::Real
     t::Real
     ft::Real
     omf::Real
 
-    function College(f::Real, t::Real, ft::Real, omf::Real)
-        return new(f, t, ft, omf)
+    function College(j::Int, f::Real, t::Real, ft::Real, omf::Real)
+        return new(j, f, t, ft, omf)
     end
 
-    function College(f::Real, t::Real, omf::Real)
-        return new(f, t, f * t, omf)
+    function College(j::Int, f::Real, t::Real, omf::Real)
+        return new(j, f, t, f * t, omf)
     end
 end
 
@@ -30,71 +31,79 @@ Produce the optimal order of application for the market `mkt` having identical
 application costs and the corresponding portfolio valuations.
 """
 function applicationorder(
-    mkt::SameCostsMarket;
-    datastructure = :heap::Symbol, verbose=false::Bool)::Tuple{Vector{Int64},Vector{Float64}}
+        mkt::SameCostsMarket;
+        datastructure = :heap::Symbol
+    )::Tuple{Vector{Int64},Vector{Float64}}
 
     apporder = zeros(Int, mkt.h)
     v = zeros(mkt.h)
     if datastructure == :heap
-        mkt_heap = MutableBinaryMaxHeap{College}()
-        for j in 1:mkt.m
-            push!(mkt_heap, College(mkt.f[j], mkt.t[j], mkt.ft[j], mkt.omf[j]))
-        end
-
-        onheap = Set(1:mkt.m)
-    
+        mkt_heap = BinaryMaxHeap{College}(collect(College(j, mkt.f[j], mkt.t[j], mkt.ft[j], mkt.omf[j]) for j in 1:mkt.m))
         for j in 1:mkt.h
-            if verbose
-                K = sort(collect(onheap))
-                println("C = ", K)
-                println("f = ", [mkt_heap[j].f for j in K])
-                println("t = ", [mkt_heap[j].t for j in K])
-                println("ft= ", [mkt_heap[j].ft for j in K])
-            end
-
-            c_k, k = top_with_handle(mkt_heap)
+            c_k = first(mkt_heap)
             v[j] = get(v, j - 1, 0) + c_k.ft
-
-            pop!(mkt_heap)
-
-            apporder[j] = k
-            delete!(onheap, k)
-            for i in onheap
-                if i < k
[…]

### 2022-03-16 src/MarketIO.jl

@@ -32,7 +32,15 @@ end
 
 """
 Contains information about a college application market with identical
-application costs.
+application costs. Fields:
+
+    `m`: Number of schools
+    `f`: Vector of admissions probabilities
+    `t`: Vector of utilitied values (must be sorted)
+    `h`: Number of schools student is allowed to apply to
+    `ft = f .* t`
+    `omf = 1 .- f`
+    `perm = sortperm(perm)`: Currently a placeholder
 """
 struct SameCostsMarket
     m::Integer
@@ -63,8 +71,19 @@ end
 
 
 """
+    VariedCostsMarket
+
 Contains information about a college application market with varying
-application costs.
+application costs. Fields:
+
+    `m`: Number of schools
+    `f`: Vector of admissions probabilities
+    `t`: Vector of utilitied values (must be sorted)
+    `g`: Vector of application costs
+    `H`: Budget to spend on applications
+    `ft = f .* t`
+    `omf = 1 .- f`
+    `perm = sortperm(perm)`: Currently a placeholder
 """
 struct VariedCostsMarket
     m::Integer

### 2022-03-16 test/runtests.jl

@@ -34,7 +34,7 @@ end
             X, vX = optimalportfolio_enumerate(mkt)
             sort!(X)
             W, vW = applicationorder(mkt; datastructure = :heap)
-            Y, vY = applicationorder(mkt; datastructure = :dict)
+            Y, vY = applicationorder(mkt; datastructure = :list)
             @test X == sort(W)
             @test vX ≈ last(vW)
             @test X == sort(Y)
@@ -48,7 +48,7 @@ end
         @testset "Exact algorithms" begin
             for _ in 1:n_markets
                 mkt = randVCM(m)
-            
+
                 X, vX = optimalportfolio_enumerate(mkt)
                 sort!(X)
                 W, VW = optimalportfolio_valuationtable(mkt)
@@ -56,7 +56,7 @@ end
                 # Without memoization
                 Z, vZ = optimalportfolio_dynamicprogram(mkt, false)
                 B, vB = optimalportfolio_branchbound(mkt)
-            
+
                 @test X == sort(W)
                 @test vX ≈ VW[mkt.m, mkt.H]
                 @test X == sort(Y)

### 2022-03-16 benchmark/benchmark.jl
updated benchmark data

@@ -7,7 +7,7 @@ using Statistics
 using Base.Threads
 import Printf: @sprintf
 
-fullscale = false
+fullscale = true
 
 # A long benchmark; tweak parameters with caution.
 # Set fullscale = false to run a smaller benchmark to check formatting etc.

### 2022-03-21 paper/paper.tex
updated benchmark data
isolate functions from dynamic programs, fix return bug, bnb improvements

@@
 			}
-			\lIf{$n'$ is not a leaf}{add $n'$ to $\mathfrak{T}$}
+			\lIf{$v_{\mathrm{LP}}^*[n] > L$ and $n'$ is not a leaf}{add $n'$ to $\mathfrak{T}$}
 		}
@@
 			}
-			\lIf{$n'$이 잎 마디 아님}{$\mathfrak{T}$에 $n'$을 추가한다}
+			\lIf{$v_{\mathrm{LP}}^*[n] > L$이고 $n'$이 잎 마디 아님}{$\mathfrak{T}$에 $n'$을 추가한다}
 		}
@@
 This means that significant fathoming operations do not occur until the algorithm has explored deep into the tree, at which point the bulk of the computational effort has already been exhausted.
-In our numerical experiments, Algorithm \ref{ellisbnb} was ineffective on instances larger than about $m=20$ schools, and therefore it does not represent a significant improvement over na\"ive enumeration.
+In our numerical experiments, Algorithm \ref{ellisbnb} was ineffective on instances larger than about $m=35$ schools, and therefore it does not represent a significant improvement over na\"ive enumeration.
 \else
@@
 이는 해법이 나무에 어느 정도 깊게 파고들어야 마디를 제외시킬 수 있음을 의미하며, 그때 계산 노력의 대부분은 이미 쏟은 것이다.
-수리 실험에서, 알고리즘 \ref{ellisbnb}은(는) 약 $m=20$개의 학교를 넘는 인스턴스에 대해 비효율적이었으며 이는 단순한 열거법에 비해 큰 개선이 아니다.
+수리 실험에서, 알고리즘 \ref{ellisbnb}은(는) 약 $m=35$개의 학교를 넘는 인스턴스에 대해 비효율적이었으며 이는 단순한 열거법에 비해 큰 개선이 아니다.
 \fi
@@
  \fi
-   16 &     0.07 &     (0.01) &        0.09 &    (0.02) \\
-   64 &     0.88 &     (0.16) &        1.18 &    (0.21) \\
-  256 &    12.51 &     (2.20) &       18.39 &    (2.90) \\
- 1024 &   281.79 &    (58.56) &      387.97 &   (96.41) \\
- 4096 &  4661.75 &  (1152.54) &     7218.04 & (1733.91)                  
+    16 &       0.07 &    (0.02) &      0.06 &    (0.02) \\
+    64 &       0.72 &    (0.21) &      0.73 &    (0.21) \\
+   256 &      10.36 &    (2.88) &     10.74 &    (3.34) \\
+  1024 &     182.14 &   (47.83) &    181.95 &   (29.10) \\
+  4096 &    2858.49 &  (123.99) &   2994.55 &  (213.49)            
 \end{tabular}

### 2022-03-21 benchmark/benchmark.jl

@@ -7,7 +7,7 @@ using Statistics
 using Base.Threads
 import Printf: @sprintf
 
-fullscale = false
+fullscale = true
 
 # A long benchmark; tweak parameters with caution.
 # Set fullscale = false to run a smaller benchmark to check formatting etc.
@@ -33,7 +33,7 @@ end
 
 randVCM(m) = VariedCostsMarket(make_correlated_market(m)...)
 function randSCM(m)
-    f, t, g, H = make_correlated_market(m)
+    f, t, _, _ = make_correlated_market(m)
     return SameCostsMarket(f, t, m ÷ 2)
 end
 
@@ -72,7 +72,7 @@ end
 function benchmark2()
     printheader("Benchmark 2: Heterogeneous-cost algorithms")
     M = fullscale ? 2 .^ (3:9) : [5, 10]
-    bnbcutoff = fullscale ? 20 : 6
+    bnbcutoff = fullscale ? 33 : 6
     epsilons = [0.5, 0.05]
 
     dtype = Union{Float64,Missing}

### 2022-03-21 src/FastSolvers.jl

@@ -138,7 +138,23 @@ function optimalportfolio_valuationtable(mkt::VariedCostsMarket)::Tuple{Vector{I
 end
 
 
-const IntTypes = [Int8, Int16, Int32, Int64]
+# Used by dynamic program below
+function V_recursor!(V_dict::Dict{Tuple{Int,Int},Float64}, j::Int, h::Int, mkt::VariedCostsMarket)
+    haskey(V_dict, (j, h)) && return V_dict[(j, h)]
+
+    if j == 0 || h == 0
+        return 0.0
+    elseif h < mkt.g[j]
+        push!(V_dict, (j, h) => V_recursor!(V_dict, j - 1, h, mkt))
+        return V_dict[(j, h)]
+    else
+        push!(V_dict, (j, h) => max(
+            V_recursor!(V_dict, j - 1, h, mkt),
+            mkt.omf[j] * V_recursor!(V_dict, j - 1, h - mkt.g[j], mkt) + mkt.ft[j]
+        ))
+        return V_dict[(j, h)]
+    end
+end
 
 
 
@@ -149,38 +165,73 @@ Use the dynamic program on application costs to produce the optimal portfolio `X
 value `v` for the market `mkt` with varying application costs. 
 """
 function optimalportfolio_dynamicprogram(mkt::VariedCostsMarket)::Tuple{Vector{Int},Float64}
-    V_dict = Dict{Tuple{Int, Int}, Float64}()
-
-    function V(j, h)
-        haskey(V_dict, (j, h)) && V_dict[(j, h)]
-
-        if j == 0 || h == 0
-            return 0.0
-        elseif h < mkt.g[j]
-            push!(V_dict, (j, h) => V(j - 1, h))
-            return V_dict[(j, h)]
-        else
-            push!(V_dict, (j, h) => max(
-                V(j - 1, h),
-                mkt.omf[j] * V(j - 1, h - mkt.g[j]) + mkt.ft[j]
-            ))
-            return V_dict[(j, h)]
-        end
-    end
-
+    V_dict = Dict{Tuple{Int,Int},Float64}()
+    v = V_recursor!(V_dict, mkt.m, mkt.H, mkt)
+    
     h = mkt.H
     X = Int[]
 
-    v = V(mkt.m, mkt.H)
-
     for j in mkt.m:-1:1
-        if V(j - 1, h) < V(j, h)
+        if V_recursor!(V_dict, j - 1, h, mkt) < V_recursor!(V_dict, j, h, mkt)
[…]

### 2022-03-21 src/MarketIO.jl

@@ -90,7 +90,7 @@ struct VariedCostsMarket
     f::Vector{Float64}
     t::Vector{Int}
     g::Vector{Int}
-    H::Real
+    H::Int
     ft::Vector{Float64}      # = f .* t
     omf::Vector{Float64}     # = 1 .- f
     perm::Vector{Int}

### 2022-03-21 src/VariedCostsBnB.jl

@@ -10,7 +10,6 @@ mutable struct Node
     H̄::Int
     v_I::Float64
     v_LP::Float64
-    # isleaf flag is actually superfluous: Can just check isempty(N)
 
     function Node(I::Set, N::Set, t̄::Dict, H̄::Int, v_I::Float64, mkt::VariedCostsMarket)
         # For generating a new node with its LP relaxation value and empty child set
@@ -119,8 +118,8 @@ end
 Use the branch-and-bound algorithm to produce the optimal portfolio for the
 market `mkt` with varying application costs. Intractable for large markets. 
 """
-function optimalportfolio_branchbound(mkt; maxit = 1000000::Int, verbose = false::Bool)
-    mkt.m ≥ 30 && @warn "Branch and bound is slow for large markets"
+function optimalportfolio_branchbound(mkt; maxit = 100000::Int, verbose = false::Bool)
+    mkt.m ≥ 33 && @warn "Branch and bound is slow for large markets"
 
     C = Set(1:length(mkt.t))
 
@@ -170,7 +169,9 @@ function optimalportfolio_branchbound(mkt; maxit = 1000000::Int, verbose = false
             end
 
             # isempty(child.N) || push!(treekeys, push!(tree, child))
-            isempty(child.N) || push!(tree, hash(child) => child)
+            if child.v_LP > LB && !isempty(child.N)
+                push!(tree, hash(child) => child)
+            end
         end
 
         if newLB

### 2022-03-21 test/runtests.jl

@@ -16,7 +16,7 @@ end
 
 randVCM(m) = VariedCostsMarket(make_correlated_market(m)...)
 function randSCM(m)
-    f, t, g, H = make_correlated_market(m)
+    f, t, _, _ = make_correlated_market(m)
     return SameCostsMarket(f, t, m ÷ 2)
 end
 
@@ -27,36 +27,36 @@ end
         for _ in 1:n_markets
             mkt = randSCM(m)
 
-            X, vX = optimalportfolio_enumerate(mkt)
-            sort!(X)
-            W, vW = applicationorder_heap(mkt)
-            Y, vY = applicationorder_list(mkt)
-            @test X == sort(W)
-            @test vX ≈ last(vW)
-            @test X == sort(Y)
-            @test vX ≈ last(vY)
+            X_enum, vX_enum = optimalportfolio_enumerate(mkt)
+            sort!(X_enum)
+            X_heap, vX_heap = applicationorder_heap(mkt)
+            X_list, vX_list = applicationorder_list(mkt)
+            @test X_enum == sort(X_heap)
+            @test vX_enum ≈ last(vX_heap)
+            @test X_enum == sort(X_list)
+            @test vX_enum ≈ last(vX_list)
         end
     end
 
     @testset verbose = true "Varied app. costs" begin
-        m = 10
+        m = 16
 
         @testset "Exact algorithms" begin
             for _ in 1:n_markets
                 mkt = randVCM(m)
 
-                X, vX = optimalportfolio_enumerate(mkt)
-                sort!(X)
-                W, VW = optimalportfolio_valuationtable(mkt)
-                Y, vY = optimalportfolio_dynamicprogram(mkt)
-                B, vB = optimalportfolio_branchbound(mkt)
+                X_enum, vX_enum = optimalportfolio_enumerate(mkt)
+                sort!(X_enum)
+                X_valtable, VX_valtable = optimalportfolio_valuationtable(mkt)
+                X_dp, vX_dp = optimalportfolio_dynamicprogram(mkt)
+                X_bnb, vX_bnb = optimalportfolio_branchbound(mkt)
 
-                @test X == sort(W)
-                @test vX ≈ VW[mkt.m, mkt.H]
-                @test X == sort(Y)
-                @test vX ≈ vY
-                @test X == sort(B)
-                @test vX ≈ vB
+                @test X_enum == sort(X_valtable)
+                @test vX_enum ≈ VX_valtable[mkt.m, mkt.H]
[…]

### 2022-03-22 paper/paper.tex
improved plot quality, type assertions

@@
 \ifen
-Pick any $h$ and let $m = 2h$. For a small constant $\varepsilon \in (0, 1)$, construct the market as follows.
+Pick any $h$ and let $m = 2h$. For a small constant $\varepsilon \in (0, 1)$, define the market as follows.
 \else
@@
 알고리즘의 출력은 $\mathtt{X} = [4, 2, 8, 1, 7, 3, 5, 6]$이며 최적 $h$-포트폴리오는 그의 첫 $h$개의 원소로 이루어진다.
-표 \ref{planetsdata}에서 등장하는 ``지원 순위'' 자료는 $\mathtt{X}$의 역순열(inverse permutation)이며 이는 해당 학교가 최적 포트폴리오에 포함되는 최소 $h$-값을 의미한다.
+표 \ref{planetsdata}에서 등장하는 ``지원 순위'' 자료는 $\mathtt{X}$의 역순열(inverse permutation)이며 이는 해당 학교가 최적 포트폴리오에 포함되는 최소한 $h$-값을 의미한다.
 그림 \ref{planetsplot}은 최적 포트폴리오의 가치를 $h$의 함수로 나타낸다.
@@
 \begin{table}[h!] \centering
+\small
 \begin{tabular}{r|lcccc}
@@
 \end{tabular}
-\caption{\label{planetsdata}
+\caption{\label{planetsdata} \normalsize
 \ifen College data and optimal application portfolios for a fictional market with $m=8$ schools.
-By the nestedness property (Theorem \ref{nestedapplication}), the optimal portfolio when the application limit is $h$ consists of the $h$ schools having priority number $h$ or less.
+By the nestedness property (Theorem \ref{nestedapplication}), the optimal portfolio when the application limit is $h$ consists of the $h$ schools having priority $h$ or less.
 \else
@@
 Our results indicate that the list implementation is faster.
-However, the ratio between the average times of the two implementations stays roughly constant at about 1.5, suggesting that a more effective heap implementation could be competitive in certain classes of problem instances.
+However, the ratio between the average times of the two implementations stays roughly constant, suggesting that a more effective heap implementation could be competitive in certain classes of problem instances.
 Overall, the rate of growth is quadratic in $m$, which accords with the $O(hm)$ time complexity result of Theorem \ref{validityofalmaalgorithm}.
@@
 실험 결과에 따라 목록 구현이 더 빠른 해법이었다.
-그러나, 두개의 방법 사이에 계산 시간 비율은 1.5같은 상수에 가까우므로 개선된 힙 해법이 어떤 문제 인스턴스에 대해 더 빠를 가능성이 있어 보인다.
+그러나, 두개의 방법 사이에 계산 시간 비율은 거의 상수이므로 개선된 힙 해법이 어떤 문제 인스턴스에 대해 더 빠를 가능성이 있어 보인다.
 전체적으로 계산 시간은 $m$의 제곱으로 늘아나며 이는 정리 \ref{validityofalmaalgorithm}에서 말한 $O(hm)$-시간과 동의한다.
@@
 \begin{table}[h!] \centering
+\small
 \begin{tabular}{r|r@{~}r|r@{~}r}
@@
  \fi
-    16 &       0.07 &    (0.02) &      0.06 &    (0.02) \\
-    64 &       0.72 &    (0.21) &      0.73 &    (0.21) \\
-   256 &      10.36 &    (2.88) &     10.74 &    (3.34) \\
-  1024 &     182.14 &   (47.83) &    181.95 &   (29.10) \\
-  4096 &    2858.49 &  (123.99) &   2994.55 &  (213.49)            
+    16 &     0.00 &  (0.00) &      0.01 &    (0.00) \\
+    64 &     0.01 &  (0.00) &      0.13 &    (0.03) \\
+   256 &     0.14 &  (0.02) &      1.56 &    (0.57) \\
+  1024 &     2.21 &  (0.35) &     41.65 &   (10.55) \\
+  4096 &    35.36 &  (4.96) &   5080.63 & (6266.66) \\
+ 16384 &   552.45 & (79.57) &  16452.20 & (5525.29)       
 \end{tabular}
-\caption{\label{experiment1results}
+\caption{\label{experiment1results} \normalsize
 \ifen Time to compute an optimal portfolio for an admissions market with homogeneous application costs using Algorithm \ref{algorithmforlargeh} when $\mathcal{C}$ is stored as a list and as a heap.
@@
 \begin{table}[h!] \centering
+\small
 \begin{tabular}{r|r@{~}r|r@{~}r|r@{~}r|r@{~}r}
@@
 \end{tabular}%
-\caption{\label{experiment2results}
+\caption{\label{experiment2results} \normalsize
 \ifen Time to compute an optimal or $(1 - \varepsilon)$-optimal portfolio for an admissions market with heterogeneous application costs using the three algorithms developed in section \ref{hetappcosts}.
@@
 
+% Another ex of nestedness.
+%Tibshirani, Robert.
+1996.
+``Regression Shrinkage and Selection via the Lasso.'' \emph{Journal of the Royal Statistical Society,.
+Series B (Methodological)} 58, no.
+1: 267--88.
+\url{https://www.jstor.org/stable/2346178}.
+
 Vazirani, Vijay. 2001. \emph{Approximation Algorithms.} Berlin: Springer.

### 2022-03-22 benchmark/benchmark.jl

@@ -6,20 +6,24 @@ using StatsBase
 using Statistics
 using Base.Threads
 import Printf: @sprintf
+using BenchmarkTools
 
-fullscale = true
+const fullscale = true
 
 # A long benchmark; tweak parameters with caution.
 # Set fullscale = false to run a smaller benchmark to check formatting etc.
 
 # Number of times to repeat each computation, where min of these is reported as time
-n_reps = fullscale ? 3 : 2
+BenchmarkTools.DEFAULT_PARAMETERS.samples = fullscale ? 5 : 2
+BenchmarkTools.DEFAULT_PARAMETERS.evals = 1 # ... which is the default
 
 # Number of markets to test at each intersection of the experimental variables
-n_markets = fullscale ? 20 : 2
+const n_markets = fullscale ? 20 : 2
+const bnbcutoff = fullscale ? 33 : 6
+const twottbnbcutoff = 2^bnbcutoff
 
 function printheader(s)
-    printstyled(s * "\n", bold = true, color = 222)
+    printstyled(s * "\n", bold=true, color=222)
 end
 
 function make_correlated_market(m)
@@ -31,6 +35,7 @@ function make_correlated_market(m)
     return f, t, g, H
 end
 
+
 randVCM(m) = VariedCostsMarket(make_correlated_market(m)...)
 function randSCM(m)
     f, t, _, _ = make_correlated_market(m)
@@ -40,9 +45,9 @@ end
 
 function benchmark1()
     printheader("Benchmark 1: Homogeneous-cost algorithms")
-    M = fullscale ? 4 .^ (2:6) : [5, 10]
+    M = fullscale ? 4 .^ (2:7) : [5, 10]
 
-    sizes = zeros(Int64, n_markets, length(M))
+    sizes = zeros(Int, n_markets, length(M))
     times_list = zeros(Float64, n_markets, length(M))
     times_heap = zeros(Float64, n_markets, length(M))
 
@@ -51,8 +56,8 @@ function benchmark1()
         for (j, m) in enumerate(M)
             mkt = randSCM(m)
             sizes[i, j] = m
-            times_list[i, j] = minimum(@elapsed applicationorder_list(mkt) for r in 1:n_reps)
-            times_heap[i, j] = minimum(@elapsed applicationorder_heap(mkt) for r in 1:n_reps)
+            times_list[i, j] = @belapsed applicationorder_list($mkt)
[…]

### 2022-03-22 src/EnumerationSolvers.jl

@@ -4,7 +4,7 @@
 Produce the optimal portfolio for the market `mkt` having identical application costs.
 """
 function optimalportfolio_enumerate(mkt::SameCostsMarket)::Tuple{Vector{Int},Float64}
-    X = zeros(Int64, mkt.h)
+    X = zeros(Int, mkt.h)
     v = 0.0
 
     for Y in multiset_combinations(1:mkt.m, mkt.h)

### 2022-03-22 src/FastSolvers.jl

@@ -5,23 +5,25 @@ Contains a college's admissions probability `f`, utility value `t`,
 their product `ft`, and `1 - f = omf`. Used only by `applicationorder()`.
 """
 struct College
-    j::Int
+    j::Int16
     f::Float64
     t::Float64
     ft::Float64
     omf::Float64
 
-    function College(j::Int, f::Float64, t::Real, ft::Float64, omf::Float64)
-        return new(j, f, t, ft, omf)
+    function College(j::Integer, f::Float64, t::Real, ft::Float64, omf::Float64)
+        return new(Int16(j), f, t, ft, omf)
     end
 
-    function College(j::Int, f::Float64, t::Real, omf::Float64)
-        return new(j, f, t, f * t, omf)
+    function College(j::Integer, f::Float64, t::Real, omf::Float64)
+        return new(Int16(j), f, t, f * t, omf)
     end
 end
 
 # Overload isless() so that the heap is ordered by expected utility.
 isless(c1::College, c2::College) = isless(c1.ft, c2.ft)
+
+const dummy_college = College(0, 1.0, -1.0, -1.0, 0.0)
 
 
 """
@@ -30,40 +32,63 @@ isless(c1::College, c2::College) = isless(c1.ft, c2.ft)
 Produce the optimal order of application for the market `mkt` having identical
 application costs and the corresponding portfolio valuations.
 """
-function applicationorder_list(mkt::SameCostsMarket)::Tuple{Vector{Int},Vector{Float64}}
+function applicationorder_list(mkt::SameCostsMarket, verbose=false::Bool)::Tuple{Vector{Int},Vector{Float64}}
     apporder = zeros(Int, mkt.h)
     v = zeros(mkt.h)
 
     mkt_list = College[College(j, mkt.f[j], mkt.t[j], mkt.ft[j], mkt.omf[j]) for j in 1:mkt.m]
     
+    c_best::College, idx_best::Int = findmax(mkt_list)
     for j in 1:mkt.h
-        c_k, k = findmax(mkt_list)
-        v[j] = get(v, j - 1, 0) + c_k.ft
-        apporder[j] = c_k.j
-
-        for i in 1:k-1
-            mkt_list[i] =
-                College(
-                    mkt_list[i].j,
-                    mkt_list[i].f,
-                    mkt_list[i].t * c_k.omf,
[…]

### 2022-03-22 src/MarketIO.jl

@@ -30,6 +30,9 @@ function isnontrivialmarket(
 end
 
 
+# TO DO: Parameterize type of m to use a smaller integer type depending on size of input.
+
+
 """
 Contains information about a college application market with identical
 application costs. Fields:

### 2022-03-22 src/VariedCostsBnB.jl

@@ -54,13 +54,13 @@ hash(nd::Node) = hash((nd.I, nd.N))
 With respect to the data in `mkt`, generates the child(ren) of node `nd` and writes 
 their hashes to `nd.children`.
 """
-function generatechildren(nd::Node, mkt::VariedCostsMarket)
+function generatechildren(nd::Node, mkt::VariedCostsMarket)::Set{Node}
     fltr = filter(j -> mkt.g[j] ≤ nd.H̄, nd.N)
 
     # No way to add any school to this: just skip to the leaf node
     if isempty(fltr)
         child = Node(nd.I, Set{Int}(), nd.t̄, nd.H̄, nd.v_I, mkt)
-        return (child,)
+        return Set((child, ))
     end
 
     # School we will branch on. In principle it can by any school in fltr.
@@ -86,12 +86,12 @@ function generatechildren(nd::Node, mkt::VariedCostsMarket)
             end
         end
         delete!(t̄1, i)
-
+    
         child1 = Node(union(nd.I, i), newN, t̄1, nd.H̄ - mkt.g[i],
             nd.v_I + mkt.f[i] * nd.t̄[i],
             mkt)
-
-        return child1, child2
+    
+        return Set((child1, child2))
     else
         # Then mkt.g[i] == nd.H̄
         # Leaf node with i
@@ -104,10 +104,10 @@ function generatechildren(nd::Node, mkt::VariedCostsMarket)
             end
         end
         delete!(t̄1, i)
-
+    
         child1 = Node(union(nd.I, i), Set{Int}(), t̄1, 0, nd.v_I + mkt.f[i] * nd.t̄[i], mkt)
-
-        return child1, child2
+    
+        return Set((child1, child2))
     end
 end
 
@@ -118,20 +118,20 @@ end
 Use the branch-and-bound algorithm to produce the optimal portfolio for the
 market `mkt` with varying application costs. Intractable for large markets. 
 """
-function optimalportfolio_branchbound(mkt; maxit = 100000::Int, verbose = false::Bool)
+function optimalportfolio_branchbound(mkt; maxit = 100000::Int, verbose = false::Bool)::Tuple{Vector{Int},Float64}
     mkt.m ≥ 33 && @warn "Branch and bound is slow for large markets"
 
     C = Set(1:length(mkt.t))
 
-    rootnode = Node(Set{Int}(), C, Dict(zip(1:mkt.m, Float64.(mkt.t))), mkt.H, 0.0, mkt)
[…]

### 2022-03-28 paper/paper.tex
conclusion, plots

@@
 \fi
+
+% This problem is not exclusively of theoretical interest. Companies sell products ...
 
@@
 The heap is ordered according to the criterion $i \geq j \iff f_i t_i \geq f_j t_j$, and by draining the heap and reheapifying at the end of each iteration, the computation time remains $O(hm)$.
-In our numerical experiments, whose results are reported in section \ref{numericalexperiments},
-we found the heap implementation to be faster than the list implementation. 
+However, in our numerical experiments, whose results are reported in section \ref{numericalexperiments},
+we found the list implementation to be much faster, because it is possible to identify the entering school $k$ as the utility parameters are updated, which all but eliminates the cost of the $\argmax\{\}$ operation.
 \else
@@
 단, 힙의 순서는 $i \geq j \iff f_i t_i \geq f_j t_j$의 조건으로 정의하며, 각 반복단계의 마지막 단계에서 힙을 비우고 다시 힙화(heapify)하면 $O(hm)$ 계산 시간을 유지할 수 있다.
-(\ref{numericalexperiments}절에서 결과를 제시할) 수리 실험에서 힙 구현으로 구성한 알고리즘은 목록보다 빨랐다.
+그러나 (\ref{numericalexperiments}절에서 결과를 제시할) 수리 실험에서 목록 구현으로 구성한 알고리즘이 더 빨랐다.
+효용 모수를 수정하면서 $k$를 구함으로 $\argmax\{\}$ 연산의 계산 비용을 거의 제거할 수 있기 때문이다.
 \fi
@@
 \ifen
-We chose to implement our algorithms in the Julia language (v1.7.0) because its system of type inference allows the compiler to optimize for differential cases such as when the $t_j$-values are integers or floating-point numbers.
+We chose to implement our algorithms in the Julia language (v1.7.0) because its system of parametric data types allows the compiler to optimize for differential cases such as when the $t_j$-values are integers or floating-point numbers.
 Julia also offers convenient macros for parallel computing, which enabled us to solve more and larger problems in the benchmark (Bezanson et al.
@@
 모든 알고리즘을 쥴리아 (Julia, v1.7.0) 언어로 실행했다.
-쥴리아는 $t_j$-값이 정수 또는 부동소수점 수와 같은 특수한 경우에 대해 컴파일러가 자동적으로 최적화하는 자료형 추론(type inference) 기능이 있어서 유용한 선택이었다.
+쥴리아는 $t_j$-값이 정수 또는 부동소수점 수와 같은 특수한 경우에 대해 컴파일러가 자동적으로 최적화하는 모수적 자료형(parametric data type) 기능이 있어서 유용한 선택이었다.
 또한 병렬 연산을 쉽게 관리할 수 있는 매크로를 제공하므로 계산 실험의 규모를 증가할 수 있었다 (Bezanson 외 2017).
@@
 The experimental procedure was as follows.
-First, to generate synthetic markets, we drew the $t_j$-values independently from an exponential distribution with a scale parameter of ten and rounding up to the nearest integer.
+First, to generate synthetic markets, we drew the $t_j$-values independently from an exponential distribution with a scale parameter of ten and rounded up to the nearest integer.
 To achieve partial negative correlation between $t_j$ and $f_j$, we then set $f_j = 1 / (t_j + 10\,Q)$, where $Q$ is drawn uniformly from the interval $[0, 1)$.
@@
 Our results indicate that the list implementation is faster.
-However, the ratio between the average times of the two implementations stays roughly constant, suggesting that a more effective heap implementation could be competitive in certain classes of problem instances.
+%However, the ratio between the average times of the two implementations stays roughly constant, suggesting that a more effective heap implementation could be competitive in certain classes of problem instances.
 Overall, the rate of growth is quadratic in $m$, which accords with the $O(hm)$ time complexity result of Theorem \ref{validityofalmaalgorithm}.
@@
 실험 결과에 따라 목록 구현이 더 빠른 해법이었다.
-그러나, 두개의 방법 사이에 계산 시간 비율은 거의 상수이므로 개선된 힙 해법이 어떤 문제 인스턴스에 대해 더 빠를 가능성이 있어 보인다.
+%그러나, 두개의 방법 사이에 계산 시간 비율은 거의 상수이므로 개선된 힙 해법이 어떤 문제 인스턴스에 대해 더 빠를 가능성이 있어 보인다.
 전체적으로 계산 시간은 $m$의 제곱으로 늘아나며 이는 정리 \ref{validityofalmaalgorithm}에서 말한 $O(hm)$-시간과 동의한다.
@@
 The branch-and-bound algorithm proved impractical for even medium-sized instances.
-%cite computational studies knapsack
 \else
@@
 For each value of $m$, \nmarkets~markets were generated, and the computation time was recorded as fastest of three repetitions of the algorithm.
-The table shows the average time (standard deviation) in milliseconds over the \nmarkets~instances.
-\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복해서 그 중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (평균편차) 시간이 밀리초 단위로 나타난다.\fi}
+The table shows the average time (standard deviation) over the \nmarkets~instances.
+\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복해서 그 중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (평균편차) 시간이 나타난다.\fi}
 
@@
 \caption{\label{experiment1results} \normalsize
-\ifen Time to compute an optimal portfolio for an admissions market with homogeneous application costs using Algorithm \ref{algorithmforlargeh} when $\mathcal{C}$ is stored as a list and as a heap.
+\ifen Time in ms to compute an optimal portfolio for an admissions market with homogeneous application costs using Algorithm \ref{algorithmforlargeh} when $\mathcal{C}$ is stored as a list and as a heap.
 \lastptofcaption~In every case, $h = m/2$.
 \else 알고리즘 \ref{algorithmforlargeh}에서 $\mathcal{C}$를 목록 또는 힙 구현으로 저장했을 때, 
-동일 지원 비용 입학 시장의 최적 지원 포트폴리오를 계산하는 시간. \lastptofcaption~모든 경우, $h = m/2$. \fi}
+동일 지원 비용 입학 시장의 최적 지원 포트폴리오를 계산하는 시간. 단위: ms. \lastptofcaption~모든 경우, $h = m/2$. \fi}
 \end{table}
@@
 \caption{\label{experiment2results} \normalsize
-\ifen Time to compute an optimal or $(1 - \varepsilon)$-optimal portfolio for an admissions market with heterogeneous application costs using the three algorithms developed in section \ref{hetappcosts}.
+\ifen Time in ms to compute an optimal or $(1 - \varepsilon)$-optimal portfolio for an admissions market with heterogeneous application costs using the three algorithms developed in section \ref{hetappcosts}.
 %branch-and-bound routine (Algorithm \ref{ellisbnb}), exact dynamic program (Algorithm \ref{ellisDP1}), and the fully polynomial-time approximation scheme (Algorithm \ref{ellisDP3}).
@@
 \else  %분지한계법 (알고리즘 \ref{ellisbnb}), 정확한 동적 계획 (알고리즘 \ref{ellisDP1}), 그리고 완전 다항 시간 근사 해법 (알고리즘 \ref{ellisDP3})을
- \ref{hetappcosts}절에서 도출한 3개의 알고리즘을 사용할 때, 다양한 지원 비용으로 갖춘 입학 시장의 최적 또는 $(1- \varepsilon)$-최적 포트폴리오를 계산하는 시간. 분지한계법은 큰 시장에서 비실용적이다. \lastptofcaption \fi}
+ \ref{hetappcosts}절에서 도출한 3개의 알고리즘을 사용할 때, 다양한 지원 비용으로 갖춘 입학 시장의 최적 또는 $(1- \varepsilon)$-최적 포트폴리오를 계산하는 시간. 단위: ms. 분지한계법은 큰 시장에서 비실용적이다. \lastptofcaption \fi}
 \end{table}
@@
 \ifen \section{Conclusion} \else \section{결론}\fi\label{conclusion}
+\ifen
+This study has introduced a novel combinatorial optimization problem that we call the college application problem.
+It can be viewed as a kind of integer portfolio optimization problem or stochastic knapsack problem.
+We showed that the special case in which each college's application cost is identical can be solved in polynomial time by a greedy algorithm because the optimal solutions are nested in the budget constraint.
+The general problem is NP-complete.
+We provided three solution algorithms; the strongest from a theoretical standpoint is an FPTAS that produces a $(1-\varepsilon)$-approximate solution in $O(m^3 / \varepsilon)$-time.
+\else
+본 연구는 새로운 조합최적화 문제를 제안했으며 이를 대학 지원 최적화 문제라고 부른다.
+정수 포트폴리오 최적화 문제 또는 확률적인 배낭 문제로 볼 수 있다.
+모든 대학의 지원 비용이 동일한 특수한 경우, 탐욕 알고리즘으로 다항 시간 안에 풀 수 있음을 보였으며 최적해가 예산 제약식에 대한 포함 사슬 관계를 가지기 때문이다.
+일반적인 문제는 NP-complete다.
+3가지 해법을 제시했으며 이론적으로 가장 효율적인 알고리즘은 $O(m^3 / \varepsilon)$-시간 안에  $(1-\varepsilon)$-근사해를 계산하는 FPTAS다.
+\fi
+
+\ifen
+Two extensions of this problem appear amenable to future research.
+First, the familiar Markowitz portfolio optimization model includes an explicit risk-aversion term, whereas our model has settled for an implicit treatment of risk, namely the tradeoff between selective schools and safety schools inherent in the ``maximax'' objective function.
+However, it is possible to augment the objective function $v(\mathcal{X}) = \operatorname{E}[X]$ to incorporate a variance penalty $\beta \geq 0$ as follows:
+\else
+이 문제를 확장할 수 있는 향후 연구 방향 2가지 생각할 수 있다.
+첫번째는, 익숙한 Markowitz 포트폴리오 최적화 모형에는 명시적인 위험 회피 항이 있으며 본 모형은 경쟁적인 학교와 안정 학교 사이의 균형으로 내제화된 위험 관리 요소만 고려했다.
+그러나 기본 목적함수인 $v(\mathcal{X}) = \operatorname{E}[X]$에 평균편차 페널티 $\beta \geq 0$를 다음처럼 더할 수 있다:
+\fi
+\begin{align}
+\begin{split}
+v_{\mathrm{\beta}}(\mathcal{X}) &=  \operatorname{E}[X] - \beta \operatorname{Var}(X) \\
+&=   \operatorname{E}[X] - \beta \left(  \operatorname{E}[X^2]  -  \operatorname{E}[X]^2 \right) \\
+& = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)
+ - \beta \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j^2 \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)
+  + \beta v(\mathcal{X})^2 \\
+  &= v(\mathcal{X}; \tau) +  \beta v(\mathcal{X}; t)^2
+\end{split}
+\end{align}
+\ifen
+where $\tau_j = t_j - \beta t_j^2$.
+Since the first term is itself a portfolio valuation function, we speculate that one of the algorithms given above could be used as a subroutine in a line-search algorithm that maximizes $\beta v(\mathcal{X}; t)^2$ subject to the budget constraint and $v(\mathcal{X}; \tau) \geq \alpha$ for various values of $\alpha \geq 0$.
+\else
+단, $\tau_j = t_j - \beta t_j^2$.
+이의 첫번째 항 그 자체가 포트폴리오 가치 함수다.
+따라서 위에서 제시한 알고리즘을 서브루틴으로 사용하는 선탐색 알고리즘으로 풀 수 있어 보인다.
+이때 다양한 $\alpha \geq 0$에 대해 예산 제약식과 $v(\mathcal{X}; \tau) \geq \alpha$ 하에서 $\beta v(\mathcal{X}; t)^2$을 최대화하는 것이 기본 틀이다.
+\fi
+
+\ifen
+Another direction of further research with immediate application in college admissions is to incorporate logical constraints into the problem.
+In the Korean admissions process, the online application form contains three multiple-choice fields, labeled \emph{ga, na,} and \emph{da} for the first three letters of the Korean alphabet, which students use to indicate the college to which they wish to apply.
+Most schools appear only in one or two of the three fields.
+Therefore, students are restricted not only in the number of applications they can submit, but in the combinations of schools that may coincide in a feasible application portfolio.
+From a portfolio optimization perspective, this is a \emph{diversification constraint,} because its primary purpose is to prevent students from applying only to top-tier schools.
+But because the \emph{ga} field is typically used to signal a student's first choice, colleges that appear in multiple tiers often impose stricter admissions criteria on students who apply in the \emph{na} or \emph{da} field, which creates a subtle strategic challenge.
+An analogous phenomenon in the American college admissions process is known as early decision, in which at the moment of application, a student commits to enrolling in a college if admitted.
+In principle, students who apply with early decision may obtain a better chance of admission by signaling their eagerness to attend, but doing so removes the college's incentive to entice the student with merit-based financial aid, altering the utility calculus.
+\else
+향후 연구에 두번째 방향은 논리 제약을 추가하는 것이다.
+한국 입학 과정의 온라인 지원 양식에는 가군, 나군, 다군이란 3가지 선다형 칸을 채워서 지원하는 학교를 선택한다.
+대부분의 학교는 모든 3개의 칸에 등장하지 않고 그 중 1~2개만 가능하다.
+따라서 학생이 지원할 수 있는 학교의 `수'일 뿐만 아니라 가능한 지원 포트폴리오에 같이 들어갈 수 있는 학교의 `조합'도 제한된다.
+이는 학생이 상위권 대학만 지원하지 않도록 만들어지기 때문에  다각화(diversification) 제약식으로 볼 수 있다.
+그러나 가군에 선택하는 학생이 가장 선호하는 학교로 여겨져 왔기 때문에 대학교가 나군이나 다군으로 지원하는 학생에게 엄격한 전형 기준을 세우는 경향이 있다.
+미묘한 전략적인 문제를 고려하게 한다.
+미국 입학 과정의 유사한 점은 조기 전형(early decision)인데 이는 학생이 지원하는 시기에 합격하면 진학한다고 약속하는 지원 방법이다.
+조기 전형으로 지원하면 입학할 관심을 시그널할 수 있으므로 합격 확률을 높일 수 있다.
+반면에 학교가 학생을 모집하려고 장학금을 제공할 동기가 없어질 수 있으므로 학생의 효용 평가가 달라진다.
+\fi
+
+\ifen
+In the integer formulation of the college application problem (Problem \ref{integernlp}), a natural way to model the diversification constraints described above without introducing too much complexity is to split $c_j$ into $c_{j+}$ and $c_{j-}$.
+The binary decision variables are $x_{j+} = 1$ if the student applies to $c_j$ with high priority (that is, in  the \emph{ga} field or with early decision) and $x_{j-} = 1$ if applying with low priority (that is, in the \emph{na} or \emph{da} field or without early decision).
+Now, assuming that the differential probabilities $f_{j+}$ and $f_{j-}$ and utilities $t_{j+}$ and $t_{j-}$ and utilities are known, adding the logical constraints
+\begin{align}
+ x_{j+} + x_{j-} \leq 1, ~~ j = 1\dots m\qquad\text{and}\qquad\sum_{j=1}^m x_{j+} \leq 1
+\end{align}
+completes the problem formulation. These constraints are conformable to those of the multiple knapsack problem (Kellerer et al. 2004, chap. 10). 
+\else
+대학 지원 문제의 정수 모형(문제 \ref{integernlp})에서 위의 다각화 제약을 어렵지 않게 도입하는 방식은  $c_j$를 $c_{j+}$와 $c_{j-}$로 나누는 것이다.
+그다음에 학생이 $c_j$에 높은 관심을 시그널하면 (즉, 가군으로 선택하거나 조기 전형으로 지원하면) $x_{j+} = 1$이며 관심을 시그널 하지 않고 지원하면 (즉 가군 혹은 나군으로 지원하거나 조기 전형 없이 지원하면) $x_{j-} = 1$이 되는 이진 결정 변수를 정의한다.
+그러면 해당 합격 확률 $f_{j+}$와 $f_{j-}$ 그리고 효용  $t_{j+}$와 $t_{j-}$가 알려져 있다고 가정하면 다음 같은 논리 제약식을 더해서 모형을 완성할 수 있다.
+\begin{align}
+ x_{j+} + x_{j-} \leq 1, ~~ j = 1\dots m\qquad\text{그리고}\qquad\sum_{j=1}^m x_{j+} \leq 1
+\end{align}
+다중 배낭 문제의 제약식과 같을 꼴이다 (Kellerer et al. 2004, 10제). 
+\fi
 

### 2022-03-28 benchmark/benchmark.jl

@@ -21,7 +21,7 @@ const n_reps = fullscale ? 5 : 2
 
 # Number of markets to test at each intersection of the experimental variables
 const n_markets = fullscale ? 50 : 2
-const bnbcutoff = fullscale ? 33 : 6
+const bnbcutoff = fullscale ? 25 : 6
 const twottbnbcutoff = 2^bnbcutoff
 
 # Sizes of markets to test
@@ -162,12 +162,12 @@ end
 
 function doeverything()
     println()
-    bm1, plts1, = benchmark1()
-    display(pretty_table(bm1, formatters=fmter))
-    for pl in plts1
-        display(pl)
-        println()
-    end
+    # bm1, plts1, = benchmark1()
+    # display(pretty_table(bm1, formatters=fmter))
+    # for pl in plts1
+    #     display(pl)
+    #     println()
+    # end
     bm2, plts2, = benchmark2()
     display(pretty_table(bm2, formatters=fmter))
     for pl in plts2

### 2022-03-28 src/FastSolvers.jl

@@ -39,7 +39,7 @@ function applicationorder_list(mkt::SameCostsMarket{T}; verbose=false::Bool)::Tu
     mkt_list = College{T}[College(T(j), mkt.f[j], mkt.t[j], mkt.ft[j], mkt.omf[j]) for j in T(1):mkt.m]
     
     c_best::College{T}, idx_best::Int = findmax(mkt_list)
-    for j in T(1):mkt.h
+    @inbounds for j in T(1):mkt.h
         if verbose
             println("Iteration $j")
             println("  ft: ", [mkt_list[i].ft for i in T(1):length(mkt_list)])
@@ -100,17 +100,17 @@ end
 Produce the optimal order of application for the market `mkt` having identical
 application costs and the corresponding portfolio valuations.
 """
-function applicationorder_heap(mkt::SameCostsMarket{T})::Tuple{Vector{Int},Vector{Float64}} where T
+function applicationorder_heap(mkt::SameCostsMarket{T})::Tuple{Vector{Int},Vector{Float64}} where {T}
     apporder = zeros(T, mkt.h)
     v = zeros(mkt.h)
-    
+
     mkt_heap = BinaryMaxHeap{College{T}}(collect(College(T(j), mkt.f[j], mkt.t[j], mkt.ft[j], mkt.omf[j]) for j in T(1):mkt.m))
-    
-    for j in T(1):mkt.h
+
+    @inbounds for j in T(1):mkt.h
         c_k = first(mkt_heap)
         v[j] = get(v, j - 1, 0) + c_k.ft
         apporder[j] = c_k.j
-    
+
         mkt_heap = BinaryMaxHeap{College{T}}(collect(
                 College(
                     c.j,
@@ -137,9 +137,9 @@ end
 Use the dynamic program on application costs to produce the optimal portfolio `X` and associated
 valuation table `V` for the market `mkt` with varying application costs.
 """
-function optimalportfolio_valuationtable(mkt::VariedCostsMarket{T})::Tuple{Vector{Int},Matrix{Float64}} where T
+function optimalportfolio_valuationtable(mkt::VariedCostsMarket{T})::Tuple{Vector{Int},Matrix{Float64}} where {T}
     V = zeros(mkt.m, mkt.H)
-    for j in T(1):mkt.m, h in 1:mkt.H
+    @inbounds for j in T(1):mkt.m, h in 1:mkt.H
         if h < mkt.g[j]
[…]

### 2022-03-28 src/VariedCostsBnB.jl

@@ -74,7 +74,7 @@ function generatechildren(nd::Node{T}, mkt::VariedCostsMarket{T})::Set{Node{T}} 
     if mkt.g[i] < nd.H̄
         # Node with i
         t̄1 = copy(nd.t̄)
-        for j in keys(t̄1)
+        @inbounds for j in keys(t̄1)
             if t̄1[j] ≤ t̄1[i]
                 t̄1[j] *= mkt.omf[i]
             else
@@ -172,7 +172,7 @@ function optimalportfolio_branchbound(mkt::VariedCostsMarket{T}; maxit = 100000:
 
         if newLB
             # for k in treekeys
-            for k in keys(tree)
+            @inbounds for k in keys(tree)
                 if tree[k].v_LP < LB
                     verbose && println("Fathoming node $k")
                     delete!(tree, k)

### 2022-03-29 paper/paper.tex
benchmarks, performance

@@
 \ifen
-We refer to this problem as the \emph{optimal college application portfolio,} as follows: Consider a college market with $m$ colleges.
+We interpret this problem as the \emph{optimal college application} problem, as follows.
+Consider a college market with $m$ colleges.
 The $j$th college is named $c_j$.
@@
 Therefore, her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which she applies.
-\else
-다음 해석에 따라 이를 `대학 지원 포트폴리오 최적화 문제'라고 부른다.
+(The problem is also conformable to other competitive matching games such as job application.)
+\else
+다음 해석에 따라 이를 `대학 지원 최적화 문제'라고 부른다.
 $m$개의 대학교를 가지는 입학 시장을 고려하자.
@@
 위 문제의 최적해가 $\mathcal{X}$일 때, 학생의 최적 대학 지원 전략은 $\mathcal{X}$에  속한 학교로 지원하는 것이다.
-\fi
-
-% This problem is not exclusively of theoretical interest. Companies sell products ...
-
-\ifen
-As Chao (2014) remarked, college application represents a somewhat subtle portfolio optimization problem.
+(취직 전략과 같은 유사한 경쟁적 매칭 게임에도 위에 문제를 적용할 수 있다.)
+\fi
+
+\ifen
+The college application problem is not solely of theoretical interest.
+In the American college consulting industry, students pay private consultants an average of \$200 per hour for assistance in preparing application materials, estimating their admissions odds, and identifying target schools (Sklarow 2018).
+In Korea, an important revenue stream for admissions consulting firms such as Megastudy (\url{megastudy.net}) and Jinhak (\url{jinhak.com}) is ``mock application'' software that claims to use artificial intelligence to optimize students' application strategy.
+To the extent that career earnings depend on college admissions outcomes, the solution of \eqref{headlineproblem} holds monetary value.
+However, although predicting admissions outcomes on a school-by-school basis is a standard benchmark for regression models (Acharya et al.
+2019), we believe our study is the first to formulate the overall application strategy as an optimization problem.
+\else
+대학 지원 최적화 문제는 이론적인 화제일 뿐만이 아니다.
+미국 입학 컨설팅 산업에서 대학 지원 서류 작성, 합격 확률 추정, 그리고 지원 학교를 선택하는 데에 자문하는 개인 상담가의 시간당 비율은 평균 200달러이다 (Sklarow 2018).
+한국에서, 메가스터디(\url{megastudy.net})와 진학(\url{jinhak.com})과 같은 입학 컨설팅 기업의 주된 수입원 중 인공지능을 활용해서 학생의 지원 전략을 최대화한다고 홍보하는 ``모의 지원'' 소프트웨어가 있다.
+미래의 소득이 대학 입학 결과로 결정되는 만큼, \eqref{headlineproblem}의 최적해는 금전적인 가치를 가지고 있다.
+그러나, 학교 개별 합격 확률을 추정하는 것은 회귀 분석 모형의 익숙한 응용 사례(Acharya et al.
+2019)이지만, 전체적인 지원 전략을 최적화 문제로 모형화한 것은 본 연구의 새로운 기여로 보인다.
+\fi
+
+\ifen
+From a methodological standpoint, as Chao (2014) remarked, college application represents a somewhat subtle asset allocation problem.
 In computational finance, traditional portfolio optimization models weigh the sum of expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Markowitz 1952; Meucci 2005).
@@
 Chao estimated her model (which considers application as a \emph{cost} rather than a constraint) by clustering the schools so that $m=8$, a scale at which enumeration is tractable.
-Our study pursues a more general solution.
-\else
-Chao (2014)가 주장한 바처럼 대학 지원 전략은 약간 미묘한 포트폴리오 최적화 문제이다.
+We pursue a more general solution.
+\else
+방법론적인 관점에서 보면, Chao (2014)가 주장한 바처럼 대학 지원 전략은 약간 미묘한 자산 배분 문제이다.
 재정학 분야에서 고전적 포트폴리오 최적화 모형은 전체 자산에 대한 기대 총이익에서 위험회피 항을 뺌으로 선형식으로 제약된 오목 최대화 문제를 이룬다 (Markowitz 1952; Meucci 2005).
@@
 2008).
-Our problem superficially resembles the preference-order knapsack problem considered by Steinberg and Parks and Carraway et al., but these models lack the college application problem's unique ``maximax'' form.
+Our problem superficially resembles the preference-order knapsack problem considered by Steinberg and Parks and Carraway et al., but these models lack the college application problem's singular ``maximax'' form.
 Additionally, unlike those models, we do not attempt to replace the real-valued objective function with a preference order over \emph{outcome distributions,} which introduces technical issues concerning competing notions of stochastic dominance (Sniedovich 1980).
-We take for granted the student's preferences over \emph{outcomes} (as encoded in the $t_j$-values), and are interested instead in an efficient computational approach to the well-defined problem above.
+We take for granted the student's preferences over \emph{outcomes} (as encoded in the $t_j$-values), and focus instead on an efficient computational approach to the well-defined problem above.
 \else
@@
 예를 들어 예산이 다 소비될 때까지 목적함수를 가장 많이 증가시키는 학교를 차례대로 추가하는 알고리즘은 일종의 탐욕 해법이다.
-탐욕 해법은 예산 $H$으로 모수화된 해의 순서를 유도하며 그의 원소들은 `포함 사슬 관계'(nestedness)로 연결한다.
+탐욕 해법은 예산 $H$로 모수화된 해의 순서를 유도하며 그의 원소들은 `포함 사슬 관계'(nestedness)로 연결된다.
 즉 $H \leq H'$일 때, 예산 $H$에 해당하는 탐욕 해는 예산 $H'$에 해당하는 탐욕 해의 부분집합이 된다.
@@
 This case mirrors the centralized college application process in Korea, where there is no application fee, but students are allowed to apply to only three schools during the main admissions cycle.
-Unfortunately, the nestedness property does not hold in the general case, nor does the greedy algorithm offers any performance guarantee.
+Unfortunately, the nestedness property does not hold in the general case, nor does the greedy algorithm offer any performance guarantee.
 Instead, we offer a branch-and-bound routine, a pseudopolynomial-time algorithm that is tractable for typical college market instances, and an approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in fully polynomial time.
@@
 \ifen It is obvious that the problem is in NP.
-\else 문제가 NP에 속하는 것은 명백하다.\fi
+\else 문제가 NP에 속하는 것은 자명하다.\fi
 
@@
 At each iteration, it selects the node $n \in \mathfrak{T}$ having the highest $v_{\mathrm{LP}}^*[n]$-value, generates its children, and removes $n$ from the candidate set.
-Next, the children $n'$ of $n$ are inspected; if one of them yields a new optimal solution, then we mark it as the best candidate and fathom any nodes $n''$ for which  $v_{\mathcal{I}}[n'] >  v_{\mathrm{LP}}^*[n'']$.
+Next, the children $n'$ of $n$ are inspected and added to the tree.
+If one of the children yields a new optimal solution, then we mark it as the best candidate and fathom any nodes $n''$ for which  $v_{\mathcal{I}}[n'] >  v_{\mathrm{LP}}^*[n'']$.
 When no nodes remain in the candidate set, the algorithm has explored every branch, so it returns the best candidate and terminates.
@@
 $n$의 새끼를 생성하고 후보 집합에서 제거한다.
-그다음에 $n$의 각 새끼 마디 $n'$을 검증한다.
-그중에 새로운 최적해를 발견하면 이를 가장 좋은 후보로 표하고  $v_{\mathcal{I}}[n'] >  v_{\mathrm{LP}}^*[n'']$인 마디 $n''$을 후보 집합에서 제거한다.
+그다음에 $n$의 각 새끼 마디 $n'$을 검증하고 나무에 추가한다.
+새끼 마디 중에 새로운 최적해를 발견하면 이를 가장 좋은 후보로 표하고  $v_{\mathcal{I}}[n'] >  v_{\mathrm{LP}}^*[n'']$인 마디 $n''$을 후보 집합에서 제거한다.
 후보 집합이 공집합이 되면 알고리즘이 모든 가지를 탐색했으므로 가장 좋은 후보를 출력하고 종료한다.
@@
 Because we cannot assume that $H \leq m$ (as was the case in Alma's problem), this represents a pseudopolynomial-time solution (Garey and Johnson 1979, \S\,4.2).
+However, it is quite effective for typical college-application instances in which the application costs are small integers.
 \else
@@
 알마의 문제와 달리, 여기에서 $H \leq m$임을 가정할 수 없으므로 이는 의사 다항 시간 해법이라고 한다 (Garey and Johnson 1979, \S\,4.2).
+그러나 지원 비용이 작은 정수가 되는 전형적인 대학 지원 문제 인스턴스에 대해 매우 효율적인 해법이다.
 \fi
@@
 with the convention that $ -\infty \cdot 0 = -\infty$.
-The corresponding optimal portfolios can be computed by observing that $\mathcal{X}[j, h]$ contains $j$ if and only if $V[j, h]> V[j-1, h]$.
+The corresponding optimal portfolios are computed by observing that $\mathcal{X}[j, h]$ contains $j$ if and only if $V[j, h]> V[j-1, h]$.
 The optimal solution is given by $\mathcal{X}[m, H]$.
@@
 단, $ -\infty \cdot 0 = -\infty$으로 처리한다.
-그러면 $\mathcal{X}[j, h]$가  $j$를 포함할 필수충분 조건이 $V[j, h]> V[j-1, h]$임을 관찰하면 대응되는 최적 포트폴리오를 쉽게 계산할 수 있으며 전역 최적해는 $\mathcal{X}[m, H]$이다.
+그러면 $\mathcal{X}[j, h]$가  $j$를 포함할 필수충분 조건이 $V[j, h]> V[j-1, h]$임을 관찰하면 대응되는 최적 포트폴리오를 계산할 수 있으며 전역 최적해는 $\mathcal{X}[m, H]$이다.
 아래 알고리즘은 이 계산을 수행하고 최적 포트폴리오 $\mathcal{X}$를 출력한다.
@@
 \ifen
-Two extensions of this problem appear amenable to future research.
+Two extensions of this problem appear amenable to future study.
 First, the familiar Markowitz portfolio optimization model includes an explicit risk-aversion term, whereas our model has settled for an implicit treatment of risk, namely the tradeoff between selective schools and safety schools inherent in the ``maximax'' objective function.
@@
 \else
-이 문제를 확장할 수 있는 향후 연구 방향 2가지 생각할 수 있다.
-첫번째는, 익숙한 Markowitz 포트폴리오 최적화 모형에는 명시적인 위험 회피 항이 있으며 본 모형은 경쟁적인 학교와 안정 학교 사이의 균형으로 내제화된 위험 관리 요소만 고려했다.
-그러나 기본 목적함수인 $v(\mathcal{X}) = \operatorname{E}[X]$에 평균편차 페널티 $\beta \geq 0$를 다음처럼 더할 수 있다:
+이 문제를 확장할 수 있는 향후 연구 방향 2가지를 제안한다.
+첫번째는, 익숙한 Markowitz 포트폴리오 최적화 모형에는 명시적인 위험 회피 항이 있으며, 본 모형은 경쟁적인 학교와 안정 학교 사이의 균형으로 내포된 위험 관리 요소만 고려했다.
+그러나 기본 목적함수인 $v(\mathcal{X}) = \operatorname{E}[X]$에 평균편차 페널티 $\beta \geq 0$를 다음처럼 도입할 수 있다:
 \fi
@@
 where $\tau_j = t_j - \beta t_j^2$.
-Since the first term is itself a portfolio valuation function, we speculate that one of the algorithms given above could be used as a subroutine in a line-search algorithm that maximizes $\beta v(\mathcal{X}; t)^2$ subject to the budget constraint and $v(\mathcal{X}; \tau) \geq \alpha$ for various values of $\alpha \geq 0$.
+Since the first term is itself a portfolio valuation function, and the second is a monotonic transformation of one, we speculate that one of the algorithms given above could be used as a subroutine to trace out the efficient frontier by maximizing $v(\mathcal{X}; t)$ subject to the budget constraint and $v(\mathcal{X}; \tau) \geq \alpha$ for various values of $\alpha \geq 0$.
 \else
 단, $\tau_j = t_j - \beta t_j^2$.
-이의 첫번째 항 그 자체가 포트폴리오 가치 함수다.
-따라서 위에서 제시한 알고리즘을 서브루틴으로 사용하는 선탐색 알고리즘으로 풀 수 있어 보인다.
-이때 다양한 $\alpha \geq 0$에 대해 예산 제약식과 $v(\mathcal{X}; \tau) \geq \alpha$ 하에서 $\beta v(\mathcal{X}; t)^2$을 최대화하는 것이 기본 틀이다.
-\fi
-
-\ifen
-Another direction of further research with immediate application in college admissions is to incorporate logical constraints into the problem.
-In the Korean admissions process, the online application form contains three multiple-choice fields, labeled \emph{ga, na,} and \emph{da} for the first three letters of the Korean alphabet, which students use to indicate the college to which they wish to apply.
+이의 첫번째 항 그 자체가 포트폴리오 가치 함수이며, 두번째 항은 포트폴리오 가치 함수의 단조 변환이다.
+따라서 위에서 제시한 알고리즘을 서브루틴으로 사용하면 효율적 투자선(efficient frontier)을 탐색할 수 있어 보인다.
+이때 다양한 $\alpha \geq 0$에 대해 예산 제약식과 $v(\mathcal{X}; \tau) \geq \alpha$ 하에서 $v(\mathcal{X}; t)$을 최대화하는 것이 알고리즘의 요점이다.
+\fi
+
+\ifen
+Another direction of further research with immediate application in college admissions is to incorporate additional signaling strategies into the problem.
+In the Korean admissions process, the online application form contains three multiple-choice fields, labeled \emph{ga, na,} and \emph{da} for the first three letters of the Korean alphabet, which students use to indicate the colleges to which they wish to apply.
 Most schools appear only in one or two of the three fields.
@@
 From a portfolio optimization perspective, this is a \emph{diversification constraint,} because its primary purpose is to prevent students from applying only to top-tier schools.
-But because the \emph{ga} field is typically used to signal a student's first choice, colleges that appear in multiple tiers often impose stricter admissions criteria on students who apply in the \emph{na} or \emph{da} field, which creates a subtle strategic challenge.
+However, in addition to overall selectivity, the three fields also indicate different evaluation schemes: Applications filed under the \emph{ga} and {na} fields are evaluated with greater emphasis on standardized-test scores than those filed under the \emph{da} field.
+Some colleges appear in multiple fields and run two parallel admissions processes, each with different evaluation criteria.
+Therefore, if a student recognizes that she has a comparative advantage in (for example) her interview skills, she may increase her chances of admission to a competitive school by applying under the \emph{ga} field.
+Clearly, the optimal application strategy in this context can be quite subtle, and are a perennial subject of debate on Korean social networks.
+
 An analogous phenomenon in the American college admissions process is known as early decision, in which at the moment of application, a student commits to enrolling in a college if admitted.
@@
 \else
-향후 연구에 두번째 방향은 논리 제약을 추가하는 것이다.
+향후 연구에 두번째 방향은 모형에 시그널 전략을 도입하는 것이다.
 한국 입학 과정의 온라인 지원 양식에는 가군, 나군, 다군이란 3가지 선다형 칸을 채워서 지원하는 학교를 선택한다.
@@
 이는 학생이 상위권 대학만 지원하지 않도록 만들어지기 때문에  다각화(diversification) 제약식으로 볼 수 있다.
-그러나 가군에 선택하는 학생이 가장 선호하는 학교로 여겨져 왔기 때문에 대학교가 나군이나 다군으로 지원하는 학생에게 엄격한 전형 기준을 세우는 경향이 있다.
-미묘한 전략적인 문제를 고려하게 한다.
-미국 입학 과정의 유사한 점은 조기 전형(early decision)인데 이는 학생이 지원하는 시기에 합격하면 진학한다고 약속하는 지원 방법이다.
-조기 전형으로 지원하면 입학할 관심을 시그널할 수 있으므로 합격 확률을 높일 수 있다.
-반면에 학교가 학생을 모집하려고 장학금을 제공할 동기가 없어질 수 있으므로 학생의 효용 평가가 달라진다.
-\fi
-
-\ifen
-In the integer formulation of the college application problem (Problem \ref{integernlp}), a natural way to model the diversification constraints described above without introducing too much complexity is to split $c_j$ into $c_{j+}$ and $c_{j-}$.
-The binary decision variables are $x_{j+} = 1$ if the student applies to $c_j$ with high priority (that is, in  the \emph{ga} field or with early decision) and $x_{j-} = 1$ if applying with low priority (that is, in the \emph{na} or \emph{da} field or without early decision).
-Now, assuming that the differential probabilities $f_{j+}$ and $f_{j-}$ and utilities $t_{j+}$ and $t_{j-}$ and utilities are known, adding the logical constraints
+그러나 대학이 어떤 군에 속하는 것은 대학 순위일 뿐더러 입학 전형 유형도 의미한다.
+가군과 나군으로 지원하는 경우, 다군보다 수능 점수의 반영 비율이 높다.
+또한 군 2개에 포함되는 학교도 있으며 입학 전형이 다르면서 평행한 모집 과정 2개를 운영한다.
+가령, 어떤 학생이 면접 능력에 비교 우위가 있으면 경쟁적인 학교에 가군으로 지원하면 합격 확률을 높일 수 있다.
+이런 맥락에서 최적의 지원 전략은 미묘한 문제이며 한국 SNS에서 자주 발생하는 화젯거리다.
+
+미국 입학 과정의 유사한 점은 조기 전형(early decision)이다. 이는 학생이 지원하는 시기에 합격하면 진학한다고 약속하는 지원 방법이다. 조기 전형으로 지원하면 입학할 관심을 시그널할 수 있으므로 합격 확률을 높일 수 있다. 반면에 학교가 학생을 모집하려고 장학금을 수여할 동기가 없어질 수 있으므로 학생의 효용 추정이 달라진다.
+\fi
+
+\ifen
+In the integer formulation of the college application problem (Problem \ref{integernlp}), a natural way to model these signaling strategies without introducing too much complexity is to split each $c_j$ into $c_{j+}$ and $c_{j-}$.
+The binary decision variables are $x_{j+} = 1$ if the student applies to $c_j$ with high priority (that is, in  the \emph{ga} or \emph{na} field or with early decision) and $x_{j-} = 1$ if applying with low priority (that is, in the \emph{da} field or without early decision).
+Now, assuming that the differential admission probabilities $f_{j+}$ and $f_{j-}$ and utilities $t_{j+}$ and $t_{j-}$ are known, adding the logical constraints
 \begin{align}
@@
 \else
-대학 지원 문제의 정수 모형(문제 \ref{integernlp})에서 위의 다각화 제약을 어렵지 않게 도입하는 방식은  $c_j$를 $c_{j+}$와 $c_{j-}$로 나누는 것이다.
+대학 지원 문제의 정수 모형(문제 \ref{integernlp})에서 위의 시그널 전략들을 어렵지 않게 도입하는 방식은 각 $c_j$를 $c_{j+}$와 $c_{j-}$로 나누는 것이다.
 그다음에 학생이 $c_j$에 높은 관심을 시그널하면 (즉, 가군으로 선택하거나 조기 전형으로 지원하면) $x_{j+} = 1$이며 관심을 시그널 하지 않고 지원하면 (즉 가군 혹은 나군으로 지원하거나 조기 전형 없이 지원하면) $x_{j-} = 1$이 되는 이진 결정 변수를 정의한다.
@@
 
+Acharya, Mohan S., Asfia Armaan, and Aneeta S.
+Antony.
+2019.
+``A Comparison of Regression Models for Prediction of Graduate Admissions.'' In \emph{Second International Conference on Computational Intelligence in Data Science.} \url{https://doi.org/10.1109/ICCIDS.2019.8862140}.
+
 Bezanson, Jeff, Alan Edelman, Stefan Karpinski, and Viral B. Shah. 2017. ``Julia: A Fresh Approach to Numerical Computing.'' \emph{SIAM Review} 59: 65–98. \url{https://doi.org/10.1137/141000671}.
@@
 \url{https://doi.org/10.1016/j.disopt.2020.100581}.
+
+Sklarow, Mark.
+2018.
+\emph{State of the Profession 2018: The 10 Trends Reshaping Independent Educational Consulting.} Technical report, Independent Eduational Consultants Association.
+\url{https://www.iecaonline.com/wp-content/uploads/2020/02/IECA-Current-Trends-2018.pdf}.
 

### 2022-03-29 benchmark/benchmark.jl

@@ -9,43 +9,30 @@ import Printf: @sprintf
 # using BenchmarkTools
 using UnicodePlots
 
-const fullscale = true
+const fullscale = false
 
 # A long benchmark; tweak parameters with caution.
 # Set fullscale = false to run a smaller benchmark to check formatting etc.
 
 # Number of times to repeat each computation, where min of these is reported as time
-const n_reps = fullscale ? 5 : 2
-# BenchmarkTools.DEFAULT_PARAMETERS.samples = n_reps
-# BenchmarkTools.DEFAULT_PARAMETERS.evals = 1 # ... which is the default
+const n_reps = fullscale ? 3 : 2
 
 # Number of markets to test at each intersection of the experimental variables
 const n_markets = fullscale ? 50 : 2
+
+# Cutoffs to exclude certain large computations
 const bnbcutoff = fullscale ? 25 : 6
 const twottbnbcutoff = 2^bnbcutoff
+const fptascutoff_m = fullscale ? 300 : 11
+const fptascutoff_eps = 0.1
 
 # Sizes of markets to test
-const marketsizes_SCM = fullscale ? 4 .^ (2:7) : [5, 10]
-const marketsizes_VCM = fullscale ? 2 .^ (3:9) : [5, 10]
+const marketsizes_SCM = fullscale ? 4 .^ (2:7) : [5, 10, 15]
+const marketsizes_VCM = fullscale ? 2 .^ (3:9) : [5, 10, 15]
+const epsilons = [0.5, 0.05]
 
-function makecorrelatedmarketdata(m)
-    t = ceil.(Int, 10 * randexp(m))
-    sort!(t)
-    f = inv.(t .+ 10 * rand(m))
-    g = rand(5:10, m)
-    H = sum(g) ÷ 2
-    return f, t, g, H
-end
-
-randVCM(m) = VariedCostsMarket(makecorrelatedmarketdata(m)...)
-
-function randSCM(m)
-    f, t, _, _ = makecorrelatedmarketdata(m)
-    return SameCostsMarket(f, t, m ÷ 2)
-end
-
-const mkts_SCM = SameCostsMarket[randSCM(m) for i in 1:n_markets, m in marketsizes_SCM]
-const mkts_VCM = VariedCostsMarket[randVCM(m) for i in 1:n_markets, m in marketsizes_VCM]
+const mkts_SCM = SameCostsMarket[SameCostsMarket(m) for m in marketsizes_SCM, i in 1:n_markets]
[…]

### 2022-03-29 paper/plots/plots.jl

@@ -1,0 +1,96 @@
+using OptimalApplication
+using Random
+using CairoMakie
+using DataFrames
+using Colors
+
+# const accentcolor = (107,142,35) ./ 255
+const accentcolor = (210, 10, 60) ./ 255
+
+#   Comparative statics example
+#   =============================
+
+function comparativestatics()
+    mkt = Market([0.39, 0.33, 0.24, 0.24, 0.05, 0.03, 0.1, 0.12],
+                collect(range(start=200, step=50, length=8)),
+                8)
+
+    apporder, v = applicationorder_list(mkt, verbose=true)
+
+    invpermapporder = invperm(apporder)
+
+    println(apporder)
+
+    df = DataFrame("f" => mkt.f, "t" => mkt.t, "order" => invpermapporder, "v" => v[invpermapporder])
+    #=
+    Row │ f        t        order  v       
+        │ Float64  Float64  Int64  Float64 
+    ─────┼──────────────────────────────────
+    1 │    0.39    200      4  230.047
+    2 │    0.33    250      2  146.7
+    3 │    0.24    300      6  281.513
+    4 │    0.24    350      1   84.0
+    5 │    0.05    400      7  288.778
+    6 │    0.03    450      8  294.106
+    7 │    0.1     500      5  257.643
+    8 │    0.12    550      3  195.096
+    =#
+    display(df)
+
+    fig1 = Figure(resolution=(600, 450))
+
+    ax = Axis(fig1[1, 1], xticks=0:8, xlabel="h", ylabel="v*")
+
+    scatterlines!(ax,
+        0:8,
+        vcat(0, v),
+        color=RGB(accentcolor...),
+        marker=:diamond,
+        markersize=12,
+        linewidth=2,
+    )
+
+    return fig1
+end
+
+#   Sample random market
+#   ======================
+
+
+function samplerandom()
+    m = 64
+    mkt = VariedCostsMarket(m)
+
+    X, v = optimalportfolio_dynamicprogram(mkt)
+
+    nX = setdiff(1:m, X)
+
+    # GR plots
+    # scal = 1.5
+    # pl = plot(size = (500, 500), xlabel = "fⱼ", ylabel = "tⱼ", legend = :topright)
[…]

### 2022-03-29 src/FastSolvers.jl

@@ -192,6 +192,8 @@ value `v` for the market `mkt` with varying application costs.
 """
 function optimalportfolio_dynamicprogram(mkt::VariedCostsMarket{T}; verbose=false::Bool)::Tuple{Vector{Int},Float64} where {T}
     V_dict = Dict{Tuple{T,Int},Float64}()
+    sizehint!(V_dict, mkt.m * mkt.m ÷ 2)
+
     v = V_recursor!(V_dict, mkt.m, mkt.H, mkt)
 
     h = mkt.H
@@ -228,15 +230,7 @@ struct ScaleParams
 
     function ScaleParams(mkt::VariedCostsMarket, ε::Float64)
         @assert 0 < ε < 1
-        P = try
-            ceil(Int, log2(Int(mkt.m)^2 / (ε * sum(mkt.ft))))
-        catch e
-            @show mkt
-            @show ε
-            @show sum(mkt.ft)
-            @show mkt.m^2 / (ε * sum(mkt.ft))
-            throw(e)
-        end
+        P = ceil(Int, log2(Int(mkt.m)^2 / (ε * sum(mkt.ft))))
         t = mkt.t .* 2^P
         ft = mkt.f .* t
         Ū = ceil(Int, sum(ft))
@@ -254,7 +248,7 @@ end
         v::Int, 
         mkt::VariedCostsMarket{T},
         sp::ScaleParams
-    )::Int where T
+    )::Int where T<:Unsigned
     haskey(G_dict, (j, v)) && return G_dict[(j, v)]
 
     if v ≤ 0
@@ -264,7 +258,7 @@ end
     else
         jmo = j - T(1)
         if mkt.f[j] < 1
-            # Clamping prevents over/underflow: for any v≤0 or v>Ū the function
+            # Clamping prevents over/underflow: for any v<0 or v≥Ū the function
             # is trivially defined, so recording any more extreme number is meaningless
             v_minus_Δ = floor(Int, clamp((v - sp.ft[j]) / mkt.omf[j], -1, sp.Ū))
 
@@ -287,10 +281,11 @@ end
 Use the fully polynomial-time approximation scheme to produce a
 `1-ε`-optimal portfolio for the market `mkt` with varying application costs. 
 """
-function optimalportfolio_fptas(mkt::VariedCostsMarket{T}, ε::Float64; verbose=false::Bool)::Tuple{Vector{Int},Float64} where T
[…]

### 2022-03-29 src/MarketIO.jl

@@ -72,6 +72,15 @@ struct SameCostsMarket{T<:Unsigned}
 
         return new{T}(m, f, t, T(h), f .* t, 1 .- f, perm)
     end
+
+    function SameCostsMarket(m::Integer)
+        T = UIntTypes[findfirst(T -> m < typemax(T), UIntTypes)]
+        t = ceil.(T, -10 * log.(rand(m)))
+        sort!(t)
+        f = inv.(t .+ 10 * rand(m))
+        perm = T.(1:m)
+        return new{T}(T(m), f, t, T(m ÷ 2), f .* t, 1 .- f, perm)
+    end
 end
 
 
@@ -115,6 +124,17 @@ struct VariedCostsMarket{T<:Unsigned}
 
         return new{T}(m, f, t, g, H, f .* t, 1 .- f, perm)
     end
+
+    function VariedCostsMarket(m::Integer)
+        T = UIntTypes[findfirst(T -> m < typemax(T), UIntTypes)]
+        t = ceil.(T, -10 * log.(rand(m)))
+        sort!(t)
+        f = inv.(t .+ 10 * rand(m))
+        g = rand(5:10, m)
+        H = sum(g) ÷ 2
+        perm = T.(1:m)    
+        return new{T}(T(m), f, t, g, H, f .* t, 1 .- f, perm)
+    end
 end
 
 """
@@ -147,7 +167,7 @@ end
 Return the valuation of the portfolio `X` on the market `mkt`, which may be either 
 a `SameCostsMarket` or a `VariedCostsMarket`.
 """
-function valuation(X::Vector{T}, mkt::Union{SameCostsMarket{T},VariedCostsMarket}) where T
+function valuation(X::Vector{T}, mkt::Union{SameCostsMarket{T},VariedCostsMarket{T}})::Float64 where T<:Unsigned
     isempty(X) && return 0.0
 
     sort!(X)

### 2022-03-29 test/runtests.jl

@@ -5,27 +5,12 @@ using Random
 # Number of markets to generate for each test
 const n_markets = 3
 
-function make_correlated_market(m)
-    t = ceil.(Int, 10 * randexp(m))
-    sort!(t)
-    f = inv.(t .+ 10 * rand(m))
-    g = rand(5:10, m)
-    H = sum(g) ÷ 2
-    return f, t, g, H
-end
-
-randVCM(m) = VariedCostsMarket(make_correlated_market(m)...)
-function randSCM(m)
-    f, t, _, _ = make_correlated_market(m)
-    return SameCostsMarket(f, t, m ÷ 2)
-end
-
 @testset verbose = true "OptimalApplication.jl" begin
     @testset verbose = true "Same app. costs" begin
         m = 20
 
         for _ in 1:n_markets
-            mkt = randSCM(m)
+            mkt = SameCostsMarket(m)
 
             X_enum, vX_enum = optimalportfolio_enumerate(mkt)
             sort!(X_enum)
@@ -43,7 +28,7 @@ end
 
         @testset "Exact algorithms" begin
             for _ in 1:n_markets
-                mkt = randVCM(m)
+                mkt = VariedCostsMarket(m)
 
                 X_enum, vX_enum = optimalportfolio_enumerate(mkt)
                 sort!(X_enum)
@@ -65,7 +50,7 @@ end
 
         @testset "FPTAS" begin
             for _ in 1:n_markets
-                mkt = randVCM(m)
+                mkt = VariedCostsMarket(m)
 
                 W, vW = optimalportfolio_fptas(mkt, ε)
                 Y, vY = optimalportfolio_dynamicprogram(mkt)
@@ -76,12 +61,12 @@ end
     end
 
     @testset verbose = true "Large problems" begin
-        mkt = randSCM(5000)
+        mkt = SameCostsMarket(5000)
 
         X, V = applicationorder_list(mkt)
         @test !isempty(X)
 
-        mkt = randVCM(500)
+        mkt = VariedCostsMarket(500)
 
         W, vW = optimalportfolio_fptas(mkt, 0.5)
         Y, vY = optimalportfolio_dynamicprogram(mkt)

### 2022-03-30 paper/paper.tex
benchmark

@@
 \ifen
-We interpret this problem as the \emph{optimal college application} problem, as follows.
-Consider a college market with $m$ colleges.
+We refer to this problem as the \emph{optimal college application} problem, as follows.
+Consider an admissions market with $m$ colleges.
 The $j$th college is named $c_j$.
@@
 In the American college consulting industry, students pay private consultants an average of \$200 per hour for assistance in preparing application materials, estimating their admissions odds, and identifying target schools (Sklarow 2018).
-In Korea, an important revenue stream for admissions consulting firms such as Megastudy (\url{megastudy.net}) and Jinhak (\url{jinhak.com}) is ``mock application'' software that claims to use artificial intelligence to optimize students' application strategy.
-To the extent that career earnings depend on college admissions outcomes, the solution of \eqref{headlineproblem} holds monetary value.
+In Korea, an important revenue stream for admissions consulting firms such as Megastudy (\url{megastudy.net}) and Jinhak (\url{jinhak.com}) is ``mock application'' software that claims to use artificial intelligence to optimize the client's application strategy.
+Due to the widespread perception that career earnings depend on college admissions outcomes, the solution of \eqref{headlineproblem} holds monetary value.
 However, although predicting admissions outcomes on a school-by-school basis is a standard benchmark for regression models (Acharya et al.
@@
 한국에서, 메가스터디(\url{megastudy.net})와 진학(\url{jinhak.com})과 같은 입학 컨설팅 기업의 주된 수입원 중 인공지능을 활용해서 학생의 지원 전략을 최대화한다고 홍보하는 ``모의 지원'' 소프트웨어가 있다.
-미래의 소득이 대학 입학 결과로 결정되는 만큼, \eqref{headlineproblem}의 최적해는 금전적인 가치를 가지고 있다.
+미래의 소득이 대학 입학 결과로 결정된다는 넓은 인식에 따라, \eqref{headlineproblem}의 최적해는 금전적인 가치를 가지고 있다.
 그러나, 학교 개별 합격 확률을 추정하는 것은 회귀 분석 모형의 익숙한 응용 사례(Acharya et al.
@@
 Finally, the combinatorial nature of the college application problem makes it difficult to solve using the gradient-based techniques used in continuous portfolio optimization.
-Chao estimated her model (which considers application as a \emph{cost} rather than a constraint) by clustering the schools so that $m=8$, a scale at which enumeration is tractable.
+Chao estimated her equilibrium model (which considers application as a \emph{cost} rather than a constraint) by clustering the schools so that $m=8$, a scale at which enumeration is tractable.
 We pursue a more general solution.
@@
 마지막, 대학 지원의 조합적인 본성으로 인해 연속적인 포트폴리오 최적화 문제에서 흔히 사용하는 기울기 해법으로는 풀기가 어렵다.
-Chao는 지원 비용을 제약 조건 대신 목적함수의 한 항으로 모형화했으며, 모형의 모수를 추정하기 위해 $m=8$이 되도록 학교들의 클러스터를 먼저 구성했다.
+Chao는 지원 비용을 제약 조건 대신 목적함수의 한 항으로 모형화했으며, 균형 모형의 모수를 추정하기 위해 $m=8$이 되도록 학교들의 클러스터를 먼저 구성했다.
 이는 열거법을 통해 문 쉽게 풀 수 있는 규모이지만 본 연구는 더 일반적인 $m$에 대한 해법을 추구한다.
@@
 \ifen \subsection{Experimental procedure} \else \subsection{실험 방법}\fi
-\def\nmarkets{20}
+\def\nmarkets{50}
 \ifen
@@
 In the second experiment, which concerns Ellis's problem, each $g_j$ is drawn uniformly from the set $\{5, \dots, 10\}$ and we set $H = \lfloor \frac{1}{2} \sum g_j \rfloor$.
+This cost distribution resembles real-world college application fees, which are often multiples of \$10 between \$50 and \$100.
 A typical instance is shown in Figure \ref{samplemarket}.
@@
 엘리스의 문제를 고려하는 두번째 실험에서, 모든  $g_j$를 $\{5, \dots, 10\}$에서 균일한 확률로 관측했으며 $H = \lfloor \frac{1}{2} \sum g_j \rfloor$으로 처리했다.
+실제 대학 지원 비용인 50~100달러 사이에 10달러의 어떤 배수이므로 합리적인 비용 분포이다.
 그림 \ref{samplemarket}에서 전형적인 인스턴스가 나타난다.
@@
 Therefore, each cell of each table below represents a statistic over {\the\numexpr 3 * \nmarkets \relax}~computations.
-Where applicable, we did not count the time required to sort the entries of $t$.
+Where applicable, we do not count the time required to sort the entries of $t$.
 \else
@@
 The results, which appear in Table \ref{experiment2results}, broadly agree with the time complexity analyses presented above.
+The branch-and-bound algorithm proved impractical for even medium-sized instances.
 Overall, we found the exact dynamic program to be the fastest algorithm, while the FPTAS was rather slow, a result that echoes the results of computational studies on knapsack problems (Martello and Toth 1990, \S\,2.10).
-The branch-and-bound algorithm proved impractical for even medium-sized instances.
+
+The strong performance of Algorithm \ref{ellisDP1} is partly attributable to the structure of our synthetic instances, in which application costs are small integers and $H$ is proportional in expectation to $m$, meaning the expected computation time is $O(m^2)$.
+However, the relative advantage of Algorithm \ref{ellisDP1} is even more pronounced in real college-application instances because the typical student's application budget accommodates at most a dozen or so schools and is constant in $m$.
 \else
@@
 정확한 동적계획 해법이 전책적으로 가장 빠르며 FPTAS가 약간 느린 것으로 나온 결과는 배낭 문제 해법을 비교한 기존 연구와 비슷하다 (Martello와 Toth 1990, \S\,2.10).
-분지한계법은 크기가 중간인 인스턴스에 대해서도 비효율적이었다.
-\fi
-
-\newcommand{\missing}{---}
+분지한계법은 크기가 중간인 인스턴스에도 비효율적이었다.
+
+알고리즘 \ref{ellisDP1}의 좋은 실험적 성능은 부분적으로 가상 인스턴스의 구조 덕이다.
+가상 인스턴스의 지원 비용이 작은 정수이고 $H$가 $m$과 선형식으로 비례하는 것은 알고리즘의 기대 계산 시간이 $O(m^2)$임을 의미한다.
+그러나 전형적인 학생의 지원 예산은 최대한 열 몇 개의 학교에 지원할 수 있게 하며 $m$에 대해 상수이므로, 실제 대학 지원 문제 인스턴스에 적용하면 알고리즘 \ref{ellisDP1}이 더욱 뚜렷한 상대적 우위를 발휘한다.
+\fi
+
 \newcommand{\lastptofcaption}{\ifen
@@
  \fi
-    16 &     0.00 &  (0.00) &      0.01 &    (0.00) \\
-    64 &     0.01 &  (0.00) &      0.13 &    (0.03) \\
-   256 &     0.14 &  (0.02) &      1.56 &    (0.57) \\
-  1024 &     2.21 &  (0.35) &     41.65 &   (10.55) \\
-  4096 &    35.36 &  (4.96) &   5080.63 & (6266.66) \\
- 16384 &   552.45 & (79.57) &  16452.20 & (5525.29)       
+    16 &           0.00 &        (0.00) &           0.02 &        (0.01) \\
+    64 &           0.03 &        (0.00) &           0.17 &        (0.05) \\
+   256 &           0.14 &        (0.03) &           2.20 &        (0.73) \\
+  1024 &           2.08 &        (0.65) &          30.51 &        (6.37) \\
+  4096 &          28.87 &       (10.60) &         485.80 &       (90.50) \\
+ 16384 &         369.79 &      (127.38) &        9780.83 &     (1888.34)  
 \end{tabular}
@@
 \fi
-   8 &          0.44 &       (0.42) &         0.06 &      (0.02) &                0.39 &             (0.16) &                 1.66 &              (0.64) \\
-  16 &       3298.08 &    (5285.45) &         0.48 &      (0.16) &                2.64 &             (0.96) &                19.20 &              (7.05) \\
-  32 &             — &          (—) &         2.30 &      (0.83) &               19.01 &             (9.18) &               184.00 &             (70.15) \\
-  64 &             — &          (—) &         8.33 &      (3.52) &              101.05 &            (59.47) &              2089.30 &            (961.43) \\
- 128 &             — &          (—) &        36.61 &     (15.36) &              481.61 &           (210.68) &              8626.29 &           (2615.38) \\
- 256 &             — &          (—) &       205.57 &     (52.98) &             3106.71 &           (957.95) &             69291.93 &          (34881.04) \\
- 512 &             — &          (—) &      1136.18 &    (396.24) &            16613.99 &          (6241.99) &            206020.31 &          (45217.42) 
+    8 &          0.05 &       (0.03) &         0.02 &      (0.01) &                0.08 &             (0.02) &                 0.28 &              (0.09) \\
+   16 &          0.37 &       (0.25) &         0.10 &      (0.03) &                0.53 &             (0.15) &                 4.39 &              (1.27) \\
+   32 &        131.58 &     (172.60) &         0.40 &      (0.10) &                3.08 &             (0.70) &                45.07 &             (15.72) \\
+   64 &             — &          (—) &         1.78 &      (0.46) &               20.91 &             (4.81) &               566.75 &            (149.18) \\
+  128 &             — &          (—) &         9.34 &      (1.98) &              121.28 &            (33.50) &              3090.48 &            (863.95) 
 \end{tabular}%
@@
 The general problem is NP-complete.
-We provided three solution algorithms; the strongest from a theoretical standpoint is an FPTAS that produces a $(1-\varepsilon)$-approximate solution in $O(m^3 / \varepsilon)$-time.
+We provided three solution algorithms.
+The strongest from a theoretical standpoint is an FPTAS that produces a $(1-\varepsilon)$-approximate solution in $O(m^3 / \varepsilon)$-time.
+However, for typical college-application instances, the dynamic program based on application expenditures is both easier to implement and substantially more time efficient.
 \else
@@
 일반적인 문제는 NP-complete다.
-3가지 해법을 제시했으며 이론적으로 가장 효율적인 알고리즘은 $O(m^3 / \varepsilon)$-시간 안에  $(1-\varepsilon)$-근사해를 계산하는 FPTAS다.
+3가지 해법을 제시했으며 이론적으로 가장 효율적인 알고리즘은 $O(m^3 / \varepsilon)$-시간 안에  $(1-\varepsilon)$-근사해를 계산하는 FPTAS이다.
+그러나 전형적인 대학 지원 문제 인스턴스에 대해서, 지원 비용 지출액 기반 동적 계획 해법은 실행하기 쉬울 뿐더러 계산 시간이 상당히 짧다.
 \fi
@@
 
-An analogous phenomenon in the American college admissions process is known as early decision, in which at the moment of application, a student commits to enrolling in a college if admitted.
+An analogous feature of the American college admissions process is known as early decision, in which at the moment of application, a student commits to enrolling in a college if admitted.
 In principle, students who apply with early decision may obtain a better chance of admission by signaling their eagerness to attend, but doing so removes the college's incentive to entice the student with merit-based financial aid, altering the utility calculus.
@@
 
-미국 입학 과정의 유사한 점은 조기 전형(early decision)이다. 이는 학생이 지원하는 시기에 합격하면 진학한다고 약속하는 지원 방법이다. 조기 전형으로 지원하면 입학할 관심을 시그널할 수 있으므로 합격 확률을 높일 수 있다. 반면에 학교가 학생을 모집하려고 장학금을 수여할 동기가 없어질 수 있으므로 학생의 효용 추정이 달라진다.
+미국 입학 과정의 유사한 특징은 조기 전형(early decision)이다. 이는 학생이 합격하면 진학한다고 약속하고 지원하는 것을 의미한다. 조기 전형으로 지원하면 입학할 관심을 시그널할 수 있으므로 합격 확률을 높일 수 있는 반면에, 학교가 학생을 모집하려고 장학금을 수여할 동기가 없어질 수 있으므로 학생의 효용 추정이 달라진다.
 \fi

### 2022-03-30 benchmark/benchmark.jl

@@ -9,7 +9,7 @@ import Printf: @sprintf
 # using BenchmarkTools
 using UnicodePlots
 
-const fullscale = false
+const fullscale = true
 
 # A long benchmark; tweak parameters with caution.
 # Set fullscale = false to run a smaller benchmark to check formatting etc.
@@ -21,14 +21,14 @@ const n_reps = fullscale ? 3 : 2
 const n_markets = fullscale ? 50 : 2
 
 # Cutoffs to exclude certain large computations
-const bnbcutoff = fullscale ? 25 : 6
+const bnbcutoff = fullscale ? 33 : 6
 const twottbnbcutoff = 2^bnbcutoff
 const fptascutoff_m = fullscale ? 300 : 11
-const fptascutoff_eps = 0.1
+const fptascutoff_eps = fullscale ? 0.0 : 0.1
 
 # Sizes of markets to test
 const marketsizes_SCM = fullscale ? 4 .^ (2:7) : [5, 10, 15]
-const marketsizes_VCM = fullscale ? 2 .^ (3:9) : [5, 10, 15]
+const marketsizes_VCM = fullscale ? 2 .^ (3:7) : [5, 10, 15]
 const epsilons = [0.5, 0.05]
 
 const mkts_SCM = SameCostsMarket[SameCostsMarket(m) for m in marketsizes_SCM, i in 1:n_markets]
@@ -43,7 +43,7 @@ function makeunicodeplots(df::DataFrame)
             ylabel=c,
             title="Time when m = $(marketsizes_SCM[end])"
         )
-        for c in setdiff(names(df), ("m", "time_bnb"))
+        for c in setdiff(names(df), ("m", "time_bnb")) if !(Inf in df[!, c])
     ]
 end
 
@@ -69,17 +69,14 @@ function benchmark1()
     printheader("Benchmark 1: Homogeneous-cost algorithms")
 
     sizes = Int[m for m in marketsizes_SCM, _ in 1:n_markets]
-    times_list = fill(Inf, n_markets, length(marketsizes_SCM))
-    times_heap = fill(Inf, n_markets, length(marketsizes_SCM))
+    times_list = fill(Inf, length(marketsizes_SCM), n_markets)
+    times_heap = fill(Inf, length(marketsizes_SCM), n_markets)
 
     @threads for j in 1:n_markets
-        println("  j = $j of $n_markets")
-        for i in 1:length(marketsizes_SCM), _ in 1:n_reps
+        println("  j = ", @sprintf("%2d", j), " of $n_markets")
[…]

### 2022-03-30 src/FastSolvers.jl

@@ -23,8 +23,6 @@ end
 # Overload isless() so that the heap is ordered by expected utility.
 isless(c1::College, c2::College) = isless(c1.ft, c2.ft)
 
-const dummy_college = College(UInt8(0), 1.0, -1.0, -1.0, 0.0)
-
 
 """
     applicationorder_list(mkt::SameCostsMarket)
@@ -37,7 +35,9 @@ function applicationorder_list(mkt::SameCostsMarket{T}; verbose=false::Bool)::Tu
     v = zeros(mkt.h)
 
     mkt_list = College{T}[College(T(j), mkt.f[j], mkt.t[j], mkt.ft[j], mkt.omf[j]) for j in T(1):mkt.m]
-    
+
+    dummy_college = College(T(0), 1.0, -1.0, -1.0, 0.0)
+
     c_best::College{T}, idx_best::Int = findmax(mkt_list)
     @inbounds for j in T(1):mkt.h
         if verbose
@@ -110,8 +110,9 @@ function applicationorder_heap(mkt::SameCostsMarket{T})::Tuple{Vector{Int},Vecto
         c_k = first(mkt_heap)
         v[j] = get(v, j - 1, 0) + c_k.ft
         apporder[j] = c_k.j
-
-        mkt_heap = BinaryMaxHeap{College{T}}(collect(
+    
+        mkt_heap = BinaryMaxHeap{College{T}}(
+            map(filter(c -> c.j != c_k.j, mkt_heap.valtree)) do c
                 College(
                     c.j,
                     c.f,
@@ -119,11 +120,7 @@ function applicationorder_heap(mkt::SameCostsMarket{T})::Tuple{Vector{Int},Vecto
                     c.t < c_k.t ? c.ft * c_k.omf : c.ft - c.f * c_k.ft,
                     c.omf
                 )
-                # This heap.valtree field is not documented in DataStructures.jl,
-                # but it contains the heap data in an arbitrary order, which is exactly 
-                # what we need. Equiv. to heap.drain!() in Rust except doesn't mutate.
-                for c in (d for d in mkt_heap.valtree if d.j != c_k.j)
-            )
+            end
         )
     end
 
@@ -165,20 +162,20 @@ end
 
 # Used by dynamic program below
 @inbounds function V_recursor!(V_dict::Dict{Tuple{T,Int},Float64}, j::T, h::Int, mkt::VariedCostsMarket{T})::Float64 where T
[…]

### 2022-03-31 paper/paper.tex
paragraph about reducing memory usage of DPs

@@
 %\end{definition}
-\ifen Here $\mathcal{C} = \{ 1 \dots m\}$ is an index set; $H > 0$ is a budget parameter; for $j = 1 \dots m$, $g_j > 0$ is a cost parameter and $Z_j$ is a random, independent Bernoulli variable with probability $f_j$; and for $j = 0\dots m$, $t_j$ is a utility parameter.
-\else 단,  $\mathcal{C} = \{ 1 \dots m\}$은 지표 집합이며 $H$는 예산을 나타내는 모수이다. 각 $j = 1 \dots m$에 대해 $g_j > 0$는 비용 모수이며 $Z_j$는 확률 $f_j$를 가지는 서로 독립적인 Bernoulli 변수이다. 각 $j = 0 \dots m$에 대해 $t_j$는 효용 모수이다. \fi
+\ifen Here $\mathcal{C} = \{ 1 \dots m\}$ is an index set; $H > 0$ is a budget parameter; for $j = 1 \dots m$, $g_j > 0$ is a cost parameter and $Z_j$ is a random, independent Bernoulli variable with probability $f_j$; and for $j = 0\dots m$, $t_j\geq 0$ is a utility parameter.
+\else 단,  $\mathcal{C} = \{ 1 \dots m\}$은 지표 집합이며 $H$는 예산을 나타내는 모수이다.
+각 $j = 1 \dots m$에 대해 $g_j > 0$는 비용 모수이며 $Z_j$는 확률 $f_j$를 가지는 서로 독립적인 Bernoulli 변수이다.
+각 $j = 0 \dots m$에 대해 $t_j\geq 0$는 효용 모수이다.
+\fi
 
@@
 대학 지원 최적화 문제는 이론적인 화제일 뿐만이 아니다.
-미국 입학 컨설팅 산업에서 대학 지원 서류 작성, 합격 확률 추정, 그리고 지원 학교를 선택하는 데에 자문하는 개인 상담가의 시간당 비율은 평균 200달러이다 (Sklarow 2018).
+미국 입학 컨설팅 산업에서 대학 지원 서류 작성, 합격 확률 추정, 그리고 지원 학교를 선택하는 데에 자문하는 개인 상담가의 시간당 급료는 평균 200달러이다 (Sklarow 2018).
 한국에서, 메가스터디(\url{megastudy.net})와 진학(\url{jinhak.com})과 같은 입학 컨설팅 기업의 주된 수입원 중 인공지능을 활용해서 학생의 지원 전략을 최대화한다고 홍보하는 ``모의 지원'' 소프트웨어가 있다.
@@
 \ifen \subsection{Structure of this paper} \else \subsection{본 논문의 구성}\fi
-\ifen Section \ref{preliminaries} introduces some additional notation and assumptions that can be imposed with trivial loss of generality. 
+\ifen Section \ref{preliminaries} introduces some additional notation and assumptions that can be imposed without loss of generality. 
 \else \ref{preliminaries}절은 표기법과 일반성을 제한하지 않는 편의적 가정을 제시한다.\fi
@@
 \ifen
-In section \ref{homogappcosts}, we consider the special case where each $g_j = 1$ and $H$ is an integer $h \leq m$.
+In Section \ref{homogappcosts}, we consider the special case where each $g_j = 1$ and $H$ is an integer $h \leq m$.
 We show that an intuitive heuristic is in fact a $1/h$-approximation algorithm.
@@
 \ifen
-In section \ref{hetappcosts}, we turn to the scenario in which colleges differ in their application fees.
+In Section \ref{hetappcosts}, we turn to the scenario in which colleges differ in their application fees.
 We show that the decision form of the portfolio optimization problem is NP-complete through a polynomial reduction from the binary knapsack problem.
@@
 \ifen
-In section \ref{numericalexperiments}, we present the results of computational experiments that confirm the validity and time complexity results established in the previous two sections.
+In Section \ref{numericalexperiments}, we present the results of computational experiments that confirm the validity and time complexity results established in the previous two sections.
 
@@
 \begin{split}
-v(\mathcal{X}_h) &= \operatorname{E}\bigl[ \max_{j \in \mathcal{X}_h}\{ t_j Z_j \}\bigr] \leq \operatorname{E}\bigl[ \sum_{j \in \mathcal{X}_h}\{ t_j Z_j \}\bigr] \leq \operatorname{E}\bigl[ \sum_{j \in \mathcal{T}_h}\{ t_j Z_j \}\bigr] \\
-&= h  \operatorname{E}\bigl[ \tfrac{1}{h} \sum_{j \in \mathcal{T}_h}\{ t_j Z_j \}\bigr]
-\leq h  \operatorname{E}\bigl[ \max_{j \in \mathcal{T}_h}\{ t_j Z_j \}\bigr]
-= h\,v(\mathcal{T}_h).
+v(\mathcal{X}_h) &= \operatorname{E}\Bigl[ \max_{j \in \mathcal{X}_h}\{ t_j Z_j \}\Bigr] \leq \operatorname{E}\Bigl[ \sum_{j \in \mathcal{X}_h}\{ t_j Z_j \}\Bigr] \leq \operatorname{E}\Bigl[ \sum_{j \in \mathcal{T}_h}\{ t_j Z_j \}\Bigr] \\
+&= h  \operatorname{E}\Bigl[ \tfrac{1}{h} \sum_{j \in \mathcal{T}_h}\{ t_j Z_j \}\Bigr]
+\leq h  \operatorname{E}\Bigl[ \max_{j \in \mathcal{T}_h}\{ t_j Z_j \}\Bigr]
+= h v(\mathcal{T}_h).
 \end{split}
@@
 \begin{proof}\ifen
-By induction on $h$. Applying Theorem \ref{assumetzerozero}, we assume that $t_0 = 0$.
+By induction on $h$.
+Applying Theorem \ref{assumetzerozero}, we assume that $t_0 = 0$.
+We also assume that each $f_j < 1$, which is without loss of generality because any school having $f_j = 1$ can be eliminated by setting $t_0 = t_j$ and reapplying the transformation from Theorem \ref{assumetzerozero}.
 
@@
 Optimality requires that
-\else$h$에 대해 귀납법을 적용한다. 정리 \ref{assumetzerozero}에 따라 $t_0 = 0$이라고 가정한다.
+\else$h$에 대해 귀납법을 적용한다. 정리 \ref{assumetzerozero}에 따라 $t_0 = 0$이라고 가정한다. 또한 모든 $f_j < 1$임을 가정한다. $f_j = 1$인 학교가 있으면 $t_0 = t_j$로 수정하고  정리 \ref{assumetzerozero}의 변환을 다시 적용해서 소거할 수 있으므로 일반성을 제한하지 않는다.
 
@@
 The heap is ordered according to the criterion $i \geq j \iff f_i t_i \geq f_j t_j$, and by draining the heap and reheapifying at the end of each iteration, the computation time remains $O(hm)$.
-However, in our numerical experiments, whose results are reported in section \ref{numericalexperiments},
+However, in our numerical experiments, whose results are reported in Section \ref{numericalexperiments},
 we found the list implementation to be much faster, because it is possible to identify the entering school $k$ as the utility parameters are updated, which all but eliminates the cost of the $\argmax\{\}$ operation.
@@
 then it is evident that the optimal portfolio for $H = 2$ is $\{1, 2\}$ while that for $H = 3$ is $\{3\}$.
-In fact, Ellis’s problem is NP-complete, as we will show by a transformation from the binary knapsack problem, which is known to be NP-complete (Garey and Johnson 1979).
+In fact, Ellis’s problem is NP-complete, as we will show by a transformation from the binary knapsack problem, which is known to be NP-complete (Garey and Johnson 1979, \S\,3.2.1).
 \else
@@
 사실은 엘리스의 문제가 NP-complete이며 이를 이진 배낭 문제에서의 변환을 통해 증명할 수 있다.
-배낭 문제의 NP-completeness가 이미 알려져 있다 (Garey and Johnson 1979).
+배낭 문제의 NP-completeness가 이미 알려져 있다 (Garey and Johnson 1979, \S\,3.2.1).
 \fi
@@
 While $H > 0$, select the school $j$ for which $f_j t_j / g_j$ is highest, set $x_j \gets \min\{1, H / g_j\}$, and set $H \gets H - g_j$ (Dantzig 1957).
+Balas and Zemel (1980, \S\,2) provide an $O(m)$ algorithm.
 \else
@@
 $x_j \gets \min\{1, H / g_j\}$ 그리고 $H \gets H - g_j$로 수정해서 반복한다 (Dantzig 1957).
+$O(m)$ 알고리즘도 알려져 있다 (Balas와 Zemel 1980, \S\,2).
 \fi
@@
 For each $j\in \mathcal{I}$,  we increment $\gamma$ by the current value of $f_j \bar t_j$, eliminate $j$ from the market, and update the remaining $\bar t_j$-values using \eqref{howtotransformfj}.
-(An example is provided below.)
+(An example is given below.)
 \else
@@
 \else
-$j = 0 \dots m$와 $h = 0 \dots H$에 대해, $\{ 1, \dots, j\}$에 속한 학교만 사용하면서 최대한 $h$의 지원 비용을 지불하는 최적 포트폴리오를 $\mathcal{X}[j, h]$라고하자.
+$j = 0 \dots m$와 $h = 0 \dots H$에 대해, $\{ 1, \dots, j\}$에 속한 학교만 사용하면서 지원 지출액이 $h$을 넘지 않는 최적 포트폴리오를 $\mathcal{X}[j, h]$라고 하자.
 또한  $V[j,h] = v(\mathcal{X}[j, h])$이다.
@@
 \else
-포티폴리오의 근사적 가치를 정확도 $P$로 구성된 고정소수점 십진수(fixed-point decimal)로 나타내자.
+포트폴리오의 근사적 가치를 정확도 $P$로 구성된 고정소수점 십진수(fixed-point decimal)로 나타내자.
 다, $P$는 소수점 뒤에 등장하는 숫자의 수다.
@@
 \ifen
-We chose to implement our algorithms in the Julia language (v1.7.0) because its system of parametric data types allows the compiler to optimize for differential cases such as when the $t_j$-values are integers or floating-point numbers.
+We chose to implement our algorithms in the Julia language (v1.8.0b1) because its system of parametric data types allows the compiler to optimize for differential cases such as when the $t_j$-values are integers or floating-point numbers.
 Julia also offers convenient macros for parallel computing, which enabled us to solve more and larger problems in the benchmark (Bezanson et al.
 2017).
-We made extensive use of the DataStructures.jl package (v0.18.11).
+We made extensive use of the DataStructures.jl package (v0.18.11, \url{github.com/JuliaCollections}).
 The code is available at \url{https://github.com/maxkapur/OptimalApplication}.
 \else
-모든 알고리즘을 쥴리아 (Julia, v1.7.0) 언어로 실행했다.
+모든 알고리즘을 쥴리아 (Julia, v1.8.0b1) 언어로 실행했다.
 쥴리아는 $t_j$-값이 정수 또는 부동소수점 수와 같은 특수한 경우에 대해 컴파일러가 자동적으로 최적화하는 모수적 자료형(parametric data type) 기능이 있어서 유용한 선택이었다.
 또한 병렬 연산을 쉽게 관리할 수 있는 매크로를 제공하므로 계산 실험의 규모를 증가할 수 있었다 (Bezanson 외 2017).
-DataStructures.jl 패키지(v0.18.11)도 활용했다.
+DataStructures.jl 패키지(v0.18.11, \url{github.com/JuliaCollections})도 활용했다.
 실험 코드는  \url{https://github.com/maxkapur/OptimalApplication}에서 공유한다.
@@
 The dynamic programs, namely Algorithms \ref{ellisDP1} and \ref{ellisDP3}, were implemented using recursive functions and dictionary memoization rather than a full lookup table.
-Our implementation of Algorithm \ref{ellisDP3} also differed from that described in subsection \ref{fptashead} in that we represented portfolio valuations in \emph{binary} rather than decimal, with the definitions of $P$ and $\mathcal{V}$ modified accordingly, and instead of fixed-point numbers, we worked in integers by multiplying each $t_j$-value by $2^P$.
+Our implementation of Algorithm \ref{ellisDP3} also differed from that described in Subsection \ref{fptashead} in that we represented portfolio valuations in \emph{binary} rather than decimal, with the definitions of $P$ and $\mathcal{V}$ modified accordingly, and instead of fixed-point numbers, we worked in integers by multiplying each $t_j$-value by $2^P$.
 These modifications yielded a substantial performance improvement without changing the fundamental algorithm design or complexity analysis.
@@
  \fi
-    16 &           0.00 &        (0.00) &           0.02 &        (0.01) \\
-    64 &           0.03 &        (0.00) &           0.17 &        (0.05) \\
-   256 &           0.14 &        (0.03) &           2.20 &        (0.73) \\
-  1024 &           2.08 &        (0.65) &          30.51 &        (6.37) \\
-  4096 &          28.87 &       (10.60) &         485.80 &       (90.50) \\
- 16384 &         369.79 &      (127.38) &        9780.83 &     (1888.34)  
+    16 &           0.00 &        (0.00) &           0.01 &        (0.00) \\
+    64 &           0.03 &        (0.00) &           0.08 &        (0.02) \\
+   256 &           0.15 &        (0.01) &           0.97 &        (0.22) \\
+  1024 &           2.31 &        (0.05) &          14.44 &        (1.66) \\
+  4096 &          37.85 &        (0.61) &         245.74 &       (17.33) \\
+ 16384 &         585.75 &        (2.11) &        4728.59 &      (552.25)
 \end{tabular}
@@
 \ifen
-\textbf{\begin{tabular}[r]{@{}r@{}}Number of\\schools $m$\end{tabular}}&\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisbnb}:\\Branch \& bound\end{tabular}}}  & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP1}:\\App.
-costs DP\end{tabular}}}  &\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.5$\end{tabular}}}  & \multicolumn{2}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.05$\end{tabular}}}   \\ \hline
+\textbf{\begin{tabular}[r]{@{}r@{}}Number of\\schools $m$\end{tabular}}&\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisbnb}:\\Branch \& bound\end{tabular}}}  & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP1}:\\Costs DP\end{tabular}}}  &\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.5$\end{tabular}}}  & \multicolumn{2}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Algorithm \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.05$\end{tabular}}}   \\ \hline
 \else
@@
 \fi
-    8 &          0.05 &       (0.03) &         0.02 &      (0.01) &                0.08 &             (0.02) &                 0.28 &              (0.09) \\
-   16 &          0.37 &       (0.25) &         0.10 &      (0.03) &                0.53 &             (0.15) &                 4.39 &              (1.27) \\
-   32 &        131.58 &     (172.60) &         0.40 &      (0.10) &                3.08 &             (0.70) &                45.07 &             (15.72) \\
-   64 &             — &          (—) &         1.78 &      (0.46) &               20.91 &             (4.81) &               566.75 &            (149.18) \\
-  128 &             — &          (—) &         9.34 &      (1.98) &              121.28 &            (33.50) &              3090.48 &            (863.95) 
+     8 &          0.04 &       (0.02) &         0.01 &      (0.00) &                0.05 &             (0.01) &                 0.21 &              (0.06) \\
+    16 &          0.22 &       (0.11) &         0.07 &      (0.01) &                0.43 &             (0.10) &                 3.15 &              (0.74) \\
+    32 &        166.20 &     (422.31) &         0.31 &      (0.05) &                2.38 &             (0.38) &                33.38 &             (11.68) \\
+    64 &             — &          (—) &         1.36 &      (0.18) &               15.32 &             (2.77) &               405.73 &            (125.98) \\
+   128 &             — &          (—) &         6.52 &      (0.72) &               84.50 &            (22.59) &              2362.19 &           (1095.11) \\
+   256 &             — &          (—) &        31.23 &      (1.91) &             1085.60 &          (1186.24) &             22129.92 &           (6588.40)
 \end{tabular}%
 \caption{\label{experiment2results} \normalsize
-\ifen Time in ms to compute an optimal or $(1 - \varepsilon)$-optimal portfolio for an admissions market with heterogeneous application costs using the three algorithms developed in section \ref{hetappcosts}.
+\ifen Time in ms to compute an optimal or $(1 - \varepsilon)$-optimal portfolio for an admissions market with heterogeneous application costs using the three algorithms developed in Section \ref{hetappcosts}.
 %branch-and-bound routine (Algorithm \ref{ellisbnb}), exact dynamic program (Algorithm \ref{ellisDP1}), and the fully polynomial-time approximation scheme (Algorithm \ref{ellisDP3}).
@@
 %\pagebreak
-\ifen \section{Conclusion} \else \section{결론}\fi\label{conclusion}
+\ifen \section{Conclusion and ideas for future research} \else \section{결론과 향후 연구}\fi\label{conclusion}
 \ifen
 This study has introduced a novel combinatorial optimization problem that we call the college application problem.
-It can be viewed as a kind of integer portfolio optimization problem or stochastic knapsack problem.
-We showed that the special case in which each college's application cost is identical can be solved in polynomial time by a greedy algorithm because the optimal solutions are nested in the budget constraint.
+It can be viewed as a kind of integer portfolio optimization problem or nonlinear knapsack problem.
+We showed that the special case in which colleges have identical application costs can be solved in polynomial time by a greedy algorithm because the optimal solutions are nested in the budget constraint.
 The general problem is NP-complete.
@@
 The strongest from a theoretical standpoint is an FPTAS that produces a $(1-\varepsilon)$-approximate solution in $O(m^3 / \varepsilon)$-time.
-However, for typical college-application instances, the dynamic program based on application expenditures is both easier to implement and substantially more time efficient.
+On the other hand, for typical college-application instances, the dynamic program based on application expenditures is both easier to implement and substantially more time efficient.
+
+Three extensions of this problem appear amenable to future study.
 \else
@@
 3가지 해법을 제시했으며 이론적으로 가장 효율적인 알고리즘은 $O(m^3 / \varepsilon)$-시간 안에  $(1-\varepsilon)$-근사해를 계산하는 FPTAS이다.
-그러나 전형적인 대학 지원 문제 인스턴스에 대해서, 지원 비용 지출액 기반 동적 계획 해법은 실행하기 쉬울 뿐더러 계산 시간이 상당히 짧다.
-\fi
-
-\ifen
-Two extensions of this problem appear amenable to future study.
-First, the familiar Markowitz portfolio optimization model includes an explicit risk-aversion term, whereas our model has settled for an implicit treatment of risk, namely the tradeoff between selective schools and safety schools inherent in the ``maximax'' objective function.
+그 반면에 전형적인 대학 지원 문제 인스턴스에 대해서, 지원 비용 지출액 기반 동적 계획 해법은 실행하기 쉬울 뿐더러 계산 시간이 상당히 짧다.
+
+이 문제를 확장할 수 있는 향후 연구 방향 3가지를 제안한다. 
+\fi
+
+\ifen \subsection{Explicit treatment of risk aversion}\else \subsection{위험 회피가 명시적인 모형}\fi
+\ifen
+The familiar Markowitz portfolio optimization model includes an explicit risk-aversion term, whereas our model has settled for an implicit treatment of risk, namely the tradeoff between selective schools and safety schools inherent in the maximax objective function.
 However, it is possible to augment the objective function $v(\mathcal{X}) = \operatorname{E}[X]$ to incorporate a variance penalty $\beta \geq 0$ as follows:
 \else
-이 문제를 확장할 수 있는 향후 연구 방향 2가지를 제안한다.
-첫번째는, 익숙한 Markowitz 포트폴리오 최적화 모형에는 명시적인 위험 회피 항이 있으며, 본 모형은 경쟁적인 학교와 안정 학교 사이의 균형으로 내포된 위험 관리 요소만 고려했다.
+익숙한 Markowitz 포트폴리오 최적화 모형에는 명시적인 위험 회피 항이 있으며, 본 모형은 maximax 목적함수에 내포된 경쟁적인 학교와 안정 학교 사이의 균형에 의한 위험 관리 요소만 고려했다.
 그러나 기본 목적함수인 $v(\mathcal{X}) = \operatorname{E}[X]$에 평균편차 페널티 $\beta \geq 0$를 다음처럼 도입할 수 있다:
@@
 
+\ifen \subsection{Signaling strategies}\else \subsection{시그런 전략}\fi
 \ifen
@@
 An analogous feature of the American college admissions process is known as early decision, in which at the moment of application, a student commits to enrolling in a college if admitted.
-In principle, students who apply with early decision may obtain a better chance of admission by signaling their eagerness to attend, but doing so removes the college's incentive to entice the student with merit-based financial aid, altering the utility calculus.
+In principle, students who apply with early decision may obtain a better chance of admission by signaling their eagerness to attend, but doing so weakens the college's incentive to entice the student with merit-based financial aid, altering the utility calculus.
 \else
@@
 
-미국 입학 과정의 유사한 특징은 조기 전형(early decision)이다. 이는 학생이 합격하면 진학한다고 약속하고 지원하는 것을 의미한다. 조기 전형으로 지원하면 입학할 관심을 시그널할 수 있으므로 합격 확률을 높일 수 있는 반면에, 학교가 학생을 모집하려고 장학금을 수여할 동기가 없어질 수 있으므로 학생의 효용 추정이 달라진다.
+미국 입학 과정의 유사한 특징은 조기 전형(early decision)이다. 이는 학생이 합격하면 진학한다고 약속하고 지원하는 것을 의미한다. 조기 전형으로 지원하면 입학할 관심을 시그널할 수 있으므로 합격 확률을 높일 수 있는 반면에, 학교가 학생을 모집하려고 장학금을 수여할 동기가 약화할 수 있으므로 학생의 효용 추정이 달라진다.
 \fi
@@
 \end{align}
-completes the problem formulation. These constraints are conformable to those of the multiple knapsack problem (Kellerer et al. 2004, chap. 10). 
+completes the formulation. These constraints are conformable to those of the multiple knapsack problem (Kellerer et al. 2004, chap. 10). 
 \else
@@
 
+\ifen \subsection{Memory-efficient dynamic programs}\else \subsection{동적 계획의 메모리 소요 절감}\fi
+\ifen
+Our numerical experiments suggest that the performance of the dynamic programming algorithms is bottlenecked by not computation time, but memory usage.
+Reducing these algorithms' storage requirements would enable us to solve considerably larger problems.
+
+Abstractly speaking, Algorithms \ref{ellisDP1} and \ref{ellisDP3} are two-dimensional dynamic programs that represent the optimal solution by $Z[N, C]$.
+Here $N$ is the number of decision variables, $C$ is a constraint parameter, and $Z[n, c]$ is the optimal objective value achievable using the first $n\leq N$ decision variables when the constraint parameter is $c \leq C$.
+The algorithm iterates on a recursion relation that expresses $Z[n, c]$ as a function of $Z[n -1, c]$ and $Z[n -1, c']$ for some $c' \leq c$.
+(In Algorithm \ref{ellisDP1}, $Z$ is the maximal portfolio valuation, $N = m$, and $C =H$; in Algorithm \ref{ellisDP3},  $Z$ is the minimal application expenditures, $N = m$, and $C = |\mathcal{V}| \propto m^2 / \varepsilon$.)
+
+When such a dynamic program is implemented using a lookup table or dictionary memoization, producing the optimal solution requires $O(NC)$-time and -space.
+Kellerer et al.
+(2004, \S\,3.3) provide a technique for transforming the dynamic program $Z$ into a divide-and-conquer algorithm that produces the optimal solution in $O(N C)$-time and $O(N + C)$-space, a significant improvement.
+However, their technique requires the objective function to be additively separable in a certain sense that appears difficult to conform to the college application problem.
+\else
+수리 실험 결과에 따르면 동적 계획 해법 성능의 목병 요소는 계산 시간이 아니라 메모리 소모량이다. 이 알고리즘의 메모리 소요를 절감하면 상당히 큰 문제를 풀 수 있다.
+
+추상적인 관점에서, 알고리즘 \ref{ellisDP1}과(와) 알고리즘 \ref{ellisDP3}은(는) 최적해를 $Z[N, C]$로 표현하는 2차원 동적 계획이다.
+이때 $N$은 결정 변수의 수, $C$는 제약식의 모수, 그리고 $Z[n, c]$는 제약 모수가 $c \leq C$일 때 첫 $n\leq N$개의 결정 변수만 이용하는 최적해이다.
+알고리즘의 구조는 $Z[n, c]$를 $Z[n -1, c]$와 어떤 $c' \leq c$에 대응하는 $Z[n -1, c']$의 함수로 표현하는 반복 관계로 탐색하는 것이다.
+(알고리즘 \ref{ellisDP1}에서 $Z$는 최대 포트폴리오 가치, $N = m$, 그리고 $C =H$이며, 알고리즘 \ref{ellisDP3}에서 $Z$는 최소한 지원 지출액, $N = m$, 그리고 $C= |\mathcal{V}| \propto m^2 / \varepsilon$이다.)
+
+이러한 형태의 동적 계획을 표 혹은 사전 기록으로 구현하면 최적해를 출력하는 것은 $O(NC)$-시간과 -공간을 소요한다.
+Kellerer외 (2004, \S\,3.3)는 동적 계획 $Z$를 $O(N C)$-시간과 $O(N + C)$-공간으로 최적해를 구하는 분할 정복(divide and conquer) 알고리즘으로 변환하는 일잔적인 방법을 제시하며 매우 유익한 기술이다.
+그러나 이를 이용할 수 있는 조건 중, 목적함수는 어떤 기술적인 의미에서 가산적으로 분할되어야 하며 이를 대학 지원 문제에 적용하기 어렵다.
+\fi
+
 \pagebreak
@@
 ``A Comparison of Regression Models for Prediction of Graduate Admissions.'' In \emph{Second International Conference on Computational Intelligence in Data Science.} \url{https://doi.org/10.1109/ICCIDS.2019.8862140}.
+
+Balas, Egon and Eitan Zemel. 1980. ``An Algorithm for Large Zero-One Knapsack Problems.'' \emph{Operations Research} 28 (5): 1130--54. \url{https://doi.org/10.1287/opre.28.5.1130}.
 

### 2022-03-31 benchmark/benchmark.jl

@@ -28,7 +28,7 @@ const fptascutoff_eps = fullscale ? 0.0 : 0.1
 
 # Sizes of markets to test
 const marketsizes_SCM = fullscale ? 4 .^ (2:7) : [5, 10, 15]
-const marketsizes_VCM = fullscale ? 2 .^ (3:7) : [5, 10, 15]
+const marketsizes_VCM = fullscale ? 2 .^ (3:8) : [5, 10, 15]
 const epsilons = [0.5, 0.05]
 
 const mkts_SCM = SameCostsMarket[SameCostsMarket(m) for m in marketsizes_SCM, i in 1:n_markets]

### 2022-04-01 paper/paper.tex
paragraph about risk aversion

@@
 \begin{align*}
-t &= \Bigl(\,\underbrace{1,  \dots, 1}_{h},\;
-\underbrace{\varepsilon^{-1}, \varepsilon^{-2}, \dots, \varepsilon^{-(h-1)}, \varepsilon^{-h}}_{h}\,\Bigr) \\
-\text{and}\quad f &= \Bigl(\,\underbrace{1,  \dots, 1}_{h},\;
-\underbrace{\varepsilon^{1}, \varepsilon^{2}, \dots, \varepsilon^{h-1}, \varepsilon^{h}}_{h}\,\Bigr)
+ f &= \Bigl(\,\underbrace{1,  \dots, 1}_{h},\;
+\underbrace{\varepsilon^{1}, \varepsilon^{2}, \dots, \varepsilon^{h-1}, \varepsilon^{h}}_{h}\,\Bigr) \\
+\text{\ifen and\else 그리고\fi}\quad t &= \Bigl(\,\underbrace{1,  \dots, 1}_{h},\;
+\underbrace{\varepsilon^{-1}, \varepsilon^{-2}, \dots, \varepsilon^{-(h-1)}, \varepsilon^{-h}}_{h}\,\Bigr) 
 \end{align*}
@@
 \begin{proof}\ifen
-By induction on $h$.
-Applying Theorem \ref{assumetzerozero}, we assume that $t_0 = 0$.
-We also assume that each $f_j < 1$, which is without loss of generality because any school having $f_j = 1$ can be eliminated by setting $t_0 = t_j$ and reapplying the transformation from Theorem \ref{assumetzerozero}.
+By induction on $h$. Applying Theorem \ref{assumetzerozero}, we assume that $t_0 = 0$.
 
@@
 Optimality requires that
-\else$h$에 대해 귀납법을 적용한다. 정리 \ref{assumetzerozero}에 따라 $t_0 = 0$이라고 가정한다. 또한 모든 $f_j < 1$임을 가정한다. $f_j = 1$인 학교가 있으면 $t_0 = t_j$로 수정하고  정리 \ref{assumetzerozero}의 변환을 다시 적용해서 소거할 수 있으므로 일반성을 제한하지 않는다.
+\else$h$에 대해 귀납법을 적용한다. 정리 \ref{assumetzerozero}에 따라 $t_0 = 0$이라고 가정한다.
 
@@
 
-\ifen Suppose $k \notin \mathcal{X}_h$. To get a contradiction, assume that $v(\mathcal{Y}_h) < v(\mathcal{X}_h)$. Then
-\else $k \notin \mathcal{X}_h$인 경우를 고려하자. 모순을 보이기 위해 $v(\mathcal{Y}_h) < v(\mathcal{X}_h)$라고 하면 \fi
+\ifen Suppose $k \notin \mathcal{X}_h$. To get a contradiction, suppose that $v(\mathcal{Y}_h) < v(\mathcal{X}_h)$ and  $v(\mathcal{X}_{h+1}) > v(\mathcal{X}_h \cup \{k\})$. Then
+\else $k \notin \mathcal{X}_h$인 경우를 고려하자. 모순을 보이기 위해 $v(\mathcal{Y}_h) < v(\mathcal{X}_h)$ 그리고 $v(\mathcal{X}_{h+1}) > v(\mathcal{X}_h \cup \{k\})$라고 가정하면 다음이 성립한다. \fi
 \begin{align}
@@
 &= (1 - f_k) v(\mathcal{Y}_h) + f_k t_k \\
-& < (1 - f_k) v(\mathcal{X}_h) + f_k \operatorname{E}\bigl[ \max\{t_k, X_h\}\bigr]\\
+&\leq (1 - f_k) v(\mathcal{X}_h) + f_k \operatorname{E}\bigl[ \max\{t_k, X_h\}\bigr]\\
 &=  v(\mathcal{X}_h\cup \{k\})
@@
 \end{align}
-\ifen contradicts the optimality of $\mathcal{X}_{h+1}$.
-\else 이는 $\mathcal{X}_{h+1}$의 최적성을 위반한다.\fi
+\ifen is a contradiction.
+\else 모순이다.\fi
 
@@
 \ifen 
-It is also possible to store $\mathcal{C}$ as a binary max heap rather than a list.
+It is possible to store $\mathcal{C}$ as a binary max heap rather than a list.
 The heap is ordered according to the criterion $i \geq j \iff f_i t_i \geq f_j t_j$, and by draining the heap and reheapifying at the end of each iteration, the computation time remains $O(hm)$.
@@
 
-\ifen \subsection{A small example} \else \subsection{작은 예} \fi
+\ifen \subsection{A small example} \else \subsection{작은 예} \fi \label{planetsexamplesection}
 \ifen 
@@
 \ifen 
-An intuitive extension of the greedy algorithm for Alma's problem is to iteratively add to $\mathcal{X}$ the school $k$ for which $[ v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) ] / g_k$ is largest.
+An intuitive extension of the greedy algorithm for Alma's problem is to iteratively add to $\mathcal{X}$ the school $k$ for which $\bigl( v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) \bigr) / g_k$ is largest.
 However, the construction above shows that the objective function of Ellis's problem can approximate that of a knapsack problem with arbitrary precision.
@@
 \else
-알마 문제를 위한 탐욕 해법의 직관적인 확장은 $[ v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) ] / g_k$-값이 가장 높은 학교 $k$를  $\mathcal{X}$에 차례대로 추가하는 것이다.
+알마 문제를 위한 탐욕 해법의 직관적인 확장은 $\bigl( v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) \bigr) / g_k$-값이 가장 높은 학교 $k$를  $\mathcal{X}$에 차례대로 추가하는 것이다.
 그러나 위에서 도출한 변환은 엘리스 문제의 목적함수를 배낭 문제의 목적함수의 원하는 만큼 정확한 근사로 만들 수 있음을 의미한다.
@@
 \else
-익숙한 Markowitz 포트폴리오 최적화 모형에는 명시적인 위험 회피 항이 있으며, 본 모형은 maximax 목적함수에 내포된 경쟁적인 학교와 안정 학교 사이의 균형에 의한 위험 관리 요소만 고려했다.
+익숙한 Markowitz 포트폴리오 최적화 모형에는 명시적인 위험 회피 항이 등장하며, 본 모형은 maximax 목적함수에 내포된 경쟁적인 학교와 안정 지원 학교 사이의 균형에 의한 위험 관리 요소만 고려했다.
 그러나 기본 목적함수인 $v(\mathcal{X}) = \operatorname{E}[X]$에 평균편차 페널티 $\beta \geq 0$를 다음처럼 도입할 수 있다:
@@
 
+A parametric treatment of risk aversion may align our model more closely to real-world applicant behavior.
+Conventional wisdom asserts that the best college application strategy combines reach, target, and safety schools in roughly equal proportion (Kim 2015).
+When the application budget is small relative to the size of the admissions market, our algorithms recommend a similar approach (see the example of Section \ref{planetsexamplesection}).
+However, inspecting equation \eqref{howtotransformfj}, which discounts school utility parameters to reflect their marginal value relative to the schools already in the portfolio, reveals that low-utility schools are penalized more harshly than high utility schools in the marginal analysis.
+Consequentially, as the application budget grows, the optimal portfolio in our model tends to favor reach schools over safety schools.
+(See Figure \ref{samplemarket} for a clear illustration of this phenomenon.)
+
+A potential application of our model is an econometric study that estimates students’ perceptions of college quality ($t_j$) from their observed application behavior ($\mathcal{X}$) under the assumption of utility maximization.
+However, empirical research indicates that risk aversion is a significant factor in human decision-making, including in the domains of educational choice and employment search (Kahneman 2011; Hartlaub and Schneider 2012; van Huizen and Alessie 2019).
+Therefore, an endogenous risk-aversion parameter would greatly enhance our model’s explanatory power in the econometric setting.
+
 \ifen \subsection{Signaling strategies}\else \subsection{시그런 전략}\fi
@@
 Therefore, if a student recognizes that she has a comparative advantage in (for example) her interview skills, she may increase her chances of admission to a competitive school by applying under the \emph{ga} field.
-Clearly, the optimal application strategy in this context can be quite subtle, and are a perennial subject of debate on Korean social networks.
+Clearly, the optimal application strategy in this context can be quite subtle, and it is a subject of perennial debate on Korean social networks.
 
@@
 가령, 어떤 학생이 면접 능력에 비교 우위가 있으면 경쟁적인 학교에 가군으로 지원하면 합격 확률을 높일 수 있다.
-이런 맥락에서 최적의 지원 전략은 미묘한 문제이며 한국 SNS에서 자주 발생하는 화젯거리다.
-
-미국 입학 과정의 유사한 특징은 조기 전형(early decision)이다. 이는 학생이 합격하면 진학한다고 약속하고 지원하는 것을 의미한다. 조기 전형으로 지원하면 입학할 관심을 시그널할 수 있으므로 합격 확률을 높일 수 있는 반면에, 학교가 학생을 모집하려고 장학금을 수여할 동기가 약화할 수 있으므로 학생의 효용 추정이 달라진다.
+이런 맥락에서 최적의 지원 전략은 미묘한 문제이며 한국 SNS에서 자주 발생하는 화젯거리이다.
+
+미국 입학 과정의 유사한 특징은 조기 전형(early decision)이다. 이는 학생이 합격하면 진학한다고 약속하고 지원하는 것을 의미한다. 조기 전형으로 지원하면 입학할 관심을 시그널할 수 있으므로 합격 확률을 높일 수 있지만, 학교가 학생을 모집하려고 장학금을 수여할 동기가 약화할 수 있으므로 학생의 효용 추정이 달라진다.
 \fi
@@
 그다음에 학생이 $c_j$에 높은 관심을 시그널하면 (즉, 가군으로 선택하거나 조기 전형으로 지원하면) $x_{j+} = 1$이며 관심을 시그널 하지 않고 지원하면 (즉 가군 혹은 나군으로 지원하거나 조기 전형 없이 지원하면) $x_{j-} = 1$이 되는 이진 결정 변수를 정의한다.
-그러면 해당 합격 확률 $f_{j+}$와 $f_{j-}$ 그리고 효용  $t_{j+}$와 $t_{j-}$가 알려져 있다고 가정하면 다음 같은 논리 제약식을 더해서 모형을 완성할 수 있다.
+그러면 해당 합격 확률 $f_{j+}$와 $f_{j-}$ 그리고 효용  $t_{j+}$와 $t_{j-}$가 알려져 있다고 가정하면 다음 같은 논리 제약식을 더하면 모형을 완성할 수 있다.
 \begin{align}
@@
 추상적인 관점에서, 알고리즘 \ref{ellisDP1}과(와) 알고리즘 \ref{ellisDP3}은(는) 최적해를 $Z[N, C]$로 표현하는 2차원 동적 계획이다.
-이때 $N$은 결정 변수의 수, $C$는 제약식의 모수, 그리고 $Z[n, c]$는 제약 모수가 $c \leq C$일 때 첫 $n\leq N$개의 결정 변수만 이용하는 최적해이다.
+이때 $N$은 결정 변수의 수, $C$는 제약식의 모수, 그리고 $Z[n, c]$는 제약 모수가 $c \leq C$일 때 첫 $n\leq N$개의 결정 변수만 이용하는 최적 목적함수값이다.
 알고리즘의 구조는 $Z[n, c]$를 $Z[n -1, c]$와 어떤 $c' \leq c$에 대응하는 $Z[n -1, c']$의 함수로 표현하는 반복 관계로 탐색하는 것이다.
@@
 이러한 형태의 동적 계획을 표 혹은 사전 기록으로 구현하면 최적해를 출력하는 것은 $O(NC)$-시간과 -공간을 소요한다.
-Kellerer외 (2004, \S\,3.3)는 동적 계획 $Z$를 $O(N C)$-시간과 $O(N + C)$-공간으로 최적해를 구하는 분할 정복(divide and conquer) 알고리즘으로 변환하는 일잔적인 방법을 제시하며 매우 유익한 기술이다.
+Kellerer외 (2004, \S\,3.3)는 동적 계획 $Z$를 $O(N C)$-시간과 $O(N + C)$-공간 안에 최적해를 구하는 분할 정복(divide and conquer) 알고리즘으로 변환하는 일잔적인 방법을 제시하며 매우 유익하다.
 그러나 이를 이용할 수 있는 조건 중, 목적함수는 어떤 기술적인 의미에서 가산적으로 분할되어야 하며 이를 대학 지원 문제에 적용하기 어렵다.
@@
 Garey, Michael and David Johnson. 1979. \emph{Computers and Intractability: A Guide to the Theory of NP-Completeness.} New York: W. H. Freeman and Company.
+
+Hartlaub, Vanessa and Thorsten Schneider.
+2012.
+“Educational Choice and Risk Aversion: How Important Is Structural vs.
+Individual Risk Aversion?” \emph{SOEPpapers on Multidisciplinary Panel Data Research,} no.
+433.
+\url{https://www.diw.de/documents/publikationen/73/diw_01.c.394455.de/diw_sp0433.pdf}.
+
+Kahneman, Daniel. 2011. \emph{Thinking, Fast and Slow.} New York: Macmillan.
 
@@
 
+Van Huizen, Thomas and Rob Alessie. 2019. ``Risk aversion and job mobility.’’ \emph{Journal of Economic Behavior \& Organization} 164: 91--106. \url{https://doi.org/10.1016/j.jebo.2019.01.021}.
+
 Vazirani, Vijay. 2001. \emph{Approximation Algorithms.} Berlin: Springer.

### 2022-04-01 paper/paper.tex
copy edits, translation improvements

@@
 The second is an dynamic program that produces an exact solution in pseudopolynomial time.
-The third is a fully polynomial-time approximation scheme.
+The third is a different dynamic program that yields fully polynomial-time approximation scheme.
+Numerical experiments demonstrate the algorithms' practicality.
 
@@
 대학의 지원 비용이 서로 다른 경우, 문제가 NP-complete함을 증명한다.
-세 가지 일반적인 해법을 도출한다.
+3가지 일반적인 해법을 도출한다.
 첫째는 분지한계 기반 해법이다.
@@
 마지막으로 다른 동적 계획 기반으로 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme)이 존재함을 보인다.
+수리적 실험을 통해서 알고리즘의 현실성을 보인다.
 
@@
 \ref{hetappcosts}절에서 각 학교의 지원 비용이 동일하지 않은 일반적인 상황을 다룬다.
-이진 배낭 문제에서의 다항 변환을 통해서 포트폴리오 최적화 문제가 NP-complete임을 증명한다.
+이진 배낭 문제에서의 다항 변환을 통해서 포트폴리오 최적화 문제가 NP-complete함을 증명한다.
 또한 3개의 일반적인 해법을 제안한다.
@@
 For the remainder of the paper, unless otherwise noted, we assume with trivial loss of generality that each $g_j \leq H$, each $f_j \in (0, 1]$, and $t_0 < t_1 \leq \cdots \leq t_m$.
-Below, we will show how to transform an arbitrary instance so that $t_0 = 0$, in which case each $t_j > 0$.
+%Below, we will show how to transform an arbitrary instance so that $t_0 = 0$, in which case each $t_j > 0$.
 \else
@@
 
-남은 논문에서 다른 언급이 없으면, 일반성을 잃지 않고 각 $g_j \leq H$, 각 $f_j \in (0, 1]$, 그리고 $t_0 < t_1 \leq \cdots \leq t_m$이라고 가정한다. 뒤에서 임의의 인스턴스를 $t_0 = 0$이 되도록 변환하는 방법을 제시하는데 이를 적용하면 $t_j > 0$인 가정을 세울 수 있다. 
+남은 논문에서 다른 언급이 없으면, 일반성을 잃지 않고 각 $g_j \leq H$, 각 $f_j \in (0, 1]$, 그리고 $t_0 < t_1 \leq \cdots \leq t_m$이라고 가정한다. 뒤에서 임의의 인스턴스를 $t_0 = 0$이 되도록 변환하는 방법을 제시한다. % 이를 적용하면 $t_j > 0$인 가정을 세울 수 있다. 
 \fi
@@
 \ifen
-Next, we show that without loss of generality, we may assume that $t_0 = 0$ (or any constant less than $t_1$).
-\else
-그다음에, 일반성을 제한하지 않고 $t_0 = 0$(또는 $t_1$보다 작은 임의의 상수)이라고 가정할 수 있음을 보이자.
+Next, we show that without loss of generality, we may assume that $t_0 = 0$.
+\else
+그다음에, 일반성을 제한하지 않고 $t_0 = 0$이라고 가정할 수 있음을 보이자.
 \fi
 \begin{theorem} \label{assumetzerozero}
-Let $\bar t_j = t_j - \gamma$ for $j = 0 \dots m$. Then $v(\mathcal{X}; \bar t_j) = v(\mathcal{X};  t_j) -  \gamma$ regardless of $\mathcal{X}$. 
+\ifen 
+For some $\gamma \leq t_0$, let $\bar t_j = t_j - \gamma$ for $j = 0 \dots m$. Then $v(\mathcal{X}; \bar t_j) = v(\mathcal{X};  t_j) -  \gamma$ regardless of $\mathcal{X}$. 
+\else
+어떤 $\gamma \leq t_0$을 선택하고 $j \dots m$에 대해 $\bar t_j = t_j - \gamma$라고 하자. 그러면 $\mathcal{X}$를 막론하고 $v(\mathcal{X}; \bar t_j) = v(\mathcal{X};  t_j) -  \gamma$가 성립한다.
+\fi
 \end{theorem}
@@
 \begin{theorem}[\ifen Nestedness of optimal application portfolios\else 최적 포트폴리오의 포함 사슬 관계\fi] \label{nestedapplication}
-\ifen There exists a sequence of portfolios $\{\mathcal{X}_h\}_{h=1}^m$ satisfying the nestedness relation \fi
+\ifen There exists a sequence of portfolios $\{\mathcal{X}_h\}_{h=1}^m$ satisfying the nestedness relation
+\else  각 $\mathcal{X}_h$가 지원 제한 $h$에 대한 최적 포트폴리오며 위의 포함 사슬 관계를 만족하는 포트톨리오 수열  $\{\mathcal{X}_h\}_{h=1}^m$가 존재한다.\fi
 \begin{equation}
-\mathcal{X}_1 \subset \mathcal{X}_2\subset \dots \subset \mathcal{X}_m.
+\mathcal{X}_1 \subset \mathcal{X}_2\subset \dots \subset \mathcal{X}_m
 \end{equation}
-\ifen such that each $\mathcal{X}_h$ is an optimal application portfolio when the application limit is $h$.
-\else  각 $\mathcal{X}_h$가 지원 제한 $h$에 대한 최적 포트폴리오며 위의 포함 사슬 관계를 만족하는 포트톨리오 수열  $\{\mathcal{X}_h\}_{h=1}^m$가 존재한다.\fi
+\ifen such that each $\mathcal{X}_h$ is an optimal application portfolio when the application limit is $h$.\fi
 \end{theorem}
@@
 %\end{proof}
+
 \ifen 
@@
 \else
-본 절에서 이 문제가 NP-complete임을 보이고 3가지 해법을 제시한다: 정확한 분지한계법 해법, 정확한 동적 계획 해법, 그리고 완전 다항 시간 근사 해법 (fully polynomial-time approximation scheme, FPTAS).
+본 절에서 이 문제가 NP-complete함을 보이고 3가지 해법을 제시한다: 정확한 분지한계법 해법, 정확한 동적 계획 해법, 그리고 완전 다항 시간 근사 해법 (fully polynomial-time approximation scheme, FPTAS).
 \fi
@@
 $f = (0.5, 0.5, 0.5)$, $t = (1, 1, 219)$, 그리고 $g = (1, 1, 3)$이면 $H = 2$에 대응하는 최적해가 $\{1, 2\}$이며 $H = 3$에 대응하는 최적해는 $\{3\}$임을 쉽게 알 수 있다.
-사실은 엘리스의 문제가 NP-complete이며 이를 이진 배낭 문제에서의 변환을 통해 증명할 수 있다.
+사실은 엘리스의 문제가 NP-complete하며 이를 이진 배낭 문제에서의 변환을 통해 증명할 수 있다.
 배낭 문제의 NP-completeness가 이미 알려져 있다 (Garey and Johnson 1979, \S\,3.2.1).
@@
 As with the knapsack problem, Ellis's problem admits a complementary dynamic program that iterates on the value of the cheapest portfolio instead of on the cost of the most valuable portfolio.
-We will use this algorithm as the basis for a fully polynomial-time approximation scheme (FPTAS) for Ellis's problem that uses $O(m^3 / \varepsilon)$-time and -space.
+We will use this algorithm as the basis for an FPTAS for Ellis's problem that uses $O(m^3 / \varepsilon)$-time and -space.
 Here we assume, with a small loss of generality, that each $t_j$ is a natural number.
@@
 \else
-엘리스의 문제는 배낭 문제와 같이, 가치가  가장 높은 포트폴리오의 비용 대신 비용이 가장 낮은 포트폴리오의 가치를 탐색하는 보완적인 동적 계획이 존재한다.
-이를 기반으로  $O(m^3 / \varepsilon)$-시간과 -공간을 사용하는 엘리스 문제의 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme, FPTAS)을 도출할 수 있다.
-여기서 일반성을 약간 제안해서 각 $t_j$가 자연수임을 가정한다.
+엘리스의 문제는 배낭 문제와 같이, 가치가  가장 높은 포트폴리오의 비용 대신 비용이 가장 낮은 포트폴리오의 가치를 탐색하는 보완적인 동적 계획이 존재한다. 이를 기반으로  $O(m^3 / \varepsilon)$-시간과 -공간을 사용하는 엘리스 문제의 FPTAS를 도출할 수 있다. 여기서 일반성을 약간 제안해서 각 $t_j$가 자연수임을 가정한다.
 \fi
@@
 포트폴리오의 근사적 가치를 정확도 $P$로 구성된 고정소수점 십진수(fixed-point decimal)로 나타내자.
-다, $P$는 소수점 뒤에 등장하는 숫자의 수다.
+다, $P$는 소수점 뒤에 등장하는 숫자의 수이다.
 이때 $x$를 가장 가까운 고정소수점 십진수로 내림한 것을 $r[x] =  10^{-P}\lfloor 10^P x \rfloor$라고 하자.
@@
 \else
-본 항 나머지에서, 다른 언급이 없으면 포트폴리오의 ``가치''라고 하면 고정소수점 환경 안에서의 가치를 의미하며 이것이 근사값임을 유념한다. 나중에 동적 계획의 타당성을 증명할 때 근사 오차를 처리한다.
+본 항 나머지에서, 다른 언급이 없으면 포트폴리오의 ``가치''라고 하면 고정소수점 환경 안에서의 가치를 의미하며 이것이 근사값임을 유념한다. 근사 오차는 나중에 동적 계획의 타당성을 증명할 때 처리한다.
 \fi
@@
 단, $\mathcal{W}[j, v]$가 존재하지 않으면 $G[j, v] = \infty$으로 처리한다.
-$v \leq 0$이면 $\mathcal{W}[j, v] = \O$이고 $G[j, h] = 0$이며, $j = 0$이고 $v > 0$이면 $G[j, h] = \infty$인 것이 분명하다.
+$v \leq 0$이면 $\mathcal{W}[j, v] = \O$이고 $G[j, h] = 0$이며, $j = 0$이고 $v > 0$이면 $G[j, h] = \infty$인 것이 자명하다.
 남은 지표(즉 $j, v > 0$)는 다음을 만족한다.
@@
 (최적성.) 알고리즘 \ref{ellisDP3}의 출력이 $\mathcal{W}$이며 실제 최적해가 $\mathcal{X}$라고 하자.
-$v(\mathcal{X}) \leq \bar U$이고, 모든 단일 원소 포트폴리오가 가능해므로  $\mathcal{X}$는 평균적인 단일 원서 포트폴리오보다 가치 높을 수 밖에 없다.
+$v(\mathcal{X}) \leq \bar U$이고, 모든 단일 원소 포트폴리오가 가능해이므로  $\mathcal{X}$는 평균적인 단일 원서 포트폴리오보다 가치 높을 수밖에 없다.
 즉,  $v(\mathcal{X}) \geq \bar U / m$이다.
@@
 \else
-\eqref{recursionrelationforcostmindp}과(와) \eqref{deltajvdef}(으)로 정의된 점화식에 따라 $\Delta_j(v)$를 내림하기 때문에, $j \in \mathcal{W}[j, v]$이면 $(1 - f_j) v\bigl(\mathcal{W}[j-1, v- \Delta_j(v)]\bigr) + f_j t_j$의 실제값은 $\mathcal{W}[j, v]$의 고정소수점값인 $v$보다 클 수 있는데 최대한 $10^{-P}$ 만큼의 차이가 발생하다.
+\eqref{recursionrelationforcostmindp}과(와) \eqref{deltajvdef}(으)로 정의된 점화식에 따라 $\Delta_j(v)$를 내림하기 때문에, $j \in \mathcal{W}[j, v]$이면 $(1 - f_j) v\bigl(\mathcal{W}[j-1, v- \Delta_j(v)]\bigr) + f_j t_j$의 실제값은 $\mathcal{W}[j, v]$의 고정소수점 가치인 $v$보다 클 수 있는데 최대한 $10^{-P}$ 만큼의 차이가 발생하다.
 또한 이 오차는 $\mathcal{W}$에 학교를 추가할 때마다 가산적으로 누적되는데 추가하는 학교의 개수는 최대한 $m$이다.
-따라서 알고리즘의 \ref{vrecordedhere}줄에서 기록하는 $\mathcal{W}$의 값이 $v'(\mathcal{W})$일 때, $v(\mathcal{W}) - v'(\mathcal{W}) \leq m 10^{-P}$이 성립한다.
+따라서 알고리즘의 \ref{vrecordedhere}줄에서 기록하는 $\mathcal{W}$의 고정소수점 가치가 $v'(\mathcal{W})$일 때, $v(\mathcal{W}) - v'(\mathcal{W}) \leq m 10^{-P}$이 성립한다.
 \fi
@@
 \else
-$\mathcal{X}$의 원소를 지표 순서대로 추가했을 때, 각 학교를 추가하면 \eqref{solvemeforvwjvminusj}에 따라 가치를 수정하고 가장 가까운 $10^{-P}$의 곱으로 내림한 값으로 $v'(\mathcal{X})$를 $v'(\mathcal{W})$와 유사하게 정의할 수 있다.
+$\mathcal{X}$의 원소를 지표 순서대로 추가했을 때, 각 학교를 추가하면 \eqref{solvemeforvwjvminusj}에 따라 가치를 수정하고 가장 가까운 $10^{-P}$의 배수로 내림한 값으로 $v'(\mathcal{X})$를 $v'(\mathcal{W})$와 유사하게 정의할 수 있다.
 위와 같은 논리에 따라 $v(\mathcal{X}) - v'(\mathcal{X}) \leq m 10^{-P}$이다.
@@
 \else
-이는 정리에서 말한 근사 계수다.
+이는 정리에서 말한 근사 계수이다.
 \fi
@@
 \else
-는 $O(m^2/ \varepsilon)$이므로 시간 복잡성을 정리와 같다.
+는 $O(m^2/ \varepsilon)$이므로 시간 복잡성은 정리와 같다.
 \fi%To prevent redundant recursive calls to $G(j, v)$, we need to record its values in a dictionary as they are computed; the space required by the dictionary is $O(|\mathcal{V}| m) = O(m^3 / \varepsilon)$.
@@
 
-알고리즘 \ref{ellisDP1}과(와) 알고리즘 \ref{ellisDP3}을(를) 표 대신 재귀 함수 기반으로 구성할 수 있다. 그러나 각 재귀 함수가 스스로를 `두번' 호출하므로 지수적으로 늘어나는 호출 수를 방화하기 위해 각 지표에 대응하는 함수값을 표에 저장하거나 다른 식으로 기록해야 한다.
+알고리즘 \ref{ellisDP1}과(와) 알고리즘 \ref{ellisDP3}을(를) 표 대신 재귀 함수 기반으로 구성할 수 있다. 그러나 각 재귀 함수가 스스로를 두번 호출하므로 지수적으로 늘어나는 호출 수를 방지하기 위해 각 지표에 대응하는 함수값을 표에 저장하거나 다른 식으로 기록해야 한다.
 \fi
@@
 
-\ifen \subsection{Implementation notes} \else \subsection{알고리즘의 실행}\fi
+\ifen \subsection{Implementation notes} \else \subsection{알고리즘의 구현}\fi
 \ifen
@@
 \else
-모든 알고리즘을 쥴리아 (Julia, v1.8.0b1) 언어로 실행했다.
+모든 알고리즘을 쥴리아 (Julia, v1.8.0b1) 언어로 구현했다.
 쥴리아는 $t_j$-값이 정수 또는 부동소수점 수와 같은 특수한 경우에 대해 컴파일러가 자동적으로 최적화하는 모수적 자료형(parametric data type) 기능이 있어서 유용한 선택이었다.
@@
 동적 계획 해법인 \ref{ellisDP1}과(와) 알고리즘 \ref{ellisDP3}을(를) 표 대신 사전 기록 및 재귀 함수로 실행했다.
-알고리즘 \ref{ellisDP3}은(는) \ref{fptashead}항에서 설명한 것과 달리, 포트폴리오의 가치를 십진수 대신 `이진수'로 나타냈으며 $P$와  $\mathcal{V}$의 정의를 적절히 수정했다.
+알고리즘 \ref{ellisDP3}은(는) \ref{fptashead}항에서 설명한 것과 달리, 포트폴리오의 가치를 십진수 대신 이진수로 나타냈으며 $P$와  $\mathcal{V}$의 정의를 적절히 수정했다.
 그리고 고정소수점 대신 $t_j$-값을 $2^P$으로 곱해서 정수로 구성했다.
-이런 변경은 기본 알고리즘 설계와 계산 시간 분석은 변하지 않는데 성능은 많이 개선한다.
+이런 변경은 기본 알고리즘 설계와 계산 시간 분석에 영향주지 않는데 성능이 많이 개선된다.
 \fi
@@
 엘리스의 문제를 고려하는 두번째 실험에서, 모든  $g_j$를 $\{5, \dots, 10\}$에서 균일한 확률로 관측했으며 $H = \lfloor \frac{1}{2} \sum g_j \rfloor$으로 처리했다.
-실제 대학 지원 비용인 50~100달러 사이에 10달러의 어떤 배수이므로 합리적인 비용 분포이다.
+실제 대학 지원 비용인 50과 100달러 사이에 10달러의 어떤 배수이므로 합리적인 비용 분포이다.
 그림 \ref{samplemarket}에서 전형적인 인스턴스가 나타난다.
@@
 \else
-두번째 실험에서 일반적 문제를 고려한다.
+두번째 실험에서 일반적인 문제를 고려한다.
 정확한 해법 (알고리즘 \ref{ellisbnb} 및 알고리즘 \ref{ellisDP1}) 그리고 허용치를 0.5와 0.05로 설정한 근사 해법 (알고리즘 \ref{ellisDP3})의 성능을 비교한다.
@@
 \lastptofcaption~In every case, $h = m/2$.
-\else 알고리즘 \ref{algorithmforlargeh}에서 $\mathcal{C}$를 목록 또는 힙 구현으로 저장했을 때, 
-동일 지원 비용 입학 시장의 최적 지원 포트폴리오를 계산하는 시간. 단위: ms. \lastptofcaption~모든 경우, $h = m/2$. \fi}
+\else (단위: ms) 알고리즘 \ref{algorithmforlargeh}에서 $\mathcal{C}$를 목록 또는 힙 구현으로 저장했을 때, 
+동일 지원 비용 입학 시장의 최적 지원 포트폴리오를 계산하는 시간. \lastptofcaption~모든 경우, $h = m/2$. \fi}
 \end{table}
@@
 \else  %분지한계법 (알고리즘 \ref{ellisbnb}), 정확한 동적 계획 (알고리즘 \ref{ellisDP1}), 그리고 완전 다항 시간 근사 해법 (알고리즘 \ref{ellisDP3})을
- \ref{hetappcosts}절에서 도출한 3개의 알고리즘을 사용할 때, 다양한 지원 비용으로 갖춘 입학 시장의 최적 또는 $(1- \varepsilon)$-최적 포트폴리오를 계산하는 시간. 단위: ms. 분지한계법은 큰 시장에서 비실용적이다. \lastptofcaption \fi}
+(단위: ms)  \ref{hetappcosts}절에서 도출한 3개의 알고리즘을 사용할 때, 다양한 지원 비용으로 갖춘 입학 시장의 최적 또는 $(1- \varepsilon)$-최적 포트폴리오를 계산하는 시간. 분지한계법은 큰 시장에서 비실용적이다. \lastptofcaption \fi}
 \end{table}
@@
 %\pagebreak
-\ifen \section{Conclusion and ideas for future research} \else \section{결론과 향후 연구}\fi\label{conclusion}
+\ifen \section{Conclusion and ideas for future research} \else \section{결론과 향후 연구 방향}\fi\label{conclusion}
 \ifen
 This study has introduced a novel combinatorial optimization problem that we call the college application problem.
-It can be viewed as a kind of integer portfolio optimization problem or nonlinear knapsack problem.
+It can be viewed as a kind of integer asset allocation problem or nonlinear knapsack problem.
 We showed that the special case in which colleges have identical application costs can be solved in polynomial time by a greedy algorithm because the optimal solutions are nested in the budget constraint.
@@
 본 연구는 새로운 조합최적화 문제를 제안했으며 이를 대학 지원 최적화 문제라고 부른다.
-정수 포트폴리오 최적화 문제 또는 확률적인 배낭 문제로 볼 수 있다.
+일종의 정수 자산 배분 문제 또는 확률적인 배낭 문제로 볼 수 있다.
 모든 대학의 지원 비용이 동일한 특수한 경우, 탐욕 알고리즘으로 다항 시간 안에 풀 수 있음을 보였으며 최적해가 예산 제약식에 대한 포함 사슬 관계를 가지기 때문이다.
-일반적인 문제는 NP-complete다.
+일반적인 문제는 NP-complete하다.
 3가지 해법을 제시했으며 이론적으로 가장 효율적인 알고리즘은 $O(m^3 / \varepsilon)$-시간 안에  $(1-\varepsilon)$-근사해를 계산하는 FPTAS이다.
@@
 
-이 문제를 확장할 수 있는 향후 연구 방향 3가지를 제안한다. 
-\fi
-
-\ifen \subsection{Explicit treatment of risk aversion}\else \subsection{위험 회피가 명시적인 모형}\fi
+이 문제를 확장할 수 있는 향후 연구 방향 3가지를 제안할 수 있다. 
+\fi
+
+\ifen \subsection{Explicit treatment of risk aversion}\else \subsection{위험 회피를 모수화한 모형}\fi
 \ifen
@@
 \else
-익숙한 Markowitz 포트폴리오 최적화 모형에는 명시적인 위험 회피 항이 등장하며, 본 모형은 maximax 목적함수에 내포된 경쟁적인 학교와 안정 지원 학교 사이의 균형에 의한 위험 관리 요소만 고려했다.
+익숙한 Markowitz 포트폴리오 최적화 모형에서는 명시적인 위험 회피 항이 등장하며, 본 모형은 maximax 목적함수에 내포된 경쟁적인 학교와 안정 지원 학교 사이의 균형에 의한 위험 관리 요소만 고려했다.
 그러나 기본 목적함수인 $v(\mathcal{X}) = \operatorname{E}[X]$에 평균편차 페널티 $\beta \geq 0$를 다음처럼 도입할 수 있다:
@@
 
+\ifen
 A parametric treatment of risk aversion may align our model more closely to real-world applicant behavior.
 Conventional wisdom asserts that the best college application strategy combines reach, target, and safety schools in roughly equal proportion (Kim 2015).
-When the application budget is small relative to the size of the admissions market, our algorithms recommend a similar approach (see the example of Section \ref{planetsexamplesection}).
+When the application budget is small relative to the size of the admissions market, our algorithms recommend a similar approach (see the example of Subsection \ref{planetsexamplesection}).
 However, inspecting equation \eqref{howtotransformfj}, which discounts school utility parameters to reflect their marginal value relative to the schools already in the portfolio, reveals that low-utility schools are penalized more harshly than high utility schools in the marginal analysis.
@@
 A potential application of our model is an econometric study that estimates students’ perceptions of college quality ($t_j$) from their observed application behavior ($\mathcal{X}$) under the assumption of utility maximization.
-However, empirical research indicates that risk aversion is a significant factor in human decision-making, including in the domains of educational choice and employment search (Kahneman 2011; Hartlaub and Schneider 2012; van Huizen and Alessie 2019).
+However, empirical research indicates that heterogeneous risk aversion is a significant source of variation in human decision-making, especially in the domains of educational choice and employment search (Kahneman 2011; Hartlaub and Schneider 2012; van Huizen and Alessie 2019).
 Therefore, an endogenous risk-aversion parameter would greatly enhance our model’s explanatory power in the econometric setting.
+\else
+위험 회피를 모수화한 모형은 실제 지원 행동을 더 정확하게 반영할 수 있다.
+``상향·소신·안정'' 지원 학교를 대략 균일하게 분산해서 지원하는 것이 최적이란 통념이다 (김민희 2015).
+입학 시장의 크기에 비한 지원 예산이 작은 경우, 본 연구에서 제시하는 알고리즘은 비슷한 전략을 추천한다 (항 \ref{planetsexamplesection}의 예제 참고).
+그러나 포트폴리오에 이미 삽입한 학교에 비한 한계 가치를 반영해서 효용 모수를 할인하는 수식 \eqref{howtotransformfj}을 살펴보면, 효용이 높은 학교보다 효용이 낮은 학교에 부과되는 페널티가 큰 것을 알 수 있다.
+그 결과는 지원 예산이 커질수록 본 모형의 최적 포트폴리오가 안정 지원 학교보다 상향 지원 학교를 편드는 것이다.
+(그림 \ref{samplemarket}에서 뚜렷히 보이는 현상이다.)
+
+학생의 지원 행동($\mathcal{X}$)을 관찰하고 학생이 인식하는 대학교의 질($t_j$)을 추정하는 경제 측정학 연구에서 본 논문의 모형을 그대로 응용할 수 있다.
+그러나 실무 연구 결과에 따르면 위험 회피의 다양성은 인간 의사 결정 다양성의 주된 원인이 되며, 교육 선택과 취업 활동은 특히 그렇다 (Kahneman 2011; Hartlaub와 Schneider 2012; van Huizen과 Alessie 2019).
+따라서 경제 측정학 맥락에서, 위험 회피 모수를 내생 변수로 도입하면 원래 모형의 묘사력을 많이 개선할 수 있다.
+\fi
 
@@
 
-미국 입학 과정의 유사한 특징은 조기 전형(early decision)이다. 이는 학생이 합격하면 진학한다고 약속하고 지원하는 것을 의미한다. 조기 전형으로 지원하면 입학할 관심을 시그널할 수 있으므로 합격 확률을 높일 수 있지만, 학교가 학생을 모집하려고 장학금을 수여할 동기가 약화할 수 있으므로 학생의 효용 추정이 달라진다.
+미국 입학 과정의 유사한 특징은 조기 전형(early decision)이라고 불린다. 이는 학생이 합격하면 진학한다고 약속하고 지원하는 것을 의미한다. 조기 전형으로 지원하면 입학할 관심을 시그널할 수 있으므로 합격 확률을 높일 수 있지만, 학교가 학생을 모집하려고 장학금을 수여할 동기가 약화할 수 있으므로 학생의 효용 추정이 달라진다.
 \fi
@@
 그다음에 학생이 $c_j$에 높은 관심을 시그널하면 (즉, 가군으로 선택하거나 조기 전형으로 지원하면) $x_{j+} = 1$이며 관심을 시그널 하지 않고 지원하면 (즉 가군 혹은 나군으로 지원하거나 조기 전형 없이 지원하면) $x_{j-} = 1$이 되는 이진 결정 변수를 정의한다.
-그러면 해당 합격 확률 $f_{j+}$와 $f_{j-}$ 그리고 효용  $t_{j+}$와 $t_{j-}$가 알려져 있다고 가정하면 다음 같은 논리 제약식을 더하면 모형을 완성할 수 있다.
+해당 합격 확률 $f_{j+}$와 $f_{j-}$ 그리고 효용  $t_{j+}$와 $t_{j-}$가 알려져 있으면, 다음 같은 논리 제약식을 더하면 모형을 완성할 수 있다.
 \begin{align}
@@
 이때 $N$은 결정 변수의 수, $C$는 제약식의 모수, 그리고 $Z[n, c]$는 제약 모수가 $c \leq C$일 때 첫 $n\leq N$개의 결정 변수만 이용하는 최적 목적함수값이다.
-알고리즘의 구조는 $Z[n, c]$를 $Z[n -1, c]$와 어떤 $c' \leq c$에 대응하는 $Z[n -1, c']$의 함수로 표현하는 반복 관계로 탐색하는 것이다.
+알고리즘의 구조는 $Z[n, c]$를 $Z[n -1, c]$와 어떤 $c' \leq c$에 대응하는 $Z[n -1, c']$의 함수로 표현하고 그 반복 관계로 탐색하는 것이다.
 (알고리즘 \ref{ellisDP1}에서 $Z$는 최대 포트폴리오 가치, $N = m$, 그리고 $C =H$이며, 알고리즘 \ref{ellisDP3}에서 $Z$는 최소한 지원 지출액, $N = m$, 그리고 $C= |\mathcal{V}| \propto m^2 / \varepsilon$이다.)
 
-이러한 형태의 동적 계획을 표 혹은 사전 기록으로 구현하면 최적해를 출력하는 것은 $O(NC)$-시간과 -공간을 소요한다.
-Kellerer외 (2004, \S\,3.3)는 동적 계획 $Z$를 $O(N C)$-시간과 $O(N + C)$-공간 안에 최적해를 구하는 분할 정복(divide and conquer) 알고리즘으로 변환하는 일잔적인 방법을 제시하며 매우 유익하다.
+이러한 형태의 동적 계획을 표 혹은 사전 기록으로 구현하면 최적해를 출력하는 것은 $O(NC)$-시간과 $O(NC)$-공간을 소요한다.
+Kellerer외 (2004, \S\,3.3)는 동적 계획 $Z$를 $O(N C)$-시간과 $O(N + C)$-공간 안에 최적해를 구하는 분할 정복(divide and conquer) 알고리즘으로 변환하는 일반적인 방법을 제시하며 매우 유익하다.
 그러나 이를 이용할 수 있는 조건 중, 목적함수는 어떤 기술적인 의미에서 가산적으로 분할되어야 하며 이를 대학 지원 문제에 적용하기 어렵다.
@@
 2018.
-\emph{State of the Profession 2018: The 10 Trends Reshaping Independent Educational Consulting.} Technical report, Independent Eduational Consultants Association.
+\emph{State of the Profession 2018: The 10 Trends Reshaping Independent Educational Consulting.} Technical report, Independent Educational Consultants Association.
 \url{https://www.iecaonline.com/wp-content/uploads/2020/02/IECA-Current-Trends-2018.pdf}.

### 2022-04-04 paper/paper.tex
table summarizing algorithms

@@
 
+\usepackage{stackengine}
+\newcommand\xrowht[2][0]{\addstackgap[.5\dimexpr#2\relax]{\vphantom{#1}}}
+
 % Language selector
 \newif\ifen
-\entrue % Change to false to export Korean
+\enfalse % Change to false to export Korean
 
@@
 마지막으로 다른 동적 계획 기반으로 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme)이 존재함을 보인다.
-수리적 실험을 통해서 알고리즘의 현실성을 보인다.
+수리적 실험을 통하여 알고리즘의 현실성을 보인다.
 
@@
 미국 입학 컨설팅 산업에서 대학 지원 서류 작성, 합격 확률 추정, 그리고 지원 학교를 선택하는 데에 자문하는 개인 상담가의 시간당 급료는 평균 200달러이다 (Sklarow 2018).
-한국에서, 메가스터디(\url{megastudy.net})와 진학(\url{jinhak.com})과 같은 입학 컨설팅 기업의 주된 수입원 중 인공지능을 활용해서 학생의 지원 전략을 최대화한다고 홍보하는 ``모의 지원'' 소프트웨어가 있다.
+한국에서, 메가스터디(\url{megastudy.net})와 진학(\url{jinhak.com})과 같은 입학 컨설팅 기업의 주된 수입원 중 인공지능을 활용하여 학생의 지원 전략을 최대화한다고 홍보하는 ``모의 지원'' 소프트웨어가 있다.
 미래의 소득이 대학 입학 결과로 결정된다는 넓은 인식에 따라, \eqref{headlineproblem}의 최적해는 금전적인 가치를 가지고 있다.
@@
 \else
-\ref{homogappcosts}절에서 각 $g_j = 1$이고 $H$가 $h \leq m$이고 자연수인 특수한 경우를 고려한다. 직관적인 휴리스틱 해법이 실제로 $1/h$-근사 해법임을 증명한다. 그다음에 최적 포트폴리오의 예산 제약에 대한 포함 사슬 관계를 보이고 이를 이용해서 $O(hm)$-시간 정확한 해법을 도출한다.
+\ref{homogappcosts}절에서 각 $g_j = 1$이고 $H$가 $h \leq m$이고 자연수인 특수한 경우를 고려한다. 직관적인 휴리스틱 해법이 실제로 $1/h$-근사 해법임을 증명한다. 그다음에 최적 포트폴리오의 예산 제약에 대한 포함 사슬 관계를 보이고 이를 이용하여 $O(hm)$-시간 정확한 해법을 도출한다.
 \fi
@@
 \ref{hetappcosts}절에서 각 학교의 지원 비용이 동일하지 않은 일반적인 상황을 다룬다.
-이진 배낭 문제에서의 다항 변환을 통해서 포트폴리오 최적화 문제가 NP-complete함을 증명한다.
+이진 배낭 문제에서의 다항 변환을 통하여 포트폴리오 최적화 문제가 NP-complete함을 증명한다.
 또한 3개의 일반적인 해법을 제안한다.
@@
 셋째는 실수값을 분수값으로 내림한 포트폴리오 가치를 탐색하는 또다른 동적 계획 해법이다.
-이를 통해서 $O(m^3 / \varepsilon)$-시간 안에 $(1 - \varepsilon)$-근사해를 출력하는 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme)을 얻는다.
+이를 통하여 $O(m^3 / \varepsilon)$-시간 안에 $(1 - \varepsilon)$-근사해를 출력하는 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme)을 얻는다.
 \fi
@@
 이 특수한 경우는 지원 비용이 없으며 국가 법에 따라 최대 $h$개의 학교에 지원할 수 있는 한국의 중심화된 대학 지원 과정과 비슷하다.
-(한국 경우,  $m=202$이고 $h=3$이다.) 정리 \ref{assumetzerozero}을(를) 적용해서 다른 언급이 없으면 $t_0 = 0$임을 가정하자.
+(한국 경우,  $m=202$이고 $h=3$이다.) 정리 \ref{assumetzerozero}을(를) 적용하여 다른 언급이 없으면 $t_0 = 0$임을 가정하자.
 본 절 내내 지원자를 `알마'라고 부르고 해당 최적화 문제를 `알마의 문제'라고 부른다.
@@
 \ifen 
-The basic error of this algorithm is that it maximizes $\operatorname{E}\left[\sum t_j Z_j \right]$ instead of $\operatorname{E}\left[\max \{t_j Z_j\} \right]$.
+\noindent This algorithm's computation time is $O(m)$ using the \small PICK \normalsize algorithm of Blum et al. (1973).
+
+The basic error of the na\"ive algorithm is that it maximizes $\operatorname{E}\left[\sum t_j Z_j \right]$ instead of $\operatorname{E}\left[\max \{t_j Z_j\} \right]$.
 The latter is what Alma is truly concerned with, since in the end she can attend only one school.
@@
 \else
-위 알고리즘의 기본 실수는 $\operatorname{E}\left[\max \{t_j Z_j\} \right]$ 대신  $\operatorname{E}\left[\sum t_j Z_j \right]$을 최대화하는 것이다. 결국에는 단 한 학교로만 진학할 수 있으므로 알마의 목적은 후자가 아닌 전자다. %학교 개별 효용의 기대 최대값이다. 
+\noindent Blum 외 (1973)의 \small PICK \normalsize 알고리즘을 사용하면 위의 해법의 계산 시간이 $O(m)$이다.
+
+나이브 알고리즘의 기본 실수는 $\operatorname{E}\left[\max \{t_j Z_j\} \right]$ 대신  $\operatorname{E}\left[\sum t_j Z_j \right]$을 최대화하는 것이다. 결국에는 단 한 학교로만 진학할 수 있으므로 알마의 목적은 후자가 아닌 전자다. %학교 개별 효용의 기대 최대값이다. 
 다음 예는 나이브 해법이 최적해가 아닌 해를 출력할 수 있음을 보인다.
@@
 \else (추론.) $\mathcal{X}_1 \subset \cdots \subset \mathcal{X}_h$라고 가정하고 $\mathcal{X}_h \subset \mathcal{X}_{h+1}$임을 보이자.
-$k = \argmax\{ t_k: k \in \mathcal{X}_{h+1}\}$으로 정의해서 $\mathcal{X}_{h+1} = \mathcal{Y}_{h} \cup \{k\}$으로 표현하자.\fi
+$k = \argmax\{ t_k: k \in \mathcal{X}_{h+1}\}$으로 정의하여 $\mathcal{X}_{h+1} = \mathcal{Y}_{h} \cup \{k\}$으로 표현하자.\fi
 
@@
 This means that
-\else 함수 $w(\mathcal{Y})$를 이용해서 $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$인 포트폴리오와 $\{k\}$ 사이의 \mbox{`보완성'을} 평가하자.
+\else 함수 $w(\mathcal{Y})$를 이용하여 $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$인 포트폴리오와 $\{k\}$ 사이의 \mbox{`보완성'을} 평가하자.
 $w(\mathcal{Y})$를 구성하기 위해, 알마가 $c_j$에 합격한 상황에서 $c_k$에 지원했을 때 $c_j$에서 받는 조건부 기대 효용을 $\tilde t_j$라고 하자.
@@
 We reduce the computation time to $O(hm)$ by taking advantage of the transformation from the inductive step in the proof of Theorem \ref{nestedapplication}.
-Once school $k$ is added to $\mathcal{X}$, we remove it from the set $\mathcal{C}\setminus \mathcal{X}$ of candidates, and update the $t_j$-values of the remaining schools according to the following transformation:
-\else
-정리 \ref{nestedapplication} 증명의 추론 단계에서 제시한 변환을 활용하면 계산 시간을 $O(hm)$으로 감소시킬 수 있다. $\mathcal{X}$에 $k$를 추가하면, 후보자 집합인 $\mathcal{C}\setminus \mathcal{X}$에서 $k$를 제거하고 남은 학교의 $t_j$-값은 다음처럼 수정한다.
-\fi
+Once school $k$ is added to $\mathcal{X}$, we eliminate it from the set $\mathcal{C}\setminus \mathcal{X}$ of candidates, and update the $t_j$-values of the remaining schools according to the following relation.
+\else
+정리 \ref{nestedapplication} 증명의 추론 단계에서 제시한 변환을 활용하면 계산 시간을 $O(hm)$으로 감소시킬 수 있다. $\mathcal{X}$에 $k$를 추가하면, 후보자 집합인 $\mathcal{C}\setminus \mathcal{X}$에서 $k$를 소거하고 남은 학교의 $t_j$-값을 다음처럼 수정한다.
+\fi
+\begin{proposition}[\ifen Eliminate $c_k$\else $c_k$ 소거\fi]
+\ifen For $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$, where
+\else $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$에 대해, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$이다. 단,\fi
 \begin{align}\label{howtotransformfj}
@@
 (1 - f_k) t_j, \quad & t_j \leq t_k \\
-t_j - f_k t_k, \quad& t_j > t_k
+t_j - f_k t_k, \quad& t_j > t_k.
 \end{cases}
 \end{align}
-\ifen
-It is easy to verify that this is the composition of the two transformations (from $t$ to $\tilde t$, and from $\tilde t$ to $\bar t$) given in the proof.
+\end{proposition}%
+\ifen
+\noindent It is easy to verify that \eqref{howtotransformfj} is the composition of the two transformations (from $t$ to $\tilde t$, and from $\tilde t$ to $\bar t$) given in the proof.
+(We will reuse this transformation as a variable-elimination technique in the branch-and-bound algorithm presented in Subsection \ref{subsectionbranchbound}.)
+
 Now, the \emph{next} school added must be the optimal singleton portfolio in the modified market.
@@
 \else
-이것이 증명에서 정의한 두개의 변환($t$에서 $\tilde t$로, 그리고 $\tilde t$에서 $\bar t$로)의 합성임을 쉽게 확인할 수 있다.
+\noindent \eqref{howtotransformfj}이(가) 증명에서 정의한 두개의 변환($t$에서 $\tilde t$로, 그리고 $\tilde t$에서 $\bar t$로)의 합성임을 쉽게 확인할 수 있다. (항 \ref{subsectionbranchbound}에서 제시하는 분지한계법에서 이 변환을 변수 소거법으로 재활용한다.)
+
 그러면 다음으로 추가하는 학교는 수정된 시장의 최적 단일 원소 포트폴리오가 되어야 한다.
 그런데 최적 단일 원소 포트폴리오는 단순히 $f_j \bar t_j$가 가장 높은 학교로 이루어진다.
-따라서 각 반복단계에서 \eqref{howtotransformfj}을(를) 이용해서 $t_j$-값을 수정하면 $v(\mathcal{X})$를 계산할 필요가 완전히 없어진다.
+따라서 각 반복단계에서 \eqref{howtotransformfj}을(를) 이용하여 $t_j$-값을 수정하면 $v(\mathcal{X})$를 계산할 필요가 완전히 없어진다.
 게다가 이 알고리즘에서 $t_j$를 배열할 필요가 없으므로 $O(m\log m)$인 배열 비용이 사라진다.
@@
 \ifen
-The algorithm below outputs a list $\mathtt{X}$ of the $h$ schools to which Alma should apply.
+Algorithm \ref{algorithmforlargeh} outputs a list $\mathtt{X}$ of the $h$ schools to which Alma should apply.
 The schools appear in the order of entry such that when the algorithm is run with $h=m$, the optimal portfolio of size $h$ is given by $\mathcal{X}_h = \{\mathtt{X[1]}, \dots, \mathtt{X[h]}\}$.
@@
 \else
-아래 알고리즘은 알마가 지원하는 $h$개의 학교를 목록 $\mathtt{X}$로 출력한다.
+알고리즘 \ref{algorithmforlargeh}은(는) 알마가 지원하는 $h$개의 학교를 목록 $\mathtt{X}$로 출력한다.
 그의 원소들은 진입하는 순서대로 등장한다.
@@
 \else
-이제 $g_j$가 $c_j$에 지원하는 비용을 나타내며 학생이 지원에 쓸 수 있는 예산이 $H$인 일반적인 문제를 고려한다. 이 문제를 풀고자 하는 학생을 엘리스라고 부르자. 본 절 내내 정리 \ref{assumetzerozero}을(를) 적용해서 $t_0 = 0$임을 가정하자.
+이제 $g_j$가 $c_j$에 지원하는 비용을 나타내며 학생이 지원에 쓸 수 있는 예산이 $H$인 일반적인 문제를 고려한다. 이 문제를 풀고자 하는 학생을 엘리스라고 부르자. 본 절 내내 정리 \ref{assumetzerozero}을(를) 적용하여 $t_0 = 0$임을 가정하자.
 \fi
@@
 즉, $\sum_{j \in \mathcal{X}} u_j \geq U$가 성립할 필수충분 조건은 $v(\mathcal{X}) \geq U -1$이다.
-$V = U-1$으로 정의해서 변환을 완성하면 증명이 완료된다.
+$V = U-1$으로 정의하여 변환을 완성하면 증명이 완료된다.
 \fi
@@
 
-\ifen \subsection{Branch-and-bound algorithm} \else \subsection{분지한계법} \fi
+\ifen \subsection{Branch-and-bound algorithm} \else \subsection{분지한계법} \fi \label{subsectionbranchbound}
 \ifen 
@@
 정수 최적화의 고전적인 기법 중 분지한계법(branch and bound)이 있다.
-이는 하나 이상의 결정 변수를 고정시키고 해당 부문제를 탐색하는 해법이며, 목적함수의 상한(bound)을 활용해서 기존 해보다 좋은 해를 생성할 수 없는 결정 나무의 가지(branch)를 제외시켜 결정 공간을 줄인다 (Martello and Toth 1990; Kellerer 외 2004).
+이는 하나 이상의 결정 변수를 고정시키고 해당 부문제를 탐색하는 해법이며, 목적함수의 상한(bound)을 활용하여 기존 해보다 좋은 해를 생성할 수 없는 결정 나무의 가지(branch)를 제외시켜 결정 공간을 줄인다 (Martello and Toth 1990; Kellerer 외 2004).
 본 항에서, 엘리스 문제의 정수 모형과 그의 상한이 되는 선형 계획을 제시한다.
-알고리즘 \ref{algorithmforlargeh}에서 개발한 $t_j$-값의 변환을 활용해서 부문제의 선형 완화 문제에 의한 상한을 더 타이트하게 만든다.
+알고리즘 \ref{algorithmforlargeh}에서 개발한 $t_j$-값의 변환을 활용하여 부문제의 선형 완화 문제에 의한 상한을 더 타이트하게 만든다.
 이 기반으로 분지한계법 알고리즘을 도출한다.
@@
 $H > 0$임이 성립하는 동안, $f_j t_j / g_j$-값이 가장 높은 학교 $j$를 택한다.
-$x_j \gets \min\{1, H / g_j\}$ 그리고 $H \gets H - g_j$로 수정해서 반복한다 (Dantzig 1957).
+$x_j \gets \min\{1, H / g_j\}$ 그리고 $H \gets H - g_j$로 수정하여 반복한다 (Dantzig 1957).
 $O(m)$ 알고리즘도 알려져 있다 (Balas와 Zemel 1980, \S\,2).
@@
 각 부문제에서 $\bar H = H - \sum_{j\in \mathcal{I}} g_j$는 잔여 예산을 의미한다.
-$\mathcal{I}$에 속한 학교에 \eqref{howtotransformfj}의 변환을 적용해서 모수 $\gamma$와 $\bar t$를 얻을 수 있다.
+$\mathcal{I}$에 속한 학교에 \eqref{howtotransformfj}의 변환을 적용하여 모수 $\gamma$와 $\bar t$를 얻을 수 있다.
 즉, 각 $j\in \mathcal{I}$에 대해 $\gamma$를 $f_j \bar t_j$ 만큼 증가시키고, 시장에서 $j$를 소거하고, \eqref{howtotransformfj}에 따라 남은 $\bar t_j$-값을 수정한다.
@@
 부문제를 계산하기 위해 먼저 $c_1$을 단순히 무시한다.
-$c_2$를 소거하기 위해 $t$에 \eqref{howtotransformfj}에 적용해서
+$c_2$를 소거하기 위해 $t$에 \eqref{howtotransformfj}에 적용하여
 $\{\bar{\bar t}_3, \bar{\bar t}_4, \bar{\bar t}_5\} = \{ (1 - f_2) t_3, (1 - f_2) t_4, (1 -f_2) t_5\} = \{30, 40, 50 \}$과 $\bar{\bar \gamma} = f_2 t_2 = 20$을 얻는다.
-$c_5$를 소거하기 위해 $\bar{\bar t}$에 \eqref{howtotransformfj}에 다시 적용해서 $\{\bar t_3, \bar t_4\} =  \{\bar{\bar t}_3 - f_5 \bar{\bar t}_5, \bar{\bar t}_4-  f_5 \bar{\bar t}_5\}= \{5, 15\}$와 $\gamma = \bar{\bar \gamma} + f_5 \bar{\bar t}_5 = 35$를 얻는다.
+$c_5$를 소거하기 위해 $\bar{\bar t}$에 \eqref{howtotransformfj}에 다시 적용하여 $\{\bar t_3, \bar t_4\} =  \{\bar{\bar t}_3 - f_5 \bar{\bar t}_5, \bar{\bar t}_4-  f_5 \bar{\bar t}_5\}= \{5, 15\}$와 $\gamma = \bar{\bar \gamma} + f_5 \bar{\bar t}_5 = 35$를 얻는다.
 마지막으로, $\bar H = H - g_2 - g_5 = 3$이다.
@@
 \else
-본 절에서 일반성을 약간 제한해서 $g_j \in \mathbb{N}$ for $j = 1\dots m$과 $H \in\mathbb{N}$임을 가정한다.
+본 절에서 일반성을 약간 제한하여 $g_j \in \mathbb{N}$ for $j = 1\dots m$과 $H \in\mathbb{N}$임을 가정한다.
 이때 엘리스의 문제를 위한 $O(Hm + m\log m)$-시간 해법을 제시한다.
@@
 \else
-엘리스의 문제는 배낭 문제와 같이, 가치가  가장 높은 포트폴리오의 비용 대신 비용이 가장 낮은 포트폴리오의 가치를 탐색하는 보완적인 동적 계획이 존재한다. 이를 기반으로  $O(m^3 / \varepsilon)$-시간과 -공간을 사용하는 엘리스 문제의 FPTAS를 도출할 수 있다. 여기서 일반성을 약간 제안해서 각 $t_j$가 자연수임을 가정한다.
+엘리스의 문제는 배낭 문제와 같이, 가치가  가장 높은 포트폴리오의 비용 대신 비용이 가장 낮은 포트폴리오의 가치를 탐색하는 보완적인 동적 계획이 존재한다. 이를 기반으로  $O(m^3 / \varepsilon)$-시간과 -공간을 사용하는 엘리스 문제의 FPTAS를 도출할 수 있다. 여기서 일반성을 약간 제한하여 각 $t_j$가 자연수임을 가정한다.
 \fi
@@
 Algorithms \ref{ellisDP1} and \ref{ellisDP3} can be written using recursive functions instead of lookup tables.
-However, since each function references itself \emph{twice,} the function values at each index must be recorded in a lookup table or otherwise memoized to prevent an exponential number of calls from forming on the stack.
+However, since each function references itself \emph{twice,} the function values at each index must be recorded in a lookup table or memoized to prevent an exponential number of calls from forming on the stack.
 \else
@@
 알고리즘 \ref{ellisDP3}은(는) \ref{fptashead}항에서 설명한 것과 달리, 포트폴리오의 가치를 십진수 대신 이진수로 나타냈으며 $P$와  $\mathcal{V}$의 정의를 적절히 수정했다.
-그리고 고정소수점 대신 $t_j$-값을 $2^P$으로 곱해서 정수로 구성했다.
+그리고 고정소수점 대신 $t_j$-값을 $2^P$으로 곱하여 정수로 구성했다.
 이런 변경은 기본 알고리즘 설계와 계산 시간 분석에 영향주지 않는데 성능이 많이 개선된다.
@@
 The table shows the average time (standard deviation) over the \nmarkets~instances.
-\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복해서 그 중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (평균편차) 시간이 나타난다.\fi}
+\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복하여 그 중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (평균편차) 시간이 나타난다.\fi}
 
@@
 \else
-\textbf{\begin{tabular}[r]{@{}r@{}}학교의\\개수 $m$\end{tabular}}&\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘  \ref{ellisbnb}:\\분지한계법\end{tabular}}}  & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP1}: 지원\\비용 동적 계획\end{tabular}}}  &\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.5$\end{tabular}}}  & \multicolumn{2}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.05$\end{tabular}}}   \\ \hline
+\textbf{\begin{tabular}[r]{@{}r@{}}학교의\\개수 $m$\end{tabular}}&\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘  \ref{ellisbnb}:\\분지한계법\end{tabular}}}  & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP1}:\\지출액 동적 계획\end{tabular}}}  &\multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.5$\end{tabular}}}  & \multicolumn{2}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}알고리즘 \ref{ellisDP3}:\\FPTAS, $\varepsilon= 0.05$\end{tabular}}}   \\ \hline
 \fi
@@
 On the other hand, for typical college-application instances, the dynamic program based on application expenditures is both easier to implement and substantially more time efficient.
-
-Three extensions of this problem appear amenable to future study.
+The algorithms discussed in this paper are summarized in Table \ref{algorithmssummary}.
 \else
@@
 3가지 해법을 제시했으며 이론적으로 가장 효율적인 알고리즘은 $O(m^3 / \varepsilon)$-시간 안에  $(1-\varepsilon)$-근사해를 계산하는 FPTAS이다.
-그 반면에 전형적인 대학 지원 문제 인스턴스에 대해서, 지원 비용 지출액 기반 동적 계획 해법은 실행하기 쉬울 뿐더러 계산 시간이 상당히 짧다.
-
+그 반면에 전형적인 대학 지원 문제 인스턴스에 대해, 지원 비용 지출액 기반 동적 계획 해법은 실행하기 쉬울 뿐더러 계산 시간이 상당히 짧다.
+표 \ref{algorithmssummary}에서 본 논문이 다운 알고리즘을 정리한다.
+\fi
+
+\ifen
+\begin{table}[h!] \centering
+\small
+\begin{tabular}{r|lllll}
+\textbf{Algorithm} & \textbf{Reference}  & \textbf{Problem} & \textbf{Restrictions} & \textbf{Exactness}       & \textbf{Computation time} \\ \hline
+\xrowht[()]{1.5em}  \begin{tabular}[r]{@{}r@{}}Na\"ive\end{tabular} & Definition \ref{naivealgorithm}                   & \begin{tabular}[l]{@{}l@{}}Homogeneous \\ costs\end{tabular}     & None                  & $(1/h)$-opt.
+& $O(m)$                    \\
+\xrowht[()]{1.5em}  Greedy                & Algorithm \ref{algorithmforlargeh} &  \begin{tabular}[l]{@{}l@{}}Homogeneous \\ costs\end{tabular}    & None                  & Exact                    & $O(hm)$                   \\
+\xrowht[()]{1.5em}  \begin{tabular}[r]{@{}r@{}}Branch and\\ bound\end{tabular}   & Algorithm \ref{ellisbnb} & General          & None                  & Exact                    & $O(2^m)$                  \\
+\xrowht[()]{1.5em}  Costs DP     &  Algorithm \ref{ellisDP1}& General          & $g_j$ integer         & Exact                    & $O(Hm + m \log m)$        \\
+\xrowht[()]{1.5em}  FPTAS      &    Algorithm \ref{ellisDP3} & General          & $t_j$ integer         & $(1 - \varepsilon)$-opt. & $O(m^3 / \varepsilon)$   
+\end{tabular}%
+\caption{\label{algorithmssummary} \normalsize
+Summary of algorithms discussed in this paper.}
+\end{table}
+\else
+\begin{table}[h!] \centering
+\small
+\begin{tabular}{r|lllll}
+\textbf{알고리즘} & \textbf{논문 내 위치}  & \textbf{문제} & \textbf{제한} & \textbf{정확도}       & \textbf{계산 시간} \\ \hline
+\xrowht[()]{1.7em}  \begin{tabular}[r]{@{}r@{}}나이브\end{tabular} & 정의 \ref{naivealgorithm}                   & \begin{tabular}[l]{@{}l@{}}동일한\\ 지원 비용\end{tabular}     & 없음                  & $(1/h)$-근사          & $O(m)$                    \\
+\xrowht[()]{1.7em}  탐욕 해법                & 알고리즘 \ref{algorithmforlargeh} &  \begin{tabular}[l]{@{}l@{}}동일한\\ 지원 비용\end{tabular}    & 없음                  & 정확                    & $O(hm)$                   \\
+\xrowht[()]{1.7em}  \begin{tabular}[r]{@{}r@{}}분지한계법\end{tabular}   & 알고리즘 \ref{ellisbnb} & 일반 문제          & 없음                  & 정확                    & $O(2^m)$                  \\
+\xrowht[()]{1.7em}   \begin{tabular}[r]{@{}r@{}}지출액\\ 동적 계획\end{tabular}     &  알고리즘 \ref{ellisDP1}& 일반 문제          & $g_j$ 정수         & 정확                    & $O(Hm + m \log m)$        \\
+\xrowht[()]{1.7em}  FPTAS      &    알고리즘 \ref{ellisDP3} & 일반 문제          & $t_j$ 정수         & $(1 - \varepsilon)$-근사 & $O(m^3 / \varepsilon)$   
+\end{tabular}%
+\caption{\label{algorithmssummary} \normalsize
+본 논문에서 다룬 해법의 요약.}
+\end{table}
+\fi
+
+\ifen 
+Three extensions of this problem appear amenable to future study.
+\else
 이 문제를 확장할 수 있는 향후 연구 방향 3가지를 제안할 수 있다. 
@@
 위험 회피를 모수화한 모형은 실제 지원 행동을 더 정확하게 반영할 수 있다.
-``상향·소신·안정'' 지원 학교를 대략 균일하게 분산해서 지원하는 것이 최적이란 통념이다 (김민희 2015).
+``상향·소신·안정'' 지원 학교를 대략 균일하게 분산하여 지원하는 것이 최적이란 통념이다 (김민희 2015).
 입학 시장의 크기에 비한 지원 예산이 작은 경우, 본 연구에서 제시하는 알고리즘은 비슷한 전략을 추천한다 (항 \ref{planetsexamplesection}의 예제 참고).
-그러나 포트폴리오에 이미 삽입한 학교에 비한 한계 가치를 반영해서 효용 모수를 할인하는 수식 \eqref{howtotransformfj}을 살펴보면, 효용이 높은 학교보다 효용이 낮은 학교에 부과되는 페널티가 큰 것을 알 수 있다.
+그러나 포트폴리오에 이미 삽입한 학교에 비한 한계 가치를 반영하여 효용 모수를 할인하는 수식 \eqref{howtotransformfj}을 살펴보면, 효용이 높은 학교보다 효용이 낮은 학교에 부과되는 페널티가 큰 것을 알 수 있다.
 그 결과는 지원 예산이 커질수록 본 모형의 최적 포트폴리오가 안정 지원 학교보다 상향 지원 학교를 편드는 것이다.
@@
 이러한 형태의 동적 계획을 표 혹은 사전 기록으로 구현하면 최적해를 출력하는 것은 $O(NC)$-시간과 $O(NC)$-공간을 소요한다.
-Kellerer외 (2004, \S\,3.3)는 동적 계획 $Z$를 $O(N C)$-시간과 $O(N + C)$-공간 안에 최적해를 구하는 분할 정복(divide and conquer) 알고리즘으로 변환하는 일반적인 방법을 제시하며 매우 유익하다.
+Kellerer 외 (2004, \S\,3.3)는 동적 계획 $Z$를 $O(N C)$-시간과 $O(N + C)$-공간 안에 최적해를 구하는 분할 정복(divide and conquer) 알고리즘으로 변환하는 일반적인 방법을 제시하며 매우 유익하다.
 그러나 이를 이용할 수 있는 조건 중, 목적함수는 어떤 기술적인 의미에서 가산적으로 분할되어야 하며 이를 대학 지원 문제에 적용하기 어렵다.
@@
 Bezanson, Jeff, Alan Edelman, Stefan Karpinski, and Viral B. Shah. 2017. ``Julia: A Fresh Approach to Numerical Computing.'' \emph{SIAM Review} 59: 65–98. \url{https://doi.org/10.1137/141000671}.
+
+ Blum,  Manuel, Robert W.
+Floyd, Vaughan Pratt, Ronald L.
+Rivest, and Robert E.
+Tarjan.
+1973.
+``Time bounds for selection.'' \emph{Journal of Computer and System Sciences} 7 (4): 448--61.
+\url{https://doi.org/10.1016/S0022-0000(73)80033-9}.
 

### 2022-04-05 paper/paper.tex
allow float values in SameCostsMarket

@@
 \begin{abstract}
-본 논문은 다수의 확률 변수로 구성된 포트폴리오의 기대 최대값을 예산 조건 하에서 최대화하는 문제를 고려한다.
+본 논문은 다수의 확률 변수로 구성된 포트폴리오의 기대 최댓값을 예산 조건 하에서 최대화하는 문제를 고려한다.
 이를 대학 지원 최적화 문제라고 부른다.
@@
 $f_j$가 (학교의 전체적인 합격률이 아니라) 바로 이 학생의 합격 확률일 때 $Z_j$의 확률적 독립성은 적절한 가정이다.
-그러면 학생의 목표는 주어진 예산 안에서 학생이 합격하는 학교들에서 얻는 효용 중 기대 최대값을 최대화하는 것이다.
+그러면 학생의 목표는 주어진 예산 안에서 학생이 합격하는 학교들에서 얻는 효용 중 기대 최댓값을 최대화하는 것이다.
 위 문제의 최적해가 $\mathcal{X}$일 때, 학생의 최적 대학 지원 전략은 $\mathcal{X}$에  속한 학교로 지원하는 것이다.
@@
 Risk management is implicit in the college application problem because, in a typical admissions market, college preferability is negatively correlated with competitiveness.
-That is, students negotiate a tradeoff between attractive, selective “reach schools” and less preferable “safety schools” where admission is a safer bet (Kim 2015).
+That is, students negotiate a tradeoff between attractive, selective “reach schools” and less preferable “safety schools” where admission is a safer bet (Jeon 2015).
 Finally, the combinatorial nature of the college application problem makes it difficult to solve using the gradient-based techniques used in continuous portfolio optimization.
@@
 또한 입학 시장에서는 전형적으로 대학의 효용과 합격 확률이 서로 반비례하므로 대학 지원 문제는 위험 관리를 포함하게 된다.
-특히 선호도가 높으며 붙기 어려운 “상향” 지원 학교(reach school)와 선호도가 낮으며 붙기 쉬운 “안정” 지원 학교(safety school) 사이의 균형을 고려해야 한다 (김민희 2015).
+특히 선호도가 높으며 붙기 어려운 “상향” 지원 학교(reach school)와 선호도가 낮으며 붙기 쉬운 “안정” 지원 학교(safety school) 사이의 균형을 고려해야 한다 (전민희 2015).
 마지막, 대학 지원의 조합적인 본성으로 인해 연속적인 포트폴리오 최적화 문제에서 흔히 사용하는 기울기 해법으로는 풀기가 어렵다.
@@
 \else
-주어진 지원 포트폴리오에 대해 학생이 $c_j$로 진학할 확률을 $p_j(\mathcal{X})$라고 하자. 이 사건이 발생하는 필수충분 조건은 $c_j$에 지원하고, $c_j$에서 합격받고, $c_j$보다 선호하는 (즉, 지표가 더 높은) 학교에서 불합격받는 것이다. 따라서  $j= 0\dots m$에 대해
+주어진 지원 포트폴리오에 대해 학생이 $c_j$로 진학할 확률을 $p_j(\mathcal{X})$라고 하자. 이 사건이 발생하는 필수충분 조건은 $c_j$에 지원하고, $c_j$에서 합격하고, $c_j$보다 선호하는 (즉, 지표가 더 높은) 학교에서 불합격는 것이다. 따라서  $j= 0\dots m$에 대해
 \fi
@@
 \else
-기대 효용 $f_j t_j$가 가장 높은 $h$개의 학교로 지원한다.
+기대 효용 $f_j t_j$가 가장 큰 $h$개의 학교로 지원한다.
 \fi
@@
 
-나이브 알고리즘의 기본 실수는 $\operatorname{E}\left[\max \{t_j Z_j\} \right]$ 대신  $\operatorname{E}\left[\sum t_j Z_j \right]$을 최대화하는 것이다. 결국에는 단 한 학교로만 진학할 수 있으므로 알마의 목적은 후자가 아닌 전자다. %학교 개별 효용의 기대 최대값이다. 
+나이브 알고리즘의 기본 실수는 $\operatorname{E}\left[\max \{t_j Z_j\} \right]$ 대신  $\operatorname{E}\left[\sum t_j Z_j \right]$을 최대화하는 것이다. 결국에는 단 한 학교로만 진학할 수 있으므로 알마의 목적은 후자가 아닌 전자다. %학교 개별 효용의 기대 최댓값이다. 
 다음 예는 나이브 해법이 최적해가 아닌 해를 출력할 수 있음을 보인다.
@@
 $v(\mathcal{T}) = 70(0.4)(1-0.4) + 80(0.4) = 48.8$.
-But $\mathcal{X} = \{2, 3\}$ with
+But $\mathcal{X} = \{2, 3\}$ with￣
 $v(\mathcal{X}) = 80(0.4)(1-0.3) + 90(0.3) = 49.4$
@@
 그러면 나이브 해법은 $\mathcal{T} = \{1, 2\}$를 선택하며 그의 기대 효용은 $v(\mathcal{T}) = 70(0.4)(1-0.4) + 80(0.4) = 48.8$이다.
-그러나 $\mathcal{X} = \{2, 3\}$를 선택하면 기대 효용은 $v(\mathcal{X}) = 80(0.4)(1-0.3) + 90(0.3) = 49.4$이고 이는 최적해다.
+그러나 $\mathcal{X} = \{2, 3\}$를 선택하면 기대 효용은 $v(\mathcal{X}) = 80(0.4)(1-0.3) + 90(0.3) = 49.4$이고 이는 최적해이다.
 \fi
@@
 \else
-지원 제한이 $h$일 때, 최적 포트폴리오가 $\mathcal{X}_h$이고 기대 효용 $f_j t_j$가 가장 높은 $h$개의 학교의 집합이 $\mathcal{T}_h$라고 하자. 그러면 $v(\mathcal{T}_h) / v(\mathcal{X}_h) \geq 1/h$이다.
+지원 제한이 $h$일 때, 최적 포트폴리오가 $\mathcal{X}_h$이고 기대 효용 $f_j t_j$가 가장 큰 $h$개의 학교의 집합이 $\mathcal{T}_h$라고 하자. 그러면 $v(\mathcal{T}_h) / v(\mathcal{X}_h) \geq 1/h$이다.
 \fi
@@
 \ifen There exists a sequence of portfolios $\{\mathcal{X}_h\}_{h=1}^m$ satisfying the nestedness relation
-\else  각 $\mathcal{X}_h$가 지원 제한 $h$에 대한 최적 포트폴리오며 위의 포함 사슬 관계를 만족하는 포트톨리오 수열  $\{\mathcal{X}_h\}_{h=1}^m$가 존재한다.\fi
+\else  각 $\mathcal{X}_h$가 지원 제한 $h$에 대한 최적 포트폴리오며 위의 포함 사슬 관계를 만족하는 포트폴리오 수열  $\{\mathcal{X}_h\}_{h=1}^m$가 존재한다.\fi
 \begin{equation}
@@
 Optimality requires that
-\else$h$에 대해 귀납법을 적용한다. 정리 \ref{assumetzerozero}에 따라 $t_0 = 0$이라고 가정한다.
+\else $h$에 대해 귀납법을 적용한다. 정리 \ref{assumetzerozero}에 따라 $t_0 = 0$이라고 가정하자.
 
@@
 Since $\bar t_0 = 0$ in the latter market, the inductive hypothesis implies that $\mathcal{Y}_{h-1} \subset \mathcal{Y}_h$, which completes the proof.\footnote{We thank Yim Seho for discovering this useful transformation.}
-\else 두번째 등호는 정리 \ref{assumetzerozero}에 따라 성립한다.
+\else 두 번째 등호는 정리 \ref{assumetzerozero}에 따라 성립한다.
 이 등호는, $\{k\}$와 최적으로 보완적인 $\mathcal{Y}_{h-1}$과 $\mathcal{Y}_h$는 목적함수가 $w(\mathcal{Y})$인 시장에서 각각 크기가 $h-1$과 $h$인 최적 포트폴리오임을 의미한다.
@@
 $t$를 배열하는 시간을 $O(m \log m)$이다.
-모든 $h$개의 반복단계에서 $k$의 후보자의 개수는 $O(m)$이며 \eqref{closedformportfoliovaluationX}을(를) 사용하면 $v(\mathcal{X}\cup \{k\})$를 계산하는 시간은 $O(h)$이다.
+모든 $h$개의 반복 단계에서 $k$의 후보자의 개수는 $O(m)$이며 \eqref{closedformportfoliovaluationX}을(를) 사용하면 $v(\mathcal{X}\cup \{k\})$를 계산하는 시간은 $O(h)$이다.
 따라서 이 해법의 계산 시간이 $O(h^2 m + m \log m)$임을 알 수 있다.\fi
@@
 \else
-\noindent \eqref{howtotransformfj}이(가) 증명에서 정의한 두개의 변환($t$에서 $\tilde t$로, 그리고 $\tilde t$에서 $\bar t$로)의 합성임을 쉽게 확인할 수 있다. (항 \ref{subsectionbranchbound}에서 제시하는 분지한계법에서 이 변환을 변수 소거법으로 재활용한다.)
+\noindent \eqref{howtotransformfj}이(가) 증명에서 정의한 2개의 변환($t$에서 $\tilde t$로, 그리고 $\tilde t$에서 $\bar t$로)의 합성임을 쉽게 확인할 수 있다. (항 \ref{subsectionbranchbound}에서 제시하는 분지한계법에서 이 변환을 변수 소거법으로 재활용한다.)
 
 그러면 다음으로 추가하는 학교는 수정된 시장의 최적 단일 원소 포트폴리오가 되어야 한다.
-그런데 최적 단일 원소 포트폴리오는 단순히 $f_j \bar t_j$가 가장 높은 학교로 이루어진다.
-따라서 각 반복단계에서 \eqref{howtotransformfj}을(를) 이용하여 $t_j$-값을 수정하면 $v(\mathcal{X})$를 계산할 필요가 완전히 없어진다.
+그런데 최적 단일 원소 포트폴리오는 단순히 $f_j \bar t_j$가 가장 큰 학교로 이루어진다.
+따라서 각 반복 단계에서 \eqref{howtotransformfj}을(를) 이용하여 $t_j$-값을 수정하면 $v(\mathcal{X})$를 계산할 필요가 완전히 없어진다.
 게다가 이 알고리즘에서 $t_j$를 배열할 필요가 없으므로 $O(m\log m)$인 배열 비용이 사라진다.
@@
 \else
-최적성은 정리 \ref{nestedapplication}의 증명에 따라 성립한다. $\mathcal{C}$를 목록 구현으로 저장한다고 하자. 그러면 각 $h$개의 반복단계에서 최적 추가 학교를 구하는 시간은 $O(m)$이며 나머지 $O(m)$개 학교의 $t_j$-값을 각 단위 시간으로 수정할 수 있다. 따라서 전체 시간 복잡도가 $O(h m)$이다.
+최적성은 정리 \ref{nestedapplication}의 증명에 따라 성립한다. $\mathcal{C}$를 목록 구현으로 저장한다고 하자. 그러면 각 $h$개의 반복 단계에서 최적 추가 학교를 구하는 시간은 $O(m)$이며 나머지 $O(m)$개 학교의 $t_j$-값을 각 단위 시간으로 수정할 수 있다. 따라서 전체 시간 복잡도가 $O(h m)$이다.
 \fi
@@
 $\mathcal{C}$는 목록 대신 이진 최대 힙 구현으로 저장하는 방법도 고려한다.
-단, 힙의 순서는 $i \geq j \iff f_i t_i \geq f_j t_j$의 조건으로 정의하며, 각 반복단계의 마지막 단계에서 힙을 비우고 다시 힙화(heapify)하면 $O(hm)$ 계산 시간을 유지할 수 있다.
+단, 힙의 순서는 $i \geq j \iff f_i t_i \geq f_j t_j$의 조건으로 정의하며, 각 반복 단계의 마지막 단계에서 힙을 비우고 다시 힙화(heapify)하면 $O(hm)$ 계산 시간을 유지할 수 있다.
 그러나 (\ref{numericalexperiments}절에서 결과를 제시할) 수리 실험에서 목록 구현으로 구성한 알고리즘이 더 빨랐다.
@@
 \ifen The first inequality follows from the optimality of $\mathcal{X}_h$, while the second follows from the fact that adding $k$ to $\mathcal{X}_{h-1}$ can only increase its valuation.
-\else 첫번째 부등식은 $\mathcal{X}_h$의 최적성에 따르며, 두번째 부등식은 $\mathcal{X}_{h-1}$에 $j$를 더하면 가치가 증가할 수 밖에 없기 때문이다.\fi
+\else 첫 번째 부등식은 $\mathcal{X}_h$의 최적성에 따르며, 두 번째 부등식은 $\mathcal{X}_{h-1}$에 $j$를 더하면 가치가 증가할 수밖에 없기 때문이다.\fi
 
@@
 \else
-아래에서 알고리즘 \ref{algorithmforlargeh}의 몇 반복단계가 나타난다.
+아래에서 알고리즘 \ref{algorithmforlargeh}의 몇 반복 단계가 나타난다.
 $f_j$, $t_j$, 그리고 그 곱의 값은 $\mathcal{C}$에 남아 있는 학교에 해당하는 값만 기록한다.
 $(f * t)_j = f_j t_j$로 정의된 $f * t$는 $f$와 $t$의 원소 당 곱을 의미한다.
-각 반복단계에서 최적해에 추가하는 학교는 $f_j t_j$-값이 가장 높은 학교이며 이를 밑줄로 강조한다.
-\fi
-\newcommand{\Iteration}{\ifen Iteration~\else 반복단계~\fi}
+각 반복 단계에서 최적해에 추가하는 학교는 $f_j t_j$-값이 가장 큰 학교이며 이를 밑줄로 강조한다.
+\fi
+\newcommand{\Iteration}{\ifen Iteration~\else 반복 단계~\fi}
 \begin{align*}
@@
 알고리즘의 출력은 $\mathtt{X} = [4, 2, 8, 1, 7, 3, 5, 6]$이며 최적 $h$-포트폴리오는 그의 첫 $h$개의 원소로 이루어진다.
-표 \ref{planetsdata}에서 등장하는 ``지원 순위'' 자료는 $\mathtt{X}$의 역순열(inverse permutation)이며 이는 해당 학교가 최적 포트폴리오에 포함되는 최소한 $h$-값을 의미한다.
+표 \ref{planetsdata}에서 등장하는 ``지원 순위'' 자료는 $\mathtt{X}$의 역 순열(inverse permutation)이며 이는 해당 학교가 최적 포트폴리오에 포함되는 최소한 $h$-값을 의미한다.
 그림 \ref{planetsplot}은 최적 포트폴리오의 가치를 $h$의 함수로 나타낸다.
@@
 Consider an instance of the knapsack problem, and we will construct an instance of Problem \ref{ellisdecisionform} that is a yes-instance if and only if the corresponding knapsack instance is a yes-instance.
-Without loss of generality, we may assume that the the objects in $\mathcal{B}$ are indexed in increasing order of $u_j$, that each $u_j >0$, and that each $w_j \leq W$.
+Without loss of generality, we may assume that the objects in $\mathcal{B}$ are indexed in increasing order of $u_j$, that each $u_j >0$, and that each $w_j \leq W$.
 \else
@@
 \else
-알마 문제를 위한 탐욕 해법의 직관적인 확장은 $\bigl( v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) \bigr) / g_k$-값이 가장 높은 학교 $k$를  $\mathcal{X}$에 차례대로 추가하는 것이다.
+알마 문제를 위한 탐욕 해법의 직관적인 확장은 $\bigl( v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) \bigr) / g_k$-값이 가장 큰 학교 $k$를  $\mathcal{X}$에 차례대로 추가하는 것이다.
 그러나 위에서 도출한 변환은 엘리스 문제의 목적함수를 배낭 문제의 목적함수의 원하는 만큼 정확한 근사로 만들 수 있음을 의미한다.
@@
 정수 최적화의 고전적인 기법 중 분지한계법(branch and bound)이 있다.
-이는 하나 이상의 결정 변수를 고정시키고 해당 부문제를 탐색하는 해법이며, 목적함수의 상한(bound)을 활용하여 기존 해보다 좋은 해를 생성할 수 없는 결정 나무의 가지(branch)를 제외시켜 결정 공간을 줄인다 (Martello and Toth 1990; Kellerer 외 2004).
+이는 하나 이상의 결정 변수를 고정하고 해당 부문제를 탐색하는 해법이며, 목적함수의 상한(bound)을 활용하여 기존 해보다 좋은 해를 생성할 수 없는 결정 나무의 가지(branch)를 제외하여 결정 공간을 줄인다 (Martello and Toth 1990; Kellerer 외 2004).
 본 항에서, 엘리스 문제의 정수 모형과 그의 상한이 되는 선형 계획을 제시한다.
@@
 문제 \ref{LPrelaxation}은(는) ``연속 배낭 문제''라고 불리며 다음 탐욕 알고리즘을 적용하면 $O(m \log m)$-시간 안에 쉽게 풀 수 있다: $x = \mathbf{0}$로 초기화한다.
-$H > 0$임이 성립하는 동안, $f_j t_j / g_j$-값이 가장 높은 학교 $j$를 택한다.
+$H > 0$임이 성립하는 동안, $f_j t_j / g_j$-값이 가장 큰 학교 $j$를 택한다.
 $x_j \gets \min\{1, H / g_j\}$ 그리고 $H \gets H - g_j$로 수정하여 반복한다 (Dantzig 1957).
@@
 분할이 결정되면 부문제의 한 쌍이 결정해진다.
-첫번째 부문제는 다음 같은 문제 \ref{integernlp}의 인스턴스다.
+첫 번째 부문제는 다음 같은 문제 \ref{integernlp}의 인스턴스다.
 \fi
@@
 \ifen The second is the corresponding instance of Problem \ref{LPrelaxation}:
-\else 두번째 부문제는 대응하는 문제 \ref{LPrelaxation} 인스턴스다.\fi
+\else 두 번째 부문제는 대응하는 문제 \ref{LPrelaxation} 인스턴스다.\fi
 \begin{align} \label{nodelprelaxation}
@@
 $j$를 $\mathcal{I}$로 이동시켜서 한 새끼를 얻고, $j$를 $\mathcal{O}$로 이동시켜서 다른 새끼를 얻는다.
-즉, 한 새끼 마디에서 $x_j = 1$로 고정시키도 다른 새끼 마디에서 $x_j = 0$으로 고정시킨다.
-일반적으로 $j$는 임의의 학교가 될 수 있지만 탐욕적인 휴리스틱으로 $f_j \bar t_j / g_j$의 비율이 가장 높은 학교를 선택한다.
-이런 식으로 새끼를 생성하면 각 마디의 $\mathcal{I}$-집합은 자신의 어미 마디의 $\mathcal{I}$-집합과 최대 한 학교로만 다르다.
-따라서 새로운 마디를 생성할 때, \eqref{howtotransformfj}을(를) 한번만 적용하면 상수 $\gamma$와 변화된 $\bar t_j$-값을 얻을 수 있다.
+즉, 한 새끼 마디에서 $x_j = 1$로 고정하고 다른 새끼 마디에서 $x_j = 0$으로 고정한다.
+일반적으로 $j$는 임의의 학교가 될 수 있지만 탐욕적인 휴리스틱으로 $f_j \bar t_j / g_j$의 비율이 가장 큰 학교를 선택한다.
+이런 식으로 새끼를 생성하면 각 마디의 $\mathcal{I}$-집합은 자기 어미 마디의 $\mathcal{I}$-집합과 최대 한 학교로만 다르다.
+따라서 새로운 마디를 생성할 때, \eqref{howtotransformfj}을(를) 한 번만 적용하면 상수 $\gamma$와 변화된 $\bar t_j$-값을 얻을 수 있다.
 \fi
@@
 이례적인 경우 2가지 있다.
-첫번째는, $\mathcal{N}$에 속한 모든 학교에 대해 $g_j > \bar H$이면 가능성을 유지하며 $\mathcal{I}$에 추가할 수 있는 학교가 없으므로 본 가지의 최적 포트폴리오는 단지 $\mathcal{I}$ 자체다.
+첫 번째는, $\mathcal{N}$에 속한 모든 학교에 대해 $g_j > \bar H$이면 가능성을 유지하며 $\mathcal{I}$에 추가할 수 있는 학교가 없으므로 본 가지의 최적 포트폴리오는 단지 $\mathcal{I}$ 자체다.
 이때 $\mathcal{N}$에 원소를 모두 $\mathcal{O}$로 이동시켜서 단일 새끼를 생성한다.
-두번째는, $\mathcal{N} = \O$인 마디는 새끼가 없다.
+두 번째는, $\mathcal{N} = \O$인 마디는 새끼가 없다.
 더 이상 분지할 수 없으므로 이 마디를 `잎 마디'(leaf)라고 한다.
@@
 $\mathcal{N}$에 속한 학교 중에 $g_j \leq \bar H$인 학교가 있으므로 $n$은 새끼 마디 2개 있다.
-$f_j \bar t_j / g_j$-비율이 가장 높은 학교는 $c_4$이므로 새끼 마디를 생성하는 법칙을 적용하면 $n_1 = (\{2, 4, 5\}, \{1\}, \{3\})$과 $n_2 = (\{2, 5\}, \{1, 4\}, \{3\})$ 같은 새끼를 얻는다.
+$f_j \bar t_j / g_j$-비율이 가장 큰 학교는 $c_4$이므로 새끼 마디를 생성하는 법칙을 적용하면 $n_1 = (\{2, 4, 5\}, \{1\}, \{3\})$과 $n_2 = (\{2, 5\}, \{1, 4\}, \{3\})$ 같은 새끼를 얻는다.
 \fi
@@
 분지한계법을 실행하기 위해 후보 마디, 즉 잎이 아니면서 새끼를 아직 생성하지 않은 마디의 집합을  $\mathfrak{T}$라고 부르자.
-마디 $n= (\mathcal{I}, \mathcal{O}, \mathcal{N})$를 생성할 때마다, $v_{\mathcal{I}}[n] = v(\mathcal{I})$과 선형 완화 문제 \eqref{nodelprelaxation}의 최대값 $v^*_{\mathrm{LP}}[n]$를 기록한다.
-$\mathcal{I}$는 예산에 가능한 포트폴리오이므로 $v_{\mathcal{I}}[n]$는 원래 목적함수 최대값의 하한이 된다.
+마디 $n= (\mathcal{I}, \mathcal{O}, \mathcal{N})$를 생성할 때마다, $v_{\mathcal{I}}[n] = v(\mathcal{I})$과 선형 완화 문제 \eqref{nodelprelaxation}의 최댓값 $v^*_{\mathrm{LP}}[n]$를 기록한다.
+$\mathcal{I}$는 예산에 가능한 포트폴리오이므로 $v_{\mathcal{I}}[n]$는 원래 목적함수 최댓값의 하한이 된다.
 또한 정리 \ref{nestedapplication}의 증명 과정에서 보였듯, \eqref{nodenlpsubproblem}의 목적함수는 함수 $v(\mathcal{I} \cup \mathcal{X})$과 동치다.
 이는 $v_{\mathrm{LP}}^*[n]$는 $\mathcal{I}$를 부분집합으로 포함하며 $\mathcal{O}$에 속한 학교를 포함하지 않은, 즉 이 가지에 있는 모든 포트폴리오의 가치에 대한 상한임을 의미한다.
-따라서 새로운 마디 $n_2$를 생성했을 때, 만약 그의 목적함수값 $v_{\mathcal{I}}[n_2]$가 어떤 다른 마디 $n_1$에 대응하는 $v_{\mathrm{LP}}^*[n_1]$보다 크다면, $n_1$과 그의 모든 후손을 무시할 수 있다.
+따라서 새로운 마디 $n_2$를 생성했을 때, 만약 그의 목적 함숫값 $v_{\mathcal{I}}[n_2]$가 어떤 다른 마디 $n_1$에 대응하는 $v_{\mathrm{LP}}^*[n_1]$보다 크다면, $n_1$과 그의 모든 후손을 무시할 수 있다.
 \fi
@@
 $\mathfrak{T}$에 뿌리 마디 $n_0 = (\O, \O, \mathcal{C})$를 넣어서 알고리즘을 초기화한다.
-각 반복단계에서 $v_{\mathrm{LP}}^*[n]$-값이 가장 높은 마디 $n \in \mathfrak{T}$을 선택한다.
+각 반복 단계에서 $v_{\mathrm{LP}}^*[n]$-값이 가장 큰 마디 $n \in \mathfrak{T}$을 선택한다.
 $n$의 새끼를 생성하고 후보 집합에서 제거한다.
@@
 
-알고리즘에서 회로가 생기지 않음을 보이기 위해 어떤 한 마디가 두번 생성되지 않는 것을 증명하면 충분하다.
+알고리즘에서 회로가 생기지 않음을 보이기 위해 어떤 한 마디가 두 번 생성되지 않는 것을 증명하면 충분하다.
 모순을 보이기 위해 같은 분할 $(\mathcal{I}_{12}, \mathcal{O}_{12}, \mathcal{N}_{12})$를 가지는 상이한 마디 $n_1$과 $n_2$가 생성된다고 하자.
@@
 $n$은 새끼 마디 하나만 가지면 그 자체가  $n_1$과 $n_2$의 공통 선조가 되며 이는 모순이므로 $n$ 새끼의 개수가 2임을 알 수 있다.
-또한 그 중 하나는 $n_1$의 선조며 다른 하나는 $n_2$의 선조다.
+또한 그중 하나는 $n_1$의 선조며 다른 하나는 $n_2$의 선조다.
 각각 $n_3 = (\mathcal{I}_{3}, \mathcal{O}_{3}, \mathcal{N}_{3})$ 그리고 $n_4 = (\mathcal{I}_{4}, \mathcal{O}_{4}, \mathcal{N}_{4})$라고 하자.
@@
 또한 $\mathcal{N}$에 학교가 많을 때, 선형 완화 문제 상한은 $v_{\mathcal{I}}[n']$보다 매우 높을 수도 있다.
-이는 해법이 나무에 어느 정도 깊게 파고들어야 마디를 제외시킬 수 있음을 의미하며, 그때 계산 노력의 대부분은 이미 쏟은 것이다.
+이는 해법이 나무에 어느 정도 깊게 파고들어야 마디를 제외할 수 있음을 의미하며, 그때 계산 노력의 대부분은 이미 쏟은 것이다.
 수리 실험에서, 알고리즘 \ref{ellisbnb}은(는) 약 $m=35$개의 학교를 넘는 인스턴스에 대해 비효율적이었으며 이는 단순한 열거법에 비해 큰 개선이 아니다.
@@
 \else
-본 항 나머지에서, 다른 언급이 없으면 포트폴리오의 ``가치''라고 하면 고정소수점 환경 안에서의 가치를 의미하며 이것이 근사값임을 유념한다. 근사 오차는 나중에 동적 계획의 타당성을 증명할 때 처리한다.
+본 항 나머지에서, 다른 언급이 없으면 포트폴리오의 ``가치''라고 하면 고정소수점 환경 안에서의 가치를 의미하며 이것이 근삿값임을 유념한다. 근사 오차는 나중에 동적 계획의 타당성을 증명할 때 처리한다.
 \fi
@@
 \else
-정수 $0 \leq j \leq m$ 그리고 $v \in [-\infty, 0) \cup \mathcal{V}$에 대해, $\{ 1, \dots, j\}$에 속한 학교만 사용하면 최소한 $v$의 가치를 가지는 포트폴리오가 존재한다면 그 중 가장 비용이 낮은 포트폴리오를  $\mathcal{W}[j, v]$라고 하자.
+정수 $0 \leq j \leq m$ 그리고 $v \in [-\infty, 0) \cup \mathcal{V}$에 대해, $\{ 1, \dots, j\}$에 속한 학교만 사용하면 최소한 $v$의 가치를 가지는 포트폴리오가 존재한다면 그중 가장 비용이 낮은 포트폴리오를  $\mathcal{W}[j, v]$라고 하자.
 그의 비용을 $G[j, v] = \sum_{j\in \mathcal{W}[j, v]} g_j$라고 하자.
@@
 각 지표에 대해 $G[j, v]$를 계산한 다음에,  $\mathcal{W}[j, v]$가 $j$를 포함할 필수충분 조건인 $G[j, v] < G[j-1, v]$를 적용하면 해당 최적 포트폴리오를 구할 수 있다.
-그러면 가능하면서 최대적인 목적함수값 $\max\{ w: G[m, w] \leq H\}$와 해당 포트폴리오를 계산하면 엘리스 문제의 근사해를 얻을 수 있다.
+그러면 가능하면서 최대한 목적 함숫값 $\max\{ w: G[m, w] \leq H\}$와 해당 포트폴리오를 계산하면 엘리스 문제의 근사해를 얻을 수 있다.
 \fi
@@
 \else
-\eqref{recursionrelationforcostmindp}과(와) \eqref{deltajvdef}(으)로 정의된 점화식에 따라 $\Delta_j(v)$를 내림하기 때문에, $j \in \mathcal{W}[j, v]$이면 $(1 - f_j) v\bigl(\mathcal{W}[j-1, v- \Delta_j(v)]\bigr) + f_j t_j$의 실제값은 $\mathcal{W}[j, v]$의 고정소수점 가치인 $v$보다 클 수 있는데 최대한 $10^{-P}$ 만큼의 차이가 발생하다.
+\eqref{recursionrelationforcostmindp}과(와) \eqref{deltajvdef}(으)로 정의된 점화식에 따라 $\Delta_j(v)$를 내림하기 때문에, $j \in \mathcal{W}[j, v]$이면 $(1 - f_j) v\bigl(\mathcal{W}[j-1, v- \Delta_j(v)]\bigr) + f_j t_j$의 실제값은 $\mathcal{W}[j, v]$의 고정소수점 가치인 $v$보다 클 수 있는데 최대한 $10^{-P}$ 만큼의 차이가 발생한다.
 또한 이 오차는 $\mathcal{W}$에 학교를 추가할 때마다 가산적으로 누적되는데 추가하는 학교의 개수는 최대한 $m$이다.
@@
 
-알고리즘 \ref{ellisDP1}과(와) 알고리즘 \ref{ellisDP3}을(를) 표 대신 재귀 함수 기반으로 구성할 수 있다. 그러나 각 재귀 함수가 스스로를 두번 호출하므로 지수적으로 늘어나는 호출 수를 방지하기 위해 각 지표에 대응하는 함수값을 표에 저장하거나 다른 식으로 기록해야 한다.
+알고리즘 \ref{ellisDP1}과(와) 알고리즘 \ref{ellisDP3}을(를) 표 대신 재귀 함수 기반으로 구성할 수 있다. 그러나 각 재귀 함수가 스스로를 두 번 호출하므로 지수적으로 늘어나는 호출 수를 방지하기 위해 각 지표에 대응하는 함숫값을 표에 저장하거나 다른 식으로 기록해야 한다.
 \fi
@@
 We made extensive use of the DataStructures.jl package (v0.18.11, \url{github.com/JuliaCollections}).
-The code is available at \url{https://github.com/maxkapur/OptimalApplication}.
+The code is available at \url{github.com/maxkapur/OptimalApplication}.
 \else
 모든 알고리즘을 쥴리아 (Julia, v1.8.0b1) 언어로 구현했다.
-쥴리아는 $t_j$-값이 정수 또는 부동소수점 수와 같은 특수한 경우에 대해 컴파일러가 자동적으로 최적화하는 모수적 자료형(parametric data type) 기능이 있어서 유용한 선택이었다.
+쥴리아는 $t_j$-값이 정수 또는 부동소수점 수와 같은 특수한 경우에 대해 컴파일러가 자동으로 최적화하는 모수적 자료형(parametric data type) 기능이 있어서 유용한 선택이었다.
 또한 병렬 연산을 쉽게 관리할 수 있는 매크로를 제공하므로 계산 실험의 규모를 증가할 수 있었다 (Bezanson 외 2017).
 DataStructures.jl 패키지(v0.18.11, \url{github.com/JuliaCollections})도 활용했다.
-실험 코드는  \url{https://github.com/maxkapur/OptimalApplication}에서 공유한다.
-\fi
-
-\ifen
-The dynamic programs, namely Algorithms \ref{ellisDP1} and \ref{ellisDP3}, were implemented using recursive functions and dictionary memoization rather than a full lookup table.
-Our implementation of Algorithm \ref{ellisDP3} also differed from that described in Subsection \ref{fptashead} in that we represented portfolio valuations in \emph{binary} rather than decimal, with the definitions of $P$ and $\mathcal{V}$ modified accordingly, and instead of fixed-point numbers, we worked in integers by multiplying each $t_j$-value by $2^P$.
-These modifications yielded a substantial performance improvement without changing the fundamental algorithm design or complexity analysis.
-\else
-동적 계획 해법인 \ref{ellisDP1}과(와) 알고리즘 \ref{ellisDP3}을(를) 표 대신 사전 기록 및 재귀 함수로 실행했다.
+실험 코드는  \url{github.com/maxkapur/OptimalApplication}에서 공유한다.
+\fi
+
+\ifen
+The dynamic programs, namely Algorithms \ref{ellisDP1} and \ref{ellisDP3}, were implemented using recursive functions and dictionary memoization, as described in Cormen et al.
+(1990, \S\,16.6).
+Our implementation of Algorithm \ref{ellisDP3} also differed from that described in Subsection \ref{fptashead} in that we represented portfolio valuations in \emph{binary} rather than decimal, with the definitions of $P$ and $\mathcal{V}$ modified accordingly, and instead of fixed-point numbers, we worked in integers by multiplying each element of $\mathcal{V}$ by $2^P$.
+These modifications yield a substantial performance improvement without changing the fundamental algorithm design or complexity analysis.
+\else
+동적 계획 해법인 \ref{ellisDP1}과(와) 알고리즘 \ref{ellisDP3}을(를) 재귀 함수와 사전 기록으로 구현했다 (Cormen 외 1990, \S\,16.6).
 알고리즘 \ref{ellisDP3}은(는) \ref{fptashead}항에서 설명한 것과 달리, 포트폴리오의 가치를 십진수 대신 이진수로 나타냈으며 $P$와  $\mathcal{V}$의 정의를 적절히 수정했다.
-그리고 고정소수점 대신 $t_j$-값을 $2^P$으로 곱하여 정수로 구성했다.
-이런 변경은 기본 알고리즘 설계와 계산 시간 분석에 영향주지 않는데 성능이 많이 개선된다.
+그리고 고정소수점 대신 $\mathcal{V}$의 모든 원소를 $2^P$으로 곱하여 정수로 변환했다.
+이런 변경은 기본 알고리즘 설계와 계산 시간 분석에 영향을 주지 않는데 성능이 많이 개선된다.
 \fi
@@
 실험 방법은 다음과 같다.
-먼저, 가산 시장을 생성하기 위해 $t_j$-값을 평균이 10인 지수 분포에서 동립적으로 관측하고 정수로 올림했다.
+먼저, 가산 시장을 생성하기 위해 $t_j$-값을 평균이 10인 지수 분포에서 독립적으로 관측하고 정수로 올림했다.
 그다음에, $t_j$와 $f_j$가 서로 부분적으로 반비례하게 만들도록 $f_j = 1 / (t_j + 10\,Q)$로 정의했다.
 단, $Q$는 $[0, 1)$에서 균일한 확률로 관측한다.
-알마의 문제를 고려하는 첫번째 실험에서, 모든 $g_j = 1$ 그리고 $H = h = \lfloor m/ 2 \rfloor$으로 처리했다.
-엘리스의 문제를 고려하는 두번째 실험에서, 모든  $g_j$를 $\{5, \dots, 10\}$에서 균일한 확률로 관측했으며 $H = \lfloor \frac{1}{2} \sum g_j \rfloor$으로 처리했다.
+알마의 문제를 고려하는 첫 번째 실험에서, 모든 $g_j = 1$ 그리고 $H = h = \lfloor m/ 2 \rfloor$으로 처리했다.
+엘리스의 문제를 고려하는 두 번째 실험에서, 모든  $g_j$를 $\{5, \dots, 10\}$에서 균일한 확률로 관측했으며 $H = \lfloor \frac{1}{2} \sum g_j \rfloor$으로 처리했다.
 실제 대학 지원 비용인 50과 100달러 사이에 10달러의 어떤 배수이므로 합리적인 비용 분포이다.
@@
 \nmarkets 개의 시장에 대한 평균과 평균편차를 표에 나타낸다.
-따라서 아래 표의 각 칸은 {\the\numexpr 3 * \nmarkets \relax}개의 계한에 대한 통계다.
+따라서 아래 표의 각 칸은 {\the\numexpr 3 * \nmarkets \relax}개의 계산에 대한 통계다.
 $t$를 배열하는 시간은 고려하지 않았다.
@@
 \else
-첫번째 실험에서, 동일한 지원 비용으로 정의된 크기가 다양한 시장에 대해 알고리즘 \ref{algorithmforlargeh}의 성능을 고려했다.
+첫 번째 실험에서, 동일한 지원 비용으로 정의된 크기가 다양한 시장에 대해 알고리즘 \ref{algorithmforlargeh}의 성능을 고려했다.
 이를 $\mathcal{C}$를 목록 혹은 $f_j \bar t_j$-값에 따라 배열된 힙 구현으로 구성하는 경우로 나눴으며 실험 결과는 표 \ref{experiment1results}에서 등장한다.
@@
 %\else
-%첫번째 실험에서, 동일한 지원 비용으로 정의된 크기가 다양한 시장에 대해 알고리즘 \ref{algorithmforlargeh}의 성능을 고려했다.
+%첫 번째 실험에서, 동일한 지원 비용으로 정의된 크기가 다양한 시장에 대해 알고리즘 \ref{algorithmforlargeh}의 성능을 고려했다.
 이를 $\mathcal{C}$를 목록 혹은 $f_j \bar t_j$-값에 따라 배열된 힙 구현으로 구성하는 경우로 나눴으며 실험 결과는 표 \ref{experiment1results}에서 등장한다.
@@
 \else
-두번째 실험에서 일반적인 문제를 고려한다.
+두 번째 실험에서 일반적인 문제를 고려한다.
 정확한 해법 (알고리즘 \ref{ellisbnb} 및 알고리즘 \ref{ellisDP1}) 그리고 허용치를 0.5와 0.05로 설정한 근사 해법 (알고리즘 \ref{ellisDP3})의 성능을 비교한다.
@@
 The table shows the average time (standard deviation) over the \nmarkets~instances.
-\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복하여 그 중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (평균편차) 시간이 나타난다.\fi}
+\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복하여 그중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (평균편차) 시간이 나타난다.\fi}
 
@@
 단, $\tau_j = t_j - \beta t_j^2$.
-이의 첫번째 항 그 자체가 포트폴리오 가치 함수이며, 두번째 항은 포트폴리오 가치 함수의 단조 변환이다.
+이의 첫 번째 항 그 자체가 포트폴리오 가치 함수이며, 두 번째 항은 포트폴리오 가치 함수의 단조 변환이다.
 따라서 위에서 제시한 알고리즘을 서브루틴으로 사용하면 효율적 투자선(efficient frontier)을 탐색할 수 있어 보인다.
@@
 A parametric treatment of risk aversion may align our model more closely to real-world applicant behavior.
-Conventional wisdom asserts that the best college application strategy combines reach, target, and safety schools in roughly equal proportion (Kim 2015).
+Conventional wisdom asserts that the best college application strategy combines reach, target, and safety schools in roughly equal proportion (Jeon 2015).
 When the application budget is small relative to the size of the admissions market, our algorithms recommend a similar approach (see the example of Subsection \ref{planetsexamplesection}).
@@
 위험 회피를 모수화한 모형은 실제 지원 행동을 더 정확하게 반영할 수 있다.
-``상향·소신·안정'' 지원 학교를 대략 균일하게 분산하여 지원하는 것이 최적이란 통념이다 (김민희 2015).
+``상향·소신·안정'' 지원 학교를 대략 균일하게 분산하여 지원하는 것이 최적이란 통념이다 (전민희 2015).
 입학 시장의 크기에 비한 지원 예산이 작은 경우, 본 연구에서 제시하는 알고리즘은 비슷한 전략을 추천한다 (항 \ref{planetsexamplesection}의 예제 참고).
@@
 \else
-향후 연구에 두번째 방향은 모형에 시그널 전략을 도입하는 것이다.
+향후 연구에 두 번째 방향은 모형에 시그널 전략을 도입하는 것이다.
 한국 입학 과정의 온라인 지원 양식에는 가군, 나군, 다군이란 3가지 선다형 칸을 채워서 지원하는 학교를 선택한다.
-대부분의 학교는 모든 3개의 칸에 등장하지 않고 그 중 1~2개만 가능하다.
+대부분의 학교는 모든 3개의 칸에 등장하지 않고 그중 1~2개만 가능하다.
 따라서 학생이 지원할 수 있는 학교의 `수'일 뿐만 아니라 가능한 지원 포트폴리오에 같이 들어갈 수 있는 학교의 `조합'도 제한된다.
@@
 추상적인 관점에서, 알고리즘 \ref{ellisDP1}과(와) 알고리즘 \ref{ellisDP3}은(는) 최적해를 $Z[N, C]$로 표현하는 2차원 동적 계획이다.
-이때 $N$은 결정 변수의 수, $C$는 제약식의 모수, 그리고 $Z[n, c]$는 제약 모수가 $c \leq C$일 때 첫 $n\leq N$개의 결정 변수만 이용하는 최적 목적함수값이다.
+이때 $N$은 결정 변수의 수, $C$는 제약식의 모수, 그리고 $Z[n, c]$는 제약 모수가 $c \leq C$일 때 첫 $n\leq N$개의 결정 변수만 이용하는 최적 목적 함숫값이다.
 알고리즘의 구조는 $Z[n, c]$를 $Z[n -1, c]$와 어떤 $c' \leq c$에 대응하는 $Z[n -1, c']$의 함수로 표현하고 그 반복 관계로 탐색하는 것이다.
@@
 \ifen \else
-김민희. 2015. ``{[대입 수시 전략]} 총 6번의 기회 $\cdots$ `상향·소신·안정' 분산 지원하라.'' 중앙일보, 8월 26일. \url{https://www.joongang.co.kr/article/18524069}.\fi
+전민희. 2015. ``{[대입 수시 전략]} 총 6번의 기회 $\cdots$ `상향·소신·안정' 분산 지원하라.'' 중앙일보, 8월 26일. \url{https://www.joongang.co.kr/article/18524069}.\fi
 
@@
 1973.
-``Time bounds for selection.'' \emph{Journal of Computer and System Sciences} 7 (4): 448--61.
+``Time Bounds for Selection.'' \emph{Journal of Computer and System Sciences} 7 (4): 448--61.
 \url{https://doi.org/10.1016/S0022-0000(73)80033-9}.
@@
 1993.
-``An algorithm for maximizing target achievement in the stochastic knapsack problem with normal returns.'' \emph{Naval Research Logistics} 40 (2): 161--73.
+``An Algorithm for Maximizing Target Achievement in the Stochastic Knapsack Problem with Normal Returns.'' \emph{Naval Research Logistics} 40 (2): 161--73.
 \url{https://doi.org/10.1002/nav.3220400203}.
+
+Cormen, Thomas, Charles Leiserson, and Ronald Rivest. 1990. \emph{Introduction to Algorithms.} Cambridge, MA: The MIT Press.
 
@@
 
-\ifen Kim, Minhee.
+\ifen Jeon, Minhee.
 2015.
@@
 
-Fisher, Marshall, George Nemhauser, and Laurence Wolsey. 1978. ``An analysis of approximations for maximizing submodular set functions—I.'' \emph{Mathematical Programming} 14: 265--94.
-
-Fredman, Michael Lawrence and Robert Tarjan.
+Fisher, Marshall, George Nemhauser, and Laurence Wolsey. 1978. ``An Analysis of Approximations for Maximizing Submodular Set Functions—I.'' \emph{Mathematical Programming} 14: 265--94.
+
+%Fredman, Michael Lawrence and Robert Tarjan.
 1987.
-``Fibonacci heaps and their uses in improved network optimization algorithms.'' \emph{Journal of the Association for Computing Machinery} 34 (3): 596--615.
+``Fibonacci Heaps and Their Uses in Improved Network Optimization Algorithms.'' \emph{Journal of the Association for Computing Machinery} 34 (3): 596--615.
 
@@
 2020.
-``The nestedness property of the convex ordered median location problem on a tree.'' \emph{Discrete Optimization} 36: 100581.
+``The Nestedness Property of the Convex Ordered Median Location Problem on a Tree.'' \emph{Discrete Optimization} 36: 100581.
 \url{https://doi.org/10.1016/j.disopt.2020.100581}.
@@
 
-Van Huizen, Thomas and Rob Alessie. 2019. ``Risk aversion and job mobility.’’ \emph{Journal of Economic Behavior \& Organization} 164: 91--106. \url{https://doi.org/10.1016/j.jebo.2019.01.021}.
+Van Huizen, Thomas and Rob Alessie. 2019. ``Risk Aversion and Job Mobility.’’ \emph{Journal of Economic Behavior \& Organization} 164: 91--106. \url{https://doi.org/10.1016/j.jebo.2019.01.021}.
 

### 2022-04-05 src/EnumerationSolvers.jl

@@ -3,7 +3,7 @@
 
 Produce the optimal portfolio for the market `mkt` having identical application costs.
 """
-function optimalportfolio_enumerate(mkt::SameCostsMarket{T})::Tuple{Vector{Int},Float64} where T
+function optimalportfolio_enumerate(mkt::SameCostsMarket{T, U})::Tuple{Vector{Int},Float64} where {T, U}
     X = zeros(T, mkt.h)
     v = 0.0
 

### 2022-04-05 src/FastSolvers.jl

@@ -30,7 +30,7 @@ isless(c1::College, c2::College) = isless(c1.ft, c2.ft)
 Produce the optimal order of application for the market `mkt` having identical
 application costs and the corresponding portfolio valuations.
 """
-function applicationorder_list(mkt::SameCostsMarket{T}; verbose=false::Bool)::Tuple{Vector{Int},Vector{Float64}} where T
+function applicationorder_list(mkt::SameCostsMarket{T, U}; verbose=false::Bool)::Tuple{Vector{Int},Vector{Float64}} where {T, U}
     apporder = zeros(T, mkt.h)
     v = zeros(mkt.h)
 
@@ -100,7 +100,7 @@ end
 Produce the optimal order of application for the market `mkt` having identical
 application costs and the corresponding portfolio valuations.
 """
-function applicationorder_heap(mkt::SameCostsMarket{T})::Tuple{Vector{Int},Vector{Float64}} where {T}
+function applicationorder_heap(mkt::SameCostsMarket{T})::Tuple{Vector{Int},Vector{Float64}} where {T, U}
     apporder = zeros(T, mkt.h)
     v = zeros(mkt.h)
 
@@ -134,7 +134,7 @@ end
 Use the dynamic program on application costs to produce the optimal portfolio `X` and associated
 valuation table `V` for the market `mkt` with varying application costs.
 """
-function optimalportfolio_valuationtable(mkt::VariedCostsMarket{T})::Tuple{Vector{Int},Matrix{Float64}} where {T}
+function optimalportfolio_valuationtable(mkt::VariedCostsMarket{T})::Tuple{Vector{Int},Matrix{Float64}} where T
     V = zeros(mkt.m, mkt.H)
     @inbounds for j in T(1):mkt.m, h in 1:mkt.H
         if h < mkt.g[j]
@@ -187,7 +187,7 @@ end
 Use the dynamic program on application costs to produce the optimal portfolio `X` and associated
 value `v` for the market `mkt` with varying application costs. 
 """
-function optimalportfolio_dynamicprogram(mkt::VariedCostsMarket{T}; verbose=false::Bool)::Tuple{Vector{Int},Float64} where {T}
+function optimalportfolio_dynamicprogram(mkt::VariedCostsMarket{T}; verbose=false::Bool)::Tuple{Vector{Int},Float64} where T
[…]

### 2022-04-05 src/MarketIO.jl

@@ -2,7 +2,7 @@ const UIntTypes = DataType[UInt8, UInt16, UInt32, UInt64]
 
 
 # TO DO: Use Julian errors rather than assertions here.
-function iscoherentmarket(f::Vector{Float64}, t::Vector{Int})
+function iscoherentmarket(f::Vector{Float64}, t::Vector{<:Real})
     @assert length(f) == length(t)
     @assert all(0 .< f .≤ 1)
     @assert issorted(t)
@@ -17,17 +17,13 @@ function iscoherentmarket(f::Vector{Float64}, t::Vector{Int}, g::Vector{Int})
 end
 
 
-function isnontrivialmarket(f::Vector{Float64}, t::Vector{Int}, h::Int)
+function isnontrivialmarket(f::Vector{Float64}, t::Vector{<:Real}, h::Int)
     iscoherentmarket(f, t)
     @assert 0 < h ≤ length(t)
 end
 
 
-function isnontrivialmarket(
-    f::Vector{Float64},
-    t::Vector{Int},
-    g::Vector{Int},
-    H::Int)
+function isnontrivialmarket(f::Vector{Float64}, t::Vector{Int}, g::Vector{Int}, H::Int)
     iscoherentmarket(f, t, g)
     @assert 0 < H ≤ sum(g)
     @assert all(0 .< g .≤ H)
@@ -46,10 +42,10 @@ application costs. Fields:
     `omf = 1 .- f`
     `perm = sortperm(perm)`: Currently a placeholder
 """
-struct SameCostsMarket{T<:Unsigned}
+struct SameCostsMarket{T<:Unsigned, U<:Real}
     m::T
     f::Vector{Float64}
-    t::Vector{Int}
+    t::Vector{U}
     h::T
     ft::Vector{Float64}      # = f .* t
     omf::Vector{Float64}     # = 1 .- f
@@ -62,7 +58,7 @@ struct SameCostsMarket{T<:Unsigned}
     admissions probabilities `f`, utility values `t`, and application limit `h`
     and return the market object.
     """
-    function SameCostsMarket(f::Vector{Float64}, t::Vector{Int}, h::Integer)
+    function SameCostsMarket(f::Vector{Float64}, t::Vector{U}, h::Integer) where U<:Real
         isnontrivialmarket(f, t, h)
         T = UIntTypes[findfirst(T -> length(f) < typemax(T), UIntTypes)]
         m = T(length(f))
[…]

### 2022-04-05 test/runtests.jl

@@ -6,11 +6,28 @@ using Random
 const n_markets = 3
 
 @testset verbose = true "OptimalApplication.jl" begin
-    @testset verbose = true "Same app. costs" begin
+    @testset verbose = true "Same app. costs, t int" begin
         m = 20
 
         for _ in 1:n_markets
             mkt = SameCostsMarket(m)
+
+            X_enum, vX_enum = optimalportfolio_enumerate(mkt)
+            sort!(X_enum)
+            X_heap, vX_heap = applicationorder_heap(mkt)
+            X_list, vX_list = applicationorder_list(mkt)
+            @test X_enum == sort(X_heap)
+            @test vX_enum ≈ last(vX_heap)
+            @test X_enum == sort(X_list)
+            @test vX_enum ≈ last(vX_list)
+        end
+    end
+
+    @testset verbose = true "Same app. costs, t float" begin
+        m = 20
+
+        for _ in 1:n_markets
+            mkt = SameCostsMarket(rand(16), sort(12 * rand(16)), m ÷ 4)
 
             X_enum, vX_enum = optimalportfolio_enumerate(mkt)
             sort!(X_enum)
@@ -82,15 +99,6 @@ const n_markets = 3
 
         @test_throws AssertionError Market(f, t, 1)
         @test_throws AssertionError Market(f, t, g, H)
-
-        # t not integer
-        f = [0.1, 0.1]
-        t = [4.0, 7.0]
-        g = [2, 2]
-        H = 3
-
-        @test_throws MethodError Market(f, t, 1)
-        @test_throws MethodError Market(f, t, g, H)
 
         # t not sorted
         f = [0.1, 0.1]

### 2022-04-07 paper/paper.tex
line edits

@@
 \newif\ifen
-\enfalse % Change to false to export Korean
+\entrue % Change to false to export Korean
 
@@
 \ifen \section{Notation and preliminary results} \else \section{표기법과 예비 결과}\fi\label{preliminaries}
-\ifen Before discussing the solution algorithms, we will introduce some additional notation and a few preliminary results that will come in handy.
-
-For the remainder of the paper, unless otherwise noted, we assume with trivial loss of generality that each $g_j \leq H$, each $f_j \in (0, 1]$, and $t_0 < t_1 \leq \cdots \leq t_m$.
+\ifen Before discussing the solution algorithms, we will introduce additional notation and a few preliminary results.
+For the remainder of the paper, unless otherwise noted, we assume with trivial loss of generality that each $g_j \leq H$, $\sum g_j > H$, each $f_j \in (0, 1]$, and $t_0 < t_1 \leq \cdots \leq t_m$.
 %Below, we will show how to transform an arbitrary instance so that $t_0 = 0$, in which case each $t_j > 0$.
@@
 해법은 의논하기 전, 본 절에서 추가적인 표기법은 명시하고 몇 가지 유용한 예비 결과를 정리하고자 한다.
-
-남은 논문에서 다른 언급이 없으면, 일반성을 잃지 않고 각 $g_j \leq H$, 각 $f_j \in (0, 1]$, 그리고 $t_0 < t_1 \leq \cdots \leq t_m$이라고 가정한다. 뒤에서 임의의 인스턴스를 $t_0 = 0$이 되도록 변환하는 방법을 제시한다. % 이를 적용하면 $t_j > 0$인 가정을 세울 수 있다. 
+남은 논문에서 다른 언급이 없으면, 일반성을 잃지 않고 각 $g_j \leq H$, $\sum g_j > H$, 각 $f_j \in (0, 1]$, 그리고 $t_0 < t_1 \leq \cdots \leq t_m$이라고 가정한다.
+뒤에서 임의의 인스턴스를 $t_0 = 0$이 되도록 변환하는 방법을 제시한다.
+% 이를 적용하면 $t_j > 0$인 가정을 세울 수 있다.
 \fi
@@
 \ifen 
-\noindent This algorithm's computation time is $O(m)$ using the \small PICK \normalsize algorithm of Blum et al. (1973).
+\noindent This algorithm's computation time is $O(m)$ using the PICK algorithm of Blum et al. (1973).
 
@@
 \else
-\noindent Blum 외 (1973)의 \small PICK \normalsize 알고리즘을 사용하면 위의 해법의 계산 시간이 $O(m)$이다.
+\noindent Blum 외 (1973)의 PICK 알고리즘을 사용하면 위의 해법의 계산 시간이 $O(m)$이다.
 
@@
 $v(\mathcal{T}) = 70(0.4)(1-0.4) + 80(0.4) = 48.8$.
-But $\mathcal{X} = \{2, 3\}$ with￣
+But $\mathcal{X} = \{2, 3\}$ with
 $v(\mathcal{X}) = 80(0.4)(1-0.3) + 90(0.3) = 49.4$
@@
 \end{example}
-
-\ifen
-In fact, the na\"ive algorithm is a $(1/h)$-approximation algorithm for Alma’s problem, as expressed in the following theorem.
-\else
-다음 정리는 나이브 해법이 알마의 문제를 위한 $(1/h)$-근사 해법임을 의미한다.
+\ifen
+\noindent In fact, the na\"ive algorithm is a $(1/h)$-approximation algorithm for Alma’s problem, as expressed in the following theorem.
+\else
+\noindent 다음 정리는 나이브 해법이 알마의 문제를 위한 $(1/h)$-근사 해법임을 의미한다.
 \fi
@@
 \end{proof}
-
-\ifen
-The following example establishes the tightness of the approximation factor. 
-\else
-다음 예는 근사 계수가 타이트(tight)함을 보인다.
+\ifen
+\noindent The following example establishes the tightness of the approximation factor. 
+\else
+\noindent 다음 예는 근사 계수가 타이트(tight)함을 보인다.
 \fi
@@
 \fi
-\begin{align*}
- f &= \Bigl(\,\underbrace{1,  \dots, 1}_{h},\;
-\underbrace{\varepsilon^{1}, \varepsilon^{2}, \dots, \varepsilon^{h-1}, \varepsilon^{h}}_{h}\,\Bigr) \\
-\text{\ifen and\else 그리고\fi}\quad t &= \Bigl(\,\underbrace{1,  \dots, 1}_{h},\;
-\underbrace{\varepsilon^{-1}, \varepsilon^{-2}, \dots, \varepsilon^{-(h-1)}, \varepsilon^{-h}}_{h}\,\Bigr) 
-\end{align*}
+%\begin{align}
+% f &= \Bigl(\,\underbrace{1,  \dots, 1}_{h},\;
+%\underbrace{\varepsilon^{1}, \varepsilon^{2}, \dots, \varepsilon^{h-1}, \varepsilon^{h}}_{h}\,\Bigr) \\
+%\text{\ifen and\else 그리고\fi}\quad t &= \Bigl(\,\underbrace{1,  \dots, 1}_{h},\;
+%\underbrace{\varepsilon^{-1}, \varepsilon^{-2}, \dots, \varepsilon^{-(h-1)}, \varepsilon^{-h}}_{h}\,\Bigr) 
+%\end{align}
+\begin{center}
+\begin{tabular}{r|cccccccc}
+$j$   & $1$      & $\cdots$ & $h$   &$h+1$         &  $h+2$ & $\cdots$ &      $m-1$  & $m$            \\ \hline
+$f_j$ & $1$     &  $\cdots$ & $1$      & $\varepsilon^{1}$ & $\varepsilon^{2}$ & $\cdots$ & $\varepsilon^{h-1}$ & $\varepsilon^{h}$ \\
+$t_j$ & $1$      &  $\cdots$ & $1$      & $\varepsilon^{-1}$ & $\varepsilon^{-2}$ & $\cdots$ & $\varepsilon^{-(h-1)}$ & $\varepsilon^{-h}$
+\end{tabular}%
+\end{center}
 \ifen
@@
 \fi
-\begin{equation*}
+\begin{equation}
 v(\mathcal{X}_h) = \sum_{j= h +1}^m \Bigl( f_j t_j \prod_{j' = j+1}^m (1 - f_{j'}) \Bigr) =  \sum_{j= 1}^h  (1 - \varepsilon)^{j} \approx h\ifen.\fi
-\end{equation*}
+\end{equation}
 \ifen
@@
 \end{proof}
-
-\ifen
-It follows that when $\mathcal{X}_h$ is the optimal $h$-portfolio for a given market, $v(\mathcal{X}_h)$ is $O(h)$.
+\ifen
+\noindent It follows that when $\mathcal{X}_h$ is the optimal $h$-portfolio for a given market, $v(\mathcal{X}_h)$ is $O(h)$.
 Example \ref{tightexampleforoneoverhopt}, in which $v(\mathcal{X}_h)$ can be made arbitrarily close to $h$, establishes the tightness of this bound.
 \else
-위 결과는 $\mathcal{X}_h$가 어떤 시장의 최적 $h$-포트폴리오일 때, $v(\mathcal{X}_h)$가 $O(h)$ 함수라고 의미한다. 예 \ref{tightexampleforoneoverhopt}에서 $v(\mathcal{X}_h)$를 $h$에 임의로 가깝게 조정할 수 있으므로 이 상한이 타이트함을 알 수 있다.
+\noindent 위 결과는 $\mathcal{X}_h$가 어떤 시장의 최적 $h$-포트폴리오일 때, $v(\mathcal{X}_h)$가 $O(h)$ 함수라고 의미한다. 예 \ref{tightexampleforoneoverhopt}에서 $v(\mathcal{X}_h)$를 $h$에 임의로 가깝게 조정할 수 있으므로 이 상한이 타이트함을 알 수 있다.
 \fi
@@
 \ifen 
-Let us consider a fictional admissions market consisting of $m=8$ schools. The school data, along with the optimal solutions for each $h \leq m$, appear in Table \ref{planetsdata}.
+Let us examine a fictional admissions market consisting of $m=8$ schools. The school data, along with the optimal solutions for each $h \leq m$, appear in Table \ref{planetsdata}.
 \else 
@@
 \begin{tabular}{r|lcccc}
-\ifen\textbf{Index $j$} & \textbf{School $c_j$} & \textbf{Utility $t_j$} & \textbf{Admit prob. $f_j$} & \textbf{Priority} & \textbf{$v(\mathcal{X}_h)$} \\ \hline
-\else 
-\textbf{지표 $j$} & \textbf{학교 $c_j$} & \textbf{효용 $t_j$} & \textbf{합격 확률 $f_j$} & \textbf{지원 순위} & \textbf{$v(\mathcal{X}_h)$} \\ \hline  \fi
+\ifen\textbf{Index $j$} & \textbf{School $c_j$} & \textbf{Admit prob. $f_j$} & \textbf{Utility $t_j$} & \textbf{Priority} & \textbf{$v(\mathcal{X}_h)$} \\ \hline
+\else 
+\textbf{지표 $j$} & \textbf{학교 $c_j$} & \textbf{합격 확률 $f_j$} & \textbf{효용 $t_j$}  & \textbf{지원 순위} & \textbf{$v(\mathcal{X}_h)$} \\ \hline  \fi
 \\[-.75em]
-1 & \ifen Mercury University   \else  수성대       \fi        & 200 & 0.39 & 4   & 230.0   \\
-2 & \ifen Venus University     \else  금성대       \fi        & 250 & 0.33 & 2   & 146.7  \\
-3 & \ifen Mars University      \else  화성대       \fi        & 300 & 0.24 & 6   & 281.5  \\
-4 & \ifen Jupiter University   \else  목성대       \fi        & 350 & 0.24 & 1   & \phantom{0}84.0  \\
-5 & \ifen Saturn University    \else  토성대       \fi        & 400 & 0.05 & 7   & 288.8  \\
-6 & \ifen Uranus University    \else  천왕성대     \fi        & 450 & 0.03 & 8   & 294.1  \\
-7 & \ifen Neptune University   \else  해왕성대     \fi        & 500 & 0.10 & 5   & 257.7  \\
-8 & \ifen Pluto College        \else  명왕성대       \fi        & 550 & 0.12 & 3   & 195.1              
+1 & \ifen Mercury University   \else  수성대  \fi   & 0.39   & 200 & 4   & 230.0   \\
+2 & \ifen Venus University     \else  금성대  \fi   & 0.33   & 250 & 2   & 146.7  \\
+3 & \ifen Mars University      \else  화성대  \fi   & 0.24   & 300 & 6   & 281.5  \\
+4 & \ifen Jupiter University   \else  목성대  \fi   & 0.24   & 350 & 1   & \phantom{0}84.0  \\
+5 & \ifen Saturn University    \else  토성대  \fi   & 0.05   & 400 & 7   & 288.8  \\
+6 & \ifen Uranus University    \else  천왕성대 \fi   & 0.03   & 450 & 8   & 294.1  \\
+7 & \ifen Neptune University   \else  해왕성대 \fi   & 0.10   & 500 & 5   & 257.7  \\
+8 & \ifen Pluto College        \else  명왕성대 \fi   & 0.12   & 550 & 3   & 195.1      
 \end{tabular}
@@
 \ifen
-In this section, we show that this problem is NP-complete, then provide three algorithmic solutions: an exact branch-and-bound routine, an exact dynamic program, and a fully polynomial-time approximation scheme (FPTAS).
-\else
-본 절에서 이 문제가 NP-complete함을 보이고 3가지 해법을 제시한다: 정확한 분지한계법 해법, 정확한 동적 계획 해법, 그리고 완전 다항 시간 근사 해법 (fully polynomial-time approximation scheme, FPTAS).
+\noindent In this section, we show that this problem is NP-complete, then provide three algorithmic solutions: an exact branch-and-bound routine, an exact dynamic program, and a fully polynomial-time approximation scheme (FPTAS).
+\else
+\noindent 본 절에서 이 문제가 NP-complete함을 보이고 3가지 해법을 제시한다: 정확한 분지한계법 해법, 정확한 동적 계획 해법, 그리고 완전 다항 시간 근사 해법 (fully polynomial-time approximation scheme, FPTAS).
 \fi
@@
 \end{example} 
-% as the following example suggests ...The construction above shows that the objective function of Ellis's problem can approximate that of a knapsack problem with arbitrary precision, and this algorithm
+\ifen
+We might expect to obtain a better approximate solution to Ellis's problem by solving the following knapsack problem (nevermind that it is NP-complete) as a surrogate.
+\begin{equation}
+\text{maximize}\quad \sum_{j \in \mathcal{X}} f_j t_j \qquad \text{subject to}\quad\mathcal{X}\subseteq\mathcal{C},~~\sum_{j \in \mathcal{X}} g_j \leq H
+\end{equation}
+However, upon inspection, this solution merely generalizes the na\"ive algorithm for Alma's problem (Definition \ref{naivealgorithm}).
+In Ellis's problem, because the number of schools in the optimal portfolio can be as large as $m-1$, the knapsack solution's approximation coefficient is $1/(m-1)$ by analogy with \eqref{oneoverhopt}.
+A tight example is as follows.
+\else
+엘리스 문제의 좀 더 좋은 근사해를 찾기 위해 다음 배낭 문제를 (NP-complete함을 무시하고) 대리 문제로 푸는 것에 매력은 있다.
+\begin{equation}
+\text{maximize}\quad \sum_{j \in \mathcal{X}} f_j t_j \qquad \text{subject to}\quad\mathcal{X}\subseteq\mathcal{C},~~\sum_{j \in \mathcal{X}} g_j \leq H
+\end{equation}
+그러나 이러한 해법은 사실 알마 문제의 나이브 해법(정의 \ref{naivealgorithm})의 단순한 일반화임을 알 수 있다. 엘리스 문제에서 최적 포트폴리오에 크기가 $m-1$이 될 수 있으므로 배낭 문제 해법의 근사 계수는 \eqref{oneoverhopt}과(와) 유사하듯이 $1 / (m-1)$이다. 타이트한 예는 다음과 같다.
+\fi
+\begin{example}
+\ifen Consider the following instance of Ellis's problem, where $H = m-1$.
+\else 엘리스 문제의 다음 인스턴스를 고려하자. 단, $H = m-1$이다. \fi
+\begin{center}
+\begin{tabular}{r|cccc}
+$j$   & $1$            & $\cdots$ & $m-1$          & $m$            \\ \hline
+$f_j$ & $1$            &  $\cdots$ & $1$            & $\frac{1}{m-1}$ \\
+$t_j$ & $\frac{1}{m-1}$ &  $\cdots$ & $\frac{1}{m-1}$ & $m-1$          \\
+$g_j$ & $1$            &  $\cdots$ & $1$            & $m-1$         
+\end{tabular}%
+\end{center}
+\ifen
+The feasible solutions $\mathcal{Y} = \{1, \dots, m-1\}$ and $\mathcal{X} = \{ m\}$ each have $\sum_{j \in \mathcal{Y}} f_j t_j = \sum_{j \in \mathcal{X}} f_j t_j =  1$, and thus the knapsack algorithm can choose $\mathcal{Y}$ with $v(\mathcal{Y}) = 1/(m-1)$.
+But the optimal solution is  $\mathcal{X}$ with $v(\mathcal{X}) =  1$.
+\else
+가능해 $\mathcal{Y} = \{1, \dots, m-1\}$와 $\mathcal{X} = \{ m\}$에 대해 $\sum_{j \in \mathcal{Y}} f_j t_j = \sum_{j \in \mathcal{X}} f_j t_j =  1$이므로 배낭 문제 해법은  $\mathcal{Y}$를 선택할 수 있다.
+그의 가치는 $v(\mathcal{Y}) = 1/(m-1)$이다.
+그러나 $\mathcal{X}$는 최적해이며 그의 가치는 $v(\mathcal{X}) =  1$이다.
+\fi
+\end{example}
+
+\ifen 
+\noindent In summary, neither a greedy algorithm nor a knapsack-based algorithm solves Ellis's problem to optimality. 
+\else
+\noindent 탐욕 해법 또는 배낭 문제 기반 해법으로 엘리스 문제의 정확한 최적해를 구할 수 없음을 의미한다.
+\fi
 
@@
 \ifen
-Since the product in $v(x)$ does not exceed one, the following LP relaxation is an upper bound on the valuation of the optimal portfolio.
-\else
-$v(x)$에서 등장하는 곱은 1을 넘을 수 없으므로 다음 선형 완화 문제는 최적 포트폴리오 가치의 상한이 된다.
+\noindent Since the product in $v(x)$ does not exceed one, the following LP relaxation is an upper bound on the valuation of the optimal portfolio.
+\else
+\noindent $v(x)$에서 등장하는 곱은 1을 넘을 수 없으므로 다음 선형 완화 문제는 최적 포트폴리오 가치의 상한이 된다.
 \fi
@@
 \ifen
-Problem \ref{LPrelaxation} is a continuous knapsack problem, which is easily solved in $O(m \log m)$-time by the following greedy algorithm: Start with $x = \mathbf{0}$.
-While $H > 0$, select the school $j$ for which $f_j t_j / g_j$ is highest, set $x_j \gets \min\{1, H / g_j\}$, and set $H \gets H - g_j$ (Dantzig 1957).
+\noindent Problem \ref{LPrelaxation} is a continuous knapsack problem, which is easily solved in $O(m \log m)$-time by a greedy algorithm (Dantzig 1957).
 Balas and Zemel (1980, \S\,2) provide an $O(m)$ algorithm.
 \else
-문제 \ref{LPrelaxation}은(는) ``연속 배낭 문제''라고 불리며 다음 탐욕 알고리즘을 적용하면 $O(m \log m)$-시간 안에 쉽게 풀 수 있다: $x = \mathbf{0}$로 초기화한다.
-$H > 0$임이 성립하는 동안, $f_j t_j / g_j$-값이 가장 큰 학교 $j$를 택한다.
-$x_j \gets \min\{1, H / g_j\}$ 그리고 $H \gets H - g_j$로 수정하여 반복한다 (Dantzig 1957).
-$O(m)$ 알고리즘도 알려져 있다 (Balas와 Zemel 1980, \S\,2).
-\fi
-
-\ifen 
-In our branch-and-bound framework, a \emph{node} is characterized by a three-way partition of schools $\mathcal{C}= \mathcal{I} \cup \mathcal{O} \cup \mathcal{N}$ satisfying $\sum_{j \in \mathcal{I}} g_j \leq H$.
+\noindent 문제 \ref{LPrelaxation}은(는) ``연속 배낭 문제''라고 불리며 탐욕 알고리즘을 적용하면 $O(m \log m)$-시간 안에 쉽게 풀 수 있다 (Dantzig 1957).  $O(m)$ 알고리즘도 알려져 있다 (Balas와 Zemel 1980, \S\,2).
+\fi
+
+\ifen In our branch-and-bound framework, a \emph{node} is characterized by a three-way partition of schools $\mathcal{C}= \mathcal{I} \cup \mathcal{O} \cup \mathcal{N}$ satisfying $\sum_{j \in \mathcal{I}} g_j \leq H$.
 $\mathcal{I}$ consists of schools that are ``in'' the application portfolio ($x_j = 1$), $\mathcal{O}$ consists of those that are ``out'' ($x_j = 0$), and $\mathcal{N}$ consists of those that are ``negotiable.''
@@
 As with the knapsack problem, Ellis's problem admits a complementary dynamic program that iterates on the value of the cheapest portfolio instead of on the cost of the most valuable portfolio.
-We will use this algorithm as the basis for an FPTAS for Ellis's problem that uses $O(m^3 / \varepsilon)$-time and -space.
-Here we assume, with a small loss of generality, that each $t_j$ is a natural number.
+We will use this algorithm as the basis for an FPTAS for Ellis's problem that uses $O(m^3 / \varepsilon)$-time and -space.% Here we assume, with a small loss of generality, that each $t_j$ is a natural number.
 %Now, we modify Algorithm \ref{ellisDP2} into a fully polynomial-time approximation scheme for Ellis's problem.
 \else
-엘리스의 문제는 배낭 문제와 같이, 가치가 가장 높은 포트폴리오의 비용 대신 비용이 가장 낮은 포트폴리오의 가치를 탐색하는 보완적인 동적 계획이 존재한다. 이를 기반으로  $O(m^3 / \varepsilon)$-시간과 -공간을 사용하는 엘리스 문제의 FPTAS를 도출할 수 있다. 여기서 일반성을 약간 제한하여 각 $t_j$가 자연수임을 가정한다.
+엘리스의 문제는 배낭 문제와 같이, 가치가 가장 높은 포트폴리오의 비용 대신 비용이 가장 낮은 포트폴리오의 가치를 탐색하는 보완적인 동적 계획이 존재한다. 이를 기반으로  $O(m^3 / \varepsilon)$-시간과 -공간을 사용하는 엘리스 문제의 FPTAS를 도출할 수 있다.% 여기서 일반성을 약간 제한하여 각 $t_j$가 자연수임을 가정한다.
 \fi
@@
 \xrowht[()]{1.5em}  Costs DP     &  Algorithm \ref{ellisDP1}& General          & $g_j$ integer         & Exact                    & $O(Hm + m \log m)$        \\
-\xrowht[()]{1.5em}  FPTAS      &    Algorithm \ref{ellisDP3} & General          & $t_j$ integer         & $(1 - \varepsilon)$-opt. & $O(m^3 / \varepsilon)$   
+\xrowht[()]{1.5em}  FPTAS      &    Algorithm \ref{ellisDP3} & General          &  None         & $(1 - \varepsilon)$-opt. & $O(m^3 / \varepsilon)$   
 \end{tabular}%
@@
 \xrowht[()]{1.7em}   \begin{tabular}[r]{@{}r@{}}지출액\\ 동적 계획\end{tabular}     &  알고리즘 \ref{ellisDP1}& 일반 문제          & $g_j$ 정수         & 정확                    & $O(Hm + m \log m)$        \\
-\xrowht[()]{1.7em}  FPTAS      &    알고리즘 \ref{ellisDP3} & 일반 문제          & $t_j$ 정수         & $(1 - \varepsilon)$-근사 & $O(m^3 / \varepsilon)$   
+\xrowht[()]{1.7em}  FPTAS      &    알고리즘 \ref{ellisDP3} & 일반 문제          &  없음    & $(1 - \varepsilon)$-근사 & $O(m^3 / \varepsilon)$   
 \end{tabular}%

### 2022-04-12 paper/paper.tex
loosen input restrictions and allow some edge cases

@@
 둘째는 의사 다항 시간 안에 정확한 해를 출력하는 동적 계획 해법이다.
-마지막으로 다른 동적 계획 기반으로 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme)이 존재함을 보인다.
+마지막으로 다른 동적 계획 기반으로 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme)을 도출한다.
 수리적 실험을 통하여 알고리즘의 현실성을 보인다.
@@
 \ifen
-From a methodological standpoint, as Chao (2014) remarked, college application represents a somewhat subtle asset allocation problem.
+From a methodological standpoint, as Fu (2014) remarked, college application represents a somewhat subtle asset allocation problem.
 In computational finance, traditional portfolio optimization models weigh the sum of expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Markowitz 1952; Meucci 2005).
@@
 Finally, the combinatorial nature of the college application problem makes it difficult to solve using the gradient-based techniques used in continuous portfolio optimization.
-Chao estimated her equilibrium model (which considers application as a \emph{cost} rather than a constraint) by clustering the schools so that $m=8$, a scale at which enumeration is tractable.
+Fu estimated her equilibrium model (which considers application as a \emph{cost} rather than a constraint) by clustering the schools so that $m=8$, a scale at which enumeration is tractable.
 We pursue a more general solution.
 \else
-방법론적인 관점에서 보면, Chao (2014)가 주장한 바처럼 대학 지원 전략은 약간 미묘한 자산 배분 문제이다.
+방법론적인 관점에서 보면, Fu (2014)가 주장한 바처럼 대학 지원 전략은 약간 미묘한 자산 배분 문제이다.
 재정학 분야에서 고전적 포트폴리오 최적화 모형은 전체 자산에 대한 기대 총이익에서 위험회피 항을 뺌으로 선형식으로 제약된 오목 최대화 문제를 이룬다 (Markowitz 1952; Meucci 2005).
@@
 마지막, 대학 지원의 조합적인 본성으로 인해 연속적인 포트폴리오 최적화 문제에서 흔히 사용하는 기울기 해법으로는 풀기가 어렵다.
-Chao는 지원 비용을 제약 조건 대신 목적함수의 한 항으로 모형화했으며, 균형 모형의 모수를 추정하기 위해 $m=8$이 되도록 학교들의 클러스터를 먼저 구성했다.
+Fu는 지원 비용을 제약 조건 대신 목적함수의 한 항으로 모형화했으며, 균형 모형의 모수를 추정하기 위해 $m=8$이 되도록 학교들의 클러스터를 먼저 구성했다.
 이는 열거법을 통해 문 쉽게 풀 수 있는 규모이지만 본 연구는 더 일반적인 $m$에 대한 해법을 추구한다.
@@
 \fi
-\begin{proposition}[\ifen Eliminate $c_k$\else $c_k$ 소거\fi]
-\ifen For $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$, where
-\else $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$에 대해, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$이다. 단,\fi
+%\begin{proposition}[\ifen Eliminate $c_k$\else $c_k$ 소거\fi]
+%\ifen For $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$, where
+%\else $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$에 대해, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$이다. 단,\fi
+%\begin{align}\label{howtotransformfj}
+%\bar t_j = 
+%\begin{cases}
+%(1 - f_k) t_j, \quad & t_j \leq t_k \\
+%t_j - f_k t_k, \quad& t_j > t_k.
+%\end{cases}
+%\end{align}
+%\end{proposition}%
 \begin{align}\label{howtotransformfj}
@@
 (1 - f_k) t_j, \quad & t_j \leq t_k \\
-t_j - f_k t_k, \quad& t_j > t_k.
+t_j - f_k t_k, \quad& t_j > t_k
 \end{cases}
 \end{align}
-\end{proposition}%
 \ifen

### 2022-04-12 src/MarketIO.jl

@@ -5,6 +5,7 @@ const UIntTypes = DataType[UInt8, UInt16, UInt32, UInt64]
 function iscoherentmarket(f::Vector{Float64}, t::Vector{<:Real})
     @assert length(f) == length(t)
     @assert all(0 .< f .≤ 1)
+    @assert all(0 .≤ t)
 end
 
 
@@ -12,19 +13,8 @@ function iscoherentmarket(f::Vector{Float64}, t::Vector{Int}, g::Vector{Int})
     @assert length(f) == length(t)
     @assert length(f) == length(g)
     @assert all(0 .< f .≤ 1)
-end
-
-
-function isnontrivialmarket(f::Vector{Float64}, t::Vector{<:Real}, h::Int)
-    iscoherentmarket(f, t)
-    @assert 0 < h ≤ length(t)
-end
-
-
-function isnontrivialmarket(f::Vector{Float64}, t::Vector{Int}, g::Vector{Int}, H::Int)
-    iscoherentmarket(f, t, g)
-    @assert 0 < H ≤ sum(g)
-    @assert all(0 .< g .≤ H)
+    @assert all(0 .≤ t)
+    @assert all(0 .≤ g)
 end
 
 
@@ -57,7 +47,7 @@ struct SameCostsMarket{T<:Unsigned, U<:Real}
     and return the market object.
     """
     function SameCostsMarket(f::Vector{Float64}, t::Vector{U}, h::Integer) where U<:Real
-        isnontrivialmarket(f, t, h)
+        iscoherentmarket(f, t)
         T = UIntTypes[findfirst(T -> length(f) < typemax(T), UIntTypes)]
 
         # We are current asserting that t is sorted; a placeholder for later work
@@ -67,7 +57,7 @@ struct SameCostsMarket{T<:Unsigned, U<:Real}
             T(length(f)),
             f[perm],
             t[perm],
-            T(h),
+            T(min(h, length(f))),
             (f .* t)[perm],
             (1 .- f)[perm],
             perm)
@@ -123,7 +113,7 @@ struct VariedCostsMarket{T<:Unsigned}
     and application budget `H` and return the market object.
     """
     function VariedCostsMarket(f::Vector{Float64}, t::Vector{Int}, g::Vector{Int}, H::Int)
-        isnontrivialmarket(f, t, g, H)
+        iscoherentmarket(f, t, g)
         T = UIntTypes[findfirst(T -> length(f) < typemax(T), UIntTypes)]
[…]

### 2022-04-12 test/runtests.jl

@@ -9,37 +9,39 @@ const n_markets = 3
 
 @testset verbose = true "OptimalApplication.jl" begin
 
-    @testset verbose = true "Same app. costs, t int" begin
-        m = 20
+    @testset verbose = true "Same app. costs" begin
+        @testset verbose = true "t int" begin
+            m = 20
 
-        for _ in 1:n_markets
-            mkt = SameCostsMarket(m)
+            for _ in 1:n_markets
+                mkt = SameCostsMarket(m)
 
-            X_enum, vX_enum = optimalportfolio_enumerate(mkt)
-            sort!(X_enum)
-            X_heap, vX_heap = applicationorder_heap(mkt)
-            X_list, vX_list = applicationorder_list(mkt)
-            @test X_enum == sort(X_heap)
-            @test vX_enum ≈ last(vX_heap)
-            @test X_enum == sort(X_list)
-            @test vX_enum ≈ last(vX_list)
+                X_enum, vX_enum = optimalportfolio_enumerate(mkt)
+                sort!(X_enum)
+                X_heap, vX_heap = applicationorder_heap(mkt)
+                X_list, vX_list = applicationorder_list(mkt)
+                @test X_enum == sort(X_heap)
+                @test vX_enum ≈ last(vX_heap)
+                @test X_enum == sort(X_list)
+                @test vX_enum ≈ last(vX_list)
+            end
         end
-    end
 
-    @testset verbose = true "Same app. costs, t float" begin
-        m = 20
+        @testset verbose = true "t float" begin
+            m = 20
 
-        for _ in 1:n_markets
-            mkt = SameCostsMarket(rand(16), sort(12 * rand(16)), m ÷ 4)
+            for _ in 1:n_markets
+                mkt = SameCostsMarket(rand(16), sort(12 * rand(16)), m ÷ 4)
 
-            X_enum, vX_enum = optimalportfolio_enumerate(mkt)
-            sort!(X_enum)
-            X_heap, vX_heap = applicationorder_heap(mkt)
-            X_list, vX_list = applicationorder_list(mkt)
-            @test X_enum == sort(X_heap)
-            @test vX_enum ≈ last(vX_heap)
-            @test X_enum == sort(X_list)
-            @test vX_enum ≈ last(vX_list)
+                X_enum, vX_enum = optimalportfolio_enumerate(mkt)
+                sort!(X_enum)
[…]

### 2022-04-14 src/EnumerationSolvers.jl
Add greedy algorithm and simulated annealing Use oneunit(T) instead of T(1) et sim Replace constant UIntTypes vector with function of same purpose

@@ -1,5 +1,5 @@
 """
-    optimalportfolio_enumerate(mkt::SameCostsMarket)
+    optimalportfolio_enumerate(mkt::SameCostsMarket) -> X, v
 
 Produce the optimal portfolio for the market `mkt` having identical application costs.
 """
@@ -10,7 +10,7 @@ function optimalportfolio_enumerate(mkt::SameCostsMarket{T,U})::Tuple{Vector{Int
     v = 0.0
 
     invp = invperm(mkt.perm)
-    for Y in multiset_combinations(T(1):T(mkt.m), Int(mkt.h))
+    for Y in multiset_combinations(oneunit(T):mkt.m, Int(mkt.h))
         if (w = valuation(Y, mkt; invp=invp)) > v
             v = w
             X[:] = Y
@@ -22,7 +22,7 @@ end
 
 
 """
-    optimalportfolio_enumerate(mkt::VariedCostsMarket)
+    optimalportfolio_enumerate(mkt::VariedCostsMarket) -> X, v
 
 Produce the optimal portfolio for the market `mkt` having varying application costs.
 """
@@ -34,7 +34,7 @@ function optimalportfolio_enumerate(mkt::VariedCostsMarket{T})::Tuple{Vector{Int
         v = 0.0
 
         invp = invperm(mkt.perm)
-        for Y in combinations(T(1):mkt.m)
+        for Y in combinations(oneunit(T):mkt.m)
             if (w = valuation(Y, mkt, invp=invp)) > v && sum(mkt.g[invp[Y]]) ≤ mkt.H
                 v = w
                 X = copy(Y)

### 2022-04-14 src/FastSolvers.jl

@@ -25,7 +25,7 @@ isless(c1::College, c2::College) = isless(c1.ft, c2.ft)
 
 
 """
-    applicationorder_list(mkt::SameCostsMarket)
+    applicationorder_list(mkt::SameCostsMarket) -> X, V
 
 Produce the optimal order of application for the market `mkt` having identical
 application costs and the corresponding portfolio valuations.
@@ -37,15 +37,15 @@ function applicationorder_list(
     apporder = zeros(T, mkt.h)
     v = zeros(mkt.h)
 
-    mkt_list = College{T}[College(T(j), mkt.f[j], mkt.t[j], mkt.ft[j], mkt.omf[j]) for j in one(T):mkt.m]
+    mkt_list = College{T}[College(j, mkt.f[j], mkt.t[j], mkt.ft[j], mkt.omf[j]) for j in oneunit(T):mkt.m]
 
-    dummy_college = College(T(0), 1.0, -1.0, -1.0, 0.0)
+    dummy_college = College(zero(T), 1.0, -1.0, -1.0, 0.0)
 
     c_best::College{T}, idx_best::Int = findmax(mkt_list)
-    @inbounds for j in one(T):mkt.h
+    @inbounds for j in oneunit(T):mkt.h
         if verbose
             println("Iteration $j")
-            println("  ft: ", [mkt_list[i].ft for i in one(T):length(mkt_list)])
+            println("  ft: ", [mkt_list[i].ft for i in oneunit(T):length(mkt_list)])
             println("  Add school $(c_best.j)")
         end
 
@@ -97,7 +97,7 @@ end
 
 
 """
-    applicationorder_heap(mkt::SameCostsMarket)
+    applicationorder_heap(mkt::SameCostsMarket) -> X, V
 
 Produce the optimal order of application for the market `mkt` having identical
 application costs and the corresponding portfolio valuations.
@@ -106,9 +106,9 @@ function applicationorder_heap(mkt::SameCostsMarket{T})::Tuple{Vector{Int},Vecto
     apporder = zeros(T, mkt.h)
     v = zeros(mkt.h)
 
-    mkt_heap = BinaryMaxHeap{College{T}}(collect(College(T(j), mkt.f[j], mkt.t[j], mkt.ft[j], mkt.omf[j]) for j in one(T):mkt.m))
+    mkt_heap = BinaryMaxHeap{College{T}}(collect(College(j, mkt.f[j], mkt.t[j], mkt.ft[j], mkt.omf[j]) for j in oneunit(T):mkt.m))
 
[…]

### 2022-04-14 src/HeuristicSolvers.jl

@@ -1,0 +1,141 @@
+"""
+    optimalportfolio_greedy(mkt::VariedCostsMarket) -> X, v
+
+Use a greedy heuristic that adds schools in decreasing order of `mkt.ft ./ mkt.g`
+to compute an approximately optimal portfolio for the market `mkt` with varying application costs.
+"""
+function optimalportfolio_greedy(mkt::VariedCostsMarket{T})::Tuple{Vector{Int},Float64} where {T}
+    priority_order = filter(
+        sortperm(mkt.ft ./ mkt.g, rev=true)
+    ) do j
+        mkt.g[j] ≤ mkt.H
+    end
+
+    X = T[]
+    H = mkt.H
+    for j in priority_order
+        if mkt.g[j] ≤ H
+            push!(X, j)
+            H -= mkt.g[j]
+        end
+
+        if H == 0
+            @goto full
+        end
+    end
+    @label full
+
+    # In the valuation we just use the identity permutation as invp
+    # To prevent wasteful permuting and then invpermuting
+    return mkt.perm[X], valuation(X, mkt; invp=oneunit(T):mkt.m)
+end
+
+
+"""
+    neighbor(X, mkt) -> X_neighbor
+
+Generate a random neighbor of `X` for the market `mkt`. If `X` is feasible,
+preserve feasibility; if not, move to a feasible solution. `X` is assumed
+to refer to the sorted indices of the market. 
+"""
+function neighbor(X::AbstractVector{T}, mkt::VariedCostsMarket{T})::Vector{T} where {T}
+    X_neighbor = copy(X)
+    Y = setdiff(oneunit(T):mkt.m, X)
+    shuffle!(X_neighbor)
+    shuffle!(Y)
+
+    H = sum(mkt.g[X])
+
+    while H ≤ mkt.H && !isempty(Y)
+        j = pop!(Y)
+        push!(X_neighbor, j)
+        H += mkt.g[j]
+    end
+
+    while H > mkt.H
+        j = popfirst!(X_neighbor)
+        H -= mkt.g[j]
+    end
+
+    return X_neighbor
+end
+
+"""
+    optimalportfolio_simulatedannealing(
+        mkt::VariedCostsMarket;
+        X0::Union{Nothing,Vector{T}}=nothing,
+        temp::Union{Nothing,Float64}=nothing,
+        nit::Integer=500,
+        red::Float64=0.95,
+        verbose::Bool=false
+    ) -> X, v
+
[…]

### 2022-04-14 src/MarketIO.jl

@@ -1,4 +1,17 @@
-const UIntTypes = DataType[UInt8, UInt16, UInt32, UInt64]
+"""
+    smallestunsigned(k::Integer)
+
+Identify the smallest unsigned integer type for which `k < typemax(T)`
+"""
+function smallestunsignedtype(k::Integer)::DataType
+    UIntTypes = DataType[UInt8, UInt16, UInt32, UInt64]
+
+    j = findfirst(UIntTypes) do T
+        k < typemax(T)
+    end
+
+    return UIntTypes[j]
+end
 
 
 # TO DO: Use Julian errors rather than assertions here.
@@ -30,7 +43,7 @@ application costs. Fields:
     `omf = 1 .- f`
     `perm = sortperm(perm)`: Currently a placeholder
 """
-struct SameCostsMarket{T<:Unsigned, U<:Real}
+struct SameCostsMarket{T<:Unsigned,U<:Real}
     m::T
     f::Vector{Float64}
     t::Vector{U}
@@ -46,25 +59,23 @@ struct SameCostsMarket{T<:Unsigned, U<:Real}
     admissions probabilities `f`, utility values `t`, and application limit `h`
     and return the market object.
     """
-    function SameCostsMarket(f::Vector{Float64}, t::Vector{U}, h::Integer) where U<:Real
+    function SameCostsMarket(f::Vector{Float64}, t::Vector{U}, h::Integer) where {U<:Real}
         iscoherentmarket(f, t)
-        T = UIntTypes[findfirst(T -> length(f) < typemax(T), UIntTypes)]
-
-        # We are current asserting that t is sorted; a placeholder for later work
+        T = smallestunsignedtype(length(f))
         perm = T.(sortperm(t))
 
-        return new{T, U}(
+        return new{T,U}(
             T(length(f)),
             f[perm],
             t[perm],
             T(min(h, length(f))),
-            (f .* t)[perm],
+            (f.*t)[perm],
             (1 .- f)[perm],
             perm)
     end
 
     function SameCostsMarket(m::Integer)
-        T = UIntTypes[findfirst(T -> m < typemax(T), UIntTypes)]
+        T = smallestunsignedtype(m)
         t = ceil.(Int, -10 * log.(rand(m)))
         f = inv.(t .+ 10 * rand(m))
         perm = T.(sortperm(t))
[…]

### 2022-04-14 src/OptimalApplication.jl

@@ -1,5 +1,6 @@
 module OptimalApplication
 
+using Random
 using Combinatorics: combinations, multiset_combinations
 using DataStructures
 import Base.isless
@@ -17,13 +18,16 @@ export
     optimalportfolio_enumerate,
     optimalportfolio_valuationtable,
     optimalportfolio_fptas,
-    optimalportfolio_branchbound
+    optimalportfolio_branchbound,
+    optimalportfolio_greedy,
+    optimalportfolio_simulatedannealing
 
 
 include("MarketIO.jl")
 include("EnumerationSolvers.jl")
 include("FastSolvers.jl")
 include("VariedCostsBnB.jl")
+include("HeuristicSolvers.jl")
 
 
 end

### 2022-04-14 src/VariedCostsBnB.jl

@@ -117,7 +117,7 @@ end
 
 
 """
-    optimalportfolio_branchbound(mkt::VariedCostsMarket; maxit=10000, verbose=false)
+    optimalportfolio_branchbound(mkt::VariedCostsMarket; maxit=10000, verbose=false) -> X, v
 
 Use the branch-and-bound algorithm to produce the optimal portfolio for the
 market `mkt` with varying application costs. Intractable for large markets. 
@@ -125,7 +125,7 @@ market `mkt` with varying application costs. Intractable for large markets.
 function optimalportfolio_branchbound(mkt::VariedCostsMarket{T}; maxit::Int=100000, verbose::Bool=false)::Tuple{Vector{Int},Float64} where {T}
     mkt.m ≥ 33 && @warn "Branch and bound is slow for large markets"
 
-    C = collect(one(T):mkt.m)
+    C = collect(oneunit(T):mkt.m)
 
     rootnode::Node{T} = Node(T[], C, Dict(zip(1:mkt.m, Float64.(mkt.t))), mkt.H, 0.0, mkt)
     LB::Float64 = 0.0

### 2022-04-14 test/runtests.jl

@@ -31,7 +31,7 @@ const n_markets = 3
             m = 20
 
             for _ in 1:n_markets
-                mkt = SameCostsMarket(rand(16), sort(12 * rand(16)), m ÷ 4)
+                mkt = SameCostsMarket(rand(16), 12 * rand(16), m ÷ 4)
 
                 X_enum, vX_enum = optimalportfolio_enumerate(mkt)
                 sort!(X_enum)
@@ -171,14 +171,16 @@ const n_markets = 3
         end
 
         @testset verbose = true "Het. costs" begin
-            # H > sum(g) should have all schools
-            mkt = VariedCostsMarket(fill(0.5, 4), fill(2, 4), fill(4, 4), 20)
+            # H ≥ sum(g) should have all schools
+            mkt = VariedCostsMarket(fill(0.5, 4), fill(2, 4), fill(4, 4), 16)
 
             @test 1:4 == sort(optimalportfolio_branchbound(mkt)[1])
             @test 1:4 == sort(optimalportfolio_valuationtable(mkt)[1])
             @test 1:4 == sort(optimalportfolio_dynamicprogram(mkt)[1])
             @test 1:4 == sort(optimalportfolio_fptas(mkt, 0.05)[1])
             @test 1:4 == sort(optimalportfolio_enumerate(mkt)[1])
+            @test 1:4 == sort(optimalportfolio_greedy(mkt)[1])
+            @test 1:4 == sort(optimalportfolio_simulatedannealing(mkt; nit=4)[1])
 
             # g[j] > H should be excluded
             mkt = VariedCostsMarket([0.5, 0.5, 1.0], [1, 2, 100], [3, 3, 5], 4)
@@ -188,7 +190,27 @@ const n_markets = 3
             @test [2] == optimalportfolio_dynamicprogram(mkt)[1]
             @test [2] == optimalportfolio_fptas(mkt, 0.25)[1]
             @test [2] == optimalportfolio_enumerate(mkt)[1]
+            @test [2] == sort(optimalportfolio_greedy(mkt)[1])
+            @test [2] == sort(optimalportfolio_simulatedannealing(mkt; nit=4)[1])
         end
+    end
+
+    @testset verbose = true "Heuristic algorithms" begin
+        mkt = VariedCostsMarket([0.34, 0.8, 0.1, 0.5], [9, 2, 2, 4], [3, 4, 4, 3], 7)
+        X, v = optimalportfolio_greedy(mkt)
[…]

### 2022-04-15 paper/paper.tex
Add greedy algorithm and simulated annealing Use oneunit(T) instead of T(1) et sim Replace constant UIntTypes vector with function of same purpose
small improvement to binary search add plot for simulated annealing, begin writeup

@@
 \ifen
-\newcommand{\thetitle}{How to apply to college}
+\newcommand{\thetitle}{The College Application Problem}
 \newcommand{\therunningtitle}{\thetitle}
@@
 \fi
-\begin{example}
+\begin{example} \label{knapsackheuristicellis}
 \ifen Consider the following instance of Ellis's problem, where $H = m-1$.
@@
 In our numerical experiments, Algorithm \ref{ellisbnb} was ineffective on instances larger than about $m=35$ schools, and therefore it does not represent a significant improvement over na\"ive enumeration.
+We speculate that the algorithm can be improved by incorporating cover inequalities (Wolsey 1998, \S\,9.3).
 \else
@@
 수리 실험에서, 알고리즘 \ref{ellisbnb}은(는) 약 $m=35$개의 학교를 넘는 인스턴스에 대해 비효율적이었으며 이는 단순한 열거법에 비해 큰 개선이 아니다.
+하지만, 커버 부등식을 도입하여 해법을 개선할 가능성이 보인다 (Wolsey 1998, \S\,9.3).
 \fi
@@
 \caption{Fully polynomial-time approximation scheme for Ellis's problem.} \label{ellisDP3}
-\KwData{Utility values $t \in \mathbb{N}^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$, tolerance $\varepsilon \in (0, 1)$.}
+\KwData{Utility values $t \in  (0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$, tolerance $\varepsilon \in (0, 1)$.}
 Index schools in ascending order by $t$\;
@@
 \caption{엘리스의 문제를 위한 완전 다항 시간 근사 해법.} \label{ellisDP3}
-\KwData{효용 모수 $t \in \mathbb{N}^m$, 합격 확률 $f \in (0, 1]^m$, 지원 비용 $g \in (0, \infty)^m$, 예산 $H \in (0, \infty)^m$, 허용치 $\varepsilon \in (0, 1)$.}
+\KwData{효용 모수 $t \in  (0, \infty)}^m$, 합격 확률 $f \in (0, 1]^m$, 지원 비용 $g \in (0, \infty)^m$, 예산 $H \in (0, \infty)^m$, 허용치 $\varepsilon \in (0, 1)$.}
 $t$의 순서대로 학교를 배열한다\;
@@
 
+\ifen \subsection{Simulated annealing} \else \subsection{Simulated annealing}\fi\label{numericalexperiments}
+Simulated annealing is a popular local-search technique for nonconvex optimization problems.
+Because of their inherent randomness, simulated-annealing algorithms offer no accuracy guarantee; however, they are easy to implement, computationally inexpensive, and often produce solutions with a small optimality gap.
+In this section, we present a simulated-annealing algorithm for Ellis's problem.
+It is based on the procedure for integer programs outlined by Wolsey (1998, \S\,12.3.2)
+
+To generate an initial solution, our algorithm adopts the greedy heuristic of adding schools in descending order by $f_j t_j / g_j$ until the budget is exhausted.
+Example \ref{knapsackheuristicellis} shows that the approximation coefficient of this solution can be as low as $1 / (m-1)$; in typical instances, it is much higher.
+Given a solution $\mathcal{X}$, we generate a neighbor $\mathcal{X}_n$ by copying $\mathcal{X}$, adding schools to $\mathcal{X}_n$ until it becomes infeasible, then removing elements of $\mathcal{X}$ from $\mathcal{X}_n$ until feasibility is restored.
+If this increases the valuation of $\mathcal{X}$, then the candidate solution is updated to equal the neighbor; if not, then the candidate solution is updated with probability $\exp\bigl(v(\mathcal{X}_n) - v(\mathcal{X}) / T\bigr)$, where $T$ is a temperature parameter.
+This is repeated $N$ times, and the algorithm returns the best $\mathcal{X}$ observed.
+
+\begin{algorithm}[h] 
+%\DontPrintSemicolon
+\caption{Simulated annealing for Ellis's problem.} \label{ellissimann}
+\KwData{Utility values $t  (0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$, initial temperature $T > 0$, temperature reduction parameter $r \in (0, 1)$, number of iterations $N$.}
+Add schools to $\mathcal{X}$ in descending order by $f_j t_j / g_j$ until the budget is exhausted\;
+\For{$i = 1 \dots N$}{
+	$\mathcal{X}_n \gets \operatorname{copy}(\mathcal{X})$\;
+	Add random schools from $\mathcal{C} \setminus \mathcal{X}$ to $\mathcal{X}_n$ until  $\mathcal{X}_n$ infeasible\;
+	Remove random schools in $\mathcal{X}$ from $\mathcal{X}_n$ until feasibility restored\;
+	$\Delta = v(\mathcal{X}_n) - v(\mathcal{X})$\;
+	\lIfElse{$\Delta \geq 0$}{$\mathcal{X} \gets \mathcal{X}_n$}
+	{$\mathcal{X} \gets \mathcal{X}_n$ with probability $\exp(\Delta / T)$}
+	$T \gets rT$\;
+}
+\Return{the best $\mathcal{X}$ found}
+\end{algorithm}
+
+\begin{proposition}
+The computation time of Algorithm \ref{ellissimann} is $O(Nm)$. 
+\end{proposition}
+
+In our numerical experiments, the output of Algorithm \ref{ellissimann} was typically quite close to optimal.
+
+%always within 10 percent of optimality.
+
 %\pagebreak
@@
 \ifen \subsection{Experimental procedure} \else \subsection{실험 방법}\fi
+We conducted three numerical experiments.
+Experiments 1 and 2 evaluate the computation times of our algorithms for Alma's problem and Ellis's problem, respectively.
+Experiment 3 investigates the empirical accuracy of the simulated-annealing heuristic as compared to an exact algorithm.
+
 \def\nmarkets{50}
 \ifen
-The experimental procedure was as follows.
-First, to generate synthetic markets, we drew the $t_j$-values independently from an exponential distribution with a scale parameter of ten and rounded up to the nearest integer.
+To generate synthetic markets, we drew the $t_j$-values independently from an exponential distribution with a scale parameter of ten and rounded up to the nearest integer.
 To achieve partial negative correlation between $t_j$ and $f_j$, we then set $f_j = 1 / (t_j + 10\,Q)$, where $Q$ is drawn uniformly from the interval $[0, 1)$.
-In the first experiment, which concerns Alma's problem, we set each $g_j = 1$ and $H = h = \lfloor m/ 2 \rfloor$.
-In the second experiment, which concerns Ellis's problem, each $g_j$ is drawn uniformly from the set $\{5, \dots, 10\}$ and we set $H = \lfloor \frac{1}{2} \sum g_j \rfloor$.
+In Experiment 1, which concerns Alma's problem, we set $h = \lfloor m/ 2 \rfloor$.
+In Experiments 2 and 3, which concern Ellis's problem, each $g_j$ is drawn uniformly from the set $\{5, \dots, 10\}$ and we set $H = \lfloor \frac{1}{2} \sum g_j \rfloor$.
 This cost distribution resembles real-world college application fees, which are often multiples of \$10 between \$50 and \$100.
@@
 \else
-실험 방법은 다음과 같다.
-먼저, 가산 시장을 생성하기 위해 $t_j$-값을 평균이 10인 지수 분포에서 독립적으로 관측하고 정수로 올림했다.
+가산 시장을 생성하기 위해 $t_j$-값을 평균이 10인 지수 분포에서 독립적으로 관측하고 정수로 올림했다.
 그다음에, $t_j$와 $f_j$가 서로 부분적으로 반비례하게 만들도록 $f_j = 1 / (t_j + 10\,Q)$로 정의했다.
 단, $Q$는 $[0, 1)$에서 균일한 확률로 관측한다.
-알마의 문제를 고려하는 첫 번째 실험에서, 모든 $g_j = 1$ 그리고 $H = h = \lfloor m/ 2 \rfloor$으로 처리했다.
-엘리스의 문제를 고려하는 두 번째 실험에서, 모든  $g_j$를 $\{5, \dots, 10\}$에서 균일한 확률로 관측했으며 $H = \lfloor \frac{1}{2} \sum g_j \rfloor$으로 처리했다.
+알마의 문제를 고려하는 실험 1에서, $h = \lfloor m/ 2 \rfloor$으로 처리했다.
+엘리스의 문제를 고려하는 실험 2와 실험 3에서, 모든  $g_j$를 $\{5, \dots, 10\}$에서 균일한 확률로 관측했으며 $H = \lfloor \frac{1}{2} \sum g_j \rfloor$으로 처리했다.
 실제 대학 지원 비용인 50과 100달러 사이에 10달러의 어떤 배수이므로 합리적인 비용 분포이다.
@@
 \ifen
-The experimental variables were the market size $m$, the choice of algorithm, and (for Algorithm \ref{ellisDP3}) the tolerance $\varepsilon$.
+The experimental variables in Experiments 1 and 2 were the market size $m$, the choice of algorithm, and (for Algorithm \ref{ellisDP3}) the tolerance $\varepsilon$.
 For each combination of the experimental variables, we generated \nmarkets~markets, and the optimal portfolio was computed three times, with the fastest of the three repetitions recorded as the computation time.
@@
 Where applicable, we do not count the time required to sort the entries of $t$.
-\else
-실험 변수로는 시장의 크기 $m$, 해법 선택, 그리고 (알고리즘 \ref{ellisDP3}인 경우) 허용치 $\varepsilon$을 고려했다.
+As the simulated annealing algorithm's computation time is negligible, it was excluded from Experiment 2.
+\else
+실험 1과 실험 2의 실험 변수로는 시장의 크기 $m$, 해법 선택, 그리고 (알고리즘 \ref{ellisDP3}인 경우) 허용치 $\varepsilon$을 고려했다.
 실험 변수의 각 조합에 대해,  \nmarkets 개의 시장을 생성했으며 최적 포트폴리오를 3번 계산 한 것 중 가장 빠른 것으로 계산 시간으로 기록했다.
@@
 $t$를 배열하는 시간은 고려하지 않았다.
-\fi
+As the simulated annealing algorithm's computation time is negligible, it was excluded from Experiment 2.
+\fi
+
+In Experiment 3, we generated 500 markets with sizes between $2^3$ and $2^11$ varying uniformly in the logarithm.
+For each market, we computed a heuristic solution using Algorithm \ref{ellissimann} and an exact solution using Algorithm \ref{ellisDP1} and calculated the optimality ratio.
 
@@
 \ifen
-In our first experiment, we compare the performance of Algorithm \ref{algorithmforlargeh} for homogeneous-cost markets of various sizes when the set of candidate schools $\mathcal{C}$ is stored as a list and as a binary max heap ordered by the $f_j \bar t_j$-values.
+Experiment 1 compared the performance of Algorithm \ref{algorithmforlargeh} for homogeneous-cost markets of various sizes when the set of candidate schools $\mathcal{C}$ is stored as a list and as a binary max heap ordered by the $f_j \bar t_j$-values.
 The results appear in Table \ref{experiment1results}.
@@
 \else
-첫 번째 실험에서, 동일한 지원 비용으로 정의된 크기가 다양한 시장에 대해 알고리즘 \ref{algorithmforlargeh}의 성능을 고려했다.
+실험 1에서, 동일한 지원 비용으로 정의된 크기가 다양한 시장에 대해 알고리즘 \ref{algorithmforlargeh}의 성능을 고려했다.
 이를 $\mathcal{C}$를 목록 혹은 $f_j \bar t_j$-값에 따라 배열된 힙 구현으로 구성하는 경우로 나눴으며 실험 결과는 표 \ref{experiment1results}에서 등장한다.
@@
 \ifen
-In the second experiment, we turn to the general problem, and compare the performance of the exact algorithms (Algorithms \ref{ellisbnb} and \ref{ellisDP1}) and the approximation scheme (Algorithm \ref{ellisDP3}) at tolerances 0.5 and 0.05.
+In Experiment 2, we turned to the general problem, and compared the performance of the exact algorithms (Algorithms \ref{ellisbnb} and \ref{ellisDP1}) and the approximation scheme (Algorithm \ref{ellisDP3}) at tolerances 0.5 and 0.05.
 The results, which appear in Table \ref{experiment2results}, broadly agree with the time complexity analyses presented above.
@@
 \else
-두 번째 실험에서 일반적인 문제를 고려한다.
+실험 2에서 일반적인 문제를 고려한다.
 정확한 해법 (알고리즘 \ref{ellisbnb} 및 알고리즘 \ref{ellisDP1}) 그리고 허용치를 0.5와 0.05로 설정한 근사 해법 (알고리즘 \ref{ellisDP3})의 성능을 비교한다.
@@
 
+Experiment 3 \dots. (WIP)
+
 \newcommand{\lastptofcaption}{\ifen
@@
 \else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복하여 그중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (평균편차) 시간이 나타난다.\fi}
+
+\begin{figure}[h!] 
+\centering
+\includegraphics[width=0.85\textwidth]{./plots/samplemarket.pdf}
+  \caption{\label{samplemarket}
+ \ifen A typical randomly-generated instance with $m=64$ schools and its optimal application portfolio.
+The application costs $g_j$ were drawn uniformly from $\{5, \dots, 10\}$.
+The optimal portfolio was computed using Algorithm \ref{ellisDP1}.
+ \else $m=64$개의 학교로 구성된 전형적인 무작위로 생성한 인스턴스와 해당 최적 포트폴리오. 지원 비용 $g_j$는 $\{5, \dots, 10\}$에서 균일한 확률로 선정했으며 기호의 넓이와 비례된다. 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
+\end{figure}
 
@@
 \caption{\label{experiment1results} \normalsize
-\ifen Time in ms to compute an optimal portfolio for an admissions market with homogeneous application costs using Algorithm \ref{algorithmforlargeh} when $\mathcal{C}$ is stored as a list and as a heap.
+\ifen (Experiment 1.) Time in ms to compute an optimal portfolio for an admissions market with homogeneous application costs using Algorithm \ref{algorithmforlargeh} when $\mathcal{C}$ is stored as a list and as a heap.
 \lastptofcaption~In every case, $h = m/2$.
-\else (단위: ms) 알고리즘 \ref{algorithmforlargeh}에서 $\mathcal{C}$를 목록 또는 힙 구현으로 저장했을 때, 
+\else (실험 1.) 단위: ms. 알고리즘 \ref{algorithmforlargeh}에서 $\mathcal{C}$를 목록 또는 힙 구현으로 저장했을 때, 
 동일 지원 비용 입학 시장의 최적 지원 포트폴리오를 계산하는 시간. \lastptofcaption~모든 경우, $h = m/2$. \fi}
@@
 \caption{\label{experiment2results} \normalsize
-\ifen Time in ms to compute an optimal or $(1 - \varepsilon)$-optimal portfolio for an admissions market with heterogeneous application costs using the three algorithms developed in Section \ref{hetappcosts}.
+\ifen (Experiment 2.) Time in ms to compute an optimal or $(1 - \varepsilon)$-optimal portfolio for an admissions market with heterogeneous application costs using the three algorithms developed in Section \ref{hetappcosts}.
 %branch-and-bound routine (Algorithm \ref{ellisbnb}), exact dynamic program (Algorithm \ref{ellisDP1}), and the fully polynomial-time approximation scheme (Algorithm \ref{ellisDP3}).
@@
 \else  %분지한계법 (알고리즘 \ref{ellisbnb}), 정확한 동적 계획 (알고리즘 \ref{ellisDP1}), 그리고 완전 다항 시간 근사 해법 (알고리즘 \ref{ellisDP3})을
-(단위: ms)  \ref{hetappcosts}절에서 도출한 3개의 알고리즘을 사용할 때, 다양한 지원 비용으로 갖춘 입학 시장의 최적 또는 $(1- \varepsilon)$-최적 포트폴리오를 계산하는 시간. 분지한계법은 큰 시장에서 비실용적이다. \lastptofcaption \fi}
+(실험 2.) 단위: ms. \ref{hetappcosts}절에서 도출한 3개의 알고리즘을 사용할 때, 다양한 지원 비용으로 갖춘 입학 시장의 최적 또는 $(1- \varepsilon)$-최적 포트폴리오를 계산하는 시간. 분지한계법은 큰 시장에서 비실용적이다. \lastptofcaption \fi}
 \end{table}
@@
 \centering
-\includegraphics[width=0.85\textwidth]{./plots/samplemarket.pdf}
+\includegraphics[width=0.95\textwidth]{./plots/accuracy_simulatedannealing.pdf}
   \caption{\label{samplemarket}
- \ifen A typical randomly-generated instance with $m=64$ schools and its optimal application portfolio.
-The application costs $g_j$ were drawn uniformly from $\{5, \dots, 10\}$.
-The optimal portfolio was computed using Algorithm \ref{ellisDP1}.
- \else $m=64$개의 학교로 구성된 전형적인 무작위로 생성한 인스턴스와 해당 최적 포트폴리오. 지원 비용 $g_j$는 $\{5, \dots, 10\}$에서 균일한 확률로 선정했으며 기호의 넓이와 비례된다. 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
+ \ifen (Experiment 3.) WIP.  The optimal portfolio was computed using Algorithm \ref{ellisDP1}.
+ \else  (실 3.) WIP.  최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
 \end{figure}
@@
 
+Wolsey, Laurence. 1998. \emph{Integer Programming.} New York: John Wiley \& Sons.
+
 \end{document}

### 2022-04-15 benchmark/benchmark.jl

@@ -9,7 +9,7 @@ import Printf: @sprintf
 # using BenchmarkTools
 using UnicodePlots
 
-const fullscale = true
+const fullscale = false
 
 # A long benchmark; tweak parameters with caution.
 # Set fullscale = false to run a smaller benchmark to check formatting etc.
@@ -18,22 +18,17 @@ const fullscale = true
 const n_reps = fullscale ? 3 : 2
 
 # Number of markets to test at each intersection of the experimental variables
-const n_markets = fullscale ? 50 : 2
+const n_markets = fullscale ? 50 : 5
 
 # Cutoffs to exclude certain large computations
-const bnbcutoff = fullscale ? 33 : 6
+const bnbcutoff = fullscale ? 33 : 10
 const twottbnbcutoff = 2^bnbcutoff
-const fptascutoff_m = fullscale ? 300 : 11
+const fptascutoff_m = fullscale ? 300 : 200
 const fptascutoff_eps = fullscale ? 0.0 : 0.1
 
-# Sizes of markets to test
-const marketsizes_SCM = fullscale ? 4 .^ (2:7) : [5, 10, 15]
-const marketsizes_VCM = fullscale ? 2 .^ (3:8) : [5, 10, 15]
-const epsilons = [0.5, 0.05]
-
-const mkts_SCM = SameCostsMarket[SameCostsMarket(m) for m in marketsizes_SCM, i in 1:n_markets]
-const mkts_VCM = VariedCostsMarket[VariedCostsMarket(m) for m in marketsizes_VCM, i in 1:n_markets]
-
+# Market sizes
+marketsizes_SCM = fullscale ? 4 .^ (2:7) : 2 .^ (3:6)
+marketsizes_VCM = fullscale ? 2 .^ (3:8) : 2 .^ (3:6)
 
 function makeunicodeplots(df::DataFrame)
     large_idx = df[!, :m] .== maximum(df[!, :m])
@@ -65,8 +60,9 @@ function printheader(s)
 end
 
 
-function benchmark1()
+function benchmark1(marketsizes_SCM = marketsizes_SCM)
     printheader("Benchmark 1: Homogeneous-cost algorithms")
+    mkts_SCM = SameCostsMarket[SameCostsMarket(m) for m in marketsizes_SCM, i in 1:n_markets]
 
     sizes = Int[m for m in marketsizes_SCM, _ in 1:n_markets]
[…]

### 2022-04-15 paper/plots/plots.jl

@@ -12,8 +12,8 @@ const accentcolor = (210, 10, 60) ./ 255
 
 function comparativestatics()
     mkt = Market([0.39, 0.33, 0.24, 0.24, 0.05, 0.03, 0.1, 0.12],
-                collect(range(start=200, step=50, length=8)),
-                8)
+        collect(range(start=200, step=50, length=8)),
+        8)
 
     apporder, v = applicationorder_list(mkt, verbose=true)
 
@@ -62,6 +62,7 @@ function samplerandom()
     mkt = VariedCostsMarket(m)
 
     X, v = optimalportfolio_dynamicprogram(mkt)
+    X[:] = invperm(mkt.perm)[X]
 
     nX = setdiff(1:m, X)
 
@@ -75,10 +76,10 @@ function samplerandom()
     scal = 3
     fig2 = Figure(resolution=(600, 600))
 
-    ax = Axis(fig2[1, 1], xlabel = "fⱼ", ylabel = "tⱼ", xlabelsize=18, ylabelsize=18)
+    ax = Axis(fig2[1, 1], xlabel="fⱼ", ylabel="tⱼ", xlabelsize=18, ylabelsize=18)
 
     scatter!(mkt.f[nX], mkt.t[nX], markersize=scal * sqrt.(mkt.g[nX]), color=RGBA(0.25, 0.25, 0.25, 0.8)) #:gray25, strokealpha=0.5)
-    scatter!(mkt.f[X], mkt.t[X], marker=:utriangle, markersize=1.6 * scal * sqrt.(mkt.g[X]), color=RGBA(accentcolor..., 0.85), label="Apply when H = $(mkt.H)")
+    scatter!(mkt.f[X], mkt.t[X], marker=:utriangle, markersize=1.6 * scal * sqrt.(mkt.g[X]), color=RGBA(accentcolor..., 0.7), label="Apply when H = $(mkt.H)")
 
     axislegend(ax, position=:rt, labelsize=16)
     text!("(Marker area: gⱼ)", position=(maximum(mkt.f), 0.9 * maximum(mkt.t)), align=(:right, :baseline), textsize=16)
@@ -86,11 +87,127 @@ function samplerandom()
     return fig2
 end
 
-fig1 = comparativestatics()
-fig2 = samplerandom()
+#   1D simulated annealing accuracy
+#   =================================
 
-save("paper/plots/h_v-example.pdf", fig1)
-save("paper/plots/h_v-example.png", fig1)
+function accuracy_simulatedannealing()
+    n_markets = 5000
[…]

### 2022-04-15 src/FastSolvers.jl

@@ -113,8 +113,9 @@ function applicationorder_heap(mkt::SameCostsMarket{T})::Tuple{Vector{Int},Vecto
         v[j] = get(v, j - 1, 0) + c_k.ft
         apporder[j] = c_k.j
 
+        # These two implementations perform almost identically
         mkt_heap = BinaryMaxHeap{College{T}}(
-            map(filter(c -> c.j != c_k.j, mkt_heap.valtree)) do c
+            College{T}[
                 College(
                     c.j,
                     c.f,
@@ -122,7 +123,17 @@ function applicationorder_heap(mkt::SameCostsMarket{T})::Tuple{Vector{Int},Vecto
                     c.t < c_k.t ? c.ft * c_k.omf : c.ft - c.f * c_k.ft,
                     c.omf
                 )
-            end
+                for c in mkt_heap.valtree if c.j != c_k.j
+            ]
+            # map(filter(c -> c.j != c_k.j, mkt_heap.valtree)) do c
+            #     College(
+            #         c.j,
+            #         c.f,
+            #         c.t < c_k.t ? c.t * c_k.omf : c.t - c_k.ft,
+            #         c.t < c_k.t ? c.ft * c_k.omf : c.ft - c.f * c_k.ft,
+            #         c.omf
+            #     )
+            # end
         )
     end
 
@@ -285,6 +296,11 @@ end
 end
 
 
+# Function I stole from Base.Sort that does midpoint using bit hacks.
+# Performance-optimized but safe only if `lo <= hi`.
+midpoint(lo::T, hi::T) where {T<:Integer} = lo + ((hi - lo) >>> 0x01)
+
+
 """
     optimalportfolio_fptas(mkt::VariedCostsMarket, ε) -> X, v
 
@@ -297,12 +313,19 @@ function optimalportfolio_fptas(mkt::VariedCostsMarket{T}, ε::Float64; verbose:
     G_dict = Dict{Tuple{T,Int},Int}()
     sizehint!(G_dict, mkt.m * mkt.m ÷ 2)
 
-    # Binary search
+    # Binary search for max{ w :  G_recursor!(G_dict, mkt.m, w, mkt, sp ≤ H }
+
+    # Should be able to use something like this with a proper by kw in searchsortedlast
+    # vs = 0:sp.Ū
+    # searchsortedlast(vs)
+
+
     v = 0
     v_UB = sp.Ū
 
-    while v + 1 < v_UB
[…]

### 2022-04-15 src/HeuristicSolvers.jl

@@ -70,9 +70,8 @@ end
         verbose::Bool=false
     ) -> X, v
 
-Use a greedy heuristic that adds schools in decreasing order of `mkt.ft ./ mkt.g`
-to compute an approximately optimal portfolio for the market `mkt` with
-varying application costs.
+Use a simulated annealing procedure to compute a heuristically optimal portfolio
+for `mkt` and its valuation. 
 """
 function optimalportfolio_simulatedannealing(
     mkt::VariedCostsMarket{T};

### 2022-04-18 benchmark/annealingparams.jl
grid search for optimal annealing params

@@ -1,0 +1,88 @@
+using OptimalApplication
+using DataFrames
+using PrettyTables
+using Plots
+using Statistics: median, quantile
+
+function plot_params()
+    Ts = 2.0 .^ (-4:2:4)
+    rs = (2.0 .^ (-5:0))
+
+    n_markets = 1000
+
+    sizerangelog2 = (3, 9)
+
+    sizes = round.(
+        Int,
+        2.0 .^ (
+            sizerangelog2[1] .+ (sizerangelog2[2] - sizerangelog2[1]) * rand(n_markets)
+        )
+    )
+
+    rats = zeros(length(rs), length(Ts), n_markets)
+
+    @time for k in 1:n_markets
+        mkt = VariedCostsMarket(sizes[k])
+        opt = optimalportfolio_dynamicprogram(mkt)[2]
+
+        Threads.@threads for j in 1:length(Ts)
+            for i in 1:length(rs)
+                rats[i, j, k] = optimalportfolio_simulatedannealing(
+                    mkt;
+                    temp=Ts[j],
+                    red=rs[i]
+                )[2]
+            end
+        end
+
+        rats[:, :, k] /= optimalportfolio_dynamicprogram(mkt)[2]
+    end
+
+    xts = 2 .^ range(sizerangelog2...)
+
+    pl = plot(
+        layout=grid(length(rs), length(Ts)),
+        link=:all,
+        xscale=:log10,
+        xticks=(xts, xts),
+        legend=nothing,
+        size=(1200, 1200),
+        xlim=2 .^ (sizerangelog2 .+ (-0.5, 0.5)),
+    )
+
+
+    avgs = zeros(length(rs), length(Ts))
+    for j in 1:length(Ts), i in 1:length(rs)
+        avgs[i, j] = quantile(rats[i, j, :], 0.1)
+    end
+    # [sum(rats[i, j, :]) / length(rats[i, j, :]) for i in 1:length(rs), j in 1:length(Ts)]
+    best_avg = maximum(avgs)
+    @show best_avg
+
+    for j in 1:length(Ts), i in 1:length(rs)
+        hline!(pl[i, j], [avgs[i, j]], c=:navy)
+        scatter!(
+            pl[i, j],
+            sizes,
+            rats[i, j, :],
+            c=avgs[i, j] ≥ best_avg ? :olivedrab : :crimson,
+            msc=:auto,
+            alpha=0.7,
[…]

### 2022-04-18 src/HeuristicSolvers.jl

@@ -76,9 +76,9 @@ for `mkt` and its valuation.
 function optimalportfolio_simulatedannealing(
     mkt::VariedCostsMarket{T};
     X0::Union{Nothing,Vector{T}}=nothing,
-    temp::Union{Nothing,Float64}=nothing,
+    temp::Float64=0.25,
     nit::Integer=500,
-    red::Float64=0.99,
+    red::Float64=0.0625,
     verbose::Bool=false
 )::Tuple{Vector{Int},Float64} where {T}
     if X0 === nothing
@@ -107,10 +107,6 @@ function optimalportfolio_simulatedannealing(
         end
     else
         X, v = copy(X0), valuation(X0, mkt)
-    end
-
-    if temp === nothing
-        temp = log(mkt.m)
     end
 
     X_best, v_best = X, v

### 2022-04-18 src/FastSolvers.jl
use break instead of goto

@@ -331,12 +331,8 @@ function optimalportfolio_fptas(mkt::VariedCostsMarket{T}, ε::Float64; verbose:
             v_UB = mid
         else
             v = mid
-            # if G_recursor!(G_dict, mkt.m, v + 1, mkt, sp) > mkt.H
-            #     @goto foundit
-            # end
         end
     end
-    @label foundit
 
     X = T[]
 

### 2022-04-18 src/HeuristicSolvers.jl

@@ -17,13 +17,10 @@ function optimalportfolio_greedy(mkt::VariedCostsMarket{T})::Tuple{Vector{Int},F
         if mkt.g[j] ≤ H
             push!(X, j)
             H -= mkt.g[j]
-        end
-
-        if H == 0
-            @goto full
+        else
+            break
         end
     end
-    @label full
 
     # In the valuation we just use the identity permutation as invp
     # To prevent wasteful permuting and then invpermuting
@@ -90,7 +87,7 @@ function optimalportfolio_simulatedannealing(
             ) do j
                 mkt.g[j] ≤ mkt.H
             end
-    
+
             X = T[]
             H = mkt.H
             for j in priority_order
@@ -98,11 +95,10 @@ function optimalportfolio_simulatedannealing(
                     push!(X, j)
                     H -= mkt.g[j]
                 else
-                    @goto full
+                    break
                 end
             end
-            @label full
-    
+
             X, valuation(X, mkt; invp=oneunit(T):mkt.m)
         end
     else

### 2022-04-18 src/VariedCostsBnB.jl

@@ -38,11 +38,9 @@ mutable struct Node{T<:Unsigned}
                 v_LP += mkt.f[j] * t̄[j]
             else
                 v_LP += mkt.f[j] * t̄[j] * H_left / mkt.g[j]
-                @goto done
+                break
             end
         end
-        # Need to use goto here for the case in which all of N fits within H̄
-        @label done
 
         return new{T}(I, N, t̄, H̄, v_I, v_LP)
     end

### 2022-04-18 paper/paper.tex
plot of simulated annealing grid search, update accuracy plot, translate writeup

@@
 \newcommand{\lIfElse}[3]{\lIf{#1}{#2 \textbf{else}~#3}}
+\SetKwInOut{KwParams}{Parameters}
 
@@
 \newif\ifen
-\entrue % Change to false to export Korean
+\enfalse% Change to false to export Korean
 
@@
 The second is an dynamic program that produces an exact solution in pseudopolynomial time.
-The third is a different dynamic program that yields fully polynomial-time approximation scheme.
+The third is a different dynamic program that yields a fully polynomial-time approximation scheme.
 Numerical experiments demonstrate the algorithms' practicality.
@@
 \ifen
-It is helpful to define the random variable $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$ as the utility achieved by the schools in the portfolio, so that when $t_0 = 0$, $v(\mathcal{X}) = \operatorname{E}[X]$.
-Similar pairs of variables with script and italic names such as $\mathcal{Y}_h$ and $Y_h$ carry an analogous meaning.
-\else
-편의상 포트폴리오에 속한 학교가 실제로 이루는 효용을 확률 변수 $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$로 정의한다.
+\noindent It is helpful to define the random variable $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$ as the utility achieved by the schools in the portfolio, so that when $t_0 = 0$, $v(\mathcal{X}) = \operatorname{E}[X]$.
+Similar pairs of variables with script and italic names such as $\mathcal{Y}_h$ and $Y_h$ will carry an analogous meaning.
+\else
+\noindent 편의상 포트폴리오에 속한 학교가 실제로 이루는 효용을 확률 변수 $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$로 정의한다.
 그러면 $t_0 = 0$이면 $v(\mathcal{X}) = \operatorname{E}[X]$임을 알 수 있다.
@@
 \caption{Optimal portfolio algorithm for Alma’s problem.} \label{algorithmforlargeh}
-\KwData{Utility values $t \in(0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application limit $h \leq m$.}
+\KwIn{Utility values $t \in(0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application limit $h \leq m$.}
 %Index schools in ascending order by $t$\;
@@
 \caption{알마의 문제를 위한 최적 포트폴리오 알고리즘.} \label{algorithmforlargeh}
-\KwData{효용 모수 $t \in(0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 제한 $h \leq m$.}
+\KwIn{효용 모수 $t \in(0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 제한 $h \leq m$.}
 %$t$의 순서대로 학교의 지표를 짓는다\;
@@
 \ifen
-\noindent In this section, we show that this problem is NP-complete, then provide three algorithmic solutions: an exact branch-and-bound routine, an exact dynamic program, and a fully polynomial-time approximation scheme (FPTAS).
-\else
-\noindent 본 절에서 이 문제가 NP-complete함을 보이고 3가지 해법을 제시한다: 정확한 분지한계법 해법, 정확한 동적 계획 해법, 그리고 완전 다항 시간 근사 해법 (fully polynomial-time approximation scheme, FPTAS).
+\noindent In this section, we show that this problem is NP-complete, then provide four algorithmic solutions: an exact branch-and-bound routine, an exact dynamic program, a fully polynomial-time approximation scheme (FPTAS), and a simulated-annealing heuristic.
+\else
+\noindent 본 절에서 이 문제가 NP-complete함을 보이고 3가지 해법을 제시한다: 정확한 분지한계법 해법, 정확한 동적 계획 해법, 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme, FPTAS), 그리고 모의 담금질(simulated annealing) 휴리스틱.
 \fi
@@
 An intuitive extension of the greedy algorithm for Alma's problem is to iteratively add to $\mathcal{X}$ the school $k$ for which $\bigl( v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) \bigr) / g_k$ is largest.
-However, the construction above shows that the objective function of Ellis's problem can approximate that of a knapsack problem with arbitrary precision.
+But the construction above shows that the objective function of Ellis's problem can approximate that of a knapsack problem with arbitrary precision.
 Therefore, in pathological examples such as the following, the greedy algorithm can achieve an arbitrarily small approximation ratio.
@@
 \fi
-\begin{example}
-\ifen
-Let $t = (10, 2021)$, $f = (1, 1)$, $g = (1, 500)$, and $H = 500$. Then the greedy approximation algorithm produces the clearly inoptimal solution $\mathcal{X} = \{1\}$. 
-\else
- $t = (10, 2021)$, $f = (1, 1)$, $g = (1, 500)$, 그리고 $H = 500$이라고 하자. 그러면 탐욕 해법은 $\mathcal{X} = \{1\}$을 출력하며 이는 분명히 최적해가 아니다.
+\begin{example} \label{greedyzeroellis}
+\ifen
+Let $t = (10, 2021)$, $f = (0.1, 0.1)$, $g = (1, 500)$, and $H = 500$. Then the greedy approximation algorithm produces the clearly inoptimal solution $\mathcal{X} = \{1\}$. 
+\else
+ $t = (10, 2021)$, $f = (0.1, 0.1)$, $g = (1, 500)$, 그리고 $H = 500$이라고 하자. 그러면 탐욕 해법은 $\mathcal{X} = \{1\}$을 출력하며 이는 분명히 최적해가 아니다.
 \fi
@@
 \ifen 
-\noindent In summary, neither a greedy algorithm nor a knapsack-based algorithm solves Ellis's problem to optimality. 
-\else
-\noindent 탐욕 해법 또는 배낭 문제 기반 해법으로 엘리스 문제의 정확한 최적해를 구할 수 없음을 의미한다.
+In summary, neither a greedy algorithm nor a knapsack-based algorithm solves Ellis's problem to optimality. 
+\else
+탐욕 해법 또는 배낭 문제 기반 해법으로 엘리스 문제의 정확한 최적해를 구할 수 없음을 의미한다.
 \fi
@@
 \caption{Branch and bound for Ellis's problem.} \label{ellisbnb}
-\KwData{Utility values $t \in(0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)$.}
+\KwIn{Utility values $t \in(0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)$.}
 Root node $n_0 \gets (\O, \O, \mathcal{C})$\;
@@
 \caption{엘리스의 문제를 위한 분지한계법.} \label{ellisbnb}
-\KwData{효용 모수 $t \in(0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 비용 $g \in (0, \infty)^m$, 예산 $H \in (0, \infty)$.}
+\KwIn{효용 모수 $t \in(0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 비용 $g \in (0, \infty)^m$, 예산 $H \in (0, \infty)$.}
 뿌리 마디 $n_0 \gets (\O, \O, \mathcal{C})$\;
@@
 		$n \gets \argmax\{ v_{\mathrm{LP}}^*[n] : n \in \mathfrak{T}\}$\;
-		$\mathfrak{T}$에서 $n$을 제거한다\;
+		$\mathfrak{T}$에서 $n$을 제거\;
 %		Generate children of $n$\;
@@
 				$L \gets v_{\mathcal{I}}[n']$\;
-				$\mathcal{X}$을 $n'$에 대응하는 $\mathcal{I}$-집합으로 수정한다\;
-			}
-			\lIf{$v_{\mathrm{LP}}^*[n] > L$이고 $n'$이 잎 마디 아님}{$\mathfrak{T}$에 $n'$을 추가한다}
+				$\mathcal{X}$을 $n'$에 대응하는 $\mathcal{I}$-집합으로 수정\;
+			}
+			\lIf{$v_{\mathrm{LP}}^*[n] > L$이고 $n'$이 잎 마디 아님}{$\mathfrak{T}$에 $n'$을 추가}
 		}
 		\For{$n'' \in \mathfrak{T}$} {
-			\lIf{$L >  v_{\mathrm{LP}}^*[n'']$}{$\mathfrak{T}$에서 $n''$을 제거한다}
+			\lIf{$L >  v_{\mathrm{LP}}^*[n'']$}{$\mathfrak{T}$에서 $n''$을 제거}
 		}
@@
 \caption{Dynamic program for Ellis's problem with integral application costs.} \label{ellisDP1}
-\KwData{Utility values $t \in(0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in \mathbb{N}^m$, budget $H \in\mathbb{N}$.}
+\KwIn{Utility values $t \in(0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in \mathbb{N}^m$, budget $H \in\mathbb{N}$.}
 Index schools in ascending order by $t$\;
@@
 \caption{정수 지원 비용의 엘리스 문제를 위한 동적 계획 해법.} \label{ellisDP1}
-\KwData{효용 모수 $t \in(0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 비용 $g \in \mathbb{N}^m$, 예산 $H \in\mathbb{N}$.}
+\KwIn{효용 모수 $t \in(0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 비용 $g \in \mathbb{N}^m$, 예산 $H \in\mathbb{N}$.}
 $t$의 순서대로 학교를 배열한다\;
@@
 \caption{Fully polynomial-time approximation scheme for Ellis's problem.} \label{ellisDP3}
-\KwData{Utility values $t \in  (0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$, tolerance $\varepsilon \in (0, 1)$.}
+\KwIn{Utility values $t \in  (0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$.}
+\KwParams{Tolerance $\varepsilon \in (0, 1)$.}
 Index schools in ascending order by $t$\;
@@
 \caption{엘리스의 문제를 위한 완전 다항 시간 근사 해법.} \label{ellisDP3}
-\KwData{효용 모수 $t \in  (0, \infty)}^m$, 합격 확률 $f \in (0, 1]^m$, 지원 비용 $g \in (0, \infty)^m$, 예산 $H \in (0, \infty)^m$, 허용치 $\varepsilon \in (0, 1)$.}
+\KwIn{효용 모수 $t \in  (0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 비용 $g \in (0, \infty)^m$, 예산 $H \in (0, \infty)^m$.}
+\KwParams{허용치 $\varepsilon \in (0, 1)$.}
 $t$의 순서대로 학교를 배열한다\;
@@
 
-\ifen \subsection{Simulated annealing} \else \subsection{Simulated annealing}\fi\label{numericalexperiments}
-Simulated annealing is a popular local-search technique for nonconvex optimization problems.
+\ifen \subsection{Simulated annealing heuristic} \else \subsection{모의 담금질 휴리스틱}\fi\label{numericalexperiments}
+\ifen
+Simulated annealing is a popular local-search heuristic for nonconvex optimization problems.
 Because of their inherent randomness, simulated-annealing algorithms offer no accuracy guarantee; however, they are easy to implement, computationally inexpensive, and often produce solutions with a small optimality gap.
 In this section, we present a simulated-annealing algorithm for Ellis's problem.
-It is based on the procedure for integer programs outlined by Wolsey (1998, \S\,12.3.2)
-
+It is based on the procedure for integer programs outlined by Wolsey (1998, \S\,12.3.2).
+\else
+모의 담금질은 비볼록 최적화 문제를 위한 인기 많은 지역 탐색 휴리스틱이다.
+선천적인 확률성 때문에 모의 담금질 해법에는 보장된 정확도가 없다.
+반면에 구현하기 쉽고 계산 비용이 작으면서 최적에 가까운 해를 구하는 경향이 있다.
+본 항에서, 엘리스의 문제를 위한 모의 담금질 해법을 제시한다.
+Wolsey (1998, \S\,12.3.2)가 제시한 정수 계획 모의 담금질 해법이 바탕이 된다.
+\fi
+
+\ifen
 To generate an initial solution, our algorithm adopts the greedy heuristic of adding schools in descending order by $f_j t_j / g_j$ until the budget is exhausted.
-Example \ref{knapsackheuristicellis} shows that the approximation coefficient of this solution can be as low as $1 / (m-1)$; in typical instances, it is much higher.
-Given a solution $\mathcal{X}$, we generate a neighbor $\mathcal{X}_n$ by copying $\mathcal{X}$, adding schools to $\mathcal{X}_n$ until it becomes infeasible, then removing elements of $\mathcal{X}$ from $\mathcal{X}_n$ until feasibility is restored.
-If this increases the valuation of $\mathcal{X}$, then the candidate solution is updated to equal the neighbor; if not, then the candidate solution is updated with probability $\exp\bigl(v(\mathcal{X}_n) - v(\mathcal{X}) / T\bigr)$, where $T$ is a temperature parameter.
+While Example \ref{greedyzeroellis} shows that the approximation coefficient of this solution can be arbitrarily small, in typical instances, it is much higher.
+Next, given a candidate solution $\mathcal{X}$, we generate a neighbor $\mathcal{X}_n$ by copying $\mathcal{X}$, adding schools to $\mathcal{X}_n$ until it becomes infeasible, then removing elements of $\mathcal{X}$ from $\mathcal{X}_n$ until feasibility is restored.
+If $v(\mathcal{X}_n) \geq v(\mathcal{X})$, then the candidate solution is updated to equal the neighbor; if not, then the candidate solution is updated with probability $\exp\bigl(v(\mathcal{X}_n) - v(\mathcal{X}) / T\bigr)$, where $T$ is a temperature parameter.
 This is repeated $N$ times, and the algorithm returns the best $\mathcal{X}$ observed.
-
+\else
+초기 해를 선정하기 위해, 예산이 고갈될 때까지 예산이 고갈될 때까지 $f_j t_j / g_j$-값이 감소하는 순서대로 학교를 추가하는 탐욕 휴리스틱을 도입한다.
+예 \ref{greedyzeroellis}에서 보였듯이 탐욕 해의 근사 계수는 무한히 작아질 수 있지만, 전형적인 인스턴스에서는 좋은 초기 해이다.
+그다음으로, 후보의 가능해 $\mathcal{X}$의 이웃 $\mathcal{X}_n$을 생성하기 위해 $\mathcal{X}$를 복사하고 불가능해질 때까지 학교를 추가하고, 가능성이 복구될 때까지 $\mathcal{X}$의 원소를 $\mathcal{X}_n$에서 제거한다.
+만약 $v(\mathcal{X}_n) \geq v(\mathcal{X})$이면, 후보를 이웃으로 대체한다.
+아니면, $\exp\bigl(v(\mathcal{X}_n) - v(\mathcal{X}) / T\bigr)$의 확률로 대체하며, 여기서 $T$는 ``온도 모수''라고 불린다.
+이를 $N$번 반복한 다음에 가장 좋은 $\mathcal{X}$를 출력한다.
+\fi
+
+\ifen
 \begin{algorithm}[h] 
@@
 \caption{Simulated annealing for Ellis's problem.} \label{ellissimann}
-\KwData{Utility values $t  (0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$, initial temperature $T > 0$, temperature reduction parameter $r \in (0, 1)$, number of iterations $N$.}
+\KwIn{Utility values $t  (0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application costs $g \in (0, \infty)^m$, budget $H \in (0, \infty)^m$.}{}
+\KwParams{Initial temperature $T \geq 0$, temperature reduction parameter $r \in (0, 1]$, number of iterations $N$.}
 Add schools to $\mathcal{X}$ in descending order by $f_j t_j / g_j$ until the budget is exhausted\;
@@
 \end{algorithm}
+\else
+\begin{algorithm}[h] 
+%\DontPrintSemicolon
+\caption{엘리스의 문제를 위한 모의 담금질.} \label{ellissimann}
+\KwIn{효용 모수 $t \in  (0, \infty)^m$, 합격 확률 $f \in (0, 1]^m$, 지원 비용 $g \in (0, \infty)^m$, 예산 $H \in (0, \infty)^m$.}
+\KwParams{초기 온도 $T \geq 0$, 온도 절감 모수 $r \in (0, 1]$, 반복 단계 개수 $N$.}
+예산이 고갈될 때까지 학교를 $f_j t_j / g_j$가 감소하는 순서대로 $\mathcal{X}$에 추가\;
+\For{$i = 1 \dots N$}{
+	$\mathcal{X}_n \gets \operatorname{copy}(\mathcal{X})$\;
+	$\mathcal{X}_n$이 불가능해질 때까지 $\mathcal{C} \setminus \mathcal{X}$에서 학교를 무작위로 추가\;
+	$\mathcal{X}_n$의 가능성이 복구될 때까지 $\mathcal{X}$에 속한 학교를 $\mathcal{X}_n$에서 제거\;
+	$\Delta = v(\mathcal{X}_n) - v(\mathcal{X})$\;
+	\lIfElse{$\Delta \geq 0$}{$\mathcal{X} \gets \mathcal{X}_n$}
+	{$\exp(\Delta / T)$의 확률로 $\mathcal{X} \gets \mathcal{X}_n$}
+	$T \gets rT$\;
+}
+\Return{가장 좋은 $\mathcal{X}$}
+\end{algorithm}
+\fi
 
 \begin{proposition}
-The computation time of Algorithm \ref{ellissimann} is $O(Nm)$. 
+\ifen The computation time of Algorithm \ref{ellissimann} is $O(Nm)$. 
+\else 알고리즘 \ref{ellissimann}의 계산 시간은 $O(Nm)$이다.\fi
 \end{proposition}
 
-In our numerical experiments, the output of Algorithm \ref{ellissimann} was typically quite close to optimal.
+\ifen
+\noindent In our numerical experiments, the output of Algorithm \ref{ellissimann} was typically quite close to the optimum.
+\else
+\noindent 계산 실험 결과에 따라, 알고리즘 \ref{ellissimann}의 출력은 보통 최적해에 아주 가깝다.
+\fi
 
@@
 \ifen \subsection{Experimental procedure} \else \subsection{실험 방법}\fi
+\ifen
 We conducted three numerical experiments.
@@
 Experiment 3 investigates the empirical accuracy of the simulated-annealing heuristic as compared to an exact algorithm.
+\else
+총 3개의 수리 실험을 진행했다. 실험 1과 실험 2에서, 각각 알마의 문제와 엘리스의 문제를 위한 해법의 비교적 계산 시간을 판단하고자 한다. 실험 3에서, 정확한 해법에 비한 모의 담금질 휴리스틱의 현실적 정확도를 탐구한다.
+\fi
 
@@
 Where applicable, we do not count the time required to sort the entries of $t$.
-As the simulated annealing algorithm's computation time is negligible, it was excluded from Experiment 2.
+As the simulated-annealing heuristic's computation time is negligible, it was excluded from Experiment 2.
 \else
@@
 $t$를 배열하는 시간은 고려하지 않았다.
-As the simulated annealing algorithm's computation time is negligible, it was excluded from Experiment 2.
-\fi
-
-In Experiment 3, we generated 500 markets with sizes between $2^3$ and $2^11$ varying uniformly in the logarithm.
+그리고 모의 담금질 휴리스틱의 계산 시간이 무시해도 될 정도이므로 실험 2에서 제외했다.
+\fi
+
+\ifen
+In Experiment 3, we generated 500 markets with sizes varying uniformly in the logarithm between $m = 2^3$ and $2^{11}$.
 For each market, we computed a heuristic solution using Algorithm \ref{ellissimann} and an exact solution using Algorithm \ref{ellisDP1} and calculated the optimality ratio.
+We chose to use $N = 500$ iterations in the simulated-annealing algorithm, and the temperature parameters $T = 1/4$ and $r = 1/16$ were chosen by a preliminary grid search.
+\else
+실험 3에서, 500개의 시장을 생성했으며, 크기는 대수가 균일하게 분포되도록 $m = 2^3$과 $2^{11}$ 사이에서 무작위로 선택했다.
+각 시장에  알고리즘 \ref{ellissimann}을(를) 적용하여 휴리스틱한 해를 구하고 알고리즘 \ref{ellisDP1}을(를) 적용하여 정확한 해를 구한 다음의 최적성 비율을 계산했다.
+모의 담금질 반복 단계의 개수는 $N = 500$이며, 격자 탐색을 이용한 예비 실험 결과에 따라 온도 모수를 $T = 1/4$ 그리고 $r = 1/16$로 선정했다.
+\fi
 
@@
 
-Experiment 3 \dots. (WIP)
+\ifen
+The results of Experiment 3, which evaluated the performance of the simulated-annealing heuristic in our synthetic instances, are plotted in Figure \ref{experiment3results}.
+In the every instance, the heuristic found a solution within 10 percent of optimality, and in the vast majority of instances, it was within 2 percent of optimality.
+The heuristic performs most poorly in small markets of size $m \leq 16$, but as exact algorithms are tractable at this scale, this result presents no cause for concern.
+\else
+실험 3은 가상 인스턴스를 통해 모의 담금질 휴리스틱의 현실 성능을 탐구했으며, 실험 결과는 그림 \ref{experiment3results}에서 나타난다.
+모든 인스턴스에서 최적의 10\% 이내에 해를 구했으며, 대다소의 경우는 2\% 이내로 달했다.
+휴리스틱의 성능은 $m \leq 16$이 되는 작은 시장에서 가장 낮았으며, 정확한 해법도 효율적으로 풀 수 있는 시장 크기이므로 염려할 이유가 없어 보인다.
+\fi
 
@@
 \includegraphics[width=0.95\textwidth]{./plots/accuracy_simulatedannealing.pdf}
-  \caption{\label{samplemarket}
- \ifen (Experiment 3.) WIP.  The optimal portfolio was computed using Algorithm \ref{ellisDP1}.
- \else  (실 3.) WIP.  최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
+  \caption{\label{experiment3results}
+ \ifen (Experiment 3.) Optimality ratio achieved by simulated annealing (Algorithm \ref{ellissimann}) in markets of varying size, with parameters $N = 500$, $T = 1/4$, and $r = 1/16$.
+Optimal portfolios were computed using Algorithm \ref{ellisDP1}.
+ \else  (실험 3.) 크기가 다양한 시장에 적용한 모의 담금질 (알고리즘 \ref{ellissimann}) 해법이 달성하는 최적성 비율. 모의 담금질의 모수는 $N = 500, T = 1/4, r = 1/16$이며, 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
 \end{figure}
@@
 The general problem is NP-complete.
-We provided three solution algorithms.
+We provided four solution algorithms.
 The strongest from a theoretical standpoint is an FPTAS that produces a $(1-\varepsilon)$-approximate solution in $O(m^3 / \varepsilon)$-time.
 On the other hand, for typical college-application instances, the dynamic program based on application expenditures is both easier to implement and substantially more time efficient.
+A heuristic simulated-annealing algorithm also exhibited strong empirical performance in our computational study.
 The algorithms discussed in this paper are summarized in Table \ref{algorithmssummary}.
@@
 일반적인 문제는 NP-complete하다.
-3가지 해법을 제시했으며 이론적으로 가장 효율적인 알고리즘은 $O(m^3 / \varepsilon)$-시간 안에  $(1-\varepsilon)$-근사해를 계산하는 FPTAS이다.
+4가지 해법을 제시했으며 이론적으로 가장 효율적인 알고리즘은 $O(m^3 / \varepsilon)$-시간 안에  $(1-\varepsilon)$-근사해를 계산하는 FPTAS이다.
 그 반면에 전형적인 대학 지원 문제 인스턴스에 대해, 지원 비용 지출액 기반 동적 계획 해법은 실행하기 쉬울 뿐더러 계산 시간이 상당히 짧다.
+계산 실험에서 모의 담금질 기반 휴리스틱 해법도 좋은 성능을 발휘했다.
 표 \ref{algorithmssummary}에서 본 논문이 다운 알고리즘을 정리한다.
@@
 \xrowht[()]{1.5em}  Costs DP     &  Algorithm \ref{ellisDP1}& General          & $g_j$ integer         & Exact                    & $O(Hm + m \log m)$        \\
-\xrowht[()]{1.5em}  FPTAS      &    Algorithm \ref{ellisDP3} & General          &  None         & $(1 - \varepsilon)$-opt. & $O(m^3 / \varepsilon)$   
+\xrowht[()]{1.5em}  FPTAS      &    Algorithm \ref{ellisDP3} & General          &  None         & $(1 - \varepsilon)$-opt. & $O(m^3 / \varepsilon)$   \\
+\xrowht[()]{1.5em}  \begin{tabular}[r]{@{}r@{}}Simulated\\annealing\end{tabular}        &    Algorithm \ref{ellissimann} & General          &  None         & $0$-opt. & $O(Nm)$   
 \end{tabular}%
@@
 \xrowht[()]{1.7em}   \begin{tabular}[r]{@{}r@{}}지출액\\ 동적 계획\end{tabular}     &  알고리즘 \ref{ellisDP1}& 일반 문제          & $g_j$ 정수         & 정확                    & $O(Hm + m \log m)$        \\
-\xrowht[()]{1.7em}  FPTAS      &    알고리즘 \ref{ellisDP3} & 일반 문제          &  없음    & $(1 - \varepsilon)$-근사 & $O(m^3 / \varepsilon)$   
+\xrowht[()]{1.7em}  FPTAS      &    알고리즘 \ref{ellisDP3} & 일반 문제          &  없음    & $(1 - \varepsilon)$-근사 & $O(m^3 / \varepsilon)$   \\
+\xrowht[()]{1.7em}  모의 담금질      &    알고리즘  \ref{ellissimann} & 일반 문제          &  없음    & $0$-근사 & $O(Nm)$   
 \end{tabular}%
@@
 A potential application of our model is an econometric study that estimates students’ perceptions of college quality ($t_j$) from their observed application behavior ($\mathcal{X}$) under the assumption of utility maximization.
-However, empirical research indicates that heterogeneous risk aversion is a significant source of variation in human decision-making, especially in the domains of educational choice and employment search (Kahneman 2011; Hartlaub and Schneider 2012; van Huizen and Alessie 2019).
+Empirical research indicates that heterogeneous risk aversion is a significant source of variation in human decision-making, especially in the domains of educational choice and employment search (Kahneman 2011; Hartlaub and Schneider 2012; van Huizen and Alessie 2019).
 Therefore, an endogenous risk-aversion parameter would greatly enhance our model’s explanatory power in the econometric setting.
@@
 학생의 지원 행동($\mathcal{X}$)을 관찰하고 학생이 인식하는 대학교의 질($t_j$)을 추정하는 경제 측정학 연구에서 본 논문의 모형을 그대로 응용할 수 있다.
-그러나 실무 연구 결과에 따르면 위험 회피의 다양성은 인간 의사 결정 다양성의 주된 원인이 되며, 교육 선택과 취업 활동은 특히 그렇다 (Kahneman 2011; Hartlaub와 Schneider 2012; van Huizen과 Alessie 2019).
+실무 연구 결과에 따르면 위험 회피의 다양성은 인간 의사 결정 다양성의 주된 원인이 되며, 교육 선택과 취업 활동은 특히 그렇다 (Kahneman 2011; Hartlaub와 Schneider 2012; van Huizen과 Alessie 2019).
 따라서 경제 측정학 맥락에서, 위험 회피 모수를 내생 변수로 도입하면 원래 모형의 묘사력을 많이 개선할 수 있다.

### 2022-04-19 paper/paper.tex
use julian default arguments reduce hashing costs line edits in korean references to simulated annealing heuristic

@@
 When colleges differ in their application fees, we show that the problem is NP-complete.
-We provide three algorithms for this more general setup.
-The first is a branch-and-bound routine.
-The second is an dynamic program that produces an exact solution in pseudopolynomial time.
-The third is a different dynamic program that yields a fully polynomial-time approximation scheme.
-Numerical experiments demonstrate the algorithms' practicality.
+We provide four algorithms for this more general setup: a branch-and-bound routine, a dynamic program that produces an exact solution in pseudopolynomial time, a different dynamic program that yields a fully polynomial-time approximation scheme, and a simulated-annealing heuristic.
+Numerical experiments demonstrate the algorithms' accuracy and efficiency.
 
@@
 대학의 지원 비용이 서로 다른 경우, 문제가 NP-complete함을 증명한다.
-3가지 일반적인 해법을 도출한다.
-첫째는 분지한계 기반 해법이다.
-둘째는 의사 다항 시간 안에 정확한 해를 출력하는 동적 계획 해법이다.
-마지막으로 다른 동적 계획 기반으로 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme)을 도출한다.
-수리적 실험을 통하여 알고리즘의 현실성을 보인다.
+일반적인 문제를 위해 4가지 해법을 제시한다.
+분지한계 기반 해법, 의사 다항 시간 안에 정확한 해를 출력하는 동적 계획 해법, 다른 동적 계획 기반으로 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme), 그리고 모의 담금질(simulated annealing)을 이용한 휴리스틱 해법.
+% 첫째는 분지한계 기반 해법이다. 둘째는 의사 다항 시간 안에 정확한 해를 출력하는 동적 계획 해법이다. 마지막으로 다른 동적 계획 기반으로 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme)을 도출한다.
+수리적 실험을 통하여 알고리즘의 정확도와 효율성을 보인다.
 
@@
 We show that the decision form of the portfolio optimization problem is NP-complete through a polynomial reduction from the binary knapsack problem.
-We provide three algorithms for this more general setup.
+We provide four algorithms for this more general setup.
 The first is a branch-and-bound routine.
@@
 It yields a fully polynomial-time approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in $O(m^3 / \varepsilon)$ time.
+The fourth is a simulated-annealing heuristic algorithm that demonstrates strong performance in our synthetic instances.
 \else
@@
 이를 통하여 $O(m^3 / \varepsilon)$-시간 안에 $(1 - \varepsilon)$-근사해를 출력하는 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme)을 얻는다.
-\fi
-
-\ifen
-In Section \ref{numericalexperiments}, we present the results of computational experiments that confirm the validity and time complexity results established in the previous two sections.
+넷째는 가상 인스턴스를 푸는 성능이 좋은 모의 담금질 기반 휴리스틱 알고리즘이다.
+\fi
+
+\ifen
+In Section \ref{numericalexperiments}, we present the results of computational experiments that confirm the validity and time complexity results established in the previous two sections. % more
 
@@
 선천적인 확률성 때문에 모의 담금질 해법에는 보장된 정확도가 없다.
-반면에 구현하기 쉽고 계산 비용이 작으면서 최적에 가까운 해를 구하는 경향이 있다.
+반면에 구현하기 쉽고 계산 비용이 적으면서 최적에 가까운 해를 구하는 경향이 있다.
 본 항에서, 엘리스의 문제를 위한 모의 담금질 해법을 제시한다.
@@
 \else
-\noindent 계산 실험 결과에 따라, 알고리즘 \ref{ellissimann}의 출력은 보통 최적해에 아주 가깝다.
+\noindent 계산 실험 결과에 따라, 알고리즘 \ref{ellissimann}의 출력은 보통 최적에 아주 가깝다.
 \fi
@@
 각 시장에 알고리즘 \ref{ellissimann}을(를) 적용하여 휴리스틱한 해를 구하고 알고리즘 \ref{ellisDP1}을(를) 적용하여 정확한 해를 구한 다음의 최적성 비율을 계산했다.
-모의 담금질 반복 단계의 개수는 $N = 500$이며, 격자 탐색을 이용한 예비 실험 결과에 따라 온도 모수를 $T = 1/4$ 그리고 $r = 1/16$로 선정했다.
+모의 담금질의 반복 단계 개수는 $N = 500$이며, 격자 탐색을 이용한 예비 실험 결과에 따라 온도 모수를 $T = 1/4$ 그리고 $r = 1/16$로 선정했다.
 \fi
@@
 실험 3은 가상 인스턴스를 통해 모의 담금질 휴리스틱의 현실 성능을 탐구했으며, 실험 결과는 그림 \ref{experiment3results}에서 나타난다.
-모든 인스턴스에서 최적의 10\% 이내에 해를 구했으며, 대다소의 경우는 2\% 이내로 달했다.
+모든 인스턴스에서 최적의 10\% 이내에 해를 구했으며, 대다수의 경우는 2\% 이내로 달했다.
 휴리스틱의 성능은 $m \leq 16$이 되는 작은 시장에서 가장 낮았으며, 정확한 해법도 효율적으로 풀 수 있는 시장 크기이므로 염려할 이유가 없어 보인다.
@@
 \else
-본 연구는 새로운 조합최적화 문제를 제안했으며 이를 대학 지원 최적화 문제라고 부른다.
+본 연구는 새로운 조합 최적화 문제를 제안했으며 이를 대학 지원 최적화 문제라고 부른다.
 일종의 정수 자산 배분 문제 또는 확률적인 배낭 문제로 볼 수 있다.
@@
 4가지 해법을 제시했으며 이론적으로 가장 효율적인 알고리즘은 $O(m^3 / \varepsilon)$-시간 안에  $(1-\varepsilon)$-근사해를 계산하는 FPTAS이다.
-그 반면에 전형적인 대학 지원 문제 인스턴스에 대해, 지원 비용 지출액 기반 동적 계획 해법은 실행하기 쉬울 뿐더러 계산 시간이 상당히 짧다.
+그 반면에 전형적인 대학 지원 문제 인스턴스에 대해, 지원 비용 지출액 기반 동적 계획 해법은 실행하기 쉬울뿐더러 계산 시간이 상당히 짧다.
 계산 실험에서 모의 담금질 기반 휴리스틱 해법도 좋은 성능을 발휘했다.
@@
 ``상향·소신·안정'' 지원 학교를 대략 균일하게 분산하여 지원하는 것이 최적이란 통념이다 (전민희 2015).
-입학 시장의 크기에 비한 지원 예산이 작은 경우, 본 연구에서 제시하는 알고리즘은 비슷한 전략을 추천한다 (항 \ref{planetsexamplesection}의 예제 참고).
-그러나 포트폴리오에 이미 삽입한 학교에 비한 한계 가치를 반영하여 효용 모수를 할인하는 수식 \eqref{howtotransformfj}을 살펴보면, 효용이 높은 학교보다 효용이 낮은 학교에 부과되는 페널티가 큰 것을 알 수 있다.
+입학 시장의 크기에 비해 지원 예산이 작은 경우, 본 연구에서 제시하는 알고리즘은 비슷한 전략을 추천한다 (항 \ref{planetsexamplesection}의 예제 참고).
+그러나 포트폴리오에 이미 삽입한 학교를 반영한 한계 가치를 나타내도록 효용 모수를 할인하는 수식 \eqref{howtotransformfj}을 살펴보면, 효용이 높은 학교보다 효용이 낮은 학교에 부과되는 페널티가 큰 것을 알 수 있다.
 그 결과는 지원 예산이 커질수록 본 모형의 최적 포트폴리오가 안정 지원 학교보다 상향 지원 학교를 편드는 것이다.
-(그림 \ref{samplemarket}에서 뚜렷히 보이는 현상이다.)
+(그림 \ref{samplemarket}에서 뚜렷이 보이는 현상이다.)
 
@@
 이는 학생이 상위권 대학만 지원하지 않도록 만들어지기 때문에 다각화(diversification) 제약식으로 볼 수 있다.
-그러나 대학이 어떤 군에 속하는 것은 대학 순위일 뿐더러 입학 전형 유형도 의미한다.
+그러나 대학이 어떤 군에 속하는 것은 대학 순위일뿐더러 입학 전형 유형도 의미한다.
 가군과 나군으로 지원하는 경우, 다군보다 수능 점수의 반영 비율이 높다.
@@
 \end{align}
-다중 배낭 문제의 제약식과 같을 꼴이다 (Kellerer et al. 2004, 10제). 
+다중 배낭 문제의 제약식과 같은 꼴이다 (Kellerer et al. 2004, 10제). 
 \fi

### 2022-04-19 src/HeuristicSolvers.jl

@@ -1,10 +1,4 @@
-"""
-    optimalportfolio_greedy(mkt::VariedCostsMarket) -> X, v
-
-Use a greedy heuristic that adds schools in decreasing order of `mkt.ft ./ mkt.g`
-to compute an approximately optimal portfolio for the market `mkt` with varying application costs.
-"""
-function optimalportfolio_greedy(mkt::VariedCostsMarket{T})::Tuple{Vector{Int},Float64} where {T}
+function optimalportfolio_greedy_nopermute(mkt::VariedCostsMarket{T})::Vector{T} where {T}
     priority_order = filter(
         sortperm(mkt.ft ./ mkt.g, rev=true)
     ) do j
@@ -21,6 +15,21 @@ function optimalportfolio_greedy(mkt::VariedCostsMarket{T})::Tuple{Vector{Int},F
             break
         end
     end
+
+    # In the valuation we just use the identity permutation as invp
+    # To prevent wasteful permuting and then invpermuting
+    return X
+end
+
+
+"""
+    optimalportfolio_greedy(mkt::VariedCostsMarket) -> X, v
+
+Use a greedy heuristic that adds schools in decreasing order of `mkt.ft ./ mkt.g`
+to compute an approximately optimal portfolio for the market `mkt` with varying application costs.
+"""
+function optimalportfolio_greedy(mkt::VariedCostsMarket{T})::Tuple{Vector{Int},Float64} where {T}
+    X = optimalportfolio_greedy_nopermute(mkt)
 
     # In the valuation we just use the identity permutation as invp
     # To prevent wasteful permuting and then invpermuting
@@ -72,38 +81,13 @@ for `mkt` and its valuation.
 """
 function optimalportfolio_simulatedannealing(
     mkt::VariedCostsMarket{T};
-    X0::Union{Nothing,Vector{T}}=nothing,
+    X0::Vector{<:Integer}=optimalportfolio_greedy_nopermute(mkt),
     temp::Float64=0.25,
     nit::Integer=500,
     red::Float64=0.0625,
     verbose::Bool=false
 )::Tuple{Vector{Int},Float64} where {T}
-    if X0 === nothing
-        # Default to the greedy solution. We code it separately so we can get
-        # X in permuted order rather than original.
-        X, v = begin
-            priority_order = filter(
-                sortperm(mkt.ft ./ mkt.g, rev=true)
-            ) do j
-                mkt.g[j] ≤ mkt.H
-            end
-
-            X = T[]
[…]

### 2022-04-19 src/MarketIO.jl

@@ -192,15 +192,12 @@ a `SameCostsMarket` or a `VariedCostsMarket`.
 function valuation(
     X::AbstractVector{<:Integer},
     mkt::Union{SameCostsMarket{T},VariedCostsMarket{T}};
-    invp::Union{Nothing,AbstractVector{T}}=nothing)::Float64 where {T<:Unsigned}
+    invp::AbstractVector{T}=invperm(mkt.perm)
+)::Float64 where {T<:Unsigned}
     isempty(X) && return 0.0
 
     X = T.(X)
 
-    # sort!(X)
-    if isnothing(invp)
-        invp = invperm(mkt.perm)
-    end
     X[:] = sort(invp[X])
 
     h = T(length(X))

### 2022-04-19 src/VariedCostsBnB.jl

@@ -46,15 +46,15 @@ mutable struct Node{T<:Unsigned}
     end
 end
 
-hash(nd::Node) = hash((nd.I, nd.N))
+
+handle(nd::Node) = Pair(hash(nd.I), hash(nd.N))
 
 
 
 """
     generatechildren(nd, mkt)
 
-With respect to the data in `mkt`, generates the child(ren) of node `nd` and writes 
-their hashes to `nd.children`.
+With respect to the data in `mkt`, generates the child(ren) of node `nd`.
 """
 function generatechildren(nd::Node{T}, mkt::VariedCostsMarket{T})::Vector{Node{T}} where {T}
     fltr = filter(j -> mkt.g[j] ≤ nd.H̄, nd.N)
@@ -137,8 +137,8 @@ function optimalportfolio_branchbound(mkt::VariedCostsMarket{T}; maxit::Int=1000
 
     # push!(treekeys, push!(tree, rootnode))
 
-    tree = Dict{UInt,Node{T}}()
-    push!(tree, hash(rootnode) => rootnode)
+    tree = Dict{Pair{UInt, UInt},Node{T}}()
+    push!(tree, handle(rootnode) => rootnode)
 
     for k in 1:maxit
         verbose && @show k, LB, length(tree)
@@ -151,10 +151,10 @@ function optimalportfolio_branchbound(mkt::VariedCostsMarket{T}; maxit::Int=1000
             # thisnode, thisnodehandle = top_with_handle(tree)
             # pop!(tree)
             # delete!(treekeys, thisnodehandle)
-
-            thisnodehandle::UInt, thisnode::Node{T} = argmax(hash_nd -> hash_nd[2].v_LP, tree)
+        
+            thisnodehandle::Pair{UInt,UInt}, thisnode::Node{T} = argmax(handle_nd -> handle_nd[2].v_LP, tree)
             delete!(tree, thisnodehandle)
-
+        
             # Another option: Select the node with best obj value. Works pretty bad. 
             # thisnodehandle, thisnode = argmax(hash_nd -> hash_nd[2].v_I, tree)
             # delete!(tree, thisnodehandle)
@@ -172,7 +172,7 @@ function optimalportfolio_branchbound(mkt::VariedCostsMarket{T}; maxit::Int=1000
 
             # isempty(child.N) || push!(treekeys, push!(tree, child))
             if child.v_LP > LB && !isempty(child.N)
-                push!(tree, hash(child) => child)
+                push!(tree, handle(child) => child)
             end
         end
 

### 2022-04-19 test/runtests.jl

@@ -7,8 +7,6 @@ using Random
 # Number of markets to generate for each test
 const n_markets = 3
 
-println("Number of threads: ", Threads.nthreads())
-
 @testset verbose = true "OptimalApplication.jl" begin
 
     @testset verbose = true "Same app. costs" begin

### 2022-04-19 benchmark/benchmark.jl
export english

@@ -8,7 +8,7 @@ using Base.Threads
 import Printf: @sprintf
 using UnicodePlots
 
-const fullscale = false
+const fullscale = true
 
 # A long benchmark; tweak parameters with caution.
 # Set fullscale = false to run a smaller benchmark to check formatting etc.

### 2022-04-22 src/VariedCostsBnB.jl
Branch and bound: use vec instead of dict for Node.t; store tree as max heap

@@ -6,49 +6,50 @@ are "negotiable" and used to generate an LP upper bound and child nodes.
 mutable struct Node
     I::Vector{Int}
     N::Vector{Int}
-    t̄::Dict{Int,Float64}
-    H̄::Int
+    t_bar::Vector{Float64}
+    H_bar::Int
     v_I::Float64
     v_LP::Float64
 
-    function Node(I::Vector{Int}, N::Vector{Int}, t̄::Dict{Int,Float64}, H̄::Int, v_I::Float64, mkt::VariedCostsMarket)
-        # For generating a new node with its LP relaxation value and empty child set
-
+    "Generate a new node with its LP relaxation value and empty child set."
+    function Node(
+        I::Vector{Int},
+        N::Vector{Int},
+        t_bar::Vector{Float64},
+        H_bar::Int,
+        v_I::Float64,
+        mkt::VariedCostsMarket
+    )
         # Leaf node
-        if isempty(N)
-            return new(I, N, t̄, H̄, v_I, v_I)
-        end
-
-        j_order = copy(N)
+        isempty(N) && new(I, N, t_bar, H_bar, v_I, v_I)
 
         sort!(
-            j_order,
+            N,
             by=function (j)
-                mkt.f[j] * t̄[j] / mkt.g[j]
+                mkt.f[j] * t_bar[j] / mkt.g[j]
             end,
             rev=true,
-            alg=InsertionSort # Because typically just one item is out of place
+            alg=InsertionSort # Typically just one item is out of place
         )
 
         v_LP = v_I
-        H_left = H̄
-        for j in j_order
+        H_left = H_bar
+        for j in N
             if mkt.g[j] ≤ H_left
                 H_left -= mkt.g[j]
-                v_LP += mkt.f[j] * t̄[j]
+                v_LP += mkt.f[j] * t_bar[j]
             else
-                v_LP += mkt.f[j] * t̄[j] * H_left / mkt.g[j]
+                v_LP += mkt.f[j] * t_bar[j] * H_left / mkt.g[j]
                 break
             end
         end
 
-        return new(I, N, t̄, H̄, v_I, v_LP)
+        return new(I, N, t_bar, H_bar, v_I, v_LP)
     end
 end
 
 
-handle(nd::Node) = Pair(hash(nd.I), hash(nd.N))
-
[…]

### 2022-04-22 src/VariedCostsBnB.jl
Branch and bound: use vec instead of dict for Node.t; store tree as max heap

@@ -34,7 +34,7 @@ mutable struct Node
 
         v_LP = v_I
         H_left = H_bar
-        for j in N
+        @inbounds for j in N
             if mkt.g[j] ≤ H_left
                 H_left -= mkt.g[j]
                 v_LP += mkt.f[j] * t_bar[j]

### 2022-04-22 benchmark/benchmark.jl
more efficient filtering in greedy

@@ -55,13 +55,8 @@ function collectmeanstd(df::DataFrame)
 end
 
 
-function printheader(s)
-    printstyled(s * "\n", bold=true, color=222)
-end
-
-
 function benchmark1(marketsizes_SCM = marketsizes_SCM)
-    printheader("Benchmark 1: Homogeneous-cost algorithms")
+    @info "Benchmark 1: Homogeneous-cost algorithms"
     mkts_SCM = SameCostsMarket[SameCostsMarket(m) for m in marketsizes_SCM, i in 1:n_markets]
 
     sizes = Int[m for m in marketsizes_SCM, _ in 1:n_markets]
@@ -90,7 +85,7 @@ end
 
 
 function benchmark2(marketsizes_VCM = marketsizes_VCM)
-    printheader("Benchmark 2: Heterogeneous-cost algorithms")
+    @info "Benchmark 2: Heterogeneous-cost algorithms"
     epsilons = [0.5, 0.05]
     mkts_VCM = VariedCostsMarket[VariedCostsMarket(m) for m in marketsizes_VCM, i in 1:n_markets]
 
@@ -144,7 +139,7 @@ end
 
 
 function doeverything()
-    println("Number of threads: ", Threads.nthreads())
+    @info "Benchmark parameters" fullscale n_markets n_reps nthreads()
     println()
     bm1, plts1, = benchmark1()
     display(pretty_table(bm1, formatters=fmter))

### 2022-04-22 src/HeuristicSolvers.jl

@@ -1,9 +1,12 @@
 function optimalportfolio_greedy_nopermute(mkt::VariedCostsMarket)::Vector{Int}
-    priority_order = filter(
-        sortperm(mkt.ft ./ mkt.g, rev=true)
-    ) do j
-        mkt.g[j] ≤ mkt.H
-    end
+    priority_order = [j for j in 1:mkt.m if mkt.g[j] ≤ mkt.H]
+    sort!(
+        priority_order,
+        by=function (j)
+            mkt.ft[j] / mkt.g[j]
+        end,
+        rev=true
+    )
 
     X = Int[]
     H = mkt.H

### 2022-04-22 test/runtests.jl

@@ -2,8 +2,6 @@ using OptimalApplication
 using Test
 using Random
 
-# "t not sorted" warnings are OK. 
-
 # Number of markets to generate for each test
 const n_markets = 3
 
@@ -54,12 +52,9 @@ const n_markets = 3
 
                 X_enum, vX_enum = optimalportfolio_enumerate(mkt)
                 sort!(X_enum)
-                # X_valtable, VX_valtable = optimalportfolio_valuationtable(mkt)
                 X_dp, vX_dp = optimalportfolio_dynamicprogram(mkt)
                 X_bnb, vX_bnb = optimalportfolio_branchbound(mkt)
 
-                # @test X_enum == sort(X_valtable)
-                # @test vX_enum ≈ VX_valtable[mkt.m, mkt.H]
                 @test X_enum == sort(X_dp)
                 @test vX_enum ≈ vX_dp
                 @test X_enum == sort(X_bnb)
@@ -82,8 +77,7 @@ const n_markets = 3
         end
     end
 
-    # These tests verify that permuting the problem data doesn't
-    # change the optimum
+    # Verify that permuting the problem data doesn't change the optimum
     @testset verbose = true "Permutation" begin
         @testset verbose = true "Same costs" begin
             m = 6
@@ -148,9 +142,6 @@ const n_markets = 3
             @test v ≈ valuation(optimalportfolio_dynamicprogram(mkt1)[1], mkt1)
             @test v ≈ valuation(optimalportfolio_dynamicprogram(mkt2)[1], mkt2)
             @test v ≈ valuation(optimalportfolio_dynamicprogram(mkt3)[1], mkt3)
-            # @test v ≈ valuation(optimalportfolio_valuationtable(mkt1)[1], mkt1)
-            # @test v ≈ valuation(optimalportfolio_valuationtable(mkt2)[1], mkt2)
-            # @test v ≈ valuation(optimalportfolio_valuationtable(mkt3)[1], mkt3)
             @test v ≈ valuation(optimalportfolio_enumerate(mkt1)[1], mkt1)
             @test v ≈ valuation(optimalportfolio_enumerate(mkt2)[1], mkt2)
 
@@ -175,7 +166,6 @@ const n_markets = 3
             mkt = VariedCostsMarket(fill(0.5, 4), fill(2, 4), fill(4, 4), 16)
 
[…]

### 2022-04-22 src/ApproximateSolvers.jl
comment garbage collection, formatting code

@@ -68,7 +68,7 @@ with valuation `v`, for the [`VariedCostsMarket`](@ref) defined by `mkt`.
 function optimalportfolio_fptas(
     mkt::VariedCostsMarket,
     ε::Float64;
-verbose::Bool=false
+    verbose::Bool=false
 )::Tuple{Vector{Int},Float64}
     sp = ScaleParams(mkt, ε)
 
@@ -76,15 +76,12 @@ verbose::Bool=false
     sizehint!(G_dict, mkt.m * mkt.m ÷ 2)
 
     # Binary search for max{ w :  G_recursor!(G_dict, mkt.m, w, mkt, sp ≤ H }
-
     # Should be able to use something like this with a proper by kw in searchsortedlast
     # vs = 0:sp.Ū
-    # searchsortedlast(vs)
-
+    # searchsortedlast(vs, by = ...)
 
     v = 0
     v_UB = sp.Ū
-
     @inbounds while v + 1 < v_UB
         # mid = (v + v_UB) ÷ 2
         mid = midpoint(v, v_UB)
@@ -97,7 +94,6 @@ verbose::Bool=false
     end
 
     X = Int[]
-
     for j in reverse(1:mkt.m)
         # G_recursor!(G_dict, j, v, mkt, sp) < sp.infty &&
         if G_recursor!(G_dict, j, v, mkt, sp) < G_recursor!(G_dict, j - 1, v, mkt, sp)
@@ -118,7 +114,5 @@ verbose::Bool=false
         display(G_table)
     end
 
-    # In the valuation we just use the identity permutation as invp
-    # To prevent wasteful permuting and then invpermuting
     return mkt.perm[X], valuation_nopermute(X, mkt)
 end

### 2022-04-22 src/EnumerationSolvers.jl

@@ -17,7 +17,6 @@ function optimalportfolio_enumerate(mkt::SameCostsMarket)::Tuple{Vector{Int},Flo
     X = zeros(Int, mkt.h)
     v = 0.0
 
-    # invp = invperm(mkt.perm)
     for Y in multiset_combinations(1:mkt.m, mkt.h)
         if (w = valuation_nopermute(Y, mkt)) > v
             v = w
@@ -49,7 +48,6 @@ function optimalportfolio_enumerate(mkt::VariedCostsMarket)::Tuple{Vector{Int},F
     X = Int[]
     v = 0.0
 
-    # invp = invperm(mkt.perm)
     for Y in combinations(1:mkt.m)
         if (w = valuation_nopermute(Y, mkt)) > v && sum(mkt.g[Y]) ≤ mkt.H
             v = w

### 2022-04-22 src/FastSolvers.jl

@@ -163,39 +163,6 @@ function applicationorder_heap(mkt::SameCostsMarket)::Tuple{Vector{Int},Vector{F
 end
 
 
-# """
-#     optimalportfolio_valuationtable(mkt::VariedCostsMarket) -> X, V
-
-# Use the dynamic program on application costs to produce the optimal portfolio `X` and associated
-# valuation table `V` for the market `mkt` with varying application costs.
-# """
-# function optimalportfolio_valuationtable(mkt::VariedCostsMarket{T})::Tuple{Vector{Int},Matrix{Float64}} where {T}
-#     V = zeros(mkt.m, mkt.H)
-#     @inbounds for j in oneunit(T):mkt.m, h in 1:mkt.H
-#         if h < mkt.g[j]
-#             V[j, h] = get(V, (j - 1, h), 0)
-#         else
-#             V[j, h] = max(
-#                 get(V, (j - 1, h), 0),
-#                 mkt.omf[j] * get(V, (j - 1, h - mkt.g[j]), 0) + mkt.ft[j]
-#             )
-#         end
-#     end
-
-#     h = mkt.H
-#     X = T[]
-#     for j in reverse(1:mkt.m)
-#         if get(V, (j - 1, h), 0) < get(V, (j, h), 0)
-#             push!(X, j)
-#             h -= mkt.g[j]
-#         end
-#     end
-
-#     issorted(mkt.perm) || @warn "t not sorted; table rows will differ from input"
-#     return mkt.perm[X], V
-# end
-
-
 # Used by dynamic program below
 @inbounds function V_recursor!(
     V_dict::Dict{Tuple{Int,Int},Float64},

### 2022-04-22 src/HeuristicSolvers.jl

@@ -43,7 +43,7 @@ Generate a random neighbor of `X` for the market `mkt`. If `X` is feasible,
 preserve feasibility; if not, move to a feasible solution. `X` is assumed
 to refer to the sorted indices of the market. 
 """
-function neighbor(X::AbstractVector{Int}, mkt::VariedCostsMarket)::Vector{Int}
+function neighbor(X::AbstractVector{<:Integer}, mkt::VariedCostsMarket)::Vector{Int}
     X_neighbor = copy(X)
     Y = setdiff(1:mkt.m, X)
     shuffle!(X_neighbor)

### 2022-04-22 src/OptimalApplication.jl

@@ -16,7 +16,6 @@ export
     applicationorder_list,
     optimalportfolio_dynamicprogram,
     optimalportfolio_enumerate,
-    # optimalportfolio_valuationtable,
     optimalportfolio_fptas,
     optimalportfolio_branchbound,
     optimalportfolio_greedy,

### 2022-04-22 src/EnumerationSolvers.jl
use maximum() instead of for loop

@@ -14,14 +14,8 @@ julia> optimalportfolio_enumerate(mkt)
 function optimalportfolio_enumerate(mkt::SameCostsMarket)::Tuple{Vector{Int},Float64}
     mkt.m ≥ 21 && @warn "Enumeration is slow for large markets"
 
-    X = zeros(Int, mkt.h)
-    v = 0.0
-
-    for Y in multiset_combinations(1:mkt.m, mkt.h)
-        if (w = valuation_nopermute(Y, mkt)) > v
-            v = w
-            X[:] = Y
-        end
+    v, X = maximum(multiset_combinations(1:mkt.m, mkt.h)) do X
+        return valuation_nopermute(X, mkt), X
     end
 
     return mkt.perm[X], v
@@ -45,14 +39,8 @@ julia> optimalportfolio_enumerate(mkt)
 function optimalportfolio_enumerate(mkt::VariedCostsMarket)::Tuple{Vector{Int},Float64}
     mkt.m ≥ 21 && @warn "Enumeration is slow for large markets"
 
-    X = Int[]
-    v = 0.0
-
-    for Y in combinations(1:mkt.m)
-        if (w = valuation_nopermute(Y, mkt)) > v && sum(mkt.g[Y]) ≤ mkt.H
-            v = w
-            X = Y
-        end
+    v, X = maximum(X for X in combinations(1:mkt.m) if sum(mkt.g[X]) ≤ mkt.H) do X
+        return valuation_nopermute(X, mkt), X
     end
 
     return mkt.perm[X], v

### 2022-04-22 paper/paper.tex
export english
docs

@@
 \newif\ifen
-\enfalse% Change to false to export Korean
+\entrue% Change to false to export Korean
 
@@
 \else
-위 논의에 따라 최적성이 성립한다. $t$를 배열하는 시간은 $O(m \log m)$이다. 목병 단계는 \ref{Vcreatedlookuptable}줄에서 $V[j, h]$의 표를 만드는 것이다. 각 원소를 단위 시간 안에 생성할 수 있으며 표의 크기가 $O(Hm)$이다.
+위 논의에 따라 최적성이 성립한다. $t$를 배열하는 시간은 $O(m \log m)$이다. 병목 단계는 \ref{Vcreatedlookuptable}줄에서 $V[j, h]$의 표를 만드는 것이다. 각 원소를 단위 시간 안에 생성할 수 있으며 표의 크기가 $O(Hm)$이다.
 \fi
@@
 \else
-(계산 시간.) \ref{createdlookuptable}줄에서 표를 채우는 것이 목병 단계이며 표의 크기는 $m \times |\mathcal{V}|$이다. 그러면
+(계산 시간.) \ref{createdlookuptable}줄에서 표를 채우는 것이 병목 단계이며 표의 크기는 $m \times |\mathcal{V}|$이다. 그러면
 \fi
@@
 실험 변수의 각 조합에 대해,  \nmarkets 개의 시장을 생성했으며 최적 포트폴리오를 3번 계산 한 것 중 가장 빠른 것으로 계산 시간으로 기록했다.
-\nmarkets 개의 시장에 대한 평균과 평균편차를 표에 나타낸다.
+\nmarkets 개의 시장에 대한 평균과 표준편차를 표에 나타낸다.
 따라서 아래 표의 각 칸은 {\the\numexpr 3 * \nmarkets \relax}개의 계산에 대한 통계다.
@@
 The table shows the average time (standard deviation) over the \nmarkets~instances.
-\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복하여 그중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (평균편차) 시간이 나타난다.\fi}
+\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복하여 그중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (표준편차) 시간이 나타난다.\fi}
 
@@
 익숙한 Markowitz 포트폴리오 최적화 모형에서는 명시적인 위험 회피 항이 등장하며, 본 모형은 maximax 목적함수에 내포된 경쟁적인 학교와 안정 지원 학교 사이의 균형에 의한 위험 관리 요소만 고려했다.
-그러나 기본 목적함수인 $v(\mathcal{X}) = \operatorname{E}[X]$에 평균편차 페널티 $\beta \geq 0$를 다음처럼 도입할 수 있다:
+그러나 기본 목적함수인 $v(\mathcal{X}) = \operatorname{E}[X]$에 표준편차 페널티 $\beta \geq 0$를 다음처럼 도입할 수 있다:
 \fi
@@
 &=   \operatorname{E}[X] - \beta \left(  \operatorname{E}[X^2]  -  \operatorname{E}[X]^2 \right) \\
-& = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)
+& = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{￣X}: \\ i > j}} (1 - f_{i}) \Bigr)
  - \beta \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j^2 \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)
@@
 \else
-수리 실험 결과에 따르면 동적 계획 해법 성능의 목병 요소는 계산 시간이 아니라 메모리 소모량이다. 이 알고리즘의 메모리 소요를 절감하면 상당히 큰 문제를 풀 수 있다.
+수리 실험 결과에 따르면 동적 계획 해법 성능의 병목 요소는 계산 시간이 아니라 메모리 소모량이다. 이 알고리즘의 메모리 소요를 절감하면 상당히 큰 문제를 풀 수 있다.
 

### 2022-04-22 docs/make.jl

@@ -11,19 +11,19 @@ makedocs(
         "Home"=>"index.md",
         "Tutorial"=>[
             "Market I/O" => "tutorial/market_io.md",
-            "Same-costs markets" => "tutorial/same_costs.md",
-            "Varied-costs markets" => "tutorial/varied_costs.md",
+            "Same-costs problem" => "tutorial/same_costs.md",
+            "Varied-costs problem" => "tutorial/varied_costs.md",
         ],
         "Public API"=>[
             "Market I/O" => "public_api/market_io.md",
             "Enumeration solvers" => "public_api/enumeration_solvers.md",
-            "Fast solvers" => "public_api/fast_solvers.md",
+            "Fast, exact solvers" => "public_api/fast_solvers.md",
             "Approximation methods" => "public_api/approximation.md",
             "Heuristic solvers" => "public_api/heuristics.md",
         ],
-        # "Examples"=>[
-        #     "Valuation curve" => "examples/valuation.md",
-        #     "Varied costs" => "examples/varied_costs.md",
-        # ],
+        "Examples"=>[
+            "Valuation curve" => "examples/valuation.md",
+            "Reach/safety schools" => "examples/reach_safety.md",
+        ],
     ],
 )

### 2022-04-22 paper/plots/plots.jl

@@ -23,17 +23,17 @@ function comparativestatics()
 
     df = DataFrame("f" => mkt.f, "t" => mkt.t, "order" => invpermapporder, "v" => v[invpermapporder])
     #=
-    Row │ f        t        order  v       
-        │ Float64  Float64  Int64  Float64 
+     Row │ f        t        order  v       
+         │ Float64  Float64  Int64  Float64 
     ─────┼──────────────────────────────────
-    1 │    0.39    200      4  230.047
-    2 │    0.33    250      2  146.7
-    3 │    0.24    300      6  281.513
-    4 │    0.24    350      1   84.0
-    5 │    0.05    400      7  288.778
-    6 │    0.03    450      8  294.106
-    7 │    0.1     500      5  257.643
-    8 │    0.12    550      3  195.096
+       1 │    0.39    200      4  230.047
+       2 │    0.33    250      2  146.7
+       3 │    0.24    300      6  281.513
+       4 │    0.24    350      1   84.0
+       5 │    0.05    400      7  288.778
+       6 │    0.03    450      8  294.106
+       7 │    0.1     500      5  257.643
+       8 │    0.12    550      3  195.096
     =#
     display(df)
 

### 2022-04-22 src/MarketIO.jl

@@ -96,9 +96,9 @@ struct SameCostsMarket{U<:Real} <: Market
     end
 
     function SameCostsMarket(m::Integer)
-        t = ceil.(Int, -10 * log.(rand(m)))
+        t = ceil.(Int, -10 * log.(rand(m))) |> sort
         f = inv.(t .+ 10 * rand(m))
-        perm = sortperm(t)
+        perm = 1:m
         return new{Int}(
             m,
             f[perm],
@@ -188,11 +188,11 @@ struct VariedCostsMarket <: Market
     end
 
     function VariedCostsMarket(m::Integer)
-        t = ceil.(Int, -10 * log.(rand(m)))
+        t = ceil.(Int, -10 * log.(rand(m))) |> sort
         f = inv.(t .+ 10 * rand(m))
         g = rand(5:10, m)
         H = sum(g) ÷ 2
-        perm = sortperm(t)
+        perm = 1:m
 
         return new(
             m,

### 2022-04-25 docs/make.jl
factor out sort in valuation_nopermute_sorted documentation remove unused deps

@@ -6,6 +6,7 @@ DocMeta.setdocmeta!(OptimalApplication, :DocTestSetup, :(using OptimalApplicatio
 makedocs(
     sitename="OptimalApplication",
     modules=[OptimalApplication],
+    format=Documenter.HTML(ansicolor=true),
     # format = Documenter.HTML(prettyurls = false), # For local browsing
     pages=Any[
         "Home"=>"index.md",
@@ -23,7 +24,7 @@ makedocs(
         ],
         "Examples"=>[
             "Valuation curve" => "examples/valuation.md",
-            "Reach/safety schools" => "examples/reach_safety.md",
+            "Reach & safety schools" => "examples/reach_safety.md",
         ],
     ],
 )

### 2022-04-25 src/ApproximateSolvers.jl

@@ -94,7 +94,7 @@ function optimalportfolio_fptas(
     end
 
     X = Int[]
-    for j in reverse(1:mkt.m)
+    @inbounds for j in reverse(1:mkt.m)
         # G_recursor!(G_dict, j, v, mkt, sp) < sp.infty &&
         if G_recursor!(G_dict, j, v, mkt, sp) < G_recursor!(G_dict, j - 1, v, mkt, sp)
             push!(X, j)
@@ -114,5 +114,5 @@ function optimalportfolio_fptas(
         display(G_table)
     end
 
-    return mkt.perm[X], valuation_nopermute(X, mkt)
+    return mkt.perm[X], valuation_nopermute_sorted(sort(X), mkt)
 end

### 2022-04-25 src/EnumerationSolvers.jl

@@ -15,7 +15,7 @@ function optimalportfolio_enumerate(mkt::SameCostsMarket)::Tuple{Vector{Int},Flo
     mkt.m ≥ 21 && @warn "Enumeration is slow for large markets"
 
     v, X = maximum(multiset_combinations(1:mkt.m, mkt.h)) do X
-        return valuation_nopermute(X, mkt), X
+        return valuation_nopermute_sorted(X, mkt), X
     end
 
     return mkt.perm[X], v
@@ -40,7 +40,7 @@ function optimalportfolio_enumerate(mkt::VariedCostsMarket)::Tuple{Vector{Int},F
     mkt.m ≥ 21 && @warn "Enumeration is slow for large markets"
 
     v, X = maximum(X for X in combinations(1:mkt.m) if sum(mkt.g[X]) ≤ mkt.H) do X
-        return valuation_nopermute(X, mkt), X
+        return valuation_nopermute_sorted(X, mkt), X
     end
 
     return mkt.perm[X], v

### 2022-04-25 src/FastSolvers.jl

@@ -43,10 +43,7 @@ function applicationorder_list(
 )::Tuple{Vector{Int},Vector{Float64}}
     apporder = zeros(Int, mkt.h)
     v = zeros(mkt.h)
-
     mkt_list = College[College(j, mkt.f[j], mkt.t[j], mkt.ft[j], mkt.omf[j]) for j in 1:mkt.m]
-
-    dummy_college = College(0, 1.0, -1.0, -1.0, 0.0)
 
     c_best::College, idx_best::Int = findmax(mkt_list)
     @inbounds for j in 1:mkt.h
@@ -59,7 +56,7 @@ function applicationorder_list(
         v[j] = get(v, j - 1, 0) + c_best.ft
         apporder[j] = c_best.j
 
-        next_c_best::College = dummy_college
+        next_c_best::College = College(0, 1.0, -1.0, -1.0, 0.0)
         next_idx_best::Int = 0
 
         for i in 1:idx_best-1
@@ -141,8 +138,8 @@ function applicationorder_heap(mkt::SameCostsMarket)::Tuple{Vector{Int},Vector{F
             #     College(
             #         c.j,
             #         c.f,
-            #         c.t < c_k.t ? c.t * c_k.omf : c.t - c_k.ft,
-            #         c.t < c_k.t ? c.ft * c_k.omf : c.ft - c.f * c_k.ft,
+            #         c.j < c_k.j ? c.t * c_k.omf : c.t - c_k.ft,
+            #         c.j < c_k.j ? c.ft * c_k.omf : c.ft - c.f * c_k.ft,
             #         c.omf
             #     )
             #     for c in mkt_heap.valtree if c.j != c_k.j
@@ -151,8 +148,8 @@ function applicationorder_heap(mkt::SameCostsMarket)::Tuple{Vector{Int},Vector{F
                 College(
                     c.j,
                     c.f,
-                    c.t < c_k.t ? c.t * c_k.omf : c.t - c_k.ft,
-                    c.t < c_k.t ? c.ft * c_k.omf : c.ft - c.f * c_k.ft,
+                    c.j < c_k.j ? c.t * c_k.omf : c.t - c_k.ft,
+                    c.j < c_k.j ? c.ft * c_k.omf : c.ft - c.f * c_k.ft,
                     c.omf
                 )
             end

### 2022-04-25 src/HeuristicSolvers.jl

@@ -32,7 +32,8 @@ defined by `mkt`.
 """
 function optimalportfolio_greedy(mkt::VariedCostsMarket)::Tuple{Vector{Int},Float64}
     X = optimalportfolio_greedy_nopermute(mkt)
-    return mkt.perm[X], valuation_nopermute(X, mkt)
+    sort!(X)
+    return mkt.perm[X], valuation_nopermute_sorted(X, mkt)
 end
 
 
@@ -86,13 +87,15 @@ function optimalportfolio_simulatedannealing(
     verbose::Bool=false
 )::Tuple{Vector{Int},Float64}
     X = optimalportfolio_greedy_nopermute(mkt)
-    v = valuation_nopermute(X, mkt)
+    sort!(X)
+    v = valuation_nopermute_sorted(X, mkt)
 
     X_best, v_best = X, v
 
     for i in 1:nit
         X_neighbor = neighbor(X, mkt)
-        v_neighbor = valuation_nopermute(X_neighbor, mkt)
+        sort!(X_neighbor)
+        v_neighbor = valuation_nopermute_sorted(X_neighbor, mkt)
 
         if verbose
             println("Iteration $i, v_best = $v_best, v_neighbor = $v_neighbor")

### 2022-04-25 src/MarketIO.jl

@@ -227,20 +227,19 @@ function Market(f::Vector{Float64}, t::Vector{Int}, g::Vector{Int}, H::Int)
 end
 
 
-function valuation_nopermute(
+function valuation_nopermute_sorted(
     X::AbstractVector{<:Integer},
     mkt::Market;
 )::Float64
     isempty(X) && return 0.0
-    X = sort(X)
+    # @assert issorted(X)
     h = length(X)
     if h > 1
-        res = 0.0
+        res = mkt.ft[X[end]]
         cp = reverse(cumprod(reverse(mkt.omf[X[2:end]])))
         for j in 1:h-1
             res += mkt.ft[X[j]] * cp[j]
         end
-        res += mkt.ft[X[end]]
         return res
     else
         return mkt.ft[X[1]]
@@ -277,5 +276,5 @@ function valuation(
     mkt::Market;
     invp::AbstractVector{<:Integer}=invperm(mkt.perm)
 )::Float64
-    return valuation_nopermute(invp[X], mkt)
+    return valuation_nopermute_sorted(sort(invp[X]), mkt)
 end

### 2022-04-25 paper/paper.tex
add appendix about submodularity; rearrange some commented text to prepare for transplant of variable elimination technique from "Alma" to "preliminary" section

@@
 Unfortunately, the nestedness property does not hold in the general case, nor does the greedy algorithm offer any performance guarantee.
-Instead, we offer a branch-and-bound routine, a pseudopolynomial-time algorithm that is tractable for typical college market instances, and an approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in fully polynomial time.
+Instead, we offer a branch-and-bound routine, a pseudopolynomial-time algorithm that is tractable for typical college market instances, and a fully polynomial-time approximation scheme (FPTAS).
 \else
@@
 그러나 일반적인 경우에서 포함 사슬 관계가 성립하지 않을뿐더러 탐욕 알고리즘이 어떤 근사 계수를 보장할 수 없을 보일 수 있다.
-대신 분지한계법, 전형적인 입학 시장 인스턴스에 대해 효율적인 의사 다항 시간 동적 계획, 그리고 완전 다항 시간 안에 $(1 - \varepsilon)$-근사해를 출력하는 알고리즘 등 3가지 해법을 제안한다.
-\fi
+대신 분지한계법, 전형적인 입학 시장 인스턴스에 대해 효율적인 의사 다항 시간 동적 계획, 그리고 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme, FPTAS)을 제시한다.
+\fi
+
+% following sentence buries the lede
+% decide how to handle FPTAS abbrev
+Although our research employs more elementary techniques, we remark that the objective function of \ref{headlineproblem} is a nondecreasing submodular function, and our approximation results are tighter than those that arise in the generic case.
+Fisher et al.'s result (1978), for example, implies that the greedy algorithm is asymptotically $(1 - 1/e)$-optimal for the $g_1 = 1$ case of the college application problem, whereas we show that the same algorithm is exact.
+Using relaxation techniques, a nearly equivalent approximation ratio is achievable when maximizing a submodular set function over knapsack constraints or matroids (Calinescu et al.
+2011; Kulik et al.
+2013).
+In the college application problem, the existence of the FPTAS supersedes these results.
 
@@
 
+%% Work text from nestedness proof into here, or probably better to just rewrite
+
+%Now, we present a variable-elimination technique that will prove useful throughout the paper.
+Suppose that the student has already resolved to apply to $c_k$, and the remainder of her decision consists of determining which \emph{other} schools to apply to.
+We can write her total application portfolio as $\mathcal{X} = \mathcal{Y} \cup \{k\}$, and we are interested in devising a function $w(\mathcal{Y})$ that orders portfolios according to how well they ``complement'' the singleton portfolio $\{k\}$.
+To construct $w(\mathcal{Y})$ ..
+%this part is like the proof
+%
+%\begin{proposition}[\ifen Eliminate $c_k$\else $c_k$ 소거\fi]
+%\ifen For $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$, where
+%\else $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$에 대해, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$이다. 단,\fi
+%\begin{align}\label{howtotransformfj}
+%\bar t_j = 
+%\begin{cases}
+%(1 - f_k) t_j, \quad & t_j \leq t_k \\
+%t_j - f_k t_k, \quad& t_j > t_k.
+%\end{cases}
+%\end{align}
+%\end{proposition}%
+% It is easy to verify that this is the composition ...
+
+%% Prove submodularity
+
 \ifen \section{Homogeneous application costs}  \else \section{동일한 지원 비용} \fi \label{homogappcosts}
@@
 
-The basic error of the na\"ive algorithm is that it maximizes $\operatorname{E}\left[\sum t_j Z_j \right]$ instead of $\operatorname{E}\left[\max \{t_j Z_j\} \right]$.
+The basic error of the na\"ive algorithm is that it maximizes $\operatorname{E}\left[\,\sum t_j Z_j\, \right]$ instead of $\operatorname{E}\left[\max \{t_j Z_j\} \right]$.
 The latter is what Alma is truly concerned with, since in the end she can attend only one school.
@@
 
-나이브 알고리즘의 기본 실수는 $\operatorname{E}\left[\max \{t_j Z_j\} \right]$ 대신  $\operatorname{E}\left[\sum t_j Z_j \right]$을 최대화하는 것이다. 결국에는 단 한 학교로만 진학할 수 있으므로 알마의 목적은 후자가 아닌 전자다. %학교 개별 효용의 기대 최댓값이다. 
+나이브 알고리즘의 기본 실수는 $\operatorname{E}\left[\max \{t_j Z_j\} \right]$ 대신  $\operatorname{E}\left[\,\sum t_j Z_j\,\right]$을 최대화하는 것이다. 결국에는 단 한 학교로만 진학할 수 있으므로 알마의 목적은 후자가 아닌 전자다. %학교 개별 효용의 기대 최댓값이다. 
 다음 예는 나이브 해법이 최적해가 아닌 해를 출력할 수 있음을 보인다.
@@
 \fi
-%\begin{proposition}[\ifen Eliminate $c_k$\else $c_k$ 소거\fi]
-%\ifen For $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$, where
-%\else $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$에 대해, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$이다. 단,\fi
-%\begin{align}\label{howtotransformfj}
-%\bar t_j = 
-%\begin{cases}
-%(1 - f_k) t_j, \quad & t_j \leq t_k \\
-%t_j - f_k t_k, \quad& t_j > t_k.
-%\end{cases}
-%\end{align}
-%\end{proposition}%
 \begin{align}\label{howtotransformfj}
@@
 For each market, we computed a heuristic solution using Algorithm \ref{ellissimann} and an exact solution using Algorithm \ref{ellisDP1} and calculated the optimality ratio.
-We chose to use $N = 500$ iterations in the simulated-annealing algorithm, and the temperature parameters $T = 1/4$ and $r = 1/16$ were chosen by a preliminary grid search.
+We chose to use $N = 500$ iterations in the simulated-annealing algorithm, and the temperature parameters $T = 1/4$ and $r = 1/16$ were selected by a preliminary grid search.
 \else
@@
 \end{align}
-completes the formulation. These constraints are conformable to those of the multiple knapsack problem (Kellerer et al. 2004, chap. 10). 
+completes the formulation.
+These are knapsack constraints; therefore, the algorithm of Kulik et al.
+(2013) can be used to obtain a $(1 - 1/e + \varepsilon)$-approximate solution for this problem.
+When the budget constraint is a cardinality constraint (as in Alma's problem), we note that adding these constraints yields a matriodal solution set, and Calinescu et al.
+(2011)'s algorithm yields an alternative solution, albeit with the same approximation coefficient of $1 - 1/e + \varepsilon$.
+%These constraints are conformable to those of the multiple knapsack problem (Kellerer et al.
+2004, chap.
+10).
 \else
@@
 \end{align}
-다중 배낭 문제의 제약식과 같은 꼴이다 (Kellerer et al. 2004, 10제). 
+These are knapsack constraints; therefore, the algorithm of Kulik et al.
+(2013) can be used to obtain a $(1 - 1/e + \varepsilon)$-approximate solution for this problem.
+When the budget constraint is a cardinality constraint (as in Alma's problem), we note that adding these constraints yields a matriodal solution set, and Calinescu et al.
+(2011)'s algorithm yields an alternative solution, albeit with the same approximation coefficient of $1 - 1/e + \varepsilon$.
+%다중 배낭 문제의 제약식과 같은 꼴이다 (Kellerer et al. 2004, 10제). 
 \fi
@@
 \fi
+
+\section{Appendix}
+
+N.B. these results will be worked into the body soon.
+
+\begin{definition}[Submodular set function]
+Given a ground set $\mathcal{C}$ and function $v : 2^{\mathcal{C}} \mapsto \mathbb{R}$, $v(\mathcal{X})$ is called a \emph{submodular set function} if and only if $v(\mathcal{X}) + v(\mathcal{Y}) \geq v(\mathcal{X}\cup\mathcal{Y}) + v(\mathcal{X}\cap\mathcal{Y})$
+for all $\mathcal{X}, \mathcal{Y} \subseteq \mathcal{C}$.
+Furthermore, if $ v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) \geq 0$ for all $\mathcal{X} \subset \mathcal{C}$ and $k \in \mathcal{C} \setminus \mathcal{X}$, $v(\mathcal{X})$ is said to be a \emph{nondecreasing} submodular set function.
+\end{definition}
+
+\begin{theorem}
+The college application portfolio valuation function
+$v(\mathcal{X})$ % = \sum_{j\in\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)\]
+is a nondecreasing submodular set function.
+\end{theorem}
+
+\begin{proof}
+It is self-evident that $v(\mathcal{X})$ is nondecreasing. To establish its submodularity, we apply proposition 2.1.iii of Nemhauser and Wolsey (1978) and show that
+\begin{equation}\label{nemhauseriii}
+v(\mathcal{X} \cup \{j\}) - v(\mathcal{X}) \geq 
+v(\mathcal{X} \cup \{j, k\}) - v(\mathcal{X} \cup \{k\})
+\end{equation}
+for $\mathcal{X} \subset \mathcal{C}$ and $j \neq k \in \mathcal{C} \setminus \mathcal{X}$.
+By repeatedly eliminating the schools in $\mathcal{X}$ according to \eqref{howtotransformfj}, we can obtain a portfolio valuation function $w(\mathcal{Y}; t)$ such that $w(\mathcal{Y}) = v(\mathcal{X} \cup \mathcal{Y}) + \text{const.}$ for any $\mathcal{Y} \subseteq \mathcal{C} \setminus \mathcal{X}$.
+Therefore, \eqref{nemhauseriii} is equivalent to
+\begin{align}
+& w(\{j\}) - w(\varnothing) \geq w(\{j, k\}) - w(\{k\}) \\
+\iff \qquad &w(\{j\})  +  w(\{k\})  \geq w(\{j, k\})  \\
+\iff \qquad &\operatorname{E}[\,\bar t_j Z_j\,] + \operatorname{E}[\,\bar t_k Z_k\,] 
+\geq \operatorname{E}\bigl[\max\{ \bar t_j Z_j, \bar t_k Z_k \} \bigr]
+\end{align}
+which is plainly true. 
+\end{proof}
 
@@
 
+% 1 - 1/e opt for the problem in the title
+Calinescu, Gruia, Chandra Chekuri, Martin Pál, and Jan Vondrák.
+2011.
+``Maximizing a Monotone Submodular Function Subject to a Matroid Constraint.'' \emph{SIAM Journal on Computing} 40 (6): 1740--66.
+\url{https://doi.org/10.1137/080733991}.
+
 Carraway, Robert, Robert Schmidt, and Lawrence Weatherford.
@@
 Kellerer, Hans, Ulrich Pferschy, and David Pisinger. 2004. \emph{Knapsack Problems.} Berlin: Springer.
+
+%1 - 1/ e for monotonic submodular with knapsack constraints, i.e. Ellis with 가나다
+Kulik, Ariel, Hadas Shachnai, and Tami Tamir.
+2013.
+``Approximations for Monotone and Nonmonotone Submodular Maximization with Knapsack Constraints.'' \emph{Mathematics of Operations Research} 38 (4): 729--39.
+\url{https://doi.org/10.1287/moor.2013.0592}.
 

### 2022-04-26 paper/paper.tex
move variable elimination technique to preliminaries section; revise nestedness proof and associated algorithm with according references; corresp translations; submodularity proof and notes on previous algorithms; slight changes to intro and conclusion reflecting submodularity

@@
 \newif\ifen
-\entrue% Change to false to export Korean
+\enfalse% Change to false to export Korean
 
@@
 특히 선호도가 높으며 붙기 어려운 “상향” 지원 학교(reach school)와 선호도가 낮으며 붙기 쉬운 “안정” 지원 학교(safety school) 사이의 균형을 고려해야 한다 (전민희 2015).
-마지막, 대학 지원의 조합적인 본성으로 인해 연속적인 포트폴리오 최적화 문제에서 흔히 사용하는 기울기 해법으로는 풀기가 어렵다.
+마지막, 대학 지원의 조합적인 본질로 인해 연속적인 포트폴리오 최적화 문제에서 흔히 사용하는 기울기 해법으로는 풀기가 어렵다.
 Fu는 지원 비용을 제약 조건 대신 목적함수의 한 항으로 모형화했으며, 균형 모형의 모수를 추정하기 위해 $m=8$이 되도록 학교들의 클러스터를 먼저 구성했다.
@@
 탐욕 알고리즘이 좋은 근사해거나 정확한 알고리즘이 되는 최적화 모형들이 알려져 있다.
-집합 크기 제약 하에서 하위 모듈 집합 함수(submodular set function)를 최대화하는 문제가 대표적인 예다 (Fisher 외 1978).
+집합 크기 제약 하에서 submodular한 집합 함수를 최대화하는 문제가 대표적인 예다 (Fisher 외 1978).
 반면에 이진 배낭 문제 같은 경우에서는 가장 직관적인 탐욕 알고리즘이 최적과 거리가 먼 해를 출력하는 예를 만들 수 있다 (Vazirani 2001).
@@
 
-% following sentence buries the lede
-% decide how to handle FPTAS abbrev
-Although our research employs more elementary techniques, we remark that the objective function of \ref{headlineproblem} is a nondecreasing submodular function, and our approximation results are tighter than those that arise in the generic case.
-Fisher et al.'s result (1978), for example, implies that the greedy algorithm is asymptotically $(1 - 1/e)$-optimal for the $g_1 = 1$ case of the college application problem, whereas we show that the same algorithm is exact.
-Using relaxation techniques, a nearly equivalent approximation ratio is achievable when maximizing a submodular set function over knapsack constraints or matroids (Calinescu et al.
+% following sentence buries the lede（
+\ifen
+Finally, we remark that the objective function of \ref{headlineproblem} is a nondecreasing submodular function.
+However, our research employs more elementary analytical techniques, and our approximation results are tighter than those associated with generic submodular maximization algorithms.
+For example, a well-known result of Fisher et al.
+(1978) implies that the greedy algorithm is asymptotically $(1 - 1/e)$-optimal for the $g_j = 1$ case of the college application problem, whereas we show that the same algorithm is exact.
+When maximizing a submodular set function over knapsack constraints or matroids, a nearly equivalent approximation ratio is achievable using relaxation techniques (Calinescu et al.
 2011; Kulik et al.
 2013).
-In the college application problem, the existence of the FPTAS supersedes these results.
+But in the general case of the college application problem, the existence of the FPTAS supersedes these results.
+\else
+마지막으로, \ref{headlineproblem}의 목적함수는 submodular하며 단조 증가하는 함수이다.
+그런데, 본 연구는 더 기초적인 분석 기술을 이용하며, 일반적인 submodular 최대화 알고리즘보다 좋은 근사 계수를 얻게 된다.
+예를 들어 Fisher 외 (1978)의 잘 알려진 결과에 따라, 대학 지원 문제의 $g_j = 1$인 경우에서 탐욕 해법의 점근적 근사 계수가 $(1 - 1/e)$이 되며, 본 논문에서 같은 해법이 정확함을 보인다.
+다중 배낭 제약식 혹은 매트로이드 위에서 submodular 집합 함수를 최대하는 문제에서, 완화 기술을 활용하면 거의 동등한 근사 계수를 이뤄 낼 수 있다 (Calinescu et al.
+2011; Kulik et al.
+2013).
+그러나 일반적인 대학 지원 문제에서, FPTAS의 존재성은 그 결과를 대체한다.
+\fi
 
@@
 \ifen Section \ref{preliminaries} introduces some additional notation and assumptions that can be imposed without loss of generality.
-\else \ref{preliminaries}절은 표기법과 일반성을 제한하지 않는 편의적 가정을 제시한다.\fi
+We also introduce a useful variable-elimination technique and prove that the objective function is submodular.
+\else \ref{preliminaries}절은 표기법과 일반성을 제한하지 않는 편의적 가정을 제시한다. 유용한 변수 소거 기술도 제시하고 목적함수가 submodularity함을 증명한다.\fi
 
@@
 \ifen
-In Section \ref{numericalexperiments}, we present the results of computational experiments that confirm the validity and time complexity results established in the previous two sections. % more
-
-A brief conclusion follows.
-\else
-\ref{numericalexperiments}절에서 위에서 도출한 타당성과 계산 복잡성을 확인하는 계산 실험 결과를 정리한다.
-
-그리고 간단한 결론을 제시한다.
+In Section \ref{numericalexperiments}, we present the results of computational experiments that confirm the validity and time complexity results established in the previous two sections.
+We also investigate the empirical accuracy of the simulated-annealing heuristic.
+
+In the conclusion, we identify possible improvements to our solution algorithms and introduce a few extensions of the model that capture the features of real-world admissions processes.
+\else
+\ref{numericalexperiments}절에서 위에서 도출한 타당성과 계산 복잡성을 확인하는 계산 실험 결과를 정리한다. 모의 담금질 휴리스틱의 정확도도 탐구한다.
+
+결론에서, 해법을 개선할 수 있는 요소를 언급하고, 실제 지원 과정의 특징을 묘사하도록 원래 모형을 확장할 수 있는 기술 몇 가지 소개한다.
 \fi
@@
 \fi
+
+\ifen \subsection{Refining the objective function} \else \subsection{목적함수의 정제} \fi
 
@@
 
-%% Work text from nestedness proof into here, or probably better to just rewrite
-
-%Now, we present a variable-elimination technique that will prove useful throughout the paper.
-Suppose that the student has already resolved to apply to $c_k$, and the remainder of her decision consists of determining which \emph{other} schools to apply to.
-We can write her total application portfolio as $\mathcal{X} = \mathcal{Y} \cup \{k\}$, and we are interested in devising a function $w(\mathcal{Y})$ that orders portfolios according to how well they ``complement'' the singleton portfolio $\{k\}$.
-To construct $w(\mathcal{Y})$ ..
-%this part is like the proof
-%
-%\begin{proposition}[\ifen Eliminate $c_k$\else $c_k$ 소거\fi]
-%\ifen For $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$, where
-%\else $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$에 대해, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$이다. 단,\fi
-%\begin{align}\label{howtotransformfj}
-%\bar t_j = 
-%\begin{cases}
-%(1 - f_k) t_j, \quad & t_j \leq t_k \\
-%t_j - f_k t_k, \quad& t_j > t_k.
-%\end{cases}
-%\end{align}
-%\end{proposition}%
-% It is easy to verify that this is the composition ...
-
-%% Prove submodularity
+\subsection{A variable-elimination technique}
+\ifen
+Now, we present a variable-elimination technique that will prove useful throughout the paper.\footnote{We thank Yim Seho for pointing out this useful transformation.} Suppose that the student has already resolved to apply to $c_k$, and the remainder of her decision consists of determining which \emph{other} schools to apply to.
+Writing her total application portfolio as $\mathcal{X} = \mathcal{Y} \cup \{k\}$, we devise a function $w(\mathcal{Y})$ that orders portfolios according to how well they ``complement'' the singleton portfolio $\{k\}$.
+Specifically, the difference between $v(\mathcal{Y} \cup\{k\})$ and $w(\mathcal{Y})$ is the constant $f_k t_k$.
+\else
+본 항에서 남은 논문에서 활용되는 변수 소거 기술을 소개한다.\footnote{이 유용한 변환을 발견한 임세호에게 감사를 표한다.} 학생이 $c_k$에 지원하기로 결심하고, 남은 결정 공간은 어떤 다른 학교에 지원하는 것으로 제한된다고 하자.
+전체 지원 포트폴리오를 $\mathcal{X} = \mathcal{Y} \cup \{k\}$일 때, $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$인 포트폴리오와 $\{k\}$ 사이의 \mbox{`보완성'을} 평가하는 함수 $w(\mathcal{Y})$를 도출할 수 있다.
+단, $v(\mathcal{Y} \cup\{k\})$와 $w(\mathcal{Y})$ 사이에서 $f_k t_k$ 같은 상수 차이가 발생한다.
+\fi
+
+\ifen
+To construct $w(\mathcal{Y})$, let $\tilde t_j$ denote the expected utility the student receives from school $c_j$ \emph{given} that she has been admitted to $c_j$ and applied to $c_k$.
+For $j < k$ (including $j = 0$), this is $\tilde t_j = (1- f_k) t_j + f_k t_k$; for $j > k $, this is $\tilde t_j = t_j$.
+This means that
+\begin{equation}\label{Vyastildet}
+v(\mathcal{Y}\cup\{k\}) = \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y}).\end{equation}
+The transformation to $\tilde t$ does not change the order of the $t_j$-values.
+Therefore, the expression on the right side of \eqref{Vyastildet} is itself a portfolio valuation function.
+In the corresponding market, $t$ is replaced by $\tilde t$ and $\mathcal{C}$ is replaced by $\mathcal{C}\setminus\{k\}$.
+To restore our convention that $t_0 = 0$, we obtain $w(\mathcal{Y})$ by taking $\bar t_j = \tilde t_j - \tilde t_0$ for all $j \neq k$ and letting
+\begin{equation}  \label{wYvXminusconst}
+w(\mathcal{Y})
+= \sum_{j \in \{0\} \cup \mathcal{Y}} \bar t_j p_j(\mathcal{Y})
+= \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y})- \tilde t_0
+= v(\mathcal{Y}\cup\{k\}) -  f_k t_k \end{equation}
+where the second equality follows from Theorem \ref{assumetzerozero}.
+The validity of this transformation is summarized in the following theorem, where we write $v(\mathcal{X}; \bar t)$ instead of $w(\mathcal{Y})$ to emphasize that $w(\mathcal{Y})$ is, in form, a portfolio valuation function.
+\else
+$w(\mathcal{Y})$를 구성하기 위해, 알마가 $c_j$에 합격한 상황에서 $c_k$에 지원했을 때 $c_j$에서 받는 조건부 기대 효용을 $\tilde t_j$라고 하자.
+$j = 0$을 포함한 $j < k$에 대해 이값은 $\tilde t_j = t_j (1- f_k) + t_k f_k$이며, $j > k $에 대해 $\tilde t_j = t_j$임을 알 수 있다.
+따라서
+\begin{equation}\label{Vyastildet}
+v(\mathcal{Y}\cup\{k\}) = \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y}).\end{equation}
+$\tilde t$로 변환하면 $t_j$의 순서가 변하지 않기 때문에 \eqref{Vyastildet}의 우변 그 자체가 포트폴리오 가치 함수이다.
+$t$를 $\tilde t$로, $\mathcal{C}$를 $\mathcal{C}\setminus \{k\}$로 대체하면 대응하는 시장이 된다.
+그다음에 $t_0 = 0$이 되도록 $\bar t_j = \tilde t_j - \tilde t_0$으로 정의하고 $w(\mathcal{Y})$를 다음처럼 얻을 수 있다.
+\begin{equation}  \label{wYvXminusconst}
+w(\mathcal{Y})
+= \sum_{j \in \{0\} \cup \mathcal{Y}} \bar t_j p_j(\mathcal{Y})
+= \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y})- \tilde t_0
+= v(\mathcal{Y}\cup\{k\}) -  f_k t_k \end{equation}
+두 번째 등호는 정리 \ref{assumetzerozero}에 따라 성립한다. 다음 정리는 이 변환의 타당성을 의미한다. 단, $w(\mathcal{Y})$가 포트폴리오 가치 함수 형태를 가지는 것을 강조하기 위해 $w(\mathcal{Y})$ 대신 $v(\mathcal{X}; \bar t)$처럼 표기한다.
+\fi
+
+\begin{theorem}[\ifen Eliminate $c_k$\else $c_k$ 소거 기술\fi] \label{eliminationtheorem}
+\ifen For $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$, where
+\else $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$에 대해, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$이다. 단,\fi
+\begin{align}\label{howtotransformtj}
+\bar t_j = 
+\begin{cases}
+(1 - f_k) t_j, \quad & t_j \leq t_k \\
+t_j - f_k t_k, \quad& t_j > t_k.
+\end{cases}
+\end{align}
+\end{theorem}
+
+\begin{proof}
+\ifen It is easy to verify that \eqref{howtotransformtj} is the composition of the two transformations (from $t$ to $\tilde t$, and from $\tilde t$ to $\bar t$) discussed above.
+\else \eqref{howtotransformtj}이(가) 위에서 논의한 2개의 변환($t$에서 $\tilde t$로, 그리고 $\tilde t$에서 $\bar t$로)의 합성임을 쉽게 확인할 수 있다. \fi
+\end{proof}
+
+\ifen
+\noindent The transformation \eqref{howtotransformtj} can be applied iteratively to accommodate the case where the student has already resolved to apply to multiple schools.
+\else
+\noindent 학생이 다수의 학교에 지원하기로 결심한 경우를 반영하기 위해, \eqref{howtotransformtj}의 변화를 반복적으로 적용할 수 있다.
+\fi
+
+\ifen \subsection{Submodularity of the objective function} \else \subsection{목적함수의 submodularity} \fi
+\ifen 
+Now, we show that the portfolio valuation function is submodular.
+This result is primarily of taxonomical interest and may be safely skipped.
+Our solution algorithms for the college application problem are better than generic algorithms for submodular maximization, and we prove their validity using elementary analysis.
+\else
+목적함수가 submodular함을 보이자. 이 결과는 주로 분류학적인 흥미를 가지며 건너뛰어도 연속성에 지장을 주지 않는다. 일반적인 submodular 최대화 알고리즘보다 본 연구가 제시하는 대학 지원 최적화 해법이 더 좋으며, 기초적인 분석을 통해 그의 타당성을 증명한다.
+\fi
+
+\ifen
+\begin{definition}[Submodular set function]
+Given a ground set $\mathcal{C}$ and function $v : 2^{\mathcal{C}} \mapsto \mathbb{R}$, $v(\mathcal{X})$ is called a \emph{submodular set function} if and only if $v(\mathcal{X}) + v(\mathcal{Y}) \geq v(\mathcal{X}\cup\mathcal{Y}) + v(\mathcal{X}\cap\mathcal{Y})$
+for all $\mathcal{X}, \mathcal{Y} \subseteq \mathcal{C}$.
+Furthermore, if $ v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) \geq 0$ for all $\mathcal{X} \subset \mathcal{C}$ and $k \in \mathcal{C} \setminus \mathcal{X}$, $v(\mathcal{X})$ is said to be a \emph{nondecreasing} submodular set function.
+\end{definition}
+\else
+\begin{definition}[Submodular한 집합 함수]
+고려 집합 $\mathcal{C}$ 그리고 함수 $v : 2^{\mathcal{C}} \mapsto \mathbb{R}$가 주어질 때, $v(\mathcal{X})$가 submodular한 집합 함수가 되는 필수충분 조건은, 모든 $\mathcal{X}, \mathcal{Y} \subseteq \mathcal{C}$에 대해 $v(\mathcal{X}) + v(\mathcal{Y}) \geq v(\mathcal{X}\cup\mathcal{Y}) + v(\mathcal{X}\cap\mathcal{Y})$이 성립하는 것이다.
+또한 모든 $\mathcal{X} \subset \mathcal{C}$와 $k \in \mathcal{C} \setminus \mathcal{X}$에 대해 $v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) \geq 0$이면, $v(\mathcal{X})$가 단조 증가하는 submodular한 집합 함수라고 부른다.
+\end{definition}
+\fi
+
+\ifen
+\begin{theorem}
+The college application portfolio valuation function
+$v(\mathcal{X})$ % = \sum_{j\in\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)\]
+is a nondecreasing submodular set function.
+\end{theorem}
+\else
+\begin{theorem}
+대학 지원의 포트폴리오 가치 함수 $v(\mathcal{X})$는 submodular한 집합 함수이다.
+\end{theorem}
+\fi
+
+\ifen
+\begin{proof}
+It is self-evident that $v(\mathcal{X})$ is nondecreasing. To establish its submodularity, we apply proposition 2.1.iii of Nemhauser and Wolsey (1978) and show that
+\begin{equation}\label{nemhauseriii}
+v(\mathcal{X} \cup \{j\}) - v(\mathcal{X}) \geq 
+v(\mathcal{X} \cup \{j, k\}) - v(\mathcal{X} \cup \{k\})
+\end{equation}
+for $\mathcal{X} \subset \mathcal{C}$ and $j \neq k \in \mathcal{C} \setminus \mathcal{X}$.
+By repeatedly eliminating the schools in $\mathcal{X}$ according to \eqref{howtotransformtj}, we can obtain a portfolio valuation function $w(\mathcal{Y})$ with parameter $\bar t$ such that $w(\mathcal{Y}) = v(\mathcal{X} \cup \mathcal{Y}) + \text{const.}$ for any $\mathcal{Y} \subseteq \mathcal{C} \setminus \mathcal{X}$.
+Therefore, \eqref{nemhauseriii} is equivalent to
+\begin{align}
+& w(\{j\}) - w(\varnothing) \geq w(\{j, k\}) - w(\{k\}) \\
+\iff \qquad &w(\{j\})  +  w(\{k\})  \geq w(\{j, k\})  \\
+\iff \qquad &\operatorname{E}[\,\bar t_j Z_j\,] + \operatorname{E}[\,\bar t_k Z_k\,] 
+\geq \operatorname{E}\bigl[\max\{ \bar t_j Z_j, \bar t_k Z_k \} \bigr]
+\end{align}
+which is plainly true. 
+\end{proof}
+\else
+\begin{proof}
+$v(\mathcal{X})$가 단조 증가하는 것은 자명하다.
+Submodularity를 보이기 위해, Nemhauser와 Wolsey (1978)의 제의 2.1.iii을 적용하여 모든 $\mathcal{X} \subset \mathcal{C}$ 그리고 $j \neq k \in \mathcal{C} \setminus \mathcal{X}$에 대해 다음이 성립함을 보이자.
+\begin{equation}\label{nemhauseriii}
+v(\mathcal{X} \cup \{j\}) - v(\mathcal{X}) \geq 
+v(\mathcal{X} \cup \{j, k\}) - v(\mathcal{X} \cup \{k\})
+\end{equation}
+\eqref{howtotransformtj}에 따라 $\mathcal{X}$에 속한 학교를 반복적으로 소거하면 모수 $\bar t$로 갖춘 포트폴리오 함수 $w(\mathcal{Y})$를 얻을 수 있으며,  모든 $\mathcal{Y} \subseteq \mathcal{C} \setminus \mathcal{X}$에 대해 $w(\mathcal{Y}) = v(\mathcal{X} \cup \mathcal{Y}) + \text{const.}$가 성립한다.
+따라서 \eqref{nemhauseriii}은(는) 다음 부등식과 동등하다.
+\begin{align}
+& w(\{j\}) - w(\varnothing) \geq w(\{j, k\}) - w(\{k\}) \\
+\iff \qquad &w(\{j\})  +  w(\{k\})  \geq w(\{j, k\})  \\
+\iff \qquad &\operatorname{E}[\,\bar t_j Z_j\,] + \operatorname{E}[\,\bar t_k Z_k\,] 
+\geq \operatorname{E}\bigl[\max\{ \bar t_j Z_j, \bar t_k Z_k \} \bigr]
+\end{align}
+마지막 부등식은 분명히 참이다.
+\end{proof}
+\fi
 
@@
 By definition, $\mathcal{Y}_{h-1}$ (respectively, $\mathcal{Y}_{h}$) maximizes the function $v(\mathcal{Y}\cup\{k\})$ over portfolios of size $h-1$ (respectively, $h$) that do not include $k$.
-That is, $\mathcal{Y}_{h-1}$ and $\mathcal{Y}_h$ are the optimal \emph{complements} to the singleton portfolio $\{k\}$.
+That is, $\mathcal{Y}_{h-1}$ and $\mathcal{Y}_h$ are the optimal complements to the singleton portfolio $\{k\}$.
+
+Applying Theorem \ref{eliminationtheorem}, we eliminate $c_k$, transform the remaining $t_j$-values to $\bar t_j$ according to \eqref{howtotransformtj}, and obtain a function $w(\mathcal{Y}) = v(\mathcal{Y} \cup \{k\}) - f_k t_k$ that grades portfolios $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$ according to how well they complement $\{k\}$.
+Since $w(\mathcal{Y})$ is itself a portfolio valuation funtion and $\bar t_0 = 0$, the inductive hypothesis implies that $\mathcal{Y}_{h-1} \subset \mathcal{Y}_h$, which completes the proof.
+
 \else 이제 $k \in \mathcal{X}_h$인 경우를 고려하자.
@@
 정의에 따라 $\mathcal{Y}_{h-1}$ ($\mathcal{Y}_{h}$)는 크기가 $h-1$ ($h$)인 포트폴리오 중에 함수 $v(\mathcal{Y}\cup\{k\})$를 최대화한다.
-다시 말해 $\mathcal{Y}_{h-1}$과 $\mathcal{Y}_h$는 각각 단일 원소 포트폴리오 $\{k\}$와 최적으로 `보완적인' 학교 집합이다.
-\fi
-
-\ifen We will use the function $w(\mathcal{Y})$ to grade portfolios $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$ according to how well they complement $\{k\}$.
+다시 말해 $\mathcal{Y}_{h-1}$과 $\mathcal{Y}_h$는 각각 단일 원소 포트폴리오 $\{k\}$와 최적으로 보완적인 학교 집합이다.
+
+정리 \ref{eliminationtheorem}을(를) 적용하여 $c_k$를 소거하고 남은 $t_j$-값을 \eqref{howtotransformtj}에 따라 $\bar t_j$로 변환한다.
+그러면 $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$인 포트폴리오와 $\{k\}$ 사이의 보완성을 평가하는 함수 $w(\mathcal{Y})$를 얻는다.
+$w(\mathcal{Y})$ 그 자체가 포트폴리오 가치 함수이며 $\bar t_0 = 0$이므로, 귀납법 가설에 따라 $\mathcal{Y}_{h-1} \subset \mathcal{Y}_h$이고 증명이 완성된다.
+\fi
+\end{proof}
+
+%\ifen We will use the function $w(\mathcal{Y})$ to grade portfolios $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$ according to how well they complement $\{k\}$.
 To construct $w(\mathcal{Y})$, let $\tilde t_j$ denote the expected utility Alma receives from school $c_j$ \emph{given} that she has been admitted to $c_j$ and applied to $c_k$.
@@
 This means that
-\else 함수 $w(\mathcal{Y})$를 이용하여 $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$인 포트폴리오와 $\{k\}$ 사이의 \mbox{`보완성'을} 평가하자.
+%\else 함수 $w(\mathcal{Y})$를 이용하여 $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$인 포트폴리오와 $\{k\}$ 사이의 \mbox{`보완성'을} 평가하자.
 $w(\mathcal{Y})$를 구성하기 위해, 알마가 $c_j$에 합격한 상황에서 $c_k$에 지원했을 때 $c_j$에서 받는 조건부 기대 효용을 $\tilde t_j$라고 하자.
@@
 따라서 \fi
-\begin{equation}\label{Vyastildet}
-v(\mathcal{Y}\cup\{k\}) = \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y}).\end{equation}
-\ifen The transformation to $\tilde t$ does not change the order of the $t_j$-values.
+%\begin{equation}\label{Vyastildet}
+%v(\mathcal{Y}\cup\{k\}) = \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y}).\end{equation}
+%\ifen The transformation to $\tilde t$ does not change the order of the $t_j$-values.
 Therefore, the expression on the right side of \eqref{Vyastildet} is itself a portfolio valuation function.
@@
 Now, we obtain $w(\mathcal{Y})$ through one more transformation: Define $\bar t_j = \tilde t_j - \tilde t_0$ so that $t_0 = 0$ and let
-\else $\tilde t$로 변환하면 $t_j$의 순서가 변하지 않기 때문에 \eqref{Vyastildet}의 우변 그 자체가 포트폴리오 가치 함수이다.
+%\else $\tilde t$로 변환하면 $t_j$의 순서가 변하지 않기 때문에 \eqref{Vyastildet}의 우변 그 자체가 포트폴리오 가치 함수이다.
 $t$를 $\tilde t$로, $\mathcal{C}$를 $\mathcal{C}\setminus \{k\}$로 대체하면 대응하는 시장이 된다.
@@
 \fi
-\begin{equation}  \label{wYvXminusconst}
-w(\mathcal{Y})
-= \sum_{j \in \{0\} \cup \mathcal{Y}} \bar t_j p_j(\mathcal{Y})
-= \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y})- \tilde t_0
-= v(\mathcal{Y}\cup\{k\}) -  f_k t_k \end{equation}
-\ifen where the second equality follows from Theorem \ref{assumetzerozero}.
+%\begin{equation}  \label{wYvXminusconst}
+%w(\mathcal{Y})
+%= \sum_{j \in \{0\} \cup \mathcal{Y}} \bar t_j p_j(\mathcal{Y})
+%= \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y})- \tilde t_0
+%= v(\mathcal{Y}\cup\{k\}) -  f_k t_k \end{equation}
+%\ifen where the second equality follows from Theorem \ref{assumetzerozero}.
 This identity says that the optimal complements to $\{k\}$, given by $\mathcal{Y}_{h-1}$ and $\mathcal{Y}_h$, are themselves optimal portfolios of size $h-1$ and $h$ for the market whose objective function is $w(\mathcal{Y})$.
 Since $\bar t_0 = 0$ in the latter market, the inductive hypothesis implies that $\mathcal{Y}_{h-1} \subset \mathcal{Y}_h$, which completes the proof.\footnote{We thank Yim Seho for discovering this useful transformation.}
-\else 두 번째 등호는 정리 \ref{assumetzerozero}에 따라 성립한다.
+%\else 두 번째 등호는 정리 \ref{assumetzerozero}에 따라 성립한다.
 이 등호는, $\{k\}$와 최적으로 보완적인 $\mathcal{Y}_{h-1}$과 $\mathcal{Y}_h$는 목적함수가 $w(\mathcal{Y})$인 시장에서 각각 크기가 $h-1$과 $h$인 최적 포트폴리오임을 의미한다.
@@
 따라서 증명이 완성된다.\footnote{이 유용한 변환을 발견한 임세호에게 감사를 표한다.} \fi
-\end{proof}
 
@@
 \ifen 
-We reduce the computation time to $O(hm)$ by taking advantage of the transformation from the inductive step in the proof of Theorem \ref{nestedapplication}.
-Once school $k$ is added to $\mathcal{X}$, we eliminate it from the set $\mathcal{C}\setminus \mathcal{X}$ of candidates, and update the $t_j$-values of the remaining schools according to the following relation.
-\else
-정리 \ref{nestedapplication} 증명의 추론 단계에서 제시한 변환을 활용하면 계산 시간을 $O(hm)$으로 감소시킬 수 있다. $\mathcal{X}$에 $k$를 추가하면, 후보자 집합인 $\mathcal{C}\setminus \mathcal{X}$에서 $k$를 소거하고 남은 학교의 $t_j$-값을 다음처럼 수정한다.
-\fi
-\begin{align}\label{howtotransformfj}
-\bar t_j = 
-\begin{cases}
-(1 - f_k) t_j, \quad & t_j \leq t_k \\
-t_j - f_k t_k, \quad& t_j > t_k
-\end{cases}
-\end{align}
-\ifen
-\noindent It is easy to verify that \eqref{howtotransformfj} is the composition of the two transformations (from $t$ to $\tilde t$, and from $\tilde t$ to $\bar t$) given in the proof.
-(We will reuse this transformation as a variable-elimination technique in the branch-and-bound algorithm presented in Subsection \ref{subsectionbranchbound}.)
-
+We reduce the computation time to $O(hm)$ by taking advantage of the transformation from Theorem \ref{eliminationtheorem}.
+Once school $k$ is added to $\mathcal{X}$, we eliminate it from the set $\mathcal{C}\setminus \mathcal{X}$ of candidates, and update the $t_j$-values of the remaining schools according to \eqref{howtotransformtj}.
 Now, the \emph{next} school added must be the optimal singleton portfolio in the modified market.
 But the optimal singleton portfolio consists simply of the school with the highest value of $f_j \bar t_j$.
-Therefore, by updating the $t_j$-values at each iteration according to \eqref{howtotransformfj}, we eliminate the need to compute $v(\mathcal{X})$ entirely.
+Therefore, by updating the $t_j$-values at each iteration according to \eqref{howtotransformtj}, we eliminate the need to compute $v(\mathcal{X})$ entirely.
 Moreover, this algorithm does not require the schools to be indexed in ascending order by $t_j$, which removes the $O(m\log m)$ sorting cost.
 \else
-\noindent \eqref{howtotransformfj}이(가) 증명에서 정의한 2개의 변환($t$에서 $\tilde t$로, 그리고 $\tilde t$에서 $\bar t$로)의 합성임을 쉽게 확인할 수 있다. (항 \ref{subsectionbranchbound}에서 제시하는 분지한계법에서 이 변환을 변수 소거법으로 재활용한다.)
-
+정리 \ref{eliminationtheorem}에서 제시한 변환을 활용하면 계산 시간을 $O(hm)$으로 감소시킬 수 있다.
+$\mathcal{X}$에 $k$를 추가하면, 후보자 집합인 $\mathcal{C}\setminus \mathcal{X}$에서 $k$를 소거하고 남은 학교의 $t_j$-값을 \eqref{howtotransformtj}에 따라 수정한다.
 그러면 다음으로 추가하는 학교는 수정된 시장의 최적 단일 원소 포트폴리오가 되어야 한다.
 그런데 최적 단일 원소 포트폴리오는 단순히 $f_j \bar t_j$가 가장 큰 학교로 이루어진다.
-따라서 각 반복 단계에서 \eqref{howtotransformfj}을(를) 이용하여 $t_j$-값을 수정하면 $v(\mathcal{X})$를 계산할 필요가 완전히 없어진다.
+따라서 각 반복 단계에서 \eqref{howtotransformtj}을(를) 이용하여 $t_j$-값을 수정하면 $v(\mathcal{X})$를 계산할 필요가 완전히 없어진다.
 게다가 이 알고리즘에서 $t_j$를 배열할 필요가 없으므로 $O(m\log m)$인 배열 비용이 사라진다.
 \fi
-
 \ifen
@@
 \end{theorem}
+
 \begin{proof}
-\ifen We will prove the equivalent expression $2 v(\mathcal{X}_h) \geq v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1})$.
-Applying Theorem \ref{nestedapplication}, we write $\mathcal{X}_h = \mathcal{X}_{h-1} \cup\{j\}$ and $\mathcal{X}_{h+1} = \mathcal{X}_{h-1} \cup\{j, k\}$.
-If $t_k \leq t_j$, then
-\else $2 v(\mathcal{X}_h) \geq v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1})$ 같은 동등한 부등식을 증명하자.
-정리 \ref{nestedapplication}을(를) 적용하면 $ \mathcal{X}_h = \mathcal{X}_{h-1} \cup\{j\}$ 그리고 $\mathcal{X}_{h+1} = \mathcal{X}_{h-1} \cup\{j, k\}$으로 표현할 수 있다.
-$t_k \leq t_j$인 경우, \fi
-\begin{align}
-\begin{split}
-2 v(\mathcal{X}_h) &= v(\mathcal{X}_{h-1} \cup\{j\}) + v(\mathcal{X}_{h-1} \cup\{j\}) \\
-&\geq v(\mathcal{X}_{h-1} \cup\{k\}) + v(\mathcal{X}_{h-1} \cup\{j\}) \\
-&= v(\mathcal{X}_{h-1} \cup\{k\}) + (1 - f_j) v(\mathcal{X}_{h-1}) + f_j \operatorname{E}[\max\{t_j, X_{h-1}\}] \\
-&= v(\mathcal{X}_{h-1} \cup\{k\}) - f_j v(\mathcal{X}_{h-1}) + f_j \operatorname{E}[\max\{t_j, X_{h-1}\}] + v(\mathcal{X}_{h-1})  \\
-&\geq v(\mathcal{X}_{h-1} \cup\{k\})  - f_j v(\mathcal{X}_{h-1}\cup\{k\}) + f_j \operatorname{E}[\max\{t_j, X_{h-1}\}]+ v(\mathcal{X}_{h-1})\\
-&= (1 - f_j) v(\mathcal{X}_{h-1} \cup\{k\})  + f_j \operatorname{E}[\max\{t_j, X_{h-1}\}]+ v(\mathcal{X}_{h-1})\\
-&=  v(\mathcal{X}_{h-1} \cup\{j, k\}) + v(\mathcal{X}_{h-1})\\
-&=  v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1}).
-\end{split} 
-\end{align}
-\ifen The first inequality follows from the optimality of $\mathcal{X}_h$, while the second follows from the fact that adding $k$ to $\mathcal{X}_{h-1}$ can only increase its valuation.
-\else 첫 번째 부등식은 $\mathcal{X}_h$의 최적성에 따르며, 두 번째 부등식은 $\mathcal{X}_{h-1}$에 $j$를 더하면 가치가 증가할 수밖에 없기 때문이다.\fi
-
-\ifen If $t_k \geq t_j$, then the steps are analogous:
-\else  $t_k \geq t_j$인 경우는 유사하다:\fi
-\begin{align}
-\begin{split}
-2 v(\mathcal{X}_h) &= v(\mathcal{X}_{h-1} \cup\{j\}) + v(\mathcal{X}_{h-1} \cup\{j\}) \\
-&\geq v(\mathcal{X}_{h-1} \cup\{k\}) + v(\mathcal{X}_{h-1} \cup\{j\}) \\
-&= (1 - f_k) v(\mathcal{X}_{h-1}) + f_k \operatorname{E}[\max\{t_k, X_{h-1}\}] +  v(\mathcal{X}_{h-1} \cup\{j\})  \\
-&= v(\mathcal{X}_{h-1}) - f_k  v(\mathcal{X}_{h-1}) + f_k \operatorname{E}[\max\{t_k, X_{h-1}\}] +  v(\mathcal{X}_{h-1} \cup\{j\})  \\
-&\geq v(\mathcal{X}_{h-1}) - f_k  v(\mathcal{X}_{h-1}\cup\{j\}) + f_k \operatorname{E}[\max\{t_k, X_{h-1}\}] +  v(\mathcal{X}_{h-1} \cup\{j\})  \\
-&= v(\mathcal{X}_{h-1}) + (1 - f_k) v(\mathcal{X}_{h-1}\cup\{j\}) + f_k \operatorname{E}[\max\{t_k, X_{h-1}\}]  \\
-&= v(\mathcal{X}_{h-1}) + v(\mathcal{X}_{h-1} \cup\{j, k\})
-\end{split}\\
-&= v(\mathcal{X}_{h-1})  + v(\mathcal{X}_{h+1}) \qedhere
-\end{align}
+\ifen Applying Theorem \ref{nestedapplication}, we write $\mathcal{X}_h = \mathcal{X}_{h-1} \cup\{j\}$ and $\mathcal{X}_{h+1} = \mathcal{X}_{h-1} \cup\{j, k\}$.
+By optimality, $v(\mathcal{X}_h) - v(\mathcal{X}_{h-1}) \geq v(\mathcal{X}_{h-1}\cup\{k\}) - v(\mathcal{X}_{h-1})$.
+By submodularity and nestedness, $v(\mathcal{X}_{h-1}\cup\{k\}) - v(\mathcal{X}_{h-1}) \geq  v(\mathcal{X}_{h}\cup\{k\}) - v(\mathcal{X}_{h}) = v(\mathcal{X}_{h+1}) - v(\mathcal{X}_{h})$.
+\else 정리 \ref{nestedapplication}을(를) 적용하면 $ \mathcal{X}_h = \mathcal{X}_{h-1} \cup\{j\}$ 그리고 $\mathcal{X}_{h+1} = \mathcal{X}_{h-1} \cup\{j, k\}$으로 표현할 수 있다.
+최적성에 따라, $v(\mathcal{X}_h) - v(\mathcal{X}_{h-1}) \geq v(\mathcal{X}_{h-1}\cup\{k\}) - v(\mathcal{X}_{h-1})$.
+Submodularity와 포함 사슬 관계 성질에 따라, $v(\mathcal{X}_{h-1}\cup\{k\}) - v(\mathcal{X}_{h-1}) \geq  v(\mathcal{X}_{h}\cup\{k\}) - v(\mathcal{X}_{h}) = v(\mathcal{X}_{h+1}) - v(\mathcal{X}_{h})$.
+\fi
 \end{proof}
 \ifen
-\noindent It follows that when $\mathcal{X}_h$ is the optimal $h$-portfolio for a given market, $v(\mathcal{X}_h)$ is $O(h)$.
+\noindent (An elementary proof is given in \S\,\ref{elementaryconcavityproof}.) It follows that when $\mathcal{X}_h$ is the optimal $h$-portfolio for a given market, $v(\mathcal{X}_h)$ is $O(h)$.
 Example \ref{tightexampleforoneoverhopt}, in which $v(\mathcal{X}_h)$ can be made arbitrarily close to $h$, establishes the tightness of this bound.
 \else
-\noindent 위 결과는 $\mathcal{X}_h$가 어떤 시장의 최적 $h$-포트폴리오일 때, $v(\mathcal{X}_h)$가 $O(h)$ 함수라고 의미한다. 예 \ref{tightexampleforoneoverhopt}에서 $v(\mathcal{X}_h)$를 $h$에 임의로 가깝게 조정할 수 있으므로 이 상한이 타이트함을 알 수 있다.
+\noindent (기초적인 증명은 \S\,\ref{elementaryconcavityproof}에서 제시한다.) 위 결과는 $\mathcal{X}_h$가 어떤 시장의 최적 $h$-포트폴리오일 때, $v(\mathcal{X}_h)$가 $O(h)$ 함수라고 의미한다.
+예 \ref{tightexampleforoneoverhopt}에서 $v(\mathcal{X}_h)$를 $h$에 임의로 가깝게 조정할 수 있으므로 이 상한이 타이트함을 알 수 있다.
 \fi
@@
 In both subproblems, $\bar H = H - \sum_{j\in \mathcal{I}} g_j$ denotes the residual budget.
-The parameters $\gamma$ and $\bar t$ are obtained by iteratively applying the transformation \eqref{howtotransformfj} to the schools in $\mathcal{I}$.
-For each $j\in \mathcal{I}$,  we increment $\gamma$ by the current value of $f_j \bar t_j$, eliminate $j$ from the market, and update the remaining $\bar t_j$-values using \eqref{howtotransformfj}.
+The parameters $\gamma$ and $\bar t$ are obtained by iteratively applying the transformation \eqref{howtotransformtj} to the schools in $\mathcal{I}$.
+For each $j\in \mathcal{I}$,  we increment $\gamma$ by the current value of $f_j \bar t_j$, eliminate $j$ from the market, and update the remaining $\bar t_j$-values using \eqref{howtotransformtj}.
 (An example is given below.)
@@
 각 부문제에서 $\bar H = H - \sum_{j\in \mathcal{I}} g_j$는 잔여 예산을 의미한다.
-$\mathcal{I}$에 속한 학교에 \eqref{howtotransformfj}의 변환을 적용하여 모수 $\gamma$와 $\bar t$를 얻을 수 있다.
-즉, 각 $j\in \mathcal{I}$에 대해 $\gamma$를 $f_j \bar t_j$ 만큼 증가시키고, 시장에서 $j$를 소거하고, \eqref{howtotransformfj}에 따라 남은 $\bar t_j$-값을 수정한다.
+$\mathcal{I}$에 속한 학교에 \eqref{howtotransformtj}의 변환을 적용하여 모수 $\gamma$와 $\bar t$를 얻을 수 있다.
+즉, 각 $j\in \mathcal{I}$에 대해 $\gamma$를 $f_j \bar t_j$ 만큼 증가시키고, 시장에서 $j$를 소거하고, \eqref{howtotransformtj}에 따라 남은 $\bar t_j$-값을 수정한다.
 (아래에서 예를 제시한다.)
@@
 In principle, any school can be chosen for $j$, but as a greedy heuristic, we choose the school for which the ratio $f_j \bar t_j / g_j$ is highest.
-Notice that this method of generating children ensures that each node's $\mathcal{I}$-set differs from its parent's by at most a single school, so the constant $\gamma$ and transformed $\bar t_j$-values for the new node can be obtained by a single application of \eqref{howtotransformfj}.
+Notice that this method of generating children ensures that each node's $\mathcal{I}$-set differs from its parent's by at most a single school, so the constant $\gamma$ and transformed $\bar t_j$-values for the new node can be obtained by a single application of \eqref{howtotransformtj}.
 \else
@@
 이런 식으로 새끼를 생성하면 각 마디의 $\mathcal{I}$-집합은 자기 어미 마디의 $\mathcal{I}$-집합과 최대 한 학교로만 다르다.
-따라서 새로운 마디를 생성할 때, \eqref{howtotransformfj}을(를) 한 번만 적용하면 상수 $\gamma$와 변화된 $\bar t_j$-값을 얻을 수 있다.
+따라서 새로운 마디를 생성할 때, \eqref{howtotransformtj}을(를) 한 번만 적용하면 상수 $\gamma$와 변화된 $\bar t_j$-값을 얻을 수 있다.
 \fi
@@
 To compute the subproblems, we first simply disregard $c_1$.
-Next, to eliminate $c_2$, we apply \eqref{howtotransformfj} to $t$ to obtain
+Next, to eliminate $c_2$, we apply \eqref{howtotransformtj} to $t$ to obtain
 $\{\bar{\bar t}_3, \bar{\bar t}_4, \bar{\bar t}_5\} = \{ (1 - f_2) t_3, (1 - f_2) t_4, (1 -f_2) t_5\} = \{30, 40, 50 \}$ and $\bar{\bar \gamma} = f_2 t_2 = 20$.
-We eliminate $c_5$ by again applying \eqref{howtotransformfj} to $\bar{ \bar t}$ to obtain $\{\bar t_3, \bar t_4\} =  \{\bar{\bar t}_3 - f_5 \bar{\bar t}_5, \bar{\bar t}_4-  f_5 \bar{\bar t}_5\}= \{5, 15\}$ and $\gamma = \bar{\bar \gamma} + f_5 \bar{\bar t}_5 = 35$.
+We eliminate $c_5$ by again applying \eqref{howtotransformtj} to $\bar{ \bar t}$ to obtain $\{\bar t_3, \bar t_4\} =  \{\bar{\bar t}_3 - f_5 \bar{\bar t}_5, \bar{\bar t}_4-  f_5 \bar{\bar t}_5\}= \{5, 15\}$ and $\gamma = \bar{\bar \gamma} + f_5 \bar{\bar t}_5 = 35$.
 Finally, $\bar H = H - g_2 - g_5 = 3$.
@@
 부문제를 계산하기 위해 먼저 $c_1$을 단순히 무시한다.
-$c_2$를 소거하기 위해 $t$에 \eqref{howtotransformfj}에 적용하여
+$c_2$를 소거하기 위해 $t$에 \eqref{howtotransformtj}에 적용하여
 $\{\bar{\bar t}_3, \bar{\bar t}_4, \bar{\bar t}_5\} = \{ (1 - f_2) t_3, (1 - f_2) t_4, (1 -f_2) t_5\} = \{30, 40, 50 \}$과 $\bar{\bar \gamma} = f_2 t_2 = 20$을 얻는다.
-$c_5$를 소거하기 위해 $\bar{\bar t}$에 \eqref{howtotransformfj}에 다시 적용하여 $\{\bar t_3, \bar t_4\} =  \{\bar{\bar t}_3 - f_5 \bar{\bar t}_5, \bar{\bar t}_4-  f_5 \bar{\bar t}_5\}= \{5, 15\}$와 $\gamma = \bar{\bar \gamma} + f_5 \bar{\bar t}_5 = 35$를 얻는다.
+$c_5$를 소거하기 위해 $\bar{\bar t}$에 \eqref{howtotransformtj}에 다시 적용하여 $\{\bar t_3, \bar t_4\} =  \{\bar{\bar t}_3 - f_5 \bar{\bar t}_5, \bar{\bar t}_4-  f_5 \bar{\bar t}_5\}= \{5, 15\}$와 $\gamma = \bar{\bar \gamma} + f_5 \bar{\bar t}_5 = 35$를 얻는다.
 마지막으로, $\bar H = H - g_2 - g_5 = 3$이다.
@@
 In this section, we present a simulated-annealing algorithm for Ellis's problem.
-It is based on the procedure for integer programs outlined by Wolsey (1998, \S\,12.3.2).
+It is patterned on the procedure for integer programs outlined by Wolsey (1998, \S\,12.3.2).
 \else
@@
 Overall, we found the exact dynamic program to be the fastest algorithm, while the FPTAS was rather slow, a result that echoes the results of computational studies on knapsack problems (Martello and Toth 1990, \S\,2.10).
-
 The strong performance of Algorithm \ref{ellisDP1} is partly attributable to the structure of our synthetic instances, in which application costs are small integers and $H$ is proportional in expectation to $m$, meaning the expected computation time is $O(m^2)$.
-However, the relative advantage of Algorithm \ref{ellisDP1} is even more pronounced in real college-application instances because the typical student's application budget accommodates at most a dozen or so schools and is constant in $m$.
+However, the relative advantage of Algorithm \ref{ellisDP1} may be even more pronounced in real college-application instances because the typical student's application budget accommodates at most a dozen or so schools and is constant in $m$.
 \else
@@
 분지한계법은 크기가 중간인 인스턴스에도 비효율적이었다.
-
 알고리즘 \ref{ellisDP1}의 좋은 실험적 성능은 부분적으로 가상 인스턴스의 구조 덕이다.
 가상 인스턴스의 지원 비용이 작은 정수이고 $H$가 $m$과 선형식으로 비례하는 것은 알고리즘의 기대 계산 시간이 $O(m^2)$임을 의미한다.
-그러나 전형적인 학생의 지원 예산은 최대한 열 몇 개의 학교에 지원할 수 있게 하며 $m$에 대해 상수이므로, 실제 대학 지원 문제 인스턴스에 적용하면 알고리즘 \ref{ellisDP1}이 더욱 뚜렷한 상대적 우위를 발휘한다.
+그러나 전형적인 학생의 지원 예산은 최대한 열 몇 개의 학교에 지원할 수 있게 하며 $m$에 대해 상수이므로, 실제 대학 지원 문제 인스턴스에 적용하면 알고리즘 \ref{ellisDP1}이 더욱 뚜렷한 상대적 우위를 발휘할 가능성이 있다.
 \fi
@@
 This study has introduced a novel combinatorial optimization problem that we call the college application problem.
-It can be viewed as a kind of integer asset allocation problem or nonlinear knapsack problem.
+It can be viewed as a kind of asset allocation or knapsack problem with a submodular objective.
 We showed that the special case in which colleges have identical application costs can be solved in polynomial time by a greedy algorithm because the optimal solutions are nested in the budget constraint.
@@
 본 연구는 새로운 조합 최적화 문제를 제안했으며 이를 대학 지원 최적화 문제라고 부른다.
-일종의 정수 자산 배분 문제 또는 확률적인 배낭 문제로 볼 수 있다.
+Submodular한 목적함수를 가지는 자산 배분 혹은 배낭 문제로 볼 수 있다.
 모든 대학의 지원 비용이 동일한 특수한 경우, 탐욕 알고리즘으로 다항 시간 안에 풀 수 있음을 보였으며 최적해가 예산 제약식에 대한 포함 사슬 관계를 가지기 때문이다.
@@
 When the application budget is small relative to the size of the admissions market, our algorithms recommend a similar approach (see the example of Subsection \ref{planetsexamplesection}).
-However, inspecting equation \eqref{howtotransformfj}, which discounts school utility parameters to reflect their marginal value relative to the schools already in the portfolio, reveals that low-utility schools are penalized more harshly than high utility schools in the marginal analysis.
+However, inspecting equation \eqref{howtotransformtj}, which discounts school utility parameters to reflect their marginal value relative to the schools already in the portfolio, reveals that low-utility schools are penalized more harshly than high utility schools in the marginal analysis.
 Consequentially, as the application budget grows, the optimal portfolio in our model tends to favor reach schools over safety schools.
@@
 입학 시장의 크기에 비해 지원 예산이 작은 경우, 본 연구에서 제시하는 알고리즘은 비슷한 전략을 추천한다 (항 \ref{planetsexamplesection}의 예제 참고).
-그러나 포트폴리오에 이미 삽입한 학교를 반영한 한계 가치를 나타내도록 효용 모수를 할인하는 수식 \eqref{howtotransformfj}을 살펴보면, 효용이 높은 학교보다 효용이 낮은 학교에 부과되는 페널티가 큰 것을 알 수 있다.
+그러나 포트폴리오에 이미 삽입한 학교를 반영한 한계 가치를 나타내도록 효용 모수를 할인하는 수식 \eqref{howtotransformtj}을 살펴보면, 효용이 높은 학교보다 효용이 낮은 학교에 부과되는 페널티가 큰 것을 알 수 있다.
 그 결과는 지원 예산이 커질수록 본 모형의 최적 포트폴리오가 안정 지원 학교보다 상향 지원 학교를 편드는 것이다.
@@
 completes the formulation.
-These are knapsack constraints; therefore, the algorithm of Kulik et al.
-(2013) can be used to obtain a $(1 - 1/e + \varepsilon)$-approximate solution for this problem.
+These are knapsack constraints; therefore, the submodular maximization algorithm of Kulik et al.
+(2013) can be used to obtain a $(1 - 1/e - \varepsilon)$-approximate solution for this problem.
 When the budget constraint is a cardinality constraint (as in Alma's problem), we note that adding these constraints yields a matriodal solution set, and Calinescu et al.
-(2011)'s algorithm yields an alternative solution, albeit with the same approximation coefficient of $1 - 1/e + \varepsilon$.
+(2011)'s algorithm yields an alternative solution, which has the same approximation coefficient of $1 - 1/e -\varepsilon$.
 %These constraints are conformable to those of the multiple knapsack problem (Kellerer et al.
@@
 \end{align}
-These are knapsack constraints; therefore, the algorithm of Kulik et al.
-(2013) can be used to obtain a $(1 - 1/e + \varepsilon)$-approximate solution for this problem.
-When the budget constraint is a cardinality constraint (as in Alma's problem), we note that adding these constraints yields a matriodal solution set, and Calinescu et al.
-(2011)'s algorithm yields an alternative solution, albeit with the same approximation coefficient of $1 - 1/e + \varepsilon$.
-%다중 배낭 문제의 제약식과 같은 꼴이다 (Kellerer et al. 2004, 10제). 
+모두 배낭 제약식이므로 Kulik 외 (2013)의 submodular 최대화 알고리즘을 통해 이 문제의 $(1 - 1/e - \varepsilon)$-근사해를 구할 수 있다.
+또한 (알마의 문제처럼) 예산 제약식이 집합 크기 제약식이 되면, 논리 제약식을 더한 모형의 해집합이 매트로이드가 되며, Calinescu 외 (2011)의 알고리즘이 대안 해법이 된다.
+단, 그의 근사 계수는 여전히 $1 - 1/e - \varepsilon$이다.
 \fi
@@
 
-\section{Appendix}
-
-N.B. these results will be worked into the body soon.
-
-\begin{definition}[Submodular set function]
-Given a ground set $\mathcal{C}$ and function $v : 2^{\mathcal{C}} \mapsto \mathbb{R}$, $v(\mathcal{X})$ is called a \emph{submodular set function} if and only if $v(\mathcal{X}) + v(\mathcal{Y}) \geq v(\mathcal{X}\cup\mathcal{Y}) + v(\mathcal{X}\cap\mathcal{Y})$
-for all $\mathcal{X}, \mathcal{Y} \subseteq \mathcal{C}$.
-Furthermore, if $ v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) \geq 0$ for all $\mathcal{X} \subset \mathcal{C}$ and $k \in \mathcal{C} \setminus \mathcal{X}$, $v(\mathcal{X})$ is said to be a \emph{nondecreasing} submodular set function.
-\end{definition}
-
-\begin{theorem}
-The college application portfolio valuation function
-$v(\mathcal{X})$ % = \sum_{j\in\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)\]
-is a nondecreasing submodular set function.
-\end{theorem}
-
+\appendix
+\ifen \section{Appendix} \else \section{부록} \fi
+
+\subsection{Elementary proof of Theorem \ref{concavityinh}} \label{elementaryconcavityproof}
 \begin{proof}
-It is self-evident that $v(\mathcal{X})$ is nondecreasing. To establish its submodularity, we apply proposition 2.1.iii of Nemhauser and Wolsey (1978) and show that
-\begin{equation}\label{nemhauseriii}
-v(\mathcal{X} \cup \{j\}) - v(\mathcal{X}) \geq 
-v(\mathcal{X} \cup \{j, k\}) - v(\mathcal{X} \cup \{k\})
-\end{equation}
-for $\mathcal{X} \subset \mathcal{C}$ and $j \neq k \in \mathcal{C} \setminus \mathcal{X}$.
-By repeatedly eliminating the schools in $\mathcal{X}$ according to \eqref{howtotransformfj}, we can obtain a portfolio valuation function $w(\mathcal{Y}; t)$ such that $w(\mathcal{Y}) = v(\mathcal{X} \cup \mathcal{Y}) + \text{const.}$ for any $\mathcal{Y} \subseteq \mathcal{C} \setminus \mathcal{X}$.
-Therefore, \eqref{nemhauseriii} is equivalent to
+\ifen We will prove the equivalent expression $2 v(\mathcal{X}_h) \geq v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1})$.
+Applying Theorem \ref{nestedapplication}, we write $\mathcal{X}_h = \mathcal{X}_{h-1} \cup\{j\}$ and $\mathcal{X}_{h+1} = \mathcal{X}_{h-1} \cup\{j, k\}$.
+If $t_k \leq t_j$, then
+\else $2 v(\mathcal{X}_h) \geq v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1})$ 같은 동등한 부등식을 증명하자.
+정리 \ref{nestedapplication}을(를) 적용하면 $ \mathcal{X}_h = \mathcal{X}_{h-1} \cup\{j\}$ 그리고 $\mathcal{X}_{h+1} = \mathcal{X}_{h-1} \cup\{j, k\}$으로 표현할 수 있다.
+$t_k \leq t_j$인 경우, \fi
 \begin{align}
-& w(\{j\}) - w(\varnothing) \geq w(\{j, k\}) - w(\{k\}) \\
-\iff \qquad &w(\{j\})  +  w(\{k\})  \geq w(\{j, k\})  \\
-\iff \qquad &\operatorname{E}[\,\bar t_j Z_j\,] + \operatorname{E}[\,\bar t_k Z_k\,] 
-\geq \operatorname{E}\bigl[\max\{ \bar t_j Z_j, \bar t_k Z_k \} \bigr]
+\begin{split}
+2 v(\mathcal{X}_h) &= v(\mathcal{X}_{h-1} \cup\{j\}) + v(\mathcal{X}_{h-1} \cup\{j\}) \\
+&\geq v(\mathcal{X}_{h-1} \cup\{k\}) + v(\mathcal{X}_{h-1} \cup\{j\}) \\
+&= v(\mathcal{X}_{h-1} \cup\{k\}) + (1 - f_j) v(\mathcal{X}_{h-1}) + f_j \operatorname{E}[\max\{t_j, X_{h-1}\}] \\
+&= v(\mathcal{X}_{h-1} \cup\{k\}) - f_j v(\mathcal{X}_{h-1}) + f_j \operatorname{E}[\max\{t_j, X_{h-1}\}] + v(\mathcal{X}_{h-1})  \\
+&\geq v(\mathcal{X}_{h-1} \cup\{k\})  - f_j v(\mathcal{X}_{h-1}\cup\{k\}) + f_j \operatorname{E}[\max\{t_j, X_{h-1}\}]+ v(\mathcal{X}_{h-1})\\
+&= (1 - f_j) v(\mathcal{X}_{h-1} \cup\{k\})  + f_j \operatorname{E}[\max\{t_j, X_{h-1}\}]+ v(\mathcal{X}_{h-1})\\
+&=  v(\mathcal{X}_{h-1} \cup\{j, k\}) + v(\mathcal{X}_{h-1})\\
+&=  v(\mathcal{X}_{h+1}) + v(\mathcal{X}_{h-1}).
+\end{split} 
 \end{align}
-which is plainly true. 
+\ifen The first inequality follows from the optimality of $\mathcal{X}_h$, while the second follows from the fact that adding $k$ to $\mathcal{X}_{h-1}$ can only increase its valuation.
+\else 첫 번째 부등식은 $\mathcal{X}_h$의 최적성에 따르며, 두 번째 부등식은 $\mathcal{X}_{h-1}$에 $j$를 더하면 가치가 증가할 수밖에 없기 때문이다.\fi
+
+\ifen If $t_k \geq t_j$, then the steps are analogous:
+\else  $t_k \geq t_j$인 경우는 유사하다:\fi
+\begin{align}
+\begin{split}
+2 v(\mathcal{X}_h) &= v(\mathcal{X}_{h-1} \cup\{j\}) + v(\mathcal{X}_{h-1} \cup\{j\}) \\
+&\geq v(\mathcal{X}_{h-1} \cup\{k\}) + v(\mathcal{X}_{h-1} \cup\{j\}) \\
+&= (1 - f_k) v(\mathcal{X}_{h-1}) + f_k \operatorname{E}[\max\{t_k, X_{h-1}\}] +  v(\mathcal{X}_{h-1} \cup\{j\})  \\
+&= v(\mathcal{X}_{h-1}) - f_k  v(\mathcal{X}_{h-1}) + f_k \operatorname{E}[\max\{t_k, X_{h-1}\}] +  v(\mathcal{X}_{h-1} \cup\{j\})  \\
+&\geq v(\mathcal{X}_{h-1}) - f_k  v(\mathcal{X}_{h-1}\cup\{j\}) + f_k \operatorname{E}[\max\{t_k, X_{h-1}\}] +  v(\mathcal{X}_{h-1} \cup\{j\})  \\
+&= v(\mathcal{X}_{h-1}) + (1 - f_k) v(\mathcal{X}_{h-1}\cup\{j\}) + f_k \operatorname{E}[\max\{t_k, X_{h-1}\}]  \\
+&= v(\mathcal{X}_{h-1}) + v(\mathcal{X}_{h-1} \cup\{j, k\})
+\end{split}\\
+&= v(\mathcal{X}_{h-1})  + v(\mathcal{X}_{h+1}) \qedhere
+\end{align}
 \end{proof}
+\ifen
+\noindent It follows that when $\mathcal{X}_h$ is the optimal $h$-portfolio for a given market, $v(\mathcal{X}_h)$ is $O(h)$.
+Example \ref{tightexampleforoneoverhopt}, in which $v(\mathcal{X}_h)$ can be made arbitrarily close to $h$, establishes the tightness of this bound.
+\else
+\noindent 위 결과는 $\mathcal{X}_h$가 어떤 시장의 최적 $h$-포트폴리오일 때, $v(\mathcal{X}_h)$가 $O(h)$ 함수라고 의미한다. 예 \ref{tightexampleforoneoverhopt}에서 $v(\mathcal{X}_h)$를 $h$에 임의로 가깝게 조정할 수 있으므로 이 상한이 타이트함을 알 수 있다.
+\fi
 
 \pagebreak
-\ifen \section{References} \else \section{참고문헌} \fi
+\ifen \section*{References} \else \section*{참고문헌} \fi
 \noindent
